### 2016-10-06 paper/paper.tex
Latex draft.

%\documentclass[preprint]{sigplanconf}
%\documentclass[preprint]{llncs}
\documentclass[a4paper,UKenglish]{lipics}

\usepackage{microtype}%if unwanted, comment out or use option "draft"
\usepackage[table,xcdraw]{xcolor}
\usepackage{color}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{theorem}
%% \usepackage{mathpartir}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}

\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{mathpartir}

% "define" code highlights for Java and Scala
\lstdefinelanguage{JavaScala}{
  morekeywords={public,int,interface,implements,default,
    abstract,case,catch,class,def,static,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,var,while,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\lstset{ %
language=JavaScala,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
morekeywords={var},
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

%%\usepackage{natbib}
%%\bibpunct();A{},
%%\let\cite=\citep

%include lhs2TeX.fmt
%include lhs2TeX.sty
%include forall.fmt

%\pagestyle{plain}

%{\theorembodyfont{\sffamily} \newtheorem{theorem}{Theorem}}
%{\theorembodyfont{\sffamily} \newtheorem{lemma}{Lemma}}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
%\newcommand{\hl}[1]{\textcolor{red}{#1}}
\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

\newcommand\mixin{\textbf{@Mixin} }
\usepackage{xspace}

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Type-Safe Modular Parsing}
\titlerunning{Type-Safe Modular Parsing}

\author[1]{John Q. Open}
\author[2]{Joan R. Access}
\affil[1]{Dummy University Computing Laboratory\\
  Address, Country\\
  \texttt{open@dummyuni.org}}
\affil[2]{Department of Informatics, Dummy College\\
  Address, Country\\
  \texttt{access@dummycollege.org}}
\authorrunning{J.\,Q. Open and J.\,R. Access}
% mandatory. First: Use abbreviated first/middle names. Second (only in severe
% cases): Use first author plus 'et. al.'

\Copyright{John Q. Open and Joan R. Access}
% mandatory, please use full first names. LIPIcs license is "CC-BY";
% http://creativecommons.org/licenses/by/3.0/

\subjclass{Dummy classification -- please refer to
  \url{http://www.acm.org/about/class/ccs98-html}}
% mandatory: Please choose ACM 1998 classifications from
% http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of
% Computation".
\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\serieslogo{}%please provide filename (without suffix)
\volumeinfo%(easychair interface)
  {Billy Editor and Bill Editors}% editors
  {2}% number of editors: 1, 2, ....
  {Conference title on which this volume is based on}% event
  {1}% volume
  {1}% issue
  {1}% starting page number
\EventShortName{}
\DOI{10.4230/LIPIcs.xxx.yyy.p}% to be completed by the volume editor
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle

\begin{abstract}

Text.

\end{abstract}

%if False

%\category{D.3.2}{Programming Languages}
%                {Language Classifications}
%                [Functional Languages]
%\category{F.3.3}{Logics and Meanings of Programs}
%                {Studies of Program Constructs}
%                []

%\terms
%Languages
%
%\keywords
%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
%interference

%endif

%===============================================================================

\input{sections/Introduction.tex}

\input{sections/Overview.tex}

\input{sections/Library.tex}

%...

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}

%===============================================================================

\bibliographystyle{plain}
\bibliography{paper}

\appendix

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "."
%%% End:

### 2016-10-06 paper/sections/CaseStudy.tex
Latex draft.

\section{Case Studies}\label{sec:casestudy}

\begin{itemize}
\item Explain the TAPL case study
\item Show various results SLOC; performance
\end{itemize}

### 2016-10-06 paper/sections/Library.tex
Latex draft.

\section{Modular Parsing Library using Delegation}\label{sec:library}

\begin{itemize}
\item Fixpoints library + explaining delegation with some examples
\item Alternative combinator + others
\item Trait Composition to do Language Composition
\end{itemize}

### 2016-10-06 paper/sections/Overview.tex
Latex draft.

\section{Overview}\label{sec:overview}

\begin{itemize}
\item Choosing the Parsing Technology
    \begin{itemize}
    \item Parser Generators : why not?
        \begin{itemize}
            \item Not type-safe
            \item No modular type-checking
            \item Not modular or no separate compilation (but we need to mention lots of work on extensible parsing here: example Language Workbenches; Rats)
        \end{itemize}
    \item Parser Combinators
        \begin{itemize}
        \item Backtracking parsers (Parsec)
            \begin{itemize}
            \item Need to remove left recursion (Problem: Transformation is not modular; if we do not know the full grammar then cannot be done)
            \item Need try/backtracking (Problem: Since we do not know the full set of rules in a modular setting, we have to assume worst case scenario and add redundant backtracking. )
            \item Mention some possible workarounds (there may be some but they still have issues)
            \end{itemize}
        \item (Non)-Backtracking/Packrat Parsers (Works well in a modular setting) (find a good name for this type of parsers?)
        \end{itemize}
    \end{itemize}
\item Adapting Parser Combinators for modularity
    \begin{itemize}
    \item Using Object Algebras
    \item Using Open Recursion
    \item Using new modular parser combinators (Library: new alternative combinator, for example)
    \end{itemize}
\item Small example
    \begin{itemize}
    \item small example of a modular parser using our technique
    \item show also the same example without modularity and write a detailed comparison
    \item Lambda Calculus is a good candidate for the example
    \item Show a small extension (adding plus and numeric literals)
    \end{itemize}
\end{itemize}

### 2016-10-06 paper/computePositions.rb

@@ -1,0 +1,69 @@
+#require 'tempfile'
+require 'fileutils'
+
+# Example 1 - Read File and close
+def scan(fileName, beginpos, endpos, fileloc)
+  file = File.new(fileName, "r")
+  counter = 1
+  while (line = file.gets)
+    if line =~ /BEGIN_(\w+)/
+      throw "Symbol '#{$1}' already defined" if fileloc[$1]
+      fileloc[$1] = fileName
+      beginpos[$1] = counter+1
+      puts "BEGIN #{$1}: #{beginpos[$1]}"
+    end
+    if line =~ /END_(\w+)/
+      endpos[$1] = counter-1
+      puts "END #{$1}: #{endpos[$1]}"
+    end
+    counter = counter + 1
+  end
+  file.close
+end
+
+def process(file, out, beginpos, endpos, fileloc)
+  while (line = file.gets)
+    if line =~ /APPLY:(\w+)=(\w+)/
+      var = $1
+      label = $2
+      line = line.sub( /#{var}=([0-9-]*)/, "#{var}=#{beginpos[label]}-#{endpos[label]}" )
+      line = line.sub( /\{[^{]*\}/, "{#{fileloc[label]}}" )
+    end
+    out.write(line)
+  end
+  file.close
+  out.close
+end
+
+fileloc = {}
+beginpos = {}
+endpos = {}
+
+Dir['../UseMixinLombok/src/*/*.java'].each do |file| 
+  scan(file, beginpos, endpos, fileloc)
+end
+
+Dir['../UseMixinLombok/src/*/*/*.java'].each do |file| 
+  scan(file, beginpos, endpos, fileloc)
+end
+
+
+
+Dir['*.tex'].each do |file| 
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+
+Dir['*/*.tex'].each do |file| 
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+

### 2016-10-06 paper/run.sh

@@ -1,0 +1,27 @@
+#!/bin/zsh
+
+if [ $1 ]
+then
+    case "$1" in
+    bib )       
+        xelatex paper.tex
+        bibtex paper     
+    ;;
+    * )
+        echo "Usage (In terminal with zsh):"
+        echo "\t\t./run.sh"
+        echo "render and open a new pdf"
+        echo "\t\t./run.sh bib"
+        echo "render and open a new pdf with bib changes"
+    esac   
+fi
+
+ruby computePositions.rb
+xelatex paper.tex
+open paper.pdf
+
+rm footempfile.txt
+rm *.bbl
+rm *.blg
+rm *.log
+rm *-old

### 2016-10-11 paper/sections/Library.tex
Some text for Section 3.

@@
-\section{Modular Parsing Library using Delegation}\label{sec:library}
+\section{Modular Parsing Library}\label{sec:library}
 
-\begin{itemize}
-\item Fixpoints library + explaining delegation with some examples
-\item Alternative combinator + others
-\item Trait Composition to do Language Composition
-\end{itemize}
+%\begin{itemize}
+%\item Fixpoints library + explaining delegation with some examples
+%\item Alternative combinator + others
+%\item Trait Composition to do Language Composition
+%\end{itemize}
+
+In this section, we will further look into the mechanism behind our modular parsing library.
+Starting from simple examples, a series of extensions and improvements are applied to illustrate
+how we achieve modularity in a type-safe way.
+
+\subsection{Open Recursion and Delegation}\label{subsec:openrecursion}
+
+Our story begins with a single parser that only parses a free variable (i.e., a string). In Scala we use the Packrat Parsing
+library, which supports lexing and direct left-recursion. To write such a parser, we need to import the library as
+well as defining the datatype for which it produces.
+
+\begin{lstlisting}
+object OpenRecursion extends StandardTokenParsers with PackratParsers {
+    class Expr
+    class Var(x : String) extends Expr
+    val pVar : PackratParser[Expr] = ident ^^ { x => new Var(x) }
+    ...
+}
+\end{lstlisting}
+Note that we are omitting lexing and demo code here. Furthermore, we would like the parser to support applications.
+Not only the implementation of that parser is needed, but also its corresponding datatype: (we omit the fact that the code for below few
+examples is written in the same object, namely \lstinline{OpenRecursion})
+\begin{lstlisting}
+class App(e1 : Expr, e2 : Expr) extends Expr
+val pApp : PackratParser[Expr] = pVar ~ pVar ^^ { case e1 ~ e2 => new App(e1, e2) }
+\end{lstlisting}
+Such a parser \lstinline{pApp} can only parse one-layer applications like \lstinline{"x y"}. Hence we revise it using recursion:
+\begin{lstlisting}
+val pApp : PackratParser[Expr] =
+    (pApp | pVar) ~ (pApp | pVar) ^^ { case e1 ~ e2 => new App(e1, e2) }
+\end{lstlisting}
+Now \lstinline{pApp} manages to parse an arbitrary number of applications like \lstinline{"x y z ..."} in a right-associative way.
+Nevertheless, such an approach does not turn out to be a modular extension. The grammar we want for an expression is actually:
+\begin{lstlisting}
+e ::= x | e ' ' e | ...
+\end{lstlisting}
+Note that such an \lstinline{"e"} in the second case should not only include all existing cases, but also take all future extensions into account (the ellipsis we used above, for instance, \lstinline{e '+'} \lstinline{e}).
+This inspires us to use \textit{open recursion}.
+To define \lstinline{pApp} in a different way, we add a parameter \lstinline{p} to it, representing the explicit ``self reference'' of our whole parser, which is open to future extensions.
+Note that \lstinline{p} should be defined as a by-name parameter using \lstinline{"=>"}.
+\begin{lstlisting}
+val pApp : (=> PackratParser[Expr]) => PackratParser[Expr] =
+    p => p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
+\end{lstlisting}
+Similarly we redefine \lstinline{pVar} also as a function using the fixpoint \lstinline{p}.
+Furthermore, we combine \lstinline{pVar} and \lstinline{pApp} together using the parser combinator \lstinline{"|"} for alternative:
+ \begin{lstlisting}
+val pVar : (=> PackratParser[Expr]) => PackratParser[Expr] =
+    p => ident ^^ { x => new Var(x) }
+val pVarApp : (=> PackratParser[Expr]) => PackratParser[Expr] =
+    p => pApp(p) | pVar(p)
+\end{lstlisting}
+But what we really want is something that has type \lstinline{PackratParser[Expr]}.
+In functional programming, we can use ``fix'' to get the fixpoint of a function, in this case we write \lstinline{fix(pVarApp)}, which parses an arbitrary number of applications.
+Such a function can easily be defined
+in Scala based on its laziness:
+\begin{lstlisting}
+def fix[A](f : (=> A) => A) : A = { lazy val a : A = f(a); a }
+\end{lstlisting}
+That is the magic of open recursion: we can define as many small components as we like, and they are implemented as functions, with the help of explicit self-reference.
+Whenever we would like to close the recursion, simply use \lstinline{fix} for their combination.
+Such a process is guaranteed to terminate when we restrict the input for parsing to be finite, and we have some base cases like \lstinline{pVar}, which terminates recursion in their branches, in addition, with the help of direct left-recursion support from Packrat parsers.
+
+Another thing which catches our attention is that, small parsers are combined using the alternative combinator.
+It is a matter of delegation: parsing failing at any position will try its next alternative, and for each recursion all the parsers are tried successively until one of them succeeds.
+It reveals its essence as a recursive descent parser.
+But it is quite easy and modular to use; with further extensions, they can be implemented separately and then appended to the combination of old ones.
+
+\subsection{Parser Combinators}\label{subsec:parsercombinators}
+
+Using Packrat parsers we obtain the access to a lot of library functions, including parser combinators.
+In the last subsection we introduced \lstinline{|} for alternation, which realizes delegation pattern in our parsing.
+Moreover, we have \lstinline{~} for sequential composition, \lstinline{rep} and \lstinline{repsep} for repetitions, \lstinline{chainl1} for left-associative grammars, and so on.
+
+Specifically, Scala parsers have a new combinator \lstinline{|||} also for alternative but with longest match composition.
+Consider the example in the last subsection, where we used \lstinline{pApp(p)} \lstinline{|} \lstinline{pVar(p)} for alternation.
+If \lstinline{pVar} is put in the front instead, the parser cannot even parse \lstinline{"x y"}, since the first case succeeds, which unexpectedly stops parsing and discards the rest.
+But \lstinline{|||} can successfully parse the whole input string.
+It implies that using \lstinline{|||} we no longer need to worry about the order of composition, which encourages its extensibility.
+
+On the other hand, practically parsers are defined as functions (as shown above), which means we do not apply \lstinline{|||} to them directly, but use \lstinline{|||} implement a new combinator for those functions.
+We will introduce it later in this paper.
+
+\subsection{Object Algebras for Different Syntax}\label{subsec:differentsyntax}
+
+Such extensibility becomes fragile when it comes to grammars with different syntax. Now suppose we would like to have types in the grammar. Firstly the new grammar is shown below:
+\begin{lstlisting}
+e ::= x | e ' ' e | \x:t.e
+t ::= "Int"
+\end{lstlisting}
+
+We proceed to implement those parsers. An issue is that we need both a \lstinline{PackratParser[Expr]} and a \lstinline{PackratParser[Type]} in the fixpoint.
+Hence we use a pair in the parameter:
+\begin{lstlisting}
+...
+\end{lstlisting}
+
+% more subsections: object algebras (different syntax, separate variant and behavior, two dimensions of extensibility, generics), lexing, traits, language composition (more extensibility)

### 2016-10-11 Scala/Parser/src/SampleCode/OpenRecursion.scala

@@ -1,0 +1,64 @@
+package Test
+
+import scala.util.parsing.combinator._
+import scala.util.parsing.combinator.syntactical._
+
+object OpenRecursion extends StandardTokenParsers with PackratParsers {
+  
+  class Expr {
+    override val toString = ""
+  }
+  
+  class Var(x : String) extends Expr {
+    override val toString = x
+  }
+  val pVar : (=> PackratParser[Expr]) => PackratParser[Expr] = p =>
+    ident ^^ { x => new Var(x) }
+  
+  class App(e1 : Expr, e2 : Expr) extends Expr {
+    override val toString = e1.toString + " " + e2.toString
+  }
+  val pApp : (=> PackratParser[Expr]) => PackratParser[Expr] = p =>
+    p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
+
+  class Type {
+    override val toString = ""
+  }
+  
+  class Int extends Type {
+    override val toString = "Int"
+  }
+  val pInt : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Type] = p =>
+    "Int" ^^ { _ => new Int() }
+    
+  class Lam(x : String, t : Type, e : Expr) extends Expr {
+    override val toString = "\\" + x + ":" + t.toString + "." + e.toString
+  }
+  val pLam : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Expr] = p => {
+    val pE = p._1
+    val pT = p._2
+    ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ { case x ~ t ~ e => new Lam(x, t, e) }
+  }
+  
+  val p : (=> (PackratParser[Expr], PackratParser[Type])) => (PackratParser[Expr], PackratParser[Type])
+    = p => (pVar(p._1) | pApp(p._1) | pLam(p), pInt(p))
+  
+  lexical.reserved += ("Int")
+  lexical.delimiters += ("\\", ":", ".")
+  def runParser(p : Parser[_]) : String => Unit = in => {
+    phrase(p)(new lexical.Scanner(in)) match {
+      case t if t.successful => println(t.get.toString())
+      case t                 => scala.sys.error(t.toString)
+    }
+  }
+  
+  def fix[A](f: (=> A) => A): A = {
+    lazy val a: A = f(a)
+    a
+  }
+  
+  def main(args : Array[String]) = {
+    runParser(fix(p)._1)("\\x:Int.x x")
+  }
+
+}

### 2016-10-11 Scala/Parser/src/SampleCode/Test.scala

@@ -1,0 +1,41 @@
+package Test
+
+import scala.util.parsing.combinator._
+import scala.util.parsing.combinator.syntactical._
+
+object Test extends StandardTokenParsers with PackratParsers {
+  
+  class Expr {
+    override val toString = ""
+  }
+  
+  class Var(x : String) extends Expr {override val toString = x}
+  val pVar : (=> PackratParser[Expr]) => PackratParser[Expr] =
+    p => ident ^^ { x => new Var(x) }
+  
+  class App(e1 : Expr, e2 : Expr) extends Expr {
+    override val toString = "[" + e1.toString + " " + e2.toString + "]"
+  }
+  val pApp : (=> PackratParser[Expr]) => PackratParser[Expr] =
+    p => p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
+    
+  val pVarApp : (=> PackratParser[Expr]) => PackratParser[Expr] =
+    p => pApp(p) | pVar(p) 
+
+  def runParser(p : Parser[_]) : String => Unit = in => {
+    phrase(p)(new lexical.Scanner(in)) match {
+      case t if t.successful => println(t.get.toString())
+      case t                 => scala.sys.error(t.toString)
+    }
+  }
+  
+  def fix[A](f: (=> A) => A): A = {
+    lazy val a: A = f(a)
+    a
+  }
+  
+  def main(args : Array[String]) = {
+    runParser(fix(pVarApp))("x y")
+  }
+
+}

### 2016-10-11 paper/sections/Library.tex
More text.

@@
 
-Such extensibility becomes fragile when it comes to grammars with different syntax. Now suppose we would like to have types in the grammar. Firstly the new grammar is shown below:
+Such extensibility becomes fragile when it comes to grammars with different syntax.
+Now suppose we would like to have types in the grammar.
+Firstly the new grammar is shown below, where the \lstinline{Int} type and lambda expression with annotation have been added:
 \begin{lstlisting}
@@
 \begin{lstlisting}
-...
+class Type
+class Int extends Type
+val pInt : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Type] =
+    p => "Int" ^^ { _ => new Int() }
+
+class Lam(x : String, t : Type, e : Expr) extends Expr
+val pLam : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Expr] =
+    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1)
+        ^^ { case x ~ t ~ e => new Lam(x, t, e) }
+
+val p : (=> (PackratParser[Expr], PackratParser[Type])) => (PackratParser[Expr], PackratParser[Type]) =
+    p => (pVar(p._1) ||| pApp(p._1) ||| pLam(p), pInt(p))
 \end{lstlisting}
+The last value \lstinline{p} integrates the four single parsers defined before, so we use \lstinline{fix(p)} to
+obtain the two parsers we want: one for expressions, the other for types. The parser \lstinline{fix(p)._1} parsers expressions
+like \lstinline{"\x:Int.x x"} successfully. Yet meanwhile, some drawbacks of this approach come into sight. With different syntax, it
+is necessary to extend the fixpoint to contain more parsers, in which case usually some structures like tuples are used,
+making code tedious and less elegant. On the other hand, the modularity is quite restricted; only parsing and datatypes are extensible.
+Suppose now we want to realize some behaviors on those data structures, for example pretty-printing, then whenever we introduce a new behavior,
+it needs implementations in every class body, requiring existing code to be modified. At this point, Object Algebras are adopted for data structures
+instead.
 
-% more subsections: object algebras (different syntax, separate variant and behavior, two dimensions of extensibility, generics), lexing, traits, language composition (more extensibility)
+Object Algebras are a nice solution to the famous \textit{Expression Problem}, achieving two dimensions of extensibility (data variants and operations)
+in a type-safe way.
+The intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design.
+We can simply add new variants for the new cases in the grammar, or realize different operations on structures, without modifying existing code, and separate compilation is supported.
+Coincidentally, Object Algebras use generics, where the type parameters perfectly support extensible different syntax, and hence simplify the code comparatively.
+Below is the code using Object Algebras, where we start from \lstinline{varE} and \lstinline{appE} only, later we extend them with \lstinline{intT} and \lstinline{lamE}:
+\begin{lstlisting}
+trait OldAlg[E] {
+    def varE(x : String) : E
+    def appE(e1 : E, e2 : E) : E
+}
+
+trait OldParser[E] {
+    val pE : OldAlg[E] => (=> PackratParser[E]) => PackratParser[E] = alg => p =>
+        ident ^^ { x => alg.varE(x) } |||
+        p ~ p ^^ { case e1 ~ e2 => alg.appE(e1, e2) }
+}
+\end{lstlisting}
+Here \lstinline{OldAlg} is called an \textit{Object Algebra interface}, which has two constructs for the AST. Its parser is
+supposed to produce a structure from an input, whereas with Object Algebras structures are implicit: an \textit{algebra}
+is immediately applied to the structure and returns the result. It means such a structure that the parser generates should have
+type \lstinline{"OldAlg[E] =>} \lstinline{E"} for some \lstinline{E}. In the above code, \lstinline{alg} which has type
+\lstinline{OldAlg[E]} is an algebra, and appears as a parameter of \lstinline{pE}, then in the body of \lstinline{pE} the algebra
+is invoked correspondingly right after parsing. Since it should be a generic algebra, the parser is enclosed by the generic trait \lstinline{OldParser}.
+
+% more: newparser, two pretty-printers OldPretty and NewPretty, they are separated to show modularity.
+% more: modify the code using List[E] and type bounds.
+% more subsections: lexing, traits, operations, language composition (more extensibility)

### 2016-10-12 paper/paper.tex
Add text for general open recursion.

@@
 title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
-escapeinside={(*}{*)},          % if you want to add a comment within your code
+escapeinside={(**}{**)},          % if you want to add a comment within your code
 keywordstyle=\ttfamily\bfseries,
@@
 
-\input{sections/Library.tex}
+\input{sections/Library/OpenRecursion.tex}
+\input{sections/Library/ObjectAlgebras.tex}
+\input{sections/Library/FirstTry.tex}
+\input{sections/Library/ParserCombinators.tex}
+\input{sections/Library/DifferentSyntax.tex}
+
+% more: two pretty-printers OldPretty and NewPretty, they are separated to show modularity.
+% subsection: modify the code using List[E] and type bounds.
+% more subsections: lexing, traits, operations, language composition (more extensibility)
 

### 2016-10-12 paper/sections/Library/DifferentSyntax.tex
Add text for general open recursion.

--- 2016-10-06 paper/sections/Library.tex
@@
-\section{Modular Parsing Library using Delegation}\label{sec:library}
+\subsection{Object Algebras for Different Syntax}\label{subsec:differentsyntax}
 
-\begin{itemize}
-\item Fixpoints library + explaining delegation with some examples
-\item Alternative combinator + others
-\item Trait Composition to do Language Composition
-\end{itemize}
+Such extensibility becomes fragile when it comes to grammars with different syntax.
+Now suppose we would like to have types in the grammar.
+Firstly the new grammar is shown below, where the \lstinline{Int} type and lambda expressions with annotation have been added:
+\begin{lstlisting}
+e ::= x | e ' ' e | \x:t.e
+t ::= "Int"
+\end{lstlisting}
+
+We proceed to implement those parsers. An issue is that we need both a \lstinline{PackratParser[Expr]} and a \lstinline{PackratParser[Type]} in the fixpoint.
+Hence we use a pair in the parameter:
+\begin{lstlisting}
+class Type
+class Int extends Type
+val pInt : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Type] =
+    p => "Int" ^^ { _ => new Int() }
+
+class Lam(x : String, t : Type, e : Expr) extends Expr
+val pLam : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Expr] =
+    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1)
+        ^^ { case x ~ t ~ e => new Lam(x, t, e) }
+
+val p : (=> (PackratParser[Expr], PackratParser[Type])) => (PackratParser[Expr], PackratParser[Type]) =
+    p => (pVar(p._1) ||| pApp(p._1) ||| pLam(p), pInt(p))
+\end{lstlisting}
+The last value \lstinline{p} integrates the four single parsers defined before, so we use \lstinline{fix(p)} to
+obtain the two parsers we want: one for expressions, the other for types. The parser \lstinline{fix(p)._1} parsers expressions
+like \lstinline{"\x:Int.x x"} successfully. Yet meanwhile, some drawbacks of this approach come into sight. With different syntax, it
+is necessary to extend the fixpoint to contain more parsers, in which case usually some structures like tuples are used,
+making code tedious and less elegant. On the other hand, the modularity is quite restricted; only parsing and datatypes are extensible.
+Suppose now we want to realize some behaviors on those data structures, for example pretty-printing, then whenever we introduce a new behavior,
+it needs implementations in every class body, requiring existing code to be modified. At this point, Object Algebras are adopted for data structures
+instead.
+
+Object Algebras are a nice solution to the famous \textit{Expression Problem}, achieving two dimensions of extensibility (data variants and operations)
+in a type-safe way.
+The intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design.
+We can simply add new variants for the new cases in the grammar, or realize different operations on structures, without modifying existing code, and separate compilation is supported.
+Coincidentally, Object Algebras use generics, where the type parameters perfectly support extensible different syntax, and hence simplify the code comparatively.
+Below is the code using Object Algebras, where we start from \lstinline{varE} and \lstinline{appE} only, later we will extend them with \lstinline{intT} and \lstinline{lamE}:
+\begin{lstlisting}
+trait OldAlg[E] {
+    def varE(x : String) : E
+    def appE(e1 : E, e2 : E) : E
+}
+
+trait OldParser[E] {
+    val pE : OldAlg[E] => (=> PackratParser[E]) => PackratParser[E] = alg => p =>
+        ident ^^ { x => alg.varE(x) } |||
+        p ~ p ^^ { case e1 ~ e2 => alg.appE(e1, e2) }
+}
+\end{lstlisting}
+Here \lstinline{OldAlg} is called an \textit{Object Algebra interface}, which has two constructs for the AST. Its parser is
+supposed to produce a structure from an input, whereas with Object Algebras structures are implicit: an \textit{algebra}
+is immediately applied to the structure and returns the result. It means such a structure that the parser generates should have
+type \lstinline{"OldAlg[E] =>} \lstinline{E"} for some \lstinline{E}. In the above code, \lstinline{alg} which has type
+\lstinline{OldAlg[E]} is an algebra, and appears as a parameter of \lstinline{pE}, then in the body of \lstinline{pE} the algebra
+is invoked correspondingly right after parsing. Since it should be a generic algebra, the parser is enclosed by the generic trait \lstinline{OldParser}.
+
+Now we move on to declare \lstinline{intT} and \lstinline{lamE} in a new trait:
+\begin{lstlisting}
+trait NewAlg[E, T] {
+    def intT() : T
+    def lamE(x : String, t : T, e : E) : E
+}
+
+trait NewParser[E, T] {
+    val pE : NewAlg[E, T] => (=> (PackratParser[E], PackratParser[T]))
+            => PackratParser[E] = alg => p =>
+        ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1)
+            ^^ { case x ~ t ~ e => alg.lamE(x, t, e) }
+    val pT : NewAlg[E, T] => (=> (PackratParser[E], PackratParser[T]))
+            => PackratParser[T] = alg => p =>
+        "Int" ^^ { _ => alg.intT() }
+}
+\end{lstlisting}
+\lstinline{NewParser} contains parsers for both expressions and types. From the code we observe that the multi-sorts of Object Algebra interface
+exactly express different syntax. At this point, to combine these two small components together we need some ``glue code''. The two ASTs \lstinline{OldAlg} and
+\lstinline{NewAlg} should be put together, but instead of defining a new trait, it is more convenient to directly use compound types in the parameter.
+\begin{lstlisting}
+trait OldNewParser[E, T] {
+    val p : (OldAlg[E] with NewAlg[E, T]) => (=> (PackratParser[E], PackratParser[T]))
+            => (PackratParser[E], PackratParser[T]) = alg => p => {
+        val pOld = new OldParser[E](){}
+        val pNew = new NewParser[E, T](){}
+        (pOld.pE(alg)(p._1) ||| pNew.pE(alg)(p), pNew.pT(alg)(p))
+    }
+  }
+\end{lstlisting}
+Parsers for expressions and types are connected using alternative, respectively. Hence the isolation for different syntax
+is guaranteed by the type system.

### 2016-10-12 paper/sections/Library/FirstTry.tex
Add text for general open recursion.

--- 2016-10-11 paper/sections/Library.tex
@@
-\section{Modular Parsing Library}\label{sec:library}
-
-%\begin{itemize}
-%\item Fixpoints library + explaining delegation with some examples
-%\item Alternative combinator + others
-%\item Trait Composition to do Language Composition
-%\end{itemize}
-
-In this section, we will further look into the mechanism behind our modular parsing library.
-Starting from simple examples, a series of extensions and improvements are applied to illustrate
-how we achieve modularity in a type-safe way.
-
-\subsection{Open Recursion and Delegation}\label{subsec:openrecursion}
+\subsection{First Try on Parsing}\label{sec:firsttry}
 
@@
 But it is quite easy and modular to use; with further extensions, they can be implemented separately and then appended to the combination of old ones.
-
-\subsection{Parser Combinators}\label{subsec:parsercombinators}
-
-Using Packrat parsers we obtain the access to a lot of library functions, including parser combinators.
-In the last subsection we introduced \lstinline{|} for alternation, which realizes delegation pattern in our parsing.
-Moreover, we have \lstinline{~} for sequential composition, \lstinline{rep} and \lstinline{repsep} for repetitions, \lstinline{chainl1} for left-associative grammars, and so on.
-
-Specifically, Scala parsers have a new combinator \lstinline{|||} also for alternative but with longest match composition.
-Consider the example in the last subsection, where we used \lstinline{pApp(p)} \lstinline{|} \lstinline{pVar(p)} for alternation.
-If \lstinline{pVar} is put in the front instead, the parser cannot even parse \lstinline{"x y"}, since the first case succeeds, which unexpectedly stops parsing and discards the rest.
-But \lstinline{|||} can successfully parse the whole input string.
-It implies that using \lstinline{|||} we no longer need to worry about the order of composition, which encourages its extensibility.
-
-On the other hand, practically parsers are defined as functions (as shown above), which means we do not apply \lstinline{|||} to them directly, but use \lstinline{|||} implement a new combinator for those functions.
-We will introduce it later in this paper.
-
-\subsection{Object Algebras for Different Syntax}\label{subsec:differentsyntax}
-
-Such extensibility becomes fragile when it comes to grammars with different syntax. Now suppose we would like to have types in the grammar. Firstly the new grammar is shown below:
-\begin{lstlisting}
-e ::= x | e ' ' e | \x:t.e
-t ::= "Int"
-\end{lstlisting}
-
-We proceed to implement those parsers. An issue is that we need both a \lstinline{PackratParser[Expr]} and a \lstinline{PackratParser[Type]} in the fixpoint.
-Hence we use a pair in the parameter:
-\begin{lstlisting}
-...
-\end{lstlisting}
-
-% more subsections: object algebras (different syntax, separate variant and behavior, two dimensions of extensibility, generics), lexing, traits, language composition (more extensibility)

### 2016-10-12 paper/sections/Library/ObjectAlgebras.tex
Add text for general open recursion.

\subsection{Object Algebras}\label{sec:objectalgebras}

### 2016-10-12 paper/sections/Library/OpenRecursion.tex
Add text for general open recursion.

\section{Modular Parsing Library}\label{sec:library}

%\begin{itemize}
%\item Fixpoints library + explaining delegation with some examples
%\item Alternative combinator + others
%\item Trait Composition to do Language Composition
%\end{itemize}

In this section, we will further look into the mechanism behind our modular parsing library.
We will first introduce some concepts independently from the process of parsing, including open recursion and Object Algebras,
then present how they are integrated in our library.
Starting from simple examples, a series of extensions and refactorings will be applied to illustrate
how we achieve modularity in a type-safe way.

\subsection{Open Recursion and Delegation}\label{sec:openrecursion}

Open recursion is known as a useful feature that one method body can ``invoke another method of the same object via a special variable called \lstinline{self} or, in some languages, \lstinline[keywords={}]{this}'' (by Ralf Hinze). The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, whereas the method that takes \lstinline{self} as a parameter can simply delegate its known cases to it.

For sure we cannot always leave it open; upon using it we have to close the recursion, and at that time we need to take a fix-point. Here one might suddenly be reminded of the \lstinline{fix} function in some functional languages like Haskell. Below is the famous Fibonacci example:
\begin{lstlisting}[language=Haskell,keywords={}]
fib :: Int -> Int
fib 0 = 0
fib 1 = 1
fib n = fib (n - 1) + fib (n - 2)
\end{lstlisting}
An alternative approach using open recursion is as follows:
\begin{lstlisting}[language=Haskell,keywords={type}]
type Fix t = t -> t

fib' :: Fix (Int -> Int)
fib' self 0 = 0
fib' self 1 = 1
fib' self n = self (n - 1) + self (n - 2)
\end{lstlisting}
where \lstinline{self} is the explicit self-reference, namely the delegator. Functions like \lstinline{fib'} which takes
an explicit self-reference as its parameter are called \textit{generators} (by Ralf Hinze). The definition of \lstinline{fix} is:
\begin{lstlisting}[language=Haskell,keywords={}]
fix f = let x = f x in x
\end{lstlisting}
If we close the recursion on \lstinline{fib'} only, the evaluation of \lstinline{fix fib'} \lstinline{2} is as follows:
\begin{lstlisting}[language=Haskell,keywords={}]
   fix fib' 2
= let x = fib' x in x 2
= let x = fib' x in fib' x 2
= let x = fib' x in (x 1) + (x 0)
= let x = fib' x in (fib' x 1) + (fib' x 0)
= let x = fib' x in 1 + 0
= 1
\end{lstlisting}
which behaves the same as \lstinline{fib 2}. The process of evaluation will terminate if there are some base cases (like the first two in
\lstinline{fib'}) which stop the recursion in their branches, and the evaluation is finally reduced to those cases.

Suppose now we want to make use of open recursion for additional operations. We first define a combinator which combines the results of two generators in a pair:
\begin{lstlisting}[language=Haskell,keywords={}]
(*) :: Fix (a -> b) -> Fix (a -> c) -> Fix (a -> (b, c))
f * g = \self x -> (f (fst . self) x, g (snd . self) x)
\end{lstlisting}
Then we define a generator which prints out a number as a string:
\begin{lstlisting}[language=Haskell,keywords={}]
show' :: Fix (Int -> String)
show' self x = show x
\end{lstlisting}
Now \lstinline{fix (fib'}\lstinline{ * show')} \lstinline{2} results in \lstinline{(1, "2")}. We have seen that both \lstinline{fib'} and \lstinline{show'} are integrated in the evaluation. In Scala, however, there is no \lstinline{fix} function, but it can be easily defined since it is based on lazy evaluation. See below:
\begin{lstlisting}
def fix[A](f : (=> A) => A) : A = {
    lazy val a : A = f(a)
    a
}
\end{lstlisting}
Next we define the general combinator \lstinline{merge} for open recursion. It is similar to the \lstinline{(*)} above, but instead of putting two values in a pair by default, \lstinline{merge} takes a parameter called \lstinline{op}, which tells the associative operation of two results.
\begin{lstlisting}
def merge[E, F, G, H[_,_]](op : F => G => H[F, G], x : (=> E) => F, y : (=> E) => G) : (=> E) => H[F, G] = e => op(x(e))(y(e))
\end{lstlisting}

### 2016-10-12 paper/sections/Library/ParserCombinators.tex
Add text for general open recursion.

--- 2016-10-06 paper/sections/Library.tex
@@
-\section{Modular Parsing Library using Delegation}\label{sec:library}
+\subsection{Parser Combinators}\label{subsec:parsercombinators}
 
-\begin{itemize}
-\item Fixpoints library + explaining delegation with some examples
-\item Alternative combinator + others
-\item Trait Composition to do Language Composition
-\end{itemize}
+Using Packrat parsers we obtain the access to a lot of library functions, including parser combinators.
+In the last subsection we introduced \lstinline{|} for alternation, which realizes delegation pattern in our parsing.
+Moreover, we have \lstinline{~} for sequential composition, \lstinline{rep} and \lstinline{repsep} for repetitions, \lstinline{chainl1} for left-associative grammars, and so on.
+
+Specifically, Scala parsers have a new combinator \lstinline{|||} also for alternative but with longest match composition.
+Consider the example in the last subsection, where we used \lstinline{pApp(p)} \lstinline{|} \lstinline{pVar(p)} for alternation.
+If \lstinline{pVar} is put in the front instead, the parser cannot even parse \lstinline{"x y"}, since the first case succeeds, which unexpectedly stops parsing and discards the rest.
+But \lstinline{|||} can successfully parse the whole input string.
+It implies that using \lstinline{|||} we no longer need to worry about the order of composition, which encourages its extensibility.
+
+On the other hand, practically parsers are defined as functions (as shown above), which means we do not apply \lstinline{|||} to them directly, but use \lstinline{|||} implement a new combinator for those functions.
+We will introduce it later in this paper.

### 2016-10-13 paper/sections/Library/DifferentSyntax.tex
More text.

@@
 Now suppose we would like to have types in the grammar.
-Firstly the new grammar is shown below, where the \lstinline{Int} type and lambda expressions with annotation have been added:
+Firstly the new grammar is shown below, where the \lstinline{IntT} type and lambda expressions with annotation have been added:
 \begin{lstlisting}
@@
 class Type
-class Int extends Type
+class IntT extends Type
 val pInt : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Type] =
-    p => "Int" ^^ { _ => new Int() }
+    p => "Int" ^^ { _ => new IntT() }
 

### 2016-10-13 paper/sections/Library/ObjectAlgebras.tex
More text.

@@
-\subsection{Object Algebras}\label{sec:objectalgebras}
+\subsection{Object Algebras}\label{subsec:objectalgebras}
+
+Object Algebras captures a design pattern to address the famous \textit{Expression Problem} nicely,
+achieving two dimensions of extensibility (data variants and operations) in a modular and type-safe way.
+Because of this, the definition of data structures is separated from behaviors on them, and future extensions
+to both sides no longer require existing code to be modified, supporting separate compilation.
+
+In Object Algebras, ASTs as recursive data structures are defined using traits, where each constructor corresponds
+to an abstract method inside. The recursive types are represented using generics. Here we show the example of a simple
+expression language:
+\begin{lstlisting}
+trait ExpAlg[E] {
+    def lit(i : Int) : E
+    def add(e1 : E, e2 : E) : E
+}
+\end{lstlisting}
+The language has two constructors: literals and additions. The trait \lstinline{ExpAlg} is called an \textit{Object Algebra interface},
+whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
+type parameter \lstinline{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
+provides implementations for all cases. Below is an example of evaluation:
+\begin{lstlisting}
+trait Eval extends ExpAlg[Int] {
+    def lit(i : Int) = i
+    def add(e1 : Int, e2 : Int) = e1 + e2
+}
+\end{lstlisting}
+Here \lstinline{Eval} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns an integer as the result.
+Furthermore, one can define a new trait to implement pretty-printing on expressions:
+\begin{lstlisting}
+trait Print extends ExpAlg[String] {
+    def lit(i : Int) = "" + i
+    def add(e1 : String, e2 : String) = e1 + " + " + e2
+}
+\end{lstlisting}
+On the other hand, the data variants can be extended by inheriting \lstinline{ExpAlg} and adding new cases only. Suppose we want to
+have subtraction in expressions, a new Object Algebra interface \lstinline{SubAlg} is defined as follows:
+\begin{lstlisting}
+trait SubAlg[E] extends ExpAlg[E] {
+    def sub(e1 : E, e2 : E) : E
+}
+\end{lstlisting}
+Now evaluation on the new language can be realized by code reuse, and without modifying existing code:
+\begin{lstlisting}
+trait SubEval extends Eval {
+    def sub(e1 : Int, e2 : Int) = e1 - e2
+}
+\end{lstlisting}
+On this, to create an expression of \lstinline{SubAlg}, a generic method is defined as follows:
+\begin{lstlisting}
+def build[E](alg : SubAlg[E]) : E =
+    alg.sub(alg.add(alg.lit(2), alg.lit(3)), alg.lit(4))
+\end{lstlisting}
+Such a method implicitly represents the expression \lstinline{"2 + 3 - 4"}. The code
+\begin{lstlisting}
+build(new Eval(){})
+\end{lstlisting}
+results in \lstinline{1}, as the result of evaluation. Now we observe that an ``object'' of expression actually
+has the function type \lstinline{SubAlg[E] =>} \lstinline{E} for generic \lstinline{E}. At this point, one may argue
+that it is illusory compared to traditional objects, and since we are struggling against the parsing problem, if a parser
+produces structures as functions, it could be hardly be further processed like objects, especially, in the design of a compiler
+there will be a lot of desugarings (or transformations). However, we argue that this is not a case, because functions can be used
+in a very clever way. For example, if one wants to refactor an \lstinline{ExpAlg} expression to a \lstinline{SubAlg} expression by
+replacing all negative numbers with subtraction, a transformation algebra can be implemented as follows:
+\begin{lstlisting}
+trait Refactor[E] extends ExpAlg[E] {
+    def alg : SubAlg[E]
+    def lit(i : Int) = if (i > 0) alg.lit(i) else alg.sub(alg.lit(0), alg.lit(-i))
+    def add(e1 : E, e2 : E) = alg.add(e1, e2)
+}
+\end{lstlisting}
+With \lstinline{Refactor}, the expression \lstinline{"1 + (-2)"} is refactored to \lstinline{"1 + (0 - 2)"}. The abstract method
+\lstinline{alg} in \lstinline{Refactor} can be any algebra of \lstinline{SubAlg}, simply for delegation. By passing a concrete implementation
+to \lstinline{alg}, \lstinline{Refactor} will be able to apply refactoring to an \lstinline{ExpAlg} expression.

### 2016-10-13 paper/sections/Library/OpenRecursion.tex
More text.

@@
 
-\subsection{Open Recursion and Delegation}\label{sec:openrecursion}
+\subsection{Open Recursion and Delegation}\label{subsec:openrecursion}
 
 Open recursion is known as a useful feature that one method body can ``invoke another method of the same object via a special variable called \lstinline{self} or, in some languages, \lstinline[keywords={}]{this}'' (by Ralf Hinze).
-The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, whereas the method that takes \lstinline{self} as a parameter can simply delegate its known cases to it.
+The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, whereas the method that takes \lstinline{self} as a parameter can simply delegate the known cases to it.
 
-For sure we cannot always leave it open; upon using it we have to close the recursion, and at that time we need to take a fix-point.
+For sure we cannot always leave it open; upon using it we have to close the recursion, at which time we need to take a fix-point.
 Here one might suddenly be reminded of the \lstinline{fix} function in some functional languages like Haskell.
@@
 It is similar to the \lstinline{(*)} above, but instead of putting two values in a pair by default, \lstinline{merge} takes a parameter called \lstinline{op}, which tells the associative operation of two results.
+The definition is straightforward:
 \begin{lstlisting}

### 2016-10-13 paper/sections/Library/ParsingLibrary.tex
More text.

\subsection{Parsing Library}\label{subsec:parsinglibrary}

### 2016-10-15 paper/sections/Library/ObjectAlgebras.tex
More text.

@@
 produces structures as functions, it could be hardly be further processed like objects, especially, in the design of a compiler
-there will be a lot of desugarings (or transformations). However, we argue that this is not a case, because functions can be used
+there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
 in a very clever way. For example, if one wants to refactor an \lstinline{ExpAlg} expression to a \lstinline{SubAlg} expression by
@@
 \lstinline{alg} in \lstinline{Refactor} can be any algebra of \lstinline{SubAlg}, simply for delegation. By passing a concrete implementation
-to \lstinline{alg}, \lstinline{Refactor} will be able to apply refactoring to an \lstinline{ExpAlg} expression.
+to \lstinline{alg}, \lstinline{Refactor} will be able to apply refactoring to an \lstinline{ExpAlg} expression before \lstinline{alg} returns the result.

### 2016-10-15 paper/sections/Library/OpenRecursion.tex
More text.

@@
 then present how they are integrated in our library.
-Starting from simple examples, a series of extensions and refactorings will be applied to illustrate
+Starting from simple examples, a series of extensions and refactoring will be applied to illustrate
 how we achieve modularity in a type-safe way.
@@
 Open recursion is known as a useful feature that one method body can ``invoke another method of the same object via a special variable called \lstinline{self} or, in some languages, \lstinline[keywords={}]{this}'' (by Ralf Hinze).
-The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, whereas the method that takes \lstinline{self} as a parameter can simply delegate the known cases to it.
+The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, while existing methods that take \lstinline{self} as a parameter can simply delegate their known cases to it.
 
@@
 \end{lstlisting}
-An alternative approach using open recursion is as follows:
+An alternative implementation using open recursion is as follows:
 \begin{lstlisting}[language=Haskell,keywords={type}]
@@
 \end{lstlisting}
-If we close the recursion on \lstinline{fib'} only, the evaluation of \lstinline{fix fib'} \lstinline{2} is as follows:
+If we close the recursion on \lstinline{fib'} only, the evaluation of \lstinline{fix fib'} \lstinline{2} is processed as follows:
 \begin{lstlisting}[language=Haskell,keywords={}]
@@
 We have seen that both \lstinline{fib'} and \lstinline{show'} are integrated in the evaluation.
-In Scala, however, there is no \lstinline{fix} function, but it can be easily defined since it is based on lazy evaluation.
+In Scala, however, there is no library \lstinline{fix} function, but it can be easily defined since it is based on lazy evaluation.
 See below:

### 2016-10-15 paper/sections/Library/ParsingLibrary.tex
More text.

@@
 \subsection{Parsing Library}\label{subsec:parsinglibrary}
+
+In the previous sections we have talked about open recursion and Object Algebras. These concepts will later be composed together for our
+initial goal, which is to build extensible parsers, but surely it cannot live without a parsing library. But we want to argue here that besides
+open recursion, which constructs the basis of extensibility in parsing, the other aspects are totally orthogonal, and hence can have their alternatives.
+For example, Object Algebras are helpful to build extensible data structures, whereas an alternative with the same function can still be used. And a better
+parsing library would introduce a lot of fancy features, nevertheless, it is independent from our key idea.
+
+In Scala we select Packrat Parsing for our prototype implementation.
+It is practical not only because of the performance, and its parser combinators, but also due to its support for direct left-recursion.
+From theoretical point of view, the algorithm behind Packrat parsers supports both direct and indirect left-recursion.
+The current version of the library is still buggy in some cases with indirect left-recursion, but we believe that they will fix it in the future, and with direct left-recursion it is already practical to use.
+
+Moreover, the Scala parsing library provides us with a number of useful parser combinators, with which the syntax for parsing looks concise and clear.
+Besides \lstinline{|} for alternation and \lstinline{~} for sequential composition, we also have \lstinline{rep} and \lstinline{repsep} for repetitions, \lstinline{chainl1} for left-associative grammar rules, and so on.
+Specifically, a new combinator \lstinline{|||} is available, also for alternative but with longest match composition.
+Imagine that we are composing a list of parsers using alternative \lstinline{|}, in an order.
+A parser in the front might be successful in parsing a substring of the input, but then it unexpectedly stops the parsing and discards the rest, while another parser might be able to parse the whole input.
+In that case, using \lstinline{|||} we no longer need to worry about the order of composition, which encourages its extensibility.
+A concrete example will be illustrated for this later.
+
+Generally, a Packrat parser has the type \lstinline{PackratParser[E]} for some \lstinline{E}.
+Whereas our modular parser has a different type, which implies that we will not apply \lstinline{|||} for composition directly, but use \lstinline{|||} to implement a new combinator for convenience.
+It will also be introduced afterwards.

### 2016-10-18 paper/sections/CaseStudy.tex
Text

@@
 \section{Case Studies}\label{sec:casestudy}
+
+\subsection{Overview}\label{subsec:overview}
+
+To demonstrate the utility of our modular parsing library, we implemented parsers of the first 18 calculi from the book \textit{Types and Programming Languages}.
+In this book, it introduces several calculi from simple ones to complex ones by gradually adding new features to syntax.
+It is suitable for our case study for mainly two reason.
+First, the calculi are arguably practical as we want to show our library could be used in real world development.
+One of the most complex calculi in our case study is System F (with several extensions such as records and variants), which could be the foundation of practical functional programming languages.
+Second, the evolution of calculi in the book reveals the advantages of modular representation of abstract syntax and modular parsing, which is the key functionality of our library.
+By decomposing those calculi into smaller components and reusing the common ones, we can obtain considerably code reuse in parsing with little performance penalty, as shown later.
+
+For better modularity and eliminate code duplication, we extract 21 reusable \textit{components} from syntax of all the calculi.
+Each component, which may contain several syntactical structures, is a reusable fragment representing a certain feature in the syntax.
+For example, \lstinline{VarApp} component represents the variable case and application case of expressions.
+In our case study, each component is represented by a Scala object which has three parts.
+\lstinline{Alg} is an object algebra interface for the abstract syntax.
+\lstinline{Print} is an object algebra implements that interface \lstinline{Alg}, for the pretty printing operation.
+\lstinline{Parser} is the modular parser for this piece of syntax.
+
+\begin{lstlisting}
+object VarApp {
+  trait Alg[E] {
+    def TmVar(x: String): E
+    def TmApp(e1: E, e2: E): E
+  }
+  trait Print extends Alg[String] {
+    def TmVar(x: String) = x
+    def TmApp(e1: String, e2: String) = "[" + e1 + " " + e2 + "]"
+  }
+  trait Parser[E, F <: {val pE : PackratParser[E]}] {
+    lexical.delimiters += ("(", ")")
+    val pVarAppE: Alg[E] => (=> F) => PackratParser[E] = alg => l => {
+      val e = l.pE
+      ident ^^ alg.TmVar |||
+      e ~ e ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
+      "(" ~> e <~ ")"
+    }
+  }
+}
+\end{lstlisting}
+
+Then each calculus could be composed directly from components and other calculi, if it has those common structures in the syntax.
+For example, the calculus \lstinline{Untyped}, representing the famous untyped lambda calculus, can be constructed from components \lstinline{VarApp} and \lstinline{UntypedAbs}.
+
+\begin{lstlisting}
+object UntypedAbs {
+  trait Alg[E] {
+    def TmAbs(x: String, e: E): E
+  }
+  trait Print extends Alg[String] {
+    def TmAbs(x: String, e: String) = "\\" + x + "." + e
+  }
+  trait Parser[E, F <: {val pE : PackratParser[E]}] {
+    val pUntypedAbsE: Alg[E] => (=> F) => PackratParser[E] = {...}
+  }
+}
+
+object Untyped {
+  trait Alg[E] extends UntypedAbs.Alg[E] with VarApp.Alg[E]
+  trait Print extends Alg[String] with UntypedAbs.Print with VarApp.Print
+  trait Parser[E, L <: {val pE : PackratParser[E]}] extends UntypedAbs.Parser[E, L] with VarApp.Parser[E, L] {
+    val pUntypedE = pUntypedAbsE | pVarAppE
+  }
+}
+\end{lstlisting}
+
+\subsection{Extensibility}\label{subsec:extensibility}
+
+\subsection{Comparison}\label{subsec:comparison}
 

### 2016-10-18 paper/paper.tex
More text.
Text

@@
     super,this,throw,trait,true,try,%
-    type,var,while,yield},
+    type,var,while,yield,with,val},
   otherkeywords={=>,<-,<\%,<:,>:,\#,@},
@@
 \input{sections/Library/ObjectAlgebras.tex}
-\input{sections/Library/FirstTry.tex}
-\input{sections/Library/ParserCombinators.tex}
+\input{sections/Library/ParsingLibrary.tex}
+\input{sections/Library/FirstTry.tex} % mention again | and ||| with concrete examples. (pApp | pVar)
 \input{sections/Library/DifferentSyntax.tex}
 
+% Object Algebras
+% type synonyms, Parser[Expr], Fix[..]
 % more: two pretty-printers OldPretty and NewPretty, they are separated to show modularity.

### 2016-10-18 paper/sections/CaseStudy.tex
Text

@@
 
-To demonstrate the utility of our modular parsing library, we implemented parsers of the first 18 calculi from the book \textit{Types and Programming Languages}.
-In this book, it introduces several calculi from simple ones to complex ones by gradually adding new features to syntax.
+To demonstrate the utility of our modular parsing library, we implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages}.
+In the book it introduces several calculi from simple to complex, by gradually adding new features to syntax.
 It is suitable for our case study for mainly two reason.
@@
 \lstinline{Print} is an object algebra implements that interface \lstinline{Alg}, for the pretty printing operation.
+It is not necessary for parsing, but we include it here for demonstration.
 \lstinline{Parser} is the modular parser for this piece of syntax.
@@
 
-Then each calculus could be composed directly from components and other calculi, if it has those common structures in the syntax.
+We use some naming conventions in our code, \lstinline{E} represents expressions, \lstinline{T} represents types and \lstinline{K} represents kinds.
+In this component \lstinline{VarApp} we only have expressions.
+The \lstinline{Alg} and \lstinline{Print} are standard object algebra operations, thus we will focus on the modular parser \lstinline{Parser}.
+Each parser carries lexical information and parsing functions.
+The lexical information includes reserved words, delimiters, etc.
+Parsing functions are similar with we described before, except that we use record types under some subtyping constraints here for the explicit self-reference.
+The record type contains a parser for expressions \lstinline{pE}, a parser for types \lstinline{pT} and a parser for kinds \lstinline{pK}, if the corresponding one is needed when parsing the syntax of this component.
+Here the type of self-reference, namely \lstinline{F}, is only required to have a parser of expressions \lstinline{pE}.
+Then we can build the parsing function \lstinline{pVarAppE}, the suffix \lstinline{E} also denotes that it will parse an expression.
+
+Each calculus could be composed directly from components and other calculi, if it has those common structures in the syntax.
 For example, the calculus \lstinline{Untyped}, representing the famous untyped lambda calculus, can be constructed from components \lstinline{VarApp} and \lstinline{UntypedAbs}.

### 2016-10-18 paper/sections/Library/DifferentSyntax.tex
Refactor code in paper.

@@
 
-We proceed to implement those parsers. An issue is that we need both a \lstinline{PackratParser[Expr]} and a \lstinline{PackratParser[Type]} in the fixpoint.
+We proceed to implement those parsers. An issue is that we need both a \lstinline{Parser[Expr]} and a \lstinline{Parser[Type]} in the fixpoint.
 Hence we use a pair in the parameter:
 \begin{lstlisting}
+type Pair = (Parser[Expr], Parser[Type])
+
 class Type
 class IntT extends Type
-val pInt : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Type] =
+val pInt : (=> Pair) => Parser[Type] =
     p => "Int" ^^ { _ => new IntT() }
@@
 class Lam(x : String, t : Type, e : Expr) extends Expr
-val pLam : (=> (PackratParser[Expr], PackratParser[Type])) => PackratParser[Expr] =
-    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1)
-        ^^ { case x ~ t ~ e => new Lam(x, t, e) }
+val pLam : (=> Pair) => Parser[Expr] =
+    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1) ^^
+        { case x ~ t ~ e => new Lam(x, t, e) }
 
-val p : (=> (PackratParser[Expr], PackratParser[Type])) => (PackratParser[Expr], PackratParser[Type]) =
-    p => (pVar(p._1) ||| pApp(p._1) ||| pLam(p), pInt(p))
+val pVar2 : (=> Pair) => Parser[Expr] = p => pVar(p._1)
+val pApp2 : (=> Pair) => Parser[Expr] = p => pApp(p._1)
+
+val pET : (=> Pair) => Pair = {
+    val pE = pVar2 ||| pApp2 ||| pLam
+    val pT = pInt
+    p => (pE(p), pT(p))
+}
 \end{lstlisting}
 The last value \lstinline{p} integrates the four single parsers defined before, so we use \lstinline{fix(p)} to
-obtain the two parsers we want: one for expressions, the other for types. The parser \lstinline{fix(p)._1} parsers expressions
+obtain a pair two parsers we want: one for expressions, the other for types. The parser \lstinline{fix(p)._1} parsers expressions
 like \lstinline{"\x:Int.x x"} successfully. Yet meanwhile, some drawbacks of this approach come into sight. With different syntax, it

### 2016-10-18 paper/sections/Library/FirstTry.tex
More text.
Refactor code in paper.

@@
-\subsection{First Try on Parsing}\label{sec:firsttry}
+\subsection{First Try on Parsing}\label{subsec:firsttry}
 
-Our story begins with a single parser that only parses a free variable (i.e., a string). In Scala we use the Packrat Parsing
-library, which supports lexing and direct left-recursion. To write such a parser, we need to import the library as
+Our story begins with a single parser that only parses a free variable (i.e., a string).
+In Scala we have Packrat parsers supporting lexing and direct left-recursion in its grammar.
+To write such a parser, we need to import the library as
 well as defining the datatype for which it produces.
@@
 object OpenRecursion extends StandardTokenParsers with PackratParsers {
+    type Parser[E] = PackratParser[E]
+
     class Expr
     class Var(x : String) extends Expr
-    val pVar : PackratParser[Expr] = ident ^^ { x => new Var(x) }
+    val pVar : Parser[Expr] = ident ^^ { x => new Var(x) }
     ...
@@
 \end{lstlisting}
-Note that we are omitting lexing and demo code here. Furthermore, we would like the parser to support applications.
+Note that we are omitting lexing and demo code here.
+And we use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} throughout the paper.
+Furthermore, we would like the parser to support applications.
 Not only the implementation of that parser is needed, but also its corresponding datatype: (we omit the fact that the code for below few
@@
 class App(e1 : Expr, e2 : Expr) extends Expr
-val pApp : PackratParser[Expr] = pVar ~ pVar ^^ { case e1 ~ e2 => new App(e1, e2) }
+val pApp : Parser[Expr] = pVar ~ pVar ^^ { case e1 ~ e2 => new App(e1, e2) }
 \end{lstlisting}
@@
 \begin{lstlisting}
-val pApp : PackratParser[Expr] =
-    (pApp | pVar) ~ (pApp | pVar) ^^ { case e1 ~ e2 => new App(e1, e2) }
+val pApp : Parser[Expr] = {
+    val p = pVar ||| pApp
+    p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
+}
 \end{lstlisting}
+Note that we are composing \lstinline{pVar} and \lstinline{pApp} using alternative.
+Alternation is a general operation to compose different parsers, since they represent different grammar rules with the logical relation ``or'' between one another.
+In this example, both sub-expressions of an application can be either a single variable or again an application.
 Now \lstinline{pApp} manages to parse an arbitrary number of applications like \lstinline{"x y z ..."} in a right-associative way.
-Nevertheless, such an approach does not turn out to be a modular extension. The grammar we want for an expression is actually:
+Nevertheless, such an approach does not turn out to be a modular extension, because the grammar we want for an expression is actually:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Note that such an \lstinline{"e"} in the second case should not only include all existing cases, but also take all future extensions into account (the ellipsis we used above, for instance, \lstinline{e '+'} \lstinline{e}).
-This inspires us to use \textit{open recursion}.
-To define \lstinline{pApp} in a different way, we add a parameter \lstinline{p} to it, representing the explicit ``self reference'' of our whole parser, which is open to future extensions.
+Note that such a recursive \lstinline{"e"} in the second case should not only include all the existing cases like \lstinline{pApp} and \lstinline{pVar}, but also take future extensions into account (the ellipsis we used above, for instance, \lstinline{e '+'} \lstinline{e}).
+This is how we are inspired to use \textbf{open recursion}.
+To define \lstinline{pApp} as an ``open'' function, we add a parameter \lstinline{p} to it, representing the explicit self-reference of our whole parser, open to future extensions.
 Note that \lstinline{p} should be defined as a by-name parameter using \lstinline{"=>"}.
 \begin{lstlisting}
-val pApp : (=> PackratParser[Expr]) => PackratParser[Expr] =
+val pApp : (=> Parser[Expr]) => Parser[Expr] =
     p => p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
 \end{lstlisting}
-Similarly we redefine \lstinline{pVar} also as a function using the fixpoint \lstinline{p}.
-Furthermore, we combine \lstinline{pVar} and \lstinline{pApp} together using the parser combinator \lstinline{"|"} for alternative:
+Here the type \lstinline{"(=>} \lstinline{Parser[Expr])} \lstinline{=>} \lstinline{Parser[Expr]"} is exactly what we want for a parser.
+Similarly \lstinline{pVar} is redefined also as a function using the explicit fix-point \lstinline{p}.
+Furthermore, we define an auxiliary combinator \lstinline{"|||"} for parsers of type \lstinline{"(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]"} with the same \lstinline{R} and \lstinline{E}.
+Such a function is overloaded and can be invoked as an infix operator when defined in an implicit class.
+It is really implemented using the general \lstinline{merge} function in Section~\ref{subsec:openrecursion}.
  \begin{lstlisting}
-val pVar : (=> PackratParser[Expr]) => PackratParser[Expr] =
-    p => ident ^^ { x => new Var(x) }
-val pVarApp : (=> PackratParser[Expr]) => PackratParser[Expr] =
-    p => pApp(p) | pVar(p)
+val pVar : (=> Parser[Expr]) => Parser[Expr] = p => ident ^^ { x => new Var(x) }
+
+implicit class ParserCombinator[E, R](x : (=> R) => Parser[E]) {
+    type H[A, B] = Parser[E]
+    def |||(y : (=> R) => Parser[E]) : (=> R) => Parser[E] = {
+        val alt : Parser[E] => Parser[E] => Parser[E] = a => b => a ||| b
+        merge[R, Parser[E], Parser[E], H](alt, x, y)
+    }
+}
 \end{lstlisting}
-But what we really want is something that has type \lstinline{PackratParser[Expr]}.
-In functional programming, we can use ``fix'' to get the fixpoint of a function, in this case we write \lstinline{fix(pVarApp)}, which parses an arbitrary number of applications.
-Such a function can easily be defined
-in Scala based on its laziness:
+Now since \lstinline{pVar} and \lstinline{pApp} are defined independently, we are able to compose them using \lstinline{|||}:
 \begin{lstlisting}
-def fix[A](f : (=> A) => A) : A = { lazy val a : A = f(a); a }
+val pVarApp : (=> Parser[Expr]) => Parser[Expr] = pVar ||| pApp
 \end{lstlisting}
-That is the magic of open recursion: we can define as many small components as we like, and they are implemented as functions, with the help of explicit self-reference.
+Here \lstinline{pVarApp} is the combination of parsers.
+But what we really want is something that has type \lstinline{Parser[Expr]}.
+One can be soon reminded that in Section~\ref{subsec:openrecursion}, we use \lstinline{fix} to get the fix-point of a function, in this case we write \lstinline{fix(pVarApp)}, which manages to parse an arbitrary number of applications.
+
+This is the magic of open recursion: we can define as many small components as we like, and they are implemented as functions, with the help of explicit self-reference.
 Whenever we would like to close the recursion, simply use \lstinline{fix} for their combination.
-Such a process is guaranteed to terminate when we restrict the input for parsing to be finite, and we have some base cases like \lstinline{pVar}, which terminates recursion in their branches, in addition, with the help of direct left-recursion support from Packrat parsers.
+Such a process is guaranteed to terminate when we restrict the input for parsing to be finite, and we have some base cases like \lstinline{pVar}, which terminates recursion in their branches, in addition, with the help of Packrat parsers with direct left-recursion support.
 
 Another thing which catches our attention is that, small parsers are combined using the alternative combinator.
-It is a matter of delegation: parsing failing at any position will try its next alternative, and for each recursion all the parsers are tried successively until one of them succeeds.
+Hence, parsing failing at any position will try its next alternative, and for each recursion all the parsers are tried successively until one of them succeeds.
 It reveals its essence as a recursive descent parser.
 But it is quite easy and modular to use; with further extensions, they can be implemented separately and then appended to the combination of old ones.
+
+Furthermore, in above code we use \lstinline{|||} for alternative instead of \lstinline{|}.
+As we have mentioned in Section~\ref{subsec:parsinglibrary}, \lstinline{|||} is more powerful with longest matching.
+Consider if we obtain \lstinline{pVarApp} from \lstinline{pVar | pApp}, the final parser \lstinline{fix(pVarApp)} cannot even parse an application \lstinline{"x y"}, since it succeeds in parsing a variable first and only returns \lstinline{"x"}.
+With \lstinline{|||}, however, we can put composition in any order and it still works as expected.

### 2016-10-18 Scala/Parser/src/SampleCode/Test.scala

@@ -34,6 +34,43 @@ object Test extends StandardTokenParsers with PackratParsers {
     a
   }
   
+  def merge[E, F, G, H[_,_]](op : F => G => H[F, G], x : (=> E) => F, y : (=> E) => G)
+    : (=> E) => H[F, G] = e => op(x(e))(y(e))
+  
+  type Parser[E] = PackratParser[E]
+  type GenParser[E] = (=> Parser[E]) => Parser[E]
+  implicit class ParserCombinator[E, R](x : (=> R) => Parser[E]) {
+    type H[A, B] = Parser[E]
+    def |||(y : (=> R) => Parser[E]) : (=> R) => Parser[E] = {
+      val alt : Parser[E] => Parser[E] => Parser[E] = a => b => a ||| b
+      merge[R, Parser[E], Parser[E], H](alt, x, y)
+    }
+  }
+  
+  type Pair = (Parser[Expr], Parser[Type])
+  
+  class Type
+  class IntT extends Type
+  val pInt : (=> Pair) => Parser[Type] =
+    p => "Int" ^^ { _ => new IntT() }
+
+  class Lam(x : String, t : Type, e : Expr) extends Expr
+  val pLam : (=> Pair) => Parser[Expr] =
+    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1) ^^
+      { case x ~ t ~ e => new Lam(x, t, e) }
+    
+  
+  val pVar2 : (=> Pair) => Parser[Expr] = p => pVar(p._1)
+  val pApp2 : (=> Pair) => Parser[Expr] = p => pApp(p._1)
+
+  val pET : (=> Pair) => Pair = {
+    val pE = pVar2 ||| pApp2 ||| pLam
+    val pT = pInt
+    p => (pE(p), pT(p))
+  }
+    
+  
+  
   def main(args : Array[String]) = {
     runParser(fix(pVarApp))("x y")
   }

### 2016-10-18 paper/sections/CaseStudy.tex
Modification

@@
 
-To demonstrate the utility of our modular parsing library, we implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages}.
+To demonstrate the utility of our modular parsing library, we implemented parsers of the first \textcolor{red}{18} calculi from book \textit{Types and Programming Languages}.
 In the book it introduces several calculi from simple to complex, by gradually adding new features to syntax.
@@
 
-For better modularity and eliminate code duplication, we extract 21 reusable \textit{components} from syntax of all the calculi.
+Using the modular parsing techniques discussed before, we extract reusable \textit{components} from syntax of all the calculi.
 Each component, which may contain several syntactical structures, is a reusable fragment representing a certain feature in the syntax.
@@
 Each calculus could be composed directly from components and other calculi, if it has those common structures in the syntax.
-For example, the calculus \lstinline{Untyped}, representing the famous untyped lambda calculus, can be constructed from components \lstinline{VarApp} and \lstinline{UntypedAbs}.
+For example, the calculus \lstinline{Untyped}, representing the famous untyped lambda calculus, can be constructed from components \lstinline{VarApp} and untyped lambda abstraction \lstinline{UntypedAbs}.
 
@@
 
+When composing the calculus from components, the \lstinline{Alg} and \lstinline{Print} can be easily combined by Scala’s \lstinline{extends...with} keywords as demonstrated in the code.
+For \lstinline{Parser}, the subtyping constraint for the explicit self-reference should satisfy all the requirement of the parsers it extends from.
+\huang{Talk about | here?}
+
+Finally, we have demo code for every calculus.
+The parsing result is a string because we provide the pretty printing algebra to the parser, and it can be changed as long as the concrete algebra is available.
+Implementing object algebra operation is another topic which is not much relevant with our library, so that in this case study we only have pretty printing as the operation.
+
+\begin{lstlisting}
+object TestUntyped {
+    to be polished
+}
+\end{lstlisting}
+
+In the next section we will show more examples and talk further about how we extend and compose these components and calculi.
+
 \subsection{Extensibility}\label{subsec:extensibility}
+
+As mentioned before, we decompose all the calculi into reusable components for modularity.
+In this case study, we have \textcolor{red}{18} components in total.
+The graph below shows the dependency of all the components, representing by circles, and calculi, representing by boxes.
 

### 2016-10-18 paper/sections/CaseStudy.tex
Some text

@@
 
-\subsection{Overview}\label{subsec:overview}
+\subsection{Overview}\label{subsec:cs-overview}
 
@@
 
-When composing the calculus from components, the \lstinline{Alg} and \lstinline{Print} can be easily combined by Scala’s \lstinline{extends...with} keywords as demonstrated in the code.
+When composing the calculus from components, the \lstinline{Alg} and \lstinline{Print} can be easily combined by Scala’s \lstinline{extends...with} keywords as shown in the code.
 For \lstinline{Parser}, the subtyping constraint for the explicit self-reference should satisfy all the requirement of the parsers it extends from.
@@
 The parsing result is a string because we provide the pretty printing algebra to the parser, and it can be changed as long as the concrete algebra is available.
-Implementing object algebra operation is another topic which is not much relevant with our library, so that in this case study we only have pretty printing as the operation.
+In this case study we only have pretty printing as the operation, since it is enough for demonstration.
 
@@
 
-\subsection{Extensibility}\label{subsec:extensibility}
+\subsection{Extensibility}\label{subsec:cs-extensibility}
 
 As mentioned before, we decompose all the calculi into reusable components for modularity.
-In this case study, we have \textcolor{red}{18} components in total.
+In this case study, we have \textcolor{red}{21} components in total.
 The graph below shows the dependency of all the components, representing by circles, and calculi, representing by boxes.
+An arrow starting from shape A to shape B denotes that B includes and thus reuses A.
 
-\subsection{Comparison}\label{subsec:comparison}
+\textcolor{red}{graph here}
+
+As shown in the graph, some components such as \lstinline{VarApp} are created from scratch, while others such as \lstinline{Typed} are extended from existing components.
+The \lstinline{Typed} component extends from \lstinline{VarApp} and has two more syntactical structures which are typed lambda abstraction and function type (arrow type).
+It is the core of many calculi in this case study, and we use it here as an example for demonstrating the extensibility of components.
+
+\begin{lstlisting}
+object Typed {
+  trait Alg[E, T] extends VarApp.Alg[E] {
+    def TmAbs(x: String, t: T, e: E): E
+    def TyArr(t1: T, t2: T): T
+  }
+  trait Print extends Alg[String, String] with VarApp.Print {
+    def TmAbs(x: String, t: String, e: String) = "\\(" + x + ":" + t + ")." + e
+    def TyArr(t1: String, t2: String) = t1 + "->" + t2
+  }
+  trait Parser[E, T, F <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends VarApp.Parser[E, F] {
+    ...
+    private val pAbsE: Alg[E, T] => (=> F) => PackratParser[E] = {...}
+    val pTypedE: Alg[E, T] => (=> F) => PackratParser[E] = pVarAppE | pAbsE
+    val pTypedT: Alg[E, T] => (=> F) => PackratParser[T] = alg => l => {
+      val t = l.pT
+      t ~ ("->" ~> t) ^^ { case t1 ~ t2 => alg.TyArr(t1, t2) } ||| "(" ~> t <~ ")"
+    }
+  }
+}
+\end{lstlisting}
+
+Except building a component by extending an existing one, here we have another extension which is adding a new sort of syntax.
+It has been discussed in section \ref{subsec:openrecursion}.
+In this example, originally \lstinline{VarApp} has only one sort of syntax which is expressions, and then \lstinline{Typed} introduces types to the syntax.
+We add an extra type parameter \lstinline{T} for it, so that it is able to be distinguished.
+The \lstinline{Parser} also requires the self-reference to have a field \lstinline{pT} for parsing types, which is represented by the subtyping constraint.
+With this support, the \lstinline{pTypedT} function parses arrow types using \lstinline{pT}.
+
+\subsection{Comparison}\label{subsec:cs-comparison}
 

### 2016-10-18 Scala/Parser/src/TAPL/SimpleBool.scala

@@ -20,14 +20,15 @@ object Typed {
   trait Parser[E, T, F <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends VarApp.Parser[E, F] {
     lexical.delimiters += ("\\", ".", "(", ")", ":", "->")
 
-    private val pAbsArrE: Alg[E, T] => (=> F) => PackratParser[E] = alg => l => {
+    private val pAbsE: Alg[E, T] => (=> F) => PackratParser[E] = alg => l => {
       val e = l.pE
       val t = l.pT
 
-      ("\\" ~> lcid) ~ (":" ~> t) ~ ("." ~> e) ^^ { case x ~ t0 ~ e0 => alg.TmAbs(x, t0, e0) } ||| "(" ~> e <~ ")"
+      ("\\" ~> lcid) ~ (":" ~> t) ~ ("." ~> e) ^^ { case x ~ t0 ~ e0 => alg.TmAbs(x, t0, e0) } |||
+        "(" ~> e <~ ")"
     }
 
-    val pTypedE: Alg[E, T] => (=> F) => PackratParser[E] = pVarAppE | pAbsArrE
+    val pTypedE: Alg[E, T] => (=> F) => PackratParser[E] = pVarAppE | pAbsE
 
     val pTypedT: Alg[E, T] => (=> F) => PackratParser[T] = alg => l => {
       val t = l.pT

### 2016-10-19 paper/sections/CaseStudy.tex
Some text

@@
 
-When composing the calculus from components, the \lstinline{Alg} and \lstinline{Print} can be easily combined by Scala’s \lstinline{extends...with} keywords as shown in the code.
+When composing the calculus from components, all the fields can be easily combined by Scala’s \lstinline{extends...with} keywords as shown in the code.
 For \lstinline{Parser}, the subtyping constraint for the explicit self-reference should satisfy all the requirement of the parsers it extends from.
@@
 With this support, the \lstinline{pTypedT} function parses arrow types using \lstinline{pT}.
+Another point worth mentioning is that we use a private local function \lstinline{pAbsE} here for parsing the typed lambda abstraction case, and we can directly combine it with the inherited \lstinline{pVarApp} to obtain our parsing function of expressions.
+
+Since calculi and components have the similar signature, each calculus can also be extended and reused directly.
+Our last example shows how to build calculus \lstinline{FullRef} from extending \lstinline{FullSimple}, another calculus.
+
+\begin{lstlisting}
+object FullRef {
+  trait Alg[E, T] extends FullSimple.Alg[E, T] with TopBot.Alg[T] with Ref.Alg[E, T]
+  trait Print extends Alg[String, String] with FullSimple.Print with TopBot.Print with Ref.Print
+  trait Parser[E, T, L <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends FullSimple.Parser[E, T, L] with TopBot.Parser[T, L] with Ref.Parser[E, T, L] {
+    val pFullRefE = pFullSimpleE | pRefE
+    val pFullRefT = pFullSimpleT | pRefT | pTopBotT
+  }
+}
+\end{lstlisting}
+
+Using our modular parsing library, dividing the syntax into small pieces and bind them with corresponding parsers to form reusable language component is convenient.
+From the dependency graph we know that the common components such as \lstinline{VarApp} are reused in lots of calculi.
+Such reuse could shorten the codes considerably.
+We will show this advantage and exam the possible performance penalty in a quantitative way in the next section.
 

### 2016-10-19 paper/sections/Library/DifferentSyntax.tex
Text.

@@
 \end{lstlisting}
-The last value \lstinline{p} integrates the four single parsers defined before, so we use \lstinline{fix(p)} to
-obtain a pair two parsers we want: one for expressions, the other for types. The parser \lstinline{fix(p)._1} parsers expressions
-like \lstinline{"\x:Int.x x"} successfully. Yet meanwhile, some drawbacks of this approach come into sight. With different syntax, it
-is necessary to extend the fixpoint to contain more parsers, in which case usually some structures like tuples are used,
-making code tedious and less elegant. On the other hand, the modularity is quite restricted; only parsing and datatypes are extensible.
+Based on the old \lstinline{pVar} and \lstinline{pApp}, we implement \lstinline{pVar2} and \lstinline{pApp2} which take the pair as their fix-points instead.
+In that case the parsers are consistent, and hence can be composed using the combinator \lstinline{|||}.
+The last value \lstinline{pET} integrates the four single parsers defined before, where parsers for expressions and parsers for types are composed respectively.
+Furthermore, we use \lstinline{fix(p)} to obtain the pair of two parsers we want: one for expressions, the other for types.
+The parser \lstinline{fix(p)._1} parsers expressions like \lstinline{"\x:Int.x x"} successfully.
+Yet meanwhile, some drawbacks of this approach come into sight.
+With different syntax, it
+is necessary to extend the fix-point to contain more parsers, in which case usually some structures like tuples are used,
+making code tedious and less elegant.
+Moreover, we need to update such a fix-point in the arguments of all previously defined parsers, so as to make parsers consistent for the combinator (things get even worse without that combinator).
+On the other hand, the modularity is quite restricted; only parsing and datatypes are extensible.
 Suppose now we want to realize some behaviors on those data structures, for example pretty-printing, then whenever we introduce a new behavior,
@@
 
-Object Algebras are a nice solution to the famous \textit{Expression Problem}, achieving two dimensions of extensibility (data variants and operations)
-in a type-safe way.
-The intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design.
-We can simply add new variants for the new cases in the grammar, or realize different operations on structures, without modifying existing code, and separate compilation is supported.
-Coincidentally, Object Algebras use generics, where the type parameters perfectly support extensible different syntax, and hence simplify the code comparatively.
-Below is the code using Object Algebras, where we start from \lstinline{varE} and \lstinline{appE} only, later we will extend them with \lstinline{intT} and \lstinline{lamE}:
+One intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
+We can simply add new variants corresponding to the new cases in the grammar, or realize different operations on structures, without modifying existing code, meanwhile separate compilation is supported.
+Another one is that Object Algebras benefit us from using generics, where we can abstract different syntax in the type parameters, and define less classes, simplifying code comparatively.
+Below we present the code using Object Algebras, where we start from \lstinline{varE} and \lstinline{appE} only, later we will extend them with \lstinline{intT} and \lstinline{lamE}:
 \begin{lstlisting}
@@
 trait OldParser[E] {
-    val pE : OldAlg[E] => (=> PackratParser[E]) => PackratParser[E] = alg => p =>
+    val pE : OldAlg[E] => (=> Parser[E]) => Parser[E] = alg => p =>
         ident ^^ { x => alg.varE(x) } |||
@@
 \end{lstlisting}
-Here \lstinline{OldAlg} is called an \textit{Object Algebra interface}, which has two constructs for the AST. Its parser is
-supposed to produce a structure from an input, whereas with Object Algebras structures are implicit: an \textit{algebra}
-is immediately applied to the structure and returns the result. It means such a structure that the parser generates should have
-type \lstinline{"OldAlg[E] =>} \lstinline{E"} for some \lstinline{E}. In the above code, \lstinline{alg} which has type
-\lstinline{OldAlg[E]} is an algebra, and appears as a parameter of \lstinline{pE}, then in the body of \lstinline{pE} the algebra
-is invoked correspondingly right after parsing. Since it should be a generic algebra, the parser is enclosed by the generic trait \lstinline{OldParser}.
+Here \lstinline{OldAlg} is an Object Algebra interface.
+An Object Algebra interface is more like a small language, where several cases (constructors) are put together.
+Since with Object Algebras structures are implicit, a parser of \lstinline{OldAlg} should generate ``structures'' with
+type \lstinline{"OldAlg[E] =>} \lstinline{E"} for any \lstinline{E}. In \lstinline{OldParser}, \lstinline{alg} of type
+\lstinline{OldAlg[E]} can be any algebra, which appears as a parameter of \lstinline{pE}, then in the body of \lstinline{pE} the algebra
+is invoked correspondingly right after parsing.
+It means we do not get any explicit objects generated from parsers, but they are immediately applied with algebras and return the results.
+To make \lstinline{alg} as a generic algebra, the enclosing trait of the parser, namely \lstinline{OldParser}, is generic.
 

### 2016-10-31 paper/sections/CaseStudy.tex
Graph in case study
Some text and half of the table

@@
 
-To demonstrate the utility of our modular parsing library, we implemented parsers of the first \textcolor{red}{18} calculi from book \textit{Types and Programming Languages}.
+To demonstrate the utility of our modular parsing library, we implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages}.
 In the book it introduces several calculi from simple to complex, by gradually adding new features to syntax.
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{depGraph}
+    \caption{Dependency graph of all calculi and components}
+    \label{fig:dependency}
+\end{figure}
+
 As mentioned before, we decompose all the calculi into reusable components for modularity.
-In this case study, we have \textcolor{red}{21} components in total.
-The graph below shows the dependency of all the components, representing by circles, and calculi, representing by boxes.
-An arrow starting from shape A to shape B denotes that B includes and thus reuses A.
-
-\textcolor{red}{graph here}
+Figure \ref{fig:dependency} shows the dependency of all the components and calculi in our case study.
+Gray boxes are calculi and white boxes are components.
+An arrow starting from box A to box B denotes that B includes and thus reuses A.
 
@@
 From the dependency graph we know that the common components such as \lstinline{VarApp} are reused in lots of calculi.
-Such reuse could shorten the codes considerably.
+Such reuse could shorten the code considerably.
 We will show this advantage and exam the possible performance penalty in a quantitative way in the next section.
@@
 
-\begin{itemize}
-\item Explain the TAPL case study
-\item Show various results SLOC; performance
-\end{itemize}
+We compared our implementation of the parsers with an implementation we found online, written by author Ilya Klyuchnikov.
+Ilya's implementation is very suitable for comparison, because it is also written in Scala using the same parser combinator library.
+Furthermore, it includes parsers of all the 18 calculi we have, but written in a non-modular way.
+Thus it is not able to reuse existing code of parsing common syntactical features.
+
+Because Ilya's implementation has extra code such as semantics of those calculi, we removed all irrelevant code and only keep abstract syntax definition, parser and pretty printer for each calculus, to ensure a fair comparison.
+
+The comparison is made from two aspects.
+First, we want to show that using our modular parsing library, the code can be reused in a significant way.
+For this propose, we measured source lines of code (SLOC) of the two implementations.
+All blank lines and comment lines are excluded, and we formatted the code of both implementations so that the length of each line does not exceed 120 characters.
+Second, users may concern the performance penalty caused by modularity.
+Thus we compared the execution time, as a major measurement of performance, between the two implementations.
+We prepared test files containing valid expressions in each calculus, then we used the corresponding parser to parse the file and the pretty printer to print the result.
+The execution time excluding reading input file was calculated.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|l|r|r|r|}
+        \hline
+        \bfseries Calculus Name & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\%
+        \csvreader{../Scala/sloc.csv}{}
+        {\\\hline \csvcoli & \csvcolii & \csvcoliii & \csvcoliv}
+        %    \csvautotabular{../Scala/sloc.csv}
+        \\\hline
+    \end{tabular}
+    \caption{Comparison of SLOC and execution time}
+    \label{tab:comparison}
+\end{table}
+
+Table \ref{tab:comparison} shows the comparison result.

### 2016-10-31 paper/sections/CaseStudy.tex
Rearrange the table and the graph

@@
     \centering
-    \includegraphics[width=\textwidth]{depGraph}
+    \includegraphics[width=\textwidth]{resources/depGraph.pdf}
     \caption{Dependency graph of all calculi and components}
@@
     \centering
-    \begin{tabular}{|l|r|r|r|}
-        \hline
-        \bfseries Calculus Name & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\%
-        \csvreader{../Scala/sloc.csv}{}
-        {\\\hline \csvcoli & \csvcolii & \csvcoliii & \csvcoliv}
-        %    \csvautotabular{../Scala/sloc.csv}
-        \\\hline
-    \end{tabular}
+    \input{resources/CaseStudyTable.tex}
     \caption{Comparison of SLOC and execution time}
@@
 
-Table \ref{tab:comparison} shows the comparison result.
+Table \ref{tab:comparison} shows data of the comparison.
+The overall result of SLOC comparison is that 60\% of code in total is reduced using our library.
+That is because the code of some common syntax structures such as variable, lambda abstraction, etc., is reused several times in the whole project.
+We can see that the first two calculi \lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's implementation, because in such two cases we do not reuse any existing components.
+But in each of the following 16 calculi, we can compose our syntax definition, parser and pretty printer from extending reusable components, resulting considerably code reduction.

### 2016-10-31 Scala/table.py
Finish the script

@@ -113,6 +113,34 @@ def write_csv(sloc1, sloc2, time1, time2):
         writer.writerow(row)
 
 
+def write_tex(file_name, sloc1, sloc2, time1, time2):
+    def cal_percent(old, new):
+        per_str = "{:+.1f}".format((new - old) / old * 100)
+        return per_str
+
+    def tex_line(name, s1, s2, t1, t2):
+        vals = [name,
+                s1, s2, cal_percent(s1, s2),
+                t1, t2, cal_percent(t1, t2)]
+        return " & ".join([str(v) for v in vals]) + " \\\\\n"
+
+    # write each case
+    print_hline()
+    print('WRITING RESULT TO', file_name)
+    print_hline()
+    with open(file_name, 'w') as f:
+        for (name, s1, s2, t1, t2) in zip(NAMES, sloc1, sloc2, time1, time2):
+            row = tex_line(name, s1, s2, t1, t2)
+            print(row, end='')
+            f.write(row)
+        # write total
+        s1, s2, t1, t2 = sum(sloc1), sum(sloc2), sum(time1), sum(time2)
+        row = tex_line('Total', s1, s2, t1, t2)
+        print(row, end='')
+        f.write('\\hline\n')
+        f.write(row)
+
+
 def main():
     # count SLOC
     sloc1 = count_theirs()
@@ -129,7 +157,9 @@ def main():
     assert(len(time1) == len(time2) == len(NAMES))
 
     # write to file
-    write_csv(sloc1, sloc2, time1, time2)
+    # write_csv(sloc1, sloc2, time1, time2)
+    tex_file = '../paper/resources/CaseStudyTable.tex'
+    write_tex(tex_file, sloc1, sloc2, time1, time2)
 
 
 if __name__ == '__main__':

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullequirec/parser.scala
All test case generated

@@ -19,6 +19,8 @@ case object TyBool extends Ty
 
 case object TyString extends Ty
 
+case object TyFloat extends Ty
+
 case object TyUnit extends Ty
 
 case class TyRec(id: String, ty: Ty) extends Ty
@@ -68,8 +70,8 @@ case class TmFix(t: Term) extends Term
 object FullEquiRecParsers extends StandardTokenParsers with PackratParsers with ImplicitConversions {
   lexical.reserved += ("Bool", "true", "false", "if", "then", "else",
     "Nat", "String", "Unit", "Float", "unit", "case", "let", "in", "succ", "pred",
-    "as", "of", "fix", "iszero", "letrec", "Rec", "_")
-  lexical.delimiters += ("\\", "(", ")", ";", "/", ".", ":", "->", "=", "<", ">", "{", "}", "=>", ",", "|")
+    "as", "of", "fix", "iszero", "letrec", "Rec", "inert")
+  lexical.delimiters += ("\\", "(", ")", ";", "/", ".", ":", "->", "=", "<", ">", "{", "}", "=>", ",", "|", "[", "]")
 
   // lower-case identifier
   lazy val lcid: PackratParser[String] = ident ^? { case id if id.charAt(0).isLower => id }
@@ -88,7 +90,8 @@ object FullEquiRecParsers extends StandardTokenParsers with PackratParsers with 
       "String" ^^ { _ => TyString } |
       "Unit" ^^ { _ => TyUnit } |
       "{" ~> fieldTypes <~ "}" ^^ { ft => TyRecord(ft) } |
-      "Nat" ^^ { _ => TyNat }
+      "Nat" ^^ { _ => TyNat } |
+      "Float" ^^ { _ => TyFloat }
 
   lazy val fieldTypes: PackratParser[List[(String, Ty)]] =
     repsep(fieldType, ",")

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullequirec/syntax.scala

@@ -51,6 +51,8 @@ object FullEquiRecPrinter {
         _).getOrElse(empty) :: "}")
     case TyNat =>
       "Nat"
+    case TyFloat =>
+      "Float"
     case tyT =>
       "(" :: ptyType(outer, tyT) :: ")"
   }

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullisorec/syntax.scala

@@ -51,6 +51,8 @@ object FullIsoRecPrinter {
         _).getOrElse(empty) :: "}")
     case TyNat =>
       "Nat"
+    case TyFloat =>
+      "Float"
     case tyT =>
       "(" :: ptyType(outer, tyT) :: ")"
   }

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullomega/parser.scala

@@ -27,6 +27,8 @@ case object TyBool extends Ty
 
 case object TyNat extends Ty
 
+case object TyFloat extends Ty
+
 case class TyApp(ty1: Ty, ty2: Ty) extends Ty
 
 case class TyAbs(v: String, k: Kind, ty: Ty) extends Ty
@@ -94,7 +96,7 @@ object FullOmegaParsers extends StandardTokenParsers with PackratParsers with Im
   lexical.reserved += ("Bool", "true", "false", "if", "then", "else",
     "Nat", "String", "Unit", "Float", "unit", "case", "let", "in", "succ", "pred",
     "as", "of", "fix", "iszero", "Star", "_", "All", "Some", "Ref", "ref")
-  lexical.delimiters += ("(", ")", ";", "/", ".", ":", "->", "=",
+  lexical.delimiters += ("(", ")", ";", "/", ".", ":", "->", "=", ":=", "!",
     "<", ">", "{", "}", "=>", ",", "|", "*", "[", "]", "\\")
 
   // lower-case identifier
@@ -133,6 +135,7 @@ object FullOmegaParsers extends StandardTokenParsers with PackratParsers with Im
       "Unit" ^^ { _ => TyUnit } |
       "{" ~> fieldTypes <~ "}" ^^ { ft => TyRecord(ft) } |
       "Nat" ^^ { _ => TyNat } |
+      "Float" ^^ { _ => TyFloat } |
       (("{" ~ "Some") ~> ucid) ~ oKind ~ ("," ~> `type` <~ "}") ^^ { case id ~ k ~ ty => TySome(id, k, ty) }
 
   lazy val fieldTypes: PackratParser[List[(String, Ty)]] =

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullpoly/syntax.scala

@@ -41,6 +41,8 @@ object FullPolyPrinter {
         _).getOrElse(empty) :: "}")
     case TyNat =>
       "Nat"
+    case TyFloat =>
+      "Float"
     case TySome(tyX, tyT2) =>
       g2("{Some " :: tyX :: "," :/: ptyType(false, tyT2) :: "}")
     case tyT =>

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullref/parser.scala

@@ -19,6 +19,8 @@ case object TyBool extends Ty
 
 case object TyString extends Ty
 
+case object TyFloat extends Ty
+
 case object TyNat extends Ty
 
 case class TyRef(ty: Ty) extends Ty
@@ -84,9 +86,9 @@ case class TmAssign(t1: Term, t2: Term) extends Term
 object FullRefParsers extends StandardTokenParsers with PackratParsers with ImplicitConversions {
   lexical.reserved += ("Bool", "true", "false", "if", "then", "else",
     "Nat", "String", "Unit", "Float", "unit", "case", "let", "in", "succ", "pred",
-    "as", "of", "fix", "iszero", "Top", "Bot", "Ref", "Source", "Sink", "ref")
-  lexical.delimiters += ("\\", "(", ")", ";", "/", ".", ":", "->", "=", "<", ">", "{", "}", "=>", "==>", ",", "|",
-    "!", ":=")
+    "as", "of", "fix", "iszero", "Top", "Bot", "Ref", "Source", "Sink", "ref", "inert")
+  lexical.delimiters += ("\\", "(", ")", ";", "/", ".", ":", "->", "=", "<", ">", "{", "}", "=>", ",", "|",
+    "!", ":=", "[", "]")
 
   // lower-case identifier
   lazy val lcid: PackratParser[String] = ident ^? { case id if id.charAt(0).isLower => id }
@@ -108,7 +110,8 @@ object FullRefParsers extends StandardTokenParsers with PackratParsers with Impl
       "String" ^^ { _ => TyString } |
       "Unit" ^^ { _ => TyUnit } |
       "{" ~> fieldTypes <~ "}" ^^ { ft => TyRecord(ft) } |
-      "Nat" ^^ { _ => TyNat }
+      "Nat" ^^ { _ => TyNat } |
+      "Float" ^^ { _ => TyFloat }
 
   lazy val fieldTypes: PackratParser[List[(String, Ty)]] =
     repsep(fieldType, ",")
@@ -167,7 +170,7 @@ object FullRefParsers extends StandardTokenParsers with PackratParsers with Impl
   lazy val cases: PackratParser[List[(String, String, Term)]] =
     rep1sep(`case`, "|")
   lazy val `case`: PackratParser[(String, String, Term)] =
[…]

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullref/syntax.scala

@@ -55,6 +55,8 @@ object FullRefPrinter {
         _).getOrElse(empty) :: "}")
     case TyNat =>
       "Nat"
+    case TyFloat =>
+      "Float"
     case tyT =>
       "(" :: ptyType(outer, tyT) :: ")"
   }

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullsimple/parser.scala

@@ -21,6 +21,8 @@ case object TyString extends Ty
 
 case object TyNat extends Ty
 
+case object TyFloat extends Ty
+
 sealed trait Term
 
 case object TmTrue extends Term
@@ -81,7 +83,8 @@ object FullSimpleParsers extends StandardTokenParsers with PackratParsers with I
       "String" ^^ { _ => TyString } |
       "Unit" ^^ { _ => TyUnit } |
       "{" ~> fieldTypes <~ "}" ^^ { ft => TyRecord(ft) } |
-      "Nat" ^^ { _ => TyNat }
+      "Nat" ^^ { _ => TyNat } |
+      "Float" ^^ { _ => TyFloat }
 
   lazy val fieldTypes: PackratParser[List[(String, Ty)]] =
     repsep(fieldType, ",")

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullsimple/syntax.scala

@@ -48,6 +48,8 @@ object FullSimplePrinter {
         _).getOrElse(empty) :: "}")
     case TyNat =>
       "Nat"
+    case TyFloat =>
+      "Float"
     case tyT =>
       "(" :: ptyType(outer, tyT) :: ")"
   }

### 2016-11-01 Scala/Parser/src/TAPLcomp/fullsub/syntax.scala

@@ -41,6 +41,8 @@ object FullSubPrinter {
         _).getOrElse(empty) :: "}")
     case TyNat =>
       "Nat"
+    case TyFloat =>
+      "Float"
     case tyT =>
       "(" :: ptyType(outer, tyT) :: ")"
   }

### 2016-11-08 paper/sections/CaseStudy.tex
Finish the script
Fix, add table header
Text

@@
     \centering
+    \begin{tabular}{|l|r|r|r|r|r|r|}
+      \hline
+        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{3}{ c| }{\bfseries Time (ms)} \\ \cline{2-7}
+        \multicolumn{1}{|c|}{} & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% \\
+      \hline
       \input{resources/CaseStudyTable.tex}
+      \hline
+    \end{tabular}
     \caption{Comparison of SLOC and execution time}
@@
 Table \ref{tab:comparison} shows data of the comparison.
-The overall result of SLOC comparison is that 60\% of code in total is reduced using our library.
+The overall result of SLOC comparison is that 66.3\% of code is reduced using our library.
 That is because the code of some common syntax structures such as variable, lambda abstraction, etc., is reused several times in the whole project.
@@
 But in each of the following 16 calculi, we can compose our syntax definition, parser and pretty printer from extending reusable components, resulting considerably code reduction.
+
+For the execution time comparison, we built a generator to randomly generate valid expressions according to the syntax of each calculus.
+Each test file corresponds with one calculus and it consists of 500 expressions randomly generated.
+The size of test files varies from 20KB to 100KB.
+After testing both implementation, we found that our version is about \huang{60\%} slower.
+Part of the reason is that our library requires more intermediate function calls, and the object algebra and self-reference needed to be passed as arguments every time.
+However, another important reason is that, because of the modular setting, we use a different alternative combinator which will try both branches and match the longest one.
+\huang{ref to somewhere before?} In contrast Ilya uses the ordinary alternative combinator which tries the second branch only if the first one fails.
+Although this difference slows our version, we want to mention that because of the memorization of Packrat parsers, we are just constant times slower, the overall complexity is still polynomial regarding to the length of the input.

### 2016-11-08 paper/sections/Overview.tex
Text

@@
 \section{Overview}\label{sec:overview}
+
+Our modular parsing library consists of four parts, and we will go through them in this section.
+The first is underlying parsing technique we used.
+Parsing has been heavily studied over the years, especially for context-free grammars and its subsets.
+However, the choice of parsing techniques is not a minor issue here, because of modularity.
+To achieve modularity, the parsers must be written in a way that can be composed, and different parsing techniques have different suitability for that.
+
+Although every parsing algorithm is able to implement by hand in principle, some of them are not friendly for that.
+A parser generator is a tool to generate parsers automatically or semi-automatically, from a grammar given by the user.
+There is no restriction on the algorithm, however most of them adopts LL and LR parsing algorithms, which are table-based.
+
+\huang{TALK ABOUT PREVIOUS WORKS USING PARSER GENERATORS, AND ALSO THE DISADVTANGES OF THEM}
+
+Comparing with the parser generators, a parser combinator takes several parsers and produce a new parser as its output.
+It is popular in functional programming, where the parsers are represented by functions and parser combinators are higher-order functions accepts them.
+Parser combinators are very suitable for our purpose, because of two reasons.
+Firstly, they are naturally modular.
+The manner of using them is to write small parsers and use combinators to composed them together.
+The construction procedure is explicit and fully controlled by the programmer.
+Secondly, each parser combinator is represented by a piece of code, and also are the parsers it takes.
+Thus in a statically typed programming language they can be type-checked during compile time.
+
+There are variants in the category of parser combinators.
+Almost all of them adopt top-down, recursive descent parsing, which uses backtracking to search through the possible branches.
+Simple backtracking parser combinators, such as the famous library Parsec in Haskell, have some drawbacks.
+One is that they cannot support left-recursive grammars.
+The common solution is to rewrite a left-recursive grammar into an equivalent one, so called left-recursion elimination.
+This solution requires the whole grammar, so that it cannot be applied in a modular setting because we only have parts of the grammar.
+Another drawback is performance.
 

### 2016-11-08 paper/sections/CaseStudy.tex
Text

@@
 \huang{ref to somewhere before?} In contrast Ilya uses the ordinary alternative combinator which tries the second branch only if the first one fails.
-Although this difference slows our version, we want to mention that because of the memorization of Packrat parsers, we are just constant times slower, the overall complexity is still polynomial regarding to the length of the input.
+Although this difference slows our version, we want to mention that because of the memorization of Packrat parsers, we are just constant times slower, the big-O complexity is still the same.

### 2016-11-08 paper/sections/Overview.tex
Text

@@
 The common solution is to rewrite a left-recursive grammar into an equivalent one, so called left-recursion elimination.
-This solution requires the whole grammar, so that it cannot be applied in a modular setting because we only have parts of the grammar.
+This solution requires information of the whole grammar, so that it cannot be applied in a modular setting because we only have parts of the grammar.
 Another drawback is performance.
+Take Parsec as an example, its choice combinator only tries the second alternative if the first fails without any token consumption.
+If two alternatives have a same non-empty prefix, an auxiliary function \lstinline{try} must be added to backtrack.
+However, if we want to extend our parsers later, we should always consider the worst case in which all alternatives share common prefixes.
+Then we need to add \lstinline{try} for all the branches, which results worst case exponential time complexity.
+
+Beyond the simple backtracking one, several advanced parser combinators were studied to resolve those issues.
+One of them is Packrat parsers, which is our choice.
+The details of Packrat parser are beyond the scope of this paper.
+In brief, Packrat parsers use memorization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated, and it supports left-recursive grammars.
+These properties are very suitable for modularity.
+Thus we use Packrat parsers as the underlying parsing technique in our library.
+
+It is worth mentioning that the choice of parser combinators will not affect the other parts of our library.
+One can choose other parser combinators like Parsec, in cases that the performance and supporting of left-recursion are not major concerns.
+
+The second part of our library is open recursion.
+It is introduced to deal with the problem of recursive calls in modular parsing.
+We use a simple example to demonstrate the problem, which is parsing a language of natural numbers.
+The language is defined below.
+
+\begin{lstlisting}[language=PlainCode]
+Expr := zero | succ Expr
+\end{lstlisting}
+
+Firstly, we need to define the abstract syntax. Here we use Scala's case classes to save spaces, this can also be done using class inheritance.
+
+\begin{lstlisting}
+abstract class Expr
+case class Zero extends Expr
+case class Succ(v: Expr) extends Expr
+\end{lstlisting}
+
+And then we write parsing functions for every cases.
+Some irrelevant details in the code are omitted to outline the main structure.
+Suppose we have \lstinline{choice} combinator for alternatives and the input tokens are global values which can be accessed by all these functions.
+
+\begin{lstlisting}
+def parseExpr(): Expr = choice(parserZero, parseSucc)
+
+def parserZero(): Expr = {...}
+
+def parseSucc(): Expr = {
+  parseToken('succ')
+  val v = parseExpr()
+  Succ(v)
+}
+\end{lstlisting}
+
+It works well, but later we want to extend the language to support addition.
+
+\begin{lstlisting}[language=PlainCode]
+Expr := zero | succ Expr | add Expr Expr
+\end{lstlisting}
+
+Extending the abstract syntax is easy.
+
+\begin{lstlisting}
+case class Add(a: Expr, b: Expr) extends Expr
+\end{lstlisting}
 

### 2016-11-09 paper/sections/Library/ParsingLibrary.tex
Organize.

--- 2016-10-13 paper/sections/Library/ParsingLibrary.tex
@@
-\subsection{Parsing Library}\label{subsec:parsinglibrary}
+\subsection{Fusion of Components/Concepts}
+
+In the previous sections we have talked about open recursion and Object Algebras. These concepts will later be composed together for our
+initial goal, which is to build extensible parsers, but surely it cannot live without a parsing library. But we want to argue here that besides
+open recursion, which constructs the basis of extensibility in parsing, the other aspects are totally orthogonal, and hence can have their alternatives.
+For example, Object Algebras are helpful to build extensible data structures, whereas an alternative with the same function can still be used. And a better
+parsing library would introduce a lot of fancy features, nevertheless, it is independent from our key idea.

### 2016-11-09 paper/sections/ObjectAlgebras.tex
Organize.

--- 2016-10-15 paper/sections/Library/ObjectAlgebras.tex
@@
-\subsection{Object Algebras}\label{subsec:objectalgebras}
+\section{Object Algebras and Parsing}\label{sec:algebrasandparsing}
 
-Object Algebras captures a design pattern to address the famous \textit{Expression Problem} nicely,
+Object Algebras, first introduced by ... et al as a solution to the famous \textit{Expression Problem}, provide extensibility on
+both data variants and operations for structures like abstract syntax trees. We integrate Object Algebras directly to enhance the
+extensibility of open parsers.
+
+\subsection{Object Algebras}\label{sec:objectalgebras}
+Object Algebras captures a design pattern to address the Expression Problem nicely,
 achieving two dimensions of extensibility (data variants and operations) in a modular and type-safe way.

### 2016-11-09 paper/sections/OpenRecursion.tex
Organize.

\section{Open Recursion and Parsing}\label{sec:openandparsing}

Our story begins with a well-known concept mostly in object-oriented programming, called \textit{open recursion}. This mechanism
constructs the basis of extensibility of our parsing. We will introduce the concept with illustration before integrating
it  to achieve halfway extensible parsers.

\subsection{Open Recursion with Delegation}\label{subsec:openrecursion}

Open recursion is known as a useful feature that one method body can ``invoke another method of the same object via a special variable called \lstinline{self} or, in some languages, \lstinline[keywords={}]{this}'' (by Ralf Hinze). The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, while existing methods that take \lstinline{self} as a parameter can simply delegate their known cases to it.

For sure we cannot always leave it open; upon using it we have to close the recursion, at which time we need to take a fix-point. Here one might suddenly be reminded of the \lstinline{fix} function in some functional languages like Haskell. In Haskell \lstinline{fix} is defined as:

\begin{lstlisting}[language=Haskell,keywords={}]
fix f = let x = f x in x
\end{lstlisting}
In Scala, however, there is no library \lstinline{fix} function, but it can be easily implemented since Scala also supports lazy evaluation. See below:
\begin{lstlisting}
def fix[A](f : (=> A) => A) : A = {
    lazy val a : A = f(a)
    a
}
\end{lstlisting}
where function \lstinline{f} has a by-name parameter of type \lstinline{"=>} \lstinline{A"}.
Here we illustrate by the famous Fibonacci example. Traditionally it can be implemented in Scala using pattern-matching:
\begin{lstlisting}
def fib : Int => Int = i => i match {
    case 0 => 0
    case 1 => 1
    case n => fib(n - 1) + fib(n - 2)
}

val x = fib(2) // 1
\end{lstlisting}
An alternative implementation using open recursion is as follows:
\begin{lstlisting}
type Fix[T] = (=> T) => T
def fib2 : Fix[Int => Int] = self => i => i match {
    case 0 => 0
    case 1 => 1
    case n => self.apply(n - 1) + self.apply(n - 2)
}

val x2 = fix(fib2)(2) // 1
\end{lstlisting}
where \lstinline{self} is the explicit self-reference, namely the delegator. Functions like \lstinline{fib2} which takes
an explicit self-reference as its parameter are called \textit{generators} (by Ralf Hinze).
If we close the recursion on \lstinline{fib2} only, the evaluation of \lstinline{fix(fib2)(2)}, namely the above \lstinline{x2} is processed as follows:
\begin{lstlisting}[language=Haskell,keywords={}]
   fix(fib2)(2)
= lazy val a = fib2(a); a.apply(2)
= lazy val a = fib2(a); fib2(a)(2)
= ... (the third case in fib2)
= lazy val a = fib2(a); a.apply(1) + a.apply(0)
= lazy val a = fib2(a); fib2(a)(1) + fib2(a)(0)
= ... (the first two cases in fib2)
= lazy val a = fib2(a); 1 + 0
= 1
\end{lstlisting}
which behaves similarly as \lstinline{fib(2)}. The process of evaluation will terminate if there are some base cases (like the first two in
\lstinline{fib2}) which stop the recursion in their branches, and the evaluation is finally reduced to those cases.

Suppose now we want to make use of open recursion for additional operations. We first define a combinator which combines the results of two generators in a pair:
\begin{lstlisting}
implicit class Compose[A, B, C](f : Fix[A => B]) {
    def *(g : Fix[A => C]) : Fix[A => (B, C)] = self => x =>
        (f(self.apply(_)._1)(x), g(self.apply(_)._2)(x))
}
\end{lstlisting}
Using Scala implicit class, \lstinline{*} can be used as an infix operator for composition in convenience.
Then we define a generator which prints out a number as a string:
\begin{lstlisting}
def show : Fix[Int => String] = self => x => x.toString
\end{lstlisting}
Now \lstinline{fix(fib2 * show)(2)} results in \lstinline{(1, "2")}. We have seen that both \lstinline{fib2} and \lstinline{show} are integrated in the evaluation.

Next we define a more general combinator \lstinline{merge} for open recursion. It is similar to the \lstinline{*} above, but instead of putting two values in a pair by default, \lstinline{merge} takes a parameter called \lstinline{op}, which tells the associative operation of two results. The definition is straightforward:
\begin{lstlisting}
def merge[E, F, G, H[_,_]](op : F => G => H[F, G], x : (=> E) => F, y : (=> E) => G) : (=> E) => H[F, G] = e => op(x(e))(y(e))
\end{lstlisting}

%\begin{itemize}
%\item Fixpoints library + explaining delegation with some examples
%\item Alternative combinator + others
%\item Trait Composition to do Language Composition
%\end{itemize}

%In this section, we will further look into the mechanism behind our modular parsing library.
%We will first introduce some concepts independently from the process of parsing, including open recursion and Object Algebras,
%then present how they are integrated in our library.
%Starting from simple examples, a series of extensions and refactoring will be applied to illustrate
%how we achieve modularity in a type-safe way.

\subsection{Parsing with Open Recursion}\label{subsec:parsingwithopen}
Open recursion potentially opens the gate of extensibility for parsing, as a parser can now be open to additional features in the future.
Our exploration starts from a simple example, where a single parser only parses a free variable (i.e., a string). Traditionally, to write
such a parser, we need to import the library as
well as defining the datatype for which it produces.

\begin{lstlisting}
type Parser[E] = PackratParser[E]

abstract class Expr
case class Var(x : String) extends Expr
val pVar : Parser[Expr] = ident ^^ Var
\end{lstlisting}
Note that we are omitting lexing and demo code here. And we use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} throughout the paper. Furthermore, we would like the parser to support applications.
Not only the implementation of that parser is needed, but also its corresponding datatype:
\begin{lstlisting}
case class App(e1 : Expr, e2 : Expr) extends Expr
val pApp : Parser[Expr] = pVar ~ pVar ^^ { case e1 ~ e2 => App(e1, e2) }
\end{lstlisting}
Such a parser \lstinline{pApp} can only parse one-layer applications like \lstinline{"x y"}. Hence we revise it using recursion:
\begin{lstlisting}
val pApp : Parser[Expr] = {
    val p = pVar ||| pApp
    p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
}
\end{lstlisting}
Note that we are composing \lstinline{pVar} and \lstinline{pApp} using alternative. Alternation is a general operation to compose different parsers, since they represent different grammar rules with the logical relation ``or'' between one another. In this example, both sub-expressions of an application can be either a single variable or again an application. Now \lstinline{pApp} manages to parse an arbitrary number of applications like \lstinline{"x y z ..."} in a right-associative way.
Nevertheless, even though we are not considering operations but the creation of objects only, such an approach does not turn out to be a modular extension, because the grammar we want for an expression is actually:
\begin{lstlisting}
e ::= x | e ' ' e | ...
\end{lstlisting}
Note that such a recursive \lstinline{"e"} in the second case should not only include all the existing cases like \lstinline{pApp} and \lstinline{pVar}, but also take future extensions into account (the ellipsis we used above, for instance, \lstinline{e '+'} \lstinline{e}). This is where we are inspired to use open recursion. To define \lstinline{pApp} as an ``open'' function, we add a parameter \lstinline{p} to it, representing the explicit self-reference of our whole parser, open to future extensions. Note that \lstinline{p} should be defined as a by-name parameter using \lstinline{"=>"}.
\begin{lstlisting}
val pApp : (=> Parser[Expr]) => Parser[Expr] =
    p => p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
\end{lstlisting}
Here the type \lstinline{"(=>} \lstinline{Parser[Expr])} \lstinline{=>} \lstinline{Parser[Expr]"} is exactly what we want for a parser. Similarly \lstinline{pVar} is redefined also as a function using the explicit fix-point \lstinline{p}. To be more convenient for composition, we define an auxiliary combinator \lstinline{"|||"} for parsers of type \lstinline{"(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]"} with the same \lstinline{R} and \lstinline{E}. Such a function is overloaded and can be invoked as an infix operator when defined in an implicit class. It is really implemented using the general \lstinline{merge} function in Section~\ref{subsec:openrecursion}.
 \begin{lstlisting}
val pVar : (=> Parser[Expr]) => Parser[Expr] = p => ident ^^ { x => new Var(x) }

implicit class Combinator[E, R](x : (=> R) => Parser[E]) {
    type H[A, B] = Parser[E]
    def |||(y : (=> R) => Parser[E]) : (=> R) => Parser[E] = {
        val alt : Parser[E] => Parser[E] => Parser[E] = a => b => a ||| b
        merge[R, Parser[E], Parser[E], H](alt, x, y)
    }
}
\end{lstlisting}
Now since \lstinline{pVar} and \lstinline{pApp} are defined independently, we are able to compose them using \lstinline{|||}:
\begin{lstlisting}
val pVarApp : (=> Parser[Expr]) => Parser[Expr] = pVar ||| pApp
\end{lstlisting}
Here \lstinline{pVarApp} is the combination of parsers. But what we really want is something that has type \lstinline{Parser[Expr]}. One can be soon reminded that in Section~\ref{subsec:openrecursion}, we use \lstinline{fix} to get the fix-point of a function, in this case we write \lstinline{fix(pVarApp)}, which manages to parse an arbitrary number of applications.

This is the magic of open recursion: we can define as many small components as we like, and they are implemented as functions, with the help of explicit self-reference. Whenever we would like to close the recursion, simply use \lstinline{fix} for their combination. Such a process is guaranteed to terminate when we restrict the input for parsing to be finite, and we have some base cases like \lstinline{pVar}, which terminates recursion in their branches, in addition, with the help of Packrat parsers with direct left-recursion support.

Another thing which catches our attention is that, small parsers are combined using the alternative combinator. Hence, parsing failing at any position will try its next alternative, and for each recursion all the parsers are tried successively until one of them succeeds. It reveals its essence as a recursive descent parser. But it is quite easy and modular to use; with further extensions, they can be implemented separately and then appended to the combination of old ones.

Furthermore, in above code we use \lstinline{|||} for alternative instead of \lstinline{|}. As we have mentioned in Section~\ref{sec:packratparsers}, \lstinline{|||} is more powerful with longest matching. Consider if we obtain \lstinline{pVarApp} from \lstinline{pVar | pApp}, the final parser \lstinline{fix(pVarApp)} cannot even parse an application \lstinline{"x y"}, since it succeeds in parsing a variable first and only returns \lstinline{"x"}. With \lstinline{|||}, however, we can put composition in any order and it still works as expected.

But our story does not tend to stop here, as we have observed some limitations of the current approach. Specifically, one is that extensibility of data operations is not guaranteed. Suppose we want new operations like pretty-printing on expressions, we may 1) add a method \lstinline{print} to all case classes with their own interpretations; 2) define a single method \lstinline{print} and do pattern-matching on case classes inside. When adding new data variants and operations, both require existing code to be modified, thus lead to the lack of code reuse and modularity.

### 2016-11-09 paper/sections/PackratParsers.tex
Organize.

\section{Scala Parsing Library: Packrat Parsers}\label{sec:packratparsers}

We have selected Packrat Parsing library for our prototype implementation. It is a library built on Scala's basic parsing features,
which opens access to high performance in parsing, convenient parser combinators with concise syntax, as well as lexers for users. Generally,
a Packrat parser has type \lstinline{PackratParser[E]} for some \lstinline{E}, which indicates the representation. Below we present a small piece of
code that uses Packrat parsing for our illustration. Note that it should be defined in an enclosing type that extends \lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers} for lexing and \lstinline{scala.util.parsing.combinator.PackratParsers} for Packrat parsers.

\begin{lstlisting}
lexical.reserved += ("str")
lexical.delimiters += ("(", ")")

val p : PackratParser[String] =
    "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toString }
\end{lstlisting}

The first two lines stand for lexing. In the third line we define a parser which can, for instance, parse \lstinline{"str(5)"} and produce the string \lstinline{"5"}. In \lstinline{StandardTokenParsers} there are a number of basic parsers, such as \lstinline{numericLit} that parsers integers, and \lstinline{stringLit} that parsers string literals for convenient use. Moreover, in Scala the syntax for parsing is quite concise and clear, as we have shown above, with the help of parser combinators. \lstinline{~>} and \lstinline{<~} only collect results from the parts that arrows point to, and \lstinline{^^} is followed by a function which is applied to a parser to return new values. Finally, we can define a generic \lstinline{runParser} function for testing:

\begin{lstlisting}
def runParser(p: Parser[_]): String => Unit = in => {
    val t = phrase(p)(new lexical.Scanner(in))
    if (t.successful) println(t.get) else scala.sys.error(t.toString)
}
\end{lstlisting}
Now \lstinline{runParser(p)("str(5)")} prints out \lstinline{"5"} as expected.

\begin{table}[t]
\centering
\begin{tabular}{l}
\hline
\begin{lstlisting}
def ~[U](q: => Parser[U]): Parser[~[T, U]]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for sequential composition. \\
\hline
\begin{lstlisting}
def ^^[U](f: (T) => U): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for function application. \\
\hline
\begin{lstlisting}
def <~[U](q: => Parser[U]): Parser[T]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
\hline
\begin{lstlisting}
def ~>[U](q: => Parser[U]): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
\hline
\begin{lstlisting}
def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
\end{lstlisting} \\
\hspace{.2in}- A parser generator for interleaved repetitions. \\
\hline
\begin{lstlisting}
def ident: Parser[String]
\end{lstlisting} \\
\hspace{.2in}- A parser which matches an identifier. \\
\hline
\begin{lstlisting}
def numericLit: Parser[String]
\end{lstlisting} \\
\hspace{.2in}- A parser which matches a numeric literal. \\
\hline
\begin{lstlisting}
def |[U >: T](q: => Parser[U]): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for alternative composition. \\
\hline
\begin{lstlisting}
def |||[U >: T](q0: => Parser[U]): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for alternative with longest match composition. \\
\hline \\
\end{tabular}
\caption{AAA}\label{tab1}
\end{table}

Furthermore, Table~\ref{tab1} shows some common parser combinators that are used in our implementation.
In particular, \lstinline{|} is common syntax for alternative parsers in many languages, whereas \lstinline{|||} is a new combinator available in Scala, which is also for alternation but with longest match composition. Imagine that we are composing a list of parsers using alternative \lstinline{|}, in an order. A parser in the front might be successful in parsing a substring of the input, but then it unexpectedly stops the parsing and discards the rest, while another parser might be able to parse the whole input. In that case, using \lstinline{|||} we no longer need to worry about the order of composition, which potentially encourages its extensibility. A concrete example will be illustrated for this later.

On the other hand, the Scala Packrat Parsing library comes to our attention not only because of its performance, but also due to its support for direct left-recursion. Left recursion has been considered as a big problem especially in recursive descent parsers, which many existing libraries cannot perfectly support. From theoretical point of view, the algorithm behind Packrat parsers supports both direct and indirect left-recursion. The current version of the library is still buggy in some cases with indirect left-recursion, but we believe that they will fix it in the future, and with direct left-recursion it is already practical to use for a large number of applications.

### 2016-11-15 paper/sections/Library/DifferentSyntax.tex
Organize.
Some changes. Will reorganize some stuff in next commit.

@@
-\subsection{Object Algebras for Different Syntax}\label{subsec:differentsyntax}
+\subsection{Parsing with Multiple Syntax}\label{subsec:differentsyntax}
 
-Such extensibility becomes fragile when it comes to grammars with different syntax.
-Now suppose we would like to have types in the grammar.
-Firstly the new grammar is shown below, where the \lstinline{IntT} type and lambda expressions with annotation have been added:
+\haoyuan{Another way to organize the paper is to introduce the lack support of class approach for multiple different syntax, and directly extend var and app with lam.}
+
+Back to the class approach, the only little extensibility is just fragile when it comes to grammars with different syntax.
+Suppose we would like to have types in the expression language.
+Firstly the new grammar is shown below, where the \lstinline{Int} type and lambda expressions with annotation have been added to class \lstinline{Expr}:
 \begin{lstlisting}
@@
 Moreover, we need to update such a fix-point in the arguments of all previously defined parsers, so as to make parsers consistent for the combinator (things get even worse without that combinator).
-On the other hand, the modularity is quite restricted; only parsing and datatypes are extensible.
-Suppose now we want to realize some behaviors on those data structures, for example pretty-printing, then whenever we introduce a new behavior,
-it needs implementations in every class body, requiring existing code to be modified. At this point, Object Algebras are adopted for data structures
-instead.
 
-One intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
+One intuition of using Object Algebras, as illustrated above, is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
 We can simply add new variants corresponding to the new cases in the grammar, or realize different operations on structures, without modifying existing code, meanwhile separate compilation is supported.
 Another one is that Object Algebras benefit us from using generics, where we can abstract different syntax in the type parameters, and define less classes, simplifying code comparatively.
-Below we present the code using Object Algebras, where we start from \lstinline{varE} and \lstinline{appE} only, later we will extend them with \lstinline{intT} and \lstinline{lamE}:
+Below we present the code using Object Algebras, where we start from \lstinline{ExprAlg} (with \lstinline{varE} and \lstinline{appE} only), and define a new Object Algebra interface with \lstinline{intT} and \lstinline{lamE}:
 \begin{lstlisting}
-trait OldAlg[E] {
-    def varE(x : String) : E
-    def appE(e1 : E, e2 : E) : E
-}
-
-trait OldParser[E] {
-    val pE : OldAlg[E] => (=> Parser[E]) => Parser[E] = alg => p =>
-        ident ^^ { x => alg.varE(x) } |||
-        p ~ p ^^ { case e1 ~ e2 => alg.appE(e1, e2) }
-}
-\end{lstlisting}
-Here \lstinline{OldAlg} is an Object Algebra interface.
-An Object Algebra interface is more like a small language, where several cases (constructors) are put together.
-Since with Object Algebras structures are implicit, a parser of \lstinline{OldAlg} should generate ``structures'' with
-type \lstinline{"OldAlg[E] =>} \lstinline{E"} for any \lstinline{E}. In \lstinline{OldParser}, \lstinline{alg} of type
-\lstinline{OldAlg[E]} can be any algebra, which appears as a parameter of \lstinline{pE}, then in the body of \lstinline{pE} the algebra
-is invoked correspondingly right after parsing.
-It means we do not get any explicit objects generated from parsers, but they are immediately applied with algebras and return the results.
-To make \lstinline{alg} as a generic algebra, the enclosing trait of the parser, namely \lstinline{OldParser}, is generic.
-
-Now we move on to declare \lstinline{intT} and \lstinline{lamE} in a new trait:
-\begin{lstlisting}
-trait NewAlg[E, T] {
+trait LamAlg[E, T] {
     def intT() : T

### 2016-11-15 paper/sections/ObjectAlgebras.tex
Some changes. Will reorganize some stuff in next commit.

@@
 
-\subsection{Object Algebras}\label{sec:objectalgebras}
+\subsection{Object Algebras}\label{subsec:objectalgebras}
 Object Algebras captures a design pattern to address the Expression Problem nicely,
@@
 In Object Algebras, ASTs as recursive data structures are defined using traits, where each constructor corresponds
-to an abstract method inside. The recursive types are represented using generics. Here we show the example of a simple
-expression language:
+to an abstract method inside. The example from Section~\ref{subsec:parsingwithopen} is again used here for illustration.
+At first the language only supports variables and applications:
 \begin{lstlisting}
-trait ExpAlg[E] {
-    def lit(i : Int) : E
-    def add(e1 : E, e2 : E) : E
+trait ExprAlg[E] {
+    def varE(x : String) : E
+    def appE(e1 : E, e2 : E) : E
 }
 \end{lstlisting}
-The language has two constructors: literals and additions. The trait \lstinline{ExpAlg} is called an \textit{Object Algebra interface},
+The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
 whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
 type parameter \lstinline{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
-provides implementations for all cases. Below is an example of evaluation:
+provides implementations for all cases. Below is an example of collecting all free variables in an expression:
 \begin{lstlisting}
-trait Eval extends ExpAlg[Int] {
-    def lit(i : Int) = i
-    def add(e1 : Int, e2 : Int) = e1 + e2
+trait FreeVars extends ExprAlg[List[String]] {
+    def varE(x : String) = List(x)
+    def appE(e1 : List[String], e2 : List[String]) = e1 ++ e2
 }
 \end{lstlisting}
-Here \lstinline{Eval} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns an integer as the result.
+Here \lstinline{FreeVars} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a list of strings as the result.
 Furthermore, one can define a new trait to implement pretty-printing on expressions:
 \begin{lstlisting}
-trait Print extends ExpAlg[String] {
-    def lit(i : Int) = "" + i
-    def add(e1 : String, e2 : String) = e1 + " + " + e2
+trait ExprPrint extends ExprAlg[String] {
+    def varE(x : String) = x
+    def appE(e1 : String, e2 : String) = "(" + e1 + " " + e2 + ")"
 }
 \end{lstlisting}
-On the other hand, the data variants can be extended by inheriting \lstinline{ExpAlg} and adding new cases only. Suppose we want to
-have subtraction in expressions, a new Object Algebra interface \lstinline{SubAlg} is defined as follows:
+On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Suppose we want to
+have literals in expressions, a new Object Algebra interface \lstinline{LitAlg} is defined as follows:
 \begin{lstlisting}
-trait SubAlg[E] extends ExpAlg[E] {
-    def sub(e1 : E, e2 : E) : E
+trait LitAlg[E] extends ExprAlg[E] {
+    def litE(n : Int) : E
 }
 \end{lstlisting}
-Now evaluation on the new language can be realized by code reuse, and without modifying existing code:
+Now pretty-printing on the new language can be realized by code reuse, without modifying existing code:
 \begin{lstlisting}
-trait SubEval extends Eval {
-    def sub(e1 : Int, e2 : Int) = e1 - e2
+trait LitPrint extends ExprPrint {
+    def litE(n : Int) = n.toString
 }
 \end{lstlisting}
-On this, to create an expression of \lstinline{SubAlg}, a generic method is defined as follows:
+On this, to create an expression of \lstinline{LitAlg}, a generic method is defined as follows:
 \begin{lstlisting}
-def build[E](alg : SubAlg[E]) : E =
-    alg.sub(alg.add(alg.lit(2), alg.lit(3)), alg.lit(4))
+def build[E](alg : LitAlg[E]) : E =
+    alg.appE(alg.varE("x"), alg.litE(3))
 \end{lstlisting}
-Such a method implicitly represents the expression \lstinline{"2 + 3 - 4"}. The code
+Such a method implicitly represents the expression \lstinline{"x 3"}. The code
 \begin{lstlisting}
-build(new Eval(){})
+build(new LitPrint(){})
 \end{lstlisting}
-results in \lstinline{1}, as the result of evaluation. Now we observe that an ``object'' of expression actually
-has the function type \lstinline{SubAlg[E] =>} \lstinline{E} for generic \lstinline{E}. At this point, one may argue
+results in \lstinline{"(x 3)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
+has the function type \lstinline{LitAlg[E] =>} \lstinline{E} for generic \lstinline{E}. At this point, one may argue
 that it is illusory compared to traditional objects, and since we are struggling against the parsing problem, if a parser
@@
 there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
-in a very clever way. For example, if one wants to refactor an \lstinline{ExpAlg} expression to a \lstinline{SubAlg} expression by
-replacing all negative numbers with subtraction, a transformation algebra can be implemented as follows:
+in a very clever way. For example, if one wants to transform an \lstinline{ExprAlg} expression to a \lstinline{LitAlg} expression by
+replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:
 \begin{lstlisting}
-trait Refactor[E] extends ExpAlg[E] {
+trait Refactor[E] extends ExprAlg[E] {
     def alg : SubAlg[E]
-    def lit(i : Int) = if (i > 0) alg.lit(i) else alg.sub(alg.lit(0), alg.lit(-i))
-    def add(e1 : E, e2 : E) = alg.add(e1, e2)
+    def env : Map[String, Int]
+    def varE(x : String) = if (env.contains(x)) alg.litE(env(x)) else alg.varE(x)
+}
+\end{lstlisting}\haoyuan{Is this example OK?}
+When \lstinline{env = Map("y"} \lstinline{-> 3)}, the \lstinline{ExprAlg} expression \lstinline{"x y"} is changed to \lstinline{LitAlg} expression \lstinline{"x 3"}. The abstract method
+\lstinline{alg} in \lstinline{Refactor} can be any algebra of \lstinline{SubAlg}, simply for delegation. By passing a concrete implementation
+to \lstinline{alg}, \lstinline{Refactor} will be able to apply transformation to an \lstinline{ExprAlg} expression before \lstinline{alg} returns the result.
+
+\subsection{Parsing with Object Algebras}\label{subsec:parsingwithoa}
+
+Just as shown above, \lstinline{ExprAlg} is defined for a small language in an Object-Algebra style. Furthermore,
+A parser can be defined for this whole language at once.
+Since Object Algebras represent ``objects'' implicitly as functions like \lstinline{ExprAlg[E] =>} \lstinline{E} for generic \lstinline{E}, such a parser should consume an algebra from its parameter, then return a value of \lstinline{Fix[Parser[E]]} as in open recursion, where \lstinline{E} is abstract.
+\begin{lstlisting}
+trait ExprParser[E] {
+    val pE : ExprAlg[E] => Fix[Parser[E]] = alg => p =>
+        ident ^^ alg.varE |||
+        p ~ p ^^ { case e1 ~ e2 => alg.appE(e1, e2) }
 }
 \end{lstlisting}
-With \lstinline{Refactor}, the expression \lstinline{"1 + (-2)"} is refactored to \lstinline{"1 + (0 - 2)"}. The abstract method
-\lstinline{alg} in \lstinline{Refactor} can be any algebra of \lstinline{SubAlg}, simply for delegation. By passing a concrete implementation
-to \lstinline{alg}, \lstinline{Refactor} will be able to apply refactoring to an \lstinline{ExpAlg} expression before \lstinline{alg} returns the result.
+It is observed that \lstinline{pE} is in fact a parser generator, where \lstinline{alg} appears as a parameter, which can be any algebra, and after \lstinline{alg} is fed, it returns a \lstinline{Fix[Parser[E]]}.
+Again \lstinline{p} is the explicit self-reference of the open parser.
+Inside the body, the algebra is invoked correspondingly for all cases right after parsing.
+Note that the two cases are combined using the original \lstinline{|||} operator.
+To make \lstinline{alg} as a generic algebra, the enclosing trait of the parser, namely \lstinline{ExprParser}, is abstracted over \lstinline{E}.
+Behaviors, or algebras that can be fed to such a parser, are expected to be defined independently from the Object Algebra interface, just like the algebra \lstinline{ExprPrint} in Section~\ref{subsec:objectalgebras}.
+
+To explain the modularity, we again add literals to expressions, but instead of using inheritance as Section~\ref{subsec:objectalgebras}, \lstinline{LitAlg} is defined independently together with its own parser and algebra:
+\begin{lstlisting}
+trait LitAlg[E] {
+    def litE(n : Int) : E
+}
+
+trait LitPrint {
+    def litE(n : Int) = n.toString
+}
+
+trait LitParser[E] {
+    val pE : LitAlg[E] => Fix[Parser[E]] = alg => p =>
+        numericLit ^^ alg.litE
+}
+\end{lstlisting}
+In this case, \lstinline{LitAlg} is yet another small language, which is self-contained. To merge the two languages together, we require their grammars
+to be merged, and we obtain a combined language, whose parser integrates the two small parsers using alternative.
+In Scala we use compound types, namely use \lstinline{"with"} for two Object Algebra interface types, to avoid polluting the namespace.
+
+\begin{lstlisting}
+trait ExprLitParser[E] {
+    val pE : (ExprAlg[E] with LitAlg[E]) => Fix[Parser[E]] = alg => p => {
+        val pExprE = new ExprParser[E](){}.pE(alg)(p)
+        val pLitE = new LitParser[E](){}.pE(alg)(p)
+        pExprE ||| pLitE
+    }
+}
+\end{lstlisting}
+
+%But we can only abstract over the whole interface type, instead of working on its type parameters, as the number of type parameters is not fixed, where we will discuss more on multiple syntax later.
+In summary, \lstinline{"|||"} produces a combined parser, which takes a compound algebra as its parameter.
+
+In client code, a user can either define a new function with type \lstinline{(ExprAlg[E] with LitAlg[E]) =>} \lstinline{Parser[E]} based on \lstinline{pE}, which implies an algebra is applied after \lstinline{fix}, or simply feed an algebra to \lstinline{pE} at first, then obtain the result from \lstinline{fix}:
+\begin{lstlisting}
+trait ExprLitPrint extends ExprPrint with LitPrint
+val parsePrint = fix(new ExprLitParser[String](){}.pE(new ExprLitPrint(){}))
+val result = runParser(parsePrint)("x 3") // prints "(x 3)"
+\end{lstlisting}
+\haoyuan{Is there a way to avoid defining ExprLitPrint?}

### 2016-11-15 paper/sections/OpenRecursion.tex
Some changes. Will reorganize some stuff in next commit.

@@
     val p = pVar ||| pApp
-    p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
+    p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
 }
@@
 Now \lstinline{pApp} manages to parse an arbitrary number of applications like \lstinline{"x y z ..."} in a right-associative way.
-Nevertheless, even though we are not considering operations but the creation of objects only, such an approach does not turn out to be a modular extension, because the grammar we want for an expression is actually:
+Nevertheless, even though we are not considering operations but the creation of objects only, such an approach does not turn out to be a modular extension, because if numeric literals are further accepted in our small language, the grammar becomes:
 \begin{lstlisting}
-e ::= x | e ' ' e | ...
+e ::= x | e ' ' e | n
 \end{lstlisting}
-Note that such a recursive \lstinline{"e"} in the second case should not only include all the existing cases like \lstinline{pApp} and \lstinline{pVar}, but also take future extensions into account (the ellipsis we used above, for instance, \lstinline{e '+'} \lstinline{e}).
+Besides writing a new parser for literals, we need to rewrite the existing \lstinline{pApp}, since applications can now contain numbers, like \lstinline{"x 3"} and so on.
+\begin{lstlisting}
+case class Lit(n : Int) extends Expr
+val pLit : Parser[Expr] = numericLit ^^ Lit
+val pApp : Parser[Expr] = {
+    val p = pVar ||| pApp ||| pLit
+    p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
+}
+\end{lstlisting}
+We have observed that \lstinline{pApp} is indeed constructed in a non-modular way, because the \lstinline{p} standing for recursive parser needs to be updated on every extension.
+Actually in the grammar, the recursive \lstinline{"e"} in the second case should not only include all the existing cases like \lstinline{pApp} and \lstinline{pVar}, but also take future extensions into account (we can have arbitrarily more cases).
 This is where we are inspired to use open recursion.
@@
 \begin{lstlisting}
-val pApp : (=> Parser[Expr]) => Parser[Expr] =
-    p => p ~ p ^^ { case e1 ~ e2 => new App(e1, e2) }
+val pApp : Fix[Parser[Expr]] = p => p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
 \end{lstlisting}
-Here the type \lstinline{"(=>} \lstinline{Parser[Expr])} \lstinline{=>} \lstinline{Parser[Expr]"} is exactly what we want for a parser.
+Here the type \lstinline{Fix[Parser[Expr]]}, namely \lstinline{(=>} \lstinline{Parser[Expr])} \lstinline{=>} \lstinline{Parser[Expr]} is exactly what we want for a parser.
 Similarly \lstinline{pVar} is redefined also as a function using the explicit fix-point \lstinline{p}.
-To be more convenient for composition, we define an auxiliary combinator \lstinline{"|||"} for parsers of type \lstinline{"(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]"} with the same \lstinline{R} and \lstinline{E}.
+To be more convenient for composition, we define an auxiliary combinator \lstinline{"|||"} for parsers of type \lstinline{(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]} with the same \lstinline{R} and \lstinline{E}.
 Such a function is overloaded and can be invoked as an infix operator when defined in an implicit class.
@@
 \begin{lstlisting}
-val pVar : (=> Parser[Expr]) => Parser[Expr] = p => ident ^^ { x => new Var(x) }
+val pVar : Fix[Parser[Expr]] = p => ident ^^ Var
 
@@
 \begin{lstlisting}
-val pVarApp : (=> Parser[Expr]) => Parser[Expr] = pVar ||| pApp
+val pVarApp : Fix[Parser[Expr]] = pVar ||| pApp
 \end{lstlisting}
-Here \lstinline{pVarApp} is the combination of parsers.
+Here \lstinline{pVarApp} is the combination of two parsers.
 But what we really want is something that has type \lstinline{Parser[Expr]}.
@@
 But our story does not tend to stop here, as we have observed some limitations of the current approach.
-Specifically, one is that extensibility of data operations is not guaranteed.
+Specifically, one is that extensibility of data operations is not guaranteed, though we obtain the extensibility of data variants by defining an arbitrarily large number of new case classes.
 Suppose we want new operations like pretty-printing on expressions, we may 1) add a method \lstinline{print} to all case classes with their own interpretations; 2) define a single method \lstinline{print} and do pattern-matching on case classes inside.
-When adding new data variants and operations, both require existing code to be modified, thus lead to the lack of code reuse and modularity.
+When adding new data variants and operations, both require existing code to be modified, thus affect code reuse and modularity.

### 2016-11-15 paper/sections/ObjectAlgebras.tex
More.

@@
 
-%But we can only abstract over the whole interface type, instead of working on its type parameters, as the number of type parameters is not fixed, where we will discuss more on multiple syntax later.
-In summary, \lstinline{"|||"} produces a combined parser, which takes a compound algebra as its parameter.
-
 In client code, a user can either define a new function with type \lstinline{(ExprAlg[E] with LitAlg[E]) =>} \lstinline{Parser[E]} based on \lstinline{pE}, which implies an algebra is applied after \lstinline{fix}, or simply feed an algebra to \lstinline{pE} at first, then obtain the result from \lstinline{fix}:
@@
 \haoyuan{Is there a way to avoid defining ExprLitPrint?}
+
+\subsection{Parsing with Multiple Syntax}\label{subsec:differentsyntax}
+
+One intuition of using Object Algebras, as illustrated above, is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
+We can simply add new variants corresponding to the new cases in the grammar, or realize different operations on structures, without modifying existing code, meanwhile separate compilation is supported.
+Another one is that Object Algebras benefit us from using generics, where we can abstract multiple syntax in the type parameters, and define less classes, simplifying code comparatively.
+Below we refactor the last example in Section~\ref{subsec:parsingwithopen}, by defining a new Object Algebra interface with two cases \lstinline{intT} and \lstinline{lamE}.
+Meanwhile a parser is implemented for that language.
+\begin{lstlisting}
+trait LamAlg[E, T] {
+    def intT() : T
+    def lamE(x : String, t : T, e : E) : E
+}
+
+trait LamParser[E, T] {
+    type Pair = (Parser[E], Parser[T])
+    val pE : LamAlg[E, T] => (=> Pair) => Parser[E] = alg => p =>
+        ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1)
+            ^^ { case x ~ t ~ e => alg.lamE(x, t, e) }
+    val pT : LamAlg[E, T] => (=> Pair) => Parser[T] = alg => p =>
+        "Int" ^^ { _ => alg.intT() }
+}
+\end{lstlisting}
+\lstinline{LamParser} contains parsers for both expressions and types. From the code we observe that the multi-sorts of Object Algebra interface
+exactly express multiple syntax.
+At this point, to combine these two small components together we need some ``glue code''.
+But instead of writing composition by hand, we reimplement the combinator \lstinline{|||}, to compose one parser with type \lstinline{Alg1 =>} \lstinline{(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]} and another one with type \lstinline{Alg2 =>} \lstinline{(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]} using alternative.
+Note that we can only abstract over the whole algebra interface type, instead of working on its type parameters, as the number of type parameters is not fixed.
+In summary, \lstinline{"|||"} produces a combined parser, which takes a compound algebra as its parameter:
+\begin{lstlisting}
+implicit class Combinator[Alg1, Alg2, R, E](x : Alg1 => (=> R) => Parser[E]) {
+    def |||(y : Alg2 => (=> R) => Parser[E]): (Alg1 with Alg2) => (=> R) => Parser[E]
+        = alg => l => x(alg)(l) ||| y(alg)(l)
+}
+\end{lstlisting}
+Now it is sufficient to merge \lstinline{ExprParser} and \lstinline{LamParser} together:
+\begin{lstlisting}
+trait ExprLamParser[E, T] {
+    val pExpr = new ExprParser[E](){}
+    val pLam = new LamParser[E, T](){}
+    val pExprE : ExprAlg[E] => (=> pLam.Pair) => Parser[E] =
+        alg => p => pExpr.pE(alg)(p._1)
+
+    val pE = pExprE ||| pLam.pE
+    val pT = pLam.pT
+}
+\end{lstlisting}
+Parsers for expressions and types are connected using \lstinline{|||}, respectively. Hence the isolation for different syntax
+is guaranteed by the type system. The only left problem is that we still need to replace the parser in \lstinline{ExprParser} with \lstinline{pExprE}
+by updating the fix-point parameter. In the next section we will address this problem, and meanwhile our implementation enhances modularity in various aspects.

### 2016-11-15 paper/sections/OpenRecursion.tex
More.

@@
 To be more convenient for composition, we define an auxiliary combinator \lstinline{"|||"} for parsers of type \lstinline{(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]} with the same \lstinline{R} and \lstinline{E}.
+It is more general using \lstinline{R} instead of \lstinline{Parser[E]}, because there might be multiple fix-points in the parameter.
 Such a function is overloaded and can be invoked as an infix operator when defined in an implicit class.
-It is really implemented using the general \lstinline{merge} function in Section~\ref{subsec:openrecursion}.
+It is implemented using the general \lstinline{merge} function in Section~\ref{subsec:openrecursion}.
 \begin{lstlisting}
@@
 When adding new data variants and operations, both require existing code to be modified, thus affect code reuse and modularity.
+
+Another weakness is that the only little extensibility is just fragile when it comes to grammars with multiple different syntax.
+For instance, in the beginning we only have variables and applications in \lstinline{Expr}, to have types in our language, the \lstinline{Int} type and lambda expressions with annotation are added.
+See the new grammar below:
+\begin{lstlisting}
+e ::= x | e ' ' e | \x:t.e
+t ::= "Int"
+\end{lstlisting}
+
+We proceed to implement those parsers. An issue is that we need both a \lstinline{Parser[Expr]} and a \lstinline{Parser[Type]} in the fixpoint.
+Hence we use a pair in the parameter:
+\begin{lstlisting}
+type Pair = (Parser[Expr], Parser[Type])
+
+abstract class Type
+case class IntT extends Type
+val pInt : (=> Pair) => Parser[Type] =
+    p => "Int" ^^ { _ => IntT() }
+
+case class Lam(x : String, t : Type, e : Expr) extends Expr
+val pLam : (=> Pair) => Parser[Expr] =
+    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1) ^^
+        { case x ~ t ~ e => Lam(x, t, e) }
+
+val pVar2 : (=> Pair) => Parser[Expr] = p => pVar(p._1)
+val pApp2 : (=> Pair) => Parser[Expr] = p => pApp(p._1)
+
+val pET : Fix[Pair] = {
+    val pE = pVar2 ||| pApp2 ||| pLam
+    val pT = pInt
+    p => (pE(p), pT(p))
+}
+\end{lstlisting}
+Based on the old \lstinline{pVar} and \lstinline{pApp}, we implement \lstinline{pVar2} and \lstinline{pApp2} which take the pair as their fix-points instead.
+In that case the parsers are consistent, and hence can be composed using the combinator \lstinline{|||}.
+The last value \lstinline{pET} integrates the four single parsers defined before, where parsers for expressions and parsers for types are composed respectively.
+Furthermore, we use \lstinline{fix(p)} to obtain the pair of two parsers we want: one for expressions, the other for types.
+The parser \lstinline{fix(p)._1} parsers expressions like \lstinline{"\x:Int.x x"} successfully.
+Yet meanwhile, some drawbacks of this approach come into sight.
+With different syntax, it
+is necessary to extend the fix-point to contain more parsers, in which case usually some structures like tuples are used,
+making code tedious and less elegant.
+Moreover, we need to update such a fix-point in the arguments of all previously defined parsers, so as to make parsers consistent for the combinator (things get even worse without that combinator).
+It introduces more boilerplate code, which highly affects code reuse.

### 2016-11-15 paper/sections/ParsingLibrary.tex
More.

\section{Implementation}\label{sec:implementation}

\subsection{Fusion of Concepts}\label{subsec:fusion}

In the previous sections we have talked about Packrat Parsing, open recursion and Object Algebras. These components fuse together in our library implementation, for the initial goal, which is to build extensible parsers. We want to argue here that those aspects are totally orthogonal, and hence can have their alternatives. For example, Object Algebras are helpful to build extensible data structures, whereas in functional programming, there has been a large amount of related work on extensible datatypes, including DTC and MRM for the Haskell language. And a better
parsing library would potentially introduce a lot of fancy features and parser combinators, run with high efficiency, or even support indirect left-recursion. 
\haoyuan{Open recursion could also have alternatives?} For now, our implementation integrates the existing ones, and has turned out to be practical in various applications.

\subsection{Framework of a Language Component}\label{subsec:framework}

\subsection{More modularity}

### 2016-11-16 paper/paper.tex
Modification
Some text and half of the table
Text
Organize.
More.
Done text for methodology and library.

@@
 
+\lstdefinelanguage{PlainCode}{}
+
 \lstset{ %
@@
 showtabs=false,                 % show tabs within strings adding particular underscores
-morekeywords={var},
+morekeywords={var, lazy},
 %  frame=single,                   % adds a frame around the code
@@
 %\newcommand{\hl}[1]{\textcolor{red}{#1}}
-\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
+\newcommand\huang[1]{\authornote{huang}{blue}{#1}}
 \newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
@@
 
-\input{sections/Library/OpenRecursion.tex}
-\input{sections/Library/ObjectAlgebras.tex}
-\input{sections/Library/ParsingLibrary.tex}
-\input{sections/Library/FirstTry.tex} % mention again | and ||| with concrete examples. (pApp | pVar)
-\input{sections/Library/DifferentSyntax.tex}
+\input{sections/PackratParsers.tex}
+
+\input{sections/OpenRecursion.tex}
+
+\input{sections/ObjectAlgebras.tex}
+\input{sections/ParsingLibrary.tex}
 
@@
 % more subsections: lexing, traits, operations, language composition (more extensibility)
-
+% use paragraphs
 %...

### 2016-11-16 paper/sections/ObjectAlgebras.tex
Done text for methodology and library.

@@
 \begin{lstlisting}
-build(new LitPrint(){})
+build(new LitPrint{})
 \end{lstlisting}
@@
     val pE : (ExprAlg[E] with LitAlg[E]) => Fix[Parser[E]] = alg => p => {
-        val pExprE = new ExprParser[E](){}.pE(alg)(p)
-        val pLitE = new LitParser[E](){}.pE(alg)(p)
+        val pExprE = new ExprParser[E]{}.pE(alg)(p)
+        val pLitE = new LitParser[E]{}.pE(alg)(p)
         pExprE ||| pLitE
@@
 
-In client code, a user can either define a new function with type \lstinline{(ExprAlg[E] with LitAlg[E]) =>} \lstinline{Parser[E]} based on \lstinline{pE}, which implies an algebra is applied after \lstinline{fix}, or simply feed an algebra to \lstinline{pE} at first, then obtain the result from \lstinline{fix}:
+In client code, a user can either define a new function with type \lstinline{(ExprAlg[E] with LitAlg[E])} \lstinline{=>} \lstinline{Parser[E]} based on \lstinline{pE}, which implies an algebra is applied after \lstinline{fix}, or simply feed an algebra to \lstinline{pE} at first, then obtain the result from \lstinline{fix}:
 \begin{lstlisting}
-trait ExprLitPrint extends ExprPrint with LitPrint
-val parsePrint = fix(new ExprLitParser[String](){}.pE(new ExprLitPrint(){}))
+val parsePrint = fix(new ExprLitParser[String]{}.pE(new ExprPrint with LitPrint{}))
 val result = runParser(parsePrint)("x 3") // prints "(x 3)"
 \end{lstlisting}
-\haoyuan{Is there a way to avoid defining ExprLitPrint?}
 
@@
     val pT : LamAlg[E, T] => (=> Pair) => Parser[T] = alg => p =>
-        "Int" ^^ { _ => alg.intT() }
+        "Int" ^^^ alg.intT
 }
 \end{lstlisting}
-\lstinline{LamParser} contains parsers for both expressions and types. From the code we observe that the multi-sorts of Object Algebra interface
+Note that we are omitting its lexer here. \lstinline{LamParser} contains parsers for both expressions and types. From the code we observe that the multi-sorts of Object Algebra interface
 exactly express multiple syntax.
@@
 trait ExprLamParser[E, T] {
-    val pExpr = new ExprParser[E](){}
-    val pLam = new LamParser[E, T](){}
+    val pExpr = new ExprParser[E]{}
+    val pLam = new LamParser[E, T]{}
     val pExprE : ExprAlg[E] => (=> pLam.Pair) => Parser[E] =

### 2016-11-16 paper/sections/ParsingLibrary.tex
Done text for methodology and library.

@@
 
+The library is made up of around twenty lines of code only, as shown in Figure~\ref{fig:sourcelibrary}.
+It consists of some types wrapped up by synonyms, basic functions \lstinline{fix} and \lstinline{runParser}, and the combinator \lstinline{|||} from Section~\ref{subsec:differentsyntax}.
+
+\begin{figure}[htbp]
+\centering
+\begin{lstlisting}
+import scala.util.parsing.combinator._
+import scala.util.parsing.combinator.syntactical._
+
+object Library extends StandardTokenParsers with PackratParsers {
+    type Fix[T] = (=> T) => T
+    type Parser[T] = PackratParser[T]
+    type OpenParser[Alg, List, T] = Alg => (=> List) => Parser[T]
+
+    def fix[A](f: Fix[A]): A = { lazy val a: A = f(a); a }
+    def runParser(p : Parser[_]) : String => Unit = in => {
+        phrase(p)(new lexical.Scanner(in)) match {
+            case t if t.successful => println(t.get)
+            case t                 => scala.sys.error(t.toString)
+        }
+    }
+
+    implicit class Combinator[Alg1, Alg2, R, E](x : Alg1 => (=> R) => Parser[E]) {
+        def |||(y : Alg2 => (=> R) => Parser[E])
+            : Alg1 with Alg2 => (=> R) => Parser[E]
+                = alg => l => x(alg)(l) ||| y(alg)(l)
+    }
+}
+\end{lstlisting}
+\caption{Source code of our library.}\label{fig:sourcelibrary}
+\end{figure}
+
 \subsection{Framework of a Language Component}\label{subsec:framework}
 
+Having pointed out that our goal is to realize modular and extensible parsers, we recommend users to write code in a modular and
+organized way. The framework we use for a language, or rather a language component in our case study, consists of the following three
+parts:
+\begin{itemize}
+\item \textbf{Object Algebra interface:} defined as a trait with several cases, corresponding to the grammar rules.
+\item \textbf{Parser:} implementation of the parser, using open recursion and Object Algebras, with its scoping trait abstracted over the type parameters.
+\item \textbf{Algebra:} some behaviors (operations) on the data structure, such as pretty-printing and so on.
+\end{itemize}
+
+Figure~\ref{fig:objectexpr} shows how to encapsulate \lstinline{ExprAlg} and its related stuff in object \lstinline{Expr} using our library.
+Note that we use type bounds to address the issue remained in Section~\ref{subsec:differentsyntax}.
+\lstinline{List[E]} is a type synonym for a record type, while the type bound \lstinline{L} \lstinline{<:} \lstinline{List[E]} in \lstinline{AlgParser} states that the fix-point of our parser is a record, which contains at least \lstinline{pE} for parsing expressions.
+Hence it saves us from updating the type of fix-points all the time.
+
+\begin{figure}[htbp]
+\centering
+\begin{lstlisting}
+object Expr {
+    trait Alg[E] {
+        def varE(x : String) : E
+        def appE(e1 : E, e2 : E) : E
+    }
+
+    type List[E] = { val pE : Parser[E] }
+
+    trait AlgParser[E, L <: List[E]] {
+        val pExprE : OpenParser[Alg[E], L, E] = alg => p => {
+            lazy val pE = p.pE
+            ident ^^ alg.varE |||
+            pE ~ pE ^^ { case e1 ~ e2 => alg.appE(e1, e2) }
+        }
+    }
+
+    trait Print extends Alg[String] {
+        def varE(x : String) = x
+        def appE(e1 : String, e2 : String) = "(" + e1 + " " + e2 + ")"
+    }
+}
+\end{lstlisting}
+\caption{Object \lstinline{Expr} for expressions.}\label{fig:objectexpr}
+\end{figure}
+
+Now \lstinline{Expr} turns out to be a self-contained small language with parsing and pretty-printing.
+To construct a different language component, we implement \lstinline{LamAlg} in Figure~\ref{fig:objectlam}, where we declare a different \lstinline{List} for multiple syntax.
+
+\begin{figure}[htbp]
+\centering
+\begin{lstlisting}
+object Lam {
+    trait Alg[E, T] {
+        def intT() : T
+        def lamE(x : String, t : T, e : E) : E
+    }
+
+    type List[E, T] = { val pE : Parser[E]; val pT : Parser[T] }
+
+    trait AlgParser[E, T, L <: List[E, T]] {
+        lexical.reserved += ("Int")
+        lexical.delimiters += ("\\", ":", ".")
+        val pLamE : OpenParser[Alg[E, T], L, E] = alg => p =>
+            ("\\" ~> ident) ~ (":" ~> p.pT) ~ ("." ~> p.pE) ^^
+                { case x ~ t ~ e => alg.lamE(x, t, e)}
+        val pLamT : OpenParser[Alg[E, T], L, T] = alg => p =>
+            "Int" ^^^ alg.intT
+    }
+
+    trait Print extends Alg[String, String] {
+        def intT() = "Int"
+        def lamE(x : String, t : String, e : String) = "\\" + x + ":" + t + "." + e
+    }
+}
+\end{lstlisting}
+\caption{Object \lstinline{Lam} for lambdas.}\label{fig:objectlam}
+\end{figure}
+
 \subsection{More modularity}
+
+We have optimized the framework for not only modular parsers, but also languages, and that is the reason why we call them language components. To achieve
+higher-order modularity, it is necessary to follow the structure when composing two small components.
+Figure~\ref{fig:objectlamexpr} presents the combination of \lstinline{Expr} and \lstinline{Lam}, as a new language, which is again modular for future compositions.
+Note that the \lstinline{AlgParser} extends both parent parsers, so as to reuse their parser functions, as well as composing lexers automatically.
+
+\begin{figure}[htbp]
+\centering
+\begin{lstlisting}
+object LamExpr {
+    type List[E, T] = Lam.List[E, T]
+    type Alg[E, T] = Expr.Alg[E] with Lam.Alg[E, T]
+    trait Print extends Expr.Print with Lam.Print
+
+    trait AlgParser[E, T, L <: List[E, T]] extends Expr.AlgParser[E, L]
+            with Lam.AlgParser[E, T, L] {
+        val pLamExprE = pExprE ||| pLamE
+        val pLamExprT = pLamT
+    }
+}
+\end{lstlisting}
+\caption{Object \lstinline{LamExpr}, composing \lstinline{Lam} and \lstinline{Expr} together.}\label{fig:objectlamexpr}
+\end{figure}
+
+Below we give an example of client code for testing \lstinline{LamExpr}. \lstinline{parse} is a generic parser generator, which takes an algebra.
+In \lstinline{parseAndPrint} we pass the printing algebra to \lstinline{parse}, so that it parses the input and apply pretty-printing, returning the result as expected.
+
+\begin{lstlisting}
+class List[E, T](pe : Parser[E], pt : Parser[T]) {
+    val pE = pe; val pT = pt
+}
+def parse[E, T](inp: String)(alg: LamExpr.Alg[E, T]) = {
+    def parser : Fix[List[E, T]] = l => {
+        val lang = new LamExpr.AlgParser[E, T, List[E, T]] {}
+        new List[E, T](lang.pLamExprE(alg)(l), lang.pLamExprT(alg)(l))
+    }
+    runParser(fix(parser).pE)(inp)
+}
+def parseAndPrint(inp: String) = parse(inp)(new LamExpr.Print {})
+
+val result = parseAndPrint("\\x:Int.x y") // prints "\\x:Int.(x y)"
+\end{lstlisting}

### 2016-11-16 paper/sections/RelatedWork.tex
Brainstorming related work.

\section{Related Work}\label{sec:relatedwork}

- extensible parsing, language workbenches: rats, modular semantic actions, (syntactic modularity, no separate compilation, modular type-checking)
(read more papers, see if they talk about this issue, some potential solutions)

- parser combinators for type-checking, previous work has not shown how to support modularity (ASTs); left-recursion and back-tracking in related techniques

- modularity: object algebras, dtc and mrm (problem with parsing, is there any related work? (PB: a paper on unfolds: build the AST))

(parsing in Javascript: using delegation, does it support modular AST)

### 2016-11-16 paper/sections/RelatedWork.tex
More.

@@
 
-- extensible parsing, language workbenches: rats, modular semantic actions, (syntactic modularity, no separate compilation, modular type-checking)
+- extensible parsing, language workbenches: rats, noa (this one already uses OA), modular semantic actions, (syntactic modularity, no separate compilation, modular type-checking)
 (read more papers, see if they talk about this issue, some potential solutions)
@@
 (parsing in Javascript: using delegation, does it support modular AST)
+
+noa, shy: shy: only override some interesting cases (transformation is tedious)
+bruijn indices: parsing + transformation

### 2016-11-23 paper/sections/Overview.tex
Text

@@
 
-Our modular parsing library consists of four parts, and we will go through them in this section.
-The first is underlying parsing technique we used.
-Parsing has been heavily studied over the years, especially for context-free grammars and its subsets.
+\subsection{The Choice of Parsing Techniques}\label{subsec:overview-parsing}
+
+Parsing techniques have been heavily studied over the years, especially for context-free grammars and its subsets.
 However, the choice of parsing techniques is not a minor issue here, because of modularity.
@@
 Although every parsing algorithm is able to implement by hand in principle, some of them are not friendly for that.
-A parser generator is a tool to generate parsers automatically or semi-automatically, from a grammar given by the user.
+A \textit{parser generator} is a tool to generate parsers automatically or semi-automatically, from a grammar given by the user.
 There is no restriction on the algorithm, however most of them adopts LL and LR parsing algorithms, which are table-based.
@@
 
-Comparing with the parser generators, a parser combinator takes several parsers and produce a new parser as its output.
+Comparing with the parser generators, a \textit{parser combinator} takes several parsers and produce a new parser as its output.
 It is popular in functional programming, where the parsers are represented by functions and parser combinators are higher-order functions accepts them.
@@
 In brief, Packrat parsers use memorization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated, and it supports left-recursive grammars.
-These properties are very suitable for modularity.
-Thus we use Packrat parsers as the underlying parsing technique in our library.
+All of these properties are very suitable for modularity, thus we use Packrat parsers as the underlying parsing technique.
+In section \ref{sec:packratparsers}, we will talk more details about the Scala Packrat parsing library we used.
 
@@
 
-The second part of our library is open recursion.
-It is introduced to deal with the problem of recursive calls in modular parsing.
-We use a simple example to demonstrate the problem, which is parsing a language of natural numbers.
-The language is defined below.
+\subsection{Modular Parsers, the First Look}\label{subsec:overview-firstlook}
+
+With Packrat parser combinators, we obtain the possibility of writing parsers in a modular way.
+Let's start by looking at how to build a parser for a simple language, which only has variables and applications at the beginning.
 
 \begin{lstlisting}[language=PlainCode]
-Expr := zero | succ Expr
+Expr := Var String | App Expr Expr
 \end{lstlisting}
 
-Firstly, we need to define the abstract syntax. Here we use Scala's case classes to save spaces, this can also be done using class inheritance.
+Firstly, we define the abstract syntax by Scala's convenient case classes. Expressions have type \lstinline{Expr}, and each of them is either \lstinline{Var} or \lstinline{App}.
 
@@
 abstract class Expr
-case class Zero extends Expr
-case class Succ(v: Expr) extends Expr
+case class Var(x: String) extends Expr
+case class App(e1: Expr, e2: Expr) extends Expr
 \end{lstlisting}
 
-And then we write parsing functions for every cases.
-Some irrelevant details in the code are omitted to outline the main structure.
-Suppose we have \lstinline{choice} combinator for alternatives and the input tokens are global values which can be accessed by all these functions.
+And then we write parsing functions for each case.
 
 \begin{lstlisting}
-def parseExpr(): Expr = choice(parserZero, parseSucc)
+val pVar: PackratParser[Expr] =
+  ident ^^ Var
 
-def parserZero(): Expr = {...}
+val pApp: (=> PackratParser[Expr]) => PackratParser[Expr] =
+  e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
+\end{lstlisting}
 
-def parseSucc(): Expr = {
-  parseToken('succ')
-  val v = parseExpr()
-  Succ(v)
+The \lstinline{PackratParser} type and the combinators are imported from the Scala Packrat parsing library.
+Some details about these structures will be discussed in later sections, so that readers could only focus on the types of these functions at the first reading.
+
+\lstinline{pVar} is straight forward.
+It parses a single identifier, and its type \lstinline{PackratParser[Expr]} indicates the result is an \lstinline{Expr}.
+However the type of \lstinline{pApp} is different.
+It takes a parser of \lstinline{Expr} as an argument, which represents the parser of the whole expression.
+In this particular case, it is essentially the parser of both two cases.
+We can write equivalent code as below.
+
+\begin{lstlisting}
+val pApp2: PackratParser[Expr] = {
+  lazy val e = pApp2 | pVar
+  e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
 }
@@
 
-It works well, but later we want to extend the language to support addition.
+\lstinline{pApp2} works fine, but it does not have the extensibility as \lstinline{pApp} does.
+Since the parser of the whole expression in \lstinline{pApp2} is hard-coded, we have to edit the existing code if we extend \lstinline{Expr} to add a new case later.
+The parameterization technique used in \lstinline{pApp} is called \textit{open recursion}, which is discussed in section \ref{sec:openandparsing}.
+
+We use a trait to glue them together, and add \lstinline{pVarApp} as a combination of the two parsing functions. This piece of code is a modular parser.
+
+\begin{lstlisting}
+trait ExprParser {
+  lexical.delimiters += ("(", ")")
+  val pVar: PackratParser[Expr] =
+    ident ^^ Var
+  val pApp: (=> PackratParser[Expr]) => PackratParser[Expr] =
+    e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
+  val pVarApp: (=> PackratParser[Expr]) => PackratParser[Expr] =
+    e => pApp(e) | pVar | "(" ~> e <~ ")"
+}
+\end{lstlisting}
+
+Finally, we are able to parse \lstinline{Expr} frome strings as shown in the code below.
+It will print \lstinline{App(Var(x),Var(y))}.
+The \lstinline{fix} function here is just the standard \textit{fixpoint}, it is explained in section \ref{subsec:openrecursion}.
+
+\begin{lstlisting}
+def parseAndPrint(inp: String) = {
+  val p = fix(new ExprParser {}.pVarApp)
+  val t = phrase(p)(new lexical.Scanner(inp))
+  if (t.successful) println(t.get) else scala.sys.error(t.toString)
+}
+
+parseAndPrint("x y")
+\end{lstlisting}
+
+\subsection{Extend and Reuse Parsers}\label{subsec:overview-extend}
+
+We have done our modular parser, and it behaves just like normal parsers.
+To demonstrate the advantages of writing parsers in that way, we extend the language to add lambda abstractions as a new case of expressions.
+The language then actually becomes the famous \textit{untyped lambda calculus}.
 
 \begin{lstlisting}[language=PlainCode]
-Expr := zero | succ Expr | add Expr Expr
+Expr := Var String | App Expr Expr | Lam String Expr
 \end{lstlisting}
 
-Extending the abstract syntax is easy.
+With Scala's case classes, extending the abstract syntax \lstinline{Expr} is easy.
 
 \begin{lstlisting}
-case class Add(a: Expr, b: Expr) extends Expr
+case class Lam(x: String, e: Expr) extends Expr
 \end{lstlisting}
+
+Since we already have the parser for variables and applications, we can reuse it and build a new parser by extending it.
+Notice that the old code is not touched, hence separate compilation is obtained.
+
+\begin{lstlisting}
+trait LambdaParser extends ExprParser {
+  lexical.delimiters += ("\\", ".")
+  val pLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
+    e => "\\" ~> ident ~ ("." ~> e) ^^ { case x ~ b => Lam(x, b) }
+  val pVarAppLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
+    e => pLam(e) | pVarApp(e)
+}
+\end{lstlisting}
+
+In the code above, we write a parsing function \lstinline{pLam} for the lambda case. And compose it with \lstinline{pVarApp} we have in the old parser.
 

### 2016-11-24 paper/sections/Overview.tex
Text

@@
 Parsing techniques have been heavily studied over the years, especially for context-free grammars and its subsets.
-However, the choice of parsing techniques is not a minor issue here, because of modularity.
-To achieve modularity, the parsers must be written in a way that can be composed, and different parsing techniques have different suitability for that.
+However, the choice of parsing techniques is not a minor issue in a modular setting.
+To achieve type-safe modularity, the parsers must be able to be extended and composed, and different parsing techniques have different suitability for that.
 
-Although every parsing algorithm is able to implement by hand in principle, some of them are not friendly for that.
-A \textit{parser generator} is a tool to generate parsers automatically or semi-automatically, from a grammar given by the user.
-There is no restriction on the algorithm, however most of them adopts LL and LR parsing algorithms, which are table-based.
+Although every parsing algorithm can be implemented by hand in principle, some of them are not friendly for that.
+A \textit{parser generator} is a tool to help users generate parsers automatically or semi-automatically from a given grammar.
+There is no restriction on the algorithm, while most of them adopts table-based LL and LR parsing algorithms.
 
-\huang{TALK ABOUT PREVIOUS WORKS USING PARSER GENERATORS, AND ALSO THE DISADVTANGES OF THEM}
+\huang{a paragraph here, refer to related work section, discuss previous works of modular parsing using parser generators and its drawbacks}
 
@@
 It is popular in functional programming, where the parsers are represented by functions and parser combinators are higher-order functions accepts them.
+
 Parser combinators are very suitable for our purpose, because of two reasons.
@@
 
-There are variants in the category of parser combinators.
-Almost all of them adopt top-down, recursive descent parsing, which uses backtracking to search through the possible branches.
-Simple backtracking parser combinators, such as the famous library Parsec in Haskell, have some drawbacks.
+Parser combinators adopt top-down, recursive descent parsing, which uses backtracking to search through the possible branches.
+However, simple backtracking parser combinators such as the famous library Parsec in Haskell, have some drawbacks.
 One is that they cannot support left-recursive grammars.
-The common solution is to rewrite a left-recursive grammar into an equivalent one, so called left-recursion elimination.
-This solution requires information of the whole grammar, so that it cannot be applied in a modular setting because we only have parts of the grammar.
+The common solution is to rewrite a left-recursive grammar into an equivalent one, so called \textit{left-recursion elimination}.
+This solution requires information of the whole grammar, so that it cannot be applied in a modular setting.
+Because the grammar may be extended in the future, essentially we only have parts of the grammar.
 Another drawback is performance.
@@
 All of these properties are very suitable for modularity, thus we use Packrat parsers as the underlying parsing technique.
-In section \ref{sec:packratparsers}, we will talk more details about the Scala Packrat parsing library we used.
+In section \ref{sec:packratparsers}, we will talk about the Packrat parsing library implemented in Scala, which we use.
 
@@
 However the type of \lstinline{pApp} is different.
-It takes a parser of \lstinline{Expr} as an argument, which represents the parser of the whole expression.
-In this particular case, it is essentially the parser of both two cases.
+It takes a parser of \lstinline{Expr} as an argument, which represents the parser of the whole expression, namely \lstinline{e} here.
+And use it to parse two consecutive expressions as an application.
+In this particular case, the parser of the whole expression is in fact the combination of \lstinline{pVar} and \lstinline{pApp}.
 We can write equivalent code as below.
@@
 \lstinline{pApp2} works fine, but it does not have the extensibility as \lstinline{pApp} does.
-Since the parser of the whole expression in \lstinline{pApp2} is hard-coded, we have to edit the existing code if we extend \lstinline{Expr} to add a new case later.
+Since the parser of the whole expression in \lstinline{pApp2} is hard-coded, it covers and only covers the variable and application cases.
+If we we extend \lstinline{Expr} later to add a third case lambda abstraction, we have to edit the existing code.
 The parameterization technique used in \lstinline{pApp} is called \textit{open recursion}, which is discussed in section \ref{sec:openandparsing}.
 
-We use a trait to glue them together, and add \lstinline{pVarApp} as a combination of the two parsing functions. This piece of code is a modular parser.
+We can use a trait to glue them together, and define \lstinline{pVarApp} as a combination of the two parsing functions. A type synonym \lstinline{Fix} is used to shorten the types.
 
 \begin{lstlisting}
+type Fix[T] = (=>T) => T
+
 trait ExprParser {
@@
     ident ^^ Var
-  val pApp: (=> PackratParser[Expr]) => PackratParser[Expr] =
+  val pApp: Fix[PackratParser[Expr]] =
     e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
-  val pVarApp: (=> PackratParser[Expr]) => PackratParser[Expr] =
+  val pVarApp: Fix[PackratParser[Expr]] =
     e => pApp(e) | pVar | "(" ~> e <~ ")"
@@
 Finally, we are able to parse \lstinline{Expr} frome strings as shown in the code below.
-It will print \lstinline{App(Var(x),Var(y))}.
 The \lstinline{fix} function here is just the standard \textit{fixpoint}, it is explained in section \ref{subsec:openrecursion}.
@@
 \begin{lstlisting}
-def parseAndPrint(inp: String) = {
-  val p = fix(new ExprParser {}.pVarApp)
-  val t = phrase(p)(new lexical.Scanner(inp))
+def parseAndPrint(p: Fix[PackratParser[Expr]])(inp: String) = {
+  val t = phrase(fix(p))(new lexical.Scanner(inp))
   if (t.successful) println(t.get) else scala.sys.error(t.toString)
@@
 
-parseAndPrint("x y")
+val p = new ExprParser {}.pVarApp
+parseAndPrint(p)("x y")
 \end{lstlisting}
+
+It will print \lstinline{App(Var(x),Var(y))} as the parsing result of string \lstinline{"x y"}.
 
@@
 
-We have done our modular parser, and it behaves just like normal parsers.
-To demonstrate the advantages of writing parsers in that way, we extend the language to add lambda abstractions as a new case of expressions.
-The language then actually becomes the famous \textit{untyped lambda calculus}.
+We have done our modular parser, and it behaves just in a normal way.
+To demonstrate the advantages of writing modular parsers, we extend the language to add lambda abstractions as a new case of expressions.
+The language then becomes the famous \textit{untyped lambda calculus}.
 
@@
 
-Since we already have the parser for variables and applications, we can reuse it and build a new parser by extending it.
+Since we already have the parser for variables and applications, we want to reuse it and build a new parser by extending it.
+It is just what the code below does.
 Notice that the old code is not touched, hence separate compilation is obtained.
@@
 
-In the code above, we write a parsing function \lstinline{pLam} for the lambda case. And compose it with \lstinline{pVarApp} we have in the old parser.
+In the new parser \lstinline{LambdaParser} extended from the old one, we write a new parsing function \lstinline{pLam} for the new case lambda abstraction.
+And compose it with \lstinline{pVarApp}, which we already have in the old parser \lstinline{ExprParser}.
+Open recursion is critical to achieve such type-safe compositions, because wherever we refer to the parser of the whole expression, it is updated to cover all of the three case we have now.
 
-\begin{itemize}
-\item Choosing the Parsing Technology
-    \begin{itemize}
-    \item Parser Generators : why not?
-        \begin{itemize}
-            \item Not type-safe
-            \item No modular type-checking
-            \item Not modular or no separate compilation (but we need to mention lots of work on extensible parsing here: example Language Workbenches; Rats)
-        \end{itemize}
-    \item Parser Combinators
-        \begin{itemize}
-        \item Backtracking parsers (Parsec)
-            \begin{itemize}
-            \item Need to remove left recursion (Problem: Transformation is not modular; if we do not know the full grammar then cannot be done)
-            \item Need try/backtracking (Problem: Since we do not know the full set of rules in a modular setting, we have to assume worst case scenario and add redundant backtracking. )
-            \item Mention some possible workarounds (there may be some but they still have issues)
-            \end{itemize}
-        \item (Non)-Backtracking/Packrat Parsers (Works well in a modular setting) (find a good name for this type of parsers?)
-        \end{itemize}
-    \end{itemize}
-\item Adapting Parser Combinators for modularity
-    \begin{itemize}
-    \item Using Object Algebras
-    \item Using Open Recursion
-    \item Using new modular parser combinators (Library: new alternative combinator, for example)
-    \end{itemize}
-\item Small example
-    \begin{itemize}
-    \item small example of a modular parser using our technique
-    \item show also the same example without modularity and write a detailed comparison
-    \item Lambda Calculus is a good candidate for the example
-    \item Show a small extension (adding plus and numeric literals)
-    \end{itemize}
-\end{itemize}
+\begin{lstlisting}
+val p = new LambdaParser {}.pVarAppLam
+parseAndPrint(p)("(\\x.x) y")
+\end{lstlisting}
+
+The code above will print result \lstinline{App(Lam(x,Var(x)),Var(y))} from parsing \lstinline{"(\\x.x)} \lstinline{y"}.
+
+\subsection{Object Algebras and More}\label{subsec:overview-oa}
+
+Instead of define the abstract syntax using Scala's case classes, we can use \textit{Object Algebras} for more extensibility.
+Object Algebras is a technique to solve the \textit{Expression Problem}, providing the possibility of extending both data variants and operations over them.
+It will be further discussed in section \ref{sec:algebrasandparsing}.
+
+Using Object Algebras, the abstract syntax of the language of variables and applications is defined as below.
+
+\begin{lstlisting}
+trait VarApp[E] {
+  def Var(x: String): E
+  def App(e1: E, e2: E): E
+}
+\end{lstlisting}
+
+% \begin{itemize}
+% \item Choosing the Parsing Technology
+%     \begin{itemize}
+%     \item Parser Generators : why not?
+%         \begin{itemize}
+%             \item Not type-safe
+%             \item No modular type-checking
+%             \item Not modular or no separate compilation (but we need to mention lots of work on extensible parsing here: example Language Workbenches; Rats)
+%         \end{itemize}
+%     \item Parser Combinators
+%         \begin{itemize}
+%         \item Backtracking parsers (Parsec)
+%             \begin{itemize}
+%             \item Need to remove left recursion (Problem: Transformation is not modular; if we do not know the full grammar then cannot be done)
+%             \item Need try/backtracking (Problem: Since we do not know the full set of rules in a modular setting, we have to assume worst case scenario and add redundant backtracking. )
+%             \item Mention some possible workarounds (there may be some but they still have issues)
+%             \end{itemize}
+%         \item (Non)-Backtracking/Packrat Parsers (Works well in a modular setting) (find a good name for this type of parsers?)
+%         \end{itemize}
+%     \end{itemize}
+% \item Adapting Parser Combinators for modularity
+%     \begin{itemize}
+%     \item Using Object Algebras
+%     \item Using Open Recursion
+%     \item Using new modular parser combinators (Library: new alternative combinator, for example)
+%     \end{itemize}
+% \item Small example
+%     \begin{itemize}
+%     \item small example of a modular parser using our technique
+%     \item show also the same example without modularity and write a detailed comparison
+%     \item Lambda Calculus is a good candidate for the example
+%     \item Show a small extension (adding plus and numeric literals)
+%     \end{itemize}
+% \end{itemize}

### 2016-11-24 paper/sections/Overview.tex
Text

@@
   lexical.delimiters += ("(", ")")
+
   val pVar: PackratParser[Expr] =
@@
 
-val p = new ExprParser {}.pVarApp
-parseAndPrint(p)("x y")
+parseAndPrint(new ExprParser {}.pVarApp)("x y")
 \end{lstlisting}
@@
   lexical.delimiters += ("\\", ".")
+
   val pLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
@@
 \begin{lstlisting}
-val p = new LambdaParser {}.pVarAppLam
-parseAndPrint(p)("(\\x.x) y")
+parseAndPrint(new LambdaParser {}.pVarAppLam)("(\\x.x) y")
 \end{lstlisting}
@@
 \begin{lstlisting}
-trait VarApp[E] {
+trait Alg[E] {
   def Var(x: String): E
@@
 }
+\end{lstlisting}
+
+Then we are able to define operations over the syntax in an extensible way. Here we write a pretty printing operation for this language.
+
+\begin{lstlisting}
+trait Print extends Alg[String] {
+  def Var(x: String) = x
+  def App(e1: String, e2: String) = "(" + e1 + " " + e2 + ")"
+}
+\end{lstlisting}
+
+And also parser for it as below.
+Notice the parsing function \lstinline{pVarApp} takes an argument of type \lstinline{Alg[E]}, that means it accepts any instance of the algebra interface \lstinline{Alg}.
+
+\begin{lstlisting}
+trait Parser[E] {
+  lexical.delimiters += ("(", ")")
+
+  val pVarApp: Alg[E] => Fix[PackratParser[E]] = alg => e =>
+    e ~ e ^^ { case e1 ~ e2 => alg.App(e1, e2) } | ident ^^ alg.Var | "(" ~> e <~ ")"
+}
+\end{lstlisting}
+
+We glue them together by defining an object.
+
+\begin{lstlisting}
+object VarApp {
+  trait Alg[E] {..}
+  trait Print extends Alg[String] {..}
+  trait Parser[E] {..}
+}
+\end{lstlisting}
+
+To use the parser, we must provide an algebra instance as the operation to construct the parsing result.
+As shown in the code below, we feed the pretty printing operation \lstinline{VarApp.Print} to the generic \lstinline{parse} function.
+It will print \lstinline{(x y)} as the result.
+
+\begin{lstlisting}
+def parse[E](inp: String)(alg: VarApp.Alg[E]): E = {
+  val p = new VarApp.Parser[E] {}.pVarApp(alg)
+  val t = phrase(fix(p))(new lexical.Scanner(inp))
+  if (t.successful) t.get else scala.sys.error(t.toString)
+}
+
+def parseAndPrint(inp: String) = println(parse(inp)(new VarApp.Print{}))
+
+parserAndPrint("x y")
 \end{lstlisting}

### 2016-11-28 paper/sections/ParsingLibrary.tex
Some text for related work.

@@
 \begin{itemize}
-\item \textbf{Object Algebra interface:} defined as a trait with several cases, corresponding to the grammar rules.
-\item \textbf{Parser:} implementation of the parser, using open recursion and Object Algebras, with its scoping trait abstracted over the type parameters.
-\item \textbf{Algebra:} some behaviors (operations) on the data structure, such as pretty-printing and so on.
+\item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax, with several methods corresponding to the production rules.
+\item \textbf{Parser:} implementation of the parser with lexing, using open recursion and Object Algebras, with its scoping trait abstracted over the type parameters.
+\item \textbf{Algebra (optional):} some behaviors (operations) on the data structure, such as pretty-printing and so on.
 \end{itemize}

### 2016-11-28 paper/sections/RelatedWork.tex
Some text for related work.

@@
 (read more papers, see if they talk about this issue, some potential solutions)
+(attribute grammars?)
 
@@
 bruijn indices: parsing + transformation
+
+Our work integrates several components including extensible parsing, parser combinators and modular datatypes. There has been a great amount of related papers
+on those hot topics, of which some inspired us of this paper and encourages us for more exploration. This section will try to lead a discussion on what difference we have made.
+
+\paragraph*{Extensible Parsing} Extensible parsing is achieved in many different areas, of which parser generators are a mainstream area specially designed for modular syntax and parsing.
+Many parser generators [OMeta, ANTLR, Rats!, noa, Ohm] \haoyuan{correct for Ohm?} support modular grammars, more specifically, they allow users to create new modules where new non-terminals and production rules can be introduced, some can even override existing rules in the old grammar modules.
+For instance, \textit{Rats!} [Rats!]
+constructs its own module system for the collection of grammars, while NOA [NOA] uses Java annotation processing to gather information together.
+Those parser combinators focus on the \textit{syntactic extensibility} of grammars, and rely on a whole compilation to generate a global parser, even with a slight modification.
+Some of them may statically check the correctness and unambiguity of grammars, but separate compilation and modular type-checking remain unsolved, together with the issue of performance.
+
+Besides, macro systems like C preprocessor, C++ templates [..] and Racket [..], and other meta-programming techniques [..] are a similar area aiming at syntactic extensibility as well, which sacrifices type safety.
+SugarJ [] is another well-known tool that conveniently introduces syntactic sugars in Java programming by library imports.
+Composition of syntactic sugars is easy for users, whereas it requires many rounds of parsing and adaption, which highly affects efficiency of compilation, moreover, the implementation was based on SDF [] and Stratego [], which focused little on separate compilation.
+Extensible compilers like JastAdd [] and Polyglot [], however, are somehow more ambitious, but they involve extensible parsing mostly by parser generators as well, and they focus more on the extensions to a host language.
+Our library is designed for modular language parsers in a type-safe way, with flexible language composition.
+Although overriding existing production rules is tricky with our approach, we could perhaps make use of embedded domain-specific languages on top of parser combinators and transformations for overriding, but it is anyway an orthogonal issue.
+\haoyuan{???} \haoyuan{BTW what about Racket?} \haoyuan{what about metafront?
+it is a macro system but does it have type safety?} \haoyuan{Extensible syntax with lexical scoping?} \haoyuan{"Extensible syntax" proposes a system for extensible syntax, where users write EDSLs in their language with concrete syntax.
+Users can write rules for type-based disambiguation.
+But separate compilation is again not mentioned.
+Shall we mention that thesis?} \haoyuan{attribute grammars?}
+
+On the other hand, extensible parsing algorithms are another area that indeed relates to separate compilation.
+Specifically, [MB] introduces \textit{parse table composition}, in which paper grammars are compiled to the generation of parse tables, which are DFAs or NFAs, later they can be composed by an algorithm, so as to provide separate compilation for parsing.
+Nevertheless, the generation of parse tables is rather expensive, and furthermore, our approach supports separate compilation as well as modular type-checking, and the idea is not restricted to Scala but applicable to many functional languages, on whose type system the safety of modular parsing can rely.
+The extensibility of parsing in our approach is further available at language composition and lexical level.
+\haoyuan{I mentioned one shortcoming of our approach here, which is we cannot override existing production rules; another one is that we do not have explicit correspondence/relationship between abstract syntax and the parser.}
+
+\paragraph*{Parser Combinators} Text.
+
+\paragraph*{Modular Datatypes} Text.

### 2016-11-29 paper/paper.tex
Abstract

@@
 \newcommand\mixin{\textbf{@Mixin} }
+\newcommand\name{\textbf{Parser}\xspace}
 \usepackage{xspace}
@@
 
-Text.
+  Over the years alot of effort has been put on solving challenging
+  extensibility problems, while retaining important software
+  engineering properties such as modular type-safety and
+  separate compilation. Most previous work focused on
+  operations that traverse and process extensible Abtract Syntax Tree
+  (AST) structures. However, there is almost no work on operations that
+  \emph{build} such extensible ASTs, including \emph{parsing}.
+
+  This paper investigates solutions for the problem of \emph{modular
+    parsing}. We focus on \emph{semantic} modularity and not just
+  \emph{syntactic} modularity. That is, the solutions should not only
+  allow complete parsers to be built out of modular parsing
+  components, but also enable the parsing components to be \emph{modularly
+  type-checked} and \emph{separately compiled}. We present \name: a parsing
+  combinator library that enables modular parsing. The library is
+  built on top of a Packrat parsing library, but adds new parsing
+  combinators to enable modular parsing. The new parsing combinators 
+  employ \emph{delegation-based} techniques and \emph{Object Algebras} 
+  to support extensibility. The choice of Packrat parsing over other
+  parsing techniques turns out to be important for achieving
+  performance in a modular setting. To evaluate \name 
+  we conduct a  case study based on the 
+  ``Types and Programming Languages'' interpreters. The case study 
+  shows that \name is effective at reusing parsing code from existing
+  interpreters, and the total parsing code is 60\% smaller than the
+  non-modular parsing code.
 

### 2016-11-29 paper/paper.tex
Draft text for the Introduction

@@
   shows that \name is effective at reusing parsing code from existing
-  interpreters, and the total parsing code is 60\% smaller than the
+  interpreters, and the total parsing code is 60\% shorter than the
   non-modular parsing code.

### 2016-11-29 paper/sections/Introduction.tex
Draft text for the Introduction

\section{Introduction}\label{sec:introduction}

The quest for improved modularity, variability and extensibility of
programs has been going on since the early days of Software
Engineering~\cite{}. Modern Programming Languages enable a certain
degree of modularity, but they have limitations as illustrated by
well-known problems such as the Expression Problem~\cite{}. To address
the limitations of Programming Languages several different approaches
have been proposed in the past. Existing approaches can be broadly
divided into two categories: \emph{syntactic} or \emph{semantic} 
modularization techniques. Syntactic modularization techniques enable 
the separate development of software components or features that can
later be combined into complete programs. However, very often, the 
composition of components in syntactic modularization techniques 
amounts to a form of textual composition of the code from multiple 
components/features. Semantic modularization
techniques go one step further, and also enable those components or
features to be modularly type-checked and separately compiled. This
enables the composition of compiled binaries, and also ensures the
type-safety of the composed code.

Many approaches use syntactic modularization techniques. Examples
include tools for developing Software-Product Lines~\cite{}, some
Language Workbenches~\cite{}, or extensible parser generators.  In
those approaches textual composition techniques such as
\emph{superimposition}~\cite{} enable the development . Syntactic
modularization techniques have also been applied to the problem of
\emph{extensible parsing}. There are several approaches that enable
the development of \emph{syntactically} modular parsers or
grammars. However these approaches do not support separate compilation
or modular type-checking that we strive for in this work.

\begin{comment}
 such as AHEAD or 
FeatureHouse, which enable the development of highly reusable 
software components that can be combined to form complete 
software products.
\end{comment}

There are also several approaches to semantic modularity. For example,
family polymorphism enables the development of highly reusable
components and can even be used to develop Software-Product Lines. 
... However, as far as we are aware, the problem of modular parsing 
has not been studied in semantic modularization
approaches.

  This paper investigates solutions for the problem of \emph{modular
    parsing}. We focus on \emph{semantic} modularity and not just
  \emph{syntactic} modularity. That is, the solutions should not only
  allow complete parsers to be built out of modular parsing
  components, but also enable the parsing components to be \emph{modularly
  type-checked} and \emph{separately compiled}. We present \name: a parsing
  combinator library that enables modular parsing. The library is
  built on top of a Packrat parsing library, but adds new parsing
  combinators to enable modular parsing. The new parsing combinators 
  employ \emph{delegation-based} techniques and \emph{Object Algebras} 
  to support extensibility. The choice of Packrat parsing over other
  parsing techniques turns out to be important for achieving
  performance in a modular setting. To evaluate \name 
  we conduct a  case study based on the 
  ``Types and Programming Languages'' interpreters. The case study 
  shows that \name is effective at reusing parsing code from existing
  interpreters, and the total parsing code is 60\% shorter than the
  non-modular parsing code.

### 2016-11-29 paper/sections/RelatedWork.tex
Related work on parser combinators.

@@
 
-\paragraph*{Parser Combinators} Text.
+\paragraph*{Parser Combinators} Since [Recursive Programming Techniques 1975] firstly introduced parser generators, and after Wadler discussed more details on backtracking in [1985], parser combinators have been more and more popularly developed and used in the research area of parsing.
+Among them many parsing libraries produces recursive descent parsers by introducing functional, or more specifically, monadic parser combinators with a foundation on [monadic parser combinators].
+Related work includes Parsec [], which is frequently used in Haskell for context-sensitive grammars with infinite lookahead.
+Nevertheless, left-recursion is known as a big issue in recursive descent parsers, in which case programmers using libraries like Parsec have to manually eliminate left-recursion in the grammar, which is cumbersome, but also it distorts the shape of its old grammar, restricting the modularity of parsing to extreme extent.
+There are certainly some approaches to afford the limitation of parser combinators, for example, we have conducted our previous experiment in Haskell, by using open recursion, MRM [] and Parsec.
+To support direct left-recursive grammars, we made use of the state monad in Parsec to ensure that a left-recursive production rule will not be applied successively in parsing.
+It turned out to be successful, yet introduced complexity in programming, hence we believe that the parsing library is obligated for left-recursion.
+
+Some more recent papers [Packrat 2006, parsing with derivatives 2011, parser combinator for ambiguous left-recursive grammars] proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion.
+As we explained before, we have selected Packrat Parsing for our prototype implementation, as the combinators are convenient to use in Scala, which is a platform where functional laziness and Object Algebras can perfectly coexist.
+Also, its support for direct left-recursion is already helpful for designing real-world parsers practically, though [packrat parsing can support left recursion] further demonstrated that general left-recursion is supported from the theoretical point of view.
+We have also mentioned that the components used in our library can have alternatives, for instance, a different set of parser combinators may support ambiguous and left-recursive grammars, or be applicable to a different subset of context-free grammars or more, or even lead us to another level of performance.
+
+Another important issue in previous papers, to the best of our knowledge, is that none of them deal with modular datatypes or abstract syntax trees.
+\haoyuan{Well, I need to go to check JS and Patrick Bahr's paper.} The modularity of data structures is yet necessary for modular parsing that supports separate compilation and modular type-checking.
+In our library, Object Algebras have been adopted for extensible ASTs, together with the behaviors on them.
 

### 2016-11-29 paper/sections/RelatedWork.tex
minor comment

@@
 
-Some more recent papers [Packrat 2006, parsing with derivatives 2011, parser combinator for ambiguous left-recursive grammars] proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion.
+Some more recent papers [Packrat 2006, parsing with derivatives 2011, parser combinator for ambiguous left-recursive grammars] \haoyuan{Pracitical, general parser combinators?} proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion.
 As we explained before, we have selected Packrat Parsing for our prototype implementation, as the combinators are convenient to use in Scala, which is a platform where functional laziness and Object Algebras can perfectly coexist.

### 2016-11-29 paper/sections/Overview.tex
Overview

@@
 
-With Packrat parser combinators, we obtain the possibility of writing parsers in a modular way.
-Let's start by looking at how to build a parser for a simple language, which only has variables and applications at the beginning.
+In this section we start demonstrating how to write modular parsers. Suppose we want to parse a simple language of variables and applications.
 
@@
 
-Firstly, we define the abstract syntax by Scala's convenient case classes. Expressions have type \lstinline{Expr}, and each of them is either \lstinline{Var} or \lstinline{App}.
+The abstract syntax can be defined by Scala's case classes, and then we write parsers for the two cases \lstinline{Var} and \lstinline{App}.
 
@@
 case class App(e1: Expr, e2: Expr) extends Expr
-\end{lstlisting}
 
-And then we write parsing functions for each case.
-
-\begin{lstlisting}
-val pVar: PackratParser[Expr] =
-  ident ^^ Var
+val pVar: PackratParser[Expr] = ident ^^ Var
 
@@
 
-The \lstinline{PackratParser} type and the combinators are imported from the Scala Packrat parsing library.
-Some details about these structures will be discussed in later sections, so that readers could only focus on the types of these functions at the first reading.
+Some structures are imported from the Scala Packrat parsing library, such as the combinator \lstinline{^^}.
+They will be discussed in later sections, so that readers could only focus on the type of parsers at the first reading.
 
-\lstinline{pVar} is straight forward.
-It parses a single identifier, and its type \lstinline{PackratParser[Expr]} indicates the result is an \lstinline{Expr}.
-However the type of \lstinline{pApp} is different.
-It takes a parser of \lstinline{Expr} as an argument, which represents the parser of the whole expression, namely \lstinline{e} here.
-And use it to parse two consecutive expressions as an application.
-In this particular case, the parser of the whole expression is in fact the combination of \lstinline{pVar} and \lstinline{pApp}.
-We can write equivalent code as below.
+\lstinline{pVar} for parsing variables is straight forward. It parses a single identifier, and its type \lstinline{PackratParser[Expr]} indicates the result is an \lstinline{Expr}.
 
-\begin{lstlisting}
-val pApp2: PackratParser[Expr] = {
-  lazy val e = pApp2 | pVar
-  e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
-}
-\end{lstlisting}
-
-\lstinline{pApp2} works fine, but it does not have the extensibility as \lstinline{pApp} does.
-Since the parser of the whole expression in \lstinline{pApp2} is hard-coded, it covers and only covers the variable and application cases.
-If we we extend \lstinline{Expr} later to add a third case lambda abstraction, we have to edit the existing code.
+The type of \lstinline{pApp} is different.
+It takes a parser of \lstinline{Expr} as an argument, and use it to parse two consecutive expressions as an application.
+This argument \lstinline{e} represents the parser of \emph{the whole expression}.
+In this particular case, it is in fact the combination of \lstinline{pVar} and \lstinline{pApp}.
+We can replace it by \lstinline{pVar | pApp} and the code still works.
+However, using an abstract argument provides extensibility to the parser.
+Because once the abstract syntax is extended later, the parser of the whole expression will change correspondingly.
+Then we have to edit existing code if it is hard-coded.
 The parameterization technique used in \lstinline{pApp} is called \textit{open recursion}, which is discussed in section \ref{sec:openandparsing}.
 
-We can use a trait to glue them together, and define \lstinline{pVarApp} as a combination of the two parsing functions. A type synonym \lstinline{Fix} is used to shorten the types.
+We can use a trait to glue them together, and define \lstinline{pVarApp} as a combination of the two parsers. A type synonym \lstinline{Fix} is used to shorten the types.
 
@@
 \begin{lstlisting}
-def parseAndPrint(p: Fix[PackratParser[Expr]])(inp: String) = {
+def parse[E](p: Fix[PackratParser[E]])(inp: String): E = {
   val t = phrase(fix(p))(new lexical.Scanner(inp))
-  if (t.successful) println(t.get) else scala.sys.error(t.toString)
+  if (t.successful) t.get else scala.sys.error(t.toString)
 }
 
-parseAndPrint(new ExprParser {}.pVarApp)("x y")
+println(parse(new ExprParser {}.pVarApp)("x y"))
 \end{lstlisting}
 
-It will print \lstinline{App(Var(x),Var(y))} as the parsing result of string \lstinline{"x y"}.
+It will print the AST as \lstinline{App(Var(x),Var(y))}.
 
@@
 \begin{lstlisting}
-parseAndPrint(new LambdaParser {}.pVarAppLam)("(\\x.x) y")
+println(parse(new LambdaParser {}.pVarAppLam)("(\\x.x) y"))
 \end{lstlisting}
 
-The code above will print result \lstinline{App(Lam(x,Var(x)),Var(y))} from parsing \lstinline{"(\\x.x)} \lstinline{y"}.
+The code above will print \lstinline{App(Lam(x,Var(x)),Var(y))} from parsing \lstinline{"(\\x.x)} \lstinline{y"}.
 
@@
 
-We glue them together by defining an object.
+We glue them together by defining an object. It actually becomes a \textit{language component} that represents a specific language feature.
 
@@
 \begin{lstlisting}
-def parse[E](inp: String)(alg: VarApp.Alg[E]): E = {
+def parseWithAlg[E](inp: String)(alg: VarApp.Alg[E]): E = {
   val p = new VarApp.Parser[E] {}.pVarApp(alg)
-  val t = phrase(fix(p))(new lexical.Scanner(inp))
-  if (t.successful) t.get else scala.sys.error(t.toString)
+  parse[E](p)(inp)
 }
 
-def parseAndPrint(inp: String) = println(parse(inp)(new VarApp.Print{}))
+println(parseWithAlg("x y")(new VarApp.Print{}))
+\end{lstlisting}
 
-parserAndPrint("x y")
+With Object Algebras, we can extend all the stuff, including abstract syntax, operations, and parsers. The code below extend the language to support lambda abstraction.
+
+\begin{lstlisting}
+object Lambda {
+  trait Alg[E] extends VarApp.Alg[E] {
+    def Lam(x: String, e: E): E
+  }
+
+  trait Print extends Alg[String] with VarApp.Print {
+    def Lam(x: String, e: String) = "\\" + x + "." + e
+  }
+
+  trait Parser[E] extends VarApp.Parser[E] {
+    lexical.delimiters += ("\\", ".")
+
+    val pLam: Alg[E] => Fix[PackratParser[E]] = alg => e =>
+      ("\\" ~> lcid) ~ ("." ~> e) ^^ { case x ~ e0 => alg.Lam(x, e0) }
+    val pVarAppLam: Alg[E] => Fix[PackratParser[E]] =
+      pVarApp | pLam
+  }
+}
+\end{lstlisting}
+
+For easily combining the parsers using Object Algebras as the representation of abstract syntax, we have a special combinator \huang{we need to rename our combinator} defined in our library.
+
+In addition to make extending operations possible, another advantage of using Object Algebras is that it supports multiple sorts of syntax.
+For example, if our language has not only expressions but also types, we can easily distinguish them in the abstract syntax by an extra type parameter.
+This topic will be discussed in section \ref{subsec:differentsyntax}.
+
+\begin{lstlisting}
+trait TypedLambda[E, T] extends VarApp.Alg[E] {
+  def Lam(x: String, t: T, e: E): E
+  def IntType(): T
+  def BoolType(): T
+  def ArrowType(a: T, b: T): T
+}
 \end{lstlisting}

### 2016-11-29 paper/paper.tex
Typos

@@
 
-  Over the years alot of effort has been put on solving challenging
+  Over the years a lot of effort has been put on solving challenging
   extensibility problems, while retaining important software
@@
   separate compilation. Most previous work focused on
-  operations that traverse and process extensible Abtract Syntax Tree
+  operations that traverse and process extensible Abstract Syntax Tree
   (AST) structures. However, there is almost no work on operations that

### 2016-11-29 paper/sections/Overview.tex
Fix

@@
 
-  val pLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
+  val pLam: Fix[PackratParser[Expr]] =
     e => "\\" ~> ident ~ ("." ~> e) ^^ { case x ~ b => Lam(x, b) }
-  val pVarAppLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
+  val pVarAppLam: Fix[PackratParser[Expr]] =
     e => pLam(e) | pVarApp(e)

### 2016-11-30 paper/sections/RelatedWork.tex
Related work: modular datatypes.

@@
 
-\paragraph*{Modular Datatypes} Text.
+\paragraph*{Modular Datatypes} There has been lots of related work on building modular data structures, especially in some functional languages with non-extensible datatypes like Haskell.
+For instance, the famous DTC paper [] represents modular ASTs using co-products of every two functors, which forms a binary tree structure.
+It supports extensible datatypes, as well as the operations on them by implementing instances of functors.
+Later Oliveira et al.
+presented MRM [] in Haskell, which is similar to DTC but adopts a list of functors in type-level lists instead.
+As mentioned above, we have conducted our early experiment with MRM for modular datatypes, whereas the complexity of library implementation was comparably high.
+
+In OO languages, however, things get much better with the mechanism of inheritance.
+Furthermore, Object Algebras [] were proposed as a design pattern for the extensibility of both data variants and operations in a modular type-safe way.
+The code is concise and expressive using Object Algebras, building a foundation for our new library with various dimensions of extensibility.
+\haoyuan{To mention: noa and modular semantic actions, and shy use Object Algebras, shy and mrm can override cases.}

### 2016-12-07 paper/sections/RelatedWork.tex
Done related work draft. Added references in bibtex.

@@
 
-- extensible parsing, language workbenches: rats, noa (this one already uses OA), modular semantic actions, (syntactic modularity, no separate compilation, modular type-checking)
-(read more papers, see if they talk about this issue, some potential solutions)
-(attribute grammars?)
-
-- parser combinators for type-checking, previous work has not shown how to support modularity (ASTs); left-recursion and back-tracking in related techniques
-
-- modularity: object algebras, dtc and mrm (problem with parsing, is there any related work? (PB: a paper on unfolds: build the AST))
-
-(parsing in Javascript: using delegation, does it support modular AST)
-
-noa, shy: shy: only override some interesting cases (transformation is tedious)
-bruijn indices: parsing + transformation
+%- extensible parsing, language workbenches: rats, noa (this one already uses OA), modular semantic actions, (syntactic modularity, no separate compilation, modular type-checking)
+%(read more papers, see if they talk about this issue, some potential solutions)
+%(attribute grammars?)
+%
+%- parser combinators for type-checking, previous work has not shown how to support modularity (ASTs); left-recursion and back-tracking in related techniques
+%
+%- modularity: object algebras, dtc and mrm (problem with parsing, is there any related work? (PB: a paper on unfolds: build the AST))
+%
+%(parsing in Javascript: using delegation, does it support modular AST)
+%
+%noa, shy: shy: only override some interesting cases (transformation is tedious)
+%bruijn indices: parsing + transformation
 
@@
 \paragraph*{Extensible Parsing} Extensible parsing is achieved in many different areas, of which parser generators are a mainstream area specially designed for modular syntax and parsing.
-Many parser generators [OMeta, ANTLR, Rats!, noa, Ohm] \haoyuan{correct for Ohm?} support modular grammars, more specifically, they allow users to create new modules where new non-terminals and production rules can be introduced, some can even override existing rules in the old grammar modules.
-For instance, \textit{Rats!} [Rats!]
-constructs its own module system for the collection of grammars, while NOA [NOA] uses Java annotation processing to gather information together.
+Many parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support modular grammars, more specifically, they allow users to create new modules where new non-terminals and production rules can be introduced, some can even override existing rules in the old grammar modules.
+For instance, \textit{Rats!}~\cite{Grimm2006}
+constructs its own module system for the collection of grammars, while NOA~\cite{Gouseti2014} uses Java annotation processing to gather information together.
 Those parser combinators focus on the \textit{syntactic extensibility} of grammars, and rely on a whole compilation to generate a global parser, even with a slight modification.
@@
 
-Besides, macro systems like C preprocessor, C++ templates [..] and Racket [..], and other meta-programming techniques [..] are a similar area aiming at syntactic extensibility as well, which sacrifices type safety.
-SugarJ [] is another well-known tool that conveniently introduces syntactic sugars in Java programming by library imports.
-Composition of syntactic sugars is easy for users, whereas it requires many rounds of parsing and adaption, which highly affects efficiency of compilation, moreover, the implementation was based on SDF [] and Stratego [], which focused little on separate compilation.
-Extensible compilers like JastAdd [] and Polyglot [], however, are somehow more ambitious, but they involve extensible parsing mostly by parser generators as well, and they focus more on the extensions to a host language.
+Besides, macro systems like C preprocessor, C++ templates and Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming techniques are a similar area aiming at syntactic extensibility as well, which sacrifices type safety.
+SugarJ~\cite{Erdweg2011} is another well-known tool that conveniently introduces syntactic sugars in Java programming by library imports.
+Composition of syntactic sugars is easy for users, whereas it requires many rounds of parsing and adaption, which highly affects efficiency of compilation, moreover, the implementation was based on SDF~\cite{Heering1989} and Stratego~\cite{Visser2001}, which focused little on separate compilation.
+Extensible compilers like JastAdd~\cite{Ekman2007} and Polyglot~\cite{Nystrom2003}, however, are somehow more ambitious, but they involve extensible parsing mostly by parser generators as well, and they focus more on the extensions to a host language.
 Our library is designed for modular language parsers in a type-safe way, with flexible language composition.
@@
 On the other hand, extensible parsing algorithms are another area that indeed relates to separate compilation.
-Specifically, [MB] introduces \textit{parse table composition}, in which paper grammars are compiled to the generation of parse tables, which are DFAs or NFAs, later they can be composed by an algorithm, so as to provide separate compilation for parsing.
+Specifically, \cite{Bravenboer2009} introduces \textit{parse table composition}, in which paper grammars are compiled to the generation of parse tables, which are DFAs or NFAs, later they can be composed by an algorithm, so as to provide separate compilation for parsing.
 Nevertheless, the generation of parse tables is rather expensive, and furthermore, our approach supports separate compilation as well as modular type-checking, and the idea is not restricted to Scala but applicable to many functional languages, on whose type system the safety of modular parsing can rely.
@@
 
-\paragraph*{Parser Combinators} Since [Recursive Programming Techniques 1975] firstly introduced parser generators, and after Wadler discussed more details on backtracking in [1985], parser combinators have been more and more popularly developed and used in the research area of parsing.
-Among them many parsing libraries produces recursive descent parsers by introducing functional, or more specifically, monadic parser combinators with a foundation on [monadic parser combinators].
-Related work includes Parsec [], which is frequently used in Haskell for context-sensitive grammars with infinite lookahead.
+\paragraph*{Parser Combinators} Since~\cite{burge1975} firstly introduced parser generators, and after Wadler discussed more details on backtracking in~\cite{Wadler1985}, parser combinators have been more and more popularly developed and used in the research area of parsing.
+Among them many parsing libraries produces recursive descent parsers by introducing functional, or more specifically, monadic parser combinators with a foundation on~\cite{nott237}.
+Related work includes Parsec~\cite{Leijen2001}, which is frequently used in Haskell for context-sensitive grammars with infinite lookahead.
 Nevertheless, left-recursion is known as a big issue in recursive descent parsers, in which case programmers using libraries like Parsec have to manually eliminate left-recursion in the grammar, which is cumbersome, but also it distorts the shape of its old grammar, restricting the modularity of parsing to extreme extent.
-There are certainly some approaches to afford the limitation of parser combinators, for example, we have conducted our previous experiment in Haskell, by using open recursion, MRM [] and Parsec.
+There are certainly some approaches to afford the limitation of parser combinators, for example, we have conducted our previous experiment in Haskell, by using open recursion, MRM~\cite{Oliveira2015} and Parsec.
 To support direct left-recursive grammars, we made use of the state monad in Parsec to ensure that a left-recursive production rule will not be applied successively in parsing.
@@
 
-Some more recent papers [Packrat 2006, parsing with derivatives 2011, parser combinator for ambiguous left-recursive grammars] \haoyuan{Pracitical, general parser combinators?} proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion.
+Some more recent papers~\cite{Ford2002,Might2011,Frost2008} \haoyuan{Pracitical, general parser combinators?} proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion.
 As we explained before, we have selected Packrat Parsing for our prototype implementation, as the combinators are convenient to use in Scala, which is a platform where functional laziness and Object Algebras can perfectly coexist.
-Also, its support for direct left-recursion is already helpful for designing real-world parsers practically, though [packrat parsing can support left recursion] further demonstrated that general left-recursion is supported from the theoretical point of view.
+Also, its support for direct left-recursion is already helpful for designing real-world parsers practically, though~\cite{warth2008} further demonstrated that general left-recursion is supported from the theoretical point of view.
 We have also mentioned that the components used in our library can have alternatives, for instance, a different set of parser combinators may support ambiguous and left-recursive grammars, or be applicable to a different subset of context-free grammars or more, or even lead us to another level of performance.
@@
 Another important issue in previous papers, to the best of our knowledge, is that none of them deal with modular datatypes or abstract syntax trees.
-\haoyuan{Well, I need to go to check JS and Patrick Bahr's paper.} The modularity of data structures is yet necessary for modular parsing that supports separate compilation and modular type-checking.
+The modularity of data structures is yet necessary for modular parsing that supports separate compilation and modular type-checking.
 In our library, Object Algebras have been adopted for extensible ASTs, together with the behaviors on them.
@@
 \paragraph*{Modular Datatypes} There has been lots of related work on building modular data structures, especially in some functional languages with non-extensible datatypes like Haskell.
-For instance, the famous DTC paper [] represents modular ASTs using co-products of every two functors, which forms a binary tree structure.
+For instance, the famous DTC paper~\cite{swierstra2008} represents modular ASTs using co-products of every two functors, which forms a binary tree structure.
 It supports extensible datatypes, as well as the operations on them by implementing instances of functors.
 Later Oliveira et al.
-presented MRM [] in Haskell, which is similar to DTC but adopts a list of functors in type-level lists instead.
+presented MRM~\cite{Oliveira2015} in Haskell, which is similar to DTC but adopts a list of functors in type-level lists instead.
 As mentioned above, we have conducted our early experiment with MRM for modular datatypes, whereas the complexity of library implementation was comparably high.
+Other related work includes~\cite{Bahr2014}, where data types are flexible with composition as well as decomposition.
 
 In OO languages, however, things get much better with the mechanism of inheritance.
-Furthermore, Object Algebras [] were proposed as a design pattern for the extensibility of both data variants and operations in a modular type-safe way.
+Furthermore, Object Algebras~\cite{Oliveira2012} were proposed as a design pattern for the extensibility of both data variants and operations in a modular type-safe way.
 The code is concise and expressive using Object Algebras, building a foundation for our new library with various dimensions of extensibility.
-\haoyuan{To mention: noa and modular semantic actions, and shy use Object Algebras, shy and mrm can override cases.}
+Moreover, in~\cite{Oliveira2012} the authors have discussed the composition of algebras.
+In our parsing approach, a parser consumes an algebra, which is delegated to return the results, during its process of parsing.
+Having a set of algebras, it requires multiple parsing with several times of invocation, which leads to redundant work.
+Instead, algebras are supposed to be composed into one before the invocation of the parser.
+Bahr et al.
+lead a similar discussion in~\cite{Bahr2011}, where queries (or \textit{catamorphisms}) and transformations (or \textit{homomorphisms}) are composable.
+They have also mentioned the dual process of folds, namely \textit{anamorphisms}.
+It is potentially related to our work, as parsing is a representative kind of unfolds, whereas they only discussed the composition of a cv-coalgebra and a term homomorphism, which differs from modular parsing.
+\haoyuan{From Object Algebras to Attribute Grammars?}
+
+Another interesting observation, as we have mentioned in~\ref{subsec:objectalgebras}, is that transformation algebras are frequently used in parsing, as concrete syntax may not always coincide with its abstract syntax, for instance, the bruijn indices in lambda calculus, as well as other syntactic sugars.
+In the Shy framework~\cite{Zhang2015}, the pattern of transformations is captured into a template for Java annotation processing, and most of boilerplate code is automatically generated, hence users only need to override a few interesting cases, conveniently.
+The idea of overriding existing cases was also proposed in MRM.
+It is potentially useful to our framework.

### 2016-12-08 paper/sections/Overview.tex
First comments on Section 2

@@
 
-\subsection{The Choice of Parsing Techniques}\label{subsec:overview-parsing}
+\subsection{Chosing the Parsing Technique}\label{subsec:overview-parsing}
 
-Parsing techniques have been heavily studied over the years, especially for context-free grammars and its subsets.
-However, the choice of parsing techniques is not a minor issue in a modular setting.
-To achieve type-safe modularity, the parsers must be able to be extended and composed, and different parsing techniques have different suitability for that.
+%A technique for type-safe modular parsing should the following 3 
+%features: \emph{modular type-checking}; \emph{separate compilation}; 
+%low performance overhead
 
-Although every parsing algorithm can be implemented by hand in principle, some of them are not friendly for that.
-A \textit{parser generator} is a tool to help users generate parsers automatically or semi-automatically from a given grammar.
-There is no restriction on the algorithm, while most of them adopts table-based LL and LR parsing algorithms.
+Although there are many parsing techniques, not all of them are
+suitable for type-safe modular parsing. In particular there are many 
+techniques which fail to provide modular type-checking and separate 
+compilation. Moreover, even if modular type-checking and separate 
+compilation are supported, efficiency is another
+concern. A parsing technique should have low overhead when applied 
+in a modular setting. In the remaider of this section, we eliminate 
+various techniques that fail to satisfy our requirements, and argue
+that that Packrat parsing~\cite{} is a suitable candidate for
+type-safe modular parsing.
 
-\huang{a paragraph here, refer to related work section, discuss previous works of modular parsing using parser generators and its drawbacks}
+\paragraph{Parser Generators} The most widely use tools for parsing 
+are parser generators~\cite{}. Parser generators help users generate parsers automatically or
+semi-automatically from a given grammar. There is no restriction on
+the algorithm, while most of them adopts table-based LL~\cite{} and LR parsing
+algorithms~\cite{}. 
+Although efficient, the main drawback of parser generators is that they do not support 
+modular type-checking and separate compilation.
 
-Comparing with the parser generators, a \textit{parser combinator} takes several parsers and produce a new parser as its output.
-It is popular in functional programming, where the parsers are represented by functions and parser combinators are higher-order functions accepts them.
+\huang{a paragraph here, refer to related work section, discuss
+  previous works of modular parsing using parser generators and its
+  drawbacks}\bruno{Please finish this.}
 
-Parser combinators are very suitable for our purpose, because of two reasons.
-Firstly, they are naturally modular.
-The manner of using them is to write small parsers and use combinators to composed them together.
-The construction procedure is explicit and fully controlled by the programmer.
-Secondly, each parser combinator is represented by a piece of code, and also are the parsers it takes.
-Thus in a statically typed programming language they can be type-checked during compile time.
+\paragraph{Parser Combinators}
+Comparing with the parser generators, a \textit{parser combinator}~\cite{}
+takes several parsers and produce a new parser as its output. Parser combinators are
+popular in functional programming, where the parsers are represented
+by functions and parser combinators are higher-order functions accepts
+them.
 
-Parser combinators adopt top-down, recursive descent parsing, which uses backtracking to search through the possible branches.
-However, simple backtracking parser combinators such as the famous library Parsec in Haskell, have some drawbacks.
-One is that they cannot support left-recursive grammars.
-The common solution is to rewrite a left-recursive grammar into an equivalent one, so called \textit{left-recursion elimination}.
-This solution requires information of the whole grammar, so that it cannot be applied in a modular setting.
-Because the grammar may be extended in the future, essentially we only have parts of the grammar.
-Another drawback is performance.
-Take Parsec as an example, its choice combinator only tries the second alternative if the first fails without any token consumption.
-If two alternatives have a same non-empty prefix, an auxiliary function \lstinline{try} must be added to backtrack.
-However, if we want to extend our parsers later, we should always consider the worst case in which all alternatives share common prefixes.
-Then we need to add \lstinline{try} for all the branches, which results worst case exponential time complexity.
+At a first look, parser combinators are very suitable for our purpose, because of two
+reasons. Firstly, they are naturally modular. The manner of using them
+is to write small parsers and use combinators to composed them
+together. The construction procedure is explicit and fully controlled
+by the programmer. Secondly, each parser combinator is represented by
+a piece of code, and also are the parsers it takes. Thus in a
+statically typed programming language they can be statically
+type-checked.
 
-Beyond the simple backtracking one, several advanced parser combinators were studied to resolve those issues.
-One of them is Packrat parsers, which is our choice.
-The details of Packrat parser are beyond the scope of this paper.
-In brief, Packrat parsers use memorization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated, and it supports left-recursive grammars.
-All of these properties are very suitable for modularity, thus we use Packrat parsers as the underlying parsing technique.
-In section \ref{sec:packratparsers}, we will talk about the Packrat parsing library implemented in Scala, which we use.
+Unfortunatelly many parser combinators have important limitations for
+modular parsing. In particular several parser combinators, 
+including the famous library Parsec~\cite{} library, require
+programmers to manually do\textit{left-recursion elimination}, and
+require significant amounts \textit{backtracking}. Both of these are
+problematic in a modular setting.
 
-It is worth mentioning that the choice of parser combinators will not affect the other parts of our library.
-One can choose other parser combinators like Parsec, in cases that the performance and supporting of left-recursion are not major concerns.
+\paragraph{Left-Recursion Elimination} The main problem with
+left-recursion elimination is that it is a \emph{global}
+transformation on a grammar. Given that the full grammar is known,
+then it is possible to remove all left-recursive cases.
+\bruno{Illustrate with a concrete example: show a simple
+  left-recursive grammar; then do left-recursion elimination. Then add
+a new case into the original left-recursive grammar; and show that 
+a very different grammar is obtained after left-recursion elimination.}
 
-\subsection{Modular Parsers, the First Look}\label{subsec:overview-firstlook}
+However, when doing modular parsing, the full grammar is not known!
 
-In this section we start demonstrating how to write modular parsers. Suppose we want to parse a simple language of variables and applications.
+Parser combinators adopt top-down, recursive descent parsing, which
+uses backtracking to search through the possible branches. However,
+simple backtracking parser combinators such as the famous library
+Parsec~\cite{} in Haskell, have some drawbacks. One is that they cannot
+support left-recursive grammars. The common solution is to rewrite a
+left-recursive grammar into an equivalent one, so called
+\textit{left-recursion elimination}. This solution requires
+information of the whole grammar, so that it cannot be applied in a
+modular setting. Because the grammar may be extended in the future,
+essentially we only have parts of the grammar.
+
+\paragraph{Backtracking} The need for backtracking is also problematic
+in a modular setting. Take Parsec as an example, its choice
+combinator only tries the second alternative if the first fails
+without any token consumption. If two alternatives have a same
+non-empty prefix, an auxiliary function \lstinline{try} must be added
+to backtrack. However, if we want to extend our parsers later, we
+should always consider the worst case in which all alternatives share
+common prefixes. Then we need to add \lstinline{try} for all the
+branches, which results worst case exponential time complexity.
+\bruno{you should again illustrate your point here with an
+  example. The example should illustrate that the addition of a new
+  case may require introducing backtracking. }
+
+\paragraph{Packrat Parsing} 
+Fortunatelly more advanced parsing techniques such as Packrat
+parsing~\cite{}, address the limitations of simple parser combinators 
+such as Parsec. Packrat parsers use
+memorization to record the result of applying each parser at each
+position of the input, so that repeated computation is eliminated, and
+they supports left-recursive grammars directly. All of these properties are very
+suitable for modularity, thus we use Packrat parsers as the underlying
+parsing technique. In section \ref{sec:packratparsers}, we will give
+an overview of the Packrat parsing library implemented in Scala, 
+which we use.
+
+It is worth mentioning that the choice of parser combinators will not
+affect the other parts of our library. One can choose other parser
+combinators like Parsec, in cases that the performance and supporting
+of left-recursion are not major concerns.
+
+\subsection{The Parsing Expression Problem}
+\bruno{You are not motivating the problem! You are going straight to
+  the solution without pointing out what the problem is first. What
+  you need to do is: First show what happens with conventional
+  parsers: at some point, if you add extensions the recursive calls
+  will be wrong. Then you show (in the next section) the solution: 
+ use delegation/open recursion.}
+
+\bruno{I think we need to set up a challenge here, similar to the challange of
+  the expression problem. The challange should be like. Build a parser
+for a simple expression language, then extend the language with both 
+a new language construct and a new operation. This section will show 
+how we can do that using traditional parsers, but without modular
+type-safety and separate compilation. The remaining sections will show
+that the techniques introduced by us, enable us to solve those two challenges.}
+
+Typical parsers defined parser combinators follow the grammar
+structure, and build a corresponding AST. If the grammar is changed to
+add a new alternative, the corresponding parser has to be changed as
+well. Thus the traditional style of developing parsers is not
+extensible, and it is at odds with object-oriented programming ASTs,
+which support the easy addition of new alternative classes. We
+illustrate this problem with a simple expression language and extension.
+
+Suppose we want to parse a simple language of variables and
+applications.
 
@@
 \end{lstlisting}
+\bruno{if you intend to write a grammar here, then please use grammar
+  (BNF) notation, and not some notation invented by you!}
 
-The abstract syntax can be defined by Scala's case classes, and then we write parsers for the two cases \lstinline{Var} and \lstinline{App}.
+\paragraph{Extensing the Parser}
+
+\paragraph{Problem: Hardcoded Recursive Calls}
+
+\subsection{Delegation-Based Parsing}\label{subsec:overview-firstlook}
+
+The problem of hardcoded recursive calls can be solved using
+delegation, which can be encoded with open recursion. The idea is
+simple: instead of making direct recursive calls, an additional
+argument, that abstracts over the recursive call, is passed to the 
+parsers. This technique enables the developement of parsers that 
+support OO-style extensibility of ASTs.
+
+\paragraph{Open Parsing}
+The example in Section \ref{} can be rewritten as follows:
 
@@
 \end{lstlisting}
+\bruno{You are copying and pasting code into the paper. Don't do this:
+code will change alot (various edits), and it is all too easy to get
+inconsistent code snippets. Instead use the small Ruby tool that we 
+have in the paper directory to import source code directly from the
+code repository. This will ensure the consistency of the code
+throughout the paper writing. Please ask Haoyuan about the tool; he
+knows how to use it!}
+\bruno{Please polish up code. You probably want to consistently use 
+the variable name ``self'' whenever you want to refer to the whole
+parser (instead of ``e'' that you use at the moment). Also you should 
+use ``Open'' (which you currently call ``Fix'') instead of using the
+longwinded ``(=> PackratParser[Expr]) => PackratParser[Expr]''} 
+\bruno{Also, for consistency, why isn't ``pVar'' open? I.e. using a
+  self reference as well? I know that it is not recursive, but giving
+  it a different treatment makes it more confusing for users.}
+\bruno{You should use a more OO approach in the example. 
+Rather than case classes show an Expr trait with pretty printing.
+This will enable you to talk about the extensibility and object
+algebras better later on.
+}
 
-Some structures are imported from the Scala Packrat parsing library, such as the combinator \lstinline{^^}.
-They will be discussed in later sections, so that readers could only focus on the type of parsers at the first reading.
+The abstract syntax is defined by Scala's case classes, and 
+we write parsers for the two cases \lstinline{Var} and \lstinline{App}.
+The parser \lstinline{pVar}, for parsing variables, is straightforward. It parses a
+single identifier, and its type \lstinline{PackratParser[Expr]}
+indicates the result is an \lstinline{Expr}.
+The parser \lstinline{pApp}, for parsing applications is more
+interesting. \lstinline{pApp} takes a parser of
+\lstinline{Expr} as an argument, and uses it to parse two consecutive
+expressions as an application. This argument \lstinline{e} represents
+the parser of \emph{the whole expression}. In this particular case, it
+is in fact the combination of \lstinline{pVar} and
+\lstinline{pApp}. We can replace it by \lstinline{pVar | pApp} and the
+code still works. However, using an abstract argument provides
+extensibility to the parser. Because once the abstract syntax is
+extended later, the parser of the whole expression will change
+correspondingly. Then we have to edit existing code if it is
+hard-coded. The parameterization technique used in \lstinline{pApp} is
+called \textit{open recursion}, which is discussed in section
+\ref{sec:openandparsing}.
 
-\lstinline{pVar} for parsing variables is straight forward. It parses a single identifier, and its type \lstinline{PackratParser[Expr]} indicates the result is an \lstinline{Expr}.
-
-The type of \lstinline{pApp} is different.
-It takes a parser of \lstinline{Expr} as an argument, and use it to parse two consecutive expressions as an application.
-This argument \lstinline{e} represents the parser of \emph{the whole expression}.
-In this particular case, it is in fact the combination of \lstinline{pVar} and \lstinline{pApp}.
-We can replace it by \lstinline{pVar | pApp} and the code still works.
-However, using an abstract argument provides extensibility to the parser.
-Because once the abstract syntax is extended later, the parser of the whole expression will change correspondingly.
-Then we have to edit existing code if it is hard-coded.
-The parameterization technique used in \lstinline{pApp} is called \textit{open recursion}, which is discussed in section \ref{sec:openandparsing}.
-
-We can use a trait to glue them together, and define \lstinline{pVarApp} as a combination of the two parsers. A type synonym \lstinline{Fix} is used to shorten the types.
+\paragraph{Parsing Components}\bruno{consider removing and/or moving
+  to later parts of the paper}
+We can use a trait to glue them together, and define
+\lstinline{pVarApp} as a combination of the two parsers. A type
+synonym \lstinline{Fix} is used to shorten the types.
 
@@
 \end{lstlisting}
+\bruno{Fix is not the right name. A better name is Open (for open
+  recursion).}
+\bruno{I'm not sure that showing the code for language components is
+  essential at this point. This can be shown latter in the paper.}
 
+\paragraph{Client Code}
 Finally, we are able to parse \lstinline{Expr} frome strings as shown in the code below.
@@
 
-\subsection{Extend and Reuse Parsers}\label{subsec:overview-extend}
+%\subsection{Extend and Reuse Parsers}\label{subsec:overview-extend}
 
-We have done our modular parser, and it behaves just in a normal way.
-To demonstrate the advantages of writing modular parsers, we extend the language to add lambda abstractions as a new case of expressions.
-The language then becomes the famous \textit{untyped lambda calculus}.
+\paragraph{Extensibility} To demonstrate the advantages of writing 
+parsers with open recursion, we extend the language to add lambda 
+abstractions as a new case of expressions. The language then 
+becomes the famous \textit{untyped lambda calculus}.
 
@@
 \end{lstlisting}
+\bruno{same comment as before: if this is supposed to be a grammar,
+  then use a grammar notation}
 
@@
 
-Since we already have the parser for variables and applications, we want to reuse it and build a new parser by extending it.
-It is just what the code below does.
-Notice that the old code is not touched, hence separate compilation is obtained.
+Since we already have the parser for variables and applications, we
+want to reuse it and build a new parser by extending it. It is just
+what the code below does. Notice that the old code is not touched,
+hence separate compilation is obtained.
 
@@
 
-In the new parser \lstinline{LambdaParser} extended from the old one, we write a new parsing function \lstinline{pLam} for the new case lambda abstraction.
-And compose it with \lstinline{pVarApp}, which we already have in the old parser \lstinline{ExprParser}.
-Open recursion is critical to achieve such type-safe compositions, because wherever we refer to the parser of the whole expression, it is updated to cover all of the three case we have now.
+In the new parser \lstinline{LambdaParser} extended from the old one,
+we write a new parsing function \lstinline{pLam} for the new case
+lambda abstraction. And compose it with \lstinline{pVarApp}, which we
+already have in the old parser \lstinline{ExprParser}. Open recursion
+is critical to achieve such type-safe compositions, because wherever
+we refer to the parser of the whole expression, it is updated to cover
+all of the three case we have now.
 
@@
 
-\subsection{Object Algebras and More}\label{subsec:overview-oa}
+\subsection{Object Algebras for full Extensibility}\label{subsec:overview-oa}
 
-Instead of define the abstract syntax using Scala's case classes, we can use \textit{Object Algebras} for more extensibility.
-Object Algebras is a technique to solve the \textit{Expression Problem}, providing the possibility of extending both data variants and operations over them.
-It will be further discussed in section \ref{sec:algebrasandparsing}.
+Although delegation enables OO extensibility, the use of an OO class
+hierarchy makes the addition of new operations over the AST
+problematic. Object Algebras~\cite{} enable us to solve this problem, and
+offer high flexibility in the choice of operations to be performed
+over the AST.
+
+Instead of define the abstract syntax using Scala's case classes, we
+can use \textit{Object Algebras} for more extensibility. Object
+Algebras is a technique to solve the \textit{Expression Problem},
+providing the possibility of extending both data variants and
+operations over them. It will be further discussed in section
+\ref{sec:algebrasandparsing}.
 

### 2016-12-11 paper/sections/PackratParsers.tex
Done text for methodology and library.
Comments on S3

@@
-\section{Scala Parsing Library: Packrat Parsers}\label{sec:packratparsers}
+\section{Background: Scala Packrat Parsing Library}\label{sec:packratparsers}
 
-We have selected Packrat Parsing library for our prototype implementation. It is a library built on Scala's basic parsing features,
-which opens access to high performance in parsing, convenient parser combinators with concise syntax, as well as lexers for users. Generally,
-a Packrat parser has type \lstinline{PackratParser[E]} for some \lstinline{E}, which indicates the representation. Below we present a small piece of
-code that uses Packrat parsing for our illustration.
-Note that it should be defined in an enclosing type that extends \lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers} for lexing and \lstinline{scala.util.parsing.combinator.PackratParsers} for Packrat parsers.
+\bruno{We need to think whether this section should be incoorporated
+  in S2}
+
+This section introduces the Scala Packrat Parsing library~\cite{}, which is
+used by \name.
+
+\paragraph{An Example}
+A Packrat parser has type \lstinline{PackratParser[E]} for some
+\lstinline{E}, which indicates the representation\bruno{the
+  representation of what?}. The code below illustrates
+Scala's Packrat parsing library.
 
@@
 
-The first two lines stand for lexing.
-In the third line we define a parser which can, for instance, parse \lstinline{"str(5)"} and produce the string \lstinline{"5"}.
-In \lstinline{StandardTokenParsers} there are a number of basic parsers, such as \lstinline{numericLit} that parsers integers, and \lstinline{stringLit} that parsers string literals for convenient use.
-Moreover, in Scala the syntax for parsing is quite concise and clear, as we have shown above, with the help of parser combinators.
-\lstinline{~>} and \lstinline{<~} only collect results from the parts that arrows point to, and \lstinline{^^} is followed by a function which is applied to a parser to return new values.
-Finally, we can define a generic \lstinline{runParser} function for testing:
+\noindent
+
+Note that the code should be defined in an enclosing type that extends
+\lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers}
+for lexing and
+\lstinline{scala.util.parsing.combinator.PackratParsers} for Packrat
+parsers.
+The first two lines are used for lexing. The third line defines a
+parser which can, for instance, parse \lstinline{"str(5)"} and produce
+the string \lstinline{"5"}. In \lstinline{StandardTokenParsers} there
+are a number of basic parsers, such as \lstinline{numericLit} that
+parses integers, and \lstinline{stringLit} that parsers string
+literals. Moreover, in Scala the syntax for parsing
+is quite concise and clear, as shown above, with the help of
+parser combinators. The combinators \lstinline{~>} and \lstinline{<~} only collect
+results from the parts that arrows point to, and the combinator \lstinline{^^} is
+followed by a function which is applied to a parser to return new
+values. Finally, a generic \lstinline{runParser}
+function is used for testing:
 
@@
 \hspace{.2in}- A parser combinator for function application. \\
+\hline
+\begin{lstlisting}
+def ^^^[U](v: => U): Parser[U]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
 \hline
@@
 \end{tabular}
-\caption{AAA}\label{tab1}
-\end{table}
+\caption{Some frequently used functions in Scala parser API.}\label{tab1}
+\end{table}\bruno{We should double-check that this table lists the
+  combinators used in the paper.}
 
-Furthermore, Table~\ref{tab1} shows some common parser combinators that are used in our implementation.
-In particular, \lstinline{|} is common syntax for alternative parsers in many languages, whereas \lstinline{|||} is a new combinator available in Scala, which is also for alternation but with longest match composition.
-Imagine that we are composing a list of parsers using alternative \lstinline{|}, in an order.
-A parser in the front might be successful in parsing a substring of the input, but then it unexpectedly stops the parsing and discards the rest, while another parser might be able to parse the whole input.
-In that case, using \lstinline{|||} we no longer need to worry about the order of composition, which potentially encourages its extensibility.
-A concrete example will be illustrated for this later.
+\paragraph{Packrat Parsing API} Table~\ref{tab1} shows some common parser combinators
+that are used in our implementation.  There are several standard
+parser combinators. For example \lstinline{|} is the combinator for
+alternative parsers. Other combinators have descriptions of their
+semantics on the table.
 
-On the other hand, the Scala Packrat Parsing library comes to our attention not only because of its performance, but also due to its support for direct left-recursion.
-Left recursion has been considered as a big problem especially in recursive descent parsers, which many existing libraries cannot perfectly support.
-From theoretical point of view, the algorithm behind Packrat parsers supports both direct and indirect left-recursion.
-The current version of the library is still buggy in some cases with indirect left-recursion, but we believe that they will fix it in the future, and with direct left-recursion it is already practical to use for a large number of applications.
+There are two features of the Packrat Parsing library that play a key role
+for modularity, which are discussed next.
+
+\paragraph{Longest Match Composition} The Packrat parsing library 
+contains an alternation combinator, which does \emph{longest match
+  composition}. This combinator has a different semantics from
+\lstinline{|}.
+Suppose that we are composing a list of parsers using alternative
+\lstinline{|}, in an order. A parser in the front might be successful
+in parsing a substring of the input, but then it unexpectedly stops
+the parsing and discards the rest, while another parser might be able
+to parse the whole input. In that case, using \lstinline{|||} we no
+longer need to worry about the order of composition, which potentially
+encourages its extensibility. A concrete example will be illustrated
+for this later. \bruno{Please show example here. }
+
+\paragraph{Left Recursion} The Scala Packrat Parsing library 
+supports direct left-recursion. Left recursion has been considered
+as a big problem especially in recursive descent parsers~\cite{}, and many
+existing parsing libraries have limited or no support for it. From theoretical point of
+view, the algorithm behind Packrat parsers supports both direct and
+indirect left-recursion. In practice, the current version of the library is still
+buggy in some cases with indirect left-recursion, but we believe that
+they will fix it in the future, and direct left-recursion is
+already practical to use for a large number of applications.

### 2016-12-12 paper/sections/Overview.tex
Done 2.2, rewriting 2.3

@@
 
-\subsection{Chosing the Parsing Technique}\label{subsec:overview-parsing}
+\subsection{Choosing the Parsing Technique}\label{subsec:overview-parsing}
 
@@
 \subsection{The Parsing Expression Problem}
-\bruno{You are not motivating the problem! You are going straight to
+\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
   the solution without pointing out what the problem is first. What
@@
 
-\bruno{I think we need to set up a challenge here, similar to the challange of
+\huang{partly done, only extend a new language construct in the example, how about leaving extension of new operations in the OA subsection?}\bruno{I think we need to set up a challenge here, similar to the challange of
   the expression problem. The challange should be like. Build a parser
@@
 
-\begin{lstlisting}[language=PlainCode]
-Expr := Var String | App Expr Expr
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <var>
+    \alt <expr> <expr>
+    \alt `(' <expr> `)'
+\end{grammar}
+
+It is straightforward to model ASTs by inheritance and write corresponding parsers for each case.
+
+\begin{lstlisting}
+abstract class Expr
+class Var(x: String) extends Expr
+class App(e1: Expr, e2: Expr) extends Expr
+
+trait BaseParser {
+  lexical.delimiters += ("(", ")")
+  val pExpr: PackratParser[Expr] = pVar ||| pApp ||| "(" ~> pExpr <~ ")"
+  val pVar: PackratParser[Expr] = ident ^^ { x => new Var(x) }
+  val pApp: PackratParser[Expr] =
+    pExpr ~ pExpr ^^ { case e1 ~ e2 => new App(e1, e2) }
+}
 \end{lstlisting}
-\bruno{if you intend to write a grammar here, then please use grammar
-  (BNF) notation, and not some notation invented by you!}
 
-\paragraph{Extensing the Parser}
+As the combination of \lstinline{pVar} and \lstinline{pApp}, \lstinline{pExpr} can parse expressions correctly.
 
-\paragraph{Problem: Hardcoded Recursive Calls}
+\paragraph{Attempt to Extend the Parser} Now consider we are extending the abstract syntax as well as the parsers.
+We add lambda abstractions as a new case of expressions.
+The language then becomes the famous \textit{untyped lambda calculus}.
+
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= ...
+    \alt `\\' <var> `.' <expr>
+\end{grammar}
+
+The ASTs can be extended trivially as below.
+
+\begin{lstlisting}
+class Lam(x: String, e: Expr) extends Expr
+\end{lstlisting}
+
+Since we already have the parser for variables and applications, we would like to build the new parser by reusing the old one.
+So we compose \lstinline{pLam}, which parses lambda abstractions, with the old parser.
+\lstinline{pExtExpr} is the new function to parse expressions.
+
+\begin{lstlisting}
+trait ExtParser extends BaseParser {
+  lexical.delimiters += ("\\", ".")
+  val pExtExpr: PackratParser[Expr] = pExpr ||| pLam ||| "(" ~> pExtExpr <~ ")"
+  val pLam: PackratParser[Expr] =
+    "\\" ~> ident ~ ("." ~> pExtExpr) ^^ { case x ~ e => new Lam(x, e) }
+}
+\end{lstlisting}
+
+Unfortunately, this extended parser does not work properly to parse all possible expressions defined by the new grammar.
+For example, it fails to parse \lstinline{(\x.x)} \lstinline{(\y.y)} which is obviously valid.
+
+\paragraph{Problem: Hard-coded Recursive Calls} The critical problem of extending parsers directly is that the recursive calls in old parsers are hard-coded.
+Specifically, \lstinline{pApp} makes two recursive calls to parse two consecutive expressions as an application.
+Before the extension, we implemented it by using \lstinline{pExpr} for those recursive calls.
+It worked fine at that time because \lstinline{pExpr} covers all cases of the old grammar.
+However, after the extension of lambda abstractions, the proper function for such recursive calls should be \lstinline{pExtExpr} instead of \lstinline{pExpr}, because the new function \lstinline{pExtExpr} covers all of the three cases in the new grammar, including the lambda case.
+
+That is the reason why the extended parser fails to parse some valid expressions such as \lstinline{(\x.x)} \lstinline{(\y.y)}.
+To resolve this issue, we need to let \lstinline{pApp} be aware of the change of grammar.
+We have some direct solutions:
+
+\begin{itemize}
+    \item Keep the old code, rewrite problematic functions like \lstinline{pApp}. \\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
+    \item Modify the old code, correct the recursive calls by proper functions.
+\\We can avoid code duplication thus get code reuse, but separate compilation is lost because of modification on the old code.
+    \item ??? \huang{third one? use dynamic languages?}
+\end{itemize}
+
+Similar to the trade-off in the Expression Problem, neither of them satisfies us.
+The challenge of building semantically modular parsers requires fundamental change of the parser structure.
+We will introduce our solution in the next sections.
 
@@
 
-The problem of hardcoded recursive calls can be solved using
-delegation, which can be encoded with open recursion. The idea is
+Hard-coded recursive calls prevent our parsers from extensibility.
+This problem can be solved using delegation, which can be encoded with \textit{open recursion}. The idea is
 simple: instead of making direct recursive calls, an additional
@@
 \begin{lstlisting}
-abstract class Expr
-case class Var(x: String) extends Expr
-case class App(e1: Expr, e2: Expr) extends Expr
+type Open[T] = (=> T) => T
 
-val pVar: PackratParser[Expr] = ident ^^ Var
+// Expr, Var and App are the same as previous section
+trait ExprParser {
+  val pVar: Open[PackratParser[Expr]] =
+    _ => ident ^^ { x => new Var(x) }
+  val pApp: Open[PackratParser[Expr]] =
+    self => self ~ self ^^ { case e1 ~ e2 => new App(e1, e2) }
+}
+\end{lstlisting}
 
-val pApp: (=> PackratParser[Expr]) => PackratParser[Expr] =
-  e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
-\end{lstlisting}
 \bruno{You are copying and pasting code into the paper. Don't do this:
@@
 knows how to use it!}
-\bruno{Please polish up code. You probably want to consistently use 
+
+\huang{done}\bruno{Please polish up code. You probably want to consistently use
 the variable name ``self'' whenever you want to refer to the whole
@@
 longwinded ``(=> PackratParser[Expr]) => PackratParser[Expr]''}
-\bruno{Also, for consistency, why isn't ``pVar'' open? I.e. using a
+
+\huang{done}\bruno{Also, for consistency, why isn't ``pVar'' open? I.e. using a
   self reference as well? I know that it is not recursive, but giving
   it a different treatment makes it more confusing for users.}
-\bruno{You should use a more OO approach in the example. 
+
+\huang{done}\bruno{You should use a more OO approach in the example.
 Rather than case classes show an Expr trait with pretty printing.
@@
 
-The abstract syntax is defined by Scala's case classes, and 
-we write parsers for the two cases \lstinline{Var} and \lstinline{App}.
-The parser \lstinline{pVar}, for parsing variables, is straightforward. It parses a
-single identifier, and its type \lstinline{PackratParser[Expr]}
-indicates the result is an \lstinline{Expr}.
-The parser \lstinline{pApp}, for parsing applications is more
-interesting. \lstinline{pApp} takes a parser of
+\lstinline{pApp} takes a parser of
 \lstinline{Expr} as an argument, and uses it to parse two consecutive
@@
 
-\begin{lstlisting}[language=PlainCode]
-Expr := Var String | App Expr Expr | Lam String Expr
-\end{lstlisting}
-\bruno{same comment as before: if this is supposed to be a grammar,
-  then use a grammar notation}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <var>
+    \alt <expr> <expr>
+    \alt `\\' <var> `.' <expr>
+\end{grammar}
 

### 2016-12-12 paper/sections/OpenRecursion.tex
Done text for methodology and library.
Using ruby tool.

@@
 In Scala, however, there is no library \lstinline{fix} function, but it can be easily implemented since Scala also supports lazy evaluation. See below:
-\begin{lstlisting}
-def fix[A](f : (=> A) => A) : A = {
-    lazy val a : A = f(a)
-    a
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=27-30]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
+
 where function \lstinline{f} has a by-name parameter of type \lstinline{"=>} \lstinline{A"}.
@@
 For instance, in the beginning we only have variables and applications in \lstinline{Expr}, to have types in our language, the \lstinline{Int} type and lambda expressions with annotation are added.
-See the new grammar below:
+See the grammar below:
 \begin{lstlisting}
@@
 abstract class Type
-case class IntT extends Type
+case class IntT() extends Type
 val pInt : (=> Pair) => Parser[Type] =
-    p => "Int" ^^ { _ => IntT() }
+    p => "Int" ^^^ IntT()
 

### 2016-12-12 paper/sections/PackratParsers.tex
Using ruby tool.

@@
 
-\begin{lstlisting}
-lexical.reserved += ("str")
-lexical.delimiters += ("(", ")")
-
-val p : PackratParser[String] =
-    "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toString }
-\end{lstlisting}
-
+\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_EXAMPLE
 \noindent
@@
 
-\begin{lstlisting}
-def runParser(p: Parser[_]): String => Unit = in => {
-    val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) println(t.get) else scala.sys.error(t.toString)
-}
-\end{lstlisting}
+\lstinputlisting[linerange=16-19]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_RUNPARSER
 Now \lstinline{runParser(p)("str(5)")} prints out \lstinline{"5"} as expected.

### 2016-12-12 Scala/Parser/src/PaperCode/SampleParser.scala

@@ -1,0 +1,33 @@
+package PaperCode
+
+import TAPL.Util._
+
+object SampleParser {
+
+//BEGIN_PACKRAT_EXAMPLE
+lexical.reserved += ("str")
+lexical.delimiters += ("(", ")")
+
+val p : PackratParser[String] =
+  "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toString }
+//END_PACKRAT_EXAMPLE
+
+//BEGIN_PACKRAT_RUNPARSER
+def runParser(p: Parser[_]): String => Unit = in => {
+  val t = phrase(p)(new lexical.Scanner(in))
+  if (t.successful) println(t.get) else scala.sys.error(t.toString)
+}
+//END_PACKRAT_RUNPARSER  
+     
+}
+
+object OpenRecursion {
+
+//BEGIN_OPENRECURSION_FIX
+def fix[A](f : (=> A) => A) : A = {
+  lazy val a : A = f(a)
+  a
+}
+//END_OPENRECURSION_FIX
+
+}

### 2016-12-12 paper/computePositions.rb

@@ -39,11 +39,11 @@ fileloc = {}
 beginpos = {}
 endpos = {}
 
-Dir['../UseMixinLombok/src/*/*.java'].each do |file| 
+Dir['../Scala/Parser/src/*/*.scala'].each do |file| 
   scan(file, beginpos, endpos, fileloc)
 end
 
-Dir['../UseMixinLombok/src/*/*/*.java'].each do |file| 
+Dir['../Scala/Parser/src/*/*/*.scala'].each do |file| 
   scan(file, beginpos, endpos, fileloc)
 end
 

### 2016-12-12 paper/sections/Overview.tex
Rewriting 2.3

@@
 It is straightforward to model ASTs by inheritance and write corresponding parsers for each case.
+\huang{This is the first time to demonstrate code in this paper.
+Readers may not know what those combinators/types are.
+Refer to the library docs here?
+move the 'scala parsing library' section to the beginning as 'background'?}
 
@@
   lexical.delimiters += ("\\", ".")
-  val pExtExpr: PackratParser[Expr] = pExpr ||| pLam ||| "(" ~> pExtExpr <~ ")"
+  val pExtExpr: PackratParser[Expr] =
+    pVar ||| pApp ||| pLam ||| "(" ~> pExtExpr <~ ")"
   val pLam: PackratParser[Expr] =
@@
 
-\paragraph{Open Parsing}
-The example in Section \ref{} can be rewritten as follows:
+\paragraph{Open Parsing} Consider the expression language of only variables and applications in the previous section. The parser can be rewritten as follows:
 
+\huang{todo: polish it}
 \begin{lstlisting}
@@
 
-// Expr, Var and App are the same as previous section
-trait ExprParser {
+// Expr, Var and App are the same as previous
+trait BaseParser {
+  lexical.delimiters += ("(", ")")
   val pVar: Open[PackratParser[Expr]] =
-    _ => ident ^^ { x => new Var(x) }
+    self => ident ^^ { x => new Var(x) }
   val pApp: Open[PackratParser[Expr]] =
     self => self ~ self ^^ { case e1 ~ e2 => new App(e1, e2) }
+  val pExpr: Open[PackratParser[Expr]] =
+    self => pVar(self) ||| pApp(self) ||| "(" ~> self <~ ")"
 }
+\end{lstlisting}
+
+According to the definition of type synonym \lstinline{Open[T]}, every parse function takes a parser of \lstinline{Expr} as a parameter, namely \lstinline{self}.
+This parameter \lstinline{self} represents \emph{the parser} of expressions, which is dynamiclly explained by the argument.
+Using an abstract parameter provides extensibility to parsers, because the recursive calls are not hard-coded here --- they are dynamic and can be changed corresponding to the grammar!
+
+We take \lstinline{pApp} as an example again.
+It uses \lstinline{self} instead of hard-coded \lstinline{pExpr} for recursive parsing.
+Depending on the real argument, \lstinline{self} can be a parser which covers variables and applications, or a parser which also supports the lambda case.
+
+\paragraph{Client Code} Since the type of parsing functions are changed, we can not apply the parsers directly.
+The \lstinline{fix} function below, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.
+
+\begin{lstlisting}
+def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
+
+def parse[E](p: Open[PackratParser[E]])(inp: String): E = {
+  val t = phrase(fix(p))(new lexical.Scanner(inp))
+  if (t.successful) t.get else scala.sys.error(t.toString)
+}
+
+println(parse(new BaseParser {}.pExpr)("x y"))
 \end{lstlisting}
@@
 
-\lstinline{pApp} takes a parser of
-\lstinline{Expr} as an argument, and uses it to parse two consecutive
-expressions as an application. This argument \lstinline{e} represents
-the parser of \emph{the whole expression}. In this particular case, it
-is in fact the combination of \lstinline{pVar} and
-\lstinline{pApp}. We can replace it by \lstinline{pVar | pApp} and the
-code still works. However, using an abstract argument provides
-extensibility to the parser. Because once the abstract syntax is
-extended later, the parser of the whole expression will change
-correspondingly. Then we have to edit existing code if it is
-hard-coded. The parameterization technique used in \lstinline{pApp} is
-called \textit{open recursion}, which is discussed in section
-\ref{sec:openandparsing}.
-
-\paragraph{Parsing Components}\bruno{consider removing and/or moving
-  to later parts of the paper}
-We can use a trait to glue them together, and define
-\lstinline{pVarApp} as a combination of the two parsers. A type
-synonym \lstinline{Fix} is used to shorten the types.
+\paragraph{Extensibility} As same as what we did in previous section, we extend the grammar by adding lambda abstractions as a new case.
+Notice that the old code is not touched, hence separate compilation is obtained.
 
 \begin{lstlisting}
-type Fix[T] = (=>T) => T
-
-trait ExprParser {
-  lexical.delimiters += ("(", ")")
-
-  val pVar: PackratParser[Expr] =
-    ident ^^ Var
-  val pApp: Fix[PackratParser[Expr]] =
-    e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
-  val pVarApp: Fix[PackratParser[Expr]] =
-    e => pApp(e) | pVar | "(" ~> e <~ ")"
-}
-\end{lstlisting}
-\bruno{Fix is not the right name. A better name is Open (for open
-  recursion).}
-\bruno{I'm not sure that showing the code for language components is
-  essential at this point. This can be shown latter in the paper.}
-
-\paragraph{Client Code}
-Finally, we are able to parse \lstinline{Expr} frome strings as shown in the code below.
-The \lstinline{fix} function here is just the standard \textit{fixpoint}, it is explained in section \ref{subsec:openrecursion}.
-
-\begin{lstlisting}
-def parse[E](p: Fix[PackratParser[E]])(inp: String): E = {
-  val t = phrase(fix(p))(new lexical.Scanner(inp))
-  if (t.successful) t.get else scala.sys.error(t.toString)
-}
-
-println(parse(new ExprParser {}.pVarApp)("x y"))
-\end{lstlisting}
-
-It will print the AST as \lstinline{App(Var(x),Var(y))}.
-
-%\subsection{Extend and Reuse Parsers}\label{subsec:overview-extend}
-
-\paragraph{Extensibility} To demonstrate the advantages of writing
-parsers with open recursion, we extend the language to add lambda
-abstractions as a new case of expressions. The language then
-becomes the famous \textit{untyped lambda calculus}.
-
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <var>
-    \alt <expr> <expr>
-    \alt `\\' <var> `.' <expr>
-\end{grammar}
-
-With Scala's case classes, extending the abstract syntax \lstinline{Expr} is easy.
-
-\begin{lstlisting}
-case class Lam(x: String, e: Expr) extends Expr
-\end{lstlisting}
-
-Since we already have the parser for variables and applications, we
-want to reuse it and build a new parser by extending it. It is just
-what the code below does. Notice that the old code is not touched,
-hence separate compilation is obtained.
-
-\begin{lstlisting}
-trait LambdaParser extends ExprParser {
+trait ExtParser extends BaseParser {
   lexical.delimiters += ("\\", ".")
-
-  val pLam: Fix[PackratParser[Expr]] =
-    e => "\\" ~> ident ~ ("." ~> e) ^^ { case x ~ b => Lam(x, b) }
-  val pVarAppLam: Fix[PackratParser[Expr]] =
-    e => pLam(e) | pVarApp(e)
+  val pLam: Open[PackratParser[Expr]] =
+    self => "\\" ~> ident ~ ("." ~> self) ^^ { case x ~ b => new Lam(x, b) }
+  val pExtExpr: Open[PackratParser[Expr]] =
+    self => pLam(self) ||| pExpr(self)
 }
@@
 
+\huang{todo: rewrite}
 In the new parser \lstinline{LambdaParser} extended from the old one,
@@
 \begin{lstlisting}
-println(parse(new LambdaParser {}.pVarAppLam)("(\\x.x) y"))
+println(parse(new ExtParser {}.pExtExpr)("(\\x.x) (\\y.y)"))
 \end{lstlisting}
 
-The code above will print \lstinline{App(Lam(x,Var(x)),Var(y))} from parsing \lstinline{"(\\x.x)} \lstinline{y"}.
+Parsing with open recursion will be discussed further in section \ref{sec:openandparsing}.
 

### 2016-12-13 paper/sections/Overview.tex
Progress

@@
 
+Modular parsing based on parser generators is supported by many libraries [].
+Users can separate the concrete/abstract syntax and related parsing code into reusable components.
+Then the corresponding parsers are built by their library utilities.
+For example, NOA [] uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers.
+However, such generation procedure requires a whole compilation after the collection of all grammar pieces.
+Once the grammar changed, even slightly, grammar information must be re-collected and the parser must be re-generated.
+Hence thoes libraries only have syntactic modularity.
+
+Generating parsers often requires full information of grammars, thus semantic modularity is difficult to achieve in this way.
+
 \huang{a paragraph here, refer to related work section, discuss
   previous works of modular parsing using parser generators and its
-  drawbacks}\bruno{Please finish this.}
+  drawbacks}\bruno{Please finish this.}\huang{done}
 
@@
 including the famous library Parsec~\cite{} library, require
-programmers to manually do\textit{left-recursion elimination}, and
+programmers to manually do \textit{left-recursion elimination}, and
 require significant amounts \textit{backtracking}. Both of these are
@@
 
-\huang{todo: polish it}
 \begin{lstlisting}
@@
 It uses \lstinline{self} instead of hard-coded \lstinline{pExpr} for recursive parsing.
-Depending on the real argument, \lstinline{self} can be a parser which covers variables and applications, or a parser which also supports the lambda case.
+Depending on the real argument, \lstinline{self} can be a parser which covers only variables and applications, or a parser which also supports the lambda case.
 
-\paragraph{Client Code} Since the type of parsing functions are changed, we can not apply the parsers directly.
-The \lstinline{fix} function below, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.
+\paragraph{Client Code} Since the type of parsing functions are changed, we can not apply them directly.
+The client code below shows how to use the new parser.
+The \lstinline{fix} function, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.
 
@@
 \paragraph{Extensibility} As same as what we did in previous section, we extend the grammar by adding lambda abstractions as a new case.
+Similarly, the new parser extends from the old one by adding \lstinline{pLam} on it.
 Notice that the old code is not touched, hence separate compilation is obtained.
@@
 
-\huang{todo: rewrite}
-In the new parser \lstinline{LambdaParser} extended from the old one,
-we write a new parsing function \lstinline{pLam} for the new case
-lambda abstraction. And compose it with \lstinline{pVarApp}, which we
-already have in the old parser \lstinline{ExprParser}. Open recursion
-is critical to achieve such type-safe compositions, because wherever
-we refer to the parser of the whole expression, it is updated to cover
-all of the three case we have now.
+All recursive calls of parsing `an expression' are called via the explicit parameter \lstinline{self}.
+Once the \lstinline{fix} function is applied, these calls will be updated appropriately.
+Parsing functions such as \lstinline{pApp} do not need to be rewritten.
+As a consequence, this parser recognizes all valid expressions of the new grammar, including \lstinline{(\\x.x)} \lstinline{(\\y.y)} which cannot be parsed in the previous section.
 
@@
 
-Parsing with open recursion will be discussed further in section \ref{sec:openandparsing}.
+As we demonstrated, delegation encoded by open recursion is the key technique to obtain semantic modularity.
+It enables type-safe code reuse and separate compilation in our parsers.
+We will discuss this topic further in section \ref{sec:openandparsing}.
 

### 2016-12-13 paper/sections/Overview.tex
Example

@@
 without any token consumption. If two alternatives have a same
-non-empty prefix, an auxiliary function \lstinline{try} must be added
+non-empty prefix, an auxiliary function \inlinecode{try} must be added
 to backtrack. However, if we want to extend our parsers later, we
 should always consider the worst case in which all alternatives share
-common prefixes. Then we need to add \lstinline{try} for all the
+common prefixes. Then we need to add \inlinecode{try} for all the
 branches, which results worst case exponential time complexity.
-\bruno{you should again illustrate your point here with an
+\huang{done}\bruno{you should again illustrate your point here with an
   example. The example should illustrate that the addition of a new
   case may require introducing backtracking. }
+
+For example, suppose that we only have \inlinecode{import..from} statements before, now we want to extend the grammar to add an \inlinecode{import..as} case.
+The third line of grammar below is the new case.
+
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<stmt> ::= `import' <ident> `from' <ident>
+    \alt ...
+    \alt `import' <ident> `as' <ident>
+\end{grammar}
+
+Because the new case \inlinecode{import..as} shares a prefix with the old case \inlinecode{import..from}, we must backtrack in order to cover all possibilities.
+
+\begin{lstlisting}[language=PlainCode]
+oldParser = parseImportFrom <|> ...
+newParser = try parseImportFrom <|> ... <|> parseImportAs
+\end{lstlisting}
 
@@
 
-As the combination of \lstinline{pVar} and \lstinline{pApp}, \lstinline{pExpr} can parse expressions correctly.
+As the combination of \inlinecode{pVar} and \inlinecode{pApp}, \inlinecode{pExpr} can parse expressions correctly.
 
@@
 Since we already have the parser for variables and applications, we would like to build the new parser by reusing the old one.
-So we compose \lstinline{pLam}, which parses lambda abstractions, with the old parser.
-\lstinline{pExtExpr} is the new function to parse expressions.
+So we compose \inlinecode{pLam}, which parses lambda abstractions, with the old parser.
+\inlinecode{pExtExpr} is the new function to parse expressions.
 
@@
 Unfortunately, this extended parser does not work properly to parse all possible expressions defined by the new grammar.
-For example, it fails to parse \lstinline{(\x.x)} \lstinline{(\y.y)} which is obviously valid.
+For example, it fails to parse \inlinecode{(\x.x)} \inlinecode{(\y.y)} which is obviously valid.
 
 \paragraph{Problem: Hard-coded Recursive Calls} The critical problem of extending parsers directly is that the recursive calls in old parsers are hard-coded.
-Specifically, \lstinline{pApp} makes two recursive calls to parse two consecutive expressions as an application.
-Before the extension, we implemented it by using \lstinline{pExpr} for those recursive calls.
-It worked fine at that time because \lstinline{pExpr} covers all cases of the old grammar.
-However, after the extension of lambda abstractions, the proper function for such recursive calls should be \lstinline{pExtExpr} instead of \lstinline{pExpr}, because the new function \lstinline{pExtExpr} covers all of the three cases in the new grammar, including the lambda case.
+Specifically, \inlinecode{pApp} makes two recursive calls to parse two consecutive expressions as an application.
+Before the extension, we implemented it by using \inlinecode{pExpr} for those recursive calls.
+It worked fine at that time because \inlinecode{pExpr} covers all cases of the old grammar.
+However, after the extension of lambda abstractions, the proper function for such recursive calls should be \inlinecode{pExtExpr} instead of \inlinecode{pExpr}, because the new function \inlinecode{pExtExpr} covers all of the three cases in the new grammar, including the lambda case.
 
-That is the reason why the extended parser fails to parse some valid expressions such as \lstinline{(\x.x)} \lstinline{(\y.y)}.
-To resolve this issue, we need to let \lstinline{pApp} be aware of the change of grammar.
+That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{(\x.x)} \inlinecode{(\y.y)}.
+To resolve this issue, we need to let \inlinecode{pApp} be aware of the change of grammar.
 We have some direct solutions:
@@
 \begin{itemize}
-    \item Keep the old code, rewrite problematic functions like \lstinline{pApp}. \\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
+    \item Keep the old code, rewrite problematic functions like \inlinecode{pApp}. \\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
     \item Modify the old code, correct the recursive calls by proper functions.
@@
 
-According to the definition of type synonym \lstinline{Open[T]}, every parse function takes a parser of \lstinline{Expr} as a parameter, namely \lstinline{self}.
-This parameter \lstinline{self} represents \emph{the parser} of expressions, which is dynamiclly explained by the argument.
+According to the definition of type synonym \inlinecode{Open[T]}, every parse function takes a parser of \inlinecode{Expr} as a parameter, namely \inlinecode{self}.
+This parameter \inlinecode{self} represents \emph{the parser} of expressions, which is dynamiclly explained by the argument.
 Using an abstract parameter provides extensibility to parsers, because the recursive calls are not hard-coded here --- they are dynamic and can be changed corresponding to the grammar!
 
-We take \lstinline{pApp} as an example again.
-It uses \lstinline{self} instead of hard-coded \lstinline{pExpr} for recursive parsing.
-Depending on the real argument, \lstinline{self} can be a parser which covers only variables and applications, or a parser which also supports the lambda case.
+We take \inlinecode{pApp} as an example again.
+It uses \inlinecode{self} instead of hard-coded \inlinecode{pExpr} for recursive parsing.
+Depending on the real argument, \inlinecode{self} can be a parser which covers only variables and applications, or a parser which also supports the lambda case.
 
@@
 The client code below shows how to use the new parser.
-The \lstinline{fix} function, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.
+The \inlinecode{fix} function, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.
 
@@
 \paragraph{Extensibility} As same as what we did in previous section, we extend the grammar by adding lambda abstractions as a new case.
-Similarly, the new parser extends from the old one by adding \lstinline{pLam} on it.
+Similarly, the new parser extends from the old one by adding \inlinecode{pLam} on it.
 Notice that the old code is not touched, hence separate compilation is obtained.
@@
 
-All recursive calls of parsing `an expression' are called via the explicit parameter \lstinline{self}.
-Once the \lstinline{fix} function is applied, these calls will be updated appropriately.
-Parsing functions such as \lstinline{pApp} do not need to be rewritten.
-As a consequence, this parser recognizes all valid expressions of the new grammar, including \lstinline{(\\x.x)} \lstinline{(\\y.y)} which cannot be parsed in the previous section.
+All recursive calls of parsing `an expression' are called via the explicit parameter \inlinecode{self}.
+Once the \inlinecode{fix} function is applied, these calls will be updated appropriately.
+Parsing functions such as \inlinecode{pApp} do not need to be rewritten.
+As a consequence, this parser recognizes all valid expressions of the new grammar, including \inlinecode{(\\x.x)} \inlinecode{(\\y.y)} which cannot be parsed in the previous section.
 
@@
 And also parser for it as below.
-Notice the parsing function \lstinline{pVarApp} takes an argument of type \lstinline{Alg[E]}, that means it accepts any instance of the algebra interface \lstinline{Alg}.
+Notice the parsing function \inlinecode{pVarApp} takes an argument of type \inlinecode{Alg[E]}, that means it accepts any instance of the algebra interface \inlinecode{Alg}.
 
@@
 To use the parser, we must provide an algebra instance as the operation to construct the parsing result.
-As shown in the code below, we feed the pretty printing operation \lstinline{VarApp.Print} to the generic \lstinline{parse} function.
-It will print \lstinline{(x y)} as the result.
+As shown in the code below, we feed the pretty printing operation \inlinecode{VarApp.Print} to the generic \inlinecode{parse} function.
+It will print \inlinecode{(x y)} as the result.
 

### 2016-12-14 paper/sections/Overview.tex
Done 2.1-2.3

@@
 Once the grammar changed, even slightly, grammar information must be re-collected and the parser must be re-generated.
-Hence thoes libraries only have syntactic modularity.
+Hence those libraries only have syntactic modularity.
 
@@
 
-Unfortunatelly many parser combinators have important limitations for
+Unfortunately many parser combinators have important limitations for
 modular parsing. In particular several parser combinators,
@@
 
-\paragraph{Left-Recursion Elimination} The main problem with
-left-recursion elimination is that it is a \emph{global}
-transformation on a grammar. Given that the full grammar is known,
-then it is possible to remove all left-recursive cases.
-\bruno{Illustrate with a concrete example: show a simple
+\paragraph{Left-Recursion Elimination} The top-down, recursive descent parsing strategy adopted by those parser combinator libraries cannot support left-recursive grammars directly.
+The common solution is to rewrite the grammar into an equivalent but not left-recursive one, so called left-recursion elimination.
+
+A left-recursive grammar of adding integers can be rewritten as below. On the right side, we got a new grammar without left-recursion.
+
+\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int> \alt <expr> `+' <int>
+\end{grammar}
+&
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int> <expr'>
+
+<expr'> ::= <empty> \alt `+' <int> <expr'>
+\end{grammar}
+\end{tabular}
+
+The main problem with left-recursion elimination is that it is a \emph{global} transformation on a grammar.
+Given that the full grammar is known, then it is possible to remove all left-recursive cases.
+However, when doing modular parsing, the full grammar is not known!
+
+For the example above, if we extend the original grammar to support subtraction, we must analyse the full grammar again to rewrite it.
+
+\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int> \alt <expr> `+' <int> \alt <expr> `-' <int>
+\end{grammar}
+&
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int> <expr'>
+
+<expr'> ::= <empty> \alt `+' <int> <expr'> \alt `-' <int> <expr'>
+\end{grammar}
+\end{tabular}
+
+Another issue of left-recursion elimination is that it requires extra bookkeeping work to retain the original semantic.
+For example, the expression $1+2-3$ is parsed as $(1+2)-3$ in the left-recursive grammar, but after rewrite the result is $1((+2)-3))$.
+The parse tree must be transformed to recover its structure.
+
+\huang{done, the example may be not very good, but i haven't found a better one}\bruno{Illustrate with a concrete example: show a simple
   left-recursive grammar; then do left-recursion elimination. Then add
@@
 a very different grammar is obtained after left-recursion elimination.}
-
-However, when doing modular parsing, the full grammar is not known!
-
-Parser combinators adopt top-down, recursive descent parsing, which
-uses backtracking to search through the possible branches. However,
-simple backtracking parser combinators such as the famous library
-Parsec~\cite{} in Haskell, have some drawbacks. One is that they cannot
-support left-recursive grammars. The common solution is to rewrite a
-left-recursive grammar into an equivalent one, so called
-\textit{left-recursion elimination}. This solution requires
-information of the whole grammar, so that it cannot be applied in a
-modular setting. Because the grammar may be extended in the future,
-essentially we only have parts of the grammar.
 
@@
 \paragraph{Packrat Parsing}
-Fortunatelly more advanced parsing techniques such as Packrat
+Fortunately more advanced parsing techniques such as Packrat
 parsing~\cite{}, address the limitations of simple parser combinators
@@
 
-\huang{partly done, only extend a new language construct in the example, how about leaving extension of new operations in the OA subsection?}\bruno{I think we need to set up a challenge here, similar to the challange of
-  the expression problem. The challange should be like. Build a parser
+\huang{partly done, only extend a new language construct in the example, how about leaving extension of new operations in the OA subsection?}\bruno{I think we need to set up a challenge here, similar to the challenge of
+  the expression problem. The challenge should be like. Build a parser
 for a simple expression language, then extend the language with both
@@
 \begin{grammar}
-<expr> ::= <var>
+<expr> ::= <ident>
     \alt <expr> <expr>
@@
 <expr> ::= ...
-    \alt `\\' <var> `.' <expr>
+    \alt `\\' <ident> `.' <expr>
 \end{grammar}

### 2016-12-14 paper/sections/Overview.tex
Remove the pasted code

@@
 
-\begin{lstlisting}
-abstract class Expr
-class Var(x: String) extends Expr
-class App(e1: Expr, e2: Expr) extends Expr
-
-trait BaseParser {
-  lexical.delimiters += ("(", ")")
-  val pExpr: PackratParser[Expr] = pVar ||| pApp ||| "(" ~> pExpr <~ ")"
-  val pVar: PackratParser[Expr] = ident ^^ { x => new Var(x) }
-  val pApp: PackratParser[Expr] =
-    pExpr ~ pExpr ^^ { case e1 ~ e2 => new App(e1, e2) }
-}
-\end{lstlisting}
+\lstinputlisting[linerange=8-18]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
 
@@
 
-\begin{lstlisting}
-class Lam(x: String, e: Expr) extends Expr
-\end{lstlisting}
+\lstinputlisting[linerange=22-22]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
 
@@
 
-\begin{lstlisting}
-trait ExtParser extends BaseParser {
-  lexical.delimiters += ("\\", ".")
-  val pExtExpr: PackratParser[Expr] =
-    pVar ||| pApp ||| pLam ||| "(" ~> pExtExpr <~ ")"
-  val pLam: PackratParser[Expr] =
-    "\\" ~> ident ~ ("." ~> pExtExpr) ^^ { case x ~ e => new Lam(x, e) }
-}
-\end{lstlisting}
+\lstinputlisting[linerange=26-32]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXT
 
@@
 
-\begin{lstlisting}
-type Open[T] = (=> T) => T
-
-// Expr, Var and App are the same as previous
-trait BaseParser {
-  lexical.delimiters += ("(", ")")
-  val pVar: Open[PackratParser[Expr]] =
-    self => ident ^^ { x => new Var(x) }
-  val pApp: Open[PackratParser[Expr]] =
-    self => self ~ self ^^ { case e1 ~ e2 => new App(e1, e2) }
-  val pExpr: Open[PackratParser[Expr]] =
-    self => pVar(self) ||| pApp(self) ||| "(" ~> self <~ ")"
-}
-\end{lstlisting}
+\lstinputlisting[linerange=53-63]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_BASE
 
@@
 
-\begin{lstlisting}
-def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
-
-def parse[E](p: Open[PackratParser[E]])(inp: String): E = {
-  val t = phrase(fix(p))(new lexical.Scanner(inp))
-  if (t.successful) t.get else scala.sys.error(t.toString)
-}
-
-println(parse(new BaseParser {}.pExpr)("x y"))
-\end{lstlisting}
-
-\bruno{You are copying and pasting code into the paper. Don't do this:
-code will change alot (various edits), and it is all too easy to get
-inconsistent code snippets. Instead use the small Ruby tool that we
-have in the paper directory to import source code directly from the
-code repository. This will ensure the consistency of the code
-throughout the paper writing. Please ask Haoyuan about the tool; he
-knows how to use it!}
+\lstinputlisting[linerange=68-75]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
 
@@
 
-\begin{lstlisting}
-trait ExtParser extends BaseParser {
-  lexical.delimiters += ("\\", ".")
-  val pLam: Open[PackratParser[Expr]] =
-    self => "\\" ~> ident ~ ("." ~> self) ^^ { case x ~ b => new Lam(x, b) }
-  val pExtExpr: Open[PackratParser[Expr]] =
-    self => pLam(self) ||| pExpr(self)
-}
-\end{lstlisting}
+\lstinputlisting[linerange=80-86]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
 
@@
 
-\begin{lstlisting}
-println(parse(new ExtParser {}.pExtExpr)("(\\x.x) (\\y.y)"))
-\end{lstlisting}
+\lstinputlisting[linerange=95-95]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
 

### 2016-12-14 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -1,0 +1,102 @@
+package PaperCode.Overview
+
+import PaperCode.Overview.Util._
+
+object Overview2 {
+
+//BEGIN_OVERVIEW_SIMPLE_EXPR
+abstract class Expr
+class Var(x: String) extends Expr
+class App(e1: Expr, e2: Expr) extends Expr
+
+trait BaseParser {
+  lexical.delimiters += ("(", ")")
+  val pExpr: PackratParser[Expr] = pVar ||| pApp ||| "(" ~> pExpr <~ ")"
+  val pVar: PackratParser[Expr] = ident ^^ { x => new Var(x) }
+  val pApp: PackratParser[Expr] =
+    pExpr ~ pExpr ^^ { case e1 ~ e2 => new App(e1, e2) }
+}
+//END_OVERVIEW_SIMPLE_EXPR
+
+//BEGIN_OVERVIEW_SIMPLE_LAM
+class Lam(x: String, e: Expr) extends Expr
+//END_OVERVIEW_SIMPLE_LAM
+
+//BEGIN_OVERVIEW_SIMPLE_EXT
+trait ExtParser extends BaseParser {
+  lexical.delimiters += ("\\", ".")
+  val pExtExpr: PackratParser[Expr] =
+    pExpr ||| pLam ||| "(" ~> pExtExpr <~ ")"
+  val pLam: PackratParser[Expr] =
+    "\\" ~> ident ~ ("." ~> pExtExpr) ^^ { case x ~ e => new Lam(x, e) }
+}
+//END_OVERVIEW_SIMPLE_EXT
+
+  def main(args: Array[String]): Unit = {
+    //val parser = new BaseParser {}.pExpr
+    //runParser(parser)("a b c")
+
+    val extParser = new ExtParser {}.pExtExpr
+    runParser(extParser)("a b c")
+    runParser(extParser)("(\\x.x) (\\y.y)")
+  }
+}
+
+object Overview3 {
+
+  abstract class Expr
+  class Var(x: String) extends Expr
+  class App(e1: Expr, e2: Expr) extends Expr
+  class Lam(x: String, e: Expr) extends Expr
+
+//BEGIN_OVERVIEW_OPEN_BASE
+type Open[T] = (=> T) => T
+
+trait BaseParser {
+  lexical.delimiters += ("(", ")")
+  val pVar: Open[PackratParser[Expr]] =
+    self => ident ^^ { x => new Var(x) }
[…]

### 2016-12-14 Scala/Parser/src/PaperCode/Overview/Util.scala

@@ -1,0 +1,23 @@
+package PaperCode.Overview
+
+import scala.util.parsing.combinator._
+import scala.util.parsing.combinator.syntactical._
+
+object Util extends StandardTokenParsers with PackratParsers {
+  val lcid = ident ^? { case id if id.charAt(0).isLower => id }
+  val ucid = ident ^? { case id if id.charAt(0).isUpper => id }
+
+  type Open[T] = (=> T) => T
+
+  implicit class Alternative[E, F, G, H](x: F => (=> H) => PackratParser[E]) {
+    def |(y: G => (=> H) => PackratParser[E]): (F with G) => (=> H) => PackratParser[E] =
+      alg => l => this.x(alg)(l) ||| y(alg)(l)
+  }
+
+  def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
+
+  def runParser(p: Parser[_]): String => Unit = in => {
+    val t = phrase(p)(new lexical.Scanner(in))
+    if (t.successful) println(t.get) else scala.sys.error(t.toString)
+  }
+}

### 2016-12-15 paper/sections/OpenRecursion.tex
Using ruby script.

@@
 Here we illustrate by the famous Fibonacci example. Traditionally it can be implemented in Scala using pattern-matching:
-\begin{lstlisting}
-def fib : Int => Int = i => i match {
-    case 0 => 0
-    case 1 => 1
-    case n => fib(n - 1) + fib(n - 2)
-}
-
-val x = fib(2) // 1
-\end{lstlisting}
+\lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
 An alternative implementation using open recursion is as follows:
-\begin{lstlisting}
-type Fix[T] = (=> T) => T
-def fib2 : Fix[Int => Int] = self => i => i match {
-    case 0 => 0
-    case 1 => 1
-    case n => self.apply(n - 1) + self.apply(n - 2)
-}
-
-val x2 = fix(fib2)(2) // 1
-\end{lstlisting}
+\lstinputlisting[linerange=44-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
 where \lstinline{self} is the explicit self-reference, namely the delegator. Functions like \lstinline{fib2} which takes
@@
 Suppose now we want to make use of open recursion for additional operations. We first define a combinator which combines the results of two generators in a pair:
-\begin{lstlisting}
-implicit class Compose[A, B, C](f : Fix[A => B]) {
-    def *(g : Fix[A => C]) : Fix[A => (B, C)] = self => x =>
-        (f(self.apply(_)._1)(x), g(self.apply(_)._2)(x))
-}
-\end{lstlisting}
+\lstinputlisting[linerange=55-58]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_COMPOSE
+
 Using Scala implicit class, \lstinline{*} can be used as an infix operator for composition in convenience.
 Then we define a generator which prints out a number as a string:
-\begin{lstlisting}
-def show : Fix[Int => String] = self => x => x.toString
-\end{lstlisting}
+\lstinputlisting[linerange=62-62]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_SHOW
+
 Now \lstinline{fix(fib2 * show)(2)} results in \lstinline{(1, "2")}. We have seen that both \lstinline{fib2} and \lstinline{show} are integrated in the evaluation.
@@
 The definition is straightforward:
-\begin{lstlisting}
-def merge[E, F, G, H[_,_]](op : F => G => H[F, G], x : (=> E) => F, y : (=> E) => G) : (=> E) => H[F, G] = e => op(x(e))(y(e))
-\end{lstlisting}
+\lstinputlisting[linerange=66-66]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_MERGE
 
@@
 
-\begin{lstlisting}
-type Parser[E] = PackratParser[E]
+\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PVAR
 
-abstract class Expr
-case class Var(x : String) extends Expr
-val pVar : Parser[Expr] = ident ^^ Var
-\end{lstlisting}
 Note that we are omitting lexing and demo code here.
@@
 Not only the implementation of that parser is needed, but also its corresponding datatype:
-\begin{lstlisting}
-case class App(e1 : Expr, e2 : Expr) extends Expr
-val pApp : Parser[Expr] = pVar ~ pVar ^^ { case e1 ~ e2 => App(e1, e2) }
-\end{lstlisting}
+
+\lstinputlisting[linerange=16-17]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PAPP
+
 Such a parser \lstinline{pApp} can only parse one-layer applications like \lstinline{"x y"}. Hence we revise it using recursion:
-\begin{lstlisting}
-val pApp : Parser[Expr] = {
-    val p = pVar ||| pApp
-    p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=29-32]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PAPP2
+
 Note that we are composing \lstinline{pVar} and \lstinline{pApp} using alternative.
@@
 Besides writing a new parser for literals, we need to rewrite the existing \lstinline{pApp}, since applications can now contain numbers, like \lstinline{"x 3"} and so on.
-\begin{lstlisting}
-case class Lit(n : Int) extends Expr
-val pLit : Parser[Expr] = numericLit ^^ Lit
-val pApp : Parser[Expr] = {
-    val p = pVar ||| pApp ||| pLit
-    p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=44-49]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PLIT
+
 We have observed that \lstinline{pApp} is indeed constructed in a non-modular way, because the \lstinline{p} standing for recursive parser needs to be updated on every extension.
@@
 Note that \lstinline{p} should be defined as a by-name parameter using \lstinline{"=>"}.
-\begin{lstlisting}
-val pApp : Fix[Parser[Expr]] = p => p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
-\end{lstlisting}
+
+\lstinputlisting[linerange=63-63]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PAPPFIX
+
 Here the type \lstinline{Fix[Parser[Expr]]}, namely \lstinline{(=>} \lstinline{Parser[Expr])} \lstinline{=>} \lstinline{Parser[Expr]} is exactly what we want for a parser.
@@
 It is implemented using the general \lstinline{merge} function in Section~\ref{subsec:openrecursion}.
-\begin{lstlisting}
-val pVar : Fix[Parser[Expr]] = p => ident ^^ Var
 
-implicit class Combinator[E, R](x : (=> R) => Parser[E]) {
-    type H[A, B] = Parser[E]
-    def |||(y : (=> R) => Parser[E]) : (=> R) => Parser[E] = {
-        val alt : Parser[E] => Parser[E] => Parser[E] = a => b => a ||| b
-        merge[R, Parser[E], Parser[E], H](alt, x, y)
-    }
-}
-\end{lstlisting}
+\lstinputlisting[linerange=67-75]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_COMBINATOR
+
 Now since \lstinline{pVar} and \lstinline{pApp} are defined independently, we are able to compose them using \lstinline{|||}:
-\begin{lstlisting}
-val pVarApp : Fix[Parser[Expr]] = pVar ||| pApp
-\end{lstlisting}
+
+\lstinputlisting[linerange=79-79]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PVARAPP
+
 Here \lstinline{pVarApp} is the combination of two parsers.
@@
 Hence we use a pair in the parameter:
-\begin{lstlisting}
-type Pair = (Parser[Expr], Parser[Type])
 
-abstract class Type
-case class IntT() extends Type
-val pInt : (=> Pair) => Parser[Type] =
-    p => "Int" ^^^ IntT()
+\lstinputlisting[linerange=83-102]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_TYPEDEXPR
 
-case class Lam(x : String, t : Type, e : Expr) extends Expr
-val pLam : (=> Pair) => Parser[Expr] =
-    p => ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1) ^^
-        { case x ~ t ~ e => Lam(x, t, e) }
-
-val pVar2 : (=> Pair) => Parser[Expr] = p => pVar(p._1)
-val pApp2 : (=> Pair) => Parser[Expr] = p => pApp(p._1)
-
-val pET : Fix[Pair] = {
-    val pE = pVar2 ||| pApp2 ||| pLam
-    val pT = pInt
-    p => (pE(p), pT(p))
-}
-\end{lstlisting}
 Based on the old \lstinline{pVar} and \lstinline{pApp}, we implement \lstinline{pVar2} and \lstinline{pApp2} which take the pair as their fix-points instead.

### 2016-12-15 Scala/Parser/src/PaperCode/SampleParser.scala

@@ -30,4 +30,39 @@ def fix[A](f : (=> A) => A) : A = {
 }
 //END_OPENRECURSION_FIX
 
+//BEGIN_OPENRECURSION_FIB
+def fib : Int => Int = i => i match {
+  case 0 => 0
+  case 1 => 1
+  case n => fib(n - 1) + fib(n - 2)
+}
+
+val x = fib(2) // 1
+//END_OPENRECURSION_FIB
+
+//BEGIN_OPENRECURSION_FIB2
+type Fix[T] = (=> T) => T
+def fib2 : Fix[Int => Int] = self => i => i match {
+  case 0 => 0
+  case 1 => 1
+  case n => self.apply(n - 1) + self.apply(n - 2)
+}
+
+val x2 = fix(fib2)(2) // 1
+//END_OPENRECURSION_FIB2
+
+//BEGIN_OPENRECURSION_COMPOSE
+implicit class Compose[A, B, C](f : Fix[A => B]) {
+  def *(g : Fix[A => C]) : Fix[A => (B, C)] = self => x =>
+    (f(self.apply(_)._1)(x), g(self.apply(_)._2)(x))
+}
+//END_OPENRECURSION_COMPOSE
+
+//BEGIN_OPENRECURSION_SHOW
+def show : Fix[Int => String] = self => x => x.toString
+//END_OPENRECURSION_SHOW
+
+//BEGIN_OPENRECURSION_MERGE
+def merge[E, F, G, H[_,_]](op : F => G => H[F, G], x : (=> E) => F, y : (=> E) => G) : (=> E) => H[F, G] = e => op(x(e))(y(e))
+//END_OPENRECURSION_MERGE
 }

### 2016-12-15 paper/sections/Overview.tex
RW

@@
 offer high flexibility in the choice of operations to be performed
-over the AST.
+over the AST. It also makes parsing with multiple sorts of syntax easier.
 
-Instead of define the abstract syntax using Scala's case classes, we
-can use \textit{Object Algebras} for more extensibility. Object
-Algebras is a technique to solve the \textit{Expression Problem},
-providing the possibility of extending both data variants and
-operations over them. It will be further discussed in section
-\ref{sec:algebrasandparsing}.
+\paragraph{Parsing with Object Algebras} Using Object Algebras, the abstract syntax of the language of variables and applications is defined as below.
 
-Using Object Algebras, the abstract syntax of the language of variables and applications is defined as below.
-
-\begin{lstlisting}
-trait Alg[E] {
-  def Var(x: String): E
-  def App(e1: E, e2: E): E
-}
-\end{lstlisting}
+\lstinputlisting[linerange=8-11]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALG
 
@@
 
-\begin{lstlisting}
-trait Print extends Alg[String] {
-  def Var(x: String) = x
-  def App(e1: String, e2: String) = "(" + e1 + " " + e2 + ")"
-}
-\end{lstlisting}
+\lstinputlisting[linerange=15-18]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
 
 And also parser for it as below.
-Notice the parsing function \inlinecode{pVarApp} takes an argument of type \inlinecode{Alg[E]}, that means it accepts any instance of the algebra interface \inlinecode{Alg}.
+Notice the parsing function \inlinecode{pExpr} takes an argument of type \inlinecode{ExprAlg[E]}, that means it accepts any instance of the algebra interface \inlinecode{ExprAlg}.
 
-\begin{lstlisting}
-trait Parser[E] {
-  lexical.delimiters += ("(", ")")
-
-  val pVarApp: Alg[E] => Fix[PackratParser[E]] = alg => e =>
-    e ~ e ^^ { case e1 ~ e2 => alg.App(e1, e2) } | ident ^^ alg.Var | "(" ~> e <~ ")"
-}
-\end{lstlisting}
-
-We glue them together by defining an object. It actually becomes a \textit{language component} that represents a specific language feature.
-
-\begin{lstlisting}
-object VarApp {
-  trait Alg[E] {..}
-  trait Print extends Alg[String] {..}
-  trait Parser[E] {..}
-}
-\end{lstlisting}
+\lstinputlisting[linerange=22-28]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PARSER
 
 To use the parser, we must provide an algebra instance as the operation to construct the parsing result.
-As shown in the code below, we feed the pretty printing operation \inlinecode{VarApp.Print} to the generic \inlinecode{parse} function.
+As shown in the code below, we feed the pretty printing operation \inlinecode{ExprPrint} to the generic \inlinecode{parse} function.
 It will print \inlinecode{(x y)} as the result.
 
-\begin{lstlisting}
-def parseWithAlg[E](inp: String)(alg: VarApp.Alg[E]): E = {
-  val p = new VarApp.Parser[E] {}.pVarApp(alg)
-  parse[E](p)(inp)
-}
+\lstinputlisting[linerange=38-43]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
 
-println(parseWithAlg("x y")(new VarApp.Print{}))
-\end{lstlisting}
+\paragraph{Extensibility of Syntax} Following previous examples, the code below shows how to extend the language to support lambda abstractions.
 
-With Object Algebras, we can extend all the stuff, including abstract syntax, operations, and parsers. The code below extend the language to support lambda abstraction.
+\lstinputlisting[linerange=48-62]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
 
-\begin{lstlisting}
-object Lambda {
-  trait Alg[E] extends VarApp.Alg[E] {
-    def Lam(x: String, e: E): E
-  }
+For easily combining parsers using Object Algebras as ASTs, we have a special combinator \huang{we need to rename our combinator} defined in our library.
 
-  trait Print extends Alg[String] with VarApp.Print {
-    def Lam(x: String, e: String) = "\\" + x + "." + e
-  }
+\paragraph{Extensibility of Operations} With Object Algebras, operations over ASTs can also be extended in a modular way.
+Here is an example of collecting free variables from an expression.
+We can use this new operation during parsing.
 
-  trait Parser[E] extends VarApp.Parser[E] {
-    lexical.delimiters += ("\\", ".")
+\lstinputlisting[linerange=75-81]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
 
-    val pLam: Alg[E] => Fix[PackratParser[E]] = alg => e =>
-      ("\\" ~> lcid) ~ ("." ~> e) ^^ { case x ~ e0 => alg.Lam(x, e0) }
-    val pVarAppLam: Alg[E] => Fix[PackratParser[E]] =
-      pVarApp | pLam
-  }
-}
-\end{lstlisting}
+\paragraph{Multiple Sorts of Syntax} Another advantage of using Object Algebras is that it supports multiple sorts of syntax easily.
+In several cases, we want to divide the syntactic elements into some groups.
+For example, the grammar below has two sorts, which are expressions and types.
 
-For easily combining the parsers using Object Algebras as the representation of abstract syntax, we have a special combinator \huang{we need to rename our combinator} defined in our library.
+\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<type> ::= `int' \alt <type> `->' <type>
+\end{grammar}
+&
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <ident> \alt <expr> <expr> \alt `\\' <ident> `.' <expr>
+\end{grammar}
+\end{tabular}
 
-In addition to make extending operations possible, another advantage of using Object Algebras is that it supports multiple sorts of syntax.
-For example, if our language has not only expressions but also types, we can easily distinguish them in the abstract syntax by an extra type parameter.
-This topic will be discussed in section \ref{subsec:differentsyntax}.
+Using Object Algebras, we can easily distinguish them in an extensible way, just by adding an extra type parameter.
 
-\begin{lstlisting}
-trait TypedLambda[E, T] extends VarApp.Alg[E] {
-  def Lam(x: String, t: T, e: E): E
-  def IntType(): T
-  def BoolType(): T
-  def ArrowType(a: T, b: T): T
-}
-\end{lstlisting}
-
-% \begin{itemize}
-% \item Choosing the Parsing Technology
-%     \begin{itemize}
-%     \item Parser Generators : why not?
-%         \begin{itemize}
-%             \item Not type-safe
-%             \item No modular type-checking
-%             \item Not modular or no separate compilation (but we need to mention lots of work on extensible parsing here: example Language Workbenches; Rats)
-%         \end{itemize}
-%     \item Parser Combinators
-%         \begin{itemize}
-%         \item Backtracking parsers (Parsec)
-%             \begin{itemize}
-%             \item Need to remove left recursion (Problem: Transformation is not modular; if we do not know the full grammar then cannot be done)
-%             \item Need try/backtracking (Problem: Since we do not know the full set of rules in a modular setting, we have to assume worst case scenario and add redundant backtracking. )
-%             \item Mention some possible workarounds (there may be some but they still have issues)
-%             \end{itemize}
-%         \item (Non)-Backtracking/Packrat Parsers (Works well in a modular setting) (find a good name for this type of parsers?)
-%         \end{itemize}
-%     \end{itemize}
-% \item Adapting Parser Combinators for modularity
-%     \begin{itemize}
-%     \item Using Object Algebras
-%     \item Using Open Recursion
-%     \item Using new modular parser combinators (Library: new alternative combinator, for example)
-%     \end{itemize}
-% \item Small example
-%     \begin{itemize}
-%     \item small example of a modular parser using our technique
-%     \item show also the same example without modularity and write a detailed comparison
-%     \item Lambda Calculus is a good candidate for the example
-%     \item Show a small extension (adding plus and numeric literals)
-%     \end{itemize}
-% \end{itemize}
+\lstinputlisting[linerange=66-70]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX

### 2016-12-15 Scala/Parser/src/PaperCode/Overview/Util.scala

@@ -4,8 +4,6 @@ import scala.util.parsing.combinator._
 import scala.util.parsing.combinator.syntactical._
 
 object Util extends StandardTokenParsers with PackratParsers {
-  val lcid = ident ^? { case id if id.charAt(0).isLower => id }
-  val ucid = ident ^? { case id if id.charAt(0).isUpper => id }
 
   type Open[T] = (=> T) => T
 

### 2016-12-16 paper/sections/OpenRecursion.tex
local changes

--- 2016-12-12 paper/sections/OpenRecursion.tex
@@
-\section{Open Recursion and Parsing}\label{sec:openandparsing}
+\section{Delegation-based Parsing}\label{sec:openandparsing}
 
-Our story begins with a well-known concept mostly in object-oriented programming, called \textit{open recursion}. This mechanism
-constructs the basis of extensibility of our parsing. We will introduce the concept with illustration before integrating
-it  to achieve halfway extensible parsers.
+The first problem that prevents modularity of parsers built with
+parser combinators is the existence of \emph{hard-coded recursive
+  calls}. To solve this problem, \name employs \emph{open recursion}:
+that is, recursive calls to parsers are parametrized. The resulting
+programming style is similar to object-oriented programming, 
+but using delegation instead of class-based inheritance.
 
-\subsection{Open Recursion with Delegation}\label{subsec:openrecursion}
+\subsection{Delegation via Open Recursion}\label{subsec:openrecursion}
 
-Open recursion is known as a useful feature that one method body can ``invoke another method of the same object via a special variable called \lstinline{self} or, in some languages, \lstinline[keywords={}]{this}'' (by Ralf Hinze).
-The interesting thing is that such a variable \lstinline{self} is late-bound, or open to the recursion, which means it can integrate some additional features defined later, while existing methods that take \lstinline{self} as a parameter can simply delegate their known cases to it.
+Open recursion is known as a useful feature that one method body can
+``invoke another method of the same object via a special variable
+called \lstinline{self} or, in some languages,
+\lstinline[keywords={}]{this}'' (by Ralf Hinze). The interesting thing
+is that such a variable \lstinline{self} is late-bound, or open to the
+recursion, which means it can integrate some additional features
+defined later, while existing methods that take \lstinline{self} as a
+parameter can simply delegate their known cases to it.
 
-For sure we cannot always leave it open; upon using it we have to close the recursion, at which time we need to take a fix-point.
-Here one might suddenly be reminded of the \lstinline{fix} function in some functional languages like Haskell.
-In Haskell \lstinline{fix} is defined as:
+For sure we cannot always leave it open; upon using it we have to
+close the recursion, at which time we need to take a fix-point. Here
+one might suddenly be reminded of the \lstinline{fix} function in some
+functional languages like Haskell. In Haskell \lstinline{fix} is
+defined as:
 
@@
 \end{lstlisting}
-In Scala, however, there is no library \lstinline{fix} function, but it can be easily implemented since Scala also supports lazy evaluation. See below:
+In Scala, however, there is no library \lstinline{fix} function, but
+it can be easily implemented since Scala also supports lazy
+evaluation. See below:
 
-\lstinputlisting[linerange=27-30]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
+\begin{lstlisting}
+def fix[A](f : (=> A) => A) : A = {
+    lazy val a : A = f(a)
+    a
+}
+\end{lstlisting}
+where function \lstinline{f} has a by-name parameter of type
+\lstinline{"=>} \lstinline{A"}.  Here we illustrate by the famous
+Fibonacci example. Traditionally it can be implemented in Scala using
+pattern-matching:
 
-where function \lstinline{f} has a by-name parameter of type \lstinline{"=>} \lstinline{A"}.
-Here we illustrate by the famous Fibonacci example. Traditionally it can be implemented in Scala using pattern-matching:
 \begin{lstlisting}

### 2016-12-17 paper/sections/OpenRecursion.tex

--- 2016-12-15 paper/sections/OpenRecursion.tex
@@
 \lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
+
 An alternative implementation using open recursion is as follows:

### 2016-12-17 paper/sections/Overview.tex

@@
 \section{Overview}\label{sec:overview}
+
+\bruno{Write a summary for this section}
+\bruno{Fill in all the missing references}
 
@@
 
-The main problem with left-recursion elimination is that it is a \emph{global} transformation on a grammar.
-Given that the full grammar is known, then it is possible to remove all left-recursive cases.
-However, when doing modular parsing, the full grammar is not known!
+The main problem with left-recursion elimination is that it is a
+\emph{global} transformation on a grammar. Given that the full grammar
+is known, then it is possible to remove all left-recursive
+cases. However, when doing modular parsing, the full grammar is not
+known!
 
-For the example above, if we extend the original grammar to support subtraction, we must analyse the full grammar again to rewrite it.
+For the example above, if we extend the original grammar to support
+subtraction, we must analyse the full grammar again to rewrite it.
 
@@
 
-Another issue of left-recursion elimination is that it requires extra bookkeeping work to retain the original semantic.
-For example, the expression $1+2-3$ is parsed as $(1+2)-3$ in the left-recursive grammar, but after rewrite the result is $1((+2)-3))$.
-The parse tree must be transformed to recover its structure.
+Another issue of left-recursion elimination is that it requires extra
+bookkeeping work to retain the original semantics. For example, the
+expression $1+2-3$ is parsed as $(1+2)-3$ in the left-recursive
+grammar, but after rewrite the result is $1((+2)-3))$. The parse tree
+must be transformed to recover its structure.
 
-\huang{done, the example may be not very good, but i haven't found a better one}\bruno{Illustrate with a concrete example: show a simple
-  left-recursive grammar; then do left-recursion elimination. Then add
-a new case into the original left-recursive grammar; and show that
-a very different grammar is obtained after left-recursion elimination.}
+\huang{done, the example may be not very good, but i haven't found a
+  better one}
+\bruno{This example does not appear to be problematic: adding ``-''
+  does not affect the rest of the grammar. Isn't the problem that
+  sometimes adding a new case to the grammar, requires changing other 
+parts? Otherwise I don't see why left-recursion is a problem, and we
+cannot simply claim that it is a problem. You have to make a better
+effort to find a problematic grammar.
+}
 
@@
 
+\bruno{The example is good, I think but the explanation is not. 
+You want to say that when you know
+  the full grammar, you can figure out where ``try'' is
+  needed. Without the full grammar you'd need assume the worst
+  case. Make an effort to make your explanation cristal clear!
+I think you want to miss the abstract explanation that you give first,
+with the explanation about the concrete example. 
+Start with ``The need for backtracking is also problematic
+in a modular setting. For example, suppose ...'' and synchronize 
+the abstract explanation and the explanation for the example.
+}
+
 \paragraph{Packrat Parsing}
@@
 such as Parsec. Packrat parsers use
-memorization to record the result of applying each parser at each
+memoization to record the result of applying each parser at each
 position of the input, so that repeated computation is eliminated, and
-they supports left-recursive grammars directly. All of these properties are very
+they support left-recursive grammars directly. All of these properties are very
 suitable for modularity, thus we use Packrat parsers as the underlying
@@
 Hard-coded recursive calls prevent our parsers from extensibility.
-This problem can be solved using delegation, which can be encoded with \textit{open recursion}. The idea is
-simple: instead of making direct recursive calls, an additional
-argument, that abstracts over the recursive call, is passed to the
-parsers. This technique enables the developement of parsers that
-support OO-style extensibility of ASTs.
+This problem can be solved using delegation, which can be encoded with
+\textit{open recursion}. The idea is simple: instead of making direct
+recursive calls, an additional argument, that abstracts over the
+recursive call, is passed to the parsers. This technique enables the
+developement of parsers that support OO-style extensibility of ASTs.
 
@@
 
-According to the definition of type synonym \inlinecode{Open[T]}, every parse function takes a parser of \inlinecode{Expr} as a parameter, namely \inlinecode{self}.
-This parameter \inlinecode{self} represents \emph{the parser} of expressions, which is dynamiclly explained by the argument.
-Using an abstract parameter provides extensibility to parsers, because the recursive calls are not hard-coded here --- they are dynamic and can be changed corresponding to the grammar!
+According to the definition of type synonym \inlinecode{Open[T]},
+every parse function takes a parser of \inlinecode{Expr} as a
+parameter, namely \inlinecode{self}. This parameter \inlinecode{self}
+represents \emph{the parser} of expressions, which is dynamiclly
+explained by the argument. Using an abstract parameter provides
+extensibility to parsers, because the recursive calls are not
+hard-coded here --- they are dynamic and can be changed corresponding
+to the grammar!
 
-We take \inlinecode{pApp} as an example again.
-It uses \inlinecode{self} instead of hard-coded \inlinecode{pExpr} for recursive parsing.
-Depending on the real argument, \inlinecode{self} can be a parser which covers only variables and applications, or a parser which also supports the lambda case.
+We take \inlinecode{pApp} as an example again. It uses
+\inlinecode{self} instead of hard-coded \inlinecode{pExpr} for
+recursive parsing. Depending on the real argument, \inlinecode{self}
+can be a parser which covers only variables and applications, or a
+parser which also supports the lambda case.
 
-\paragraph{Client Code} Since the type of parsing functions are changed, we can not apply them directly.
-The client code below shows how to use the new parser.
-The \inlinecode{fix} function, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.
+\paragraph{Client Code} Since the type of parsing functions are
+changed, we can not apply them directly. The client code below shows
+how to use the new parser. The \inlinecode{fix} function, which is the
+the standard \textit{fixpoint combinator}, is used to `close' the open
+recursion. \bruno{Too much detail is being given here about open
+  recursion. In Section 2 we should talk only about \emph{how} to use 
+open recursion, and not how open recursion is implemented. We 
+can summarize, for example the API related to open recursion. 
+Section 4 is where the implementation and detailed explanation 
+about open recursion should come.
+}
 

### 2016-12-17 paper/sections/PackratParsers.tex

@@
 A Packrat parser has type \lstinline{PackratParser[E]} for some
-\lstinline{E}, which indicates the representation\bruno{the
-  representation of what?}. The code below illustrates
+\lstinline{E}, which indicates the type of results it produces.
+%\bruno{the representation of what?}
+The code below illustrates
 Scala's Packrat parsing library.
@@
 longer need to worry about the order of composition, which potentially
-encourages its extensibility. A concrete example will be illustrated
-for this later. \bruno{Please show example here. }
+encourages its extensibility. As a concrete example, consider the grammar
+\begin{grammar}
+<expr> ::= <int> \alt <expr> `+' <int>
+\end{grammar}
+which might be implemented in Packrat by \lstinline{pExpr = pInt | pAdd}, where \lstinline{pInt}
+and \lstinline{pAdd} handle the two cases respectively. It does not work as expected even for input \lstinline{"1 + 2"},
+as \lstinline{pInt}, which is applied first, parses \lstinline{1} and returns. Whereas \lstinline{pExpr = pInt ||| pAdd}
+can successfully parse the whole string. Without the longest match composition, adjusting the order of alternatives is possible
+yet inconvenient with extensible parsing.
+
+%\bruno{Please show example here. }
 

### 2016-12-17 paper/sections/RelatedWork.tex

@@
 
-Our work integrates several components including extensible parsing, parser combinators and modular datatypes. There has been a great amount of related papers
-on those hot topics, of which some inspired us of this paper and encourages us for more exploration. This section will try to lead a discussion on what difference we have made.
+Our work integrates several components including extensible parsing,
+parser combinators and extensibility techniques. There has been a
+great amount of related papers on those hot topics, of which some
+inspired us of this paper and encourages us for more exploration. This
+section will try to lead a discussion on what difference we have made.
 
-\paragraph*{Extensible Parsing} Extensible parsing is achieved in many different areas, of which parser generators are a mainstream area specially designed for modular syntax and parsing.
-Many parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support modular grammars, more specifically, they allow users to create new modules where new non-terminals and production rules can be introduced, some can even override existing rules in the old grammar modules.
-For instance, \textit{Rats!}~\cite{Grimm2006}
-constructs its own module system for the collection of grammars, while NOA~\cite{Gouseti2014} uses Java annotation processing to gather information together.
-Those parser combinators focus on the \textit{syntactic extensibility} of grammars, and rely on a whole compilation to generate a global parser, even with a slight modification.
-Some of them may statically check the correctness and unambiguity of grammars, but separate compilation and modular type-checking remain unsolved, together with the issue of performance.
+\paragraph*{Syntactically Extensible Parsing} Extensible parsing is achieved in many
+different areas, of which parser generators are a mainstream area
+specially designed for modular syntax and parsing. Many parser
+generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
+modular grammars. They allow users to create new
+modules where new non-terminals and production rules can be
+introduced, some can even override existing rules in the old grammar
+modules. For instance, \textit{Rats!}~\cite{Grimm2006} constructs its
+own module system for the collection of grammars, while
+NOA~\cite{Gouseti2014} uses Java annotation processing to gather
+information together. Those parser combinators focus on the
+\textit{syntactic extensibility} of grammars, and rely on a whole
+compilation to generate a global parser, even with a slight
+modification. Some of them may statically check the correctness and
+unambiguity of grammars, but separate compilation and modular
+type-checking remain unsolved, together with the issue of performance.
+\bruno{what performance issue?}
 
-Besides, macro systems like C preprocessor, C++ templates and Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming techniques are a similar area aiming at syntactic extensibility as well, which sacrifices type safety.
-SugarJ~\cite{Erdweg2011} is another well-known tool that conveniently introduces syntactic sugars in Java programming by library imports.
-Composition of syntactic sugars is easy for users, whereas it requires many rounds of parsing and adaption, which highly affects efficiency of compilation, moreover, the implementation was based on SDF~\cite{Heering1989} and Stratego~\cite{Visser2001}, which focused little on separate compilation.
-Extensible compilers like JastAdd~\cite{Ekman2007} and Polyglot~\cite{Nystrom2003}, however, are somehow more ambitious, but they involve extensible parsing mostly by parser generators as well, and they focus more on the extensions to a host language.
-Our library is designed for modular language parsers in a type-safe way, with flexible language composition.
-Although overriding existing production rules is tricky with our approach, we could perhaps make use of embedded domain-specific languages on top of parser combinators and transformations for overriding, but it is anyway an orthogonal issue.
-\haoyuan{???} \haoyuan{BTW what about Racket?} \haoyuan{what about metafront?
-it is a macro system but does it have type safety?} \haoyuan{Extensible syntax with lexical scoping?} \haoyuan{"Extensible syntax" proposes a system for extensible syntax, where users write EDSLs in their language with concrete syntax.
-Users can write rules for type-based disambiguation.
-But separate compilation is again not mentioned.
-Shall we mention that thesis?} \haoyuan{attribute grammars?}
+Macro systems like C preprocessor, C++ templates and
+Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming
+techniques are a similar area aiming at syntactic extensibility. 
+SugarJ~\cite{Erdweg2011} conveniently introduces syntactic sugar
+for Java using library imports. Composition of syntactic
+sugar is easy for users, but it requires many rounds of parsing
+and adaption. This significantly affects efficiency of compilation,
+moreover, the implementation was based on SDF~\cite{Heering1989} and
+Stratego~\cite{Visser2001}, which focused little on separate
+compilation\bruno{focused little or did not have it at all?}. 
+Extensible compilers like JastAdd~\cite{Ekman2007} and
+Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
+is mostly done using parser generators. Moreover 
+they focus on the extensions to a host language. Our library
+is designed for modular language parsers in a type-safe way, with
+flexible language composition. Although overriding existing production
+rules is tricky with our approach, we could perhaps make use of
+embedded domain-specific languages on top of parser combinators and
+transformations for overriding, but it is anyway an orthogonal issue.
+\bruno{Not sure that I understand the last sentence}
+\haoyuan{???} \haoyuan{BTW what about Racket?} \bruno{Have you read 
+Languages as Libraries? That is probably an important reference, which
+I think we should cite.}\haoyuan{what about
+  metafront? it is a macro system but does it have type safety?}
+\haoyuan{Extensible syntax with lexical scoping?} \haoyuan{"Extensible
+  syntax" proposes a system for extensible syntax, where users write
+  EDSLs in their language with concrete syntax. Users can write rules
+  for type-based disambiguation. But separate compilation is again not
+  mentioned. Shall we mention that thesis?} \haoyuan{attribute
+  grammars?}
 
-On the other hand, extensible parsing algorithms are another area that indeed relates to separate compilation.
-Specifically, \cite{Bravenboer2009} introduces \textit{parse table composition}, in which paper grammars are compiled to the generation of parse tables, which are DFAs or NFAs, later they can be composed by an algorithm, so as to provide separate compilation for parsing.
-Nevertheless, the generation of parse tables is rather expensive, and furthermore, our approach supports separate compilation as well as modular type-checking, and the idea is not restricted to Scala but applicable to many functional languages, on whose type system the safety of modular parsing can rely.
-The extensibility of parsing in our approach is further available at language composition and lexical level.
-\haoyuan{I mentioned one shortcoming of our approach here, which is we cannot override existing production rules; another one is that we do not have explicit correspondence/relationship between abstract syntax and the parser.}
+\paragraph{Extensible Parsing Algorithms}
+Extensible parsing algorithms are another area that indeed relates to
+separate compilation. Specifically, \cite{Bravenboer2009} introduces
+\textit{parse table composition}, in which paper grammars\bruno{what
+  is a paper grammar?} are compiled
+to the generation of parse tables\bruno{I don't understand the last
+  part; please rephrase. Also please break the sentence into multple
+  sentences. Generally speaking go over the text you wrote and watch
+  out for long sentences. You have a tendency to write very long
+  sentences. }, which are DFAs or NFAs, later they
+can be composed by an algorithm, so as to provide separate compilation
+for parsing. Nevertheless, the generation of parse tables is rather
+expensive\bruno{expensive in terms of what? why is it expensive?}. 
+Furthermore, \name supports \emph{both} separate compilation
+as well as \emph{modular type-checking}. The extensibility of
+parsing in our approach is further available at language composition
+and lexical level. \haoyuan{I mentioned one shortcoming of our
+  approach here, which is we cannot override existing production
+  rules; another one is that we do not have explicit
+  correspondence/relationship between abstract syntax and the parser.}
 
-\paragraph*{Parser Combinators} Since~\cite{burge1975} firstly introduced parser generators, and after Wadler discussed more details on backtracking in~\cite{Wadler1985}, parser combinators have been more and more popularly developed and used in the research area of parsing.
-Among them many parsing libraries produces recursive descent parsers by introducing functional, or more specifically, monadic parser combinators with a foundation on~\cite{nott237}.
-Related work includes Parsec~\cite{Leijen2001}, which is frequently used in Haskell for context-sensitive grammars with infinite lookahead.
-Nevertheless, left-recursion is known as a big issue in recursive descent parsers, in which case programmers using libraries like Parsec have to manually eliminate left-recursion in the grammar, which is cumbersome, but also it distorts the shape of its old grammar, restricting the modularity of parsing to extreme extent.
-There are certainly some approaches to afford the limitation of parser combinators, for example, we have conducted our previous experiment in Haskell, by using open recursion, MRM~\cite{Oliveira2015} and Parsec.
-To support direct left-recursive grammars, we made use of the state monad in Parsec to ensure that a left-recursive production rule will not be applied successively in parsing.
-It turned out to be successful, yet introduced complexity in programming, hence we believe that the parsing library is obligated for left-recursion.
+\paragraph*{Parser Combinators} Since~\cite{burge1975} firstly
+introduced parser generators, and after Wadler discussed more details
+on backtracking in~\cite{Wadler1985}, parser combinators have been
+more and more popularly developed and used in the research area of
+parsing. Among them many parsing libraries produces recursive descent
+parsers by introducing functional, or more specifically, monadic
+parser combinators with a foundation on~\cite{nott237}. Related work
+includes Parsec~\cite{Leijen2001}, which is frequently used in Haskell
+for context-sensitive grammars with infinite lookahead. Nevertheless,
+left-recursion is known as a big issue in recursive descent parsers,
+in which case programmers using libraries like Parsec have to manually
+eliminate left-recursion in the grammar, which is cumbersome, but also
+it distorts the shape of its old grammar, restricting the modularity
+of parsing to extreme extent. It is possible to avoid some of  
+ the limitations of parser combinators. 
+\bruno{I'm not sure that th following text should be here. It could 
+be in the paper, but somewhere else.}
+For example, we have
+conducted our previous experiment in Haskell, by using open recursion,
+MRM~\cite{Oliveira2015} and Parsec. To support direct left-recursive
+grammars, we made use of the state monad in Parsec to ensure that a
+left-recursive production rule will not be applied successively in
+parsing. It turned out to be successful, yet introduced complexity in
+programming, hence we believe that the parsing library is obligated
+for left-recursion.
 
-Some more recent papers~\cite{Ford2002,Might2011,Frost2008} \haoyuan{Pracitical, general parser combinators?} proposed a series of novel parsing techniques, moreover, they did put an eye on the issue of left-recursion.
-As we explained before, we have selected Packrat Parsing for our prototype implementation, as the combinators are convenient to use in Scala, which is a platform where functional laziness and Object Algebras can perfectly coexist.
-Also, its support for direct left-recursion is already helpful for designing real-world parsers practically, though~\cite{warth2008} further demonstrated that general left-recursion is supported from the theoretical point of view.
-We have also mentioned that the components used in our library can have alternatives, for instance, a different set of parser combinators may support ambiguous and left-recursive grammars, or be applicable to a different subset of context-free grammars or more, or even lead us to another level of performance.
+Some more recent papers~\cite{Ford2002,Might2011,Frost2008}
+\haoyuan{Pracitical, general parser combinators?} proposed a series of
+novel parsing techniques that addressed the issue
+of left-recursion. We have selected Packrat
+Parsing for our prototype implementation, as the combinators are
+convenient to use in Scala, which is a platform where functional
+laziness and Object Algebras can perfectly coexist. Also, its support
+for direct left-recursion is already helpful for designing real-world
+parsers practically, though~\cite{warth2008} further demonstrated that
+general left-recursion is supported from the theoretical point of
+view. We have also mentioned that the components used in our library
+can have alternatives, for instance, a different set of parser
+combinators may support ambiguous and left-recursive grammars, or be
+applicable to a different subset of context-free grammars or more, or
+even lead us to another level of performance.
 
-Another important issue in previous papers, to the best of our knowledge, is that none of them deal with modular datatypes or abstract syntax trees.
-The modularity of data structures is yet necessary for modular parsing that supports separate compilation and modular type-checking.
-In our library, Object Algebras have been adopted for extensible ASTs, together with the behaviors on them.
+To the best of our knowledge, no previous work on parser combinators 
+deals with semantic modularity and
+extensibility of parsing. The modularity of abstract syntax trees is
+necessary for modular parsing that supports separate compilation and
+modular type-checking. In our library, Object Algebras have been
+adopted for extensible ASTs, together with the behaviors on them.
 
-\paragraph*{Modular Datatypes} There has been lots of related work on building modular data structures, especially in some functional languages with non-extensible datatypes like Haskell.
-For instance, the famous DTC paper~\cite{swierstra2008} represents modular ASTs using co-products of every two functors, which forms a binary tree structure.
-It supports extensible datatypes, as well as the operations on them by implementing instances of functors.
-Later Oliveira et al.
-presented MRM~\cite{Oliveira2015} in Haskell, which is similar to DTC but adopts a list of functors in type-level lists instead.
-As mentioned above, we have conducted our early experiment with MRM for modular datatypes, whereas the complexity of library implementation was comparably high.
-Other related work includes~\cite{Bahr2014}, where data types are flexible with composition as well as decomposition.
+\paragraph*{Extensibility} There has been alot of work on building
+modular programs, in both functional and object-oriented programming .
+For instance, the famous DTC paper~\cite{swierstra2008} represents
+modular ASTs using co-products of every two functors, which forms a
+binary tree structure. It supports extensible datatypes, as well as
+the operations on them by implementing instances of functors. Later
+Oliveira et al. presented MRM~\cite{Oliveira2015} in Haskell, which is
+similar to DTC but adopts a list of functors in type-level lists
+instead. As mentioned above, we have conducted our early experiment
+with MRM for modular datatypes, whereas the complexity of library
+implementation was comparably high. Other related work
+includes~\cite{Bahr2014}, where data types are flexible with
+composition as well as decomposition.
 
-In OO languages, however, things get much better with the mechanism of inheritance.
-Furthermore, Object Algebras~\cite{Oliveira2012} were proposed as a design pattern for the extensibility of both data variants and operations in a modular type-safe way.
-The code is concise and expressive using Object Algebras, building a foundation for our new library with various dimensions of extensibility.
-Moreover, in~\cite{Oliveira2012} the authors have discussed the composition of algebras.
-In our parsing approach, a parser consumes an algebra, which is delegated to return the results, during its process of parsing.
-Having a set of algebras, it requires multiple parsing with several times of invocation, which leads to redundant work.
-Instead, algebras are supposed to be composed into one before the invocation of the parser.
-Bahr et al.
-lead a similar discussion in~\cite{Bahr2011}, where queries (or \textit{catamorphisms}) and transformations (or \textit{homomorphisms}) are composable.
-They have also mentioned the dual process of folds, namely \textit{anamorphisms}.
-It is potentially related to our work, as parsing is a representative kind of unfolds, whereas they only discussed the composition of a cv-coalgebra and a term homomorphism, which differs from modular parsing.
-\haoyuan{From Object Algebras to Attribute Grammars?}
+In OO languages, however, things get much better with the mechanism of
+inheritance. Furthermore, Object Algebras~\cite{Oliveira2012} were
+proposed as a design pattern for the extensibility of both data
+variants and operations in a modular type-safe way. The code is
+concise and expressive using Object Algebras, building a foundation
+for our new library with various dimensions of extensibility.
+Moreover, in~\cite{Oliveira2012} the authors have discussed the
+composition of algebras. In our parsing approach, a parser consumes an
+algebra, which is delegated to return the results, during its process
+of parsing. Having a set of algebras, it requires multiple parsing
+with several times of invocation, which leads to redundant work.
+Instead, algebras are supposed to be composed into one before the
+invocation of the parser. Bahr et al. lead a similar discussion
+in~\cite{Bahr2011}, where queries (or \textit{catamorphisms}) and
+transformations (or \textit{homomorphisms}) are composable. They have
+also mentioned the dual process of folds, namely
+\textit{anamorphisms}. It is potentially related to our work, as
+parsing is a representative kind of unfolds, whereas they only
+discussed the composition of a cv-coalgebra and a term homomorphism,
+which differs from modular parsing. \haoyuan{From Object Algebras to
+  Attribute Grammars?}
 
-Another interesting observation, as we have mentioned in~\ref{subsec:objectalgebras}, is that transformation algebras are frequently used in parsing, as concrete syntax may not always coincide with its abstract syntax, for instance, the bruijn indices in lambda calculus, as well as other syntactic sugars.
-In the Shy framework~\cite{Zhang2015}, the pattern of transformations is captured into a template for Java annotation processing, and most of boilerplate code is automatically generated, hence users only need to override a few interesting cases, conveniently.
-The idea of overriding existing cases was also proposed in MRM.
-It is potentially useful to our framework.
+Another interesting observation, as we have mentioned
+in~\ref{subsec:objectalgebras}, is that transformation algebras are
+frequently used in parsing, as concrete syntax may not always coincide
+with its abstract syntax, for instance, the bruijn indices in lambda
+calculus, as well as other syntactic sugars. In the Shy
+framework~\cite{Zhang2015}, the pattern of transformations is captured
+into a template for Java annotation processing, and most of
+boilerplate code is automatically generated, hence users only need to
+override a few interesting cases, conveniently. The idea of overriding
+existing cases was also proposed in MRM. It is potentially useful to
+our framework.\bruno{The discussion on Extensibility can be
+  considerably shorter. We just need to point out that there are many
+  techniques for extensibility (without discussing alot of details),
+  but most of that work does not address the problem of 
+\emph{producing extensible ASTs} (rather than consuming those. Then
+you can talk about Baht's work, and have the discussion on
+transformations. Generally speaking, go over your text, polish it, and
+trim it. You are a bit too ``chatty'' (that is you talk too much): go
+straight to the point.}

### 2016-12-18 paper/paper.tex
First comments on Section 2
Done 2.2, rewriting 2.3
work on the intro

@@
 \usepackage{theorem}
-%% \usepackage{mathpartir}
+\usepackage{syntax}
 \usepackage{bbm}
@@
 %\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
+\newcommand\inlinecode[1]{\lstinline[language=PlainCode]{#1}}
 \newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
@@
 \usepackage{xspace}
+\renewcommand{\paragraph}[1]{\vspace{2pt}\noindent {\bf #1}}
 
@@
   allow complete parsers to be built out of modular parsing
-  components, but also enable the parsing components to be \emph{modularly
-  type-checked} and \emph{separately compiled}. We present \name: a parsing
-  combinator library that enables modular parsing. The library is
-  built on top of a Packrat parsing library, but adds new parsing
+  components, but also enable the parsing components to be
+  \emph{modularly type-checked} and \emph{separately compiled}. We
+  present \name: a parsing combinator library that enables modular
+  parsing. Interestingly, the modularity requirements for modular
+  parsing rule out several existing parsing combinator approaches,
+  which rely on some non-modular techniques. Fortunatelly Packrat
+  parsing techniques, provide solutions for such modularity problems,
+  and enable reasonable performance in a modular setting.  Thus \name
+  is built on top of a Packrat parsing library, but adds new parsing
   combinators to enable modular parsing. The new parsing combinators
   employ \emph{delegation-based} techniques and \emph{Object Algebras}
-  to support extensibility. The choice of Packrat parsing over other
-  parsing techniques turns out to be important for achieving
-  performance in a modular setting. To evaluate \name 
-  we conduct a  case study based on the 
-  ``Types and Programming Languages'' interpreters. The case study 
-  shows that \name is effective at reusing parsing code from existing
-  interpreters, and the total parsing code is 60\% shorter than the
-  non-modular parsing code.
+  to support extensibility. To evaluate \name we conduct a case study
+  based on the ``Types and Programming Languages'' interpreters. The
+  case study shows that \name is effective at reusing parsing code
+  from existing interpreters, and the total parsing code is 60\%
+  shorter than the non-modular parsing code.
 

### 2016-12-18 paper/sections/Introduction.tex
work on the intro

@@
 degree of modularity, but they have limitations as illustrated by
-well-known problems such as the Expression Problem~\cite{}. To address
-the limitations of Programming Languages several different approaches
-have been proposed in the past. Existing approaches can be broadly
-divided into two categories: \emph{syntactic} or \emph{semantic} 
-modularization techniques. Syntactic modularization techniques enable 
-the separate development of software components or features that can
-later be combined into complete programs. However, very often, the 
-composition of components in syntactic modularization techniques 
-amounts to a form of textual composition of the code from multiple 
-components/features. Semantic modularization
-techniques go one step further, and also enable those components or
-features to be modularly type-checked and separately compiled. This
-enables the composition of compiled binaries, and also ensures the
-type-safety of the composed code.
+well-known problems such as the Expression Problem~\cite{}. The
+Expression Problem refers to the difficulty of writing data
+abstractions that can be easily extended with both new operations and
+new data variants. Traditionally the kinds of data abstraction found
+in functional languages can be extended with new operations, but
+adding new data variants is difficult. The traditional object-oriented
+approach to data abstraction facilitates adding new data variants
+(classes), while adding new operations is more difficult.
 
-Many approaches use syntactic modularization techniques. Examples
-include tools for developing Software-Product Lines~\cite{}, some
-Language Workbenches~\cite{}, or extensible parser generators.  In
-those approaches textual composition techniques such as
-\emph{superimposition}~\cite{} enable the development . Syntactic
-modularization techniques have also been applied to the problem of
-\emph{extensible parsing}. There are several approaches that enable
-the development of \emph{syntactically} modular parsers or
-grammars. However these approaches do not support separate compilation
-or modular type-checking that we strive for in this work.
+A reason why a solution to the Expression Problem is important in
+practice is that it is necessary for the development of
+\emph{Software-Product Lines} (SPLs)~\cite{}. A software-product line
+is a reusable set of components, which can be combined in multiple ways
+to obtain different programs. Programming languages offer a concrete
+example for SPLs. A SPL for programming languages would allow us to
+model various typical operations of programming languages (such as
+evaluation, compilation, or parsing) for various different language
+constructs (such as binding, arithmetic, conditional or loops)
+independently and separately. For example, evaluation components could be defined
+independently for binding and arithmetic constructs. If the language
+to be implemented is the pure lambda calculus, only evaluation of
+binding constructs is necessary. However, more realistic programming
+languages will include arithmetic constructs, and will require 
+evaluation for such constructs as well. In this case 
+both the component for evaluation of binders and arithmetic 
+expressions can be combined to implement the desired functionality.
 
 \begin{comment}
- such as AHEAD or 
-FeatureHouse, which enable the development of highly reusable 
-software components that can be combined to form complete 
-software products.
+Most programming languages share alot of features in
+common.
+
+For example, most languages have language constructs for:
+binding (such as variables, functions, and function applications);
+basic arithmetic operations; basic logic and conditional operations;
+loops; as well as various other features. For each language construct,
+various operations (such as evaluation, compilation, or parsing) need
+to be implemented. It is reasonable to wonder whether we can simply
+implement those features independently of a particular implementation
+of a programming language. Evaluation could be defined independently 
+for binding and arithmetic constructs. If the language to be
+implemented is the pure lambda calculus, only evaluation of binding 
+constructs is necessary. Thus only the component that implements 
+evaluation for binding needs to be used in such an implementation.
+However, more realistic programming languages 
+will include arithmetic constructs, and will require an evaluation
+function for those.
+
+Then it would be possible to \emph{reuse}
+some of those features in \emph{multiple} different implementations of
+programming languages. Essentially, this would enable a SPL for
+programming languages, where all
+
+A solution to the Expression Problem could ena
+
+A concrete 
+example that illustrates this issue is 
 \end{comment}
 
-There are also several approaches to semantic modularity. For example,
-family polymorphism enables the development of highly reusable
-components and can even be used to develop Software-Product Lines. 
-... However, as far as we are aware, the problem of modular parsing 
-has not been studied in semantic modularization
-approaches.
+To address the modularity limitations of Programming Languages several
+different approaches have been proposed in the past. Existing
+approaches can be broadly divided into two categories:
+\emph{syntactic} or \emph{semantic} modularization
+techniques. Syntactic modularization techniques are quite popular in
+practice. Examples include many tools for developing Software-Product
+Lines~\cite{}, some Language Workbenches~\cite{}, or extensible parser
+generators~\cite{}.  Most syntactic approaches employ textual
+composition techniques such as \emph{superimposition}~\cite{} to
+enable the development modular program features. Such textual
+composition techniques collect the code for multiple features and
+merge it together when a concrete combination of features is needed
+for a particular program. The typical drawback of such techniques is
+the lack of modular type-checking and separate compilation:
+``\emph{most feature-oriented implementation mechanisms lack proper
+  interfaces and support neither modular type checking nor separate
+  compilation}''~\cite{kastner11road}\bruno{reference to ``The road to
+  Feature Modularity''}. Syntactic modularization techniques have also
+been applied to the problem of \emph{extensible parsing}. There are
+several approaches~\cite{} that enable the development of
+\emph{syntactically} modular parsers or grammars. However these
+approaches do not support separate compilation or modular
+type-checking either.\bruno{mention that a reason why syntactic 
+modularization techniques are popular is simplicity (in implementation
+and also use).}
 
-  This paper investigates solutions for the problem of \emph{modular
-    parsing}. We focus on \emph{semantic} modularity and not just
-  \emph{syntactic} modularity. That is, the solutions should not only
+Semantic modularization techniques move away
+from syntactic composition techniques that rely on the textual source
+code. This allows us to go one step further in terms of modularity,
+and also enable components or features to be modularly type-checked
+and separately compiled. Modular type-checking and separate
+compilation are desirable properties to have from a Software
+Engineering point-of-view, and enable the composition of compiled
+binaries as well as ensuring the type-safety of the code composed of
+multiple components. Examples of semantic modularization techniques 
+include various approaches to \emph{family polymorphism}~\cite{}, as 
+well as several techniques for solving the Expression Problem. 
+\bruno{strenghten? mention dynamic languages (separate compilation 
+but no modular type-checking), and techniques that weaken type-safety 
+requirements.} Most semantic
+modularization techniques have focused on operations that traverse or 
+or process extensible datastructures, such as ASTs.
+\bruno{Object Algebras here?} However, as far as
+we know there is little work on operations that build/produce ASTs. 
+In particular the problem of modular parsing has not been studied in
+semantic modularization approaches. This is a shame because, to
+realise the vision of software product-lines for programming
+languages, modular parsing is a necessity.
+
+\paragraph{Modular and Extensible Parsing}
+  This paper investigates presents \name: a parsing
+  combinator library that enables modular parsing.
+  \name provides a solution for the problem of \emph{semantic modular
+    parsing}. That is, the solutions should not only
   allow complete parsers to be built out of modular parsing
   components, but also enable the parsing components to be \emph{modularly
-  type-checked} and \emph{separately compiled}. We present \name: a parsing
-  combinator library that enables modular parsing. The library is
+  type-checked} and \emph{separately compiled}. \name is
   built on top of a Packrat parsing library, but adds new parsing
@@
   parsing techniques turns out to be important for achieving
-  performance in a modular setting. To evaluate \name 
-  we conduct a  case study based on the 
-  ``Types and Programming Languages'' interpreters. The case study 
-  shows that \name is effective at reusing parsing code from existing
+  performance in a modular setting. \bruno{left recursion, and
+    backtracking removal.}
+
+  By analising the \emph{full} grammar it is possible to remove
+  backtracking, which would otherwise increase parsing times. Many
+  parsing combinator libraries routinely use backtracting elimination
+  to achieve performance. However, in a modular setting this technique
+  cannot be used, because the full grammar is not known. Thus we have
+  to be very conservative at eliminating backtracting. Unfortunatelly,
+  this has a severe impact on performance.
+
+  To evaluate \name we conduct a case study based on the ``Types and
+  Programming Languages'' (TAPL) interpreters.  The case study shows
+  that \name is effective at reusing parsing code from existing
   interpreters, and the total parsing code is 60\% shorter than the
-  non-modular parsing code.
+  non-modular parsing code.\bruno{comment on the efficiency}
+
+In summary our contributions are:
+
+\begin{itemize}
+
+\item {\name:} A parser combinator library that allows the development 
+of modular parser.
+
+\item {{\bf TAPL case study:}} We conduct a case study with 18 interpreters
+  from the TAPL book. The case study show the effectiveness of modular 
+  parsing in terms of reuse.
+
+\end{itemize}

### 2016-12-19 paper/sections/Introduction.tex
contributions

@@
 \item {\name:} A parser combinator library that allows the development 
-of modular parser.
+of modular parser. The library uses \emph{delegation} and
+\emph{Object-Algebras} to achieve modularity and extensibility.
+
+\item {{\bf A Parsing Technique for OO ASTs:}} A simplified version of
+  our technique also enables parsing OO-style ASTs, where new language
+  constructs can be easily added.
+
+\item {{\bf Limitations of existing Parser Combinator Techniques:}}
+\bruno{Improve and write something here.}
 

### 2016-12-20 paper/sections/Overview.tex
Rewrite, add references

@@
 various techniques that fail to satisfy our requirements, and argue
-that that Packrat parsing~\cite{} is a suitable candidate for
+that that Packrat parsing~\cite{Ford2002} is a suitable candidate for
 type-safe modular parsing.
@@
 \paragraph{Parser Generators} The most widely use tools for parsing
-are parser generators~\cite{}. Parser generators help users generate parsers automatically or
+are parser generators. Parser generators help users generate parsers automatically or
 semi-automatically from a given grammar. There is no restriction on
-the algorithm, while most of them adopts table-based LL~\cite{} and LR parsing
-algorithms~\cite{}.
+the algorithm, while most of them adopts table-based LL~\cite{lewis1968syntax} and LR~\cite{knuth1965translation} parsing
+algorithms.
 Although efficient, the main drawback of parser generators is that they do not support
@@
 
-Modular parsing based on parser generators is supported by many libraries [].
-Users can separate the concrete/abstract syntax and related parsing code into reusable components.
+Modular parsing based on parser generators is supported by many libraries~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.
+Users can separate the syntax definition and related parsing code into reusable components.
 Then the corresponding parsers are built by their library utilities.
-For example, NOA [] uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers.
+For example, NOA~\cite{Gouseti2014} uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers.
 However, such generation procedure requires a whole compilation after the collection of all grammar pieces.
@@
 
-\huang{a paragraph here, refer to related work section, discuss
-  previous works of modular parsing using parser generators and its
-  drawbacks}\bruno{Please finish this.}\huang{done}
-
 \paragraph{Parser Combinators}
-Comparing with the parser generators, a \textit{parser combinator}~\cite{}
+Comparing with the parser generators, a \textit{parser combinator}~\cite{burge1975,Wadler1985}
 takes several parsers and produce a new parser as its output. Parser combinators are
@@
 modular parsing. In particular several parser combinators,
-including the famous library Parsec~\cite{} library, require
+including the famous library Parsec~\cite{Leijen2001} library, require
 programmers to manually do \textit{left-recursion elimination}, and
@@
 branches, which results worst case exponential time complexity.
-\huang{done}\bruno{you should again illustrate your point here with an
-  example. The example should illustrate that the addition of a new
-  case may require introducing backtracking. }
 
@@
 Fortunately more advanced parsing techniques such as Packrat
-parsing~\cite{}, address the limitations of simple parser combinators
+parsing~\cite{Ford2002}, address the limitations of simple parser combinators
 such as Parsec. Packrat parsers use
@@
 position of the input, so that repeated computation is eliminated, and
-they support left-recursive grammars directly. All of these properties are very
+they support left-recursive grammars directly~\cite{warth2008}. All of these properties are very
 suitable for modularity, thus we use Packrat parsers as the underlying
@@
 Unfortunately, this extended parser does not work properly to parse all possible expressions defined by the new grammar.
-For example, it fails to parse \inlinecode{(\x.x)} \inlinecode{(\y.y)} which is obviously valid.
+For example, it fails to parse \inlinecode{(\\x.x)} \inlinecode{(\\y.y)} which is obviously valid.
 
@@
 
-That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{(\x.x)} \inlinecode{(\y.y)}.
+That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{(\\x.x)} \inlinecode{(\\y.y)}.
 To resolve this issue, we need to let \inlinecode{pApp} be aware of the change of grammar.
@@
 \paragraph{Client Code} Since the type of parsing functions are
-changed, we can not apply them directly. The client code below shows
-how to use the new parser. The \inlinecode{fix} function, which is the
+changed, we can not apply them directly. A \inlinecode{fix} function, which is the
 the standard \textit{fixpoint combinator}, is used to `close' the open
-recursion. \bruno{Too much detail is being given here about open
+recursion. After that, it can be used as an ordinary parser.
+The client code below demonstrates the usage of new parsers.
+
+\huang{done}\bruno{Too much detail is being given here about open
   recursion. In Section 2 we should talk only about \emph{how} to use
@@
 
-\lstinputlisting[linerange=68-75]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
+\lstinputlisting[linerange=68-72]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
 
@@
 
-\lstinputlisting[linerange=80-86]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
+\lstinputlisting[linerange=77-83]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
 
@@
 
-\lstinputlisting[linerange=95-95]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
+\lstinputlisting[linerange=89-89]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
 
@@
 hierarchy makes the addition of new operations over the AST
-problematic. Object Algebras~\cite{} enable us to solve this problem, and
+problematic. Object Algebras~\cite{Oliveira2012} enable us to solve this problem, and
 offer high flexibility in the choice of operations to be performed
@@
 
-To use the parser, we must provide an algebra instance as the operation to construct the parsing result.
-As shown in the code below, we feed the pretty printing operation \inlinecode{ExprPrint} to the generic \inlinecode{parse} function.
-It will print \inlinecode{(x y)} as the result.
-
-\lstinputlisting[linerange=38-43]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
-
 \paragraph{Extensibility of Syntax} Following previous examples, the code below shows how to extend the language to support lambda abstractions.
 
-\lstinputlisting[linerange=48-62]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
+\lstinputlisting[linerange=32-46]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
 
@@
 
+To use the parser, we must provide an algebra instance as the operation to construct the parsing results.
+In the code below, we use the pretty printing operation \inlinecode{LambdaPrint} as the algebra instance, so that the parsing result is a pretty printed AST.
+
+\lstinputlisting[linerange=55-60]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
+
 \paragraph{Extensibility of Operations} With Object Algebras, operations over ASTs can also be extended in a modular way.
 Here is an example of collecting free variables from an expression.
-We can use this new operation during parsing.
+We can use this operation during parsing and obtain a set of free variables.
 
-\lstinputlisting[linerange=75-81]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
+\lstinputlisting[linerange=74-80]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
 
@@
 
-\lstinputlisting[linerange=66-70]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=65-69]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX

### 2016-12-20 Scala/Parser/src/PaperCode/Overview/OA.scala

@@ -28,22 +28,6 @@ trait ExprParser[E] {
 }
 //END_OVERVIEW_OA_PARSER
 
-
-  def parse[E](p: Open[PackratParser[E]])(inp: String): E = {
-    val t = phrase(fix(p))(new lexical.Scanner(inp))
-    if (t.successful) t.get else scala.sys.error(t.toString)
-  }
-/*
-//BEGIN_OVERVIEW_OA_USE
-def parseWithAlg[E](inp: String)(alg: ExprAlg[E]): E = {
-  val p = new ExprParser[E] {}.pExpr(alg)
-  parse[E](p)(inp)
-}
-
-println(parseWithAlg("x y")(new ExprPrint{}))
-//END_OVERVIEW_OA_USE
-*/
-
 //BEGIN_OVERVIEW_OA_EXT
 trait LambdaAlg[E] extends ExprAlg[E] {
   def Lam(x: String, e: E): E
@@ -62,6 +46,21 @@ trait LambdaParser[E] extends ExprParser[E] {
 }
 //END_OVERVIEW_OA_EXT
 
+  def openParse[E](p: Open[PackratParser[E]])(inp: String): E = {
+    val t = phrase(fix(p))(new lexical.Scanner(inp))
+    if (t.successful) t.get else scala.sys.error(t.toString)
+  }
+
+//BEGIN_OVERVIEW_OA_USE
+def parseLambdaAlg[E](inp: String)(alg: LambdaAlg[E]): E = {
+  val p = new LambdaParser[E] {}.pExtExpr(alg)
+  openParse[E](p)(inp)
+}
+
+parseLambdaAlg("(\\x.x) (\\y.y)")(new LambdaPrint{})
+//END_OVERVIEW_OA_USE
+
+
 //BEGIN_OVERVIEW_OA_MULTI_SYNTAX
 trait TypedLambdaAlg[E, T] extends ExprAlg[E] {
   def Lam(x: String, t: T, e: E): E
@@ -70,7 +69,7 @@ trait TypedLambdaAlg[E, T] extends ExprAlg[E] {
 }
 //END_OVERVIEW_OA_MULTI_SYNTAX
 
-/*
+
 //BEGIN_OVERVIEW_OA_EXT_OP
 trait CollectFreeVars extends LambdaAlg[Set[String]] {
   def Var(x: String): Set[String] = Set(x)
[…]

### 2016-12-20 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -26,7 +26,7 @@ class Lam(x: String, e: Expr) extends Expr
 trait ExtParser extends BaseParser {
   lexical.delimiters += ("\\", ".")
   val pExtExpr: PackratParser[Expr] =
-    pExpr ||| pLam ||| "(" ~> pExtExpr <~ ")"
+    pVar ||| pApp ||| pLam ||| "(" ~> pExtExpr <~ ")"
   val pLam: PackratParser[Expr] =
     "\\" ~> ident ~ ("." ~> pExtExpr) ^^ { case x ~ e => new Lam(x, e) }
 }
@@ -37,8 +37,8 @@ trait ExtParser extends BaseParser {
     //runParser(parser)("a b c")
 
     val extParser = new ExtParser {}.pExtExpr
-    runParser(extParser)("a b c")
-    runParser(extParser)("(\\x.x) (\\y.y)")
+    parse(extParser)("a b c")
+    parse(extParser)("(\\x.x) (\\y.y)")
   }
 }
 
@@ -63,18 +63,15 @@ trait BaseParser {
 }
 //END_OVERVIEW_OPEN_BASE
 
-  def use(): Unit = {
+/*
 //BEGIN_OVERVIEW_OPEN_USE
-def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
+def parse[E](p: PackratParser[E]): String => E = {...}
+def fix[T](f: Open[T]): T = {...}
+def openParse[E](p: Open[PackratParser[E]]): String => E = parse[E](fix(p))
 
-def parse[E](p: Open[PackratParser[E]])(inp: String): E = {
-  val t = phrase(fix(p))(new lexical.Scanner(inp))
-  if (t.successful) t.get else scala.sys.error(t.toString)
-}
-
-println(parse(new BaseParser {}.pExpr)("x y"))
+openParse(new BaseParser {}.pExpr)("x y")
 //END_OVERVIEW_OPEN_USE
-  }
+*/
 
 //BEGIN_OVERVIEW_OPEN_EXT
 trait ExtParser extends BaseParser {
@@ -87,13 +84,14 @@ trait ExtParser extends BaseParser {
 //END_OVERVIEW_OPEN_EXT
 
   def use2(): Unit = {
-    def parse[E](p: Open[PackratParser[E]])(inp: String): E = {
-      val t = phrase(fix(p))(new lexical.Scanner(inp))
[…]

### 2016-12-20 Scala/Parser/src/PaperCode/Overview/Util.scala

@@ -14,8 +14,8 @@ object Util extends StandardTokenParsers with PackratParsers {
 
   def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
 
-  def runParser(p: Parser[_]): String => Unit = in => {
+  def parse[E](p: Parser[E]): String => E = in => {
     val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) println(t.get) else scala.sys.error(t.toString)
+    if (t.successful) t.get else scala.sys.error(t.toString)
   }
 }

### 2016-12-20 paper/sections/Overview.tex
Summary, rewrite one example

@@
 
-\bruno{Write a summary for this section}
-\bruno{Fill in all the missing references}
+This section gives an overview of our library \name, from a user's perspective.
+Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions.
+We start from section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers.
+Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse.
+Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls.
+In section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers.
+Finally, section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.
+
+\huang{done} \bruno{Write a summary for this section}
+
+\huang{done}\bruno{Fill in all the missing references}
 
@@
 
-\huang{done, the example may be not very good, but i haven't found a
-  better one}
+\huang{todo}
 \bruno{This example does not appear to be problematic: adding ``-''
@@
 \paragraph{Backtracking} The need for backtracking is also problematic
-in a modular setting. Take Parsec as an example, its choice
-combinator only tries the second alternative if the first fails
-without any token consumption. If two alternatives have a same
-non-empty prefix, an auxiliary function \inlinecode{try} must be added
-to backtrack. However, if we want to extend our parsers later, we
-should always consider the worst case in which all alternatives share
-common prefixes. Then we need to add \inlinecode{try} for all the
-branches, which results worst case exponential time complexity.
-
+in a modular setting.
 For example, suppose that we only have \inlinecode{import..from} statements before, now we want to extend the grammar to add an \inlinecode{import..as} case.
@@
 
-Because the new case \inlinecode{import..as} shares a prefix with the old case \inlinecode{import..from}, we must backtrack in order to cover all possibilities.
+Because the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning.
+Take Parsec as an example, its choice
+combinator \inlinecode{<|>} only tries the second alternative if the first fails
+without any token consumption. An auxiliary function \inlinecode{try} is used for explicit backtracking.
 
 \begin{lstlisting}[language=PlainCode]
-oldParser = parseImportFrom <|> ...
-newParser = try parseImportFrom <|> ... <|> parseImportAs
+oldParser = parseImpFrom <|> parseA <|> parseB <|> ...
+newParser = try parseImpFrom <|> parseA <|> parseB <|> ... <|> parseImpAs
 \end{lstlisting}
 
-\bruno{The example is good, I think but the explanation is not.
+Given the full grammar, we can decide where to put \inlinecode{try} for backtracking.
+However, in a modular setting we should always consider the worst case, that all alternatives may share common prefixes with future cases.
+Then we need to backtrack for all the branches, which results worst case exponential time complexity.
+In Parsec, we must add \inlinecode{try} everywhere:
+
+\begin{lstlisting}[language=PlainCode]
+parser = try parseImpFrom <|> try parseA <|> try parseB <|> ... <|> try parseImpAs
+\end{lstlisting}
+
+\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
 You want to say that when you know
@@
 
-\subsection{The Parsing Expression Problem}
+\subsection{The Parsing Expression Problem}\label{subsec:overview-problem}
 \huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
@@
 
-\subsection{Delegation-Based Parsing}\label{subsec:overview-firstlook}
+\subsection{Delegation-Based Parsing}\label{subsec:overview-delegation}
 

### 2016-12-20 paper/paper.tex
contributions
Some comments on S4

@@
   parsing rule out several existing parsing combinator approaches,
-  which rely on some non-modular techniques. Fortunatelly Packrat
+  which rely on some non-modular techniques. We show that Packrat
   parsing techniques, provide solutions for such modularity problems,
-  and enable reasonable performance in a modular setting.  Thus \name
-  is built on top of a Packrat parsing library, but adds new parsing
+  and enable reasonable performance in a modular setting.  \name
+  builts on top of a Packrat parsing library, but adds new parsing
   combinators to enable modular parsing. The new parsing combinators

### 2016-12-20 paper/sections/OpenRecursion.tex
Some comments on S4

@@
 but using delegation instead of class-based inheritance.
+This programming style enables parsing OO-style ASTs, which 
+can be naturally extended with new language constructs.
 
@@
 called \lstinline{self} or, in some languages,
-\lstinline[keywords={}]{this}'' (by Ralf Hinze). The interesting thing
+\lstinline[keywords={}]{this}'' (by Ralf Hinze)\bruno{this is not how we cite references!}. The interesting thing
 is that such a variable \lstinline{self} is late-bound, or open to the
@@
 
-For sure we cannot always leave it open; upon using it we have to
-close the recursion, at which time we need to take a fix-point. Here
-one might suddenly be reminded of the \lstinline{fix} function in some
-functional languages like Haskell. In Haskell \lstinline{fix} is
-defined as:
+\paragraph{Example of Open Recursion} To illustrate the use and
+usefulness of open recursion, we first consider a standard recursive 
+implementation of the Fibonacci function:
 
-\begin{lstlisting}[language=Haskell,keywords={}]
-fix f = let x = f x in x
-\end{lstlisting}
-In Scala, however, there is no library \lstinline{fix} function, but it can be easily implemented since Scala also supports lazy evaluation. See below:
+\lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
+
+\noindent An alternative implementation using open recursion is:
+
+\lstinputlisting[linerange=44-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
+\bruno{Fix should be replaces by Open. Please coordinate with Huang to
+ensure that your code and his code are coherent (i.e you use same
+names and same conventions),}
+\noindent This alternative implementation has an extra argument
+\lstinline{self}, which serves as an explicit self-reference. In the 
+definition of the function, direct recursive calls are replaced by 
+calls to the self reference.
+
+\paragraph{Closing the Open Recursion} An implementation of
+ standard the standard Fibonacci function is recovered by closing 
+the open recursion using a \emph{lazy fixpoint combinator}:
+\bruno{definition of fib with fix here.}
+
+In Scala the lazy fixpoint combinator is straightforwardly implemented~\cite{} as follows:
 
 \lstinputlisting[linerange=27-30]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
+\bruno{can be defined in a single line.}
+\noindent Note that the function \lstinline{f} has a by-name parameter 
+of type \lstinline{"=>} \lstinline{A"}. This is necessary to ensure
+termination of well-founded recursive calls. In such cases, 
+there are some base cases (like the first two in
+\lstinline{fib2}) which stop the recursion in their branches, and the 
+evaluation is finally reduced to those cases.
 
-where function \lstinline{f} has a by-name parameter of type \lstinline{"=>} \lstinline{A"}.
-Here we illustrate by the famous Fibonacci example. Traditionally it can be implemented in Scala using pattern-matching:
-\lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
-
-An alternative implementation using open recursion is as follows:
-\lstinputlisting[linerange=44-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
-where \lstinline{self} is the explicit self-reference, namely the delegator. Functions like \lstinline{fib2} which takes
-an explicit self-reference as its parameter are called \textit{generators} (by Ralf Hinze).
+\begin{comment} 
 If we close the recursion on \lstinline{fib2} only, the evaluation of \lstinline{fix(fib2)(2)}, namely the above \lstinline{x2} is processed as follows:
@@
 \lstinline{fib2}) which stop the recursion in their branches, and the evaluation is finally reduced to those cases.
+\end{comment}
 
+\bruno{We may want to use a better example than what is next.}
 Suppose now we want to make use of open recursion for additional operations. We first define a combinator which combines the results of two generators in a pair:
@@
 \subsection{Parsing with Open Recursion}\label{subsec:parsingwithopen}
-Open recursion potentially opens the gate of extensibility for parsing, as a parser can now be open to additional features in the future.
+
+Open recursion potentially opens the gate of extensibility for
+parsing, as a parser can now be open to additional features in the future.
+
+\bruno{Most of the text below overlaps with what Huang was already
+  written. I think you and Huang need to sit down, look at the text 
+both of you wrote to motivate the problem, and rewrite that in the
+Overview Section.}
+
+\bruno{BEGIN: Text to be deleted/moved/rewritten in S2}
 Our exploration starts from a simple example, where a single parser only parses a free variable (i.e., a string). Traditionally, to write
@@
 \lstinputlisting[linerange=44-49]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PLIT
+\bruno{END: Text to be deleted/moved/rewritten in S2}
 
-We have observed that \lstinline{pApp} is indeed constructed in a non-modular way, because the \lstinline{p} standing for recursive parser needs to be updated on every extension.
-Actually in the grammar, the recursive \lstinline{"e"} in the second case should not only include all the existing cases like \lstinline{pApp} and \lstinline{pVar}, but also take future extensions into account (we can have arbitrarily more cases).
-This is where we are inspired to use open recursion.
-To define \lstinline{pApp} as an ``open'' function, we add a parameter \lstinline{p} to it, representing the explicit self-reference of our whole parser, open to future extensions.
-Note that \lstinline{p} should be defined as a by-name parameter using \lstinline{"=>"}.
+\bruno{Now, here you can refer back to the example which is used in S2
+to illustrate the problem. But your \emph{main goal} is to talk about
+the implementation of the combinators that are helpful for Open
+Recursion and parsing OO ASTs. So you need some major rewrite of the 
+text that comes next. First, building on open recursion, explain the
+implementation of the new parsing combinators. Then, finish of the
+section with a good example of how to parse OO ASTs, perhaps
+extending the example discussed in S2 a bit more.}
+
+We have observed that \lstinline{pApp} is indeed constructed in a
+non-modular way, because the \lstinline{p} standing for recursive
+parser needs to be updated on every extension. Actually in the
+grammar, the recursive \lstinline{"e"} in the second case should not
+only include all the existing cases like \lstinline{pApp} and
+\lstinline{pVar}, but also take future extensions into account (we can
+have arbitrarily more cases). This is where we are inspired to use
+open recursion. To define \lstinline{pApp} as an ``open'' function, we
+add a parameter \lstinline{p} to it, representing the explicit
+self-reference of our whole parser, open to future extensions. Note
+that \lstinline{p} should be defined as a by-name parameter using
+\lstinline{"=>"}.
 
@@
 We proceed to implement those parsers. An issue is that we need both a \lstinline{Parser[Expr]} and a \lstinline{Parser[Type]} in the fixpoint.
-Hence we use a pair in the parameter:
+Hence we use a pair in the parameter:\bruno{I don't think the approach
+below, adding pairs, is very modular: it requires changing your }
 

### 2016-12-21 paper/paper.tex
Update the tex template. Minor comments on Overview.

@@
 %\documentclass[preprint]{llncs}
-\documentclass[a4paper,UKenglish]{lipics}
+\documentclass[a4paper,UKenglish]{lipics-v2016}
 
@@
 
+%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
 \author[1]{John Q. Open}
 \author[2]{Joan R. Access}
-\affil[1]{Dummy University Computing Laboratory\\
-  Address, Country\\
-  \texttt{open@dummyuni.org}}
-\affil[2]{Department of Informatics, Dummy College\\
-  Address, Country\\
+\affil[1]{Dummy University Computing Laboratory, Address/City, Country\\
+  \texttt{open@dummyuniversity.org}}
+\affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
   \texttt{access@dummycollege.org}}
-\authorrunning{J.\,Q. Open and J.\,R. Access}
-% mandatory. First: Use abbreviated first/middle names. Second (only in severe
-% cases): Use first author plus 'et. al.'
+\authorrunning{J.\,Q. Open and J.\,R. Access} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
 
-\Copyright{John Q. Open and Joan R. Access}
-% mandatory, please use full first names. LIPIcs license is "CC-BY";
-% http://creativecommons.org/licenses/by/3.0/
+\Copyright{John Q. Open and Joan R. Access}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 
-\subjclass{Dummy classification -- please refer to
-  \url{http://www.acm.org/about/class/ccs98-html}}
-% mandatory: Please choose ACM 1998 classifications from
-% http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of
-% Computation".
+\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
+E.g., cite as "F.1.1 Models of Computation".
 \keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
@@
 %Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\serieslogo{}%please provide filename (without suffix)
-\volumeinfo%(easychair interface)
-  {Billy Editor and Bill Editors}% editors
-  {2}% number of editors: 1, 2, ....
-  {Conference title on which this volume is based on}% event
-  {1}% volume
-  {1}% issue
-  {1}% starting page number
-\EventShortName{}
-\DOI{10.4230/LIPIcs.xxx.yyy.p}% to be completed by the volume editor
+\EventEditors{John Q. Open and Joan R. Acces}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
 % Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   parsing techniques, provide solutions for such modularity problems,
-  and enable reasonable performance in a modular setting.  \name
-  builts on top of a Packrat parsing library, but adds new parsing
+  and enable reasonable performance in a modular setting.  Thus \name
+  is built on top of a Packrat parsing library, but adds new parsing
   combinators to enable modular parsing. The new parsing combinators

### 2016-12-21 paper/sections/Overview.tex
Update the tex template. Minor comments on Overview.

@@
 Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions.
-We start from section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers.
+We start from Section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers.
 Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse.
 Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls.
-In section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers.
-Finally, section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.
+In Section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers.
+Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.
 
@@
 suitable for modularity, thus we use Packrat parsers as the underlying
-parsing technique. In section \ref{sec:packratparsers}, we will give
+parsing technique. In Section \ref{sec:packratparsers}, we will give
 an overview of the Packrat parsing library implemented in Scala,
@@
 
-Typical parsers defined parser combinators follow the grammar
-structure, and build a corresponding AST. If the grammar is changed to
-add a new alternative, the corresponding parser has to be changed as
+Typical parsers defined by parser combinators follow the grammar
+structure, they parse an input to build a corresponding AST. If the grammar is extended
+with a new alternative, the corresponding parser has to be changed as
 well. Thus the traditional style of developing parsers is not
@@
 
-It is straightforward to model ASTs by inheritance and write corresponding parsers for each case.
+It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases. Note that we have imported
+the Scala Packrat Parsing library in code below, and use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} throughtout the paper.
 \huang{This is the first time to demonstrate code in this paper.
@@
 
-As the combination of \inlinecode{pVar} and \inlinecode{pApp}, \inlinecode{pExpr} can parse expressions correctly.
+\haoyuan{trait Expr with pretty printing? Shall we add client code here?}
+
+\noindent As the combination of \inlinecode{pVar}, \inlinecode{pApp} and \inlinecode{pParen}, \inlinecode{pExpr} can parse expressions like \lstinline{"x (y z)"} correctly.
 
 \paragraph{Attempt to Extend the Parser} Now consider we are extending the abstract syntax as well as the parsers.
-We add lambda abstractions as a new case of expressions.
-The language then becomes the famous \textit{untyped lambda calculus}.
+Hence we introduce lambda abstractions as a new case, which refers to the famous \textit{untyped lambda calculus}.
 
@@
 
-The ASTs can be extended trivially as below.
+The corresponding AST is extended as below.
 
@@
 
-\lstinputlisting[linerange=26-32]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXT
+\lstinputlisting[linerange=26-31]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXT
 
@@
 It worked fine at that time because \inlinecode{pExpr} covers all cases of the old grammar.
-However, after the extension of lambda abstractions, the proper function for such recursive calls should be \inlinecode{pExtExpr} instead of \inlinecode{pExpr}, because the new function \inlinecode{pExtExpr} covers all of the three cases in the new grammar, including the lambda case.
+However, after the extension of lambda abstractions, the proper function for such recursive calls should be \inlinecode{pExtExpr} instead of \inlinecode{pExpr}, because the new function \inlinecode{pExtExpr} includes the lambda case in the new grammar.
+\lstinline{pParen} has the similar problem.
 
 That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{(\\x.x)} \inlinecode{(\\y.y)}.
-To resolve this issue, we need to let \inlinecode{pApp} be aware of the change of grammar.
+To resolve this issue, \inlinecode{pApp} and \inlinecode{pParen} need to be aware of the change of grammar.
 We have some direct solutions:
@@
 \begin{itemize}
-    \item Keep the old code, rewrite problematic functions like \inlinecode{pApp}. \\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
+    \item Keep the old code, rewrite non-extensible functions like \inlinecode{pApp} and \inlinecode{pParen}.
+\\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
     \item Modify the old code, correct the recursive calls by proper functions.
-\\We can avoid code duplication thus get code reuse, but separate compilation is lost because of modification on the old code.
+\\We can avoid code duplication so as to get code reuse, but separate compilation is sacrificed because of modification on the old code.
     \item ??? \huang{third one? use dynamic languages?}
@@
 
-Similar to the trade-off in the Expression Problem, neither of them satisfies us.
+Because of the trade-off, neither of them are satisfying.
 The challenge of building semantically modular parsers requires fundamental change of the parser structure.
@@
 
-Hard-coded recursive calls prevent our parsers from extensibility.
+Hard-coded recursive calls prevent our parsers from extensions.
 This problem can be solved using delegation, which can be encoded with
@@
 recursive call, is passed to the parsers. This technique enables the
-developement of parsers that support OO-style extensibility of ASTs.
+development of parsers that support OO-style extensibility of ASTs.
 
@@
 
-\lstinputlisting[linerange=53-63]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_BASE
+\lstinputlisting[linerange=52-62]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_BASE
 
 According to the definition of type synonym \inlinecode{Open[T]},
-every parse function takes a parser of \inlinecode{Expr} as a
+every parsing function takes a parser of \inlinecode{Expr} as a
 parameter, namely \inlinecode{self}. This parameter \inlinecode{self}
@@
 
-\paragraph{Client Code} Since the type of parsing functions are
+\paragraph{Client Code} Since the type of parsing functions is
 changed, we can not apply them directly. A \inlinecode{fix} function, which is the
@@
 
-\lstinputlisting[linerange=68-72]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
+\lstinputlisting[linerange=67-71]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
 
@@
 
-\paragraph{Extensibility} As same as what we did in previous section, we extend the grammar by adding lambda abstractions as a new case.
-Similarly, the new parser extends from the old one by adding \inlinecode{pLam} on it.
+\paragraph{Extensibility} Similarly, we extend the grammar by adding lambda abstractions as a new case.
+The new parser also extends the old one by adding \inlinecode{pLam} on it.
 Notice that the old code is not touched, hence separate compilation is obtained.
 
-\lstinputlisting[linerange=77-83]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
+\lstinputlisting[linerange=76-82]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
+
+\haoyuan{Code without combinators?}
 
@@
 
-\lstinputlisting[linerange=89-89]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
+\lstinputlisting[linerange=88-88]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
 
@@
 It enables type-safe code reuse and separate compilation in our parsers.
-We will discuss this topic further in section \ref{sec:openandparsing}.
+We will discuss this topic further in Section \ref{sec:openandparsing}.
 

### 2016-12-21 paper/sections/RelatedWork.tex
Update the tex template. Minor comments on Overview.

@@
 
-\paragraph*{Syntactically Extensible Parsing} Extensible parsing is achieved in many
+\paragraph{Syntactically Extensible Parsing} Extensible parsing is achieved in many
 different areas, of which parser generators are a mainstream area
@@
 unambiguity of grammars, but separate compilation and modular
-type-checking remain unsolved, together with the issue of performance.
-\bruno{what performance issue?}
+type-checking remain unsolved, together with their affects on the performance.
+\bruno{what performance issue?}\haoyuan{Revised}
 
@@
 Stratego~\cite{Visser2001}, which focused little on separate
-compilation\bruno{focused little or did not have it at all?}. 
+compilation\bruno{focused little or did not have it at all?}.\haoyuan{The authors have
+mentioned separate compilation as a possible yet unimportant feature; they decided to discard it.}
 Extensible compilers like JastAdd~\cite{Ekman2007} and
@@
 separate compilation. Specifically, \cite{Bravenboer2009} introduces
-\textit{parse table composition}, in which paper grammars\bruno{what
-  is a paper grammar?} are compiled
+\textit{parse table composition}, where grammars\bruno{what
+  is a paper grammar?}\haoyuan{Revised} are compiled
 to the generation of parse tables\bruno{I don't understand the last
@@
 expensive\bruno{expensive in terms of what? why is it expensive?}.
+\haoyuan{in terms of performance. The sentence is copied from original paper.}
 Furthermore, \name supports \emph{both} separate compilation
@@
 
-\paragraph*{Parser Combinators} Since~\cite{burge1975} firstly
+\paragraph{Parser Combinators} Since~\cite{burge1975} firstly
 introduced parser generators, and after Wadler discussed more details
@@
 
-\paragraph*{Extensibility} There has been alot of work on building
+\paragraph{Extensibility} There has been alot of work on building
 modular programs, in both functional and object-oriented programming .

### 2016-12-21 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -11,10 +11,10 @@ class App(e1: Expr, e2: Expr) extends Expr
 
 trait BaseParser {
   lexical.delimiters += ("(", ")")
-  val pExpr: PackratParser[Expr] = pVar ||| pApp ||| "(" ~> pExpr <~ ")"
-  val pVar: PackratParser[Expr] = ident ^^ { x => new Var(x) }
-  val pApp: PackratParser[Expr] =
-    pExpr ~ pExpr ^^ { case e1 ~ e2 => new App(e1, e2) }
+  val pExpr: Parser[Expr] = pVar ||| pApp ||| pParen
+  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
+  val pApp: Parser[Expr] = pExpr ~ pExpr ^^ { case e1 ~ e2 => new App(e1, e2) }
+  val pParen: Parser[Expr] = "(" ~> pExpr <~ ")"
 }
 //END_OVERVIEW_SIMPLE_EXPR
 
@@ -25,9 +25,8 @@ class Lam(x: String, e: Expr) extends Expr
 //BEGIN_OVERVIEW_SIMPLE_EXT
 trait ExtParser extends BaseParser {
   lexical.delimiters += ("\\", ".")
-  val pExtExpr: PackratParser[Expr] =
-    pVar ||| pApp ||| pLam ||| "(" ~> pExtExpr <~ ")"
-  val pLam: PackratParser[Expr] =
+  val pExtExpr: Parser[Expr] = pExpr ||| pLam
+  val pLam: Parser[Expr] =
     "\\" ~> ident ~ ("." ~> pExtExpr) ^^ { case x ~ e => new Lam(x, e) }
 }
 //END_OVERVIEW_SIMPLE_EXT
@@ -54,20 +53,20 @@ type Open[T] = (=> T) => T
 
 trait BaseParser {
   lexical.delimiters += ("(", ")")
-  val pVar: Open[PackratParser[Expr]] =
+  val pVar: Open[Parser[Expr]] =
     self => ident ^^ { x => new Var(x) }
-  val pApp: Open[PackratParser[Expr]] =
+  val pApp: Open[Parser[Expr]] =
     self => self ~ self ^^ { case e1 ~ e2 => new App(e1, e2) }
-  val pExpr: Open[PackratParser[Expr]] =
+  val pExpr: Open[Parser[Expr]] =
[…]

### 2016-12-22 paper/sections/OpenRecursion.tex
Update the tex template. Minor comments on Overview.
Fix

@@
 called \lstinline{self} or, in some languages,
-\lstinline[keywords={}]{this}'' (by Ralf Hinze)\bruno{this is not how we cite references!}. The interesting thing
+\lstinline[keywords={}]{this}'' (by Ralf Hinze)\bruno{this is not how we cite references!}
+\haoyuan{It's a talk by him, I'm not sure how to cite it, see URL: http://www.cs.ox.ac.uk/ralf.hinze/talks/}. The interesting thing
 is that such a variable \lstinline{self} is late-bound, or open to the
@@
 
-\lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
+\lstinputlisting[linerange=31-37]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
 
@@
 
-\lstinputlisting[linerange=44-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
+\lstinputlisting[linerange=41-47]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
 \bruno{Fix should be replaces by Open. Please coordinate with Huang to
@@
 \paragraph{Closing the Open Recursion} An implementation of
- standard the standard Fibonacci function is recovered by closing 
+ the standard Fibonacci function is recovered by closing
 the open recursion using a \emph{lazy fixpoint combinator}:
@@
 
+\lstinputlisting[linerange=51-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB_FIX
+
 In Scala the lazy fixpoint combinator is straightforwardly implemented~\cite{} as follows:
 
-\lstinputlisting[linerange=27-30]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
+\lstinputlisting[linerange=27-27]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
 \bruno{can be defined in a single line.}
 \noindent Note that the function \lstinline{f} has a by-name parameter
-of type \lstinline{"=>} \lstinline{A"}. This is necessary to ensure
+of type \lstinline{"=>} \lstinline{T"}. This is necessary to ensure
 termination of well-founded recursive calls. In such cases,
@@
 The definition is straightforward:
-\lstinputlisting[linerange=66-66]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_MERGE
+\lstinputlisting[linerange=66-67]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_MERGE
 

### 2016-12-22 paper/sections/PackratParsers.tex
Fix

@@
 followed by a function which is applied to a parser to return new
-values. Finally, a generic \lstinline{runParser}
+values. Finally, a generic \inlinecode{parse}
 function is used for testing:
@@
 \lstinputlisting[linerange=16-19]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_RUNPARSER
-Now \lstinline{runParser(p)("str(5)")} prints out \lstinline{"5"} as expected.
+Now \lstinline{parse(p)("str(5)")} produces \lstinline{"5"} as expected.
 

### 2016-12-22 Scala/Parser/src/PaperCode/Overview/Util.scala

@@ -6,6 +6,8 @@ import scala.util.parsing.combinator.syntactical._
 object Util extends StandardTokenParsers with PackratParsers {
 
   type Open[T] = (=> T) => T
+  
+  type Parser[T] = PackratParser[T]
 
   implicit class Alternative[E, F, G, H](x: F => (=> H) => PackratParser[E]) {
     def |(y: G => (=> H) => PackratParser[E]): (F with G) => (=> H) => PackratParser[E] =

### 2016-12-22 Scala/Parser/src/PaperCode/SampleParser.scala

@@ -5,17 +5,17 @@ import TAPL.Util._
 object SampleParser {
 
 //BEGIN_PACKRAT_EXAMPLE
-lexical.reserved += ("str")
+lexical.reserved += "str"
 lexical.delimiters += ("(", ")")
 
-val p : PackratParser[String] =
+val p: PackratParser[String] =
   "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toString }
 //END_PACKRAT_EXAMPLE
 
 //BEGIN_PACKRAT_RUNPARSER
-def runParser(p: Parser[_]): String => Unit = in => {
+def parse[E](p: Parser[E]): String => E = in => {
   val t = phrase(p)(new lexical.Scanner(in))
-  if (t.successful) println(t.get) else scala.sys.error(t.toString)
+  if (t.successful) t.get else scala.sys.error(t.toString)
 }
 //END_PACKRAT_RUNPARSER  
      
@@ -24,45 +24,46 @@ def runParser(p: Parser[_]): String => Unit = in => {
 object OpenRecursion {
 
 //BEGIN_OPENRECURSION_FIX
-def fix[A](f : (=> A) => A) : A = {
-  lazy val a : A = f(a)
-  a
-}
+def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
 //END_OPENRECURSION_FIX
 
 //BEGIN_OPENRECURSION_FIB
-def fib : Int => Int = i => i match {
+def fib: Int => Int = {
   case 0 => 0
   case 1 => 1
   case n => fib(n - 1) + fib(n - 2)
 }
 
-val x = fib(2) // 1
+fib(2)  // 1
 //END_OPENRECURSION_FIB
 
 //BEGIN_OPENRECURSION_FIB2
-type Fix[T] = (=> T) => T
-def fib2 : Fix[Int => Int] = self => i => i match {
+type Open[T] = (=> T) => T
+
+def fib2: Open[Int => Int] = self => {
   case 0 => 0
   case 1 => 1
   case n => self.apply(n - 1) + self.apply(n - 2)
 }
-
-val x2 = fix(fib2)(2) // 1
 //END_OPENRECURSION_FIB2
 
+//BEGIN_OPENRECURSION_FIB_FIX
+fix(fib2)(2)  // 1
+//END_OPENRECURSION_FIB_FIX
+
 //BEGIN_OPENRECURSION_COMPOSE
[…]

### 2016-12-22 paper/sections/ParsingLibrary.tex
Comments on Sec 6

@@
 \subsection{Fusion of Concepts}\label{subsec:fusion}
+
+\huang{These alternatives could be discussed in the discussion section, not here. 6.1 could only talks about library code. And the titles should be changed as well}
 
@@
 It consists of some types wrapped up by synonyms, basic functions \lstinline{fix} and \lstinline{runParser}, and the combinator \lstinline{|||} from Section~\ref{subsec:differentsyntax}.
+
+\huang{code needs to be polished}
 
@@
 
+\huang{The point of introducing Language Component could be: modular parsers enable us to build modular language features.
+We can abstract language features, compose them and thus reuse them.
+A language feature covers all the related stuffs we need, as a combination of syntax+parsing+operations.
+Rather than: we recommend users to ...}
+
 Having pointed out that our goal is to realize modular and extensible parsers, we recommend users to write code in a modular and
@@
     \item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax, with several methods corresponding to the production rules.
+\huang{Production rule is about the \emph{concrete} syntax, The methods correspond to \emph{abstract} syntax.
+}
 \item \textbf{Parser:} implementation of the parser with lexing, using open recursion and Object Algebras, with its scoping trait abstracted over the type parameters.
@@
 
+\huang{the title could be more concrete, as all of the content is about the usage of langauge components}
+
+\huang{need to refactor the code as well}
+
 We have optimized the framework for not only modular parsers, but also languages, and that is the reason why we call them language components. To achieve

### 2016-12-22 paper/sections/Overview.tex
Choosing the parsing technique.

@@
-\section{Overview}\label{sec:overview}
+\section{To be discussed in Introduction}
 
-This section gives an overview of our library \name, from a user's perspective.
-Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions.
-We start from Section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers.
-Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse.
-Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls.
-In Section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers.
-Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.
-
-\huang{done} \bruno{Write a summary for this section}
-
-\huang{done}\bruno{Fill in all the missing references}
-
-\subsection{Choosing the Parsing Technique}\label{subsec:overview-parsing}
-
-%A technique for type-safe modular parsing should the following 3
-%features: \emph{modular type-checking}; \emph{separate compilation};
-%low performance overhead
+\haoyuan{I suggest this part to be moved to Introduction, and only discuss why Packrat is selected among
+different parser combinators in Overview.}
 
@@
 
-Unfortunately many parser combinators have important limitations for
-modular parsing. In particular several parser combinators,
-including the famous library Parsec~\cite{Leijen2001} library, require
-programmers to manually do \textit{left-recursion elimination}, and
-require significant amounts \textit{backtracking}. Both of these are
+\section{Overview}\label{sec:overview}
+
+This section gives an overview of our library \name, from a user's perspective.
+Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions.
+We start from Section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers.
+Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse.
+Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls.
+In Section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers.
+Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.
+
+\subsection{Choosing the Parsing Technique}\label{subsec:overview-parsing}
+
+%A technique for type-safe modular parsing should the following 3
+%features: \emph{modular type-checking}; \emph{separate compilation};
+%low performance overhead
+
+In the last section, we have argued that parser combinators are a suitable parsing technique for
+our purpose, as they naturally build modular parsers for type-checking.
+Unfortunately many parser combinators have important limitations.
+In particular several parser combinators,
+including the famous Parsec~\cite{Leijen2001} library, require
+programmers to manually do \textit{left-recursion elimination} and \textit{longest match composition}, and
+require significant amounts of \textit{backtracking}. All of them are
 problematic in a modular setting.
@@
 
+\paragraph{Longest Match Composition}
+Some parsing libraries provide combinators for alternative composition with longest match,
+while others do not. Now consider the non-left-recursive grammar:
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int> \alt <int> `+' <expr>
+\end{grammar}
+In Parsec, for instance, the parser
+\begin{lstlisting}[language=PlainCode]
+parser = parseInt <|> parseAdd
+\end{lstlisting}
+will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
+and terminates parsing. With traditional alternative
+composition in recursive descent parsers, the parser in the front successfully parses a substring of the input, then it unexpectedly
+finishes parsing and discards the rest, in spite that latter parsers might be able to parse the whole input.
+In contrast, \inlinecode{parser = parseAdd <|> parseInt} works as expected, yet manual composition for longest match is quite
+inconvenient, especially in a modular setting. When the grammar is extended with new rules, programmers are supposed to adjust
+the order of parsers, which is tedious.
+
 \paragraph{Backtracking} The need for backtracking is also problematic
-in a modular setting.
-For example, suppose that we only have \inlinecode{import..from} statements before, now we want to extend the grammar to add an \inlinecode{import..as} case.
-The third line of grammar below is the new case.
+in a modular setting. For example, suppose that we only have \inlinecode{import..from} statements before,
+but now we want to extend the grammar with the \inlinecode{import..as} case, as shown in the third line of grammar below.
 
@@
 
-Because the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning.
+Since the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning.
 Take Parsec as an example, its choice
@@
 Given the full grammar, we can decide where to put \inlinecode{try} for backtracking.
-However, in a modular setting we should always consider the worst case, that all alternatives may share common prefixes with future cases.
-Then we need to backtrack for all the branches, which results worst case exponential time complexity.
-In Parsec, we must add \inlinecode{try} everywhere:
+However, with modular parsing we are unable to overlook the full grammar, hence the worst case should be considered, namely all alternatives may share common prefixes with future cases.
+In that case we need to backtrack for all the branches.
+To avoid failures in the future, we have to add \inlinecode{try} everywhere:
 
@@
 \end{lstlisting}
+which results in the worst exponential time complexity.
 
-\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
-You want to say that when you know
-  the full grammar, you can figure out where ``try'' is
-  needed. Without the full grammar you'd need assume the worst
-  case. Make an effort to make your explanation cristal clear!
-I think you want to miss the abstract explanation that you give first,
-with the explanation about the concrete example.
-Start with ``The need for backtracking is also problematic
-in a modular setting. For example, suppose ...'' and synchronize
-the abstract explanation and the explanation for the example.
-}
+%\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
+%You want to say that when you know
+%  the full grammar, you can figure out where ``try'' is
+%  needed. Without the full grammar you'd need assume the worst
+%  case. Make an effort to make your explanation cristal clear!
+%I think you want to miss the abstract explanation that you give first,
+%with the explanation about the concrete example.
+%Start with ``The need for backtracking is also problematic
+%in a modular setting. For example, suppose ...'' and synchronize
+%the abstract explanation and the explanation for the example.
+%}
 
 \paragraph{Packrat Parsing}
-Fortunately more advanced parsing techniques such as Packrat
+Fortunately some more advanced parsing techniques such as Packrat
 parsing~\cite{Ford2002}, address the limitations of simple parser combinators
@@
 memoization to record the result of applying each parser at each
-position of the input, so that repeated computation is eliminated, and
-they support left-recursive grammars directly~\cite{warth2008}. All of these properties are very
-suitable for modularity, thus we use Packrat parsers as the underlying
-parsing technique. In Section \ref{sec:packratparsers}, we will give
-an overview of the Packrat parsing library implemented in Scala,
-which we use.
+position of the input, so that repeated computation is eliminated.
+Moreover, theoretically the algorithm behind Packrat parsers
+supports both direct and indirect left-recursion ~\cite{warth2008}.
+The current version of the library is still
+buggy with indirect left-recursion, but we believe that
+they will fix it in the future, and direct left-recursion is
+already practical to use for a large number of applications. All of these properties are very
+suitable for modularity, thus we decided to use Packrat parsers as the underlying
+parsing technique in \name.
 
@@
 combinators like Parsec, in cases that the performance and supporting
-of left-recursion are not major concerns.
+of left-recursion are not major concerns. A different library can even build a new
+\name with fancy features or higher efficiency.
 

### 2016-12-22 paper/sections/ObjectAlgebras.tex
Code format

@@
 
-Object Algebras, first introduced by ... et al as a solution to the famous \textit{Expression Problem}, provide extensibility on
-both data variants and operations for structures like abstract syntax trees. We integrate Object Algebras directly to enhance the
-extensibility of open parsers.
+Object Algebras~\cite{Oliveira2012} is a solution to the famous \textit{Expression Problem}, provide extensibility on
+both data variants and operations. It is suitable for structures like abstract syntax trees. We integrate Object Algebras directly to enhance the extensibility of \name.
 
@@
 At first the language only supports variables and applications:
-\begin{lstlisting}
-trait ExprAlg[E] {
-    def varE(x : String) : E
-    def appE(e1 : E, e2 : E) : E
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=4-8]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_EXPR
+
 The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
@@
 provides implementations for all cases. Below is an example of collecting all free variables in an expression:
-\begin{lstlisting}
-trait FreeVars extends ExprAlg[List[String]] {
-    def varE(x : String) = List(x)
-    def appE(e1 : List[String], e2 : List[String]) = e1 ++ e2
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=12-15]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_FREE_VARS
+
 Here \lstinline{FreeVars} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a list of strings as the result.
 Furthermore, one can define a new trait to implement pretty-printing on expressions:
-\begin{lstlisting}
-trait ExprPrint extends ExprAlg[String] {
-    def varE(x : String) = x
-    def appE(e1 : String, e2 : String) = "(" + e1 + " " + e2 + ")"
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=19-22]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_EXPR_PRINT
+
 On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Suppose we want to
 have literals in expressions, a new Object Algebra interface \lstinline{LitAlg} is defined as follows:
-\begin{lstlisting}
-trait LitAlg[E] extends ExprAlg[E] {
-    def litE(n : Int) : E
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=26-28]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT
+
 Now pretty-printing on the new language can be realized by code reuse, without modifying existing code:
-\begin{lstlisting}
-trait LitPrint extends ExprPrint {
-    def litE(n : Int) = n.toString
-}
-\end{lstlisting}
+
+\lstinputlisting[linerange=32-34]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT_PRINT
+
 On this, to create an expression of \lstinline{LitAlg}, a generic method is defined as follows:
-\begin{lstlisting}
-def build[E](alg : LitAlg[E]) : E =
-    alg.appE(alg.varE("x"), alg.litE(3))
-\end{lstlisting}
+
+\huang{todo: better name for build}
+\lstinputlisting[linerange=48-48]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_BUILD
+
 Such a method implicitly represents the expression \lstinline{"x 3"}. The code
+
 \begin{lstlisting}
@@
 \end{lstlisting}
+
 results in \lstinline{"(x 3)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
@@
 replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:
-\begin{lstlisting}
-trait Refactor[E] extends ExprAlg[E] {
-    def alg : SubAlg[E]
-    def env : Map[String, Int]
-    def varE(x : String) = if (env.contains(x)) alg.litE(env(x)) else alg.varE(x)
-}
-\end{lstlisting}\haoyuan{Is this example OK?}
+
+\lstinputlisting[linerange=38-42]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_REFACTOR
+
 When \lstinline{env = Map("y"} \lstinline{-> 3)}, the \lstinline{ExprAlg} expression \lstinline{"x y"} is changed to \lstinline{LitAlg} expression \lstinline{"x 3"}. The abstract method

### 2016-12-26 paper/sections/Overview.tex
Polishing overview & trying a more concise example.

@@
 
-This section gives an overview of our library \name, from a user's perspective.
+This section gives an overview of our library \name, and the problem that motivates our work.
 Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions.
@@
 In Section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers.
-Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.
+Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.\haoyuan{TODO}
 
@@
 Moreover, theoretically the algorithm behind Packrat parsers
-supports both direct and indirect left-recursion ~\cite{warth2008}.
+supports both direct and indirect left-recursion~\cite{warth2008}.
 The current version of the library is still
@@
 
+\subsection{Packrat Parsing and Our Library \name}\label{subsec:overview-library-api}
+
+Our prototype library is implemented in Scala, with its foundation on Packrat Parsing. In the current version of Scala (2.11), Packrat Parsing is included
+in the Scala standard parser combinator library, which is an external jar file. The library provides a couple of
+parser combinators, together with some basic parsers. Below we present an example to illustrate Scala parsers.
+
+\paragraph{An Example}
+A Packrat parser has type \lstinline{PackratParser[E]} for some
+\lstinline{E}, which indicates the type of results it produces.
+%\bruno{the representation of what?}
+And we use \lstinline{lexical} for its lexing.
+
+\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_EXAMPLE
+\noindent
+Note that the code should be defined in an enclosing type that extends
+\lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers}
+for lexing and
+\lstinline{scala.util.parsing.combinator.PackratParsers}.
+The first two lines are used for lexing. The third line defines a
+parser which can, for instance, parse \lstinline{"str(5)"} and produce
+the integer \lstinline{5}, where \lstinline{numericLit} is just a basic parser
+in the API. Table~\ref{tab:packrat} gives a closer look at part of the API, which covers our experiments.
+
+\begin{table}[t]
+\begin{tabular}{l}
+\hline
+\begin{lstlisting}
+def ~[U](q: => Parser[U]): Parser[~[T, U]]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator for sequential composition. \\
+\hline
+\begin{lstlisting}
+def ^^[U](f: (T) => U): Parser[U]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator for function application. \\
+\hline
+\begin{lstlisting}
+def ^^^[U](v: => U): Parser[U]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
+\hline
+\begin{lstlisting}
+def <~[U](q: => Parser[U]): Parser[T]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
+\hline
+\begin{lstlisting}
+def ~>[U](q: => Parser[U]): Parser[U]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
+\hline
+\begin{lstlisting}
+def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
+\end{lstlisting} \\
+\hspace{.2in}- A parser generator for interleaved repetitions. \\
+\hline
+\begin{lstlisting}
+def ident: Parser[String]
+\end{lstlisting} \\
+\hspace{.2in}- A parser which matches an identifier. \\
+\hline
+\begin{lstlisting}
+def numericLit: Parser[String]
+\end{lstlisting} \\
+\hspace{.2in}- A parser which matches a numeric literal. \\
+\hline
+\begin{lstlisting}
+def |[U >: T](q: => Parser[U]): Parser[U]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator for alternative composition. \\
+\hline
+\begin{lstlisting}
+def |||[U >: T](q0: => Parser[U]): Parser[U]
+\end{lstlisting} \\
+\hspace{.2in}- A parser combinator for alternative with longest match composition. \\
+\hline \\
+\end{tabular}
+\caption{Part of Scala Parser API.\bruno{We should double-check that this table lists the
+  combinators used in the paper.}}\label{tab:packrat}
+\end{table}
+
+\begin{table}[t]
+\begin{tabular}{l}
+\hline
+\begin{lstlisting}
+type OpenParser[Alg, R, E] = Alg => (=> R) => PackratParser[E]
+\end{lstlisting} \\
+\hspace{.2in}- The type constructor for modular parsers. \\
+\hline
+\begin{lstlisting}
+def parse(p: Parser[_]): String => Unit
+\end{lstlisting} \\
+\hspace{.2in}- A function that applies a parser to an input, and prints the result as a string. \\
+\hline
+\begin{lstlisting}
+implicit def <|>(x: OpenParser[Alg1, R, E], y: OpenParser[Alg2, R, E])
+    : OpenParser[Alg1 with Alg2, R, E]
+\end{lstlisting} \\
+\hspace{.2in}- Open parser combinator for alternative with longest match composition. \\
+\hline \\
+\end{tabular}
+\caption{Our parser combinator library \name API.\haoyuan{I don't know if I can present <|> like this.}}\label{tab:parser}
+\end{table}
+
+Finally, a generic \inlinecode{parse}
+function is used for testing:
+
+\lstinputlisting[linerange=16-19]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+Now \lstinline{parse(p)("str(5)")} produces \lstinline{5} as expected.
+
+\paragraph{\name} Our library is constructed by only around twenty lines of code. We have wrapped up the type of parsers
+to make them modular and extensible, and we use our custom combinator \lstinline{<|>} for alternative composition of those
+modular ``open parsers'' with longest match, built on \lstinline{|||}. Table~\ref{tab:parser} shows the API that \name provides for modular parsing.
+
 \subsection{The Parsing Expression Problem}\label{subsec:overview-problem}
-\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
-  the solution without pointing out what the problem is first. What
-  you need to do is: First show what happens with conventional
-  parsers: at some point, if you add extensions the recursive calls
-  will be wrong. Then you show (in the next section) the solution:
- use delegation/open recursion.}
+%\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
+%  the solution without pointing out what the problem is first. What
+%  you need to do is: First show what happens with conventional
+%  parsers: at some point, if you add extensions the recursive calls
+%  will be wrong. Then you show (in the next section) the solution:
+% use delegation/open recursion.}
 
-\huang{partly done, only extend a new language construct in the example, how about leaving extension of new operations in the OA subsection?}\bruno{I think we need to set up a challenge here, similar to the challenge of
-  the expression problem. The challenge should be like. Build a parser
-for a simple expression language, then extend the language with both
-a new language construct and a new operation. This section will show
-how we can do that using traditional parsers, but without modular
-type-safety and separate compilation. The remaining sections will show
-that the techniques introduced by us, enable us to solve those two challenges.}
+%\huang{partly done, only extend a new language construct in the example, how about leaving extension of new operations in the OA subsection?}\bruno{I think we need to set up a challenge here, similar to the challenge of
+%  the expression problem. The challenge should be like. Build a parser
+%for a simple expression language, then extend the language with both
+%a new language construct and a new operation. This section will show
+%how we can do that using traditional parsers, but without modular
+%type-safety and separate compilation. The remaining sections will show
+%that the techniques introduced by us, enable us to solve those two challenges.}
 
+In the following text we are going to illustrate the problem that motivates our work.
 Typical parsers defined by parser combinators follow the grammar
@@
 which support the easy addition of new alternative classes. We
-illustrate this problem with a simple expression language and extension.
+illustrate this with a simple expression language and extensions.
 
-Suppose we want to parse a simple language of variables and
-applications.
+Suppose we want to parse a simple language of literals and
+additions. The concrete syntax is as follows:
 
@@
 \begin{grammar}
-<expr> ::= <ident>
-    \alt <expr> <expr>
-    \alt `(' <expr> `)'
+<expr> ::= <int>
+    \alt <expr> `+' <expr>
 \end{grammar}
+%\alt `(' <expr> `)'
 
-It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases. Note that we have imported
-the Scala Packrat Parsing library in code below, and use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} throughtout the paper.
-\huang{This is the first time to demonstrate code in this paper.
+This time we allow both parts of an addition to be expressions.
+It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases:
+%\huang{This is the first time to demonstrate code in this paper.
 Readers may not know what those combinators/types are.
@@
 
-\lstinputlisting[linerange=8-18]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
+\lstinputlisting[linerange=8-24]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
 
-\haoyuan{trait Expr with pretty printing? Shall we add client code here?}
+\noindent Note that we use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} in the rest of paper.
+As the combination of \inlinecode{pVar}, \inlinecode{pApp}, \inlinecode{pExpr} can parse expressions like \lstinline{"1+2"} correctly.
+At this point, one would like such code to be extensible in two dimensions, for example:
+\begin{itemize}
+\item \textbf{Challenge 1}: adding a new case (or rather, a new construct) to \lstinline{Expr};
+\item \textbf{Challenge 2}: adding a new operation to \lstinline{Expr}.
+\end{itemize}
 
-\noindent As the combination of \inlinecode{pVar}, \inlinecode{pApp} and \inlinecode{pParen}, \inlinecode{pExpr} can parse expressions like \lstinline{"x (y z)"} correctly.
-
-\paragraph{Attempt to Extend the Parser} Now consider we are extending the abstract syntax as well as the parsers.
-Hence we introduce lambda abstractions as a new case, which refers to the famous \textit{untyped lambda calculus}.
+\paragraph{Attempt to Extend the Parser} Now we firstly consider challenge 1, namely we are extending the syntax as well as the parser. Hence we introduce variables as a new case.
 
@@
 <expr> ::= ...
-    \alt `\\' <ident> `.' <expr>
+   \alt <ident>
 \end{grammar}
@@
 
-\lstinputlisting[linerange=22-22]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
+\lstinputlisting[linerange=28-30]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
 
-Since we already have the parser for variables and applications, we would like to build the new parser by reusing the old one.
-So we compose \inlinecode{pLam}, which parses lambda abstractions, with the old parser.
+Since we already have the parser for literals and additions, we would like to build the new parser by reusing the old one.
+So we compose \inlinecode{pVar}, which parses variables, with the old parser.
 \inlinecode{pExtExpr} is the new function to parse expressions.
 
-\lstinputlisting[linerange=26-31]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXT
+\lstinputlisting[linerange=34-37]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXT
 
-Unfortunately, this extended parser does not work properly to parse all possible expressions defined by the new grammar.
-For example, it fails to parse \inlinecode{(\\x.x)} \inlinecode{(\\y.y)} which is obviously valid.
+Unfortunately, this extended parser does not work properly to parse all possible expressions defined by the new grammar. For example, it fails to parse \inlinecode{"1 + x"} which is obviously valid.
 
 \paragraph{Problem: Hard-coded Recursive Calls} The critical problem of extending parsers directly is that the recursive calls in old parsers are hard-coded.
-Specifically, \inlinecode{pApp} makes two recursive calls to parse two consecutive expressions as an application.
+Specifically, \inlinecode{pAdd} makes two recursive calls to parse two sub-expressions around \lstinline{"+"}.
 Before the extension, we implemented it by using \inlinecode{pExpr} for those recursive calls.
 It worked fine at that time because \inlinecode{pExpr} covers all cases of the old grammar.
-However, after the extension of lambda abstractions, the proper function for such recursive calls should be \inlinecode{pExtExpr} instead of \inlinecode{pExpr}, because the new function \inlinecode{pExtExpr} includes the lambda case in the new grammar.
-\lstinline{pParen} has the similar problem.
+However, after the extension of variables, the proper function for such recursive calls should be \inlinecode{pExtExpr} instead of \inlinecode{pExpr}, because the new function \inlinecode{pExtExpr} includes the new extension.
 
-That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{(\\x.x)} \inlinecode{(\\y.y)}.
-To resolve this issue, \inlinecode{pApp} and \inlinecode{pParen} need to be aware of the change of grammar.
-We have some direct solutions:
+That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{"1 + x"}.
+To resolve this issue, \inlinecode{pAdd} needs to be aware of the change of grammar.
+There are some possible solutions:
 
 \begin{itemize}
-    \item Keep the old code, rewrite non-extensible functions like \inlinecode{pApp} and \inlinecode{pParen}.
-\\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
+    \item Keep the old code, rewrite the non-extensible function \inlinecode{pAdd}. \\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
     \item Modify the old code, correct the recursive calls by proper functions.
-\\We can avoid code duplication so as to get code reuse, but separate compilation is sacrificed because of modification on the old code.
+\\We can avoid duplication for code reuse, but separate compilation is sacrificed because of modification on the existing code.
     \item ??? \huang{third one? use dynamic languages?}
@@
 
-Hard-coded recursive calls prevent our parsers from extensions.
+Hard-coded recursive calls prevent parsers from extensions.
 This problem can be solved using delegation, which can be encoded with
@@
 
-\paragraph{Open Parsing} Consider the expression language of only variables and applications in the previous section. The parser can be rewritten as follows:
+\paragraph{Open Parsing} Consider the expression language of only literals and additions in the previous section. The parser can be rewritten as follows:
 
-\lstinputlisting[linerange=52-62]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_BASE
+\lstinputlisting[linerange=58-68]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_BASE
 
@@
 represents \emph{the parser} of expressions, which is dynamiclly
-explained by the argument. Using an abstract parameter provides
+explained by the argument. Using such an abstract ``self-reference'' provides
 extensibility to parsers, because the recursive calls are not
@@
 to the grammar!
-
-We take \inlinecode{pApp} as an example again. It uses
+We take \inlinecode{pAdd} as an example again. It uses
 \inlinecode{self} instead of hard-coded \inlinecode{pExpr} for
 recursive parsing. Depending on the real argument, \inlinecode{self}
-can be a parser which covers only variables and applications, or a
-parser which also supports the lambda case.
+can be a parser which covers only literals and additions, or a
+parser which also supports variables.
 
@@
 
-\huang{done}\bruno{Too much detail is being given here about open
-  recursion. In Section 2 we should talk only about \emph{how} to use
-open recursion, and not how open recursion is implemented. We
-can summarize, for example the API related to open recursion.
-Section 4 is where the implementation and detailed explanation
-about open recursion should come.
-}
+%\huang{done}\bruno{Too much detail is being given here about open
+%  recursion. In Section 2 we should talk only about \emph{how} to use
+%open recursion, and not how open recursion is implemented. We
+%can summarize, for example the API related to open recursion.
+%Section 4 is where the implementation and detailed explanation
+%about open recursion should come.
+%}
 
-\lstinputlisting[linerange=67-71]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
+\lstinputlisting[linerange=73-77]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
 
-\huang{done}\bruno{Please polish up code. You probably want to consistently use
-the variable name ``self'' whenever you want to refer to the whole
-parser (instead of ``e'' that you use at the moment). Also you should
-use ``Open'' (which you currently call ``Fix'') instead of using the
-longwinded ``(=> PackratParser[Expr]) => PackratParser[Expr]''}
-
-\huang{done}\bruno{Also, for consistency, why isn't ``pVar'' open? I.e. using a
-  self reference as well? I know that it is not recursive, but giving
-  it a different treatment makes it more confusing for users.}
-
-\huang{done}\bruno{You should use a more OO approach in the example.
-Rather than case classes show an Expr trait with pretty printing.
-This will enable you to talk about the extensibility and object
-algebras better later on.
-}
+%\huang{done}\bruno{Please polish up code. You probably want to consistently use
+%the variable name ``self'' whenever you want to refer to the whole
+%parser (instead of ``e'' that you use at the moment). Also you should
+%use ``Open'' (which you currently call ``Fix'') instead of using the
+%longwinded ``(=> PackratParser[Expr]) => PackratParser[Expr]''}
+%
+%\huang{done}\bruno{Also, for consistency, why isn't ``pVar'' open? I.e. using a
+%  self reference as well? I know that it is not recursive, but giving
+%  it a different treatment makes it more confusing for users.}
+%
+%\huang{done}\bruno{You should use a more OO approach in the example.
+%Rather than case classes show an Expr trait with pretty printing.
+%This will enable you to talk about the extensibility and object
+%algebras better later on.
+%}
 
@@
 
-\lstinputlisting[linerange=76-82]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
+\lstinputlisting[linerange=82-88]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
 
@@
 
-\lstinputlisting[linerange=88-88]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
+\lstinputlisting[linerange=94-94]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
 

### 2016-12-26 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -5,29 +5,35 @@ import PaperCode.Overview.Util._
 object Overview2 {
 
 //BEGIN_OVERVIEW_SIMPLE_EXPR
-abstract class Expr
-class Var(x: String) extends Expr
-class App(e1: Expr, e2: Expr) extends Expr
+// Abstract syntax
+trait Expr { def print : String }
+class Lit(x: Int) extends Expr {
+  def print = x.toString
+}
+class Add(e1: Expr, e2: Expr) extends Expr {
+  def print = "(" + e1.print + "+" + e2.print + ")"
+}
 
+// Parser with lexing
 trait BaseParser {
-  lexical.delimiters += ("(", ")")
-  val pExpr: Parser[Expr] = pVar ||| pApp ||| pParen
-  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
-  val pApp: Parser[Expr] = pExpr ~ pExpr ^^ { case e1 ~ e2 => new App(e1, e2) }
-  val pParen: Parser[Expr] = "(" ~> pExpr <~ ")"
+  lexical.delimiters += ("+")
+  val pLit: Parser[Expr] = numericLit ^^ { x => new Lit(x.toInt) }
+  val pAdd: Parser[Expr] =
+    pExpr ~ ("+" ~> pExpr) ^^{ case e1 ~ e2 => new Add(e1, e2) }
+  val pExpr: Parser[Expr] = pLit ||| pAdd
 }
 //END_OVERVIEW_SIMPLE_EXPR
 
 //BEGIN_OVERVIEW_SIMPLE_LAM
-class Lam(x: String, e: Expr) extends Expr
+class Var(x: String) extends Expr {
+  def print = x
+}
 //END_OVERVIEW_SIMPLE_LAM
 
 //BEGIN_OVERVIEW_SIMPLE_EXT
 trait ExtParser extends BaseParser {
-  lexical.delimiters += ("\\", ".")
-  val pExtExpr: Parser[Expr] = pExpr ||| pLam
-  val pLam: Parser[Expr] =
-    "\\" ~> ident ~ ("." ~> pExtExpr) ^^ { case x ~ e => new Lam(x, e) }
+  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
+  val pExtExpr: Parser[Expr] = pExpr ||| pVar
 }
 //END_OVERVIEW_SIMPLE_EXT
 
@@ -52,13 +58,13 @@ object Overview3 {
[…]

### 2016-12-26 Scala/Parser/src/PaperCode/SampleParser.scala

@@ -8,8 +8,8 @@ object SampleParser {
 lexical.reserved += "str"
 lexical.delimiters += ("(", ")")
 
-val p: PackratParser[String] =
-  "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toString }
+val p: PackratParser[Int] =
+  "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toInt }
 //END_PACKRAT_EXAMPLE
 
 //BEGIN_PACKRAT_RUNPARSER

### 2016-12-26 paper/sections/CaseStudy.tex
Some more comments on Case studies and Implementation

@@
 \section{Case Studies}\label{sec:casestudy}
+
+To demonstrate the utility of our modular parsing library, we
+implemented parsers of the first 18 calculi from book \textit{Types
+  and Programming Languages} (TAPL). TAPL introduces several
+calculi from simple to complex, by gradually adding new features to
+syntax. \bruno{mention that you compare against an existing
+  non-modular implementation.}
 
@@
 
-To demonstrate the utility of our modular parsing library, we implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages}.
-In the book it introduces several calculi from simple to complex, by gradually adding new features to syntax.
-It is suitable for our case study for mainly two reason.
-First, the calculi are arguably practical as we want to show our library could be used in real world development.
-One of the most complex calculi in our case study is System F (with several extensions such as records and variants), which could be the foundation of practical functional programming languages.
-Second, the evolution of calculi in the book reveals the advantages of modular representation of abstract syntax and modular parsing, which is the key functionality of our library.
-By decomposing those calculi into smaller components and reusing the common ones, we can obtain considerably code reuse in parsing with little performance penalty, as shown later.
+The calculi in the TAPL book are suitable for our case study for mainly two reasons.
+Firstly, the calculi are arguably capture many of the language features 
+required in realistic programming languages. Secondly, the evolution of
+calculi in the book reveals the advantages of modular representation
+of abstract syntax and modular parsing, which is the key functionality
+of our library. By decomposing those calculi into smaller components
+and reusing the common ones, we can obtain considerably code reuse in
+parsing with little performance penalty, as shown later.
 
-Using the modular parsing techniques discussed before, we extract reusable \textit{components} from syntax of all the calculi.
-Each component, which may contain several syntactical structures, is a reusable fragment representing a certain feature in the syntax.
-For example, \lstinline{VarApp} component represents the variable case and application case of expressions.
-In our case study, each component is represented by a Scala object which has three parts.
-\lstinline{Alg} is an object algebra interface for the abstract syntax.
-\lstinline{Print} is an object algebra implements that interface \lstinline{Alg}, for the pretty printing operation.
-It is not necessary for parsing, but we include it here for demonstration.
+\paragraph{Extracting Language Components}
+Using the modular parsing techniques discussed before, we extract
+reusable \textit{components} from syntax of all the calculi. Each
+component, which may contain several syntactical structures, is a
+reusable fragment representing a certain feature in the syntax. For
+example, \lstinline{VarApp} component represents the variable case and
+application case of expressions. In our case study, each component is
+represented by a Scala object which has three parts. \lstinline{Alg}
+is an object algebra interface for the abstract syntax.
+\lstinline{Print} is an object algebra implements that interface
+\lstinline{Alg}, for the pretty printing operation. It is not
+necessary for parsing, but we include it here for demonstration.
 \lstinline{Parser} is the modular parser for this piece of syntax.
@@
 
-We use some naming conventions in our code, \lstinline{E} represents expressions, \lstinline{T} represents types and \lstinline{K} represents kinds.
-In this component \lstinline{VarApp} we only have expressions.
-The \lstinline{Alg} and \lstinline{Print} are standard object algebra operations, thus we will focus on the modular parser \lstinline{Parser}.
-Each parser carries lexical information and parsing functions.
-The lexical information includes reserved words, delimiters, etc.
-Parsing functions are similar with we described before, except that we use record types under some subtyping constraints here for the explicit self-reference.
-The record type contains a parser for expressions \lstinline{pE}, a parser for types \lstinline{pT} and a parser for kinds \lstinline{pK}, if the corresponding one is needed when parsing the syntax of this component.
-Here the type of self-reference, namely \lstinline{F}, is only required to have a parser of expressions \lstinline{pE}.
-Then we can build the parsing function \lstinline{pVarAppE}, the suffix \lstinline{E} also denotes that it will parse an expression.
+We use some naming conventions in our code, \lstinline{E} represents
+expressions, \lstinline{T} represents types and \lstinline{K}
+represents kinds. In this component \lstinline{VarApp} we only have
+expressions. The \lstinline{Alg} and \lstinline{Print} are standard
+object algebra operations, thus we will focus on the modular parser
+\lstinline{Parser}. Each parser carries lexical information and
+parsing functions. The lexical information includes reserved words,
+delimiters, etc. Parsing functions are similar with we described
+before, except that we use record types under some subtyping
+constraints here for the explicit self-reference. The record type
+contains a parser for expressions \lstinline{pE}, a parser for types
+\lstinline{pT} and a parser for kinds \lstinline{pK}, if the
+corresponding one is needed when parsing the syntax of this component.
+Here the type of self-reference, namely \lstinline{F}, is only
+required to have a parser of expressions \lstinline{pE}. Then we can
+build the parsing function \lstinline{pVarAppE}, the suffix
+\lstinline{E} also denotes that it will parse an expression.
 
-Each calculus could be composed directly from components and other calculi, if it has those common structures in the syntax.
-For example, the calculus \lstinline{Untyped}, representing the famous untyped lambda calculus, can be constructed from components \lstinline{VarApp} and untyped lambda abstraction \lstinline{UntypedAbs}.
+\paragraph{Composing Language Components}
+Each calculus could be composed directly from components and other
+calculi, if it has those common structures in the syntax. For example,
+the calculus \lstinline{Untyped}, representing the famous untyped
+lambda calculus, can be constructed from components \lstinline{VarApp}
+and untyped lambda abstraction \lstinline{UntypedAbs}.
 
@@
 
-When composing the calculus from components, all the fields can be easily combined by Scala’s \lstinline{extends...with} keywords as shown in the code.
-For \lstinline{Parser}, the subtyping constraint for the explicit self-reference should satisfy all the requirement of the parsers it extends from.
-\huang{Talk about | here?}
+When composing the calculus from components, all the fields can be
+easily combined by Scala’s \lstinline{extends...with} keywords as
+shown in the code. For \lstinline{Parser}, the subtyping constraint
+for the explicit self-reference should satisfy all the requirement of
+the parsers it extends from. \huang{Talk about | here?}
 
-Finally, we have demo code for every calculus.
-The parsing result is a string because we provide the pretty printing algebra to the parser, and it can be changed as long as the concrete algebra is available.
-In this case study we only have pretty printing as the operation, since it is enough for demonstration.
+Finally, we have demo code for every calculus. The parsing result is a
+string because we provide the pretty printing algebra to the parser,
+and it can be changed as long as the concrete algebra is available. In
+this case study we only have pretty printing as the operation, since
+it is enough for demonstration.
 
@@
 
-In the next section we will show more examples and talk further about how we extend and compose these components and calculi.
+In the next section we will show more examples and talk further about
+how we extend and compose these components and calculi.
 
@@
 
-As mentioned before, we decompose all the calculi into reusable components for modularity.
-Figure \ref{fig:dependency} shows the dependency of all the components and calculi in our case study.
-Gray boxes are calculi and white boxes are components.
-An arrow starting from box A to box B denotes that B includes and thus reuses A.
+As mentioned before, we decompose all the calculi into reusable
+components for modularity. Figure \ref{fig:dependency} shows the
+dependency of all the components and calculi in our case study. Gray
+boxes are calculi and white boxes are components. An arrow starting
+from box A to box B denotes that B includes and thus reuses A.
 
-As shown in the graph, some components such as \lstinline{VarApp} are created from scratch, while others such as \lstinline{Typed} are extended from existing components.
-The \lstinline{Typed} component extends from \lstinline{VarApp} and has two more syntactical structures which are typed lambda abstraction and function type (arrow type).
-It is the core of many calculi in this case study, and we use it here as an example for demonstrating the extensibility of components.
+As shown in the graph, some components such as \lstinline{VarApp} are
+created from scratch, while others such as \lstinline{Typed} are
+extended from existing components. The \lstinline{Typed} component
+extends from \lstinline{VarApp} and has two more syntactical
+structures which are typed lambda abstraction and function type (arrow
+type). It is the core of many calculi in this case study, and we use
+it here as an example for demonstrating the extensibility of
+components.
 
@@
 
-Except building a component by extending an existing one, here we have another extension which is adding a new sort of syntax.
-It has been discussed in section \ref{subsec:openrecursion}.
-In this example, originally \lstinline{VarApp} has only one sort of syntax which is expressions, and then \lstinline{Typed} introduces types to the syntax.
-We add an extra type parameter \lstinline{T} for it, so that it is able to be distinguished.
-The \lstinline{Parser} also requires the self-reference to have a field \lstinline{pT} for parsing types, which is represented by the subtyping constraint.
-With this support, the \lstinline{pTypedT} function parses arrow types using \lstinline{pT}.
-Another point worth mentioning is that we use a private local function \lstinline{pAbsE} here for parsing the typed lambda abstraction case, and we can directly combine it with the inherited \lstinline{pVarApp} to obtain our parsing function of expressions.
+Except building a component by extending an existing one, here we have
+another extension which is adding a new sort of syntax. It has been
+discussed in section \ref{subsec:openrecursion}. In this example,
+originally \lstinline{VarApp} has only one sort of syntax which is
+expressions, and then \lstinline{Typed} introduces types to the
+syntax. We add an extra type parameter \lstinline{T} for it, so that
+it is able to be distinguished. The \lstinline{Parser} also requires
+the self-reference to have a field \lstinline{pT} for parsing types,
+which is represented by the subtyping constraint. With this support,
+the \lstinline{pTypedT} function parses arrow types using
+\lstinline{pT}. Another point worth mentioning is that we use a
+private local function \lstinline{pAbsE} here for parsing the typed
+lambda abstraction case, and we can directly combine it with the
+inherited \lstinline{pVarApp} to obtain our parsing function of
+expressions.
 
-Since calculi and components have the similar signature, each calculus can also be extended and reused directly.
-Our last example shows how to build calculus \lstinline{FullRef} from extending \lstinline{FullSimple}, another calculus.
+Since calculi and components have the similar signature, each calculus
+can also be extended and reused directly. Our last example shows how
+to build calculus \lstinline{FullRef} from extending
+\lstinline{FullSimple}, another calculus.
 
@@
 
-Using our modular parsing library, dividing the syntax into small pieces and bind them with corresponding parsers to form reusable language component is convenient.
-From the dependency graph we know that the common components such as \lstinline{VarApp} are reused in lots of calculi.
-Such reuse could shorten the code considerably.
-We will show this advantage and exam the possible performance penalty in a quantitative way in the next section.
+Using our modular parsing library, dividing the syntax into small
+pieces and bind them with corresponding parsers to form reusable
+language component is convenient. From the dependency graph we know
+that the common components such as \lstinline{VarApp} are reused in
+lots of calculi. Such reuse could shorten the code considerably. We
+will show this advantage and exam the possible performance penalty in
+a quantitative way in the next section.
 
@@
 
-We compared our implementation of the parsers with an implementation we found online, written by author Ilya Klyuchnikov.
-Ilya's implementation is very suitable for comparison, because it is also written in Scala using the same parser combinator library.
-Furthermore, it includes parsers of all the 18 calculi we have, but written in a non-modular way.
-Thus it is not able to reuse existing code of parsing common syntactical features.
+We compared our implementation of the parsers with an implementation
+by  Ilya Klyuchnikov, available online. Ilya's
+implementation is suitable for comparison, because it is also
+written in Scala using the same parser combinator library.
+Furthermore, it includes parsers of all the 18 calculi we have, but
+written in a non-modular way. Thus it is not able to reuse existing
+code of parsing common syntactical features.
 
-Because Ilya's implementation has extra code such as semantics of those calculi, we removed all irrelevant code and only keep abstract syntax definition, parser and pretty printer for each calculus, to ensure a fair comparison.
+Because Ilya's implementation has extra code such as semantics of
+those calculi, we removed all irrelevant code and only keep abstract
+syntax definition, parser and pretty printer for each calculus, to
+ensure a fair comparison.
 
-The comparison is made from two aspects.
-First, we want to show that using our modular parsing library, the code can be reused in a significant way.
-For this propose, we measured source lines of code (SLOC) of the two implementations.
-All blank lines and comment lines are excluded, and we formatted the code of both implementations so that the length of each line does not exceed 120 characters.
-Second, users may concern the performance penalty caused by modularity.
-Thus we compared the execution time, as a major measurement of performance, between the two implementations.
-We prepared test files containing valid expressions in each calculus, then we used the corresponding parser to parse the file and the pretty printer to print the result.
+The comparison is made from two aspects. First, we want to show that
+using our modular parsing library, the code can be reused in a
+significant way. For this propose, we measured source lines of code
+(SLOC) of the two implementations. All blank lines and comment lines
+are excluded, and we formatted the code of both implementations so
+that the length of each line does not exceed 120 characters. Second,
+we are interested to assess the performance penalty caused by modularity. Thus
+we compared the execution time, as a major measurement of performance,
+between the two implementations. We prepared test files containing
+valid expressions in each calculus,\bruno{say something about what
+  those expressions are, size, and how they were generated.} then we used the corresponding
+parser to parse the file and the pretty printer to print the result.
 The execution time excluding reading input file was calculated.
@@
 
-Table \ref{tab:comparison} shows data of the comparison.
-The overall result of SLOC comparison is that 66.3\% of code is reduced using our library.
-That is because the code of some common syntax structures such as variable, lambda abstraction, etc., is reused several times in the whole project.
-We can see that the first two calculi \lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's implementation, because in such two cases we do not reuse any existing components.
-But in each of the following 16 calculi, we can compose our syntax definition, parser and pretty printer from extending reusable components, resulting considerably code reduction.
+Table \ref{tab:comparison} shows data of the comparison. The overall
+result of SLOC comparison is that 66.3\% of code is reduced using our
+library. That is because the code of some common syntax structures
+such as variables, lambda abstractions, etc., are reused several times in
+the whole project. We can see that in the first two calculi
+\lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's
+implementation, because in such two cases we do not reuse any existing
+components. But in each of the following 16 calculi, we can compose
+our syntax definition, parser and pretty printer from extending
+reusable components, resulting considerably code reduction.
 
-For the execution time comparison, we built a generator to randomly generate valid expressions according to the syntax of each calculus.
-Each test file corresponds with one calculus and it consists of 500 expressions randomly generated.
-The size of test files varies from 20KB to 100KB.
-After testing both implementation, we found that our version is about \huang{60\%} slower.
-Part of the reason is that our library requires more intermediate function calls, and the object algebra and self-reference needed to be passed as arguments every time.
-However, another important reason is that, because of the modular setting, we use a different alternative combinator which will try both branches and match the longest one.
-\huang{ref to somewhere before?} In contrast Ilya uses the ordinary alternative combinator which tries the second branch only if the first one fails.
-Although this difference slows our version, we want to mention that because of the memorization of Packrat parsers, we are just constant times slower, the big-O complexity is still the same.
+\paragraph{Performance}
+For the execution time comparison, we built a generator to randomly
+generate valid expressions according to the syntax of each calculus.
+Each test file corresponds with one calculus and it consists of 500
+expressions randomly generated. The size of test files varies from
+20KB to 100KB. After testing both implementation, we found that our
+version is about \huang{60\%} slower. \bruno{A table showing a summary
+of results is missing here. We discussed this before.}
+ Part of the reason is that our
+library requires more intermediate function calls, and the object
+algebra and self-references need to be passed as arguments every
+time. However, another important reason is that, because of the
+modular setting, we use a different alternative combinator which will
+try both branches and match the longest one. \huang{ref to somewhere
+  before?} In contrast Ilya uses the ordinary alternative combinator
+which tries the second branch only if the first one fails. Although
+this difference slows our version, we want to mention that because of
+the memoization of Packrat parsers, we are just constant times
+slower, the big-O complexity is still the same.

### 2016-12-27 paper/sections/Overview.tex
Overview ready for a second review.

@@
 
-\paragraph{Extensibility} Similarly, we extend the grammar by adding lambda abstractions as a new case.
-The new parser also extends the old one by adding \inlinecode{pLam} on it.
+\paragraph{Extensibility} Similarly, we extend the grammar by adding variables as a new case.
+The new parser also extends the old one by adding \inlinecode{pVar} on it.
 Notice that the old code is not touched, hence separate compilation is obtained.
 
-\lstinputlisting[linerange=82-88]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
+\lstinputlisting[linerange=82-85]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
 
@@
 
-All recursive calls of parsing `an expression' are called via the explicit parameter \inlinecode{self}.
+All recursive calls of parsing `an expression' are realized via the explicit parameter \inlinecode{self}.
 Once the \inlinecode{fix} function is applied, these calls will be updated appropriately.
-Parsing functions such as \inlinecode{pApp} do not need to be rewritten.
-As a consequence, this parser recognizes all valid expressions of the new grammar, including \inlinecode{(\\x.x)} \inlinecode{(\\y.y)} which cannot be parsed in the previous section.
+Parsing functions such as \inlinecode{pAdd} do not need to be rewritten.
+As a consequence, this parser recognizes all valid expressions of the new grammar, including \inlinecode{"1 + x"} which cannot be parsed previously.
 
-\lstinputlisting[linerange=94-94]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
+\lstinputlisting[linerange=91-91]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
 
@@
 hierarchy makes the addition of new operations over the AST
-problematic. Object Algebras~\cite{Oliveira2012} enable us to solve this problem, and
+problematic. Modification on existing code breaks the modularity. This
+refers to the famous \textit{Expression Problem} []. For challenge 2, suppose now we want to
+support evaluation on expressions, in Scala one attempt would be extending \lstinline{Expr} with
+evaluation:
+
+\lstinputlisting[linerange=111-111]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_ATTEMPT_EXPRWITHEVAL
+But then the programmer has to define new classes for literals and additions that extend old ones and \lstinline{EvalExpr}. Specifically,
+the critical point is that existing parsing code only generates \lstinline{Expr} objects with pretty printing, so modification on parsing is again required.
+
+In contrast, Object Algebras~\cite{Oliveira2012} enable us to solve this problem, as it separates data variants and
+operations, to
 offer high flexibility in the choice of operations to be performed
@@
 
-\paragraph{Parsing with Object Algebras} Using Object Algebras, the abstract syntax of the language of variables and applications is defined as below.
+\paragraph{Parsing with Object Algebras} Using Object Algebras, the abstract syntax of the expression language is defined as below.
 
@@
 
-Then we are able to define operations over the syntax in an extensible way. Here we write a pretty printing operation for this language.
+Then we are able to define operations over the syntax in a modular way. For instance, pretty printing operation can be realized as:
 
@@
 And also parser for it as below.
-Notice the parsing function \inlinecode{pExpr} takes an argument of type \inlinecode{ExprAlg[E]}, that means it accepts any instance of the algebra interface \inlinecode{ExprAlg}.
+Notice the parsing function \inlinecode{pExpr} now takes an argument of type \inlinecode{ExprAlg[E]}, which means it accepts any instance of \inlinecode{ExprAlg}.
+Such an instance can for example be \lstinline{Print}, which takes charge of operations on expressions, and parsing code only does delegation.
 
-\lstinputlisting[linerange=22-28]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PARSER
+\lstinputlisting[linerange=22-27]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PARSER
 
-\paragraph{Extensibility of Syntax} Following previous examples, the code below shows how to extend the language to support lambda abstractions.
+\haoyuan{I'm trying to avoid using terminology like algebra and algebra interface.}
 
-\lstinputlisting[linerange=32-46]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
+\paragraph{Extensibility of Syntax} Following previous examples, the code below shows how to extend the language with variables.
 
-For easily combining parsers using Object Algebras as ASTs, we have a special combinator \huang{we need to rename our combinator} defined in our library.
+\lstinputlisting[linerange=31-44]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
+Note that we use the new combinator \lstinline{<|>} defined in our library, to compose two parsing functions into one.
+The function \lstinline{pExtExpr} has the compound type \lstinline{ExprAlg[E] with VarAlg[E]} in its argument, where two language ASTs are combined to represent the new language.
 
 To use the parser, we must provide an algebra instance as the operation to construct the parsing results.
-In the code below, we use the pretty printing operation \inlinecode{LambdaPrint} as the algebra instance, so that the parsing result is a pretty printed AST.
+In the code below, we use the combination of \lstinline{Print} and \lstinline{PrintVar}, so that the parsing result is a pretty printing of the AST.
 
-\lstinputlisting[linerange=55-60]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
+\lstinputlisting[linerange=53-58]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
 
@@
 Here is an example of collecting free variables from an expression.
-We can use this operation during parsing and obtain a set of free variables.
+We can feed this operation to the parser and obtain a set of free variables.
 
-\lstinputlisting[linerange=74-80]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
+\lstinputlisting[linerange=72-78]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
+
+\haoyuan{I think in our library language ASTs are composed using intersection, not inheritance.}
 
@@
 
-\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
+\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
 \setlength{\grammarindent}{5em}
@@
 \begin{grammar}
-<expr> ::= <ident> \alt <expr> <expr> \alt `\\' <ident> `.' <expr>
+<expr> ::=  `\\' <ident> `:' <type> `.' <expr>
 \end{grammar}
@@
 
-Using Object Algebras, we can easily distinguish them in an extensible way, just by adding an extra type parameter.
+Using Object Algebras, we can easily distinguish them just by adding an extra type parameter. The abstract syntax is
+presented below:
 
-\lstinputlisting[linerange=65-69]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=63-67]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+In the next sections, we will discuss them with more details.

### 2016-12-27 Scala/Parser/src/PaperCode/Overview/OA.scala

@@ -6,43 +6,41 @@ object OA {
 
 //BEGIN_OVERVIEW_OA_ALG
 trait ExprAlg[E] {
-  def Var(x: String): E
-  def App(e1: E, e2: E): E
+  def lit(x: Int): E
+  def add(e1: E, e2: E): E
 }
 //END_OVERVIEW_OA_ALG
 
 //BEGIN_OVERVIEW_OA_PRINT
-trait ExprPrint extends ExprAlg[String] {
-  def Var(x: String): String = x
-  def App(e1: String, e2: String): String = "(" + e1 + " " + e2 + ")"
+trait Print extends ExprAlg[String] {
+  def lit(x: Int) = x.toString
+  def add(e1: String, e2: String) = "(" + e1 + " + " + e2 + ")"
 }
 //END_OVERVIEW_OA_PRINT
 
 //BEGIN_OVERVIEW_OA_PARSER
-trait ExprParser[E] {
-  lexical.delimiters += ("(", ")")
-  val pExpr: ExprAlg[E] => Open[PackratParser[E]] = alg => self =>
-    ident ^^ alg.Var |||
-      self ~ self ^^ { case e1 ~ e2 => alg.App(e1, e2) } |||
-      "(" ~> self <~ ")"
+trait BaseParser[E] {
+  lexical.delimiters += ("+")
+  val pExpr: ExprAlg[E] => Open[Parser[E]] = alg => self =>
+    numericLit ^^ alg.lit |||
+    self => self ~ ("+" ~> self) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
 }
 //END_OVERVIEW_OA_PARSER
 
 //BEGIN_OVERVIEW_OA_EXT
-trait LambdaAlg[E] extends ExprAlg[E] {
-  def Lam(x: String, e: E): E
+trait VarAlg[E] {
+  def varE(x: String): E
 }
 
-trait LambdaPrint extends LambdaAlg[String] with ExprPrint {
-  def Lam(x: String, e: String): String = "\\" + x + "." + e
+trait PrintVar extends VarAlg[String] {
+  def varE(x: String) = x
 }
 
-trait LambdaParser[E] extends ExprParser[E] {
-  lexical.delimiters += ("\\", ".")
-  val pLam: LambdaAlg[E] => Open[PackratParser[E]] = alg => self =>
[…]

### 2016-12-27 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -6,12 +6,12 @@ object Overview2 {
 
 //BEGIN_OVERVIEW_SIMPLE_EXPR
 // Abstract syntax
-trait Expr { def print : String }
+trait Expr { def print: String }
 class Lit(x: Int) extends Expr {
   def print = x.toString
 }
 class Add(e1: Expr, e2: Expr) extends Expr {
-  def print = "(" + e1.print + "+" + e2.print + ")"
+  def print = "(" + e1.print + " + " + e2.print + ")"
 }
 
 // Parser with lexing
@@ -62,7 +62,7 @@ trait BaseParser {
   val pLit: Open[Parser[Expr]] =
     self => numericLit ^^ { x => new Lit(x.toInt) }
   val pAdd: Open[Parser[Expr]] =
-    self => self ~ ("+" ~> self) ^^{ case e1 ~ e2 => new Add(e1, e2) }
+    self => self ~ ("+" ~> self) ^^ { case e1 ~ e2 => new Add(e1, e2) }
   val pExpr: Open[Parser[Expr]] =
     self => pVar(self) ||| pAdd(self)
 }
@@ -74,24 +74,21 @@ def parse[E](p: Parser[E]): String => E = {...}
 def fix[T](f: Open[T]): T = {...}
 def openParse[E](p: Open[Parser[E]]): String => E = parse[E](fix(p))
 
-openParse(new BaseParser {}.pExpr)("1 + x")
+openParse(new BaseParser {}.pExpr)("1 + 2").print // "(1 + 2)"
 //END_OVERVIEW_OPEN_USE
 */
 
 //BEGIN_OVERVIEW_OPEN_EXT
 trait ExtParser extends BaseParser {
-  lexical.delimiters += ("\\", ".")
-  val pExtExpr: Open[Parser[Expr]] =
-    self => pLam(self) ||| pExpr(self)
-  val pLam: Open[Parser[Expr]] =
-    self => "\\" ~> ident ~ ("." ~> self) ^^ { case x ~ b => new Lam(x, b) }
+  val pVar: Open[Parser[Expr]] = self => ident ^^ { x => new Var(x) }
+  val pExtExpr: Open[Parser[Expr]] = self => pExpr(self) ||| pVar(self)
 }
 //END_OVERVIEW_OPEN_EXT
 
   def use2(): Unit = {
[…]

### 2016-12-28 paper/sections/Introduction.tex
some text

@@
 techniques. Syntactic modularization techniques are quite popular in
-practice. Examples include many tools for developing Software-Product
+practice, due to their simplicity of implementation and use. 
+Examples include many tools for developing Software-Product
 Lines~\cite{}, some Language Workbenches~\cite{}, or extensible parser
@@
 approaches do not support separate compilation or modular
-type-checking either.\bruno{mention that a reason why syntactic 
-modularization techniques are popular is simplicity (in implementation
-and also use).}
+type-checking either.
 
-Semantic modularization techniques move away
-from syntactic composition techniques that rely on the textual source
-code. This allows us to go one step further in terms of modularity,
+Semantic modularization techniques go one step further in terms of modularity,
 and also enable components or features to be modularly type-checked
 and separately compiled. Modular type-checking and separate
-compilation are desirable properties to have from a Software
-Engineering point-of-view, and enable the composition of compiled
+compilation are desirable properties to have from a software
+engineering point-of-view. Furthermore they enable the composition of compiled
 binaries as well as ensuring the type-safety of the code composed of
 multiple components. Examples of semantic modularization techniques 
-include various approaches to \emph{family polymorphism}~\cite{}, as 
-well as several techniques for solving the Expression Problem. 
+include various approaches to \emph{family polymorphism}~\cite{},
+virtual classes~\cite{}, as 
+well as several other techniques for solving the Expression Problem~\cite{}. 
 \bruno{strenghten? mention dynamic languages (separate compilation 
 but no modular type-checking), and techniques that weaken type-safety 
-requirements.} Most semantic
+requirements.}
+
+Semantic modularization techniques are less widely used in practice 
+than syntactic techniques. This is partly due to the perceived need
+for sophisticated type systems, which are not available in mainstream 
+languages and may require more knowledge from users. However, recently 
+several lightweight modularization techniques have been shown to work 
+in mainstream programming languages like Java or Scala. Object
+Algebras are one such technique, which 
+Most semantic
 modularization techniques have focused on operations that traverse or 

### 2016-12-28 paper/sections/Introduction.tex
Revising and Splitting Overview

--- 2016-12-19 paper/sections/Introduction.tex
@@
 merge it together when a concrete combination of features is needed
-for a particular program. The typical drawback of such techniques is
-the lack of modular type-checking and separate compilation:
+for a particular program. As Kastner~\cite{} notes, the typical
+drawback of such techniques is that
 ``\emph{most feature-oriented implementation mechanisms lack proper
   interfaces and support neither modular type checking nor separate
-  compilation}''~\cite{kastner11road}\bruno{reference to ``The road to
+  compilation}''\bruno{reference to ``The road to
   Feature Modularity''}. Syntactic modularization techniques have also
@@
 \end{itemize}
+
+\haoyuan{I suggest this part to be moved to Introduction, and only discuss why Packrat is selected among
+different parser combinators in Overview.}
+
+\begin{comment}
+Although there are many parsing techniques, not all of them are
+suitable for type-safe modular parsing. In particular there are many
+techniques which fail to provide modular type-checking and separate
+compilation. Moreover, even if modular type-checking and separate
+compilation are supported, efficiency is another
+concern. A parsing technique should have low overhead when applied
+in a modular setting. In the remaider of this section, we eliminate
+various techniques that fail to satisfy our requirements, and argue
+that that Packrat parsing~\cite{Ford2002} is a suitable candidate for
+type-safe modular parsing.
+
+\paragraph{Parser Generators} The most widely use tools for parsing
+are parser generators. Parser generators help users generate parsers automatically or
+semi-automatically from a given grammar. There is no restriction on
+the algorithm, while most of them adopts table-based LL~\cite{lewis1968syntax} and LR~\cite{knuth1965translation} parsing
+algorithms.
+Although efficient, the main drawback of parser generators is that they do not support
+modular type-checking and separate compilation.
+
+Modular parsing based on parser generators is supported by many libraries~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.
+Users can separate the syntax definition and related parsing code into reusable components.
+Then the corresponding parsers are built by their library utilities.
+For example, NOA~\cite{Gouseti2014} uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers.
+However, such generation procedure requires a whole compilation after the collection of all grammar pieces.
+Once the grammar changed, even slightly, grammar information must be re-collected and the parser must be re-generated.
+Hence those libraries only have syntactic modularity.
+
+Generating parsers often requires full information of grammars, thus semantic modularity is difficult to achieve in this way.
+
+\paragraph{Parser Combinators}
+Comparing with the parser generators, a \textit{parser combinator}~\cite{burge1975,Wadler1985}
+takes several parsers and produce a new parser as its output. Parser combinators are
+popular in functional programming, where the parsers are represented
+by functions and parser combinators are higher-order functions accepts
+them.
+
+At a first look, parser combinators are very suitable for our purpose, because of two
+reasons. Firstly, they are naturally modular. The manner of using them
+is to write small parsers and use combinators to composed them
+together. The construction procedure is explicit and fully controlled
+by the programmer. Secondly, each parser combinator is represented by
+a piece of code, and also are the parsers it takes. Thus in a
+statically typed programming language they can be statically
+type-checked.
+\end{comment}

### 2016-12-28 paper/sections/Overview.tex
Revising and Splitting Overview

@@
-\section{To be discussed in Introduction}
-
-\haoyuan{I suggest this part to be moved to Introduction, and only discuss why Packrat is selected among
-different parser combinators in Overview.}
-
-Although there are many parsing techniques, not all of them are
-suitable for type-safe modular parsing. In particular there are many
-techniques which fail to provide modular type-checking and separate
-compilation. Moreover, even if modular type-checking and separate
-compilation are supported, efficiency is another
-concern. A parsing technique should have low overhead when applied
-in a modular setting. In the remaider of this section, we eliminate
-various techniques that fail to satisfy our requirements, and argue
-that that Packrat parsing~\cite{Ford2002} is a suitable candidate for
-type-safe modular parsing.
-
-\paragraph{Parser Generators} The most widely use tools for parsing
-are parser generators. Parser generators help users generate parsers automatically or
-semi-automatically from a given grammar. There is no restriction on
-the algorithm, while most of them adopts table-based LL~\cite{lewis1968syntax} and LR~\cite{knuth1965translation} parsing
-algorithms.
-Although efficient, the main drawback of parser generators is that they do not support
-modular type-checking and separate compilation.
-
-Modular parsing based on parser generators is supported by many libraries~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.
-Users can separate the syntax definition and related parsing code into reusable components.
-Then the corresponding parsers are built by their library utilities.
-For example, NOA~\cite{Gouseti2014} uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers.
-However, such generation procedure requires a whole compilation after the collection of all grammar pieces.
-Once the grammar changed, even slightly, grammar information must be re-collected and the parser must be re-generated.
-Hence those libraries only have syntactic modularity.
-
-Generating parsers often requires full information of grammars, thus semantic modularity is difficult to achieve in this way.
-
-\paragraph{Parser Combinators}
-Comparing with the parser generators, a \textit{parser combinator}~\cite{burge1975,Wadler1985}
-takes several parsers and produce a new parser as its output. Parser combinators are
-popular in functional programming, where the parsers are represented
-by functions and parser combinators are higher-order functions accepts
-them.
-
-At a first look, parser combinators are very suitable for our purpose, because of two
-reasons. Firstly, they are naturally modular. The manner of using them
-is to write small parsers and use combinators to composed them
-together. The construction procedure is explicit and fully controlled
-by the programmer. Secondly, each parser combinator is represented by
-a piece of code, and also are the parsers it takes. Thus in a
-statically typed programming language they can be statically
-type-checked.
-
 \section{Overview}\label{sec:overview}
@@
 
+\begin{comment}
 \subsection{Choosing the Parsing Technique}\label{subsec:overview-parsing}
@@
 problematic in a modular setting.
-
-\paragraph{Left-Recursion Elimination} The top-down, recursive descent parsing strategy adopted by those parser combinator libraries cannot support left-recursive grammars directly.
-The common solution is to rewrite the grammar into an equivalent but not left-recursive one, so called left-recursion elimination.
-
-A left-recursive grammar of adding integers can be rewritten as below. On the right side, we got a new grammar without left-recursion.
-
-\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> \alt <expr> `+' <int>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> <expr'>
-
-<expr'> ::= <empty> \alt `+' <int> <expr'>
-\end{grammar}
-\end{tabular}
-
-The main problem with left-recursion elimination is that it is a
-\emph{global} transformation on a grammar. Given that the full grammar
-is known, then it is possible to remove all left-recursive
-cases. However, when doing modular parsing, the full grammar is not
-known!
-
-For the example above, if we extend the original grammar to support
-subtraction, we must analyse the full grammar again to rewrite it.
-
-\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> \alt <expr> `+' <int> \alt <expr> `-' <int>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> <expr'>
-
-<expr'> ::= <empty> \alt `+' <int> <expr'> \alt `-' <int> <expr'>
-\end{grammar}
-\end{tabular}
-
-Another issue of left-recursion elimination is that it requires extra
-bookkeeping work to retain the original semantics. For example, the
-expression $1+2-3$ is parsed as $(1+2)-3$ in the left-recursive
-grammar, but after rewrite the result is $1((+2)-3))$. The parse tree
-must be transformed to recover its structure.
-
-\huang{todo}
-\bruno{This example does not appear to be problematic: adding ``-''
-  does not affect the rest of the grammar. Isn't the problem that
-  sometimes adding a new case to the grammar, requires changing other
-parts? Otherwise I don't see why left-recursion is a problem, and we
-cannot simply claim that it is a problem. You have to make a better
-effort to find a problematic grammar.
-}
-
-\paragraph{Longest Match Composition}
-Some parsing libraries provide combinators for alternative composition with longest match,
-while others do not. Now consider the non-left-recursive grammar:
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> \alt <int> `+' <expr>
-\end{grammar}
-In Parsec, for instance, the parser
-\begin{lstlisting}[language=PlainCode]
-parser = parseInt <|> parseAdd
-\end{lstlisting}
-will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
-and terminates parsing. With traditional alternative
-composition in recursive descent parsers, the parser in the front successfully parses a substring of the input, then it unexpectedly
-finishes parsing and discards the rest, in spite that latter parsers might be able to parse the whole input.
-In contrast, \inlinecode{parser = parseAdd <|> parseInt} works as expected, yet manual composition for longest match is quite
-inconvenient, especially in a modular setting. When the grammar is extended with new rules, programmers are supposed to adjust
-the order of parsers, which is tedious.
-
-\paragraph{Backtracking} The need for backtracking is also problematic
-in a modular setting. For example, suppose that we only have \inlinecode{import..from} statements before,
-but now we want to extend the grammar with the \inlinecode{import..as} case, as shown in the third line of grammar below.
-
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<stmt> ::= `import' <ident> `from' <ident>
-    \alt ...
-    \alt `import' <ident> `as' <ident>
-\end{grammar}
-
-Since the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning.
-Take Parsec as an example, its choice
-combinator \inlinecode{<|>} only tries the second alternative if the first fails
-without any token consumption. An auxiliary function \inlinecode{try} is used for explicit backtracking.
-
-\begin{lstlisting}[language=PlainCode]
-oldParser = parseImpFrom <|> parseA <|> parseB <|> ...
-newParser = try parseImpFrom <|> parseA <|> parseB <|> ... <|> parseImpAs
-\end{lstlisting}
-
-Given the full grammar, we can decide where to put \inlinecode{try} for backtracking.
-However, with modular parsing we are unable to overlook the full grammar, hence the worst case should be considered, namely all alternatives may share common prefixes with future cases.
-In that case we need to backtrack for all the branches.
-To avoid failures in the future, we have to add \inlinecode{try} everywhere:
-
-\begin{lstlisting}[language=PlainCode]
-parser = try parseImpFrom <|> try parseA <|> try parseB <|> ... <|> try parseImpAs
-\end{lstlisting}
-which results in the worst exponential time complexity.
-
-%\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
-%You want to say that when you know
-%  the full grammar, you can figure out where ``try'' is
-%  needed. Without the full grammar you'd need assume the worst
-%  case. Make an effort to make your explanation cristal clear!
-%I think you want to miss the abstract explanation that you give first,
-%with the explanation about the concrete example.
-%Start with ``The need for backtracking is also problematic
-%in a modular setting. For example, suppose ...'' and synchronize
-%the abstract explanation and the explanation for the example.
-%}
 
@@
 \name with fancy features or higher efficiency.
-
-\subsection{Packrat Parsing and Our Library \name}\label{subsec:overview-library-api}
-
-Our prototype library is implemented in Scala, with its foundation on Packrat Parsing. In the current version of Scala (2.11), Packrat Parsing is included
-in the Scala standard parser combinator library, which is an external jar file. The library provides a couple of
-parser combinators, together with some basic parsers. Below we present an example to illustrate Scala parsers.
-
-\paragraph{An Example}
-A Packrat parser has type \lstinline{PackratParser[E]} for some
-\lstinline{E}, which indicates the type of results it produces.
-%\bruno{the representation of what?}
-And we use \lstinline{lexical} for its lexing.
-
-\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_EXAMPLE
-\noindent
-Note that the code should be defined in an enclosing type that extends
-\lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers}
-for lexing and
-\lstinline{scala.util.parsing.combinator.PackratParsers}.
-The first two lines are used for lexing. The third line defines a
-parser which can, for instance, parse \lstinline{"str(5)"} and produce
-the integer \lstinline{5}, where \lstinline{numericLit} is just a basic parser
-in the API. Table~\ref{tab:packrat} gives a closer look at part of the API, which covers our experiments.
-
-\begin{table}[t]
-\begin{tabular}{l}
-\hline
-\begin{lstlisting}
-def ~[U](q: => Parser[U]): Parser[~[T, U]]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition. \\
-\hline
-\begin{lstlisting}
-def ^^[U](f: (T) => U): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for function application. \\
-\hline
-\begin{lstlisting}
-def ^^^[U](v: => U): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
-\hline
-\begin{lstlisting}
-def <~[U](q: => Parser[U]): Parser[T]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
-\hline
-\begin{lstlisting}
-def ~>[U](q: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
-\hline
-\begin{lstlisting}
-def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
-\end{lstlisting} \\
-\hspace{.2in}- A parser generator for interleaved repetitions. \\
-\hline
-\begin{lstlisting}
-def ident: Parser[String]
-\end{lstlisting} \\
-\hspace{.2in}- A parser which matches an identifier. \\
-\hline
-\begin{lstlisting}
-def numericLit: Parser[String]
-\end{lstlisting} \\
-\hspace{.2in}- A parser which matches a numeric literal. \\
-\hline
-\begin{lstlisting}
-def |[U >: T](q: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for alternative composition. \\
-\hline
-\begin{lstlisting}
-def |||[U >: T](q0: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for alternative with longest match composition. \\
-\hline \\
-\end{tabular}
-\caption{Part of Scala Parser API.\bruno{We should double-check that this table lists the
-  combinators used in the paper.}}\label{tab:packrat}
-\end{table}
-
-\begin{table}[t]
-\begin{tabular}{l}
-\hline
-\begin{lstlisting}
-type OpenParser[Alg, R, E] = Alg => (=> R) => PackratParser[E]
-\end{lstlisting} \\
-\hspace{.2in}- The type constructor for modular parsers. \\
-\hline
-\begin{lstlisting}
-def parse(p: Parser[_]): String => Unit
-\end{lstlisting} \\
-\hspace{.2in}- A function that applies a parser to an input, and prints the result as a string. \\
-\hline
-\begin{lstlisting}
-implicit def <|>(x: OpenParser[Alg1, R, E], y: OpenParser[Alg2, R, E])
-    : OpenParser[Alg1 with Alg2, R, E]
-\end{lstlisting} \\
-\hspace{.2in}- Open parser combinator for alternative with longest match composition. \\
-\hline \\
-\end{tabular}
-\caption{Our parser combinator library \name API.\haoyuan{I don't know if I can present <|> like this.}}\label{tab:parser}
-\end{table}
-
-Finally, a generic \inlinecode{parse}
-function is used for testing:
-
-\lstinputlisting[linerange=16-19]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_RUNPARSER
-Now \lstinline{parse(p)("str(5)")} produces \lstinline{5} as expected.
+\end{comment}
 

### 2016-12-28 paper/sections/PackratParsing.tex
Revising and Splitting Overview

\section{Packrat Parsing for Modularity}\label{sec:packrat}

This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators based on Packrat Parsing 
are suitable to address those challenges.

\subsection{Algorithmic Challenges of Modularity}
At a first look, parser combinators are very suitable for modular parsing, because of two reasons. Firstly, they are naturally modular. The manner of using them is to write small parsers and use combinators to composed them together. The construction procedure is explicit and fully controlled by the programmer. Secondly, each parser combinator is represented by a piece of code, and also are the parsers it takes. Thus in a statically typed programming language they can be statically type-checked. 
Unfortunately many parser combinators have important limitations.
In particular several parser combinators,
including the famous Parsec~\cite{Leijen2001} library, require
programmers to manually do \textit{left-recursion elimination}, \textit{longest match composition}, and
require significant amounts of \textit{backtracking}. All of them are
problematic in a modular setting.

\paragraph{Left-Recursion Elimination} The top-down, recursive descent parsing strategy adopted by those parser combinator libraries cannot support left-recursive grammars directly. The common solution is to rewrite the grammar into an equivalent but not left-recursive one, so called left-recursion elimination.

A left-recursive grammar for a simple arithmetic expression language can be written in a straightforward way. An alternative grammar, that eliminates left recursion is also presented on the right side:

\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= <int> \alt <expr> `+' <int>
\end{grammar}
&
\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= <int> <expr'>

<expr'> ::= <empty> \alt `+' <int> <expr'>
\end{grammar}
\end{tabular}

The main problem with left-recursion elimination is that it is a
\emph{global} transformation on a grammar. Given that the full grammar
is known, then it is possible to remove all left-recursive
cases. However, when doing modular parsing, the full grammar is not
known!

For the example above, if we extend the original grammar to support
subtraction, we must analyse the full grammar again to rewrite it.

\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= <int> \alt <expr> `+' <int> \alt <expr> `-' <int>
\end{grammar}
&
\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= <int> <expr'>

<expr'> ::= <empty> \alt `+' <int> <expr'> \alt `-' <int> <expr'>
\end{grammar}
\end{tabular}

Another issue of left-recursion elimination is that it requires extra
bookkeeping work to retain the original semantics. For example, the
expression $1+2-3$ is parsed as $(1+2)-3$ in the left-recursive
grammar, but after rewrite the result is $1((+2)-3))$. The parse tree
must be transformed to recover its structure.

\bruno{Huang, please fix this example ASAP.}
\bruno{This example does not appear to be problematic: adding ``-''
  does not affect the rest of the grammar. Isn't the problem that
  sometimes adding a new case to the grammar, requires changing other
parts? Otherwise I don't see why left-recursion is a problem, and we
cannot simply claim that it is a problem. You have to make a better
effort to find a problematic grammar.
}

\paragraph{Longest Match Composition} Another problematic issue 
in parser combinator libraries is the need for manually ordering 
alternatives in a grammar.
Consider the grammar:
\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= <int> \alt <int> `+' <expr>
\end{grammar}
In Parsec, for instance, the parser
\begin{lstlisting}[language=PlainCode]
parser = parseInt <|> parseAdd
\end{lstlisting}
will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
and terminates parsing. With traditional alternative
composition in recursive descent parsers, the parser in the front successfully parses a substring of the input, then it unexpectedly
finishes parsing and discards the rest, in spite that latter parsers might be abl e to parse the whole input.
%\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
%You want to say that when you know
%  the full grammar, you can figure out where ``try'' is
%  needed. Without the full grammar you'd need assume the worst
%  case. Make an effort to make your explanation cristal clear!
%I think you want to miss the abstract explanation that you give first,
%with the explanation about the concrete example.
%Start with ``The need for backtracking is also problematic
%in a modular setting. For example, suppose ...'' and synchronize
%the abstract explanation and the explanation for the example.
%}
In contrast, \inlinecode{parser = parseAdd <|> parseInt} works as expected. 
In this case reordering the components in the alternatives ensures that 
the \emph{longest match} for a string is parsed. Yet manual composition for longest match is inconvenient, and worst still, it is essentially non-modular. When the grammar is extended with new rules, programmers are supposed to \emph{manually} adjust
the order of parsers, which is requires rewritting previously written code.

\paragraph{Backtracking} The need for backtracking is also problematic
in a modular setting. For example, consider a grammar that includes 
import statements such as \inlinecode{import..from},
but now we want to extend the grammar with an \inlinecode{import..as} construct, as shown in the third line of grammar below.

\setlength{\grammarindent}{5em}
\begin{grammar}
<stmt> ::= `import' <ident> `from' <ident>
    \alt ...
    \alt `import' <ident> `as' <ident>
\end{grammar}

Since the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning. Take Parsec as an example, its choice
combinator \inlinecode{<|>} only tries the second alternative if the first fails
without any token consumption. An auxiliary function \inlinecode{try} is used for explicit backtracking.

\begin{lstlisting}[language=PlainCode]
oldParser = parseImpFrom <|> parseA <|> parseB <|> ...
newParser = try parseImpFrom <|> parseA <|> parseB <|> ... <|> parseImpAs
\end{lstlisting}

Given the full grammar, we can decide where to put \inlinecode{try} for backtracking. However, with modular parsing we are unable to have a global view of the full grammar, hence the worst case should be considered. Namely all alternatives may share common prefixes with future cases. In that case we need to backtrack for all the branches. To avoid failures in the future, we have to add \inlinecode{try} everywhere:

\begin{lstlisting}[language=PlainCode]
parser = try parseImpFrom <|> try parseA <|> try parseB <|> ... <|> try parseImpAs
\end{lstlisting}

\noindent Unfortunatelly, this results in the worst exponential time complexity!

\subsection{Packrat Parsing}
Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002}, address the limitations of simple parser combinators such as Parsec. Packrat parsers use memoization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated.  Moreover, theoretically the algorithm behind Packrat parsers supports both direct and indirect left-recursion~\cite{warth2008}. All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in our modular parsing library.
\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}
In the current version of Scala (2.11), Packrat Parsing is included
in the Scala standard parser combinator library, which is an external jar file. The library provides a number of
parser combinators, together with some basic parsers. Below we present an example to illustrate Scala parsers.

\bruno{So, we make a big fuss identifying the problems with other 
parser combinators in the previous subsection and now we present 
an example that does not suffer from any of the previous problems!
What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that 
there's no need for the user to do manual backtracking. Just use 
the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations 
discussed previously applies!
}

\paragraph{An Example}
A Packrat parser has type \lstinline{PackratParser[E]} for some
\lstinline{E}, which indicates the type of results it produces.
%\bruno{the representation of what?}
And we use \lstinline{lexical} for its lexing.

\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_EXAMPLE
\noindent
Note that the code should be defined in an enclosing type that extends
\lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers}
for lexing and
\lstinline{scala.util.parsing.combinator.PackratParsers}.
\bruno{Well, if we need to present the lexing code, then it is better 
to present the code for the enclosing type as well. Instead of 
just saying this in the text.}
The first two lines are used for lexing. The third line defines a
parser which can, for instance, parse \lstinline{"str(5)"} and produce
the integer \lstinline{5}, where \lstinline{numericLit} is just a basic parser
in the API. Table~\ref{tab:packrat} gives a closer look at part of the API that 
we have used in our work.

\begin{table}[t]
\begin{tabular}{l}
\hline
\begin{lstlisting}
def ~[U](q: => Parser[U]): Parser[~[T, U]]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for sequential composition. \\
\hline
\begin{lstlisting}
def ^^[U](f: (T) => U): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for function application. \\
\hline
\begin{lstlisting}
def ^^^[U](v: => U): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
\hline
\begin{lstlisting}
def <~[U](q: => Parser[U]): Parser[T]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
\hline
\begin{lstlisting}
def ~>[U](q: => Parser[U]): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
\hline
\begin{lstlisting}
def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
\end{lstlisting} \\
\hspace{.2in}- A parser generator for interleaved repetitions. \\
\hline
\begin{lstlisting}
def ident: Parser[String]
\end{lstlisting} \\
\hspace{.2in}- A parser which matches an identifier. \\
\hline
\begin{lstlisting}
def numericLit: Parser[String]
\end{lstlisting} \\
\hspace{.2in}- A parser which matches a numeric literal. \\
\hline
\begin{lstlisting}
def |[U >: T](q: => Parser[U]): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for alternative composition. \\
\hline
\begin{lstlisting}
def |||[U >: T](q0: => Parser[U]): Parser[U]
\end{lstlisting} \\
\hspace{.2in}- A parser combinator for alternative with longest match composition. \\
\hline \\
\end{tabular}
\caption{Part of Scala Parser API.\bruno{We should double-check that this table lists the
  combinators used in the paper.}}\label{tab:packrat}
\end{table}

\begin{table}[t]
\begin{tabular}{l}
\hline
\begin{lstlisting}
type OpenParser[Alg, R, E] = Alg => (=> R) => PackratParser[E]
\end{lstlisting} \\
\hspace{.2in}- The type constructor for modular parsers. \\
\hline
\begin{lstlisting}
def parse(p: Parser[_]): String => Unit
\end{lstlisting} \\
\hspace{.2in}- A function that applies a parser to an input, and prints the result as a string. \\
\hline
\begin{lstlisting}
implicit def <|>(x: OpenParser[Alg1, R, E], y: OpenParser[Alg2, R, E])
    : OpenParser[Alg1 with Alg2, R, E]
\end{lstlisting} \\
\hspace{.2in}- Open parser combinator for alternative with longest match composition. \\
\hline \\
\end{tabular}
\caption{Our parser combinator library \name API.\haoyuan{I don't know if I can present <|> like this.}}\label{tab:parser}
\end{table}

Finally, a generic \inlinecode{parse}
function is used for testing:

\lstinputlisting[linerange=16-19]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_RUNPARSER
Now \lstinline{parse(p)("str(5)")} produces \lstinline{5} as expected.

### 2016-12-28 paper/sections/Overview.tex
Partly revised S3

@@
-\section{Overview}\label{sec:overview}
+\section{An Overview of \name}\label{sec:overview}
 
-This section gives an overview of our library \name, and the problem that motivates our work.
+This section gives an overview of our library \name, motivates
+additional problems that occur when parsing modular ASTs.  The first
+problem is that traditional parsers using parser combinators use
+hard-coded recursive calls. This problem is solved using
+delegation-based techniques, which enable parsing OO ASTs.  The second
+problem is how to achieve dual extensibility of ASTs (that is: how to
+allow for both new language constructs and new operations over the
+ASTs). This problem is solved using Object Algebras.
+
+\begin{comment}
 Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions.
@@
 Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.\haoyuan{TODO}
+\end{comment}
 
@@
 
-\paragraph{\name} Our library is constructed by only around twenty lines of code. We have wrapped up the type of parsers
-to make them modular and extensible, and we use our custom combinator \lstinline{<|>} for alternative composition of those
-modular ``open parsers'' with longest match, built on \lstinline{|||}. Table~\ref{tab:parser} shows the API that \name provides for modular parsing.
-
 \subsection{The Parsing Expression Problem}\label{subsec:overview-problem}
@@
 
-In the following text we are going to illustrate the problem that motivates our work.
-Typical parsers defined by parser combinators follow the grammar
-structure, they parse an input to build a corresponding AST. If the grammar is extended
-with a new alternative, the corresponding parser has to be changed as
-well. Thus the traditional style of developing parsers is not
-extensible, and it is at odds with object-oriented programming ASTs,
-which support the easy addition of new alternative classes. We
-illustrate this with a simple expression language and extensions.
+In the following text we are going to illustrate the problems that
+motivate our work via a minimalistic example inspired by Wadler's
+Expression Problem~\cite{}.
 
-Suppose we want to parse a simple language of literals and
+\paragraph{Parsing a simple Arithmetic Language}
+Suppose we want to parse a simple language with literals and
 additions. The concrete syntax is as follows:
+\bruno{Here's a good opportunity to both improve the paper and make it
+more compact! This is the example that should be illustrated and
+discussed at the end of Section 2. Then we can just create a figure to
+refer to the code in Section 2 and refer again to the figure here. No
+need for two different examples.}
 
@@
 
-This time we allow both parts of an addition to be expressions.
-It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases:
-%\huang{This is the first time to demonstrate code in this paper.
-Readers may not know what those combinators/types are.
-Refer to the library docs here?
-move the 'scala parsing library' section to the beginning as 'background'?}
+\noindent It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases:
+%\huang{This is the first time to demonstrate code in this
+%paper. Readers may not know what those combinators/types are. Refer
+%to the library docs here? move the 'scala parsing library' section to
+%the beginning as 'background'?}
 
@@
 
-\noindent Note that we use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} in the rest of paper.
-As the combination of \inlinecode{pVar}, \inlinecode{pApp}, \inlinecode{pExpr} can parse expressions like \lstinline{"1+2"} correctly.
+\noindent Note that we use \lstinline{Parser} as a type synonym for
+\lstinline{PackratParser} in the rest of paper. 
+The parser \inlinecode{pExpr} is able to parse 
+expressions like \lstinline{"1+2"} correctly.
+
+\paragraph{Dimensions of Extensibility}
 At this point, one would like such code to be extensible in two dimensions, for example:
 \begin{itemize}
-\item \textbf{Challenge 1}: adding a new case (or rather, a new construct) to \lstinline{Expr};
-\item \textbf{Challenge 2}: adding a new operation to \lstinline{Expr}.
+\item \textbf{Extension 1}: adding a new case (or rather, a new construct) to \lstinline{Expr};
+\item \textbf{Extension 2}: adding a new operation to \lstinline{Expr}.
 \end{itemize}
 
-\paragraph{Attempt to Extend the Parser} Now we firstly consider challenge 1, namely we are extending the syntax as well as the parser. Hence we introduce variables as a new case.
+\paragraph{Attempt to Extend the Parser} Now we first consider 
+extension 1, namely we are extending the syntax as well as the parser. Hence we introduce variables as a new case.
 
@@
 
-The corresponding AST is extended as below.
+As in an OO solution to the Expression Problem, it is easy to extend
+the corresponding AST:
 
@@
 
-Since we already have the parser for literals and additions, we would like to build the new parser by reusing the old one.
-So we compose \inlinecode{pVar}, which parses variables, with the old parser.
-\inlinecode{pExtExpr} is the new function to parse expressions.
+Unfortunatelly, extending the parser for the AST is not as easy.
+Since we already have the parser for literals and additions, we would
+like to build the new parser by reusing the old one. So we compose
+\inlinecode{pVar}, which parses variables, with the old
+parser. \inlinecode{pExtExpr} is the new function to parse
+expressions.
 
@@
 
-Unfortunately, this extended parser does not work properly to parse all possible expressions defined by the new grammar. For example, it fails to parse \inlinecode{"1 + x"} which is obviously valid.
+\noindent This extended parser does not work properly to parse all possible expressions defined by the new grammar. For example, it fails to parse \inlinecode{"1 + x"} which is obviously valid.
 
-\paragraph{Problem: Hard-coded Recursive Calls} The critical problem of extending parsers directly is that the recursive calls in old parsers are hard-coded.
-Specifically, \inlinecode{pAdd} makes two recursive calls to parse two sub-expressions around \lstinline{"+"}.
-Before the extension, we implemented it by using \inlinecode{pExpr} for those recursive calls.
-It worked fine at that time because \inlinecode{pExpr} covers all cases of the old grammar.
-However, after the extension of variables, the proper function for such recursive calls should be \inlinecode{pExtExpr} instead of \inlinecode{pExpr}, because the new function \inlinecode{pExtExpr} includes the new extension.
+\paragraph{Problem: Hard-coded Recursive Calls} 
+Typical parsers defined by parser combinators follow the grammar
+structure, they parse an input to build a corresponding AST. If the grammar is extended
+with a new alternative, the corresponding parser has to be changed as
+well. Thus the traditional style of developing parsers is not
+extensible, and it is at odds with object-oriented programming ASTs,
+which support the easy addition of new alternative classes.
+The critical problem
+of extending parsers directly is that the recursive calls in old
+parsers are hard-coded. Specifically, \inlinecode{pAdd} makes two
+recursive calls to parse two sub-expressions around
+\lstinline{"+"}. Before the extension, we implemented it by using
+\inlinecode{pExpr} for those recursive calls. It worked fine at that
+time because \inlinecode{pExpr} covers all cases of the old
+grammar. However, after the extension of variables, the proper
+function for such recursive calls should be \inlinecode{pExtExpr}
+instead of \inlinecode{pExpr}, because the new function
+\inlinecode{pExtExpr} includes the new extension.
 
-That is the reason why the extended parser fails to parse some valid expressions such as \inlinecode{"1 + x"}.
-To resolve this issue, \inlinecode{pAdd} needs to be aware of the change of grammar.
-There are some possible solutions:
+That is the reason why the extended parser fails to parse some valid
+expressions such as \inlinecode{"1 + x"}. To resolve this issue,
+\inlinecode{pAdd} needs to be aware of the change of grammar. There
+are some possible solutions:
 
@@
 In the next sections, we will discuss them with more details.
+
+\bruno{Put the following text somewhere?}
+\paragraph{\name} Our library is constructed by only around twenty lines of code. We have wrapped up the type of parsers
+to make them modular and extensible, and we use our custom combinator \lstinline{<|>} for alternative composition of those
+modular ``open parsers'' with longest match, built on \lstinline{|||}. Table~\ref{tab:parser} shows the API that \name provides for modular parsing.

### 2016-12-29 paper/sections/Inheritance.tex
Added S3, inheritance-based approach.

--- 2016-12-28 paper/sections/Overview.tex
@@
-\section{An Overview of \name}\label{sec:overview}
+\section{Modularizing Parsers with Inheritance}\label{sec:inheritance}
 
-This section gives an overview of our library \name, motivates
-additional problems that occur when parsing modular ASTs.  The first
-problem is that traditional parsers using parser combinators use
-hard-coded recursive calls. This problem is solved using
-delegation-based techniques, which enable parsing OO ASTs.  The second
-problem is how to achieve dual extensibility of ASTs (that is: how to
-allow for both new language constructs and new operations over the
-ASTs). This problem is solved using Object Algebras.
+This section introduces the problem that motivates our work. In Figure~\ref{fig:packrat-arith} we have observed
+that Packrat parsers can be defined in a convenient way, with recursive calls. The example is actually inspired
+by Walder's Expression Problem. Similarly, our goal is to build parsers that are extensible in some dimensions.
+
+\subsection{Dimensions of Extensibility}
+We continue with the expression language of literals and additions. At this point,
+one would like to have two kinds of extensions:
+
+%\section{An Overview of \name}\label{sec:overview}
+%
+%This section gives an overview of our library \name, motivates
+%additional problems that occur when parsing modular ASTs.  The first
+%problem is that traditional parsers using parser combinators use
+%hard-coded recursive calls. This problem is solved using
+%delegation-based techniques, which enable parsing OO ASTs.  The second
+%problem is how to achieve dual extensibility of ASTs (that is: how to
+%allow for both new language constructs and new operations over the
+%ASTs). This problem is solved using Object Algebras.
 
@@
 
-\subsection{The Parsing Expression Problem}\label{subsec:overview-problem}
 %\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
@@
 
-In the following text we are going to illustrate the problems that
-motivate our work via a minimalistic example inspired by Wadler's
-Expression Problem~\cite{}.
-
-\paragraph{Parsing a simple Arithmetic Language}
-Suppose we want to parse a simple language with literals and
-additions. The concrete syntax is as follows:
-\bruno{Here's a good opportunity to both improve the paper and make it
-more compact! This is the example that should be illustrated and
-discussed at the end of Section 2. Then we can just create a figure to
-refer to the code in Section 2 and refer again to the figure here. No
-need for two different examples.}
-
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int>
-    \alt <expr> `+' <expr>
-\end{grammar}
-%\alt `(' <expr> `)'
-
-\noindent It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases:
-%\huang{This is the first time to demonstrate code in this
-%paper. Readers may not know what those combinators/types are. Refer
-%to the library docs here? move the 'scala parsing library' section to
-%the beginning as 'background'?}
-
-\lstinputlisting[linerange=8-24]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
-
-\noindent Note that we use \lstinline{Parser} as a type synonym for
-\lstinline{PackratParser} in the rest of paper. 
-The parser \inlinecode{pExpr} is able to parse 
-expressions like \lstinline{"1+2"} correctly.
-
-\paragraph{Dimensions of Extensibility}
-At this point, one would like such code to be extensible in two dimensions, for example:
 \begin{itemize}
@@
 
-\lstinputlisting[linerange=28-30]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
+\lstinputlisting[linerange=95-97]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
 
-Unfortunatelly, extending the parser for the AST is not as easy.
-Since we already have the parser for literals and additions, we would
-like to build the new parser by reusing the old one. So we compose
-\inlinecode{pVar}, which parses variables, with the old
-parser. \inlinecode{pExtExpr} is the new function to parse
-expressions.
+And since we already have the parser for literals and additions, we would
+like to build the new parser by reusing the old one. Here one may quickly come up
+with the following attempt, where a new parser is defined for \lstinline{Var}, then composed
+with \lstinline{pExpr}:
 
-\lstinputlisting[linerange=34-37]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXT
+\lstinputlisting[linerange=62-63]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=BAD_ATTEMPT
 
-\noindent This extended parser does not work properly to parse all possible expressions defined by the new grammar. For example, it fails to parse \inlinecode{"1 + x"} which is obviously valid.
+Unfortunately, it fails to parse some expressions like \lstinline{"1 + x"}, which is obviously valid in the new grammar.
+The reason is that \lstinline{pAdd} makes two recursive calls to parse sub-expressions, by using \lstinline{pExpr}, which
+covers both cases in the old grammar. But the newly added case \lstinline{pVar} is not observed by the recursive \lstinline{pExpr},
+hence the parser does not work as expected. Still, one may continue to change the recursive call in \lstinline{pAdd} with \lstinline{pExtExpr},
+but modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.
 
-\paragraph{Problem: Hard-coded Recursive Calls} 
-Typical parsers defined by parser combinators follow the grammar
-structure, they parse an input to build a corresponding AST. If the grammar is extended
-with a new alternative, the corresponding parser has to be changed as
-well. Thus the traditional style of developing parsers is not
-extensible, and it is at odds with object-oriented programming ASTs,
-which support the easy addition of new alternative classes.
-The critical problem
-of extending parsers directly is that the recursive calls in old
-parsers are hard-coded. Specifically, \inlinecode{pAdd} makes two
-recursive calls to parse two sub-expressions around
-\lstinline{"+"}. Before the extension, we implemented it by using
-\inlinecode{pExpr} for those recursive calls. It worked fine at that
-time because \inlinecode{pExpr} covers all cases of the old
-grammar. However, after the extension of variables, the proper
-function for such recursive calls should be \inlinecode{pExtExpr}
-instead of \inlinecode{pExpr}, because the new function
-\inlinecode{pExtExpr} includes the new extension.
+\subsection{Inheritance-Based Approach}\label{subsec:inheritance-approach}
 
-That is the reason why the extended parser fails to parse some valid
-expressions such as \inlinecode{"1 + x"}. To resolve this issue,
-\inlinecode{pAdd} needs to be aware of the change of grammar. There
-are some possible solutions:
+Fortunately in Scala we can make use of inheritance for modular parsing. Specifically,
+the new parser can be defined in an enclosing trait that extends
+the old one:
 
-\begin{itemize}
-    \item Keep the old code, rewrite the non-extensible function \inlinecode{pAdd}. \\Separate compilation is obtained but code reuse is lost, because every function may contain such recursive calls.
-    \item Modify the old code, correct the recursive calls by proper functions.
-\\We can avoid duplication for code reuse, but separate compilation is sacrificed because of modification on the existing code.
-    \item ??? \huang{third one? use dynamic languages?}
-\end{itemize}
+\lstinputlisting[linerange=68-73]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=INHERITANCE_APPROACH
+Now \lstinline{ExtParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
+method overriding in inheritance. When the input \lstinline{"1 + x"} is fed to the parser \lstinline{this.pExpr}, it firstly delegates
+the work to \lstinline{super.pExpr}, which parses literals and additions. However, the recursive call \lstinline{pExpr} in \lstinline{pAdd}
+actually refers to \lstinline{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
+elegant inheritance-based approach opens the recursion of parsers in a modular way.
 
-Because of the trade-off, neither of them are satisfying.
-The challenge of building semantically modular parsers requires fundamental change of the parser structure.
-We will introduce our solution in the next sections.
+Notice that Scala has a linearized-style multiple inheritance for traits. It can be very helpful when composing several languages. Suppose now
+we want to compose the parsers for expressions from pre-defined languages \lstinline{LanguageA} and \lstinline{LanguageB} using alternative.
+The use of keyword \lstinline{super} is able to
+specify the implementation from inheritance, like:
 
-\subsection{Delegation-Based Parsing}\label{subsec:overview-delegation}
+\lstinputlisting[linerange=78-84]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 
-Hard-coded recursive calls prevent parsers from extensions.
-This problem can be solved using delegation, which can be encoded with
-\textit{open recursion}. The idea is simple: instead of making direct
-recursive calls, an additional argument, that abstracts over the
-recursive call, is passed to the parsers. This technique enables the
-development of parsers that support OO-style extensibility of ASTs.
-
-\paragraph{Open Parsing} Consider the expression language of only literals and additions in the previous section. The parser can be rewritten as follows:
-
-\lstinputlisting[linerange=58-68]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_BASE
-
-According to the definition of type synonym \inlinecode{Open[T]},
-every parsing function takes a parser of \inlinecode{Expr} as a
-parameter, namely \inlinecode{self}. This parameter \inlinecode{self}
-represents \emph{the parser} of expressions, which is dynamiclly
-explained by the argument. Using such an abstract ``self-reference'' provides
-extensibility to parsers, because the recursive calls are not
-hard-coded here --- they are dynamic and can be changed corresponding
-to the grammar!
-We take \inlinecode{pAdd} as an example again. It uses
-\inlinecode{self} instead of hard-coded \inlinecode{pExpr} for
-recursive parsing. Depending on the real argument, \inlinecode{self}
-can be a parser which covers only literals and additions, or a
-parser which also supports variables.
-
-\paragraph{Client Code} Since the type of parsing functions is
-changed, we can not apply them directly. A \inlinecode{fix} function, which is the
-the standard \textit{fixpoint combinator}, is used to `close' the open
-recursion. After that, it can be used as an ordinary parser.
-The client code below demonstrates the usage of new parsers.
-
-%\huang{done}\bruno{Too much detail is being given here about open
-%  recursion. In Section 2 we should talk only about \emph{how} to use
-%open recursion, and not how open recursion is implemented. We
-%can summarize, for example the API related to open recursion.
-%Section 4 is where the implementation and detailed explanation
-%about open recursion should come.
-%}
-
-\lstinputlisting[linerange=73-77]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_USE
-
-%\huang{done}\bruno{Please polish up code. You probably want to consistently use
-%the variable name ``self'' whenever you want to refer to the whole
-%parser (instead of ``e'' that you use at the moment). Also you should
-%use ``Open'' (which you currently call ``Fix'') instead of using the
-%longwinded ``(=> PackratParser[Expr]) => PackratParser[Expr]''}
-%
-%\huang{done}\bruno{Also, for consistency, why isn't ``pVar'' open? I.e. using a
-%  self reference as well? I know that it is not recursive, but giving
-%  it a different treatment makes it more confusing for users.}
-%
-%\huang{done}\bruno{You should use a more OO approach in the example.
-%Rather than case classes show an Expr trait with pretty printing.
-%This will enable you to talk about the extensibility and object
-%algebras better later on.
-%}
-
-\paragraph{Extensibility} Similarly, we extend the grammar by adding variables as a new case.
-The new parser also extends the old one by adding \inlinecode{pVar} on it.
-Notice that the old code is not touched, hence separate compilation is obtained.
-
-\lstinputlisting[linerange=82-85]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT
-
-\haoyuan{Code without combinators?}
-
-All recursive calls of parsing `an expression' are realized via the explicit parameter \inlinecode{self}.
-Once the \inlinecode{fix} function is applied, these calls will be updated appropriately.
-Parsing functions such as \inlinecode{pAdd} do not need to be rewritten.
-As a consequence, this parser recognizes all valid expressions of the new grammar, including \inlinecode{"1 + x"} which cannot be parsed previously.
-
-\lstinputlisting[linerange=91-91]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_OPEN_EXT_USE
-
-As we demonstrated, delegation encoded by open recursion is the key technique to obtain semantic modularity.
-It enables type-safe code reuse and separate compilation in our parsers.
-We will discuss this topic further in Section \ref{sec:openandparsing}.
+As we demonstrated, inheritance is the key technique to obtain semantic modularity.
+It enables type-safe code reuse and separate compilation for parsing OO style ASTs with extension 1.
 
@@
 
-\lstinputlisting[linerange=111-111]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_ATTEMPT_EXPRWITHEVAL
+\lstinputlisting[linerange=178-178]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_ATTEMPT_EXPRWITHEVAL
 But then the programmer has to define new classes for literals and additions that extend old ones and \lstinline{EvalExpr}. Specifically,
@@
 In the next sections, we will discuss them with more details.
-
-\bruno{Put the following text somewhere?}
-\paragraph{\name} Our library is constructed by only around twenty lines of code. We have wrapped up the type of parsers
-to make them modular and extensible, and we use our custom combinator \lstinline{<|>} for alternative composition of those
-modular ``open parsers'' with longest match, built on \lstinline{|||}. Table~\ref{tab:parser} shows the API that \name provides for modular parsing.

### 2016-12-29 paper/sections/OpenRecursion.tex
Added S3, inheritance-based approach.

@@
 
-\lstinputlisting[linerange=31-37]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
+\lstinputlisting[linerange=24-30]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
 
@@
 
-\lstinputlisting[linerange=41-47]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
+\lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
 \bruno{Fix should be replaces by Open. Please coordinate with Huang to
@@
 
-\lstinputlisting[linerange=51-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB_FIX
+\lstinputlisting[linerange=44-44]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB_FIX
 
@@
 
-\lstinputlisting[linerange=27-27]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
+\lstinputlisting[linerange=20-20]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
 \bruno{can be defined in a single line.}
@@
 Suppose now we want to make use of open recursion for additional operations. We first define a combinator which combines the results of two generators in a pair:
-\lstinputlisting[linerange=55-58]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_COMPOSE
+\lstinputlisting[linerange=48-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_COMPOSE
 
@@
 Then we define a generator which prints out a number as a string:
-\lstinputlisting[linerange=62-62]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_SHOW
+\lstinputlisting[linerange=55-55]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_SHOW
 
@@
 The definition is straightforward:
-\lstinputlisting[linerange=66-67]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_MERGE
+\lstinputlisting[linerange=59-60]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_MERGE
 

### 2016-12-29 paper/sections/PackratParsing.tex
Added S3, inheritance-based approach.

@@
 
-This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators based on Packrat Parsing 
+This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators from Packrat Parsing
 are suitable to address those challenges.
@@
 In the current version of Scala (2.11), Packrat Parsing is included
-in the Scala standard parser combinator library, which is an external jar file. The library provides a number of
+in the Scala standard parser combinator library. The library provides a number of
 parser combinators, together with some basic parsers. Below we present an example to illustrate Scala parsers.
+%
+%\bruno{So, we make a big fuss identifying the problems with other
+%parser combinators in the previous subsection and now we present
+%an example that does not suffer from any of the previous problems!
+%What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that
+%there's no need for the user to do manual backtracking. Just use
+%the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations
+%discussed previously applies!
+%}
 
-\bruno{So, we make a big fuss identifying the problems with other 
-parser combinators in the previous subsection and now we present 
-an example that does not suffer from any of the previous problems!
-What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that 
-there's no need for the user to do manual backtracking. Just use 
-the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations 
-discussed previously applies!
-}
+\paragraph{Parsing a simple Arithmetic Language}
+Suppose we want to parse a simple language with literals and
+additions. The concrete syntax is as follows:
+%\bruno{Here's a good opportunity to both improve the paper and make it
+%more compact! This is the example that should be illustrated and
+%discussed at the end of Section 2. Then we can just create a figure to
+%refer to the code in Section 2 and refer again to the figure here. No
+%need for two different examples.}
 
-\paragraph{An Example}
-A Packrat parser has type \lstinline{PackratParser[E]} for some
-\lstinline{E}, which indicates the type of results it produces.
-%\bruno{the representation of what?}
-And we use \lstinline{lexical} for its lexing.
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int>
+    \alt <expr> `+' <expr>
+\end{grammar}
+%\alt `(' <expr> `)'
 
-\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_EXAMPLE
-\noindent
-Note that the code should be defined in an enclosing type that extends
-\lstinline{scala.util.parsing.combinator.syntactical.StandardTokenParsers}
-for lexing and
-\lstinline{scala.util.parsing.combinator.PackratParsers}.
-\bruno{Well, if we need to present the lexing code, then it is better 
-to present the code for the enclosing type as well. Instead of 
-just saying this in the text.}
-The first two lines are used for lexing. The third line defines a
-parser which can, for instance, parse \lstinline{"str(5)"} and produce
-the integer \lstinline{5}, where \lstinline{numericLit} is just a basic parser
-in the API. Table~\ref{tab:packrat} gives a closer look at part of the API that 
-we have used in our work.
+It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases.
+Note that a parser has type \lstinline{PackratParser[E]} for some
+\lstinline{E}, which indicates the type of results it produces. Figure~\ref{fig:packrat-arith} shows the abstract syntax and
+the parser code that generates expressions as ASTs.
+
+\begin{figure}[t]
+\centering
+\lstinputlisting[linerange=7-27]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
+\caption{Packrat Parsing for an arithmetic language.}\label{fig:packrat-arith}
+\end{figure}
+
+In the trait \lstinline{BaseParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} can parse an integer for \lstinline{Lit},
+whereas \lstinline{pAdd} handles addition and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
+recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the alternative combinator with longest match.
+Table~\ref{tab:packrat} gives a closer look at part of the API that we have used in our work.
 
@@
 
-\begin{table}[t]
-\begin{tabular}{l}
-\hline
-\begin{lstlisting}
-type OpenParser[Alg, R, E] = Alg => (=> R) => PackratParser[E]
-\end{lstlisting} \\
-\hspace{.2in}- The type constructor for modular parsers. \\
-\hline
-\begin{lstlisting}
-def parse(p: Parser[_]): String => Unit
-\end{lstlisting} \\
-\hspace{.2in}- A function that applies a parser to an input, and prints the result as a string. \\
-\hline
-\begin{lstlisting}
-implicit def <|>(x: OpenParser[Alg1, R, E], y: OpenParser[Alg2, R, E])
-    : OpenParser[Alg1 with Alg2, R, E]
-\end{lstlisting} \\
-\hspace{.2in}- Open parser combinator for alternative with longest match composition. \\
-\hline \\
-\end{tabular}
-\caption{Our parser combinator library \name API.\haoyuan{I don't know if I can present <|> like this.}}\label{tab:parser}
-\end{table}
+Finally, a generic \inlinecode{parse} function is used for testing:
 
-Finally, a generic \inlinecode{parse}
-function is used for testing:
+\lstinputlisting[linerange=52-57]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=PACKRAT_RUNPARSER
 
-\lstinputlisting[linerange=16-19]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=PACKRAT_RUNPARSER
-Now \lstinline{parse(p)("str(5)")} produces \lstinline{5} as expected.
+It is worth mentioning that such a grammar is left-recursive, yet well supported by Packrat parsers. It also employs automatic
+longest match composition by using the library function \lstinline{|||}. Finally, Packrat parsers do not suffer from the backtracking
+problem, as the technique uses memoization to record intermediate parsing results. We will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} in the rest of paper.

### 2016-12-29 Scala/Parser/src/PaperCode/SampleParser.scala

@@ -11,13 +11,6 @@ lexical.delimiters += ("(", ")")
 val p: PackratParser[Int] =
   "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toInt }
 //END_PACKRAT_EXAMPLE
-
-//BEGIN_PACKRAT_RUNPARSER
-def parse[E](p: Parser[E]): String => E = in => {
-  val t = phrase(p)(new lexical.Scanner(in))
-  if (t.successful) t.get else scala.sys.error(t.toString)
-}
-//END_PACKRAT_RUNPARSER  
      
 }
 

### 2016-12-29 paper/sections/Inheritance.tex
First part of Object Algebras.

@@
 
-\subsection{Dimensions of Extensibility}
+\subsection{Dimensions of Extensibility}\label{subsec:dimensions}
 We continue with the expression language of literals and additions. At this point,
@@
 
-\lstinputlisting[linerange=95-97]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
+\lstinputlisting[linerange=98-100]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
 
@@
 
-Although delegation enables OO extensibility, the use of an OO class
-hierarchy makes the addition of new operations over the AST
-problematic. Modification on existing code breaks the modularity. This
-refers to the famous \textit{Expression Problem} []. For challenge 2, suppose now we want to
-support evaluation on expressions, in Scala one attempt would be extending \lstinline{Expr} with
-evaluation:
-
-\lstinputlisting[linerange=178-178]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_ATTEMPT_EXPRWITHEVAL
-But then the programmer has to define new classes for literals and additions that extend old ones and \lstinline{EvalExpr}. Specifically,
-the critical point is that existing parsing code only generates \lstinline{Expr} objects with pretty printing, so modification on parsing is again required.
-
-In contrast, Object Algebras~\cite{Oliveira2012} enable us to solve this problem, as it separates data variants and
-operations, to
-offer high flexibility in the choice of operations to be performed
-over the AST. It also makes parsing with multiple sorts of syntax easier.
-
-\paragraph{Parsing with Object Algebras} Using Object Algebras, the abstract syntax of the expression language is defined as below.
-
-\lstinputlisting[linerange=8-11]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALG
-
-Then we are able to define operations over the syntax in a modular way. For instance, pretty printing operation can be realized as:
-
-\lstinputlisting[linerange=15-18]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
-
 And also parser for it as below.
@@
 
-\lstinputlisting[linerange=22-27]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PARSER
+\lstinputlisting[linerange=29-34]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PARSER
 
@@
 
-\lstinputlisting[linerange=31-44]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
+\lstinputlisting[linerange=38-51]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
 Note that we use the new combinator \lstinline{<|>} defined in our library, to compose two parsing functions into one.
@@
 
-\lstinputlisting[linerange=53-58]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
+\lstinputlisting[linerange=60-65]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
 
@@
 
-\lstinputlisting[linerange=72-78]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
+\lstinputlisting[linerange=79-85]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
 
@@
 
-\lstinputlisting[linerange=63-67]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=70-74]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 In the next sections, we will discuss them with more details.

### 2016-12-29 paper/sections/ObjectAlgebras.tex
Added S3, inheritance-based approach.
First part of Object Algebras.

@@
-\section{Object Algebras and Parsing}\label{sec:algebrasandparsing}
+\section{Extensible ASTs with Object Algebras}\label{sec:algebrasandparsing}
 
-Object Algebras~\cite{Oliveira2012} is a solution to the famous \textit{Expression Problem}, provide extensibility on
-both data variants and operations. It is suitable for structures like abstract syntax trees. We integrate Object Algebras directly to enhance the extensibility of \name.
+The inheritance-based approach benefits us from building extensible parsers, based on an OO class hierarchy.
+Nevertheless, the addition of new operations over the AST becomes problematic. Often it requires modification on
+the existing code, which breaks modularity. In this section, we will have a closer look at how it motivates us to
+use Object Algebras for full extensibility.
+
+\subsection{Problem with Traditional OO ASTs}\label{subsec:problemwithoutoa}
+
+Now it comes to extension 2 in Section~\ref{subsec:dimensions}. In the previous code, we have presented pretty-printing as
+an example of operations on the AST. Suppose now we also want to collect free variables on the expression language, which models
+literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \lstinline{Expr} with the new
+operation directly:
+
+\lstinputlisting[linerange=89-89]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
+It refers to one Scala solution to the Expression Problem []. But then programmers have to define new classes for literals, additions and variables
+by extending the old ones together with \lstinline{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
+types, again it is a modularity issue, since it is a bad idea to modify existing code again and again with extensions.
+Other solutions including modelling constructors as functions introduce unnecessary complexity.
+
+In contrast, Object Algebras~\cite{Oliveira2012} enables us to solve this problem. As a nice solution to the Expression Problem, it separates data variants
+and operations, and offers high flexibility in the choice of operations to be performed over the AST. Before adopting it for parsing, the next section introduces the background of Object Algebras.
 
@@
 In Object Algebras, ASTs as recursive data structures are defined using traits, where each constructor corresponds
-to an abstract method inside. The example from Section~\ref{subsec:parsingwithopen} is again used here for illustration.
-At first the language only supports variables and applications:
+to an abstract method inside. The example from Section~\ref{subsec:packratparsing} is again used here for illustration.
+At first the language only supports literals and additions:
 
-\lstinputlisting[linerange=4-8]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_EXPR
+\lstinputlisting[linerange=8-11]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALG
 
@@
 type parameter \lstinline{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
-provides implementations for all cases. Below is an example of collecting all free variables in an expression:
+provides implementations for all cases. Below is an example of pretty-printing an expression:
 
-\lstinputlisting[linerange=12-15]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_FREE_VARS
+\lstinputlisting[linerange=15-18]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
 
-Here \lstinline{FreeVars} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a list of strings as the result.
-Furthermore, one can define a new trait to implement pretty-printing on expressions:
+Here \lstinline{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
+Furthermore, one can define a new trait to implement evaluation on arithmetic expressions:
 
-\lstinputlisting[linerange=19-22]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_EXPR_PRINT
+\lstinputlisting[linerange=18-21]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_EXPR_PRINT
 
-On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Suppose we want to
-have literals in expressions, a new Object Algebra interface \lstinline{LitAlg} is defined as follows:
+On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Now we extend the language
+with variables. A new Object Algebra interface \lstinline{VarAlg} is defined as follows:\haoyuan{Updates till here.}
 
-\lstinputlisting[linerange=26-28]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT
+\lstinputlisting[linerange=25-27]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT
 
@@
 
-\lstinputlisting[linerange=32-34]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT_PRINT
+\lstinputlisting[linerange=31-33]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT_PRINT
 
@@
 \huang{todo: better name for build}
-\lstinputlisting[linerange=48-48]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_BUILD
+\lstinputlisting[linerange=47-47]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_BUILD
 
@@
 
-\lstinputlisting[linerange=38-42]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_REFACTOR
+\lstinputlisting[linerange=37-41]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_REFACTOR
 

### 2016-12-29 paper/sections/PackratParsing.tex
First part of Object Algebras.

@@
 
-\subsection{Packrat Parsing}
+\subsection{Packrat Parsing}\label{subsec:packratparsing}
 Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002}, address the limitations of simple parser combinators such as Parsec.
@@
 \lstinline{E}, which indicates the type of results it produces. Figure~\ref{fig:packrat-arith} shows the abstract syntax and
-the parser code that generates expressions as ASTs.
+the parser code that generates expressions as ASTs. Meanwhile, pretty-printing shows up as an operation on the AST.
 

### 2016-12-29 Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala

@@ -2,10 +2,9 @@ package PaperCode.ObjectAlg
 
 //BEGIN_OBJECT_ALG_EXPR
 trait ExprAlg[E] {
-  def Var(x: String): E
-  def App(e1: E, e2: E): E
+  def lit(n: Int): E
+  def add(e1: E, e2: E): E
 }
-
 //END_OBJECT_ALG_EXPR
 
 //BEGIN_OBJECT_ALG_FREE_VARS

### 2016-12-29 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -85,6 +85,9 @@ trait LanguageC extends LanguageA with LanguageB {
 //END_MULTIPLE_INHERITANCE
 */
 
+//BEGIN_ATTEMPT_EXPRWITHFREEVARS
+trait FreeVarsExpr extends Expr { def freeVars: Set[String] }
+//END_ATTEMPT_EXPRWITHFREEVARS
 }
 
 object Overview2 {

### 2017-01-01 paper/sections/Inheritance.tex
Revised S4.

@@
 covers both cases in the old grammar. But the newly added case \lstinline{pVar} is not observed by the recursive \lstinline{pExpr},
-hence the parser does not work as expected. Still, one may continue to change the recursive call in \lstinline{pAdd} with \lstinline{pExtExpr},
+hence the parser does not work as expected. Still, one may continue to change the recursive call in \lstinline{pAdd} with \lstinline{pVarExpr},
 but modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.
@@
 \lstinputlisting[linerange=68-73]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=INHERITANCE_APPROACH
-Now \lstinline{ExtParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
+Now \lstinline{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
 method overriding in inheritance. When the input \lstinline{"1 + x"} is fed to the parser \lstinline{this.pExpr}, it firstly delegates
@@
 It enables type-safe code reuse and separate compilation for parsing OO style ASTs with extension 1.
-
-\subsection{Object Algebras for full Extensibility}\label{subsec:overview-oa}
-
-And also parser for it as below.
-Notice the parsing function \inlinecode{pExpr} now takes an argument of type \inlinecode{ExprAlg[E]}, which means it accepts any instance of \inlinecode{ExprAlg}.
-Such an instance can for example be \lstinline{Print}, which takes charge of operations on expressions, and parsing code only does delegation.
-
-\lstinputlisting[linerange=29-34]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PARSER
-
-\haoyuan{I'm trying to avoid using terminology like algebra and algebra interface.}
-
-\paragraph{Extensibility of Syntax} Following previous examples, the code below shows how to extend the language with variables.
-
-\lstinputlisting[linerange=38-51]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT
-Note that we use the new combinator \lstinline{<|>} defined in our library, to compose two parsing functions into one.
-The function \lstinline{pExtExpr} has the compound type \lstinline{ExprAlg[E] with VarAlg[E]} in its argument, where two language ASTs are combined to represent the new language.
-
-To use the parser, we must provide an algebra instance as the operation to construct the parsing results.
-In the code below, we use the combination of \lstinline{Print} and \lstinline{PrintVar}, so that the parsing result is a pretty printing of the AST.
-
-\lstinputlisting[linerange=60-65]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_USE
-
-\paragraph{Extensibility of Operations} With Object Algebras, operations over ASTs can also be extended in a modular way.
-Here is an example of collecting free variables from an expression.
-We can feed this operation to the parser and obtain a set of free variables.
-
-\lstinputlisting[linerange=79-85]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXT_OP
-
-\haoyuan{I think in our library language ASTs are composed using intersection, not inheritance.}
-
-\paragraph{Multiple Sorts of Syntax} Another advantage of using Object Algebras is that it supports multiple sorts of syntax easily.
-In several cases, we want to divide the syntactic elements into some groups.
-For example, the grammar below has two sorts, which are expressions and types.
-
-\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<type> ::= `int' \alt <type> `->' <type>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::=  `\\' <ident> `:' <type> `.' <expr>
-\end{grammar}
-\end{tabular}
-
-Using Object Algebras, we can easily distinguish them just by adding an extra type parameter. The abstract syntax is
-presented below:
-
-\lstinputlisting[linerange=70-74]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
-In the next sections, we will discuss them with more details.

### 2017-01-01 paper/sections/ObjectAlgebras.tex
Revised S4.

@@
 \lstinputlisting[linerange=8-11]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALG
-
 The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
@@
 \lstinputlisting[linerange=15-18]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
-
 Here \lstinline{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
@@
 
-\lstinputlisting[linerange=18-21]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_EXPR_PRINT
+\lstinputlisting[linerange=22-25]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
 On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Now we extend the language
-with variables. A new Object Algebra interface \lstinline{VarAlg} is defined as follows:\haoyuan{Updates till here.}
+with variables. A new Object Algebra interface \lstinline{VarAlg} is defined as follows:
 
-\lstinputlisting[linerange=25-27]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT
+\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+Now pretty-printing on the new language can be realized without modifying existing code:
 
-Now pretty-printing on the new language can be realized by code reuse, without modifying existing code:
+\lstinputlisting[linerange=35-37]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
+It offers programmers with great code reuse, in the sense that existing cases can also be
+overridden in \lstinline{VarExprPrint} using \lstinline{"override def"}.
 
-\lstinputlisting[linerange=31-33]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_LIT_PRINT
+On this, to create an expression of \lstinline{VarExprAlg}, a generic method is defined as follows:
 
-On this, to create an expression of \lstinline{LitAlg}, a generic method is defined as follows:
-
-\huang{todo: better name for build}
-\lstinputlisting[linerange=47-47]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_BUILD
-
-Such a method implicitly represents the expression \lstinline{"x 3"}. The code
+\huang{todo: better name for build}\haoyuan{makeExp}
+\lstinputlisting[linerange=41-41]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+Such a method implicitly represents the expression \lstinline{"1 + x"}. The code
 
 \begin{lstlisting}
-build(new LitPrint{})
+makeExp(new VarExprPrint {})
 \end{lstlisting}
-
-results in \lstinline{"(x 3)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
-has the function type \lstinline{LitAlg[E] =>} \lstinline{E} for generic \lstinline{E}. At this point, one may argue
+results in \lstinline{"(1 + x)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
+has the function type \lstinline{VarExprAlg[E] =>} \lstinline{E} for generic \lstinline{E}. At this point, one may argue
 that it is illusory compared to traditional objects, and since we are struggling against the parsing problem, if a parser
@@
 there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
-in a very clever way. For example, if one wants to transform an \lstinline{ExprAlg} expression to a \lstinline{LitAlg} expression by
-replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:
+in a very clever way. For example, if one wants to transform a \lstinline{VarExprAlg} expression to an \lstinline{ExprAlg} expression by
+replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}
 
-\lstinputlisting[linerange=37-41]{../Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala}% APPLY:linerange=OBJECT_ALG_REFACTOR
-
-When \lstinline{env = Map("y"} \lstinline{-> 3)}, the \lstinline{ExprAlg} expression \lstinline{"x y"} is changed to \lstinline{LitAlg} expression \lstinline{"x 3"}. The abstract method
-\lstinline{alg} in \lstinline{Refactor} can be any algebra of \lstinline{SubAlg}, simply for delegation. By passing a concrete implementation
-to \lstinline{alg}, \lstinline{Refactor} will be able to apply transformation to an \lstinline{ExprAlg} expression before \lstinline{alg} returns the result.
+\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+Here the \lstinline{VarExprAlg} expression \lstinline{"1 + x"} is changed to \lstinline{ExprAlg} expression \lstinline{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
+\lstinline{alg} in \lstinline{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
+to \lstinline{alg}, \lstinline{Refactor} will be able to transform a \lstinline{VarExprAlg} expression before we obtain the result.
 
@@
 A parser can be defined for this whole language at once.
-Since Object Algebras represent ``objects'' implicitly as functions like \lstinline{ExprAlg[E] =>} \lstinline{E} for generic \lstinline{E}, such a parser should consume an algebra from its parameter, then return a value of \lstinline{Fix[Parser[E]]} as in open recursion, where \lstinline{E} is abstract.
-\begin{lstlisting}
-trait ExprParser[E] {
-    val pE : ExprAlg[E] => Fix[Parser[E]] = alg => p =>
-        ident ^^ alg.varE |||
-        p ~ p ^^ { case e1 ~ e2 => alg.appE(e1, e2) }
-}
-\end{lstlisting}
-It is observed that \lstinline{pE} is in fact a parser generator, where \lstinline{alg} appears as a parameter, which can be any algebra, and after \lstinline{alg} is fed, it returns a \lstinline{Fix[Parser[E]]}.
-Again \lstinline{p} is the explicit self-reference of the open parser.
-Inside the body, the algebra is invoked correspondingly for all cases right after parsing.
-Note that the two cases are combined using the original \lstinline{|||} operator.
-To make \lstinline{alg} as a generic algebra, the enclosing trait of the parser, namely \lstinline{ExprParser}, is abstracted over \lstinline{E}.
-Behaviors, or algebras that can be fed to such a parser, are expected to be defined independently from the Object Algebra interface, just like the algebra \lstinline{ExprPrint} in Section~\ref{subsec:objectalgebras}.
+Since Object Algebras represent ``objects'' implicitly as functions like \lstinline{ExprAlg[E] =>} \lstinline{E} for generic \lstinline{E}, such a parser should have type \lstinline{ExprAlg[E] =>} \lstinline{Parser[E]}, where \lstinline{E} can be abstracted by the enclosing type.
 
-To explain the modularity, we again add literals to expressions, but instead of using inheritance as Section~\ref{subsec:objectalgebras}, \lstinline{LitAlg} is defined independently together with its own parser and algebra:
-\begin{lstlisting}
-trait LitAlg[E] {
-    def litE(n : Int) : E
-}
+\lstinputlisting[linerange=60-68]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+Such a parser works fine, yet it does not appear to be extensible.
+For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \lstinline{pExpr} for an extended syntax, but now \lstinline{pExpr} will have a different type, like \lstinline{VarExprAlg[E] =>} \lstinline{Parser[E]} where the difference appears in the argument.
+The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}.
+\haoyuan{Shall we mention this?}
 
-trait LitPrint {
-    def litE(n : Int) = n.toString
-}
-
-trait LitParser[E] {
-    val pE : LitAlg[E] => Fix[Parser[E]] = alg => p =>
-        numericLit ^^ alg.litE
-}
-\end{lstlisting}
-In this case, \lstinline{LitAlg} is yet another small language, which is self-contained. To merge the two languages together, we require their grammars
-to be merged, and we obtain a combined language, whose parser integrates the two small parsers using alternative.
-In Scala we use compound types, namely use \lstinline{"with"} for two Object Algebra interface types, to avoid polluting the namespace.
-
-\begin{lstlisting}
-trait ExprLitParser[E] {
-    val pE : (ExprAlg[E] with LitAlg[E]) => Fix[Parser[E]] = alg => p => {
-        val pExprE = new ExprParser[E]{}.pE(alg)(p)
-        val pLitE = new LitParser[E]{}.pE(alg)(p)
-        pExprE ||| pLitE
-    }
-}
-\end{lstlisting}
-
-In client code, a user can either define a new function with type \lstinline{(ExprAlg[E] with LitAlg[E])} \lstinline{=>} \lstinline{Parser[E]} based on \lstinline{pE}, which implies an algebra is applied after \lstinline{fix}, or simply feed an algebra to \lstinline{pE} at first, then obtain the result from \lstinline{fix}:
-\begin{lstlisting}
-val parsePrint = fix(new ExprLitParser[String]{}.pE(new ExprPrint with LitPrint{}))
-val result = runParser(parsePrint)("x 3") // prints "(x 3)"
-\end{lstlisting}
+One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
+\lstinputlisting[linerange=72-81]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER
+This is precisely the pattern that we advocate for modular parsing.
+Note that \lstinline{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance.
+The reason why we use an extra and seemingly redundant \lstinline{pExprE}, is that \lstinline{pE} can no longer be referred to by the \lstinline{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \lstinline{pExprE} as the specific \lstinline{pE} in \lstinline{ExprOAParser}.
+And we still want code reuse for the composition of \lstinline{pLit} and \lstinline{pAdd}.
+Now it is worth trying on this pattern with an extension, like \lstinline{VarExprAlg} which we have been using:
+\lstinputlisting[linerange=85-96]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXT_OA_PARSER
+In the client code shown above, an instance of algebra is fed to \lstinline{VarExprOAParser} for pretty-printing expressions.
+Then the trait exports \lstinline{pE}, parses the input and applies the algebra.
+The result shows that such a pattern is a nice solution to modular parsing.
+At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice.
+This is generally solved in the original Object Algebras paper [] by composing two algebras into one in advance.
 
@@
 Another one is that Object Algebras benefit us from using generics, where we can abstract multiple syntax in the type parameters, and define less classes, simplifying code comparatively.
-Below we refactor the last example in Section~\ref{subsec:parsingwithopen}, by defining a new Object Algebra interface with two cases \lstinline{intT} and \lstinline{lamE}.
-Meanwhile a parser is implemented for that language.
-\begin{lstlisting}
-trait LamAlg[E, T] {
-    def intT() : T
-    def lamE(x : String, t : T, e : E) : E
-}
+Below we further extend the language with types, by introducing an extra type parameter in the algebra interface.
+The extended grammar is:\\
 
-trait LamParser[E, T] {
-    type Pair = (Parser[E], Parser[T])
-    val pE : LamAlg[E, T] => (=> Pair) => Parser[E] = alg => p =>
-        ("\\" ~> ident) ~ (":" ~> p._2) ~ ("." ~> p._1)
-            ^^ { case x ~ t ~ e => alg.lamE(x, t, e) }
-    val pT : LamAlg[E, T] => (=> Pair) => Parser[T] = alg => p =>
-        "Int" ^^^ alg.intT
-}
-\end{lstlisting}
-Note that we are omitting its lexer here. \lstinline{LamParser} contains parsers for both expressions and types. From the code we observe that the multi-sorts of Object Algebra interface
-exactly express multiple syntax.
-At this point, to combine these two small components together we need some ``glue code''.
-But instead of writing composition by hand, we reimplement the combinator \lstinline{|||}, to compose one parser with type \lstinline{Alg1 =>} \lstinline{(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]} and another one with type \lstinline{Alg2 =>} \lstinline{(=>} \lstinline{R)} \lstinline{=>} \lstinline{Parser[E]} using alternative.
-Note that we can only abstract over the whole algebra interface type, instead of working on its type parameters, as the number of type parameters is not fixed.
-In summary, \lstinline{"|||"} produces a combined parser, which takes a compound algebra as its parameter:
-\begin{lstlisting}
-implicit class Combinator[Alg1, Alg2, R, E](x : Alg1 => (=> R) => Parser[E]) {
-    def |||(y : Alg2 => (=> R) => Parser[E]): (Alg1 with Alg2) => (=> R) => Parser[E]
-        = alg => l => x(alg)(l) ||| y(alg)(l)
-}
-\end{lstlisting}
-Now it is sufficient to merge \lstinline{ExprParser} and \lstinline{LamParser} together:
-\begin{lstlisting}
-trait ExprLamParser[E, T] {
-    val pExpr = new ExprParser[E]{}
-    val pLam = new LamParser[E, T]{}
-    val pExprE : ExprAlg[E] => (=> pLam.Pair) => Parser[E] =
-        alg => p => pExpr.pE(alg)(p._1)
+\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<type> ::= ... \alt `int' \alt <type> `->' <type>
+\end{grammar}
+&
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::=  ... \alt `\\' <ident> `:' <type> `.' <expr>
+\end{grammar}
+\end{tabular}
 
-    val pE = pExprE ||| pLam.pE
-    val pT = pLam.pT
-}
-\end{lstlisting}
-Parsers for expressions and types are connected using \lstinline{|||}, respectively. Hence the isolation for different syntax
-is guaranteed by the type system. The only left problem is that we still need to replace the parser in \lstinline{ExprParser} with \lstinline{pExprE}
-by updating the fix-point parameter. In the next section we will address this problem, and meanwhile our implementation enhances modularity in various aspects.
+Hence we write code for the extended AST, pretty-printing operation and the extended parser:
+
+\lstinputlisting[linerange=100-128]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+Besides lexing, the trait \lstinline{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
+we use \lstinline{pTypedLamT} and \lstinline{pTypedLamE} as copies of current \lstinline{pT} and \lstinline{pE}, recursively. \lstinline{pT} and \lstinline{pE} are actually the parsers for recursion.
+
+From the code we observe that the multi-sorts of Object Algebra interface
+exactly express multiple syntax. The isolation for different syntax
+is guaranteed by the type system. Now it is straightforward to show some client code
+for the new parser:
+
+\lstinputlisting[linerange=132-134]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+Pretty-printing the parsed results just works as expected.

### 2017-01-01 Scala/Parser/src/PaperCode/Overview/OA.scala

@@ -25,66 +25,114 @@ trait Eval extends ExprAlg[Int] {
 }
 //END_OVERVIEW_OA_EVAL
 
-//BEGIN_OVERVIEW_OA_PARSER
-trait BaseParser[E] {
-  lexical.delimiters += ("+")
-  val pExpr: ExprAlg[E] => Open[Parser[E]] = alg => self =>
-    numericLit ^^ alg.lit |||
-    self => self ~ ("+" ~> self) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
-}
-//END_OVERVIEW_OA_PARSER
-
-//BEGIN_OVERVIEW_OA_EXT
-trait VarAlg[E] {
+//BEGIN_OVERVIEW_OA_ALGEXT
+trait VarExprAlg[E] extends ExprAlg[E] {
   def varE(x: String): E
 }
+//END_OVERVIEW_OA_ALGEXT
 
-trait PrintVar extends VarAlg[String] {
+//BEGIN_OVERVIEW_OA_EXTPRINT
+trait VarExprPrint extends VarExprAlg[String] with Print {
   def varE(x: String) = x
 }
+//END_OVERVIEW_OA_EXTPRINT
 
-trait ExtParser[E] extends BaseParser[E] {
-  val pVar: VarAlg[E] => Open[Parser[E]] = alg => self =>
-    ident ^^ alg.varE
-  val pExtExpr: ExprAlg[E] with VarAlg[E] => Open[Parser[E]] =
-    pExpr <|> pVar
+//BEGIN_OVERVIEW_OA_MAKEEXP
+def makeExp[E](alg: VarExprAlg[E]): E = alg.add(alg.lit(1), alg.varE("x"))
+//END_OVERVIEW_OA_MAKEEXP
+
+//BEGIN_OVERVIEW_OA_REFACTOR
+trait Refactor[E] extends VarExprAlg[E] {
+  val alg: ExprAlg[E]
+  val env: Map[String, Int]
+  def lit(n: Int) = alg.lit(n)
+  def add(e1: E, e2: E) = alg.add(e1, e2)
+  def varE(x: String) = alg.lit(env(x))
 }
-//END_OVERVIEW_OA_EXT
 
-  def openParse[E](p: Open[PackratParser[E]])(inp: String): E = {
-    val t = phrase(fix(p))(new lexical.Scanner(inp))
-    if (t.successful) t.get else scala.sys.error(t.toString)
+makeExp(new Refactor {
[…]

### 2017-01-01 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -17,7 +17,7 @@ class Add(e1: Expr, e2: Expr) extends Expr {
 }
 
 // Parser with lexing
-trait BaseParser {
+trait ExprParser {
   lexical.delimiters += ("+")
   val pLit: PackratParser[Expr] =
     numericLit ^^ { x => new Lit(x.toInt) }
@@ -54,23 +54,23 @@ def parse[E](p: Parser[E]): String => E = in => {
   if (t.successful) t.get else scala.sys.error(t.toString)
 }
 
-val result: String = parse(new BaseParser {}.pExpr)("1 + 2").print // "(1 + 2)"
+val result: String = parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
 //END_PACKRAT_RUNPARSER  
 
 /*
 //BEGIN_BAD_ATTEMPT
 val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
-val pExtExpr = pExpr ||| pVar
+val pVarExpr = pExpr ||| pVar
 //END_BAD_ATTEMPT
 */
 
 //BEGIN_INHERITANCE_APPROACH
-trait ExtParser extends BaseParser {
+trait VarExprParser extends ExprParser {
   val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
   override val pExpr = super.pExpr ||| pVar
 }
 
-val result2: String = parse(new ExtParser {}.pExpr)("1 + x").print // "(1 + x)"
+val result2: String = parse(new VarExprParser {}.pExpr)("1 + x").print // "(1 + x)"
 //END_INHERITANCE_APPROACH
 
 /*

### 2017-01-03 paper/sections/CaseStudy.tex
Revise section 6

@@
 
-To demonstrate the utility of our modular parsing library, we
+To demonstrate the utility of our modular parsing approach, we
 implemented parsers of the first 18 calculi from book \textit{Types
-  and Programming Languages} (TAPL). TAPL introduces several
+  and Programming Languages} (TAPL) []. TAPL introduces several
 calculi from simple to complex, by gradually adding new features to
-syntax. \bruno{mention that you compare against an existing
+syntax.
+We compared our implementation with a non-modular implementation we found online [], which is also written in Scala and uses the same Packrat parsing library.
+The comparison is made from two aspects, source lines of code (SLOC) and execution time.
+\huang{done}\bruno{mention that you compare against an existing
   non-modular implementation.}
 
-\subsection{Overview}\label{subsec:cs-overview}
+\subsection{Implementation}
 
@@
 of abstract syntax and modular parsing, which is the key functionality
-of our library. By decomposing those calculi into smaller components
-and reusing the common ones, we can obtain considerably code reuse in
-parsing with little performance penalty, as shown later.
+of our approach. By extracting common components from those calculi
+and reusing them, we obtain considerably code reuse with reasonable performance penalty, as shown later.
 
@@
 Using the modular parsing techniques discussed before, we extract
-reusable \textit{components} from syntax of all the calculi. Each
-component, which may contain several syntactical structures, is a
-reusable fragment representing a certain feature in the syntax. For
-example, \lstinline{VarApp} component represents the variable case and
-application case of expressions. In our case study, each component is
-represented by a Scala object which has three parts. \lstinline{Alg}
-is an object algebra interface for the abstract syntax.
-\lstinline{Print} is an object algebra implements that interface
-\lstinline{Alg}, for the pretty printing operation. It is not
-necessary for parsing, but we include it here for demonstration.
-\lstinline{Parser} is the modular parser for this piece of syntax.
+reusable components from syntax of all the calculi. Each
+component, which may contain several syntactical structures,
+represents a certain feature of the language. For
+example, the \lstinline{VarApp} component below represents variables and
+function applications.
 
+\huang{todo: move the code, rename PackratParser}
 \begin{lstlisting}
@@
   trait Print extends Alg[String] {
-    def TmVar(x: String) = x
-    def TmApp(e1: String, e2: String) = "[" + e1 + " " + e2 + "]"
+    def TmVar(x: String): String = x
+    def TmApp(e1: String, e2: String): String = "[" + e1 + " " + e2 + "]"
   }
-  trait Parser[E, F <: {val pE : PackratParser[E]}] {
+  trait Parser[E] extends EParser[E] {
     lexical.delimiters += ("(", ")")
-    val pVarAppE: Alg[E] => (=> F) => PackratParser[E] = alg => l => {
-      val e = l.pE
-      ident ^^ alg.TmVar |||
-      e ~ e ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
-      "(" ~> e <~ ")"
+    val alg: Alg[E]
+    val pVarAppE: PackratParser[E] = {
+      lcid ^^ alg.TmVar |||
+        pE ~ pE ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
+        "(" ~> pE <~ ")"
     }
@@
 
-We use some naming conventions in our code, \lstinline{E} represents
+In our case study, each component is represented by a Scala object which has three parts.
+\lstinline{Alg} is an object algebra interface for the abstract syntax.
+\lstinline{Print} is an object algebra implements that interface
+\lstinline{Alg}, for the pretty printing operation. They are standard object algebra operations.
+\lstinline{Parser} is the modular parser for this piece of syntax.
+
+We have some naming conventions in our code, \lstinline{E} represents
 expressions, \lstinline{T} represents types and \lstinline{K}
-represents kinds. In this component \lstinline{VarApp} we only have
-expressions. The \lstinline{Alg} and \lstinline{Print} are standard
-object algebra operations, thus we will focus on the modular parser
-\lstinline{Parser}. Each parser carries lexical information and
-parsing functions. The lexical information includes reserved words,
-delimiters, etc. Parsing functions are similar with we described
-before, except that we use record types under some subtyping
-constraints here for the explicit self-reference. The record type
-contains a parser for expressions \lstinline{pE}, a parser for types
-\lstinline{pT} and a parser for kinds \lstinline{pK}, if the
-corresponding one is needed when parsing the syntax of this component.
-Here the type of self-reference, namely \lstinline{F}, is only
-required to have a parser of expressions \lstinline{pE}. Then we can
-build the parsing function \lstinline{pVarAppE}, the suffix
-\lstinline{E} also denotes that it will parse an expression.
+represents kinds. They are the three sorts of syntax in our case study.
+In this particular component \lstinline{VarApp} we only have expressions.
+
+Similar as before, each parser carries lexical information, an object algebra field, and
+parsing functions. We have some helper traits for convenience, such as \inlinecode{EParser}
+containing \inlinecode{pE} for parsing expressions.
+
+\begin{lstlisting}
+trait EParser[E] {
+  val pE: PackratParser[E]
+}
+\end{lstlisting}
 
@@
 Each calculus could be composed directly from components and other
-calculi, if it has those common structures in the syntax. For example,
-the calculus \lstinline{Untyped}, representing the famous untyped
-lambda calculus, can be constructed from components \lstinline{VarApp}
+calculi as needed. For example, the calculus \lstinline{Untyped} in our case study,
+representing the famous untyped lambda calculus, is constructed from components \lstinline{VarApp}
 and untyped lambda abstraction \lstinline{UntypedAbs}.
@@
   trait Print extends Alg[String] {
-    def TmAbs(x: String, e: String) = "\\" + x + "." + e
+    def TmAbs(x: String, e: String): String = "\\" + x + "." + e
   }
-  trait Parser[E, F <: {val pE : PackratParser[E]}] {
-    val pUntypedAbsE: Alg[E] => (=> F) => PackratParser[E] = {...}
+  trait Parser[E] extends EParser[E] {
+    lexical.delimiters += ("\\", ".", "(", ")")
+    val alg: Alg[E]
+    val pUntypedAbsE: PackratParser[E] =
+      ("\\" ~> lcid) ~ ("." ~> pE) ^^ { case x ~ e0 => alg.TmAbs(x, e0) } |||
+      "(" ~> pE <~ ")"
   }
@@
   trait Print extends Alg[String] with UntypedAbs.Print with VarApp.Print
-  trait Parser[E, L <: {val pE : PackratParser[E]}] extends UntypedAbs.Parser[E, L] with VarApp.Parser[E, L] {
-    val pUntypedE = pUntypedAbsE | pVarAppE
+  trait Parser[E] extends UntypedAbs.Parser[E] with VarApp.Parser[E] {
+    override val alg: Alg[E]
+    val pUntypedE: PackratParser[E] = pUntypedAbsE ||| pVarAppE
+    override val pE: PackratParser[E] = pUntypedE
   }
@@
 
-When composing the calculus from components, all the fields can be
-easily combined by Scala’s \lstinline{extends...with} keywords as
-shown in the code. For \lstinline{Parser}, the subtyping constraint
-for the explicit self-reference should satisfy all the requirement of
-the parsers it extends from. \huang{Talk about | here?}
+When composing the calculus from components, all the traits can be
+easily combined by Scala's \inlinecode{extends...with} keywords.
+In \lstinline{Parser}, we need to override the object algebra interface and
+parsing functions accordingly.
 
-Finally, we have demo code for every calculus. The parsing result is a
-string because we provide the pretty printing algebra to the parser,
-and it can be changed as long as the concrete algebra is available. In
-this case study we only have pretty printing as the operation, since
-it is enough for demonstration.
+\paragraph{Demo code}
+We have demo code for every calculus in order to test and evaluate our implementation.
+We feed the pretty printing algebra to the parser and print out the result.
+The parsing result can be changed as long as the concrete algebra is available.
 
@@
 object TestUntyped {
-    to be polished
+  def work[E](inp: String)(a: Untyped.Alg[E]): E = {
+    val lang = new Untyped.Parser[E] {
+      override val alg: Untyped.Alg[E] = a
+    }
+    parse(lang.pE)(inp)
+  }
+  def parseAndPrint(inp: String): Unit = println(work(inp)(new Untyped.Print {}))
+  def main(args: Array[String]): Unit = {
+    parseAndPrint("(\\x.x) (\\y.y)")
+  }
 }
@@
 
-In the next section we will show more examples and talk further about
-how we extend and compose these components and calculi.
-
-\subsection{Extensibility}\label{subsec:cs-extensibility}
+\paragraph{Dependency Overview}
+Figure \ref{fig:dependency} shows the
+dependency of all the components and calculi in our case study. Gray
+boxes are calculi and white boxes are components. An arrow starting
+from box A to box B denotes that B includes and thus reuses A.
 
@@
 
-As mentioned before, we decompose all the calculi into reusable
-components for modularity. Figure \ref{fig:dependency} shows the
-dependency of all the components and calculi in our case study. Gray
-boxes are calculi and white boxes are components. An arrow starting
-from box A to box B denotes that B includes and thus reuses A.
-
 As shown in the graph, some components such as \lstinline{VarApp} are
 created from scratch, while others such as \lstinline{Typed} are
-extended from existing components. The \lstinline{Typed} component
-extends from \lstinline{VarApp} and has two more syntactical
-structures which are typed lambda abstraction and function type (arrow
-type). It is the core of many calculi in this case study, and we use
-it here as an example for demonstrating the extensibility of
-components.
+extended from existing components. Since calculi and components have the similar signature, each calculus
+can also be extended and reused directly. For example, calculus \lstinline{FullRef} extends from
+calculus \lstinline{FullSimple}.
 
-\begin{lstlisting}
-object Typed {
-  trait Alg[E, T] extends VarApp.Alg[E] {
-    def TmAbs(x: String, t: T, e: E): E
-    def TyArr(t1: T, t2: T): T
-  }
-  trait Print extends Alg[String, String] with VarApp.Print {
-    def TmAbs(x: String, t: String, e: String) = "\\(" + x + ":" + t + ")." + e
-    def TyArr(t1: String, t2: String) = t1 + "->" + t2
-  }
-  trait Parser[E, T, F <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends VarApp.Parser[E, F] {
-    ...
-    private val pAbsE: Alg[E, T] => (=> F) => PackratParser[E] = {...}
-    val pTypedE: Alg[E, T] => (=> F) => PackratParser[E] = pVarAppE | pAbsE
-    val pTypedT: Alg[E, T] => (=> F) => PackratParser[T] = alg => l => {
-      val t = l.pT
-      t ~ ("->" ~> t) ^^ { case t1 ~ t2 => alg.TyArr(t1, t2) } ||| "(" ~> t <~ ")"
-    }
-  }
-}
-\end{lstlisting}
-
-Except building a component by extending an existing one, here we have
-another extension which is adding a new sort of syntax. It has been
-discussed in section \ref{subsec:openrecursion}. In this example,
-originally \lstinline{VarApp} has only one sort of syntax which is
-expressions, and then \lstinline{Typed} introduces types to the
-syntax. We add an extra type parameter \lstinline{T} for it, so that
-it is able to be distinguished. The \lstinline{Parser} also requires
-the self-reference to have a field \lstinline{pT} for parsing types,
-which is represented by the subtyping constraint. With this support,
-the \lstinline{pTypedT} function parses arrow types using
-\lstinline{pT}. Another point worth mentioning is that we use a
-private local function \lstinline{pAbsE} here for parsing the typed
-lambda abstraction case, and we can directly combine it with the
-inherited \lstinline{pVarApp} to obtain our parsing function of
-expressions.
-
-Since calculi and components have the similar signature, each calculus
-can also be extended and reused directly. Our last example shows how
-to build calculus \lstinline{FullRef} from extending
-\lstinline{FullSimple}, another calculus.
-
-\begin{lstlisting}
-object FullRef {
-  trait Alg[E, T] extends FullSimple.Alg[E, T] with TopBot.Alg[T] with Ref.Alg[E, T]
-  trait Print extends Alg[String, String] with FullSimple.Print with TopBot.Print with Ref.Print
-  trait Parser[E, T, L <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends FullSimple.Parser[E, T, L] with TopBot.Parser[T, L] with Ref.Parser[E, T, L] {
-    val pFullRefE = pFullSimpleE | pRefE
-    val pFullRefT = pFullSimpleT | pRefT | pTopBotT
-  }
-}
-\end{lstlisting}
-
-Using our modular parsing library, dividing the syntax into small
-pieces and bind them with corresponding parsers to form reusable
-language component is convenient. From the dependency graph we know
-that the common components such as \lstinline{VarApp} are reused in
+Using our modular parsing approach, it is convenient to extract language
+features as components and reuse them. From the dependency graph, we know
+that common components such as \lstinline{VarApp} are reused in
 lots of calculi. Such reuse could shorten the code considerably. We
 will show this advantage and exam the possible performance penalty in
-a quantitative way in the next section.
+a quantitative way in the next subsection.
 
@@
 We compared our implementation of the parsers with an implementation
-by  Ilya Klyuchnikov, available online. Ilya's
+by Ilya Klyuchnikov, available online []. Ilya's
 implementation is suitable for comparison, because it is also
@@
 written in a non-modular way. Thus it is not able to reuse existing
-code of parsing common syntactical features.
+code when those calculi share common features.
 
-Because Ilya's implementation has extra code such as semantics of
-those calculi, we removed all irrelevant code and only keep abstract
+The comparison is made from two aspects. First, we want to discover
+the amount of code reuse using our modular parsing approach.
+For this propose, we measured source lines of code (SLOC) of two implementations.
+Second, we are interested to assess the performance penalty caused by modularity.
+Thus we compared the execution time of parsing random expressions between two implementations.
+
+\paragraph{Standard of Comparison}
+In the SLOC comparison, all blank lines and comment lines are excluded,
+and we formatted the code of both implementations to guarantee that
+the length of each line does not exceed 120 characters. Furthermore,
+because Ilya's implementation has extra code such as semantics,
+we removed all irrelevant code and only keep abstract
 syntax definition, parser and pretty printer for each calculus, to
@@
 
-The comparison is made from two aspects. First, we want to show that
-using our modular parsing library, the code can be reused in a
-significant way. For this propose, we measured source lines of code
-(SLOC) of the two implementations. All blank lines and comment lines
-are excluded, and we formatted the code of both implementations so
-that the length of each line does not exceed 120 characters. Second,
-we are interested to assess the performance penalty caused by modularity. Thus
-we compared the execution time, as a major measurement of performance,
-between the two implementations. We prepared test files containing
-valid expressions in each calculus,\bruno{say something about what
-  those expressions are, size, and how they were generated.} then we used the corresponding
-parser to parse the file and the pretty printer to print the result.
-The execution time excluding reading input file was calculated.
+For the comparison of execution time, we built a generator to randomly
+generate valid expressions of each calculus, according to the syntax. These expressions are
+written to test files, one file per calculus. Each test file consists of 500
+expressions randomly generated, and the size of test files varies from 20KB to 100KB.
+Then we run the corresponding parser to parse the file and the pretty printer to print the result.
+The average execution time of 5 runs excluding reading input file was calculated, in milliseconds.
+
+\huang{done}\bruno{say something about what those expressions are, size, and how they were generated.}
 
@@
 
-Table \ref{tab:comparison} shows data of the comparison. The overall
-result of SLOC comparison is that 66.3\% of code is reduced using our
-library. That is because the code of some common syntax structures
-such as variables, lambda abstractions, etc., are reused several times in
-the whole project. We can see that in the first two calculi
-\lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's
+\paragraph{Comparison Results}
+Table \ref{tab:comparison} shows results of the comparison.
+The overall result is that 69.2\% of code is reduced using our
+library, and our implementation is 42.7\% slower.
+
+The good SLOC result is because of that the code of common language features
+such as variables, lambda abstractions, etc., are reused lots of times in
+the whole case study. We can see that in the first two calculi
+\lstinline{Arith} and \lstinline{Untyped} we are not better than Ilya's
 implementation, because in such two cases we do not reuse any existing
-components. But in each of the following 16 calculi, we can compose
-our syntax definition, parser and pretty printer from extending
-reusable components, resulting considerably code reduction.
+components. However in the following 16 calculi, we can reuse
+language components, resulting considerably code reduction. In particular,
+the calculi \inlinecode{EquiRec}, \inlinecode{Recon} and some others are only 22 lines
+in our implementation, because we only compose existing codes in such cases.
 
-\paragraph{Performance}
-For the execution time comparison, we built a generator to randomly
-generate valid expressions according to the syntax of each calculus.
-Each test file corresponds with one calculus and it consists of 500
-expressions randomly generated. The size of test files varies from
-20KB to 100KB. After testing both implementation, we found that our
-version is about \huang{60\%} slower. \bruno{A table showing a summary
-of results is missing here. We discussed this before.}
- Part of the reason is that our
-library requires more intermediate function calls, and the object
-algebra and self-references need to be passed as arguments every
-time. However, another important reason is that, because of the
-modular setting, we use a different alternative combinator which will
-try both branches and match the longest one. \huang{ref to somewhere
-  before?} In contrast Ilya uses the ordinary alternative combinator
-which tries the second branch only if the first one fails. Although
-this difference slows our version, we want to mention that because of
-the memoization of Packrat parsers, we are just constant times
-slower, the big-O complexity is still the same.
+To discover the reasons of slower execution time, we made some more experiments
+on two possible factors which could affect the performance.
+They are object algebra and the longest matching alternative combinator.
+We use object algebra for ASTs and the longest matching alternative combinator \inlinecode{|||} for parsing,
+while Ilya's implementation uses case class and the ordinary alternative combinator.
+
+We implemented two more versions. One is a modified version of our implementation,
+with object algebra replaced by case class for the ASTs.
+The other is a modified version of Ilya's implementation, using the longest matching
+alternative combinator instead of the ordinary one.
+
+\huang{done}\bruno{A table showing a summary of results is missing here. We discussed this before.}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|l|r|r|r|r|r|r|r|}
+      \hline
+        \multirow{2}{*}{\bfseries Calculus Name} & \bfseries Ilya's & \multicolumn{2}{ c| }{\bfseries Ours} & \multicolumn{2}{ c| }{\bfseries Ilya's with \inlinecode{|||}} & \multicolumn{2}{ c| }{\bfseries Ours with class} \\ \cline{3-8}
+        \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{\bfseries Time} & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% \\
+      \hline
+        Arith & 741 & 913 & +23.2 & 793 & +7.0 & 932 & +25.8 \\
+        Untyped & 770 & 1018 & +32.2 & 821 & +6.6 & 1007 & +30.8 \\
+      \hline
+        \multicolumn{1}{|c|}{\dots} & \multicolumn{7}{c|}{\dots} \\
+      \hline
+        FullRecon & 1094 & 1645 & +50.4 & 1161 & +6.1 & 1652 & +51.0 \\
+        FullPoly & 1398 & 2086 & +49.2 & 1511 & +8.1 & 2019 & +44.4 \\
+        FullOmega & 1451 & 2352 & +62.1 & 1582 & +9.0 & 2308 & +59.1 \\
+      \hline
+        Total & 21746 & 31024 & +42.7 & 23260 & +7.0 & 30795 & +41.6 \\
+      \hline
+    \end{tabular}
+    \caption{Execution time of four versions}
+    \label{tab:ext-comparison}
+\end{table}
+
+Table \ref{tab:ext-comparison} shows part of the comparison results of these four versions,
+full table is available online []. It suggests that the difference of running time between
+using object algebra and class is little, roughly 1\%.
+The usage of longest matching combinator slows the performance by 7\%. The main reason of slower
+execution time may be the overall structure of the modular parsing approach, because we indeed have
+more intermediate function calls and method overriding. However, it is worth metioning that
+because of the memoization technique of Packrat parsers, we are only constant times
+slower, the algorithmic complexity is still the same.

### 2017-01-03 paper/sections/CaseStudy.tex
Changes

@@
 implemented parsers of the first 18 calculi from book \textit{Types
-  and Programming Languages} (TAPL) []. TAPL introduces several
+  and Programming Languages} (TAPL) \cite{pierce2002types}. TAPL introduces several
 calculi from simple to complex, by gradually adding new features to
-syntax.
-We compared our implementation with a non-modular implementation we found online [], which is also written in Scala and uses the same Packrat parsing library.
-The comparison is made from two aspects, source lines of code (SLOC) and execution time.
-\huang{done}\bruno{mention that you compare against an existing
-  non-modular implementation.}
+syntax. We compared our implementation with a non-modular implementation
+we found online, which is also written in Scala and uses the same Packrat parsing library.
+We counted source lines of code (SLOC) and measured execution time for both implementations.
+Our implementation saves almost 70\% code comparing with that non-modular one.
+
+\huang{done}\bruno{mention that you compare against an existing non-modular implementation.}
 
@@
 We compared our implementation of the parsers with an implementation
-by Ilya Klyuchnikov, available online []. Ilya's
-implementation is suitable for comparison, because it is also
+by Ilya Klyuchnikov, available online\footnote{https://github.com/ilya-klyuchnikov/tapl-scala/}.
+Ilya's implementation is suitable for comparison, because it is also
 written in Scala using the same parser combinator library.

### 2017-01-03 paper/sections/PackratParsing.tex
Revised S4.
Changes

@@
 composition in recursive descent parsers, the parser in the front successfully parses a substring of the input, then it unexpectedly
-finishes parsing and discards the rest, in spite that latter parsers might be abl e to parse the whole input.
+finishes parsing and discards the rest, in spite that latter parsers might be able to parse the whole input.
 %\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
@@
 
-\noindent Unfortunatelly, this results in the worst exponential time complexity!
+\noindent However, this results in the worst case exponential time complexity!
 
@@
 \end{comment}
-In the current version of Scala (2.11), Packrat Parsing is included
-in the Scala standard parser combinator library. The library provides a number of
-parser combinators, together with some basic parsers. Below we present an example to illustrate Scala parsers.
+Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
+which implements Packrat parsers.
+The library provides a number of parser combinators, together with some basic parsers.
+Below we present an example to illustrate Scala parsers.
 %
@@
 
-\paragraph{Parsing a simple Arithmetic Language}
+\paragraph{Parsing a Simple Arithmetic Language}
 Suppose we want to parse a simple language with literals and
@@
 
+\huang{doesn't compile}
+
 \begin{figure}[t]
@@
 
-In the trait \lstinline{BaseParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} can parse an integer for \lstinline{Lit},
+In the trait \lstinline{ExprParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} can parse an integer for \lstinline{Lit},
 whereas \lstinline{pAdd} handles addition and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}

### 2017-01-03 paper/sections/Discussion.tex
Revised S5. Related work to be revised.

\section{Discussion}\label{sec:discussion}

The use of interface-based approach and Object Algebras enables us to do modular parsing. Programmers can easily follow the pattern
captured in the previous section, build modular ASTs and hence have access to multiple dimensions of extensibility. This section
explores more interesting features of this pattern, including overriding existing parsing rules, and framework of language components
for modularity over hierarchies.

\subsection{Overriding Existing Rules}

Just as many parser generators do, our pattern also supports overriding part of existing parsers easily, but in a type-safe way.
We can simply update existing production rules, namely the concrete syntax, with new implementations, or eliminate them in extended parsers.
To illustrate, suppose we have an untyped lambda case in a base parser, but then we replace it with a typed one in its extended parser:

\lstinputlisting[linerange=139-144]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
The above parser shows part of the code, for the space reason we omit details. With OO inheritance, we can actually override \lstinline{pLam}
in a subtype of \lstinline{BaseParser}, and give a different implementation. It will make a difference during runtime due to dynamic
dispatch.

\lstinputlisting[linerange=148-154]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
On the other hand, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
One can even ``eliminate'' a production rule in the extension, by actually overriding it with a failure parser.

It is natural to keep overriding existing parsers, whereas another potential use of this pattern is to reuse old versions of a parser. For instance,
we have a \lstinline{NewParser} that extends \lstinline{ExtParser}, with some more cases added, but for the lambda case we want to go back to the untyped one.
It can be done by copying the code from \lstinline{BaseParser} again, yet does not achieve code reuse. Instead, we can instantiate an instance of \lstinline{BaseParser} to obtain the old \lstinline{pLam}:
\haoyuan{I don't know if this works or makes sense.}

\lstinputlisting[linerange=158-164]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM

\subsection{Framework of a Language Component}\label{subsec:framework}

Having pointed out that our goal is to realize modular parsers, we expect the code written in an extensible way. We have been focusing on how to extend a language with some additional cases, just as many extensible compilers do, nevertheless, in other cases, languages are defined separately, and the composition of them is again a matter of code reuse. In this section, we propose a framework for defining small languages as modular components, in order for better composition.

The framework we use for a language, or rather a language component in our case study, consists of three parts: syntax, parsing and operations on the AST. More
specifically:
\begin{itemize}
    \item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax. The type parameters represent multiple sorts of syntax, and the methods are constructs.
\item \textbf{Parser:} implementation of the parser with lexing, using inheritance and Object Algebras, with its scoping trait abstracted over the type parameters.
\item \textbf{Algebra (optional):} some behaviors (operations) on the data structure, such as pretty-printing and so on.
\end{itemize}

Taking again the example in Section~\ref{subsec:parsingwithoa}. The language constructs consists of literals, additions and variables. First we define the
language component in a Scala object, shown in the left part of Figure~\ref{fig:lng-components}.

\begin{figure}[t]
\begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
\lstinputlisting[linerange=7-22]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
&
\lstinputlisting[linerange=26-41]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
\end{tabular}
\caption{Language \lstinline{VarExpr} (left) and language \lstinline{TypedLam} (right).}\label{fig:lng-components}
\end{figure}

For the example in Section~\ref{subsec:differentsyntax}, instead of building inheritance directly, we define the language separately,
which consists of types and typed lambda expressions. Shape of the code is presented in the right of Figure~\ref{fig:lng-components}.

At this point, one may want those two language components to be composed into one, and meanwhile ensure that the one is still a modular
component ready for future composition. In that case modularity is realized over higher-order hierarchies. The composed language \lstinline{VarLamExpr}
is shown below:

\lstinputlisting[linerange=45-54]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
Note that lexing code is automatically combined under multiple inheritance. The combined language shows perfect modularity. The only
drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty printers of
\lstinline{VarExpr} and \lstinline{TypedLam} respectively. Such a pattern refers to \textit{family polymorphism}[]. Unfortunately Scala does not
support this feature, nonetheless, one can avoid such boilerplate by using metaprogramming techniques.

\subsection{Fusion of Concepts}\label{subsec:fusion}

\haoyuan{Not sure if this section should be here, or in the related work. I tend to have a big discussion section.}
\huang{These alternatives could be discussed in the discussion
  section, not here. 6.1 could only talks about library code. And the
  titles should be changed as well}\bruno{I think this subsection is
  unnecessary. The code will be presented already at the end of S5, I
  presume. }

In the previous sections we have talked about Packrat Parsing, open recursion and Object Algebras. These components fuse together in our library implementation, for the initial goal, which is to build extensible parsers. We want to argue here that those aspects are totally orthogonal, and hence can have their alternatives. For example, Object Algebras are helpful to build extensible data structures, whereas in functional programming, there has been a large amount of related work on extensible datatypes, including DTC and MRM for the Haskell language. And a better
parsing library would potentially introduce a lot of fancy features and parser combinators, run with high efficiency, or even support indirect left-recursion.
\haoyuan{Open recursion could also have alternatives?} For now, our implementation integrates the existing ones, and has turned out to be practical in various applications.

### 2017-01-03 paper/sections/ObjectAlgebras.tex
Revised S5. Related work to be revised.

@@
 On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Now we extend the language
-with variables. A new Object Algebra interface \lstinline{VarAlg} is defined as follows:
+with variables. A new Object Algebra interface \lstinline{VarAlg} is defined as follows:\haoyuan{Can we use var instead of varE?}
 
@@
 
-\lstinputlisting[linerange=100-128]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=100-129]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 Besides lexing, the trait \lstinline{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
@@
 
-\lstinputlisting[linerange=132-134]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=133-135]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 Pretty-printing the parsed results just works as expected.

### 2017-01-03 paper/sections/OpenRecursion.tex
Revised S5. Related work to be revised.

@@
 
-\lstinputlisting[linerange=24-30]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_FIB
 
@@
 
-\lstinputlisting[linerange=34-40]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB2
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_FIB2
 \bruno{Fix should be replaces by Open. Please coordinate with Huang to
@@
 
-\lstinputlisting[linerange=44-44]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIB_FIX
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_FIB_FIX
 
@@
 
-\lstinputlisting[linerange=20-20]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_FIX
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_FIX
 \bruno{can be defined in a single line.}
@@
 Suppose now we want to make use of open recursion for additional operations. We first define a combinator which combines the results of two generators in a pair:
-\lstinputlisting[linerange=48-51]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_COMPOSE
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_COMPOSE
 
@@
 Then we define a generator which prints out a number as a string:
-\lstinputlisting[linerange=55-55]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_SHOW
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_SHOW
 
@@
 The definition is straightforward:
-\lstinputlisting[linerange=59-60]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=OPENRECURSION_MERGE
+\lstinputlisting[linerange=-]{}% APPLY:linerange=OPENRECURSION_MERGE
 

### 2017-01-03 Scala/Parser/src/PaperCode/Overview/OA.scala

@@ -115,7 +115,8 @@ trait TypedLamOAParser[E, T] extends VarExprOAParser[E] {
   override val alg: TypedLamAlg[E, T]
   
   val pIntT: Parser[T] = "int" ^^^ alg.intT
-  val pArrowT: Parser[T] = pT ~ ("->" ~> pT) ^^ { case t1 ~ t2 => alg.arrowT(t1, t2) }
+  val pArrowT: Parser[T] = pT ~ ("->" ~> pT) ^^
+    { case t1 ~ t2 => alg.arrowT(t1, t2) }
   val pTypedLamT: Parser[T] = pIntT ||| pArrowT
   
   val pLam: Parser[E] = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^
@@ -133,6 +134,35 @@ val result2 = parse(new TypedLamOAParser[String, String] {
   override val alg = new TypedLamPrint {}
 }.pE)("\\x:int->int. 1 + x") // "\x : int -> int. (1 + x)"
 //END_OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+
+//BEGIN_BASEPARSER_UNTYPEDLAM
+trait BaseParser[E] {
+  ...
+  val alg: ...
+  val pLam: Parser[E] = ("\\" ~> ident) ~ ("." ~> pE) ^^ ...
+  val pE: Parser[E] = ...
+}
+//END_BASEPARSER_UNTYPEDLAM
+
+//BEGIN_EXTPARSER_TYPEDLAM
+trait ExtParser[E, T] extends BaseParser[E] {
+  ...
+  override val alg: ...
+  override val pLam = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ ...
+  val pT: Parser[T] = ...
+  override val pE: Parser[E] = ...
+}
+//END_EXTPARSER_TYPEDLAM
+
+//BEGIN_NEWPARSER_UNTYPEDLAM
+trait NewParser[E, T] extends ExtParser[E, T] {
+  ...
+  override val alg: ...
+  val tempAlg = alg
+  override val pLam = new BaseParser[E] { override val alg = tempAlg }.pLam
+  ...
+}
+//END_NEWPARSER_UNTYPEDLAM
 
   def main(args: Array[String]): Unit = {
 

### 2017-01-03 Scala/Parser/src/PaperCode/SampleParser.scala

@@ -2,6 +2,59 @@ package PaperCode
 
 import TAPL.Util._
 
+/*
+//BEGIN_LANGUAGE_COMPONENTS_VAREXPR
+object VarExpr {
+  // Abstract syntax
+  trait Alg[E] {
+    def lit(n: Int): E
+    def add(e1: E, e2: E): E
+    def varE(x: String): E
+  }
+  
+  // Parser
+  trait Parser[E] { ... }
+  
+  // Pretty-printer
+  trait Print extends Alg[String] {
+    ...
+  }
+}
+//END_LANGUAGE_COMPONENTS_VAREXPR
+
+//BEGIN_LANGUAGE_COMPONENTS_TYPEDLAM
+object TypedLam {
+  // Abstract syntax
+  trait Alg[E, T] {
+    def intT(): T
+    def arrowT(t1: T, t2: T): T
+    def lam(x: String, t: T, e: E): E
+  }
+  
+  // Parser
+  trait Parser[E] { ... }
+  
+  // Pretty-printer
+  trait Print extends Alg[String, String] {
+    ...
+  }
+}
+//END_LANGUAGE_COMPONENTS_TYPEDLAM
+
+//BEGIN_LANGUAGE_COMPONENTS_VARLAMEXPR
+object VarLamExpr {
+  trait Alg[E, T] extends VarExpr.Alg[E] with TypedLam.Alg[E, T]
+  
+  trait Parser[E, T] extends VarExpr.Parser[E] with TypedLam.Parser[E, T] {
+    override val alg: Alg[E, T]
+    override val pE = ...
+  }
+  
+  trait Print extends VarExpr.Print with TypedLam.Print
+}
+//END_LANGUAGE_COMPONENTS_VARLAMEXPR
+/*
+
 object SampleParser {
 
 //BEGIN_PACKRAT_EXAMPLE
@@ -12,51 +65,4 @@ val p: PackratParser[Int] =
   "str" ~> ("(" ~> numericLit <~ ")") ^^ { x => x.toInt }
 //END_PACKRAT_EXAMPLE
      
-}
-
-object OpenRecursion {
-
-//BEGIN_OPENRECURSION_FIX
-def fix[T](f: Open[T]): T = { lazy val a: T = f(a); a }
-//END_OPENRECURSION_FIX
-
-//BEGIN_OPENRECURSION_FIB
-def fib: Int => Int = {
[…]

### 2017-01-07 paper/sections/CaseStudy.tex
Some changes

@@
 on two possible factors which could affect the performance.
-They are object algebra and the longest matching alternative combinator.
-We use object algebra for ASTs and the longest matching alternative combinator \inlinecode{|||} for parsing,
+They are object algebra and the longest match alternative combinator.
+We use object algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
 while Ilya's implementation uses case class and the ordinary alternative combinator.
@@
 with object algebra replaced by case class for the ASTs.
-The other is a modified version of Ilya's implementation, using the longest matching
+The other is a modified version of Ilya's implementation, using the longest match
 alternative combinator instead of the ordinary one.
@@
 using object algebra and class is little, roughly 1\%.
-The usage of longest matching combinator slows the performance by 7\%. The main reason of slower
+The usage of longest match combinator slows the performance by 7\%. The main reason of slower
 execution time may be the overall structure of the modular parsing approach, because we indeed have

### 2017-01-07 paper/sections/Discussion.tex
Some changes

@@
 
-\lstinputlisting[linerange=139-144]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=106-111]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
 The above parser shows part of the code, for the space reason we omit details. With OO inheritance, we can actually override \lstinline{pLam}
@@
 
-\lstinputlisting[linerange=148-154]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\lstinputlisting[linerange=115-121]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
 On the other hand, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
@@
 
-\lstinputlisting[linerange=158-164]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=125-131]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
 

### 2017-01-07 paper/sections/Inheritance.tex
Some changes

@@
 \begin{itemize}
-\item \textbf{Extension 1}: adding a new case (or rather, a new construct) to \lstinline{Expr};
-\item \textbf{Extension 2}: adding a new operation to \lstinline{Expr}.
+\item \textbf{Extension 1}: adding a new case (or rather, a new construct) to \inlinecode{Expr};
+\item \textbf{Extension 2}: adding a new operation to \inlinecode{Expr}.
 \end{itemize}
@@
 
-\lstinputlisting[linerange=98-100]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_LAM
+\lstinputlisting[linerange=30-32]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
 
@@
 like to build the new parser by reusing the old one. Here one may quickly come up
-with the following attempt, where a new parser is defined for \lstinline{Var}, then composed
-with \lstinline{pExpr}:
+with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
+with \inlinecode{pExpr}:
 
-\lstinputlisting[linerange=62-63]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=BAD_ATTEMPT
+\huang{Extending from old parser is necessory to get the `pExpr', added a surrounding trait.}
+\lstinputlisting[linerange=43-46]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
 
-Unfortunately, it fails to parse some expressions like \lstinline{"1 + x"}, which is obviously valid in the new grammar.
-The reason is that \lstinline{pAdd} makes two recursive calls to parse sub-expressions, by using \lstinline{pExpr}, which
-covers both cases in the old grammar. But the newly added case \lstinline{pVar} is not observed by the recursive \lstinline{pExpr},
-hence the parser does not work as expected. Still, one may continue to change the recursive call in \lstinline{pAdd} with \lstinline{pVarExpr},
+Unfortunately, it fails to parse some expressions like \inlinecode{"1 + x"}, which is obviously valid in the new grammar.
+The reason is that \inlinecode{pAdd} makes two recursive calls to parse sub-expressions, by using \inlinecode{pExpr}, which
+covers both cases in the old grammar. But the newly added case \inlinecode{pVar} is not observed by the recursive \inlinecode{pExpr},
+hence the parser does not work as expected. Still, one may continue to change the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr},
 but modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.
@@
 \subsection{Inheritance-Based Approach}\label{subsec:inheritance-approach}
+
+\huang{It seems a little bit weird since we were already using inheritance in the `bad attempt'. I think we could focus on overriding here}
 
@@
 
-\lstinputlisting[linerange=68-73]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=INHERITANCE_APPROACH
-Now \lstinline{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
-method overriding in inheritance. When the input \lstinline{"1 + x"} is fed to the parser \lstinline{this.pExpr}, it firstly delegates
-the work to \lstinline{super.pExpr}, which parses literals and additions. However, the recursive call \lstinline{pExpr} in \lstinline{pAdd}
-actually refers to \lstinline{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
+\lstinputlisting[linerange=50-55]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
+Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
+method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
+the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
+actually refers to \inlinecode{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
 elegant inheritance-based approach opens the recursion of parsers in a modular way.
@@
 Notice that Scala has a linearized-style multiple inheritance for traits. It can be very helpful when composing several languages. Suppose now
-we want to compose the parsers for expressions from pre-defined languages \lstinline{LanguageA} and \lstinline{LanguageB} using alternative.
-The use of keyword \lstinline{super} is able to
+we want to compose the parsers for expressions from pre-defined languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using alternative.
+The use of keyword \inlinecode{super} is able to
 specify the implementation from inheritance, like:
 
-\lstinputlisting[linerange=78-84]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
+\lstinputlisting[linerange=60-66]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 

### 2017-01-07 paper/sections/ObjectAlgebras.tex
Some changes

@@
 The inheritance-based approach benefits us from building extensible parsers, based on an OO class hierarchy.
-Nevertheless, the addition of new operations over the AST becomes problematic. Often it requires modification on
+Nevertheless, the addition of new operations over the ASTs becomes problematic. Often it requires modification on
 the existing code, which breaks modularity. In this section, we will have a closer look at how it motivates us to
@@
 Now it comes to extension 2 in Section~\ref{subsec:dimensions}. In the previous code, we have presented pretty-printing as
-an example of operations on the AST. Suppose now we also want to collect free variables on the expression language, which models
+an example of operations on the ASTs. Suppose now we also want to collect free variables on the expression language, which models
 literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \lstinline{Expr} with the new
@@
 
-\lstinputlisting[linerange=89-89]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
+\lstinputlisting[linerange=19-19]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
 It refers to one Scala solution to the Expression Problem []. But then programmers have to define new classes for literals, additions and variables
@@
 
-\lstinputlisting[linerange=8-11]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALG
+\lstinputlisting[linerange=7-10]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
 The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
@@
 
-\lstinputlisting[linerange=15-18]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
+\lstinputlisting[linerange=14-17]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
 Here \lstinline{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
@@
 
-\lstinputlisting[linerange=22-25]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
+\lstinputlisting[linerange=21-24]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
@@
 
-\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+\lstinputlisting[linerange=28-30]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
 Now pretty-printing on the new language can be realized without modifying existing code:
 
-\lstinputlisting[linerange=35-37]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+\lstinputlisting[linerange=34-36]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
 We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
@@
 \huang{todo: better name for build}\haoyuan{makeExp}
-\lstinputlisting[linerange=41-41]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+\lstinputlisting[linerange=8-8]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
 Such a method implicitly represents the expression \lstinline{"1 + x"}. The code
@@
 
-\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+\lstinputlisting[linerange=12-23]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
 Here the \lstinline{VarExprAlg} expression \lstinline{"1 + x"} is changed to \lstinline{ExprAlg} expression \lstinline{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
@@
 
-\lstinputlisting[linerange=60-68]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+\lstinputlisting[linerange=27-35]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
 Such a parser works fine, yet it does not appear to be extensible.
@@
 One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
-\lstinputlisting[linerange=72-81]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER
+\lstinputlisting[linerange=39-48]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
@@
 Now it is worth trying on this pattern with an extension, like \lstinline{VarExprAlg} which we have been using:
-\lstinputlisting[linerange=85-96]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXT_OA_PARSER
+\lstinputlisting[linerange=52-63]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXT_OA_PARSER
 In the client code shown above, an instance of algebra is fed to \lstinline{VarExprOAParser} for pretty-printing expressions.
@@
 
-\lstinputlisting[linerange=100-129]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=67-96]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 Besides lexing, the trait \lstinline{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
@@
 
-\lstinputlisting[linerange=133-135]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=100-102]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 Pretty-printing the parsed results just works as expected.

### 2017-01-07 paper/sections/PackratParsing.tex
Some changes

@@
 
-This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators from Packrat Parsing
+This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators from Packrat parsing
 are suitable to address those challenges.
@@
 Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
-which implements Packrat parsers.
+\cite{moors2008parser} which implements Packrat parsers.
 The library provides a number of parser combinators, together with some basic parsers.
@@
 
-\huang{doesn't compile}
-
 \begin{figure}[t]
 \centering
-\lstinputlisting[linerange=7-27]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=OVERVIEW_SIMPLE_EXPR
+\lstinputlisting[linerange=6-29]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
 \caption{Packrat Parsing for an arithmetic language.}\label{fig:packrat-arith}
@@
 whereas \lstinline{pAdd} handles addition and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
-recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the alternative combinator with longest match.
+recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the longest match alternative combinator \inlinecode{|||}.
 Table~\ref{tab:packrat} gives a closer look at part of the API that we have used in our work.
@@
 
-Finally, a generic \inlinecode{parse} function is used for testing:
+It is worth mentioning that such a grammar is left-recursive, yet well supported by Packrat parsers. It also employs automatic
+longest match composition by using the combinator \lstinline{|||} in the library. Furthermore, Packrat parsers do not suffer from the backtracking
+problem, as the technique uses memoization to record intermediate parsing results.
 
-\lstinputlisting[linerange=52-57]{../Scala/Parser/src/PaperCode/Overview/Overview.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+To shorten our code, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing in the rest of paper.
+The last line demonstrates how we parse a valid expression \inlinecode{1 + 2} using our parser.
 
-It is worth mentioning that such a grammar is left-recursive, yet well supported by Packrat parsers. It also employs automatic
-longest match composition by using the library function \lstinline{|||}. Finally, Packrat parsers do not suffer from the backtracking
-problem, as the technique uses memoization to record intermediate parsing results. We will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} in the rest of paper.
+\lstinputlisting[linerange=36-43]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER

### 2017-01-07 Scala/Parser/src/PaperCode/Overview/OA.scala

@@ -4,39 +4,6 @@ import PaperCode.Overview.Util._
 
 object OA {
 
-//BEGIN_OVERVIEW_OA_ALG
-trait ExprAlg[E] {
-  def lit(n: Int): E
-  def add(e1: E, e2: E): E
-}
-//END_OVERVIEW_OA_ALG
-
-//BEGIN_OVERVIEW_OA_PRINT
-trait Print extends ExprAlg[String] {
-  def lit(n: Int) = x.toString
-  def add(e1: String, e2: String) = "(" + e1 + " + " + e2 + ")"
-}
-//END_OVERVIEW_OA_PRINT
-
-//BEGIN_OVERVIEW_OA_EVAL
-trait Eval extends ExprAlg[Int] {
-  def lit(n: Int) = n
-  def add(e1: Int, e2: Int) = e1 + e2
-}
-//END_OVERVIEW_OA_EVAL
-
-//BEGIN_OVERVIEW_OA_ALGEXT
-trait VarExprAlg[E] extends ExprAlg[E] {
-  def varE(x: String): E
-}
-//END_OVERVIEW_OA_ALGEXT
-
-//BEGIN_OVERVIEW_OA_EXTPRINT
-trait VarExprPrint extends VarExprAlg[String] with Print {
-  def varE(x: String) = x
-}
-//END_OVERVIEW_OA_EXTPRINT
-
 //BEGIN_OVERVIEW_OA_MAKEEXP
 def makeExp[E](alg: VarExprAlg[E]): E = alg.add(alg.lit(1), alg.varE("x"))
 //END_OVERVIEW_OA_MAKEEXP

### 2017-01-07 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -1,171 +1,5 @@
 package PaperCode.Overview
 
-import PaperCode.Overview.Util._
-
-/*
-//BEGIN_OVERVIEW_SIMPLE_EXPR
-import scala.util.parsing.combinator.syntactical.StandardTokenParsers
-import scala.util.parsing.combinator.PackratParsers
-
-// Abstract syntax
-trait Expr { def print: String }
-class Lit(x: Int) extends Expr {
-  def print = x.toString
-}
-class Add(e1: Expr, e2: Expr) extends Expr {
-  def print = "(" + e1.print + " + " + e2.print + ")"
-}
-
-// Parser with lexing
-trait ExprParser {
-  lexical.delimiters += ("+")
-  val pLit: PackratParser[Expr] =
-    numericLit ^^ { x => new Lit(x.toInt) }
-  val pAdd: PackratParser[Expr] =
-    pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-  val pExpr: PackratParser[Expr] = pLit ||| pAdd
-}
-//END_OVERVIEW_SIMPLE_EXPR
-*/
-
-object Latest {
-
-type Parser[E] = PackratParser[E]
-
-trait Expr { def print: String }
-class Lit(x: Int) extends Expr {
-  def print = x.toString
-}
-class Add(e1: Expr, e2: Expr) extends Expr {
-  def print = "(" + e1.print + " + " + e2.print + ")"
-}
-
-trait BaseParser {
-  lexical.delimiters += ("+")
-  val pLit: Parser[Expr] = numericLit ^^ { x => new Lit(x.toInt) }
-  val pAdd: Parser[Expr] =
-    pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-  val pExpr: Parser[Expr] = pLit ||| pAdd
-}
-
-//BEGIN_PACKRAT_RUNPARSER
-def parse[E](p: Parser[E]): String => E = in => {
-  val t = phrase(p)(new lexical.Scanner(in))
-  if (t.successful) t.get else scala.sys.error(t.toString)
-}
-
-val result: String = parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
-//END_PACKRAT_RUNPARSER  
-
-/*
[…]

### 2017-01-07 Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala

@@ -1,0 +1,31 @@
+package PaperCode.Sec2Packrat
+
+
+object Code1 {
+//BEGIN_PACKRAT_SIMPLE_EXPR
+import scala.util.parsing.combinator.syntactical._
+import scala.util.parsing.combinator._
+
+// Abstract syntax
+trait Expr {
+  def print: String
+}
+class Lit(x: Int) extends Expr {
+  def print = x.toString
+}
+class Add(e1: Expr, e2: Expr) extends Expr {
+  def print = "(" + e1.print + " + " + e2.print + ")"
+}
+
+// Parser with lexing
+trait ExprParser extends StandardTokenParsers with PackratParsers {
+  lexical.delimiters += "+"
+  val pLit: PackratParser[Expr] =
+    numericLit ^^ { x => new Lit(x.toInt) }
+  val pAdd: PackratParser[Expr] =
+    pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
+  val pExpr: PackratParser[Expr] =
+    pLit ||| pAdd
+}
+//END_PACKRAT_SIMPLE_EXPR
+}

### 2017-01-07 Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala

@@ -1,0 +1,48 @@
+package PaperCode.Sec2Packrat
+
+import scala.util.parsing.combinator.syntactical._
+import scala.util.parsing.combinator._
+
+
+object Code2 extends StandardTokenParsers with PackratParsers {
+
+  // Abstract syntax
+  trait Expr {
+    def print: String
+  }
+
+  class Lit(x: Int) extends Expr {
+    def print = x.toString
+  }
+
+  class Add(e1: Expr, e2: Expr) extends Expr {
+    def print = "(" + e1.print + " + " + e2.print + ")"
+  }
+
+  // Parser with lexing
+  trait ExprParser {
+    lexical.delimiters += "+"
+    val pLit: PackratParser[Expr] =
+      numericLit ^^ { x => new Lit(x.toInt) }
+    val pAdd: PackratParser[Expr] =
+      pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
+    val pExpr: PackratParser[Expr] =
+      pLit ||| pAdd
+  }
+
+  def main(args: Array[String]): Unit = {
+
+//BEGIN_PACKRAT_RUNPARSER
+type Parser[E] = PackratParser[E]
+
+def parse[E](p: Parser[E]): String => E = in => {
+  val t = phrase(p)(new lexical.Scanner(in))
+  if (t.successful) t.get else scala.sys.error(t.toString)
+}
+
+parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
+//END_PACKRAT_RUNPARSER
+
+    println(parse(new ExprParser {}.pExpr)("1 + 2").print)
+  }
+}

### 2017-01-07 Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala

@@ -1,0 +1,69 @@
+package PaperCode.Sec3Inheritance
+
+import scala.util.parsing.combinator.syntactical._
+import scala.util.parsing.combinator._
+
+
+object Code1 extends StandardTokenParsers with PackratParsers {
+
+  trait Expr {
+    def print: String
+  }
+  class Lit(x: Int) extends Expr {
+    def print = x.toString
+  }
+  class Add(e1: Expr, e2: Expr) extends Expr {
+    def print = "(" + e1.print + " + " + e2.print + ")"
+  }
+
+  trait ExprParser {
+    lexical.delimiters += "+"
+    val pLit: PackratParser[Expr] =
+      numericLit ^^ { x => new Lit(x.toInt) }
+    val pAdd: PackratParser[Expr] =
+      pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
+    val pExpr: PackratParser[Expr] =
+      pLit ||| pAdd
+  }
+
+//BEGIN_INHERITANCE_SIMPLE_LAM
+class Var(x: String) extends Expr {
+  def print = x
+}
+//END_INHERITANCE_SIMPLE_LAM
+
+  type Parser[E] = PackratParser[E]
+
+  def parse[E](p: Parser[E]): String => E = in => {
+    val t = phrase(p)(new lexical.Scanner(in))
+    if (t.successful) t.get else scala.sys.error(t.toString)
+  }
+
+//BEGIN_INHERITANCE_BAD_ATTEMPT
+object Attempt extends ExprParser {
+  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
+  val pVarExpr: Parser[Expr] = pExpr ||| pVar
+}
+//END_INHERITANCE_BAD_ATTEMPT
+
+//BEGIN_INHERITANCE_APPROACH
+trait VarExprParser extends ExprParser {
+  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
+  override val pExpr: Parser[Expr] = super.pExpr ||| pVar
+}
+
+parse(new VarExprParser {}.pExpr)("1 + x").print // "(1 + x)"
+//END_INHERITANCE_APPROACH
+
+/*
+//BEGIN_MULTIPLE_INHERITANCE
+trait LanguageA {...}
+
+trait LanguageB {...}
+
+trait LanguageC extends LanguageA with LanguageB {
[…]

### 2017-01-07 paper/paper.tex
Polishing overview & trying a more concise example.
Revising and Splitting Overview
Added S3, inheritance-based approach.
Revised S4.
Revised S5. Related work to be revised.
Fix

@@
     type,var,while,yield,with,val},
-  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
+  otherkeywords={=>,<\%,<:,>:,\#,@},
   sensitive=true,
@@
 
-\input{sections/Overview.tex}
+\input{sections/PackratParsing.tex}
 
-\input{sections/PackratParsers.tex}
-
-\input{sections/OpenRecursion.tex}
+\input{sections/Inheritance.tex}
 
 \input{sections/ObjectAlgebras.tex}
-\input{sections/ParsingLibrary.tex}
 
-% Object Algebras
-% type synonyms, Parser[Expr], Fix[..]
-% more: two pretty-printers OldPretty and NewPretty, they are separated to show modularity.
-% subsection: modify the code using List[E] and type bounds.
-% more subsections: lexing, traits, operations, language composition (more extensibility)
-% use paragraphs
-%...
+\input{sections/Discussion.tex}
 

### 2017-01-07 paper/sections/PackratParsing.tex
Fix

@@
 
-A left-recursive grammar for a simple arithmetic expression language can be written in a straightforward way.
-An alternative grammar, that eliminates left recursion is also presented on the right side:
+The grammar below represents a simple arithmetic language containing only integers. Its corresponding parser is shown on the right side, written in Parsec.
 
-\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
+\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
 \setlength{\grammarindent}{5em}
 \begin{grammar}
-<expr> ::= <int> \alt <expr> `+' <int>
+<expr> ::= <int>
 \end{grammar}
 &
+\begin{lstlisting}[language=PlainCode]
+parseExpr = parseInt
+\end{lstlisting}
+\end{tabular}
+
+Consider we extended the language by adding subtractions to the grammar.
+If we still write its parser by directly following the grammar structure, the new left-recursive case will get the parser into an infinite loop.
+Namely, \inlinecode{parseExpr} and \inlinecode{parseSub} call each other and never stop.
+
+\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::= <int> \alt <expr> `-' <int>
+\end{grammar}
+&
+\begin{lstlisting}[language=PlainCode]
+parseExpr = parseSub <|> parseInt
+parseSub = do
+  e <- parseExpr
+  ...
+\end{lstlisting}
+\end{tabular}
+
+To solve this issue, we have to rewrite the grammar as below to eliminate left-recursion and build the parser based on this new grammar.
+
 \setlength{\grammarindent}{5em}
@@
 
-<expr'> ::= <empty> \alt `+' <int> <expr'>
+<expr'> ::= <empty> \alt `-' <int> <expr'>
 \end{grammar}
-\end{tabular}
 
-The main problem with left-recursion elimination is that it is a
-\emph{global} transformation on a grammar. Given that the full grammar
-is known, then it is possible to remove all left-recursive
-cases. However, when doing modular parsing, the full grammar is not
-known!
-
-For the example above, if we extend the original grammar to support
-subtraction, we must analyse the full grammar again to rewrite it.
-
-\begin{tabular}{m{0.4\linewidth}m{0.4\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> \alt <expr> `+' <int> \alt <expr> `-' <int>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> <expr'>
-
-<expr'> ::= <empty> \alt `+' <int> <expr'> \alt `-' <int> <expr'>
-\end{grammar}
-\end{tabular}
+After left-recursion elimination, the structure of grammar is changed, as well as its corresponding parser.
+In a modular setting, it is possible but unnecessorily compilcated to analyse the grammar and rewrite it when doing extension.
+Anticipating that every non-terminal has left-recursive rules, even though some do not, is helpful for extension.
+However, it is inconvenient and introduces extra complexity for representation of grammar and implementation of parser.
 
@@
 bookkeeping work to retain the original semantics. For example, the
-expression $1+2-3$ is parsed as $(1+2)-3$ in the left-recursive
-grammar, but after rewrite the result is $1((+2)-3))$. The parse tree
+expression $1-2-3$ is parsed as $(1-2)-3$ in the left-recursive
+grammar, but after rewrite the result is $1((-2)-3))$. The parse tree
 must be transformed to recover its structure.
-
-\bruno{Huang, please fix this example ASAP.}
-\bruno{This example does not appear to be problematic: adding ``-''
-  does not affect the rest of the grammar. Isn't the problem that
-  sometimes adding a new case to the grammar, requires changing other
-parts? Otherwise I don't see why left-recursion is a problem, and we
-cannot simply claim that it is a problem. You have to make a better
-effort to find a problematic grammar.
-}
 

### 2017-01-07 Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala

@@ -40,7 +40,7 @@ class Var(x: String) extends Expr {
   }
 
 //BEGIN_INHERITANCE_BAD_ATTEMPT
-object Attempt extends ExprParser {
+trait Attempt extends ExprParser {
   val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
   val pVarExpr: Parser[Expr] = pExpr ||| pVar
 }

### 2017-01-07 paper/sections/Discussion.tex
Organise code into sections, some fixes

@@
 
-\lstinputlisting[linerange=106-111]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=7-12]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
 The above parser shows part of the code, for the space reason we omit details. With OO inheritance, we can actually override \lstinline{pLam}
@@
 
-\lstinputlisting[linerange=115-121]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\lstinputlisting[linerange=16-22]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
 On the other hand, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
@@
 
-\lstinputlisting[linerange=125-131]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=26-32]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
 
@@
 \begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
-\lstinputlisting[linerange=7-22]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=44-59]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
 &
-\lstinputlisting[linerange=26-41]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\lstinputlisting[linerange=63-78]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
 \end{tabular}
@@
 
-\lstinputlisting[linerange=45-54]{../Scala/Parser/src/PaperCode/SampleParser.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=82-91]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 Note that lexing code is automatically combined under multiple inheritance. The combined language shows perfect modularity. The only

### 2017-01-07 paper/sections/ObjectAlgebras.tex
Organise code into sections, some fixes

@@
 an example of operations on the ASTs. Suppose now we also want to collect free variables on the expression language, which models
-literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \lstinline{Expr} with the new
+literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \inlinecode{Expr} with the new
 operation directly:
@@
 It refers to one Scala solution to the Expression Problem []. But then programmers have to define new classes for literals, additions and variables
-by extending the old ones together with \lstinline{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
+by extending the old ones together with \inlinecode{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
 types, again it is a modularity issue, since it is a bad idea to modify existing code again and again with extensions.
@@
 
-\lstinputlisting[linerange=7-10]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
-The language has two constructors: literals and additions. The trait \lstinline{ExprAlg} is called an \textit{Object Algebra interface},
+\lstinputlisting[linerange=5-8]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
+The language has two constructors: literals and additions. The trait \inlinecode{ExprAlg} is called an \textit{Object Algebra interface},
 whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
-type parameter \lstinline{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
+type parameter \inlinecode{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
 provides implementations for all cases. Below is an example of pretty-printing an expression:
 
-\lstinputlisting[linerange=14-17]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
-Here \lstinline{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
+\lstinputlisting[linerange=12-15]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
+Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
 Furthermore, one can define a new trait to implement evaluation on arithmetic expressions:
 
-\lstinputlisting[linerange=21-24]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
+\lstinputlisting[linerange=19-22]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
-On the other hand, the data variants can be extended by inheriting \lstinline{ExprAlg} and adding new cases only. Now we extend the language
-with variables. A new Object Algebra interface \lstinline{VarAlg} is defined as follows:\haoyuan{Can we use var instead of varE?}
+On the other hand, the data variants can be extended by inheriting \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
+with variables. A new Object Algebra interface \inlinecode{VarAlg} is defined as follows:\haoyuan{Can we use var instead of varE?}
 
-\lstinputlisting[linerange=28-30]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+\lstinputlisting[linerange=26-28]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
 Now pretty-printing on the new language can be realized without modifying existing code:
 
-\lstinputlisting[linerange=34-36]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+\lstinputlisting[linerange=32-34]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
 We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
 It offers programmers with great code reuse, in the sense that existing cases can also be
-overridden in \lstinline{VarExprPrint} using \lstinline{"override def"}.
+overridden in \inlinecode{VarExprPrint} using \inlinecode{"override def"}.
 
-On this, to create an expression of \lstinline{VarExprAlg}, a generic method is defined as follows:
+On this, to create an expression of \inlinecode{VarExprAlg}, a generic method is defined as follows:
 
-\huang{todo: better name for build}\haoyuan{makeExp}
-\lstinputlisting[linerange=8-8]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
-Such a method implicitly represents the expression \lstinline{"1 + x"}. The code
+\lstinputlisting[linerange=41-41]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+Such a method implicitly represents the expression \inlinecode{"1 + x"}. The code
 
@@
 \end{lstlisting}
-results in \lstinline{"(1 + x)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
-has the function type \lstinline{VarExprAlg[E] =>} \lstinline{E} for generic \lstinline{E}. At this point, one may argue
+results in \inlinecode{"(1 + x)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
+has the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{E} for generic \inlinecode{E}. At this point, one may argue
 that it is illusory compared to traditional objects, and since we are struggling against the parsing problem, if a parser
@@
 there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
-in a very clever way. For example, if one wants to transform a \lstinline{VarExprAlg} expression to an \lstinline{ExprAlg} expression by
+in a very clever way. For example, if one wants to transform a \inlinecode{VarExprAlg} expression to an \inlinecode{ExprAlg} expression by
 replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}
 
-\lstinputlisting[linerange=12-23]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
-Here the \lstinline{VarExprAlg} expression \lstinline{"1 + x"} is changed to \lstinline{ExprAlg} expression \lstinline{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
-\lstinline{alg} in \lstinline{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
-to \lstinline{alg}, \lstinline{Refactor} will be able to transform a \lstinline{VarExprAlg} expression before we obtain the result.
+\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+Here the \inlinecode{VarExprAlg} expression \inlinecode{"1 + x"} is changed to \inlinecode{ExprAlg} expression \inlinecode{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
+\inlinecode{alg} in \inlinecode{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
+to \inlinecode{alg}, \inlinecode{Refactor} will be able to transform a \inlinecode{VarExprAlg} expression before we obtain the result.
 
@@
 
-Just as shown above, \lstinline{ExprAlg} is defined for a small language in an Object-Algebra style. Furthermore,
+Just as shown above, \inlinecode{ExprAlg} is defined for a small language in an Object-Algebra style. Furthermore,
 A parser can be defined for this whole language at once.
-Since Object Algebras represent ``objects'' implicitly as functions like \lstinline{ExprAlg[E] =>} \lstinline{E} for generic \lstinline{E}, such a parser should have type \lstinline{ExprAlg[E] =>} \lstinline{Parser[E]}, where \lstinline{E} can be abstracted by the enclosing type.
+Since Object Algebras represent ``objects'' implicitly as functions like \inlinecode{ExprAlg[E] =>} \inlinecode{E} for generic \inlinecode{E}, such a parser should have type \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}, where \inlinecode{E} can be abstracted by the enclosing type.
 
-\lstinputlisting[linerange=27-35]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+\lstinputlisting[linerange=9-17]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
 Such a parser works fine, yet it does not appear to be extensible.
-For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \lstinline{pExpr} for an extended syntax, but now \lstinline{pExpr} will have a different type, like \lstinline{VarExprAlg[E] =>} \lstinline{Parser[E]} where the difference appears in the argument.
+For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \inlinecode{pExpr} for an extended syntax, but now \inlinecode{pExpr} will have a different type, like \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} where the difference appears in the argument.
 The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}.
@@
 One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
-\lstinputlisting[linerange=39-48]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=BASE_OA_PARSER
+\lstinputlisting[linerange=24-34]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
-Note that \lstinline{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance.
-The reason why we use an extra and seemingly redundant \lstinline{pExprE}, is that \lstinline{pE} can no longer be referred to by the \lstinline{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \lstinline{pExprE} as the specific \lstinline{pE} in \lstinline{ExprOAParser}.
-And we still want code reuse for the composition of \lstinline{pLit} and \lstinline{pAdd}.
-Now it is worth trying on this pattern with an extension, like \lstinline{VarExprAlg} which we have been using:
-\lstinputlisting[linerange=52-63]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=EXT_OA_PARSER
-In the client code shown above, an instance of algebra is fed to \lstinline{VarExprOAParser} for pretty-printing expressions.
-Then the trait exports \lstinline{pE}, parses the input and applies the algebra.
+Note that \inlinecode{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance.
+The reason why we use an extra and seemingly redundant \inlinecode{pExpr}, is that \inlinecode{pE} can no longer be referred to by the \inlinecode{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \inlinecode{pExpr} as the specific \inlinecode{pE} in \inlinecode{ExprOAParser}.
+And we still want code reuse for the composition of \inlinecode{pLit} and \inlinecode{pAdd}.
+Now it is worth trying on this pattern with an extension, like \inlinecode{VarExprAlg} which we have been using:
+\lstinputlisting[linerange=38-49]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+In the client code shown above, an instance of algebra is fed to \inlinecode{VarExprOAParser} for pretty-printing expressions.
+Then the trait exports \inlinecode{pE}, parses the input and applies the algebra.
 The result shows that such a pattern is a nice solution to modular parsing.
@@
 
-One intuition of using Object Algebras, as illustrated above, is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
-We can simply add new variants corresponding to the new cases in the grammar, or realize different operations on structures, without modifying existing code, meanwhile separate compilation is supported.
-Another one is that Object Algebras benefit us from using generics, where we can abstract multiple syntax in the type parameters, and define less classes, simplifying code comparatively.
+As illustrated above, one intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
+We can simply add new variants corresponding to the new cases in the grammar, and realize different operations on structures, with both code reuse and separate compilation supported.
+
+Another advantage is that Object Algebras benefit us from using generics, where we can abstract multiple syntax in the type parameters, and define less classes, simplifying code comparatively.
 Below we further extend the language with types, by introducing an extra type parameter in the algebra interface.
@@
 \begin{grammar}
-<type> ::= ... \alt `int' \alt <type> `->' <type>
+<type> ::= `int' \alt <type> `->' <type>
 \end{grammar}
@@
 
-Hence we write code for the extended AST, pretty-printing operation and the extended parser:
+We write code for the extended AST, pretty-printing operation and the extended parser:
 
-\lstinputlisting[linerange=67-96]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
-Besides lexing, the trait \lstinline{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
-we use \lstinline{pTypedLamT} and \lstinline{pTypedLamE} as copies of current \lstinline{pT} and \lstinline{pE}, recursively. \lstinline{pT} and \lstinline{pE} are actually the parsers for recursion.
+\lstinputlisting[linerange=32-60]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
+we use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, recursively.
+\inlinecode{pT} and \inlinecode{pE} are actually the parsers for recursion.
 
@@
 
-\lstinputlisting[linerange=100-102]{../Scala/Parser/src/PaperCode/Overview/OA.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=64-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 Pretty-printing the parsed results just works as expected.

### 2017-01-07 paper/sections/PackratParsing.tex
Organise code into sections, some fixes

@@
 
-To solve this issue, we have to rewrite the grammar as below to eliminate left-recursion and build the parser based on this new grammar.
+To solve this issue, we have to rewrite the grammar as below to eliminate left-recursion and build the parser based on this new grammar.\\
 
+\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
 \setlength{\grammarindent}{5em}
@@
 \end{grammar}
+&
+\end{tabular}
 
@@
 
-To shorten our code, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing in the rest of paper.
+For more concise demonstration, in the rest of paper we assume that all the code are in an enclosing environment of traits \inlinecode{StandardTokenParsers} and \inlinecode{PackratParsers}.
+Futhermore, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing as below.
 The last line demonstrates how we parse a valid expression \inlinecode{1 + 2} using our parser.
 
-\lstinputlisting[linerange=36-43]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+\lstinputlisting[linerange=34-41]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER

### 2017-01-07 Scala/Parser/src/PaperCode/ObjectAlg/OA1.scala

@@ -1,50 +1,0 @@
-package PaperCode.ObjectAlg
-
-//BEGIN_OBJECT_ALG_EXPR
-trait ExprAlg[E] {
-  def lit(n: Int): E
-  def add(e1: E, e2: E): E
-}
-//END_OBJECT_ALG_EXPR
-
-//BEGIN_OBJECT_ALG_FREE_VARS
-trait FreeVars extends ExprAlg[Set[String]] {
-  def Var(x: String) = Set(x)
-  def App(e1: Set[String], e2: Set[String]): Set[String] = e1 ++ e2
-}
-//END_OBJECT_ALG_FREE_VARS
-
-//BEGIN_OBJECT_ALG_EXPR_PRINT
-trait ExprPrint extends ExprAlg[String] {
-  def Var(x: String): String = x
-  def App(e1: String, e2: String): String = "(" + e1 + " " + e2 + ")"
-}
-//END_OBJECT_ALG_EXPR_PRINT
-
-//BEGIN_OBJECT_ALG_LIT
-trait LitAlg[E] extends ExprAlg[E] {
-  def Lit(n: Int): E
-}
-//END_OBJECT_ALG_LIT
-
-//BEGIN_OBJECT_ALG_LIT_PRINT
-trait LitPrint extends ExprPrint {
-  def Lit(n: Int): String = n.toString
-}
-//END_OBJECT_ALG_LIT_PRINT
-
-//BEGIN_OBJECT_ALG_REFACTOR
-trait Refactor[E] extends ExprAlg[E] {
-  val alg: LitAlg[E]
-  val env: Map[String, Int]
-  def Var(x: String): E = if (env.contains(x)) alg.Lit(env(x)) else alg.Var(x)
-}
-//END_OBJECT_ALG_REFACTOR
-
-object OA1 {
-
-//BEGIN_OBJECT_ALG_BUILD
-def build[E](alg: LitAlg[E]): E = alg.App(alg.Var("x"), alg.Lit(3))
-//END_OBJECT_ALG_BUILD
-
-}

### 2017-01-07 Scala/Parser/src/PaperCode/Overview/OA.scala

@@ -1,138 +1,0 @@
-package PaperCode.Overview
-
-import PaperCode.Overview.Util._
-
-object OA {
-
-//BEGIN_OVERVIEW_OA_MAKEEXP
-def makeExp[E](alg: VarExprAlg[E]): E = alg.add(alg.lit(1), alg.varE("x"))
-//END_OVERVIEW_OA_MAKEEXP
-
-//BEGIN_OVERVIEW_OA_REFACTOR
-trait Refactor[E] extends VarExprAlg[E] {
-  val alg: ExprAlg[E]
-  val env: Map[String, Int]
-  def lit(n: Int) = alg.lit(n)
-  def add(e1: E, e2: E) = alg.add(e1, e2)
-  def varE(x: String) = alg.lit(env(x))
-}
-
-makeExp(new Refactor {
-  override val alg = new Print {}
-  override val env = Map("x" -> 2)
-}) // "(1 + 2)"
-//END_OVERVIEW_OA_REFACTOR
-
-//BEGIN_BASE_OA_PARSER_BAD
-trait ExprOAParser[E] {
-  lexical.delimiters += ("+")
-  val pLit: ExprAlg[E] => Parser[E] = alg => numericLit ^^ { x => alg.lit(x.toInt) }
-  val pAdd: ExprAlg[E] => Parser[E] = alg => {
-    val p = pExpr(alg)
-    p ~ ("+" ~> p) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
-  }
-  val pExpr: ExprAlg[E] => Parser[E] = alg => pLit(alg) ||| pAdd(alg)
-}
-//END_BASE_OA_PARSER_BAD
-
-//BEGIN_BASE_OA_PARSER
-trait ExprOAParser[E] {
-  lexical.delimiters += ("+")
-  val alg: ExprAlg[E]
-  
-  val pLit: Parser[E] = numericLit ^^ { x => alg.lit(x.toInt) }
-  val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
-  val pExprE: Parser[E] = pLit ||| pAdd
-  
-  val pE: Parser[E] = pLitAdd
-}
-//END_BASE_OA_PARSER
-
-//BEGIN_EXT_OA_PARSER
-trait VarExprOAParser[E] extends ExprOAParser[E] {
-  override val alg: VarExprAlg[E]
-  
-  val pVar: Parser[E] = ident ^^ alg.varE
[…]

### 2017-01-07 Scala/Parser/src/PaperCode/Overview/Overview.scala

@@ -1,24 +1,0 @@
-package PaperCode.Overview
-
-
-object Overview {
-
-trait Expr { def print : String }
-class Lit(x: Int) extends Expr {
-  def print = x.toString
-}
-class Add(e1: Expr, e2: Expr) extends Expr {
-  def print = "(" + e1.print + " + " + e2.print + ")"
-}
-
-//BEGIN_OVERVIEW_ATTEMPT_EXPRWITHEVAL
-trait EvalExpr extends Expr { def eval: Int }
-//END_OVERVIEW_ATTEMPT_EXPRWITHEVAL
-class EvalLit(x: Int) extends Lit(x) with EvalExpr {
-  def eval = x
-}
-class EvalAdd(e1: EvalExpr, e2: EvalExpr) extends Add(e1, e2) with EvalExpr {
-  def eval = e1.eval + e2.eval
-}
-
-}

### 2017-01-07 Scala/Parser/src/PaperCode/ParsingWithOR.scala

@@ -1,105 +1,0 @@
-package PaperCode
-
-import TAPL.Util._
-
-object ParsingWithOR {
-
-//BEGIN_PARSING_PVAR
-type Parser[E] = PackratParser[E]
-
-abstract class Expr
-case class Var(x : String) extends Expr
-val pVar : Parser[Expr] = ident ^^ Var
-//END_PARSING_PVAR
-
-//BEGIN_PARSING_PAPP
-case class App(e1 : Expr, e2 : Expr) extends Expr
-val pApp : Parser[Expr] = pVar ~ pVar ^^ { case e1 ~ e2 => App(e1, e2) }
-//END_PARSING_PAPP
-
-}
-
-object ParsingWithOR2 {
-  
-class Expr
-case class App(e1 : Expr, e2 : Expr) extends Expr
-val pVar : Parser[Expr] = ???
-  
-//BEGIN_PARSING_PAPP2
-val pApp : Parser[Expr] = {
-  val p = pVar ||| pApp
-  p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
-}
-//END_PARSING_PAPP2
-
-}
-
-object ParsingWithOR3 {
-
-class Expr
-case class App(e1 : Expr, e2 : Expr) extends Expr
-val pVar : Parser[Expr] = ???
-
-//BEGIN_PARSING_PLIT
-case class Lit(n : Int) extends Expr
-val pLit : Parser[Expr] = numericLit ^^ { x => Lit(x.toInt) }
-val pApp : Parser[Expr] = {
-  val p = pVar ||| pApp ||| pLit
-  p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
-}
-//END_PARSING_PLIT
-  
-}
-
-object ParsingWithOR4 {
-  
-type Fix[T] = (=> T) => T
-class Expr
-case class App(e1 : Expr, e2 : Expr) extends Expr
-case class Var(x : String) extends Expr
-def merge[E, F, G, H[_,_]](op : F => G => H[F, G], x : (=> E) => F, y : (=> E) => G) : (=> E) => H[F, G] = e => op(x(e))(y(e))
-
-//BEGIN_PARSING_PAPPFIX
-val pApp : Fix[Parser[Expr]] = p => p ~ p ^^ { case e1 ~ e2 => App(e1, e2) }
-//END_PARSING_PAPPFIX
-
-//BEGIN_PARSING_COMBINATOR
[…]

### 2017-01-07 Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala

@@ -30,8 +30,6 @@ object Code2 extends StandardTokenParsers with PackratParsers {
       pLit ||| pAdd
   }
 
-  def main(args: Array[String]): Unit = {
-
 //BEGIN_PACKRAT_RUNPARSER
 type Parser[E] = PackratParser[E]
 
@@ -40,9 +38,10 @@ def parse[E](p: Parser[E]): String => E = in => {
   if (t.successful) t.get else scala.sys.error(t.toString)
 }
 
-parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
+val r = parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
 //END_PACKRAT_RUNPARSER
 
-    println(parse(new ExprParser {}.pExpr)("1 + 2").print)
+  def main(args: Array[String]): Unit = {
+    println(r)
   }
 }

### 2017-01-07 Scala/Parser/src/PaperCode/Sec4OA/Code2.scala

@@ -1,8 +1,6 @@
 package PaperCode.Sec4OA
 
 
-object Code2 {
-
 //BEGIN_OVERVIEW_OA_ALG
 trait ExprAlg[E] {
   def lit(n: Int): E
@@ -36,4 +34,29 @@ trait VarExprPrint extends VarExprAlg[String] with Print {
 }
 //END_OVERVIEW_OA_EXTPRINT
 
+
+object Code2 {
+
+//BEGIN_OVERVIEW_OA_MAKEEXP
+def makeExp[E](alg: VarExprAlg[E]): E = alg.add(alg.lit(1), alg.varE("x"))
+//END_OVERVIEW_OA_MAKEEXP
+
+//BEGIN_OVERVIEW_OA_REFACTOR
+trait Refactor[E] extends VarExprAlg[E] {
+  val alg: ExprAlg[E]
+  val env: Map[String, Int]
+  def lit(n: Int) = alg.lit(n)
+  def add(e1: E, e2: E) = alg.add(e1, e2)
+  def varE(x: String) = alg.lit(env(x))
+}
+
+val r = makeExp(new Refactor[String] {
+  override val alg = new Print {}
+  override val env = Map("x" -> 2)
+}) // "(1 + 2)"
+//END_OVERVIEW_OA_REFACTOR
+
+  def main(args: Array[String]): Unit = {
+    println(r)
+  }
 }

### 2017-01-07 Scala/Parser/src/PaperCode/Sec4OA/Code3.scala

@@ -1,0 +1,60 @@
+package PaperCode.Sec4OA
+
+import scala.util.parsing.combinator.syntactical._
+import scala.util.parsing.combinator._
+
+
+object Code30 extends StandardTokenParsers with PackratParsers {
+//BEGIN_BASE_OA_PARSER_BAD
+trait ExprOAParser[E] {
+  lexical.delimiters += "+"
+  val pLit: ExprAlg[E] => Parser[E] = alg => numericLit ^^ { x => alg.lit(x.toInt) }
+  val pAdd: ExprAlg[E] => Parser[E] = alg => {
+    val p = pExpr(alg)
+    p ~ ("+" ~> p) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
+  }
+  val pExpr: ExprAlg[E] => Parser[E] = alg => pLit(alg) ||| pAdd(alg)
+}
+//END_BASE_OA_PARSER_BAD
+}
+
+
+object Code3 extends StandardTokenParsers with PackratParsers {
+//BEGIN_BASE_OA_PARSER
+trait ExprOAParser[E] {
+  lexical.delimiters += "+"
+
+  val alg: ExprAlg[E]
+
+  val pLit: Parser[E] = numericLit ^^ { x => alg.lit(x.toInt) }
+  val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
+  val pExpr: Parser[E] = pLit ||| pAdd
+
+  val pE: Parser[E] = pExpr
+}
+//END_BASE_OA_PARSER
+
+//BEGIN_EXT_OA_PARSER
+trait VarExprOAParser[E] extends ExprOAParser[E] {
+  override val alg: VarExprAlg[E]
+
+  val pVar: Parser[E] = ident ^^ alg.varE
+  val pVarExpr: Parser[E] = pExpr ||| pVar
+
+  override val pE = pVarExpr
+}
+
+val r = parse(new VarExprOAParser[String] {
+  override val alg = new VarExprPrint {}
+}.pE)("1 + x") // "(1 + x)"
+//END_EXT_OA_PARSER
+
+  def parse[E](p: Parser[E]): String => E = in => {
+    val t = phrase(p)(new lexical.Scanner(in))
+    if (t.successful) t.get else scala.sys.error(t.toString)
+  }
+
+  def main(args: Array[String]): Unit = {
+    println(r)
+  }
+}

### 2017-01-07 Scala/Parser/src/PaperCode/Sec4OA/Code4.scala

@@ -1,0 +1,77 @@
+package PaperCode.Sec4OA
+
+import scala.util.parsing.combinator.syntactical._
+import scala.util.parsing.combinator._
+
+
+object Code4 extends StandardTokenParsers with PackratParsers {
+
+  trait ExprOAParser[E] {
+    lexical.delimiters += "+"
+
+    val alg: ExprAlg[E]
+
+    val pLit: Parser[E] = numericLit ^^ { x => alg.lit(x.toInt) }
+    val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
+    val pExpr: Parser[E] = pLit ||| pAdd
+
+    val pE: Parser[E] = pExpr
+  }
+
+  trait VarExprOAParser[E] extends ExprOAParser[E] {
+    override val alg: VarExprAlg[E]
+
+    val pVar: Parser[E] = ident ^^ alg.varE
+    val pVarExpr: Parser[E] = pExpr ||| pVar
+
+    override val pE = pVarExpr
+  }
+
+
+//BEGIN_OVERVIEW_OA_MULTI_SYNTAX
+trait TypedLamAlg[E, T] extends VarExprAlg[E] {
+  def intT(): T
+  def arrT(t1: T, t2: T): T
+  def lam(x: String, t: T, e: E): E
+}
+
+trait TypedLamPrint extends TypedLamAlg[String, String] with VarExprPrint {
+  def intT() = "int"
+  def arrT(t1: String, t2: String) = t1 + " -> " + t2
+  def lam(x: String, t: String, e: String) = "\\" + x + " : " + t + ". " + e
+}
+
+trait TypedLamOAParser[E, T] extends VarExprOAParser[E] {
+  lexical.reserved += "int"
+  lexical.delimiters += ("->", "\\", ":", ".")
+
+  override val alg: TypedLamAlg[E, T]
+
+  val pIntT: Parser[T] = "int" ^^^ alg.intT
+  val pArrT: Parser[T] = pT ~ ("->" ~> pT) ^^ { case t1 ~ t2 => alg.arrT(t1, t2) }
+  val pTypedLamT: Parser[T] = pIntT ||| pArrT
+
+  val pLam: Parser[E] = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^
+    { case x ~ t ~ e => alg.lam(x, t, e) }
[…]

### 2017-01-08 paper/sections/Inheritance.tex
Organise code into sections, some fixes
Code

@@
 
-\lstinputlisting[linerange=30-32]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
+\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
 
@@
 \huang{Extending from old parser is necessory to get the `pExpr', added a surrounding trait.}
-\lstinputlisting[linerange=43-46]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
+\lstinputlisting[linerange=35-38]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
 
@@
 
-\lstinputlisting[linerange=50-55]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
+\lstinputlisting[linerange=42-47]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
 Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
@@
 
-\lstinputlisting[linerange=60-66]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
+\lstinputlisting[linerange=56-62]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 

### 2017-01-08 paper/sections/ObjectAlgebras.tex
Code

@@
 
-\lstinputlisting[linerange=9-17]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+\lstinputlisting[linerange=8-16]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
 Such a parser works fine, yet it does not appear to be extensible.
@@
 One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
-\lstinputlisting[linerange=24-34]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
+\lstinputlisting[linerange=23-33]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
@@
 Now it is worth trying on this pattern with an extension, like \inlinecode{VarExprAlg} which we have been using:
-\lstinputlisting[linerange=38-49]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+\lstinputlisting[linerange=37-48]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
 In the client code shown above, an instance of algebra is fed to \inlinecode{VarExprOAParser} for pretty-printing expressions.
@@
 
-\lstinputlisting[linerange=32-60]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
@@
 
-\lstinputlisting[linerange=64-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 Pretty-printing the parsed results just works as expected.

### 2017-01-08 Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala

@@ -24,7 +24,7 @@ trait ExprParser extends StandardTokenParsers with PackratParsers {
     numericLit ^^ { x => new Lit(x.toInt) }
   val pAdd: PackratParser[Expr] =
     pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-  val pExpr: PackratParser[Expr] =
+  def pExpr: PackratParser[Expr] =
     pLit ||| pAdd
 }
 //END_PACKRAT_SIMPLE_EXPR

### 2017-01-08 Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala

@@ -1,10 +1,9 @@
 package PaperCode.Sec3Inheritance
 
-import scala.util.parsing.combinator.syntactical._
-import scala.util.parsing.combinator._
+import PaperCode.Common
 
 
-object Code1 extends StandardTokenParsers with PackratParsers {
+object Code1 extends Common {
 
   trait Expr {
     def print: String
@@ -22,7 +21,7 @@ object Code1 extends StandardTokenParsers with PackratParsers {
       numericLit ^^ { x => new Lit(x.toInt) }
     val pAdd: PackratParser[Expr] =
       pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-    val pExpr: PackratParser[Expr] =
+    def pExpr: PackratParser[Expr] =
       pLit ||| pAdd
   }
 
@@ -31,13 +30,6 @@ class Var(x: String) extends Expr {
   def print = x
 }
 //END_INHERITANCE_SIMPLE_LAM
-
-  type Parser[E] = PackratParser[E]
-
-  def parse[E](p: Parser[E]): String => E = in => {
-    val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) t.get else scala.sys.error(t.toString)
-  }
 
 //BEGIN_INHERITANCE_BAD_ATTEMPT
 trait Attempt extends ExprParser {
@@ -49,7 +41,7 @@ trait Attempt extends ExprParser {
 //BEGIN_INHERITANCE_APPROACH
 trait VarExprParser extends ExprParser {
   val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
-  override val pExpr: Parser[Expr] = super.pExpr ||| pVar
+  override def pExpr: Parser[Expr] = super.pExpr ||| pVar
 }
 
 val r = parse(new VarExprParser {}.pExpr)("1 + x").print // "(1 + x)"

### 2017-01-08 Scala/Parser/src/PaperCode/Sec4OA/Code3.scala

@@ -1,10 +1,9 @@
 package PaperCode.Sec4OA
 
-import scala.util.parsing.combinator.syntactical._
-import scala.util.parsing.combinator._
+import PaperCode.Common
 
 
-object Code30 extends StandardTokenParsers with PackratParsers {
+object Code30 extends Common {
 //BEGIN_BASE_OA_PARSER_BAD
 trait ExprOAParser[E] {
   lexical.delimiters += "+"
@@ -19,7 +18,7 @@ trait ExprOAParser[E] {
 }
 
 
-object Code3 extends StandardTokenParsers with PackratParsers {
+object Code3 extends Common {
 //BEGIN_BASE_OA_PARSER
 trait ExprOAParser[E] {
   lexical.delimiters += "+"
@@ -48,11 +47,6 @@ val r = parse(new VarExprOAParser[String] {
   override val alg = new VarExprPrint {}
 }.pE)("1 + x") // "(1 + x)"
 //END_EXT_OA_PARSER
-
-  def parse[E](p: Parser[E]): String => E = in => {
-    val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) t.get else scala.sys.error(t.toString)
-  }
 
   def main(args: Array[String]): Unit = {
     println(r)

### 2017-01-08 Scala/Parser/src/PaperCode/Sec4OA/Code4.scala

@@ -1,10 +1,9 @@
 package PaperCode.Sec4OA
 
-import scala.util.parsing.combinator.syntactical._
-import scala.util.parsing.combinator._
+import PaperCode.Common
 
 
-object Code4 extends StandardTokenParsers with PackratParsers {
+object Code4 extends Common {
 
   trait ExprOAParser[E] {
     lexical.delimiters += "+"
@@ -65,11 +64,6 @@ val r = parse(new TypedLamOAParser[String, String] {
   override val alg = new TypedLamPrint {}
 }.pE)("\\x:int->int. 1 + x") // "\x : int -> int. (1 + x)"
 //END_OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-
-  def parse[E](p: Parser[E]): String => E = in => {
-    val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) t.get else scala.sys.error(t.toString)
-  }
 
   def main(args: Array[String]): Unit = {
     println(r)

### 2017-01-08 paper/paper.tex
Revised abstract for submission

@@
   \emph{modularly type-checked} and \emph{separately compiled}. We
-  present \name: a parsing combinator library that enables modular
+  present a technique based parsing combinators that enables modular
   parsing. Interestingly, the modularity requirements for modular
@@
   parsing techniques, provide solutions for such modularity problems,
-  and enable reasonable performance in a modular setting.  Thus \name
-  is built on top of a Packrat parsing library, but adds new parsing
-  combinators to enable modular parsing. The new parsing combinators
-  employ \emph{delegation-based} techniques and \emph{Object Algebras}
-  to support extensibility. To evaluate \name we conduct a case study
-  based on the ``Types and Programming Languages'' interpreters. The
-  case study shows that \name is effective at reusing parsing code
-  from existing interpreters, and the total parsing code is 60\%
-  shorter than the non-modular parsing code.
+  and enable reasonable performance in a modular setting.
+  Extensibility is achieved using \emph{Object Algebras}.  To evaluate
+  the approach we conduct a case study based on the ``Types and
+  Programming Languages'' interpreters. The case study the
+  effectiveness at reusing parsing code from existing interpreters,
+  and the total parsing code is 68\% shorter than the non-modular
+  parsing code.
 

### 2017-01-08 paper/sections/PackratParsing.tex
Changes

@@
 \end{grammar}
-In Parsec, for instance, the parser
+
+In Parsec, for instance, the parser below will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
+and terminates parsing.
+
 \begin{lstlisting}[language=PlainCode]
@@
 \end{lstlisting}
-will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
-and terminates parsing. With traditional alternative
-composition in recursive descent parsers, the parser in the front successfully parses a substring of the input, then it unexpectedly
-finishes parsing and discards the rest, in spite that latter parsers might be able to parse the whole input.
+
+Using traditional alternative composition, when a preceding parser successfully parses a prefix of the input, it will finish parsing and return the result, in spite that subsequent parsers may be able to parse the whole input.
 %\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
@@
 %}
-In contrast, \inlinecode{parser = parseAdd <|> parseInt} works as expected.
-In this case reordering the components in the alternatives ensures that
-the \emph{longest match} for a string is parsed.
-Yet manual composition for longest match is inconvenient, and worst still, it is essentially non-modular.
+In contrast with the parser above, \inlinecode{parser = parseAdd <|> parseInt} works as expected with the two cases swapped.
+
+In this case, reordering the alternatives ensures that
+the \emph{longest match} is picked among the possible results.
+However, manual reording for the longest match is inconvenient, and worst still, it is essentially non-modular.
 When the grammar is extended with new rules, programmers are supposed to \emph{manually} adjust
-the order of parsers, which is requires rewritting previously written code.
+the order of parsers, which requires rewritting previously written code.
 
@@
 in a modular setting. For example, consider a grammar that includes
-import statements such as \inlinecode{import..from},
-but now we want to extend the grammar with an \inlinecode{import..as} construct, as shown in the third line of grammar below.
+``import'' statements including \inlinecode{import..from}.
+Now we want to extend the grammar with an \inlinecode{import..as} case, as shown in the third line of grammar below.
 
@@
 Given the full grammar, we can decide where to put \inlinecode{try} for backtracking.
-However, with modular parsing we are unable to have a global view of the full grammar, hence the worst case should be considered.
-Namely all alternatives may share common prefixes with future cases.
+However, with modular parsing we are unable to have a global view of the full grammar.
+Hence the worst case should be considered that all alternatives may share common prefixes with future cases.
 In that case we need to backtrack for all the branches.
@@
 
-\noindent However, this results in the worst case exponential time complexity!
+\noindent However, this results in the worst case exponential time complexity, if the library does not have related optimization.
 
 \subsection{Packrat Parsing}\label{subsec:packratparsing}
-Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002}, address the limitations of simple parser combinators such as Parsec.
+Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002} have been developed to address limitations of simple parser combinators.
 Packrat parsers use memoization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated.
-Moreover, theoretically the algorithm behind Packrat parsers supports both direct and indirect left-recursion~\cite{warth2008}.
+Moreover, Packrat parsing supports both direct and indirect left-recursion~\cite{warth2008} theoretically.
 All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in our modular parsing library.
@@
 \cite{moors2008parser} which implements Packrat parsers.
-The library provides a number of parser combinators, together with some basic parsers.
+The library provides a number of parser combinators, including the longeset match alternative combinator.
 Below we present an example to illustrate Scala parsers.
@@
 
-In the trait \lstinline{ExprParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} can parse an integer for \lstinline{Lit},
-whereas \lstinline{pAdd} handles addition and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
+In the trait \lstinline{ExprParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} parses an integer for the literal case.
+\lstinline{pAdd} handles the addition case and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
 recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the longest match alternative combinator \inlinecode{|||}.
@@
 
-It is worth mentioning that such a grammar is left-recursive, yet well supported by Packrat parsers. It also employs automatic
-longest match composition by using the combinator \lstinline{|||} in the library. Furthermore, Packrat parsers do not suffer from the backtracking
-problem, as the technique uses memoization to record intermediate parsing results.
+It is worth mentioning that such a grammar is left-recursive in the addition case.
+Since left-recursion is well supported by Packrat parsers, we do not need extra code for it.
+Our parser also employs longest match composition by using the combinator \lstinline{|||} in the library.
+Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees the efficiency.
 

### 2017-01-08 Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala

@@ -51,6 +51,20 @@ val r = parse(new VarExprParser {}.pExpr)("1 + x").print // "(1 + x)"
     println(r)
   }
 
+  // ---------------------------------------------------
+
+  trait LanguageA {
+    def pExpr: Parser[String] = ident
+  }
+
+  trait LanguageB {
+    def pExpr: Parser[String] = ident
+  }
+
+  trait LanguageC extends LanguageA with LanguageB {
+    override def pExpr = super[LanguageA].pExpr ||| super[LanguageB].pExpr
+  }
+
 /*
 //BEGIN_MULTIPLE_INHERITANCE
 trait LanguageA {...}
@@ -58,7 +72,7 @@ trait LanguageA {...}
 trait LanguageB {...}
 
 trait LanguageC extends LanguageA with LanguageB {
-  override val pExpr = super[LanguageA].pExpr ||| super[LanguageB].pExpr
+  override def pExpr = super[LanguageA].pExpr ||| super[LanguageB].pExpr
 }
 //END_MULTIPLE_INHERITANCE
 */

### 2017-01-09 paper/sections/Inheritance.tex
Changes
Changes on section 3, 4

@@
-\section{Modularizing Parsers with Inheritance}\label{sec:inheritance}
-
-This section introduces the problem that motivates our work. In Figure~\ref{fig:packrat-arith} we have observed
-that Packrat parsers can be defined in a convenient way, with recursive calls. The example is actually inspired
-by Walder's Expression Problem. Similarly, our goal is to build parsers that are extensible in some dimensions.
-
-\subsection{Dimensions of Extensibility}\label{subsec:dimensions}
-We continue with the expression language of literals and additions. At this point,
-one would like to have two kinds of extensions:
-
-%\section{An Overview of \name}\label{sec:overview}
-%
-%This section gives an overview of our library \name, motivates
-%additional problems that occur when parsing modular ASTs.  The first
-%problem is that traditional parsers using parser combinators use
-%hard-coded recursive calls. This problem is solved using
-%delegation-based techniques, which enable parsing OO ASTs.  The second
-%problem is how to achieve dual extensibility of ASTs (that is: how to
-%allow for both new language constructs and new operations over the
-%ASTs). This problem is solved using Object Algebras.
-
 \begin{comment}
@@
 \begin{comment}
-\subsection{Choosing the Parsing Technique}\label{subsec:overview-parsing}
-
-%A technique for type-safe modular parsing should the following 3
-%features: \emph{modular type-checking}; \emph{separate compilation};
-%low performance overhead
-
-In the last section, we have argued that parser combinators are a suitable parsing technique for
-our purpose, as they naturally build modular parsers for type-checking.
-Unfortunately many parser combinators have important limitations.
-In particular several parser combinators,
-including the famous Parsec~\cite{Leijen2001} library, require
-programmers to manually do \textit{left-recursion elimination} and \textit{longest match composition}, and
-require significant amounts of \textit{backtracking}. All of them are
-problematic in a modular setting.
-
-\paragraph{Packrat Parsing}
-Fortunately some more advanced parsing techniques such as Packrat
-parsing~\cite{Ford2002}, address the limitations of simple parser combinators
-such as Parsec. Packrat parsers use
-memoization to record the result of applying each parser at each
-position of the input, so that repeated computation is eliminated.
-Moreover, theoretically the algorithm behind Packrat parsers
-supports both direct and indirect left-recursion~\cite{warth2008}.
-The current version of the library is still
-buggy with indirect left-recursion, but we believe that
-they will fix it in the future, and direct left-recursion is
-already practical to use for a large number of applications. All of these properties are very
-suitable for modularity, thus we decided to use Packrat parsers as the underlying
-parsing technique in \name.
-
 It is worth mentioning that the choice of parser combinators will not
@@
 
-%\huang{done, added the attempt/failure/reason of extending conventional parsers}\bruno{You are not motivating the problem! You are going straight to
-%  the solution without pointing out what the problem is first. What
-%  you need to do is: First show what happens with conventional
-%  parsers: at some point, if you add extensions the recursive calls
-%  will be wrong. Then you show (in the next section) the solution:
-% use delegation/open recursion.}
+\section{Modularizing Parsers with Inheritance}\label{sec:inheritance}
 
-%\huang{partly done, only extend a new language construct in the example, how about leaving extension of new operations in the OA subsection?}\bruno{I think we need to set up a challenge here, similar to the challenge of
-%  the expression problem. The challenge should be like. Build a parser
-%for a simple expression language, then extend the language with both
-%a new language construct and a new operation. This section will show
-%how we can do that using traditional parsers, but without modular
-%type-safety and separate compilation. The remaining sections will show
-%that the techniques introduced by us, enable us to solve those two challenges.}
+This section introduces the problem of semantic modular parsing which motivates our work, and an initial solution using only standard inheritance in OO.
+
+\subsection{Modular Parsing Problem}\label{subsec:parsingproblem}
+\huang{or "Parsing Extensible ASTs" ?}
+The extensibility issues of AST structures and operations that process them, can be illustrated by the famous Expression Problem~\cite{wadler1998expression}. There are two dimensions of extensibility:
 
 \begin{itemize}
-\item \textbf{Extension 1}: adding a new case (or rather, a new construct) to \inlinecode{Expr};
-\item \textbf{Extension 2}: adding a new operation to \inlinecode{Expr}.
+\item \textbf{Extension 1}: adding a new data variant (or rather, a new constructor of ASTs).
+\item \textbf{Extension 2}: adding a new operation over ASTs.
 \end{itemize}
 
-\paragraph{Attempt to Extend the Parser} Now we first consider
-extension 1, namely we are extending the syntax as well as the parser. Hence we introduce variables as a new case.
+We already have solutions such Object Algebras~\cite{Oliveira2012} for this problem.
+However, when ASTs evolve, the corresponding parsers which produce such structures should also change accordingly.
+Our motivation is to build modular parsers, which can be extended together with ASTs.
+Futhermore, we focus on semantic modularity, that means we expect parsers to be modularly type-checked and separately compiled.
+
+In the rest of Section~\ref{sec:inheritance}, we will introduce an solution for modular parsing with Extension 1 above, using only standard inheritance in OO.
+Extension 2 will be discussed in Section~\ref{sec:algebrasandparsing}.
+
+\subsection{Inheritance-Based Parser}
+
+Let's firstly consider Extension 1, which is to extend the syntax as well as the parser.
+As an example, we continue with the expression language of literals and additions in Section~\ref{subsec:packratparsing}.
+We introduce variables as a new case as below.
 
@@
 
-As in an OO solution to the Expression Problem, it is easy to extend
-the corresponding AST:
+It is easy to extend the corresponding AST in an OO way:
 
@@
 
-And since we already have the parser for literals and additions, we would
-like to build the new parser by reusing the old one. Here one may quickly come up
+Since we already have the parser for literals and additions, we would
+like to build the new parser by reusing the old one. In Scala we can take advantage of inheritance for such reuse. Specifically,
+the new parser can be defined in an enclosing trait that extends
+the old one. Here one may quickly come up
 with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
@@
 
-\huang{Extending from old parser is necessory to get the `pExpr', added a surrounding trait.}
 \lstinputlisting[linerange=35-38]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
@@
 covers both cases in the old grammar. But the newly added case \inlinecode{pVar} is not observed by the recursive \inlinecode{pExpr},
-hence the parser does not work as expected. Still, one may continue to change the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr},
-but modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.
+hence the parser does not work as expected. It is possible to build the correct parser by replacing the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr}.
+However, modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.
 
-\subsection{Inheritance-Based Approach}\label{subsec:inheritance-approach}
+\subsection{Overriding for Extensibility}
 
-\huang{It seems a little bit weird since we were already using inheritance in the `bad attempt'. I think we could focus on overriding here}
-
-Fortunately in Scala we can make use of inheritance for modular parsing. Specifically,
-the new parser can be defined in an enclosing trait that extends
-the old one:
+It is actually quite simple to let \inlinecode{pExpr} cover the newly extended case without modifying existing code.
+Method overriding is a standard feature which often comes with inheritance, and it allows us to explain inherited method again, such as \inlinecode{pExpr}.
+We can build the new parser which correctly parses \inlinecode{"1 + x"} as below.
 
 \lstinputlisting[linerange=42-47]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
+
 Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
@@
 the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
-actually refers to \inlinecode{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
-elegant inheritance-based approach opens the recursion of parsers in a modular way.
+actually refers to \inlinecode{this.pExpr} again due to dynamic dispatch, and it covers the variable case. Similarly, all recursive calls can be updated to include new extensions if needed.
 
@@
 
-\lstinputlisting[linerange=56-62]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
+\lstinputlisting[linerange=70-76]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 
 As we demonstrated, inheritance is the key technique to obtain semantic modularity.
-It enables type-safe code reuse and separate compilation for parsing OO style ASTs with extension 1.
+It enables type-safe code reuse and separate compilation for parsing OO style ASTs with Extension 1.

### 2017-01-09 paper/sections/ObjectAlgebras.tex
Changes on section 3, 4

@@
 The inheritance-based approach benefits us from building extensible parsers, based on an OO class hierarchy.
-Nevertheless, the addition of new operations over the ASTs becomes problematic. Often it requires modification on
-the existing code, which breaks modularity. In this section, we will have a closer look at how it motivates us to
-use Object Algebras for full extensibility.
+Nevertheless, the addition of new operations over ASTs is problematic using traditional OO classes.
+Often it requires modification on existing code, which breaks modularity.
+In this section, we will have a closer look at how Object Algebras is applied for full extensibility.
 
@@
 
-Now it comes to extension 2 in Section~\ref{subsec:dimensions}. In the previous code, we have presented pretty-printing as
-an example of operations on the ASTs. Suppose now we also want to collect free variables on the expression language, which models
+Now it comes to Extension 2 in Section~\ref{subsec:parsingproblem}. In the previous code, we have presented pretty-printing as
+an example of operations on ASTs. Suppose we also want to collect free variables on the expression language which models
 literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \inlinecode{Expr} with the new
@@
 \lstinputlisting[linerange=19-19]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
-It refers to one Scala solution to the Expression Problem []. But then programmers have to define new classes for literals, additions and variables
+
+It refers to one Scala solution to the Expression Problem.\huang{what's the solution? is it successful?} But then programmers have to define new classes for literals, additions and variables
 by extending the old ones together with \inlinecode{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
@@
 
-In contrast, Object Algebras~\cite{Oliveira2012} enables us to solve this problem. As a nice solution to the Expression Problem, it separates data variants
-and operations, and offers high flexibility in the choice of operations to be performed over the AST. Before adopting it for parsing, the next section introduces the background of Object Algebras.
+In contrast, Object Algebras~\cite{Oliveira2012} enables us to solve this problem. As a nice solution, it separates data variants
+and operations, and offers high flexibility in the choice of operations to be performed over ASTs. Before adopting it for parsing, the next section introduces the background of Object Algebras.
 
 \subsection{Object Algebras}\label{subsec:objectalgebras}
-Object Algebras captures a design pattern to address the Expression Problem nicely,
+\huang{to polish}Object Algebras captures a design pattern to address the Expression Problem nicely,
 achieving two dimensions of extensibility (data variants and operations) in a modular and type-safe way.
@@
 
-In Object Algebras, ASTs as recursive data structures are defined using traits, where each constructor corresponds
-to an abstract method inside. The example from Section~\ref{subsec:packratparsing} is again used here for illustration.
+Using Object Algebras in Scala, ASTs as recursive data structures are defined by traits, where each constructor corresponds to an abstract method inside.
+The example from Section~\ref{subsec:packratparsing} is used here again for illustration.
 At first the language only supports literals and additions:
@@
 \lstinputlisting[linerange=5-8]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
-The language has two constructors: literals and additions. The trait \inlinecode{ExprAlg} is called an \textit{Object Algebra interface},
+We have two constructors \inlinecode{lit} and \inlinecode{add}, corresponding to literals and additions. The trait \inlinecode{ExprAlg} is called an \textit{Object Algebra interface},
 whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
@@
 Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
-Furthermore, one can define a new trait to implement evaluation on arithmetic expressions:
+One can define evaluation operation as a new trait on this expression langauge:
 
@@
 
-On the other hand, the data variants can be extended by inheriting \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
-with variables. A new Object Algebra interface \inlinecode{VarAlg} is defined as follows:\haoyuan{Can we use var instead of varE?}
+Futhermore, the data variants can be extended by inheriting \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
+with variables. A new Object Algebra interface \inlinecode{VarAlg} is defined as follows:
 
@@
 We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
-It offers programmers with great code reuse, in the sense that existing cases can also be
-overridden in \inlinecode{VarExprPrint} using \inlinecode{"override def"}.
+It offers programmers great code reuse, in the sense that existing cases can also be
+overridden in \inlinecode{VarExprPrint}.
 
-On this, to create an expression of \inlinecode{VarExprAlg}, a generic method is defined as follows:
+To create an expression of \inlinecode{VarExprAlg}, a generic method is defined as follows:
 

### 2017-01-09 paper/sections/Discussion.tex
Polishing section 4 and 5

@@
-\section{Discussion}\label{sec:discussion}
+\section{More Features}
 
-The use of interface-based approach and Object Algebras enables us to do modular parsing. Programmers can easily follow the pattern
-captured in the previous section, build modular ASTs and hence have access to multiple dimensions of extensibility. This section
-explores more interesting features of this pattern, including overriding existing parsing rules, and framework of language components
-for modularity over hierarchies.
+The use of inheritance-based approach and Object Algebras enables us to build modular parsers, which are able to evolve with syntax together.
+This section explores more interesting features, including overriding existing parsing rules, language components for abstracting language features, and alternative techniques under the whole framework.
 
@@
 \lstinputlisting[linerange=16-22]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
-On the other hand, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
+Futhermore, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
 One can even ``eliminate'' a production rule in the extension, by actually overriding it with a failure parser.
@@
 
-\subsection{Framework of a Language Component}\label{subsec:framework}
+\subsection{Language Component}
 
@@
 \item \textbf{Parser:} implementation of the parser with lexing, using inheritance and Object Algebras, with its scoping trait abstracted over the type parameters.
-\item \textbf{Algebra (optional):} some behaviors (operations) on the data structure, such as pretty-printing and so on.
+\item \textbf{Algebras (optional):} concrete operations on ASTs, such as pretty-printing.
 \end{itemize}
@@
 
-\subsection{Fusion of Concepts}\label{subsec:fusion}
-
-\haoyuan{Not sure if this section should be here, or in the related work. I tend to have a big discussion section.}
-\huang{These alternatives could be discussed in the discussion
-  section, not here. 6.1 could only talks about library code. And the
-  titles should be changed as well}\bruno{I think this subsection is
-  unnecessary. The code will be presented already at the end of S5, I
-  presume. }
-
+\subsection{Alternative Techniques}
+\huang{to rewrite}
 In the previous sections we have talked about Packrat Parsing, open recursion and Object Algebras.

### 2017-01-09 paper/sections/ObjectAlgebras.tex
Polishing section 4 and 5

@@
-\section{Extensible ASTs with Object Algebras}\label{sec:algebrasandparsing}
+\section{Full Extensibility with Object Algebras}\label{sec:algebrasandparsing}
 
@@
 
-Now it comes to Extension 2 in Section~\ref{subsec:parsingproblem}. In the previous code, we have presented pretty-printing as
+Now it comes to Extension 2 in Section~\ref{subsec:parsingproblem}. In the previous code, we have pretty-printing as
 an example of operations on ASTs. Suppose we also want to collect free variables on the expression language which models
@@
 
-It refers to one Scala solution to the Expression Problem.\huang{what's the solution? is it successful?} But then programmers have to define new classes for literals, additions and variables
+It refers to one Scala solution to the Expression Problem.\huang{what's the solution?
+does it have the later "critical issue"?
+if so, why is it a "solution"} But then programmers have to define new classes for literals, additions and variables
 by extending the old ones together with \inlinecode{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
@@
 \subsection{Object Algebras}\label{subsec:objectalgebras}
-\huang{to polish}Object Algebras captures a design pattern to address the Expression Problem nicely,
-achieving two dimensions of extensibility (data variants and operations) in a modular and type-safe way.
-Because of this, the definition of data structures is separated from behaviors on them, and future extensions
-to both sides no longer require existing code to be modified, supporting separate compilation.
+Object Algebras captures a design pattern to address the Expression Problem nicely, achieving two dimensions of extensibility (data variants and operations) in a modular and type-safe way.
+The definition of data structures is separated from their behaviors, and future extensions on both dimensions no longer require existing code to be modified, supporting separate compilation.
 
@@
 \lstinputlisting[linerange=32-34]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
-We observe that only the new case is implemented for pretty-printing, and the others have been inherited.
+An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
 It offers programmers great code reuse, in the sense that existing cases can also be
@@
 
-To create an expression of \inlinecode{VarExprAlg}, a generic method is defined as follows:
+To create an expression representing \inlinecode{1 + x}, a generic method is defined as follows:
 
 \lstinputlisting[linerange=41-41]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
-Such a method implicitly represents the expression \inlinecode{"1 + x"}. The code
+
+The code below results in \inlinecode{"(1 + x)"}, as the result of pretty-printing that expression.
 
@@
 \end{lstlisting}
-results in \inlinecode{"(1 + x)"}, as the result of pretty-printing. Now we observe that an ``object'' of expression actually
-has the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{E} for generic \inlinecode{E}. At this point, one may argue
-that it is illusory compared to traditional objects, and since we are struggling against the parsing problem, if a parser
-produces structures as functions, it could be hardly be further processed like objects, especially, in the design of a compiler
+
+Now we observe that an ``object'' of expression actually
+has the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{E} for generic \inlinecode{E}.
+Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
+Especially, in the design of a compiler
 there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
@@
 
-Just as shown above, \inlinecode{ExprAlg} is defined for a small language in an Object-Algebra style. Furthermore,
-A parser can be defined for this whole language at once.
-Since Object Algebras represent ``objects'' implicitly as functions like \inlinecode{ExprAlg[E] =>} \inlinecode{E} for generic \inlinecode{E}, such a parser should have type \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}, where \inlinecode{E} can be abstracted by the enclosing type.
+Just as shown above, \inlinecode{ExprAlg} is defined for a small language using Object Algebras.
+Since Object Algebras represent ``objects'' implicitly as functions like \inlinecode{ExprAlg[E] =>} \inlinecode{E} for abstract \inlinecode{E}, we first try to build the corresponding parser by type \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.
 
@@
 Such a parser works fine, yet it does not appear to be extensible.
-For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \inlinecode{pExpr} for an extended syntax, but now \inlinecode{pExpr} will have a different type, like \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} where the difference appears in the argument.
-The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}.
-\haoyuan{Shall we mention this?}
+For example, we have demonstrated in Section~\ref{subsec:overriding} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
+However, now \inlinecode{pExpr} has type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}, where the type of Object Algebra interface appears in \emph{contravariant} position.
+This difference prohibits method overriding, because Object Algebra interfaces extended from \inlinecode{VarExprAlg} are subtypes of it, but the types of their parsers are supertypes of \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}.
 
-One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
+This problem refers to \textit{binary methods}. \haoyuan{Shall we mention this?}\huang{???}
+
+One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. We rewrite the parsing code as follows:
+
 \lstinputlisting[linerange=23-33]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
-Note that \inlinecode{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance.
-The reason why we use an extra and seemingly redundant \inlinecode{pExpr}, is that \inlinecode{pE} can no longer be referred to by the \inlinecode{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \inlinecode{pExpr} as the specific \inlinecode{pE} in \inlinecode{ExprOAParser}.
-And we still want code reuse for the composition of \inlinecode{pLit} and \inlinecode{pAdd}.
-Now it is worth trying on this pattern with an extension, like \inlinecode{VarExprAlg} which we have been using:
+Note that \inlinecode{pE} is the one that is used for recursive calls, exported by the trait and considered to be overridden during inheritance.
+The reason why we use an extra and seemingly redundant \inlinecode{pExpr}, is a subtle issue caused by Scala language and its parser combinator library.
+There is a restriction of \inlinecode{super} keyword in Scala that \inlinecode{super} can only use fields defined by keyword \inlinecode{def}, can not use those defined by \inlinecode{val}, while the parser combinator library suggests using \inlinecode{val} to define parsers, especially for left-recursive ones.
+Our workaround is that we use different synonyms for \inlinecode{pE} in different traits, so that we can directly distinguish them by names without using \inlinecode{super}.
+
+Now let's try this pattern with the extension of varibles, which we have been using:
 \lstinputlisting[linerange=37-48]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
@@
 At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice.
-This is generally solved in the original Object Algebras paper [] by composing two algebras into one in advance.
+However, that is generally solved~\cite{Oliveira2012} \huang{it was "original object algebra paper", i added the citation here, is it the correct one?
+i'll check later} by composing two algebras into one in advance.
 
-\subsection{Parsing with Multiple Syntax}\label{subsec:differentsyntax}
+\subsection{Parsing Multi-sort Syntax}\label{subsec:differentsyntax}
 
-As illustrated above, one intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
+\huang{to polish}As illustrated above, one intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
 We can simply add new variants corresponding to the new cases in the grammar, and realize different operations on structures, with both code reuse and separate compilation supported.
 
-Another advantage is that Object Algebras benefit us from using generics, where we can abstract multiple syntax in the type parameters, and define less classes, simplifying code comparatively.
-Below we further extend the language with types, by introducing an extra type parameter in the algebra interface.
-The extended grammar is:\\
+Another advantage of using Object Algebras is that we benefit from using generics.
+If the syntax contains multiple sorts, we can distinguish them by different type parameters.
+Thus fewer classes are defined and code is simplified.
+As an example, we extend the language of literals, additions and variables by introducing types as a new sort.
+Now it has two sorts in the syntax, types and expressions.
+We certainly do not want to mix them together.\\
 
@@
 
-We write code for the extended AST, pretty-printing operation and the extended parser:
+When extending the Object Algebra interface, pretty-printing operation and parser, we use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
+They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
 
@@
 \lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-Pretty-printing the parsed results just works as expected.
+It works just as expected to parse and pretty-print the result.

### 2017-01-09 Scala/Parser/src/PaperCode/Sec4OA/Code2.scala

@@ -45,9 +45,9 @@ def makeExp[E](alg: VarExprAlg[E]): E = alg.add(alg.lit(1), alg.varE("x"))
 trait Refactor[E] extends VarExprAlg[E] {
   val alg: ExprAlg[E]
   val env: Map[String, Int]
-  def lit(n: Int) = alg.lit(n)
-  def add(e1: E, e2: E) = alg.add(e1, e2)
-  def varE(x: String) = alg.lit(env(x))
+  def lit(n: Int): E = alg.lit(n)
+  def add(e1: E, e2: E): E = alg.add(e1, e2)
+  def varE(x: String): E = alg.lit(env(x))
 }
 
 val r = makeExp(new Refactor[String] {

### 2017-01-09 Scala/Parser/src/PaperCode/Sec4OA/Code3.scala

@@ -40,7 +40,7 @@ trait VarExprOAParser[E] extends ExprOAParser[E] {
   val pVar: Parser[E] = ident ^^ alg.varE
   val pVarExpr: Parser[E] = pExpr ||| pVar
 
-  override val pE = pVarExpr
+  override val pE: Parser[E] = pVarExpr
 }
 
 val r = parse(new VarExprOAParser[String] {

### 2017-01-09 Scala/Parser/src/PaperCode/Sec5/Code1.scala

@@ -64,7 +64,7 @@ object TypedLam {
   // Abstract syntax
   trait Alg[E, T] {
     def intT(): T
-    def arrowT(t1: T, t2: T): T
+    def arrT(t1: T, t2: T): T
     def lam(x: String, t: T, e: E): E
   }
 

### 2017-01-09 paper/sections/Conclusion.tex
Related work ready for review.

\section{Conclusion}\label{sec:conclusion}

Huang: to write a couple of sentences for a summary.

\noindent Haoyuan: to discuss future work. (some boilerplate can be avoided via metaprogramming.)

### 2017-01-09 paper/sections/Discussion.tex
Related work ready for review.

@@
 \subsection{Alternative Techniques}
-\huang{to rewrite}
-In the previous sections we have talked about Packrat Parsing, open recursion and Object Algebras.
-These components fuse together in our library implementation, for the initial goal, which is to build extensible parsers.
+In the previous sections we have talked about Packrat Parsing, inheritance and Object Algebras.
+These components fuse together in our library implementation, for building extensible parsers.
 We want to argue here that those aspects are totally orthogonal, and hence can have their alternatives.
-For example, Object Algebras are helpful to build extensible data structures, whereas in functional programming, there has been a large amount of related work on extensible datatypes, including DTC and MRM for the Haskell language.
-And a better
+
+For example, we have conducted some previous experiments of modular paring on the Haskell language, based on open recursion, MRM [] and Parsec.
+\haoyuan{I mention our previous work here.
+} The delegation-based open recursion can be an alternative for inheritance.
+It exposes recursion explicitly by a parameter for ``self-reference''.
+Furthermore, there has been a lot of related work that realizes extensible data types in functional programming, including MRM using a list of functors.
+We used Parsec as the parsing library, but suffered from the issues already illustrated in Section~\ref{subsec:challenges}.
+As a workaround, we made use of state monad in Parsec to ensure that a left-recursive production rule will not be applied successively in parsing.
+Whereas a better
 parsing library would potentially introduce a lot of fancy features and parser combinators, run with high efficiency, or even support indirect left-recursion.
-\haoyuan{Open recursion could also have alternatives?} For now, our implementation integrates the existing ones, and has turned out to be practical in various applications.

### 2017-01-09 paper/sections/PackratParsing.tex
Changes on section 3, 4
Polishing section 4 and 5
Related work ready for review.

@@
 
-\subsection{Algorithmic Challenges of Modularity}
+\subsection{Algorithmic Challenges of Modularity}\label{subsec:challenges}
 At a first look, parser combinators are very suitable for modular parsing, because of two reasons.
@@
 
-\noindent However, this results in the worst case exponential time complexity, if the library does not have related optimization.
+\noindent However, this results in the worst case exponential time complexity in Parsec, because it does not have related optimization.
 

### 2017-01-09 paper/sections/RelatedWork.tex
Related work ready for review.

@@
 parser combinators and extensibility techniques. There has been a
-great amount of related papers on those hot topics, of which some
-inspired us of this paper and encourages us for more exploration. This
+great amount of related papers on those hot topics. Some
+inspired us of this paper and encourage us for more exploration. This
 section will try to lead a discussion on what difference we have made.
@@
 \paragraph{Syntactically Extensible Parsing} Extensible parsing is achieved in many
-different areas, of which parser generators are a mainstream area
+different areas, among them parser generators are a mainstream area
 specially designed for modular syntax and parsing. Many parser
@@
 moreover, the implementation was based on SDF~\cite{Heering1989} and
-Stratego~\cite{Visser2001}, which focused little on separate
-compilation\bruno{focused little or did not have it at all?}.\haoyuan{The authors have
+Stratego~\cite{Visser2001}, which did not focus on separate
+compilation.\bruno{focused little or did not have it at all?}.\haoyuan{The authors have
 mentioned separate compilation as a possible yet unimportant feature; they decided to discard it.}
-Extensible compilers like JastAdd~\cite{Ekman2007} and
+Racket adopts macro system for library-based language extensibility. It is powerful with attributed
+ASTs for contextual information, and extensions can be integrated in a modular way. Whereas
+such modularity is not flexible enough for language unification, as the syntax is only built from extensions.
+Moreover, existing macros cannot be further changed after definition. \haoyuan{correct?}
+On the other hand, extensible compilers like JastAdd~\cite{Ekman2007} and
 Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
@@
 is designed for modular language parsers in a type-safe way, with
-flexible language composition. Although overriding existing production
-rules is tricky with our approach, we could perhaps make use of
-embedded domain-specific languages on top of parser combinators and
-transformations for overriding, but it is anyway an orthogonal issue.
-\bruno{Not sure that I understand the last sentence}
-\haoyuan{???} \haoyuan{BTW what about Racket?} \bruno{Have you read
-Languages as Libraries? That is probably an important reference, which
-I think we should cite.}\haoyuan{what about
-  metafront? it is a macro system but does it have type safety?}
-\haoyuan{Extensible syntax with lexical scoping?} \haoyuan{"Extensible
-  syntax" proposes a system for extensible syntax, where users write
-  EDSLs in their language with concrete syntax. Users can write rules
-  for type-based disambiguation. But separate compilation is again not
-  mentioned. Shall we mention that thesis?} \haoyuan{attribute
-  grammars?}
+flexible language composition.
+%\bruno{Have you read
+%Languages as Libraries? That is probably an important reference, which
+%I think we should cite.}
+%\haoyuan{what about
+%  metafront? it is a macro system but does it have type safety?}
+%\haoyuan{Extensible syntax with lexical scoping?} \haoyuan{"Extensible
+%  syntax" proposes a system for extensible syntax, where users write
+%  EDSLs in their language with concrete syntax. Users can write rules
+%  for type-based disambiguation. But separate compilation is again not
+%  mentioned. Shall we mention that thesis?} \haoyuan{attribute
+%  grammars?}
 
@@
 separate compilation. Specifically, \cite{Bravenboer2009} introduces
-\textit{parse table composition}, where grammars\bruno{what
-  is a paper grammar?}\haoyuan{Revised} are compiled
-to the generation of parse tables\bruno{I don't understand the last
-  part; please rephrase. Also please break the sentence into multple
-  sentences. Generally speaking go over the text you wrote and watch
-  out for long sentences. You have a tendency to write very long
-  sentences. }, which are DFAs or NFAs, later they
+\textit{parse table composition}, where grammars are compiled
+to the generation of parse tables. Those parse tables are expressed as DFAs or NFAs, later they
 can be composed by an algorithm, so as to provide separate compilation
 for parsing. Nevertheless, the generation of parse tables is rather
-expensive\bruno{expensive in terms of what? why is it expensive?}.
-\haoyuan{in terms of performance. The sentence is copied from original paper.}
-Furthermore, \name supports \emph{both} separate compilation
+expensive in terms of performance.
+Furthermore, our approach supports \emph{both} separate compilation
 as well as \emph{modular type-checking}. The extensibility of
-parsing in our approach is further available at language composition
-and lexical level. \haoyuan{I mentioned one shortcoming of our
-  approach here, which is we cannot override existing production
-  rules; another one is that we do not have explicit
-  correspondence/relationship between abstract syntax and the parser.}
+parsing is further available at language composition
+and lexical level.
+%\haoyuan{we do not have explicit
+%  correspondence/relationship between abstract syntax and the parser.}
 
@@
 for context-sensitive grammars with infinite lookahead. Nevertheless,
-left-recursion is known as a big issue in recursive descent parsers,
-in which case programmers using libraries like Parsec have to manually
-eliminate left-recursion in the grammar, which is cumbersome, but also
-it distorts the shape of its old grammar, restricting the modularity
-of parsing to extreme extent. It is possible to avoid some of
- the limitations of parser combinators.
-\bruno{I'm not sure that th following text should be here. It could
-be in the paper, but somewhere else.}
-For example, we have
-conducted our previous experiment in Haskell, by using open recursion,
-MRM~\cite{Oliveira2015} and Parsec. To support direct left-recursive
-grammars, we made use of the state monad in Parsec to ensure that a
-left-recursive production rule will not be applied successively in
-parsing. It turned out to be successful, yet introduced complexity in
-programming, hence we believe that the parsing library is obligated
-for left-recursion.
+Parsec users suffer from manual left-recursion elimination, back-tracking and
+longest match composition issues, as we discussed before.
 
-Some more recent papers~\cite{Ford2002,Might2011,Frost2008}
-\haoyuan{Pracitical, general parser combinators?} proposed a series of
+Some more recent papers~\cite{Ford2002,Might2011,Frost2008} proposed a series of
 novel parsing techniques that addressed the issue
@@
 parsers practically, though~\cite{warth2008} further demonstrated that
-general left-recursion is supported from the theoretical point of
-view. We have also mentioned that the components used in our library
-can have alternatives, for instance, a different set of parser
-combinators may support ambiguous and left-recursive grammars, or be
-applicable to a different subset of context-free grammars or more, or
-even lead us to another level of performance.
+general left-recursion is supported theoretically.
+Yet we have mentioned that the components used in our approach
+can have alternatives to achieve better modular parsing.
 
-To the best of our knowledge, no previous work on parser combinators
-deals with semantic modularity and
-extensibility of parsing. The modularity of abstract syntax trees is
-necessary for modular parsing that supports separate compilation and
-modular type-checking. In our library, Object Algebras have been
-adopted for extensible ASTs, together with the behaviors on them.
-
-\paragraph{Extensibility} There has been alot of work on building
-modular programs, in both functional and object-oriented programming .
+\paragraph{Extensibility} To the best of our knowledge, none of the existing techniques
+address the problem of producing extensible ASTs. Certainly, there has been a lot of work on extensible
+data types, in both functional and object-oriented programming.
 For instance, the famous DTC paper~\cite{swierstra2008} represents
 modular ASTs using co-products of every two functors, which forms a
-binary tree structure. It supports extensible datatypes, as well as
-the operations on them by implementing instances of functors. Later
+binary tree structure. Later
 Oliveira et al. presented MRM~\cite{Oliveira2015} in Haskell, which is
 similar to DTC but adopts a list of functors in type-level lists
-instead. As mentioned above, we have conducted our early experiment
-with MRM for modular datatypes, whereas the complexity of library
-implementation was comparably high. Other related work
+instead. Other related work
 includes~\cite{Bahr2014}, where data types are flexible with
@@
 discussed the composition of a cv-coalgebra and a term homomorphism,
-which differs from modular parsing. \haoyuan{From Object Algebras to
-  Attribute Grammars?}
+which differs from modular parsing.
 
@@
 boilerplate code is automatically generated, hence users only need to
-override a few interesting cases, conveniently. The idea of overriding
-existing cases was also proposed in MRM. It is potentially useful to
-our framework.\bruno{The discussion on Extensibility can be
-  considerably shorter. We just need to point out that there are many
-  techniques for extensibility (without discussing alot of details),
-  but most of that work does not address the problem of
-\emph{producing extensible ASTs} (rather than consuming those. Then
-you can talk about Baht's work, and have the discussion on
-transformations. Generally speaking, go over your text, polish it, and
-trim it. You are a bit too ``chatty'' (that is you talk too much): go
-straight to the point.}
+override a few interesting cases, conveniently. It is potentially useful
+in our case. By the way, the idea of overriding
+existing cases was also proposed in MRM. Our approach allows existing parsing
+code to be overridden during inheritance.

### 2017-01-10 paper/sections/RelatedWork.tex
Mention some future work in conclusion.

@@
 expensive in terms of performance.
-Furthermore, our approach supports \emph{both} separate compilation
+Furthermore, our approach supports both \emph{separate compilation}
 as well as \emph{modular type-checking}. The extensibility of
@@
 Parsing for our prototype implementation, as the combinators are
-convenient to use in Scala, which is a platform where functional
-laziness and Object Algebras can perfectly coexist. Also, its support
+convenient to use in Scala, and Object Algebras can be naturally realized with
+OO inheritance. Also, its support
 for direct left-recursion is already helpful for designing real-world
@@
 
-\paragraph{Extensibility} To the best of our knowledge, none of the existing techniques
+\paragraph{Extensibility} To the best of our knowledge, none of the existing parsing techniques
 address the problem of producing extensible ASTs. Certainly, there has been a lot of work on extensible

### 2017-01-10 paper/sections/PackratParsing.tex
Revised S2

@@
 
-After left-recursion elimination, the structure of grammar is changed, as well as its corresponding parser.
-In a modular setting, it is possible but unnecessorily compilcated to analyse the grammar and rewrite it when doing extension.
-Anticipating that every non-terminal has left-recursive rules, even though some do not, is helpful for extension.
-However, it is inconvenient and introduces extra complexity for representation of grammar and implementation of parser.
+After left-recursion elimination, the structure of grammar is changed,
+as well as its corresponding parser. In a modular setting, it is
+possible but unnecessarily compilcated to analyse the grammar and
+rewrite it when doing extension. Anticipating that every non-terminal
+has left-recursive rules is helpful for extensibility but overkill,
+since it is inconvenient and introduces extra complexity for
+representation of grammar and implementation of parser.
 
@@
 \subsection{Packrat Parsing}\label{subsec:packratparsing}
-Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002} have been developed to address limitations of simple parser combinators.
-Packrat parsers use memoization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated.
-Moreover, Packrat parsing supports both direct and indirect left-recursion~\cite{warth2008} theoretically.
-All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in our modular parsing library.
+Fortunately some more advanced parsing techniques such as Packrat
+parsing~\cite{Ford2002} have been developed to address limitations of
+simple parser combinators. Packrat parsers use memoization to record
+the result of applying each parser at each position of the input, so
+that repeated computation is eliminated. Moreover, Packrat parsing, in
+theory, supports both direct and indirect left-recursion~\cite{warth2008}.
+All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in for modular parsing.
 \begin{comment}
@@
 \lstinline{E}, which indicates the type of results it produces. Figure~\ref{fig:packrat-arith} shows the abstract syntax and
-the parser code that generates expressions as ASTs. Meanwhile, pretty-printing shows up as an operation on the AST.
+the parser code that generates expressions as ASTs. The AST supports
+pretty-printing via the \lstinline{print} method.
 
@@
 
-It is worth mentioning that such a grammar is left-recursive in the addition case.
+It is worth mentioning that the grammar is left-recursive in the addition case.
 Since left-recursion is well supported by Packrat parsers, we do not need extra code for it.

### 2017-01-10 paper/sections/Inheritance.tex
Polishing section 4 and 5
Revised S3

@@
 
-\section{Modularizing Parsers with Inheritance}\label{sec:inheritance}
+\section{Parsing OO ASTs with Multiple Inheritance}\label{sec:inheritance}
 
+Before we address the problem of full modular parsing, we first
+address a simpler problem: how to parse Object-Oriented ASTs. To solve
+this problem we employ multiple inheritance, which is supported in
+Scala via traits.
+
+\subsection{The Mismatch Between OO ASTs and Grammars}
+Object-oriented ASTs are ASTs encoded via standard OO techniques, 
+such as the {\sc Composite} or {\sc Interpreter} patterns~\cite{gamma94design}. Such ASTs 
+fix the operations required in the AST, but allow the easy addition of new 
+language constructs. This is in contrast with ASTs modelled using 
+functional programming algebraic datatypes or OO style Visitors, where 
+adding new operations is easy, but adding new language constructs is hard.
+The AST in Figure~\ref{} provides a simple example of an OO AST. The same 
+AST could be encoded using algebraic datatypes (or alternatively Visitors) 
+as:
+
+\begin{lstlisting} 
+data Exp = Lit Int | Add Exp Exp
+\end{lstlisting}
+
+Algebraic datatypes and grammars fit well together. Indeed there are
+obvious similarities in declaring a grammar, and declaring an
+algebraic datatype. It is difficult to add new language
+constructs to both grammars and ASTs based on algebraic datatypes. However,
+ASTs modelled with algebraic datatypes naturally allow for the easy
+addition of operations. So using such ASTs at least preserves one
+dimension of extensibility.
+
+Unfortunatelly, when parsing OO style ASTs, there are difficulties with
+the two dimensions of extensibility. On the one hand it is naturally hard to 
+add new operations to the AST, as usual in the OO style. On the hand 
+grammars are not modularly extensible, thus the addition of new
+language constructs is problematic. Therefore,
+it is perhaps not a suprise that parser generators that also generate 
+AST code (auch as ANTLR~\cite{}), normally generate code based on 
+visitors. Clearly this seems a better choice since generating code 
+based on the {\sc Interpreter} or {\sc Composite} pattern would 
+simply be too inflexible, due to the dual limitation on extensibility.
+
+In the remainder of this section we show a technique for 
+defining \emph{extensible} parsing code. That is new 
+language constructs can be modularly added without changing 
+existing parsing code. The technique employs standard OO 
+style mechanisms such as subtyping and inheritance, as well as, 
+Packrat parsing combinators.
+
+\begin{comment}
 This section introduces the problem of semantic modular parsing which motivates our work, and an initial solution using only standard inheritance in OO.
@@
 Extension 2 will be discussed in Section~\ref{sec:algebrasandparsing}.
+\end{comment}
 
-\subsection{Inheritance-Based Parser}
+\subsection{Extensible Parsing via Inheritance}
 
-Let's firstly consider Extension 1, which is to extend the syntax as well as the parser.
-As an example, we continue with the expression language of literals and additions in Section~\ref{subsec:packratparsing}.
-We introduce variables as a new case as below.
+To illustrate how to modularly add new language constructs and the 
+corresponding parsing code, we continue with the expression language of literals and additions in Section~\ref{subsec:packratparsing}. We introduce variables as a new case:
 
@@
 
-It is easy to extend the corresponding AST in an OO way:
+It is easy to extend the corresponding OO AST in a modular way:
 
@@
 
-Unfortunately, it fails to parse some expressions like \inlinecode{"1 + x"}, which is obviously valid in the new grammar.
+Unfortunately, this fails to parse some expressions like \inlinecode{"1 + x"}, which are obviously valid in the new grammar.
 The reason is that \inlinecode{pAdd} makes two recursive calls to parse sub-expressions, by using \inlinecode{pExpr}, which
@@
 
-\subsection{Overriding for Extensibility}
+\subsection{Overriding for Extensibility}\label{subsec:overriding}
 
@@
 
-Notice that Scala has a linearized-style multiple inheritance for traits. It can be very helpful when composing several languages. Suppose now
-we want to compose the parsers for expressions from pre-defined languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using alternative.
+\subsection{Multiple Extensions and Independent Extensibility}
+A nice feature of Scala is its support for a form of multiple inheritance.
+Scala has a linearized-style multiple inheritance for traits~\cite{}. 
+This can be very helpful when composing several languages, which 
+may have been independently developed. Suppose now
+we want to compose the parsers for expressions from pre-defined
+languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
+alternative.\bruno{Don't be abstract! Use concrete examples. You 
+already have Var, maybe you can add another language extension for 
+boolean literals. Show the code for the boolean literals parsing as
+well as the composition code.}
 The use of keyword \inlinecode{super} is able to
@@
 
+Interestingly note that, due to the use of multiple inheritance, we 
+need two different super calls.\bruno{expand here. People may not 
+be familiar with scala super calls, you have to explain what the 
+syntax does.}
 As we demonstrated, inheritance is the key technique to obtain semantic modularity.
-It enables type-safe code reuse and separate compilation for parsing OO style ASTs with Extension 1.
+It enables type-safe code reuse and separate compilation for parsing OO style ASTs.

### 2017-01-10 paper/sections/PackratParsing.tex
Revised S3

@@
 
-This section discusses the algorithmic challenges introduced by modular parsing and argues that parser combinators from Packrat parsing
-are suitable to address those challenges.
+This section discusses the algorithmic challenges introduced by
+modular parsing and argues that Packrat parser combinators 
+are suitable to address those challenges. The algorithm challenges 
+are important because they dictate how modular parsers should be
+written. Generally speaking, the methodology we propose is that:
+
+\begin{itemize}
+
+\item {\bf Modular parsers should be left-recursive;}
+
+\item {\bf Modular parsers should use a longest match composition operator.}
+
+\end{itemize}
+
+Moreover, the underlying parsing formalism should make backtracking 
+cheap, since \emph{backtracking is pervasive in modular parsing}.
 
@@
 programmers to manually do \textit{left-recursion elimination}, \textit{longest match composition}, and
-require significant amounts of \textit{backtracking}. All of them are
+require significant amounts of \textit{backtracking}. All of those are
 problematic in a modular setting.

### 2017-01-10 paper/sections/Inheritance.tex
Parsing

@@
 languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
-alternative.\bruno{Don't be abstract! Use concrete examples. You 
+alternative.\bruno{You can present the abstract example, as you do
+  here, but you should also present a concrete example. You 
 already have Var, maybe you can add another language extension for 

### 2017-01-11 paper/sections/ObjectAlgebras.tex
Some revision for S4

@@
 
-The inheritance-based approach benefits us from building extensible parsers, based on an OO class hierarchy.
-Nevertheless, the addition of new operations over ASTs is problematic using traditional OO classes.
-Often it requires modification on existing code, which breaks modularity.
-In this section, we will have a closer look at how Object Algebras is applied for full extensibility.
+The inheritance-based approach allows building extensible parsers, based on an OO class hierarchy.
+Nevertheless, the addition of new operations over ASTs is problematic using traditional OO ASTs.
+In this section, we
+show how to support both the forms of extensibility for ASTs (easy 
+addition of language contructs, and easy addition of operations)
+using Object Algebras~\cite{}.
 
@@
 
-Now it comes to Extension 2 in Section~\ref{subsec:parsingproblem}. In the previous code, we have pretty-printing as
-an example of operations on ASTs. Suppose we also want to collect free variables on the expression language which models
+OO ASTs allow easily adding new language constructs, but it is hard 
+to add new operations over the AST. The OO AST in Figure~\ref{}, 
+supports a pretty-printing operation. Suppose we also want to collect free variables on the expression language which models
 literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \inlinecode{Expr} with the new
@@
 
+\bruno{The following text is not very good, needs to be revised. I think 
+we need to show the full code and where it breaks. Since I removed text 
+about the Expression problem in S3, I think we need to talk more about it here.}
 It refers to one Scala solution to the Expression Problem.\huang{what's the solution?
@@
 \subsection{Object Algebras}\label{subsec:objectalgebras}
-Object Algebras captures a design pattern to address the Expression Problem nicely, achieving two dimensions of extensibility (data variants and operations) in a modular and type-safe way.
+Object Algebras captures a design pattern to address the Expression Problem, achieving two dimensions of extensibility (language constructs and operations) in a modular and type-safe way.
 The definition of data structures is separated from their behaviors, and future extensions on both dimensions no longer require existing code to be modified, supporting separate compilation.
@@
 Using Object Algebras in Scala, ASTs as recursive data structures are defined by traits, where each constructor corresponds to an abstract method inside.
+Essentially, Object Algebras generalize the {\sc Abstract Factory} pattern~\cite{}, and promote the use of factory methods, instead of constructors, for instantiating objects.
 The example from Section~\ref{subsec:packratparsing} is used here again for illustration.
@@
 \lstinputlisting[linerange=5-8]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
-We have two constructors \inlinecode{lit} and \inlinecode{add}, corresponding to literals and additions. The trait \inlinecode{ExprAlg} is called an \textit{Object Algebra interface},
-whereas as a factory, it cannot produces objects for the expressions directly like traditional approaches, but abstract the results in its
-type parameter \inlinecode{E} instead. To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
+We have two constructors \inlinecode{lit} and \inlinecode{add}, corresponding to literals and additions.
+The trait \inlinecode{ExprAlg} is called an \textit{Object Algebra interface}.
+The trait is parametrized by the type
+\inlinecode{E}, which abstracts over the concrete type of the AST.
+
+\paragraph{Adding New Operations}
+To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
 provides implementations for all cases. Below is an example of pretty-printing an expression:
@@
 
-Futhermore, the data variants can be extended by inheriting \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
+\paragraph{Adding New AST Constructs}
+Futhermore, new language constructs can be added by extending \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
 with variables. A new Object Algebra interface \inlinecode{VarAlg} is defined as follows:
@@
 An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
-It offers programmers great code reuse, in the sense that existing cases can also be
-overridden in \inlinecode{VarExprPrint}.
+In other words, existing code was reused and was not modified!
 
@@
 
-The code below results in \inlinecode{"(1 + x)"}, as the result of pretty-printing that expression.
+Note how the construction of the asbtract syntax happens throught 
+the use of factory methods, instead of constructors.
+The pretty print the expression, resulting in \inlinecode{"(1 + x)"},
+we pass an insteance of \inlinecode{VarExprPrint} to \inlinecode{makeExp}.
 
@@
 
+\bruno{This is a good discussion, but it is out-of-place. We haven't 
+yet talking about parsing with object algebras, so I think this discussion 
+should come later.}
 Now we observe that an ``object'' of expression actually
@@
 
-Just as shown above, \inlinecode{ExprAlg} is defined for a small language using Object Algebras.
+Parsing builds the AST as a result. When Object Algebras are used 
+to build ASTs, an Object Algebra containing the constructor/factory
+methods has to be passed to the parsing function.
+
+\inlinecode{ExprAlg} is defined for a small language using Object Algebras.
 Since Object Algebras represent ``objects'' implicitly as functions like \inlinecode{ExprAlg[E] =>} \inlinecode{E} for abstract \inlinecode{E}, we first try to build the corresponding parser by type \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.

### 2017-01-11 paper/sections/PackratParsing.tex
Some revision for S4

@@
 are suitable to address those challenges. The algorithm challenges 
-are important because they dictate how modular parsers should be
-written. Generally speaking, the methodology we propose is that:
+are important because they rule out various common techniques 
+used by non-modular code using parsing combinators. 
+To avoid pitfalls related to those algortihmic challenges,
+we propose a the following methodology:
 
@@
 cheap, since \emph{backtracking is pervasive in modular parsing}.
+Although we chose Packrat parsing (due to its immediate availability
+and support), any other parsing formalism that provides similar
+features should be adequate for modular parsing.
 

### 2017-01-11 paper/sections/Inheritance.tex
Some revision for S4
Fix 3.4

--- 2017-01-10 paper/sections/Inheritance.tex
@@
 obvious similarities in declaring a grammar, and declaring an
-algebraic datatype. It is difficult to add new language
+algebraic datatype. It is difficult to modularly add new language
 constructs to both grammars and ASTs based on algebraic datatypes. However,
@@
 To illustrate how to modularly add new language constructs and the
-corresponding parsing code, we continue with the expression language of literals and additions in Section~\ref{subsec:packratparsing}. We introduce variables as a new case:
+corresponding parsing code, we continue with the expression language
+of literals and additions in Section~\ref{subsec:packratparsing}. We
+introduce variables as a new case:
 
@@
 languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
-alternative.\bruno{Don't be abstract! Use concrete examples. You 
+alternative. The new parser can be built by inheriting both parsers
+at the same time:
+
+\lstinputlisting[linerange=67-73]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
+
+\huang{added explanation of the `super' keyword}
+
+The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}.
+Under multiple inheritance, it can be used to distinguish the method of the same name.
+Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.
+
+\huang{added a concrete example}
+
+As a concrete example, we compose the language of integers and additions with a language of boolean values and if-then-else expressions, to build a new language.
+
+\lstinputlisting[linerange=79-95]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
+
+\begin{comment}
+\bruno{You can present the abstract example, as you do
+  here, but you should also present a concrete example. You
 already have Var, maybe you can add another language extension for
@@
 well as the composition code.}
-The use of keyword \inlinecode{super} is able to
-specify the implementation from inheritance, like:
-
-\lstinputlisting[linerange=70-76]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 
@@
 syntax does.}
-As we demonstrated, inheritance is the key technique to obtain semantic modularity.
+\end{comment}
+
+Then the new parser, which contains both integer and boolean expressions, is able to parse hybrid expressions such as \inlinecode{"if true then 1 + 2 else 3 + 4"}.
+
+As we demonstrated, inheritance with method overriding is the key technique to obtain semantic modularity.
 It enables type-safe code reuse and separate compilation for parsing OO style ASTs.

### 2017-01-11 paper/sections/Discussion.tex
Polishing 5.1

@@
 
-Just as many parser generators do, our pattern also supports overriding part of existing parsers easily, but in a type-safe way.
-We can simply update existing production rules, namely the concrete syntax, with new implementations, or eliminate them in extended parsers.
-To illustrate, suppose we have an untyped lambda case in a base parser, but then we replace it with a typed one in its extended parser:
+Just as many parser generators do\huang{???}, our pattern also supports overriding part of existing parsers, but in a type-safe way.
+We can easily update existing parsing rules with new implementations, or eliminate them in extended parsers.
+As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below.
+\inlinecode{pLam} parses a lambda symbol, an identifier, a dot and an expression in sequence.
+Some irrelevant details are omitted.
 
 \lstinputlisting[linerange=7-12]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
-The above parser shows part of the code, for the space reason we omit details. With OO inheritance, we can actually override \lstinline{pLam}
-in a subtype of \lstinline{BaseParser}, and give a different implementation. It will make a difference during runtime due to dynamic
-dispatch.
+
+Then we want to replace the untyped lambda abstractions by typed lambdas.
+With inheritance and method overriding, it is easy to only change the implementation of \lstinline{pLam} in the extended parser.
+The new \inlinecode{pLam} parses a colon and a type in addition, between the identifier and the dot.
+Due to dynamic dispatch, our new explanation of lambdas will make a difference without affecting the other parts of the parser.
 
 \lstinputlisting[linerange=16-22]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
-Futhermore, the lexer can also be updated, since keywords and delimiters are represented by sets of strings.
+
 One can even ``eliminate'' a production rule in the extension, by actually overriding it with a failure parser.
+The lexer can also be updated, since keywords and delimiters are represented by sets of strings.
 

### 2017-01-11 Scala/Parser/src/PaperCode/Sec5/Code1.scala

@@ -16,7 +16,7 @@ trait BaseParser[E] {
 trait ExtParser[E, T] extends BaseParser[E] {
   ...
   override val alg: ...
-  override val pLam = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ ...
+  override val pLam: Parser[E] = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ ...
   val pT: Parser[T] = ...
   override val pE: Parser[E] = ...
 }

### 2017-01-12 paper/sections/ObjectAlgebras.tex
More revisions on S4

@@
 to build ASTs, an Object Algebra containing the constructor/factory
-methods has to be passed to the parsing function.
+methods has to be used by the parsing function. Thus, a first attempt 
+at defining the parser for the small arithmetic language is:
 
+\begin{comment}
 \inlinecode{ExprAlg} is defined for a small language using Object Algebras.
 Since Object Algebras represent ``objects'' implicitly as functions like \inlinecode{ExprAlg[E] =>} \inlinecode{E} for abstract \inlinecode{E}, we first try to build the corresponding parser by type \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.
+\end{comment}
 
 \lstinputlisting[linerange=8-16]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
-Such a parser works fine, yet it does not appear to be extensible.
+Such a parser works fine, but it is not extensible!
 For example, we have demonstrated in Section~\ref{subsec:overriding} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
-However, now \inlinecode{pExpr} has type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}, where the type of Object Algebra interface appears in \emph{contravariant} position.
-This difference prohibits method overriding, because Object Algebra interfaces extended from \inlinecode{VarExprAlg} are subtypes of it, but the types of their parsers are supertypes of \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}.
+However, trying to do a similar method override for \inlinecode{pExpr} would require a type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}, where the type of the \emph{extended} Object Algebra interface appears in \emph{contravariant} position.
+This difference prohibits method/field overriding.
+Scala supports type-refinement of fields if the overriding
+field has a type which is a \emph{subtype} of the original field.
+Unfortunatelly, in this case, although \inlinecode{VarExprAlg} is a subtype 
+of \inlinecode{ExprAlg}, the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} is a supertype of \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.
+Thus it is not possible to have overriding.
 
-This problem refers to \textit{binary methods}. \haoyuan{Shall we mention this?}\huang{???}
+% because Object Algebra interfaces extended from \inlinecode{VarExprAlg} %are subtypes of it, but the types of their parsers are supertypes of %\inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}.
 
-One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. We rewrite the parsing code as follows:
+\paragraph{A Solution}
+A solution to this problem is instead have a field with the Object Algebra
+in \lstinline{ExprOAParser}. This approach supports overriding as well. 
+The parsing code is as follows:
 
@@
 This is precisely the pattern that we advocate for modular parsing. 
-Note that \inlinecode{pE} is the one that is used for recursive calls, exported by the trait and considered to be overridden during inheritance.
-The reason why we use an extra and seemingly redundant \inlinecode{pExpr}, is a subtle issue caused by Scala language and its parser combinator library.
+One important remark is we introduce \inlinecode{pE} for recursive calls. 
+The reason why we use an extra and seemingly redundant field, is due to a subtle issue caused by Scala language and its parser combinator library.
 There is a restriction of \inlinecode{super} keyword in Scala that \inlinecode{super} can only use fields defined by keyword \inlinecode{def}, can not use those defined by \inlinecode{val}, while the parser combinator library suggests using \inlinecode{val} to define parsers, especially for left-recursive ones.
@@
 
-Now let's try this pattern with the extension of varibles, which we have been using:
+\paragraph{Extensions}
+Now let's try this pattern with the variables extension:
 \lstinputlisting[linerange=37-48]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
-In the client code shown above, an instance of algebra is fed to \inlinecode{VarExprOAParser} for pretty-printing expressions.
-Then the trait exports \inlinecode{pE}, parses the input and applies the algebra.
-The result shows that such a pattern is a nice solution to modular parsing.
+\bruno{break code into 2 parts: the definition of the trait; and the client 
+code.}
+\bruno{the following text should be for explaining the definition of the trait}
+\noindent The type of the object algebra field \lstinline{alg} is first refined 
+to \lstinline{VarExpAlg[E]}, to allow calling the additional factory method 
+for variables. Note that, unlike the previous attempt, \lstinline{VarExpAlg[E]}
+is a subtype of \lstinline{ExpAlg[E]}, thus the type-refinement of the 
+field is allowed. Now, the code for parsing variables (\lstinline{pVar}) can 
+call \lstinline{alg.varE}. The following code illustrates how to use 
+the parser from a client perspective:
+
+\bruno{put the client code here.}
+
+\bruno{have we defined the ``parse'' method?? I don't think we have!
+We need to first present the parse method. }
+In the client code shown above, we first need to chose a concrete 
+object algebra to initialize the \lstinline{alg} field in \lstinline{VarExprOAParser}. Here we pick the 
+pretty-printing algebra \lstinline{VarExprPrint}, but any other object 
+algebra that implements \lstinline{VarExprAlg} would work. 
+With an instance of \lstinline{VarExprOAParser} in hand, we can call 
+\lstinline{pE} to obtain the parser to feed to the \lstinline{parse} method. 
+The result shows that such a pattern provides modular parsing.
+
+\begin{comment}
 At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice.
@@
 i'll check later} by composing two algebras into one in advance.
+\end{comment}
 

### 2017-01-12 paper/sections/ObjectAlgebras.tex
Section 4

@@
 
-\huang{to polish}As illustrated above, one intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable.
-We can simply add new variants corresponding to the new cases in the grammar, and realize different operations on structures, with both code reuse and separate compilation supported.
+\begin{comment}
+As illustrated above, using Object Algebras separates data structures from behaviors, thus enabling more modularity and reuse.
+New language constructs correspond to the new cases in the algebra.
+Different operations
+ on structures, with both code reuse and separate compilation supported.
+\end{comment}
 
-Another advantage of using Object Algebras is that we benefit from using generics.
+Another advantage of using Object Algebras is that is is easy to 
+model multi-sorted languages.
 If the syntax contains multiple sorts, we can distinguish them by different type parameters.
-Thus fewer classes are defined and code is simplified.
-As an example, we extend the language of literals, additions and variables by introducing types as a new sort.
-Now it has two sorts in the syntax, types and expressions.
-We certainly do not want to mix them together.\\
+As an example, we extend the language of literals, additions and variables by introducing types as a new kind of type argument, and we add lambda abstractions to expressions.
+Now the language has two sorts in the syntax, types and expressions:\\
 
@@
 
-When extending the Object Algebra interface, pretty-printing operation and parser, we use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
+\begin{figure}
+\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
+\label{fig:multi}
+\end{figure}
+
+Figure~\ref{fig:multi} illustrates the corresponding Scala code 
+that extends the Object Algebra interface, pretty-printing operation and parser.
+We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
 They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
-
-\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
-Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
-we use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, recursively.
-\inlinecode{pT} and \inlinecode{pE} are actually the parsers for recursion.
+Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. 
+We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue 
+with \lstinline{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
 
 From the code we observe that the multi-sorts of Object Algebra interface
-exactly express multiple syntax. The isolation for different syntax
-is guaranteed by the type system. Now it is straightforward to show some client code
+express multiple syntax. The isolation for different syntax
+is guaranteed by the type system. The following illustrates some client code
 for the new parser:
@@
 \lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-It works just as expected to parse and pretty-print the result.
+The code works as expected to parse and pretty-print the result.
+
+Note that, similarly to the approach in Section~\ref{sec:inheritance} independent extensibility is also supported via multiple trait inheritance. 
+Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance} we omit that code here.

### 2017-01-12 paper/sections/RelatedWork.tex
Some revision on Related Work

@@
 
-Our work integrates several components including extensible parsing,
-parser combinators and extensibility techniques. There has been a
-great amount of related papers on those hot topics. Some
+Our work touches upon several topics including extensible parsing,
+parser combinators and extensibility techniques. However, as far as we
+know there's no work that discusses how to do statically type-safe and 
+separately compilable modular parsing.
+
+\begin{comment}
+There has been a
+great amount of related papers on those topics. Some
 inspired us of this paper and encourage us for more exploration. This
 section will try to lead a discussion on what difference we have made.
+\end{comment}
 
-\paragraph{Syntactically Extensible Parsing} Extensible parsing is achieved in many
-different areas, among them parser generators are a mainstream area
-specially designed for modular syntax and parsing. Many parser
+\paragraph{Syntactically Extensible Parsing} Extensible parsing is
+used in many different areas. Extensible parser generators are a
+mainstream area specially designed for modular syntax and
+parsing. Many parser
 generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
-modular grammars. They allow users to create new
-modules where new non-terminals and production rules can be
-introduced, some can even override existing rules in the old grammar
-modules. For instance, \textit{Rats!}~\cite{Grimm2006} constructs its
-own module system for the collection of grammars, while
-NOA~\cite{Gouseti2014} uses Java annotation processing to gather
-information together. Those parser combinators focus on the
-\textit{syntactic extensibility} of grammars, and rely on a whole
-compilation to generate a global parser, even with a slight
-modification. Some of them may statically check the correctness and
-unambiguity of grammars, but separate compilation and modular
-type-checking remain unsolved, together with their affects on the performance.
-\bruno{what performance issue?}\haoyuan{Revised}
+modular grammars. They allow users to create new modules where new
+non-terminals and production rules can be introduced, some can even
+override existing rules in the old grammar modules. For instance,
+\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
+the collection of grammars.  NOA~\cite{Gouseti2014} uses Java
+annotation processing to gather information about grammars in Java
+code. After all information is collected NOA merges all grammar
+components, produce an ANTLR~\cite{} grammar, and generates
+corresponding parsing code from the ANTLR grammar.  Those parser
+generators focus on the \textit{syntactic extensibility} of grammars:
+they rely on whole compilation to generate a global parser, even if
+there is only a slight modification in the grammar. Some of those
+parser generators may statically check the correctness and unambiguity
+of grammars. In constrast, because our approach is based on parser
+combinators, there is no support for ambiguity checking.  However, as
+far as we are aware, no extensible parser generators support separate
+compilation or modular type-checking.
 
-Macro systems like C preprocessor, C++ templates and
+Macro systems like the C preprocessor, C++ templates and
 Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming
 techniques are a similar area aiming at syntactic extensibility.
-SugarJ~\cite{Erdweg2011} conveniently introduces syntactic sugar
-for Java using library imports. Composition of syntactic
-sugar is easy for users, but it requires many rounds of parsing
-and adaption. This significantly affects efficiency of compilation,
-moreover, the implementation was based on SDF~\cite{Heering1989} and
-Stratego~\cite{Visser2001}, which did not focus on separate
-compilation.\bruno{focused little or did not have it at all?}.\haoyuan{The authors have
-mentioned separate compilation as a possible yet unimportant feature; they decided to discard it.}
-Racket adopts macro system for library-based language extensibility. It is powerful with attributed
-ASTs for contextual information, and extensions can be integrated in a modular way. Whereas
-such modularity is not flexible enough for language unification, as the syntax is only built from extensions.
-Moreover, existing macros cannot be further changed after definition. \haoyuan{correct?}
-On the other hand, extensible compilers like JastAdd~\cite{Ekman2007} and
+SugarJ~\cite{Erdweg2011} conveniently introduces syntactic sugar for
+Java using library imports. Composition of syntactic sugar is easy for
+users, but it requires many rounds of parsing and adaption. This
+significantly affects efficiency of compilation. Since the
+implementation was based on SDF~\cite{Heering1989} and
+Stratego~\cite{Visser2001}, the approach does not support separate
+compilation. Racket adopts a macro system for library-based language
+extensibility~\cite{Tobin-Hochstadt2011}. This is quite powerful and it uses 
+attributed ASTs for contextual
+information. Moreover extensions can be integrated in a modular
+way. However such modularity is not flexible enough for language
+unification, as the syntax is only built from extensions.  Moreover,
+existing macros cannot be further changed after
+definition. \haoyuan{correct?}  Extensible
+compilers like JastAdd~\cite{Ekman2007} and
 Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
-is mostly done using parser generators. Moreover
-they focus on the extensions to a host language. Our library
-is designed for modular language parsers in a type-safe way, with
-flexible language composition.
+is mostly done using parser generators. They focus on the
+extensions to a host language. Our library is designed for type-safe 
+modular language parsers (not just language 
+extensions), with very flexible parser composition.
+
 %\bruno{Have you read
@@
 \paragraph{Extensible Parsing Algorithms}
-Extensible parsing algorithms are another area that indeed relates to
-separate compilation. Specifically, \cite{Bravenboer2009} introduces
-\textit{parse table composition}, where grammars are compiled
-to the generation of parse tables. Those parse tables are expressed as DFAs or NFAs, later they
-can be composed by an algorithm, so as to provide separate compilation
-for parsing. Nevertheless, the generation of parse tables is rather
-expensive in terms of performance.
-Furthermore, our approach supports both \emph{separate compilation}
-as well as \emph{modular type-checking}. The extensibility of
-parsing is further available at language composition
-and lexical level.
+Extensible parsing algorithms are another area that relates to
+separate compilation. 
+\textit{Parse table composition}~\cite{Bravenboer2009} 
+is an approach where grammars are compiled to
+modular parse tables. Those parse tables are expressed as DFAs
+or NFAs, and later they can be composed by an algorithm, to provide
+separate compilation for parsing. The generation of parse tables can
+be quite expensive in terms of performance. The approach 
+is quite different from ours, since it uses parse
+ tables, wheres we use parser combinators. 
+Our approach supports both
+\emph{separate compilation} as well as \emph{modular
+  type-checking}. Moreover, the extensibility of parsing is further
+available at language composition and lexical level.
+
 %\haoyuan{we do not have explicit
@@
 
-\paragraph{Parser Combinators} Since~\cite{burge1975} firstly
-introduced parser generators, and after Wadler discussed more details
-on backtracking in~\cite{Wadler1985}, parser combinators have been
+\paragraph{Parser Combinators} Since Burge~\cite{burge1975} firstly
+introduced parser generators, and after Wadler popularized them 
+in functional programming~\cite{Wadler1985}, parser combinators have been
 more and more popularly developed and used in the research area of
-parsing. Among them many parsing libraries produces recursive descent
-parsers by introducing functional, or more specifically, monadic
-parser combinators with a foundation on~\cite{nott237}. Related work
-includes Parsec~\cite{Leijen2001}, which is frequently used in Haskell
+parsing. Many parsing libraries produce recursive descent
+parsers by introducing functional monadic
+parser combinators~\cite{nott237}. Parsec~\cite{Leijen2001} is 
+perhaps the most popular parser combinator library in this line. 
+It is widely used in Haskell (with various ``clones'' in other languages)
 for context-sensitive grammars with infinite lookahead. Nevertheless,
-Parsec users suffer from manual left-recursion elimination, back-tracking and
-longest match composition issues, as we discussed before.
+Parsec users suffer from manual left-recursion elimination, 
+high cost for backtracking and longest match composition issues, 
+as we discussed in Section~\ref{}. Those limitation make Parsec 
+inadequate for modular parsing.
 
-Some more recent papers~\cite{Ford2002,Might2011,Frost2008} proposed a series of
-novel parsing techniques that addressed the issue
-of left-recursion. We have selected Packrat
-Parsing for our prototype implementation, as the combinators are
-convenient to use in Scala, and Object Algebras can be naturally realized with
-OO inheritance. Also, its support
-for direct left-recursion is already helpful for designing real-world
-parsers practically, though~\cite{warth2008} further demonstrated that
-general left-recursion is supported theoretically.
-Yet we have mentioned that the components used in our approach
-can have alternatives to achieve better modular parsing.
+Some more recent work on parser
+combinators~\cite{Ford2002,Might2011,Frost2008} proposed a series of
+novel parsing techniques that address the issue of
+left-recursion. We have selected Packrat parsing for our approach, 
+as the combinators are convenient to use in Scala, and 
+integrate well with multiple trait inheritance and
+Object Algebras. 
+%Also,
+%its support for direct left-recursion is already helpful for designing
+%real-world parsers practically, though~\cite{warth2008} further
+%demonstrated that general left-recursion is supported theoretically.
+Despite our choice of Packrat parsing, as argued in Section~\ref{}, 
+any parser combinator approach which supports left-recursion, longest 
+match composition, and makes the cost of backtracking reasonable,
+should be applicable to modular parsing.
+%Yet we have mentioned that the components used in our approach can
+%have alternatives to achieve better modular parsing.
 

### 2017-01-12 paper/sections/Discussion.tex
Rewrite 5.3

@@
 \subsection{Alternative Techniques}
-In the previous sections we have talked about Packrat Parsing, inheritance and Object Algebras.
-These components fuse together in our library implementation, for building extensible parsers.
-We want to argue here that those aspects are totally orthogonal, and hence can have their alternatives.
+Under our modular parsing framework, we use Packrat parsing as the underlying parsing technique, OO inheritance with method overriding for composing and extending parsers, and Object Algebras for parsing extensible ASTs.
+However, our framework is more general and thus more powerful, because those aspects are orthogonal to each other, and hence can have their alternatives.
 
-For example, we have conducted some previous experiments of modular paring on the Haskell language, based on open recursion, MRM [] and Parsec.
-\haoyuan{I mention our previous work here.
-} The delegation-based open recursion can be an alternative for inheritance.
-It exposes recursion explicitly by a parameter for ``self-reference''.
-Furthermore, there has been a lot of related work that realizes extensible data types in functional programming, including MRM using a list of functors.
-We used Parsec as the parsing library, but suffered from the issues already illustrated in Section~\ref{subsec:challenges}.
-As a workaround, we made use of state monad in Parsec to ensure that a left-recursive production rule will not be applied successively in parsing.
-Whereas a better
-parsing library would potentially introduce a lot of fancy features and parser combinators, run with high efficiency, or even support indirect left-recursion.
+\begin{itemize}
+
+\item {\bf Parsing Technique}
+
+Our approach does not depend on a particular parser combinator library or parsing algorithm.
+We demonstrated that Packrat parsing has some advantages and is suitable in a modular setting.
+However, under certain circumstances or in other programming languages, it may not be the best choice.
+One can use other parser combinator libraries as long as it meets his requirements.
+
+\item {\bf Composing and Extending Parsers}
+
+We use traits to model parsers, OO inheritance to compose them, and method overriding for their extensibility.
+These language features may not be supported in other programming languages, especially in functional languages such as Haskell.
+Nevertheless, \textit{open recursion} could be used as an alternative.
+An explict ``self-reference'' parameter is able to explain the recursive calls dynamically, by the real argument passed at runtime.
+
+\item {\bf Extensible ASTs}
+
+Besides Object Algebras, many other techniques including \textit{Data types à la carte} (DTC) [] and \textit{Modular reifiable matching} (MRM) [] also support extensible data structures.
+These alternatives can also be adopted to build ASTs and corresponding parsers.
+
+\end{itemize}
+
+As we demonstrated, the whole modular parsing framework itself is modular, which means it can be customized easily.
+Actually, we conducted some previous experiments of modular paring in Haskell, based on open recursion, MRM and Parsec.
+However, we suffered from issues of Parsec, as illustrated in Section~\ref{subsec:challenges}.

### 2017-01-12 paper/sections/RelatedWork.tex
Finished revision of Related Work

@@
 high cost for backtracking and longest match composition issues, 
-as we discussed in Section~\ref{}. Those limitation make Parsec 
-inadequate for modular parsing.
+as we discussed in Section~\ref{}. Those limitations make Parsec 
+(and similar parsing techniques) inadequate for modular parsing.
 
@@
 
-\paragraph{Extensibility} To the best of our knowledge, none of the existing parsing techniques
-address the problem of producing extensible ASTs. Certainly, there has been a lot of work on extensible
-data types, in both functional and object-oriented programming.
-For instance, the famous DTC paper~\cite{swierstra2008} represents
-modular ASTs using co-products of every two functors, which forms a
-binary tree structure. Later
-Oliveira et al. presented MRM~\cite{Oliveira2015} in Haskell, which is
-similar to DTC but adopts a list of functors in type-level lists
-instead. Other related work
-includes~\cite{Bahr2014}, where data types are flexible with
-composition as well as decomposition.
+\paragraph{Extensibility} Various design patterns~\cite{}, in multiple
+languages, have been proposed over the years to address extensibility
+problems, such as the Expression Problem~\cite{}. However, to the best
+of our knowledge, none of the existing extensibility techniques
+address the problem of \emph{modularly 
+parsing extensible ASTs}.  There has been
+a lot of work on extensible ASTs, in both functional and
+object-oriented programming.  For instance, the famous ``Datatypes \`a
+la Carte'' (DTC)
+paper~\cite{swierstra2008} represents modular ASTs using co-products
+of every two functors, which forms a binary tree structure. 
+Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}
+All of that work essentially covers how to traverse and 
+consume extensible ASTs. 
+Only in Bahr's~\cite{Bahr2011} work a brief mention of \emph{unfolds} 
+(which are a class of algorithms that \emph{build} ASTs) are mentioned. 
+However Bahr's work does not cover parsing.
 
-In OO languages, however, things get much better with the mechanism of
-inheritance. Furthermore, Object Algebras~\cite{Oliveira2012} were
-proposed as a design pattern for the extensibility of both data
-variants and operations in a modular type-safe way. The code is
-concise and expressive using Object Algebras, building a foundation
-for our new library with various dimensions of extensibility.
+In OO languages, there are also many design patterns that achieve 
+type-safe extensibility~\cite{torgerson,odersky,Oliveira09,Oliveira12,Wang16}. 
+We chose Object Algebras~\cite{Oliveira2012} because the pattern is 
+relatively lightweight and makes good use of existing OO features,
+such as inheritance, generics and subtyping. As seen throughout the paper, 
+the parsing code is concise and expressive using Object Algebras.
+In principle our approach is not coupled to the use of Object
+Algebras, and variants of our approach could be used with other 
+extensibility techniques. Similarly to the work in functional
+programming, the extensibility techniques in OOP have focused 
+on traversing and processing ASTs. We are unware of any work on OOP
+that covers how to do modular parsing for extensible ASTs.
+
+It is worth mentioning that Scala case classes~\cite{} provide a near
+solution to the Expression Problem. Case classes are similar to
+algebraic datatypes but, unlike algebraic datatypes, new cases can be
+modularly added. The reason why case classes are not a complete
+solution to the Expression Problem is that they do not enforce
+exhaustiveness of pattern matching for extensible operations. In other
+words run-time pattern matching errors can happen when writting
+extensible code with case classes.  So full static type-safety is not
+ensured. Nevertheless case classes are a pragmatic approach, which is
+widely used in practice. Therefore modular parsing techniques may be
+of value for extensible code using case classes.  The approach we
+presented in Section 3~\ref{} can readily be adapted to case classes:
+all that the users need to do is to use case classes, instead of
+standard OO classes in their code.
+
+\begin{comment}
 Moreover, in~\cite{Oliveira2012} the authors have discussed the
@@
 which differs from modular parsing.
+\end{comment}
 
+\bruno{There's a problem with the text in 4.2. We need to first 
+fix the problem in S4.2 and then come back to the text below and 
+adapt it accordingly.}
 Another interesting observation, as we have mentioned
@@
 boilerplate code is automatically generated, hence users only need to
-override a few interesting cases, conveniently. It is potentially useful
-in our case. By the way, the idea of overriding
-existing cases was also proposed in MRM. Our approach allows existing parsing
-code to be overridden during inheritance.
+override a few interesting cases, conveniently. It is potentially
+useful in our case. The idea of overriding existing cases was also
+proposed in MRM. Our approach allows existing parsing code to be
+overridden during inheritance.

### 2017-01-12 paper/sections/Discussion.tex
Revised 5.2

@@
 
-Having pointed out that our goal is to realize modular parsers, we expect the code written in an extensible way.
-We have been focusing on how to extend a language with some additional cases, just as many extensible compilers do, nevertheless, in other cases, languages are defined separately, and the composition of them is again a matter of code reuse.
-In this section, we propose a framework for defining small languages as modular components, in order for better composition.
+Modular parsing not only enables us to build a corresponding parser which evolves with the language together, but also allows us to abstract language features as reusable, independent components.
+Generally, a language feature includes related abstract syntax, methods to \textit{build} the syntax (parsing), and methods to \textit{process} the syntax (evaluation, pretty-printing, etc.).
+From this perspective, not only one language, but many languages can be developed in a modular way, with common language features reused.
 
-The framework we use for a language, or rather a language component in our case study, consists of three parts: syntax, parsing and operations on the AST. More
-specifically:
+Instead of design and build a language from scratch, we can easily add a new feature by reusing the corresponding language component.
+For example, if a language is composed with a component of boolean expressions, including if-then-else, it immediately knows how to parse, traverse, and pretty-print if-then-else structure!
+It could be very useful for rapid developing DSLs.
+
+For implementation, a language component is represented by a Scala object, and it consists of three parts: Object Algebra interface, parser, and Object Algebras.
+
 \begin{itemize}
     \item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax. The type parameters represent multiple sorts of syntax, and the methods are constructs.
-\item \textbf{Parser:} implementation of the parser with lexing, using inheritance and Object Algebras, with its scoping trait abstracted over the type parameters.
-\item \textbf{Algebras (optional):} concrete operations on ASTs, such as pretty-printing.
+    \item \textbf{Parser:} corresponding parser of the abstract syntax, written in a modular way as we demonstrated in previous sections.
+    \item \textbf{Object Algebras (optional):} concrete operations on ASTs, such as pretty-printing.
 \end{itemize}
 
-Taking again the example in Section~\ref{subsec:parsingwithoa}. The language constructs consists of literals, additions and variables. First we define the
-language component in a Scala object, shown in the left part of Figure~\ref{fig:lng-components}.
+We take the example in Section~\ref{subsec:parsingwithoa} again, which is a language of literals, additions and variables. It is defined as a
+language component on the left side of Figure~\ref{fig:lng-components}.
+Then for the extension of types and lambda abstractions in Section~\ref{subsec:differentsyntax}, instead of inheriting from the previous language directly, we define it as another independent language component.
+The code is presented on the right side of Figure~\ref{fig:lng-components}.
 
@@
 
-For the example in Section~\ref{subsec:differentsyntax}, instead of building inheritance directly, we define the language separately,
-which consists of types and typed lambda expressions. Shape of the code is presented in the right of Figure~\ref{fig:lng-components}.
-
-At this point, one may want those two language components to be composed into one, and meanwhile ensure that the one is still a modular
-component ready for future composition. In that case modularity is realized over higher-order hierarchies. The composed language \lstinline{VarLamExpr}
-is shown below:
+As shown below, those two language components can be merged into one to obtain the language we want. Futhermore, the new language is still a modular
+component ready for future composition. In that case modularity is realized over higher-order hierarchies.
 
 \lstinputlisting[linerange=82-91]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
-Note that lexing code is automatically combined under multiple inheritance. The combined language shows perfect modularity. The only
+Such composition shows perfect modularity. The only
 drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty printers of
@@
 \subsection{Alternative Techniques}
+
 Under our modular parsing framework, we use Packrat parsing as the underlying parsing technique, OO inheritance with method overriding for composing and extending parsers, and Object Algebras for parsing extensible ASTs.

### 2017-01-12 paper/sections/Discussion.tex
Revised 5.1

@@
 
-Just as many parser generators do\huang{???}, our pattern also supports overriding part of existing parsers, but in a type-safe way.
+As many syntactically extensible parsers, our approach also supports changing part of existing parsers, but in a type-safe way.
 We can easily update existing parsing rules with new implementations, or eliminate them in extended parsers.
@@
 
-\lstinputlisting[linerange=7-12]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=76-81]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
 
@@
 
-\lstinputlisting[linerange=16-22]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\lstinputlisting[linerange=85-90]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
 
@@
 we have a \lstinline{NewParser} that extends \lstinline{ExtParser}, with some more cases added, but for the lambda case we want to go back to the untyped one.
-It can be done by copying the code from \lstinline{BaseParser} again, yet does not achieve code reuse.
-Instead, we can instantiate an instance of \lstinline{BaseParser} to obtain the old \lstinline{pLam}:
-\haoyuan{I don't know if this works or makes sense.}
+It can be achieved by instantiating an instance of \lstinline{BaseParser} to obtain the old \lstinline{pLam}.
+However, we must have \inlinecode{lazy} modifier for all the \inlinecode{pLam} starting from \inlinecode{BaseParser}.
 
-\lstinputlisting[linerange=26-32]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
+\huang{i finally managed to make it work, with two overriding.
+i think this case may not be very common and the solution is kind of unly.
+not sure whether to keep this case or not}\haoyuan{I don't know if this works or makes sense.}
+
+\lstinputlisting[linerange=94-102]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
 
@@
 \begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
-\lstinputlisting[linerange=44-59]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=113-128]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
 &
-\lstinputlisting[linerange=63-78]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\lstinputlisting[linerange=132-147]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
 \end{tabular}
@@
 
-\lstinputlisting[linerange=82-91]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=151-160]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 Such composition shows perfect modularity. The only

### 2017-01-12 Scala/Parser/src/PaperCode/Sec5/Code1.scala

@@ -1,7 +1,76 @@
 package PaperCode.Sec5
 
+import PaperCode.Common
 
-object Code1 {
+
+object Code1 extends Common {
+  trait Alg[E] {
+    def varE(x: String): E
+
+    def lam(x: String, e: E): E
+  }
+
+  trait ExtAlg[E, T] extends Alg[E] {
+    def tyInt(): T
+
+    def tyLam(x: String, t: T, e: E): E
+  }
+
+  trait Print extends ExtAlg[String, String] {
+    override def varE(x: String) = x
+
+    override def lam(x: String, e: String) = "\\" + x + "." + e
+
+    override def tyInt() = "Int"
+
+    override def tyLam(x: String, t: String, e: String) = "\\" + x + ":" + t + "." + e
+  }
+
+  trait BaseParser[E] {
+    lexical.delimiters += ("\\", ".")
+    val alg: Alg[E]
+    val pVar: Parser[E] = ident ^^ alg.varE
+    lazy val pLam: Parser[E] = ("\\" ~> ident) ~ ("." ~> pE) ^^ { case x ~ e => alg.lam(x, e) }
+    val pE: Parser[E] = pLam ||| pVar
+  }
+
+  trait ExtParser[E, T] extends BaseParser[E] {
+    lexical.delimiters += ("\\", ":", ".")
+    lexical.reserved += "Int"
+    override val alg: ExtAlg[E, T]
+    override lazy val pLam: Parser[E] = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ { case x ~ t ~ e => alg.tyLam(x, t, e) }
+    val pT: Parser[T] = "Int" ^^^ alg.tyInt()
+  }
+
+  trait NewAlg[E, T] extends ExtAlg[E, T] {
+    def num(x: Int): E
+  }
+
+  trait NewPrint extends NewAlg[String, String] with Print {
+    def num(x: Int): String = x.toString
+  }
+
+  trait NewParser[E, T] extends ExtParser[E, T] {
+    override val alg: NewAlg[E, T]
+    override lazy val pLam: Parser[E] = new BaseParser[E] {
+      override val alg = NewParser.this.alg
+      override val pE: Parser[E] = NewParser.this.pE
+    }.pLam
+    val pNum: Parser[E] = numericLit ^^ { x => alg.num(x.toInt) }
[…]

### 2017-01-12 paper/sections/ObjectAlgebras.tex
Fix 4.3

@@
 Now let's try this pattern with the variables extension:
-\lstinputlisting[linerange=37-48]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
-\bruno{break code into 2 parts: the definition of the trait; and the client 
+\lstinputlisting[linerange=37-44]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+\huang{done}\bruno{break code into 2 parts: the definition of the trait; and the client
 code.}
@@
 
-\bruno{put the client code here.}
+\lstinputlisting[linerange=48-51]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
 
-\bruno{have we defined the ``parse'' method?? I don't think we have!
-We need to first present the parse method. }
-In the client code shown above, we first need to chose a concrete 
+In the client code shown above, we first need to choose a concrete
 object algebra to initialize the \lstinline{alg} field in \lstinline{VarExprOAParser}. Here we pick the
@@
 We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue
-with \lstinline{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
+with \inlinecode{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
 
@@
 
-\lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=63-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 The code works as expected to parse and pretty-print the result.

### 2017-01-12 Scala/Parser/src/PaperCode/Sec4OA/Code4.scala

@@ -60,9 +60,10 @@ trait TypedLamOAParser[E, T] extends VarExprOAParser[E] {
 //END_OVERVIEW_OA_MULTI_SYNTAX
 
 //BEGIN_OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-val r = parse(new TypedLamOAParser[String, String] {
+val p = new TypedLamOAParser[String, String] {
   override val alg = new TypedLamPrint {}
-}.pE)("\\x:int->int. 1 + x") // "\x : int -> int. (1 + x)"
+}
+val r = parse(p.pE)("\\x:int->int. 1 + x") // "\x : int -> int. (1 + x)"
 //END_OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 
   def main(args: Array[String]): Unit = {

### 2017-01-13 paper/sections/PackratParsing.tex
Fix Section 2, added object for code demonstration

@@
 Using traditional alternative composition, when a preceding parser successfully parses a prefix of the input, it will finish parsing and return the result, in spite that subsequent parsers may be able to parse the whole input.
-%\huang{I've rewritten this paragraph}\bruno{The example is good, I think but the explanation is not.
-%You want to say that when you know
-%  the full grammar, you can figure out where ``try'' is
-%  needed. Without the full grammar you'd need assume the worst
-%  case. Make an effort to make your explanation cristal clear!
-%I think you want to miss the abstract explanation that you give first,
-%with the explanation about the concrete example.
-%Start with ``The need for backtracking is also problematic
-%in a modular setting. For example, suppose ...'' and synchronize
-%the abstract explanation and the explanation for the example.
-%}
 In contrast with the parser above, \inlinecode{parser = parseAdd <|> parseInt} works as expected with the two cases swapped.
@@
 Below we present an example to illustrate Scala parsers.
-%
-%\bruno{So, we make a big fuss identifying the problems with other
-%parser combinators in the previous subsection and now we present
-%an example that does not suffer from any of the previous problems!
-%What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that
-%there's no need for the user to do manual backtracking. Just use
-%the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations
-%discussed previously applies!
-%}
+
+\paragraph{Code Demonstration}
+For more concise demonstration, we assume that all the Scala code in the rest of this paper are in the object \inlinecode{Code}, as shown in Figure~\ref{fig:papercode}.
+It extends traits \inlinecode{StandardTokenParsers} and \inlinecode{PackratParsers} from the Scala parser combinator library.
+Futhermore, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing.
+
+\begin{figure}[t]
+\centering
+\lstinputlisting[linerange=42-53]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_PAPERCODE
+\caption{Helper object for code demonstration in this paper.}\label{fig:papercode}
+\end{figure}
 
@@
 additions. The concrete syntax is as follows:
-%\bruno{Here's a good opportunity to both improve the paper and make it
-%more compact! This is the example that should be illustrated and
-%discussed at the end of Section 2. Then we can just create a figure to
-%refer to the code in Section 2 and refer again to the figure here. No
-%need for two different examples.}
 
@@
 \end{grammar}
-%\alt `(' <expr> `)'
 
-It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases.
-Note that a parser has type \lstinline{PackratParser[E]} for some
-\lstinline{E}, which indicates the type of results it produces. Figure~\ref{fig:packrat-arith} shows the abstract syntax and
-the parser code that generates expressions as ASTs. The AST supports
-pretty-printing via the \lstinline{print} method.
+It is straightforward to model the abstract syntax by classes as below. The ASTs support pretty-printing via the \lstinline{print} method.
 
-\begin{figure}[t]
-\centering
-\lstinputlisting[linerange=6-29]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
-\caption{Packrat Parsing for an arithmetic language.}\label{fig:packrat-arith}
-\end{figure}
+\lstinputlisting[linerange=9-17]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
+
+Then we write corresponding parsers for all cases.
+Note that a parser has type \lstinline{Parser[E]} for some
+\lstinline{E}, which indicates the type of results it produces.
+
+\lstinputlisting[linerange=21-29]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
 
@@
 recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the longest match alternative combinator \inlinecode{|||}.
-Table~\ref{tab:packrat} gives a closer look at part of the API that we have used in our work.
+Table~\ref{tab:packrat} gives a closer look at part of the API used in our work.
 
@@
 
-For more concise demonstration, in the rest of paper we assume that all the code are in an enclosing environment of traits \inlinecode{StandardTokenParsers} and \inlinecode{PackratParsers}.
-Futhermore, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing as below.
 The last line demonstrates how we parse a valid expression \inlinecode{1 + 2} using our parser.
 
-\lstinputlisting[linerange=34-41]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+\lstinputlisting[linerange=33-33]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER

### 2017-01-13 Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala

@@ -1,12 +1,11 @@
 package PaperCode.Sec2Packrat
 
+import PaperCode.Common
 
-object Code1 {
-//BEGIN_PACKRAT_SIMPLE_EXPR
-import scala.util.parsing.combinator.syntactical._
-import scala.util.parsing.combinator._
 
-// Abstract syntax
+object Code1 extends Common {
+
+//BEGIN_PACKRAT_OO_AST
 trait Expr {
   def print: String
 }
@@ -16,15 +15,16 @@ class Lit(x: Int) extends Expr {
 class Add(e1: Expr, e2: Expr) extends Expr {
   def print = "(" + e1.print + " + " + e2.print + ")"
 }
+//END_PACKRAT_OO_AST
 
-// Parser with lexing
-trait ExprParser extends StandardTokenParsers with PackratParsers {
+//BEGIN_PACKRAT_SIMPLE_EXPR
+trait ExprParser {
   lexical.delimiters += "+"
-  val pLit: PackratParser[Expr] =
+  val pLit: Parser[Expr] =
     numericLit ^^ { x => new Lit(x.toInt) }
-  val pAdd: PackratParser[Expr] =
+  val pAdd: Parser[Expr] =
     pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-  def pExpr: PackratParser[Expr] =
+  def pExpr: Parser[Expr] =
     pLit ||| pAdd
 }
 //END_PACKRAT_SIMPLE_EXPR

### 2017-01-13 Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala

@@ -1,10 +1,9 @@
 package PaperCode.Sec2Packrat
 
-import scala.util.parsing.combinator.syntactical._
-import scala.util.parsing.combinator._
+import PaperCode.Common
 
 
-object Code2 extends StandardTokenParsers with PackratParsers {
+object Code2 extends Common {
 
   // Abstract syntax
   trait Expr {
@@ -31,13 +30,6 @@ object Code2 extends StandardTokenParsers with PackratParsers {
   }
 
 //BEGIN_PACKRAT_RUNPARSER
-type Parser[E] = PackratParser[E]
-
-def parse[E](p: Parser[E]): String => E = in => {
-  val t = phrase(p)(new lexical.Scanner(in))
-  if (t.successful) t.get else scala.sys.error(t.toString)
-}
-
 val r = parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
 //END_PACKRAT_RUNPARSER
 
@@ -45,3 +37,18 @@ val r = parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
     println(r)
   }
 }
+
+//BEGIN_PACKRAT_PAPERCODE
+import scala.util.parsing.combinator.syntactical.StandardTokenParsers
+import scala.util.parsing.combinator.PackratParsers
+
+object Code extends StandardTokenParsers with PackratParsers {
+  type Parser[E] = PackratParser[E]
+
+  def parse[E](p: Parser[E]): String => E = in => {
+    val t = phrase(p)(new lexical.Scanner(in))
+    if (t.successful) t.get else scala.sys.error(t.toString)
+  }
+  // Any Scala code in the paper comes here
+}
+//END_PACKRAT_PAPERCODE

### 2017-01-13 Scala/Parser/src/PaperCode/Sec5/Code1.scala

@@ -106,7 +106,7 @@ trait NewParser[E, T] extends ExtParser[E, T] {
 }
 
 
-object Code2 {
+object Code2 extends Common {
 
 /*
 //BEGIN_LANGUAGE_COMPONENTS_VAREXPR
@@ -119,7 +119,11 @@ object VarExpr {
   }
 
   // Parser
-  trait Parser[E] { ... }
+  trait Parsing[E] {
+    val alg: Alg[E]
+    val pE: Parser[E] = ...
+    ...
+  }
 
   // Pretty-printer
   trait Print extends Alg[String] {
@@ -138,7 +142,12 @@ object TypedLam {
   }
 
   // Parser
-  trait Parser[E] { ... }
+  trait Parsing[E, T] {
+    val alg: Alg[E, T]
+    val pE: Parser[E] = ...
+    val pT: Parser[T] = ...
+    ...
+  }
 
   // Pretty-printer
   trait Print extends Alg[String, String] {
@@ -151,9 +160,10 @@ object TypedLam {
 object VarLamExpr {
   trait Alg[E, T] extends VarExpr.Alg[E] with TypedLam.Alg[E, T]
 
-  trait Parser[E, T] extends VarExpr.Parser[E] with TypedLam.Parser[E, T] {
+  trait Parsing[E, T] extends VarExpr.Parsing[E] with TypedLam.Parsing[E, T] {
     override val alg: Alg[E, T]
-    override val pE = ...
+    override val pE: Parser[E] = ...
+    ...
   }
 
   trait Print extends VarExpr.Print with TypedLam.Print

### 2017-01-13 paper/paper.tex

@@
   and enable reasonable performance in a modular setting.
-  Extensibility is achieved using \emph{Object Algebras}.  To evaluate
+  Extensibility is achieved using \emph{multiple inheritance} and 
+  \emph{Object Algebras}.  To evaluate
   the approach we conduct a case study based on the ``Types and
@@
   effectiveness at reusing parsing code from existing interpreters,
-  and the total parsing code is 68\% shorter than the non-modular
-  parsing code.
+  and the total parsing code is 69\% shorter than an existing code
+  based using non-modular parsing code.
 

### 2017-01-13 paper/sections/Conclusion.tex

@@
 
-Huang: to write a couple of sentences for a summary.
+\haoyuan{Huang: to write a couple of sentences for a summary.}
 
-\noindent Haoyuan: to discuss future work. (some boilerplate can be avoided via metaprogramming.)
+For a better user experience, there are certainly some aspects in our pattern for modular parsing that needs
+improvements. Specifically, in Section~\ref{subsec:implementation} we have observed that the glue code in \lstinline{Untyped}
+appears to be boilerplate, as we need to compose the data structures, parsers and operations respectively from the same set of super types.
+Such an issue refers to \textit{family polymorphism} [], and we may rely it on a language feature or meta-programming techniques.
+Moreover, to make better use of Object Algebras, we can possibly adopt some patterns in Shy, together with the composition of algebras, to reduce boilerplate by using code generation.
+
+For future work, we will experiment more on how open recursion contributes to extensible parsing in functional languages, by making use of laziness.
+It is also interesting to see that parsing, to some extent, can be viewed as a special example of unfolds.
+So it is worthwhile considering to generalize the composition of unfolds under certain circumstances.

### 2017-01-13 paper/sections/Discussion.tex

@@
 \begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
-\lstinputlisting[linerange=113-128]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=113-132]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
 &
-\lstinputlisting[linerange=132-147]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\lstinputlisting[linerange=136-156]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
 \end{tabular}
@@
 
-\lstinputlisting[linerange=151-160]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=160-170]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 Such composition shows perfect modularity. The only

### 2017-01-13 paper/sections/CaseStudy.tex
Mention some future work in conclusion.
Done 6.1

@@
 We counted source lines of code (SLOC) and measured execution time for both implementations.
-Our implementation saves almost 70\% code comparing with that non-modular one.
+Our implementation saves 69\% code comparing with that non-modular one.
 
-\huang{done}\bruno{mention that you compare against an existing non-modular implementation.}
-
-\subsection{Implementation}
+\subsection{Implementation}\label{subsec:implementation}
 
 The calculi in the TAPL book are suitable for our case study for mainly two reasons.
-Firstly, the calculi are arguably capture many of the language features
-required in realistic programming languages. Secondly, the evolution of
+Firstly, they capture many of the language features
+required in realistic programming languages, such as lambdas, records and variants. Secondly, the evolution of
 calculi in the book reveals the advantages of modular representation
@@
 of our approach. By extracting common components from those calculi
-and reusing them, we obtain considerably code reuse with reasonable performance penalty, as shown later.
+and reusing them, we obtain considerably code reuse as shown later.
 
 \paragraph{Extracting Language Components}
-Using the modular parsing techniques discussed before, we extract
-reusable components from syntax of all the calculi. Each
+Using the pattern demonstrated in Section~\ref{subsec:language-component}, we extract
+reusable components from all the calculi. Each
 component, which may contain several syntactical structures,
@@
 example, the \lstinline{VarApp} component below represents variables and
-function applications.
+function applications, shown in Figure~\ref{fig:casestudy-varapp}.
 
-\huang{todo: move the code, rename PackratParser}
-\begin{lstlisting}
-object VarApp {
-  trait Alg[E] {
-    def TmVar(x: String): E
-    def TmApp(e1: E, e2: E): E
-  }
-  trait Print extends Alg[String] {
-    def TmVar(x: String): String = x
-    def TmApp(e1: String, e2: String): String = "[" + e1 + " " + e2 + "]"
-  }
-  trait Parser[E] extends EParser[E] {
-    lexical.delimiters += ("(", ")")
-    val alg: Alg[E]
-    val pVarAppE: PackratParser[E] = {
-      lcid ^^ alg.TmVar |||
-        pE ~ pE ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
-        "(" ~> pE <~ ")"
-    }
-  }
-}
-\end{lstlisting}
+\begin{figure}[ht]
+\lstinputlisting[linerange=15-33]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
+\caption{The \inlinecode{VarApp} component in case study.}\label{fig:casestudy-varapp}
+\end{figure}
 
-In our case study, each component is represented by a Scala object which has three parts.
-\lstinline{Alg} is an object algebra interface for the abstract syntax.
-\lstinline{Print} is an object algebra implements that interface
-\lstinline{Alg}, for the pretty printing operation. They are standard object algebra operations.
-\lstinline{Parser} is the modular parser for this piece of syntax.
+Similar as before, each component is represented by a Scala object which includes \lstinline{Alg} for the abstract syntax, \lstinline{Print} for pretty-printing, and \lstinline{Parse} for parsing.
 
@@
 represents kinds. They are the three sorts of syntax in our case study.
-In this particular component \lstinline{VarApp} we only have expressions.
+In the component \lstinline{VarApp} we only have expressions.
+We use some helper traits for eliminating duplicate definitions, such as \inlinecode{EParser} containing \inlinecode{pE} for parsing expressions.
 
-Similar as before, each parser carries lexical information, an object algebra field, and
-parsing functions. We have some helper traits for convenience, such as \inlinecode{EParser}
-containing \inlinecode{pE} for parsing expressions.
-
-\begin{lstlisting}
-trait EParser[E] {
-  val pE: PackratParser[E]
-}
-\end{lstlisting}
+\lstinputlisting[linerange=9-11]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
 
@@
 
-\begin{lstlisting}
-object UntypedAbs {
-  trait Alg[E] {
-    def TmAbs(x: String, e: E): E
-  }
-  trait Print extends Alg[String] {
-    def TmAbs(x: String, e: String): String = "\\" + x + "." + e
-  }
-  trait Parser[E] extends EParser[E] {
-    lexical.delimiters += ("\\", ".", "(", ")")
-    val alg: Alg[E]
-    val pUntypedAbsE: PackratParser[E] =
-      ("\\" ~> lcid) ~ ("." ~> pE) ^^ { case x ~ e0 => alg.TmAbs(x, e0) } |||
-      "(" ~> pE <~ ")"
-  }
-}
+The code of building \lstinline{Untyped} is presented in Figure~\ref{fig:casestudy-untyped}. Note that in the parser \inlinecode{Parse}, we need to override the object algebra interface and
+parsing functions accordingly.
 
-object Untyped {
-  trait Alg[E] extends UntypedAbs.Alg[E] with VarApp.Alg[E]
-  trait Print extends Alg[String] with UntypedAbs.Print with VarApp.Print
-  trait Parser[E] extends UntypedAbs.Parser[E] with VarApp.Parser[E] {
-    override val alg: Alg[E]
-    val pUntypedE: PackratParser[E] = pUntypedAbsE ||| pVarAppE
-    override val pE: PackratParser[E] = pUntypedE
-  }
-}
-\end{lstlisting}
-
-When composing the calculus from components, all the traits can be
-easily combined by Scala's \inlinecode{extends...with} keywords.
-In \lstinline{Parser}, we need to override the object algebra interface and
-parsing functions accordingly.
+\begin{figure}[ht]
+\lstinputlisting[linerange=37-64]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
+\caption{Build the \inlinecode{Untyped} calculus by composing to language components.}\label{fig:casestudy-untyped}
+\end{figure}
 
@@
 
-\begin{lstlisting}
-object TestUntyped {
-  def work[E](inp: String)(a: Untyped.Alg[E]): E = {
-    val lang = new Untyped.Parser[E] {
-      override val alg: Untyped.Alg[E] = a
-    }
-    parse(lang.pE)(inp)
-  }
-  def parseAndPrint(inp: String): Unit = println(work(inp)(new Untyped.Print {}))
-  def main(args: Array[String]): Unit = {
-    parseAndPrint("(\\x.x) (\\y.y)")
-  }
-}
-\end{lstlisting}
+\lstinputlisting[linerange=68-78]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_DEMO
 
@@
     \includegraphics[width=\textwidth]{resources/depGraph.pdf}
-    \caption{Dependency graph of all calculi and components}
+    \caption{Dependency graph of all calculi and components in case study.}
     \label{fig:dependency}
@@
 
-Using our modular parsing approach, it is convenient to extract language
-features as components and reuse them. From the dependency graph, we know
+From the dependency graph, we know
 that common components such as \lstinline{VarApp} are reused in

### 2017-01-13 Scala/Parser/src/PaperCode/Common.scala

@@ -8,6 +8,9 @@ trait Common extends StandardTokenParsers with PackratParsers {
 
   type Parser[E] = PackratParser[E]
 
+  val lcid: Parser[String] = ident ^? { case id if id.charAt(0).isLower => id }
+  val ucid: Parser[String] = ident ^? { case id if id.charAt(0).isUpper => id }
+
   def parse[E](p: Parser[E]): String => E = in => {
     val t = phrase(p)(new lexical.Scanner(in))
     if (t.successful) t.get else scala.sys.error(t.toString)

### 2017-01-13 Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala

@@ -1,0 +1,80 @@
+package PaperCode.Sec6CaseStudy
+
+import PaperCode.Common
+
+
+object Code1 extends Common {
+
+//BEGIN_CASESTUDY_EPARSER
+trait EParser[E] {
+  val pE: Parser[E]
+}
+//END_CASESTUDY_EPARSER
+
+//BEGIN_CASESTUDY_VARAPP
+object VarApp {
+  trait Alg[E] {
+    def TmVar(x: String): E
+    def TmApp(e1: E, e2: E): E
+  }
+
+  trait Print extends Alg[String] {
+    def TmVar(x: String): String = x
+    def TmApp(e1: String, e2: String): String = "[" + e1 + " " + e2 + "]"
+  }
+
+  trait Parse[E] extends EParser[E] {
+    lexical.delimiters += ("(", ")")
+    val alg: Alg[E]
+    val pVarAppE: Parser[E] = lcid ^^ alg.TmVar |||
+      pE ~ pE ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
+      "(" ~> pE <~ ")"
+  }
+}
+//END_CASESTUDY_VARAPP
+
+//BEGIN_CASESTUDY_UNTYPED
+object UntypedAbs {
+  trait Alg[E] {
+    def TmAbs(x: String, e: E): E
+  }
+
+  trait Print extends Alg[String] {
+    def TmAbs(x: String, e: String): String = "\\" + x + "." + e
+  }
+
+  trait Parse[E] extends EParser[E] {
+    lexical.delimiters += ("\\", ".")
+    val alg: Alg[E]
+    val pUntypedAbsE: Parser[E] =
+      ("\\" ~> lcid) ~ ("." ~> pE) ^^ { case x ~ e0 => alg.TmAbs(x, e0) }
+  }
+}
+
+object Untyped {
+  trait Alg[E] extends UntypedAbs.Alg[E] with VarApp.Alg[E]
+
+  trait Print extends Alg[String] with UntypedAbs.Print with VarApp.Print
+
+  trait Parse[E] extends UntypedAbs.Parse[E] with VarApp.Parse[E] {
+    override val alg: Alg[E]
+    val pUntypedE: Parser[E] = pUntypedAbsE ||| pVarAppE
+    override val pE: Parser[E] = pUntypedE
+  }
+}
+//END_CASESTUDY_UNTYPED
+
+//BEGIN_CASESTUDY_DEMO
[…]

### 2017-01-13 paper/sections/CaseStudy.tex
Text

@@
 To demonstrate the utility of our modular parsing approach, we
-implemented parsers of the first 18 calculi from book \textit{Types
-  and Programming Languages} (TAPL) \cite{pierce2002types}. TAPL introduces several
-calculi from simple to complex, by gradually adding new features to
-syntax. We compared our implementation with a non-modular implementation
+implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages} (TAPL) \cite{pierce2002types}. We compared our implementation with a non-modular implementation
 we found online, which is also written in Scala and uses the same Packrat parsing library.
 We counted source lines of code (SLOC) and measured execution time for both implementations.
-Our implementation saves 69\% code comparing with that non-modular one.
+The result suggests that our implementation saves 69\% code comparing with that non-modular one.
 
@@
 
-The calculi in the TAPL book are suitable for our case study for mainly two reasons.
+TAPL introduces several calculi from simple to complex, by gradually adding new features to syntax.
+These calculi are suitable for our case study for mainly two reasons.
 Firstly, they capture many of the language features

### 2017-01-13 paper/sections/CaseStudy.tex

@@
 We have demo code for every calculus in order to test and evaluate our implementation.
-We feed the pretty printing algebra to the parser and print out the result.
+We feed the pretty-printing algebra to the parser and print out the result.
 The parsing result can be changed as long as the concrete algebra is available.
@@
 
-We compared our implementation of the parsers with an implementation
-by Ilya Klyuchnikov, available online\footnote{https://github.com/ilya-klyuchnikov/tapl-scala/}.
-Ilya's implementation is suitable for comparison, because it is also
+\newcommand\ourimpl{$\texttt{Mod}_{\texttt{OA}}$}
+\newcommand\ilyaimpl{\texttt{NonMod}}
+\newcommand\ourclass{$\texttt{Mod}_{\texttt{CLASS}}$}
+\newcommand\ilyalongest{$\texttt{NonMod}_{\texttt{|||}}$}
+
+We compared our implementation (named \ourimpl{}) with an implementation
+by Ilya Klyuchnikov (named \ilyaimpl{}), available online\footnote{https://github.com/ilya-klyuchnikov/tapl-scala/}.
+\ilyaimpl{} is suitable for comparison, because it is also
 written in Scala using the same parser combinator library.
@@
 the length of each line does not exceed 120 characters. Furthermore,
-because Ilya's implementation has extra code such as semantics,
+because \ilyaimpl{} has extra code such as semantics,
 we removed all irrelevant code and only keep abstract
-syntax definition, parser and pretty printer for each calculus, to
+syntax definition, parser and pretty-printer for each calculus, to
 ensure a fair comparison.
@@
 expressions randomly generated, and the size of test files varies from 20KB to 100KB.
-Then we run the corresponding parser to parse the file and the pretty printer to print the result.
+Then we run the corresponding parser to parse the file and the pretty-printer to print the result.
 The average execution time of 5 runs excluding reading input file was calculated, in milliseconds.
-
-\huang{done}\bruno{say something about what those expressions are, size, and how they were generated.}
 
@@
         \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{3}{ c| }{\bfseries Time (ms)} \\ \cline{2-7}
-        \multicolumn{1}{|c|}{} & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% \\
+        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% \\
       \hline
@@
 the whole case study. We can see that in the first two calculi
-\lstinline{Arith} and \lstinline{Untyped} we are not better than Ilya's
-implementation, because in such two cases we do not reuse any existing
+\lstinline{Arith} and \lstinline{Untyped} we are not better than \ilyaimpl{}, because in such two cases we do not reuse any existing
 components. However in the following 16 calculi, we can reuse
@@
 We use object algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
-while Ilya's implementation uses case class and the ordinary alternative combinator.
+while \ilyaimpl{} uses case class and the ordinary alternative combinator.
 
-We implemented two more versions. One is a modified version of our implementation,
-with object algebra replaced by case class for the ASTs.
-The other is a modified version of Ilya's implementation, using the longest match
-alternative combinator instead of the ordinary one.
-
-\huang{done}\bruno{A table showing a summary of results is missing here. We discussed this before.}
+We implemented two more versions. One is a modified version of our implementation, named \ourclass{}, with object algebra replaced by case class for the ASTs.
+The other is a modified version of \ilyaimpl{}, named \ilyalongest{}, using the longest match alternative combinator instead of the ordinary one.
 
@@
       \hline
-        \multirow{2}{*}{\bfseries Calculus Name} & \bfseries Ilya's & \multicolumn{2}{ c| }{\bfseries Ours} & \multicolumn{2}{ c| }{\bfseries Ilya's with \inlinecode{|||}} & \multicolumn{2}{ c| }{\bfseries Ours with class} \\ \cline{3-8}
+        \multirow{2}{*}{\bfseries Calculus Name} & \ilyaimpl{} & \multicolumn{2}{ c| }{\ourimpl{}} & \multicolumn{2}{ c| }{\ilyalongest{}} & \multicolumn{2}{ c| }{\ourclass{}} \\ \cline{2-8}
         \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{\bfseries Time} & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% \\

### 2017-01-13 paper/sections/Discussion.tex
Done 6.1

@@
 
-\subsection{Language Component}
+\subsection{Language Component}\label{subsec:language-component}
 
@@
 \end{tabular}
-\caption{Language \lstinline{VarExpr} (left) and language \lstinline{TypedLam} (right).}\label{fig:lng-components}
+\caption{Language components \lstinline{VarExpr} (left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
 \end{figure}
@@
 Such composition shows perfect modularity. The only
-drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty printers of
+drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty-printers of
 \lstinline{VarExpr} and \lstinline{TypedLam} respectively. Such a pattern refers to \textit{family polymorphism}[]. Unfortunately Scala does not

### 2017-01-13 paper/sections/Introduction.tex
Progress on Intro

@@
 programs has been going on since the early days of Software
-Engineering~\cite{}. Modern Programming Languages enable a certain
+Engineering~\cite{McIlroy68}. Modern Programming Languages enable a certain
 degree of modularity, but they have limitations as illustrated by
-well-known problems such as the Expression Problem~\cite{}. The
+well-known problems such as the Expression Problem~\cite{wadler1998expression}. The
 Expression Problem refers to the difficulty of writing data
@@
 
+\begin{comment}
 A reason why a solution to the Expression Problem is important in
@@
 expressions can be combined to implement the desired functionality.
+\end{comment}
 
@@
 merge it together when a concrete combination of features is needed
-for a particular program. As Kastner~\cite{} notes, the typical
+for a particular program. As Kastner et. al~\cite{Kastner11road} note, the typical
 drawback of such techniques is that
@@
   interfaces and support neither modular type checking nor separate
-  compilation}''\bruno{reference to ``The road to
-  Feature Modularity''}. Syntactic modularization techniques have also
-been applied to the problem of \emph{extensible parsing}. There are
-several approaches~\cite{} that enable the development of
-\emph{syntactically} modular parsers or grammars. However these
-approaches do not support separate compilation or modular
-type-checking either.
+  compilation}''.
+
+Syntactic modularization techniques have also been applied to the
+problem of \emph{extensible parsing}. There are several approaches
+that enable the development of \emph{syntactically} modular parsers or
+grammars. Many parser
+generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
+modular grammars. For instance, \textit{Rats!}~\cite{Grimm2006} has
+its own module system for the collection of grammars.  Extensible
+compilers like JastAdd~\cite{Ekman2007} and
+Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
+is mostly done ultimately resorting to standard (non-modular) parser
+generators. Various techniques supporting languages that can extend
+their own syntax, such as SugarJ~\cite{Erdweg2011}, also offer a form
+of extensible parsing. However these syntactic approaches do not
+support separate compilation and/or modular type-checking of parsing
+code either.
 
@@
 compilation are desirable properties to have from a software
-engineering point-of-view. Furthermore they enable the composition of compiled
-binaries as well as ensuring the type-safety of the code composed of
-multiple components. Examples of semantic modularization techniques 
-include various approaches to \emph{family polymorphism}~\cite{},
-virtual classes~\cite{}, as 
-well as several other techniques for solving the Expression Problem~\cite{}. 
-\bruno{strenghten? mention dynamic languages (separate compilation 
-but no modular type-checking), and techniques that weaken type-safety 
-requirements.}
+engineering point-of-view. Modular type-checking can report errors 
+earlier and in terms of the modular code programmers have written 
+in the first place. Separate compilation avoids global compilation
+steps, which can be very costly. Furthermore semantic modularization 
+enables the composition of compiled binaries as well as ensuring the 
+type-safety of the code composed of multiple components. Examples of semantic modularization techniques 
+include various approaches to \emph{family polymorphism}~\cite{ernst01FP},
+virtual classes~\cite{Ernst:2006}, as 
+well as several other techniques for solving the Expression
+Problem~\cite{Oliveira:2012}\bruno{more here; yanlin; Torgersen; Odersky...}.
 
@@
 in mainstream programming languages like Java or Scala. Object
-Algebras are one such technique, which 
+Algebras~\cite{Oliveira:2012} are one such technique, which 
 Most semantic
@@
     backtracking removal.}
+
+\begin{itemize}
+
+\item {\bf Algorithmic challenges:} answer Packrat parsing
+
+\item {\bf Typing and Reusability Challenges:} answer Multiple
+  inheritance and Object Algebras.
+
+\end{itemize}
 

### 2017-01-13 paper/sections/RelatedWork.tex
Progress on Intro

@@
 code. After all information is collected NOA merges all grammar
-components, produce an ANTLR~\cite{} grammar, and generates
+components, produce an ANTLR~\cite{antlr1995} grammar, and generates
 corresponding parsing code from the ANTLR grammar.  Those parser

### 2017-01-13 paper/sections/Discussion.tex
Section 4.1 revised. Section 5.1 is a newly-added subsection. Please review.

@@
 This section explores more interesting features, including overriding existing parsing rules, language components for abstracting language features, and alternative techniques under the whole framework.
+
+\subsection{Composition of Algebras}
+
+\haoyuan{I add a new subsection here, please review.}
+By adopting Object Algebras, we also inherit some advanced features from the pattern.
+Specifically, in the original paper~\cite{Oliveira2012}, there are two kinds of Object Algebras: \textit{queries} and \textit{transformations}.
+A query traverses the abstract syntax tree, collects information and returns the result.
+The aforementioned \lstinline{Print} and \lstinline{Eval} are traditional queries.
+But as we discussed before, an algebra is first fed to the parser, then \lstinline{parse} is invoked to do the parsing.
+It is tedious to fed \lstinline{Print} and \lstinline{Eval} separately, as the same input is parsed twice.
+Instead, we can firstly compose two queries into one, then feed it to the parser.
+For space reasons we omit the code here.
+
+Another interesting observation, if we look back at the end of Section~\ref{subsec:objectalgebras}, is that \lstinline{makeExp} represents an ``object'' of the expression AST.
+Or rather, what OA parsers generate are ``implicit objects'', which are actually functions of type \lstinline{Alg[E]} \lstinline{=>} \lstinline{E} for generic \lstinline{E}.
+Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
+Especially, in the design of a compiler
+there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
+in a very clever way. Continuing from the old example, if one wants to transform a \inlinecode{VarExprAlg} expression to an \inlinecode{ExprAlg} expression by
+replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}
+
+\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+Here the \inlinecode{VarExprAlg} expression \inlinecode{"1 + x"} is changed to \inlinecode{ExprAlg} expression \inlinecode{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
+\inlinecode{alg} in \inlinecode{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
+to \inlinecode{alg}, \inlinecode{Refactor} will be able to transform a \inlinecode{VarExprAlg} expression before we obtain the result.
+Furthermore, transformations can be composed in a linear pipeline-style, before a query algebra is fed ~\cite{Zhang2015}.
 

### 2017-01-13 Scala/Parser/src/PaperCode/Sec4OA/Code1.scala

@@ -16,7 +16,7 @@ object Code1 {
   }
 
 //BEGIN_ATTEMPT_EXPRWITHFREEVARS
-trait FreeVarsExpr extends Expr { def freeVars: Set[String] }
+trait FVarsExpr extends Expr { def fVars: Set[String] }
 //END_ATTEMPT_EXPRWITHFREEVARS
 
 }

### 2017-01-13 paper/sections/ObjectAlgebras.tex

@@
 OO ASTs allow easily adding new language constructs, but it is hard
-to add new operations over the AST. The OO AST in Figure~\ref{},
+to add new operations over the AST. The code in Section~\ref{subsec:packratparsing},
 supports a pretty-printing operation. Suppose we also want to collect free variables on the expression language which models
@@
 
-\bruno{The following text is not very good, needs to be revised. I think
-we need to show the full code and where it breaks. Since I removed text
-about the Expression problem in S3, I think we need to talk more about it here.}
-It refers to one Scala solution to the Expression Problem.\huang{what's the solution?
+But then programmers have to define new classes for all the constructs, including literals, additions and variables, by extending the old ones together with \lstinline{FVarsExpr}, and implementing the corresponding \lstinline{fVars} respectively.
+
+\begin{lstlisting}
+class FVarsLit(x: Int) extends Lit(x) with FVarsExpr {
+def fVars = Set()
+}
+class FVarsAdd(e1: FVarsExpr, e2: FVarsExpr) extends Add(e1, e2) with FVarsExpr {
+def fVars = e1.fVars ++ e2.fVars
+}
+class FVarsVar(x: String) extends Var(x) with FVarsExpr {
+def fVars = Set(x)
+}
+\end{lstlisting}
+The above code already appears as boilerplate, especially, more classes have to be defined along with future extensions.
+It pollutes the namespace heavily.
+Moreover, the \lstinline{ExprParser} code is no longer useful; to generate expressions that support free variables, we have to substitute \lstinline{FVarsExpr}, \lstinline{FVarsLit}, \lstinline{FVarsAdd} and \lstinline{FVarsVar} for the old \lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var} correspondingly.
+Again, to avoid modifying existing code, we have to sacrifice code reuse.
+In fact, one would not like to make a choice from them, but expect both.
+We have to seek for a different approach to build ASTs.
+\haoyuan{Please check text.}
+
+%\bruno{The following text is not very good, needs to be revised. I think
+%we need to show the full code and where it breaks. Since I removed text
+%about the Expression problem in S3, I think we need to talk more about it %here.}
+%It refers to one Scala solution to the Expression Problem.\huang{what's the %solution?
 does it have the later "critical issue"?
-if so, why is it a "solution"} But then programmers have to define new classes for literals, additions and variables
-by extending the old ones together with \inlinecode{FreeVarsExpr}. The critical issue is that, the existing parsing code only generates objects of the old
-types, again it is a modularity issue, since it is a bad idea to modify existing code again and again with extensions.
-Other solutions including modelling constructors as functions introduce unnecessary complexity.
+if so, why is it a %"solution"} But then programmers have to define new classes for literals, %additions and variables
+%by extending the old ones together with \inlinecode{FreeVarsExpr}. The %critical issue is that, the existing parsing code only generates objects of %the old
+%types, again it is a modularity issue, since it is a bad idea to modify %existing code again and again with extensions.
+Other solutions including %modelling constructors as functions introduce unnecessary complexity.
 
-In contrast, Object Algebras~\cite{Oliveira2012} enables us to solve this problem. As a nice solution, it separates data variants
+Fortunately, Object Algebras~\cite{Oliveira2012} enables us to solve this problem. As a nice solution, it separates data variants
 and operations, and offers high flexibility in the choice of operations to be performed over ASTs. Before adopting it for parsing, the next section introduces the background of Object Algebras.
@@
 
-Note how the construction of the asbtract syntax happens throught
+Note how the construction of the abstract syntax happens through
 the use of factory methods, instead of constructors.
-The pretty print the expression, resulting in \inlinecode{"(1 + x)"},
-we pass an insteance of \inlinecode{VarExprPrint} to \inlinecode{makeExp}.
+To pretty print the expression,
+we pass an instance of \inlinecode{VarExprPrint} to \inlinecode{makeExp}:
 
@@
 \end{lstlisting}
-
-\bruno{This is a good discussion, but it is out-of-place. We haven't
-yet talking about parsing with object algebras, so I think this discussion
-should come later.}
-Now we observe that an ``object'' of expression actually
-has the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{E} for generic \inlinecode{E}.
-Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
-Especially, in the design of a compiler
-there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
-in a very clever way. For example, if one wants to transform a \inlinecode{VarExprAlg} expression to an \inlinecode{ExprAlg} expression by
-replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}
-
-\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
-Here the \inlinecode{VarExprAlg} expression \inlinecode{"1 + x"} is changed to \inlinecode{ExprAlg} expression \inlinecode{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
-\inlinecode{alg} in \inlinecode{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
-to \inlinecode{alg}, \inlinecode{Refactor} will be able to transform a \inlinecode{VarExprAlg} expression before we obtain the result.
+which results in \inlinecode{"(1 + x)"} as expected.
 

### 2017-01-13 paper/sections/Introduction.tex
More Intro

@@
 programs has been going on since the early days of Software
-Engineering~\cite{McIlroy68}. Modern Programming Languages enable a certain
+Engineering~\cite{McIlroy68}. Modern Programming Languages (PLs) enable a certain
 degree of modularity, but they have limitations as illustrated by
@@
 include various approaches to \emph{family polymorphism}~\cite{ernst01FP},
-virtual classes~\cite{Ernst:2006}, as 
+\emph{virtual classes}~\cite{Ernst:2006}, as 
 well as several other techniques for solving the Expression
@@
 Semantic modularization techniques are less widely used in practice 
-than syntactic techniques. This is partly due to the perceived need
-for sophisticated type systems, which are not available in mainstream 
+than syntactic techniques. This is partly due to the need
+for more sophisticated type systems, which are not available in mainstream 
 languages and may require more knowledge from users. However, recently 
@@
 in mainstream programming languages like Java or Scala. Object
-Algebras~\cite{Oliveira:2012} are one such technique, which 
-Most semantic
-modularization techniques have focused on operations that traverse or 
-or process extensible datastructures, such as ASTs.
-\bruno{Object Algebras here?} However, as far as
-we know there is little work on operations that build/produce ASTs. 
-In particular the problem of modular parsing has not been studied in
-semantic modularization approaches. This is a shame because, to
-realise the vision of software product-lines for programming
-languages, modular parsing is a necessity.
+Algebras~\cite{Oliveira:2012} are one such technique, which works in
+Java-like languages and uses simple generics only. So far research on semantic
+modularization techniques have focused on operations that \emph{traverse} or 
+or \emph{process} extensible datastructures, such as ASTs. Indeed 
+many documented applications of semantic modularization techniques
+focus on modularizing various aspects of PL implementations.
+However, as far as we know there is little work on operations that build/produce ASTs. 
+In particular the problem of how to modularize parsing has not been studied in
+semantic modularization approaches. This is a shame because parsing
+is a fundamental part of PL implementations, and it ought to be
+modularized as well.
 
-\paragraph{Modular and Extensible Parsing}
-  This paper investigates presents \name: a parsing
-  combinator library that enables modular parsing.
-  \name provides a solution for the problem of \emph{semantic modular
-    parsing}. That is, the solutions should not only
-  allow complete parsers to be built out of modular parsing
-  components, but also enable the parsing components to be \emph{modularly
-  type-checked} and \emph{separately compiled}. \name is
-  built on top of a Packrat parsing library, but adds new parsing
-  combinators to enable modular parsing. The new parsing combinators 
-  employ \emph{delegation-based} techniques and \emph{Object Algebras} 
-  to support extensibility. The choice of Packrat parsing over other
-  parsing techniques turns out to be important for achieving
-  performance in a modular setting. \bruno{left recursion, and
-    backtracking removal.}
+%\paragraph{Modular and Extensible Parsing}
+
+This paper investigates presents a technique for doing semantically
+modular parsing.  That is, our approach not only allows complete
+parsers to be built out of modular parsing components, but also enables
+those parsing components to be \emph{modularly type-checked} and
+\emph{separately compiled}. In developing our techniques we encontered 
+two different classes of challenges:
 
@@
 
-\item {\bf Algorithmic challenges:} answer Packrat parsing
+\item {\bf Algorithmic challenges:} A first challenge was todo with
+  the parsing algorithms themselves. Since many parsing formalisms
+  were not designed with extensibility in mind, some of the employed 
+  techniques were non-modular, or performed poorly in a modular setting.
 
-\item {\bf Typing and Reusability Challenges:} answer Multiple
-  inheritance and Object Algebras.
+\item {\bf Typing and Reusability Challenges:} The second class of
+  challenges were problems related to modularity, reusability and
+  typing.
 
 \end{itemize}
+
+\name is built on top of a Packrat parsing
+library, but adds new parsing combinators to enable modular
+parsing. The new parsing combinators employ \emph{delegation-based}
+techniques and \emph{Object Algebras} to support extensibility. The
+choice of Packrat parsing over other parsing techniques turns out to
+be important for achieving performance in a modular
+setting. \bruno{left recursion, and backtracking removal.}
 

### 2017-01-13 paper/sections/Introduction.tex
Paper

@@
 Semantic modularization techniques are less widely used in practice
-than syntactic techniques. This is partly due to the need
-for more sophisticated type systems, which are not available in mainstream 
+than syntactic techniques. This is partly due to the need for more
+sophisticated type systems, which are not available in mainstream
 languages and may require more knowledge from users. However, recently
@@
 Algebras~\cite{Oliveira:2012} are one such technique, which works in
-Java-like languages and uses simple generics only. So far research on semantic
-modularization techniques have focused on operations that \emph{traverse} or 
-or \emph{process} extensible datastructures, such as ASTs. Indeed 
-many documented applications of semantic modularization techniques
-focus on modularizing various aspects of PL implementations.
-However, as far as we know there is little work on operations that build/produce ASTs. 
-In particular the problem of how to modularize parsing has not been studied in
-semantic modularization approaches. This is a shame because parsing
-is a fundamental part of PL implementations, and it ought to be
-modularized as well.
+Java-like languages and uses simple generics only. So far research on
+semantic modularization techniques have focused on operations that
+\emph{traverse} or or \emph{process} extensible datastructures, such
+as ASTs. Indeed many documented applications of semantic
+modularization techniques focus on modularizing various aspects of PL
+implementations.  However, as far as we know there is little work on
+operations that build/produce ASTs.  In particular the problem of how
+to modularize parsing has not been studied in semantic modularization
+approaches. This is a shame because parsing is a fundamental part of
+PL implementations, and it ought to be modularized as well.
 

### 2017-01-13 paper/sections/ObjectAlgebras.tex
Revised 4.1

@@
 addition of language contructs, and easy addition of operations)
-using Object Algebras~\cite{}.
+using Object Algebras~\cite{Oliveira2012}.
 
@@
 
-OO ASTs allow easily adding new language constructs, but it is hard
-to add new operations over the AST. The code in Section~\ref{subsec:packratparsing},
-supports a pretty-printing operation. Suppose we also want to collect free variables on the expression language which models
-literals, additions and variables. Based on the traditional OO class hierarchy, one attempt would be extending \inlinecode{Expr} with the new
-operation directly:
+The Expression Problem illustrates the difficulty of extending data structures or ASTs in two dimensions.
+For traditional OO ASTs, adding new language constructs is easy, but adding new operations is hard.
+In the last section, we have a simple language containing literals, additions and variables.
+The corresponding ASTs are built using classes with a pretty-printing operation.
+Now suppose we want to collect free variables over the ASTs.
+By traditional OO approach, one attempt would be extending \inlinecode{Expr} with the new operation to obtain a new abstract type for ASTs:
 
-\lstinputlisting[linerange=19-19]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=ATTEMPT_EXPRWITHFREEVARS
+\lstinputlisting[linerange=23-23]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=FVARSEXPR
 
-But then programmers have to define new classes for all the constructs, including literals, additions and variables, by extending the old ones together with \lstinline{FVarsExpr}, and implementing the corresponding \lstinline{fVars} respectively.
+But then programmers have to define a new class for each construct by extending the old class with \lstinline{FVarsExpr}, and implementing the corresponding \lstinline{fVars} method:
 
-\begin{lstlisting}
-class FVarsLit(x: Int) extends Lit(x) with FVarsExpr {
-def fVars = Set()
-}
-class FVarsAdd(e1: FVarsExpr, e2: FVarsExpr) extends Add(e1, e2) with FVarsExpr {
-def fVars = e1.fVars ++ e2.fVars
-}
-class FVarsVar(x: String) extends Var(x) with FVarsExpr {
-def fVars = Set(x)
-}
-\end{lstlisting}
-The above code already appears as boilerplate, especially, more classes have to be defined along with future extensions.
-It pollutes the namespace heavily.
-Moreover, the \lstinline{ExprParser} code is no longer useful; to generate expressions that support free variables, we have to substitute \lstinline{FVarsExpr}, \lstinline{FVarsLit}, \lstinline{FVarsAdd} and \lstinline{FVarsVar} for the old \lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var} correspondingly.
+\lstinputlisting[linerange=27-37]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=FVAR_CLASSES
+
+Besides introducing boilerplate code, more classes have to be defined along with future extensions.
+They pollutes the namespace heavily.
+More importantly, the \lstinline{VarExprParser} we defined is no longer suitable for parsing the new ASTs!
+All the methods in the previous parser has type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
+
+Therefore, in order to generate new AST structures, we have to substitute \lstinline{FVarsExpr}, \lstinline{FVarsLit}, \lstinline{FVarsAdd} and \lstinline{FVarsVar} for the old \lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var} correspondingly.
 Again, to avoid modifying existing code, we have to sacrifice code reuse.
-In fact, one would not like to make a choice from them, but expect both.
-We have to seek for a different approach to build ASTs.
-\haoyuan{Please check text.}
+In fact, there is a trade-off between separate compilation and type-safe code reuse.
+For semantic modularity, we need a different approach for building ASTs.
 

### 2017-01-13 Scala/Parser/src/PaperCode/Sec4OA/Code1.scala

@@ -15,8 +15,26 @@ object Code1 {
     def print = "(" + e1.print + " + " + e2.print + ")"
   }
 
-//BEGIN_ATTEMPT_EXPRWITHFREEVARS
+  class Var(x: String) extends Expr {
+    def print = x
+  }
+
+//BEGIN_FVARSEXPR
 trait FVarsExpr extends Expr { def fVars: Set[String] }
-//END_ATTEMPT_EXPRWITHFREEVARS
+//END_FVARSEXPR
+
+//BEGIN_FVAR_CLASSES
+class FVarsLit(x: Int) extends Lit(x) with FVarsExpr {
+  def fVars = Set()
+}
+
+class FVarsAdd(e1: FVarsExpr, e2: FVarsExpr) extends Add(e1, e2) with FVarsExpr {
+  def fVars = e1.fVars ++ e2.fVars
+}
+
+class FVarsVar(x: String) extends Var(x) with FVarsExpr {
+  def fVars = Set(x)
+}
+//END_FVAR_CLASSES
 
 }

### 2017-01-13 paper/sections/Discussion.tex
Swap 4.4 for 5.1

--- 2017-01-13 paper/sections/Discussion.tex
@@
 The use of inheritance-based approach and Object Algebras enables us to build modular parsers, which are able to evolve with syntax together.
-This section explores more interesting features, including overriding existing parsing rules, language components for abstracting language features, and alternative techniques under the whole framework.
+This section explores more interesting features, including parsing multi-sorted syntax, overriding existing parsing rules, language components for abstracting language features, and alternative techniques under the whole framework.
+
+\subsection{Parsing Multi-sorted Syntax}\label{subsec:differentsyntax}
+
+\begin{comment}
+As illustrated above, using Object Algebras separates data structures from behaviors, thus enabling more modularity and reuse.
+New language constructs correspond to the new cases in the algebra.
+Different operations
+ on structures, with both code reuse and separate compilation supported.
+\end{comment}
+
+Another advantage of using Object Algebras is that is is easy to
+model multi-sorted languages.
+If the syntax contains multiple sorts, we can distinguish them by different type parameters.
+As an example, we extend the language of literals, additions and variables by introducing types as a new kind of type argument, and we add lambda abstractions to expressions.
+Now the language has two sorts in the syntax, types and expressions:\\
+
+\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<type> ::= `int' \alt <type> `->' <type>
+\end{grammar}
+&
+\setlength{\grammarindent}{5em}
+\begin{grammar}
+<expr> ::=  ... \alt `\\' <ident> `:' <type> `.' <expr>
+\end{grammar}
+\end{tabular}
+
+\begin{figure}
+\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
+\label{fig:multi}
+\end{figure}
+
+Figure~\ref{fig:multi} illustrates the corresponding Scala code
+that extends the Object Algebra interface, pretty-printing operation and parser.
+We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
+They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
+Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions.
+We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue
+with \inlinecode{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
+
+From the code we observe that the multi-sorts of Object Algebra interface
+express multiple syntax. The isolation for different syntax
+is guaranteed by the type system. The following illustrates some client code
+for the new parser:
+
+\lstinputlisting[linerange=63-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+The code works as expected to parse and pretty-print the result.
 

### 2017-01-13 paper/sections/ObjectAlgebras.tex
Swap 4.4 for 5.1

@@
 \lstinputlisting[linerange=37-44]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
-\huang{done}\bruno{break code into 2 parts: the definition of the trait; and the client
-code.}
-\bruno{the following text should be for explaining the definition of the trait}
+
 \noindent The type of the object algebra field \lstinline{alg} is first refined
@@
 
-\begin{comment}
-At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice.
-However, that is generally solved~\cite{Oliveira2012} \huang{it was "original object algebra paper", i added the citation here, is it the correct one?
-i'll check later} by composing two algebras into one in advance.
-\end{comment}
-
-\subsection{Parsing Multi-sort Syntax}\label{subsec:differentsyntax}
-
-\begin{comment}
-As illustrated above, using Object Algebras separates data structures from behaviors, thus enabling more modularity and reuse.
-New language constructs correspond to the new cases in the algebra.
-Different operations
- on structures, with both code reuse and separate compilation supported.
-\end{comment}
-
-Another advantage of using Object Algebras is that is is easy to
-model multi-sorted languages.
-If the syntax contains multiple sorts, we can distinguish them by different type parameters.
-As an example, we extend the language of literals, additions and variables by introducing types as a new kind of type argument, and we add lambda abstractions to expressions.
-Now the language has two sorts in the syntax, types and expressions:\\
-
-\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<type> ::= `int' \alt <type> `->' <type>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::=  ... \alt `\\' <ident> `:' <type> `.' <expr>
-\end{grammar}
-\end{tabular}
-
-\begin{figure}
-\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
-\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
-\label{fig:multi}
-\end{figure}
-
-Figure~\ref{fig:multi} illustrates the corresponding Scala code
-that extends the Object Algebra interface, pretty-printing operation and parser.
-We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
-They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
-Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions.
-We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue
-with \inlinecode{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
-
-From the code we observe that the multi-sorts of Object Algebra interface
-express multiple syntax. The isolation for different syntax
-is guaranteed by the type system. The following illustrates some client code
-for the new parser:
-
-\lstinputlisting[linerange=63-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-The code works as expected to parse and pretty-print the result.
-
 Note that, similarly to the approach in Section~\ref{sec:inheritance} independent extensibility is also supported via multiple trait inheritance.
 Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance} we omit that code here.
+
+\subsection{Composition of Algebras}
+
+By adopting Object Algebras, we also inherit some advanced features from the pattern.
+Specifically, in the original paper~\cite{Oliveira2012}, there are two kinds of Object Algebras: \textit{queries} and \textit{transformations}.
+A query traverses the abstract syntax tree, collects information and returns the result.
+The aforementioned \lstinline{Print} and \lstinline{Eval} are traditional queries.
+But as we discussed before, an algebra is first fed to the parser, then \lstinline{parse} is invoked to do the parsing.
+It is tedious to fed \lstinline{Print} and \lstinline{Eval} separately, as the same input is parsed twice.
+Instead, we can firstly compose two queries into one, then feed it to the parser.
+For space reasons we omit the code here.
+
+Another interesting observation, if we look back at the end of Section~\ref{subsec:objectalgebras}, is that \lstinline{makeExp} represents an ``object'' of the expression AST.
+Or rather, what OA parsers generate are ``implicit objects'', which are actually functions of type \lstinline{Alg[E]} \lstinline{=>} \lstinline{E} for generic \lstinline{E}.
+Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
+Especially, in the design of a compiler
+there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
+in a very clever way. Continuing from the old example, if one wants to transform a \inlinecode{VarExprAlg} expression to an \inlinecode{ExprAlg} expression by
+replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}
+
+\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+Here the \inlinecode{VarExprAlg} expression \inlinecode{"1 + x"} is changed to \inlinecode{ExprAlg} expression \inlinecode{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
+\inlinecode{alg} in \inlinecode{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
+to \inlinecode{alg}, \inlinecode{Refactor} will be able to transform a \inlinecode{VarExprAlg} expression before we obtain the result.
+Furthermore, transformations can be composed in a linear pipeline-style, before a query algebra is fed ~\cite{Zhang2015}.

### 2017-01-13 paper/sections/Inheritance.tex
Fix refs in Section 2 & 3

@@
 Object-oriented ASTs are ASTs encoded via standard OO techniques,
-such as the {\sc Composite} or {\sc Interpreter} patterns~\cite{gamma94design}. Such ASTs
+such as the {\sc Composite} or {\sc Interpreter} patterns~\cite{gamma1995design}. Such ASTs
 fix the operations required in the AST, but allow the easy addition of new
@@
 adding new operations is easy, but adding new language constructs is hard.
-The AST in Figure~\ref{} provides a simple example of an OO AST. The same
-AST could be encoded using algebraic datatypes (or alternatively Visitors)
+In Section~\ref{subsec:packratparsing} we provide a simple example of OO ASTs. The same abstract syntax could be encoded using algebraic datatypes (or alternatively Visitors)
 as:
@@
 
-Unfortunatelly, when parsing OO style ASTs, there are difficulties with
+Unfortunately, when parsing OO style ASTs, there are difficulties with
 the two dimensions of extensibility. On the one hand it is naturally hard to
@@
 language constructs is problematic. Therefore,
-it is perhaps not a suprise that parser generators that also generate
-AST code (auch as ANTLR~\cite{}), normally generate code based on
+it is perhaps not a surprise that parser generators that also generate
+AST code (such as ANTLR~\cite{parr2013definitive}), normally generate code based on
 visitors. Clearly this seems a better choice since generating code
@@
 hence the parser does not work as expected. It is possible to build the correct parser by replacing the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr}.
-However, modification on existing code sacrifices separate compilation, as mentioned by the Expression Problem.
+However, modification on existing code sacrifices separate compilation, and hence breaks semantic modularity.
 
@@
 A nice feature of Scala is its support for a form of multiple inheritance.
-Scala has a linearized-style multiple inheritance for traits~\cite{}.
+Scala has a linearized-style multiple inheritance for traits~\cite{}\huang{what's the proper citation here?}.
 This can be very helpful when composing several languages, which
@@
 
-\huang{added explanation of the `super' keyword}
-
 The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}.
@@
 Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.
-
-\huang{added a concrete example}
 

### 2017-01-13 paper/sections/PackratParsing.tex
Fix refs in Section 2 & 3

@@
 used by non-modular code using parsing combinators.
-To avoid pitfalls related to those algortihmic challenges,
-we propose a the following methodology:
+To avoid pitfalls related to those algorithmic challenges,
+we propose the following methodology:
 
@@
 The construction procedure is explicit and fully controlled by the programmer.
-Secondly, each parser combinator is represented by a piece of code, and also are the parsers it takes.
+Secondly, each parser combinator is represented by a piece of code, also are the parsers it takes.
 Thus in a statically typed programming language they can be statically type-checked.
 Unfortunately many parser combinators have important limitations.
-In particular several parser combinators,
-including the famous Parsec~\cite{Leijen2001} library, require
+In particular, several parser combinators including the famous Parsec~\cite{Leijen2001} library, require
 programmers to manually do \textit{left-recursion elimination}, \textit{longest match composition}, and
@@
 as well as its corresponding parser. In a modular setting, it is
-possible but unnecessarily compilcated to analyse the grammar and
+possible but unnecessarily complicated to analyse the grammar and
 rewrite it when doing extension. Anticipating that every non-terminal
@@
 the \emph{longest match} is picked among the possible results.
-However, manual reording for the longest match is inconvenient, and worst still, it is essentially non-modular.
+However, manual reordering for the longest match is inconvenient, and worst still, it is essentially non-modular.
 When the grammar is extended with new rules, programmers are supposed to \emph{manually} adjust
-the order of parsers, which requires rewritting previously written code.
+the order of parsers, which requires rewriting previously written code.
 
@@
 \subsection{Packrat Parsing}\label{subsec:packratparsing}
-Fortunately some more advanced parsing techniques such as Packrat
+Fortunately, some more advanced parsing techniques such as Packrat
 parsing~\cite{Ford2002} have been developed to address limitations of
@@
 It extends traits \inlinecode{StandardTokenParsers} and \inlinecode{PackratParsers} from the Scala parser combinator library.
-Futhermore, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing.
+Furthermore, we will use \lstinline{Parser} as a type synonym for \lstinline{PackratParser} and a generic \inlinecode{parse} function for testing.
 
@@
 recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the longest match alternative combinator \inlinecode{|||}.
-Table~\ref{tab:packrat} gives a closer look at part of the API used in our work.
+Table~\ref{tab:packrat} gives a closer look at the parser combinators used in our work.
 
@@
 \end{tabular}
-\caption{Part of Scala Parser API.\bruno{We should double-check that this table lists the
-  combinators used in the paper.}}\label{tab:packrat}
+\caption{Combinators used in our work, from the Scala standard parser combinator library.}\label{tab:packrat}
 \end{table}

### 2017-01-13 paper/sections/CaseStudy.tex
Fix sec 4 5 6

@@
 Figure \ref{fig:dependency} shows the
-dependency of all the components and calculi in our case study. Gray
+dependency of all the components and calculi in our case study. Grey
 boxes are calculi and white boxes are components. An arrow starting
@@
 execution time may be the overall structure of the modular parsing approach, because we indeed have
-more intermediate function calls and method overriding. However, it is worth metioning that
+more intermediate function calls and method overriding. However, it is worth mentioning that
 because of the memoization technique of Packrat parsers, we are only constant times

### 2017-01-13 paper/sections/Discussion.tex
Fix sec 4 5 6

@@
 
-Another advantage of using Object Algebras is that is is easy to
+Another advantage of using Object Algebras is that it is easy to
 model multi-sorted languages.
@@
 
-From the code we observe that the multi-sorts of Object Algebra interface
+From the code, we observe that the multi-sorts of Object Algebra interface
 express multiple syntax. The isolation for different syntax
@@
 
-One can even ``eliminate'' a production rule in the extension, by actually overriding it with a failure parser.
-The lexer can also be updated, since keywords and delimiters are represented by sets of strings.
+One can even ``eliminate'' a production rule in the extension, by overriding it with a failure parser. The lexer can also be updated, since keywords and delimiters are represented by sets of strings.
 
@@
 
-As shown below, those two language components can be merged into one to obtain the language we want. Futhermore, the new language is still a modular
+As shown below, those two language components can be merged into one to obtain the language we want. Furthermore, the new language is still a modular
 component ready for future composition. In that case modularity is realized over higher-order hierarchies.
@@
 drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty-printers of
-\lstinline{VarExpr} and \lstinline{TypedLam} respectively. Such a pattern refers to \textit{family polymorphism}[]. Unfortunately Scala does not
-support this feature, nonetheless, one can avoid such boilerplate by using metaprogramming techniques.
+\lstinline{VarExpr} and \lstinline{TypedLam} respectively.
+Such a pattern refers to \textit{family polymorphism}~\cite{ernst01FP} which is unfortunately not supported in Scala.
+Nonetheless, one can avoid such boilerplate by using metaprogramming techniques.
 
@@
 Nevertheless, \textit{open recursion} could be used as an alternative.
-An explict ``self-reference'' parameter is able to explain the recursive calls dynamically, by the real argument passed at runtime.
+An explicit ``self-reference'' parameter is able to explain the recursive calls dynamically, by the real argument passed at runtime.
 
@@
 
-Besides Object Algebras, many other techniques including \textit{Data types à la carte} (DTC) [] and \textit{Modular reifiable matching} (MRM) [] also support extensible data structures.
+Besides Object Algebras, many other techniques including \textit{Data types à la carte} (DTC)~\cite{swierstra2008data} and \textit{Modular reifiable matching} (MRM)~\cite{Oliveira2015} also support extensible data structures.
 These alternatives can also be adopted to build ASTs and corresponding parsers.

### 2017-01-13 paper/sections/ObjectAlgebras.tex
Fix sec 4 5 6

@@
 show how to support both the forms of extensibility for ASTs (easy
-addition of language contructs, and easy addition of operations)
+addition of language constructs, and easy addition of operations)
 using Object Algebras~\cite{Oliveira2012}.
@@
 
-The Expression Problem illustrates the difficulty of extending data structures or ASTs in two dimensions.
+The Expression Problem~\cite{wadler1998expression} illustrates the difficulty of extending data structures or ASTs in two dimensions.
 For traditional OO ASTs, adding new language constructs is easy, but adding new operations is hard.
@@
 Object Algebras captures a design pattern to address the Expression Problem, achieving two dimensions of extensibility (language constructs and operations) in a modular and type-safe way.
-The definition of data structures is separated from their behaviors, and future extensions on both dimensions no longer require existing code to be modified, supporting separate compilation.
+The definition of data structures is separated from their behaviours, and future extensions on both dimensions no longer require existing code to be modified, supporting separate compilation.
 
 Using Object Algebras in Scala, ASTs as recursive data structures are defined by traits, where each constructor corresponds to an abstract method inside.
-Essentially, Object Algebras generalize the {\sc Abstract Factory} pattern~\cite{}, and promote the use of factory methods, instead of constructors, for instantiating objects.
+Essentially, Object Algebras generalize the {\sc Abstract Factory} pattern~\cite{gamma1995design}, and promote the use of factory methods, instead of constructors, for instantiating objects.
 The example from Section~\ref{subsec:packratparsing} is used here again for illustration.
@@
 Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
-One can define evaluation operation as a new trait on this expression langauge:
+One can define evaluation operation as a new trait on this expression language:
 
@@
 \paragraph{Adding New AST Constructs}
-Futhermore, new language constructs can be added by extending \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
+Furthermore, new language constructs can be added by extending \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
 with variables. A new Object Algebra interface \inlinecode{VarAlg} is defined as follows:
@@
 
-Parsing builds the AST as a result. When Object Algebras are used
+Parsing produces ASTs as the result. When Object Algebras are used
 to build ASTs, an Object Algebra containing the constructor/factory
@@
 field has a type which is a \emph{subtype} of the original field.
-Unfortunatelly, in this case, although \inlinecode{VarExprAlg} is a subtype
+Unfortunately, in this case, although \inlinecode{VarExprAlg} is a subtype
 of \inlinecode{ExprAlg}, the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} is a supertype of \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.
@@
 \subsection{Composition of Algebras}
-
+\huang{to revise}
 By adopting Object Algebras, we also inherit some advanced features from the pattern.

### 2017-01-13 paper/sections/RelatedWork.tex
Fix sec 4 5 6

@@
 la Carte'' (DTC)
-paper~\cite{swierstra2008} represents modular ASTs using co-products
+paper~\cite{swierstra2008data} represents modular ASTs using co-products
 of every two functors, which forms a binary tree structure.

### 2017-01-13 paper/sections/ObjectAlgebras.tex
Fix refs, spell check

@@
 addition of language constructs, and easy addition of operations)
-using Object Algebras~\cite{Oliveira2012}.
+using Object Algebras~\cite{Oliveira:2012}.
 
@@
 
-Fortunately, Object Algebras~\cite{Oliveira2012} enables us to solve this problem. As a nice solution, it separates data variants
+Fortunately, Object Algebras~\cite{Oliveira:2012} enables us to solve this problem. As a nice solution, it separates data variants
 and operations, and offers high flexibility in the choice of operations to be performed over ASTs. Before adopting it for parsing, the next section introduces the background of Object Algebras.
@@
 
-\subsection{Composition of Algebras}
+\subsection{Composition of Algebras}\label{subsec:more-oa}
 \huang{to revise}
 By adopting Object Algebras, we also inherit some advanced features from the pattern.
-Specifically, in the original paper~\cite{Oliveira2012}, there are two kinds of Object Algebras: \textit{queries} and \textit{transformations}.
+Specifically, in the original paper~\cite{Oliveira:2012}, there are two kinds of Object Algebras: \textit{queries} and \textit{transformations}.
 A query traverses the abstract syntax tree, collects information and returns the result.

### 2017-01-13 paper/sections/RelatedWork.tex
Fix refs, spell check

--- 2017-01-13 paper/sections/RelatedWork.tex
@@
 parser generators may statically check the correctness and unambiguity
-of grammars. In constrast, because our approach is based on parser
+of grammars. In contrast, because our approach is based on parser
 combinators, there is no support for ambiguity checking.  However, as
@@
 way. However such modularity is not flexible enough for language
-unification, as the syntax is only built from extensions.  Moreover,
-existing macros cannot be further changed after
-definition. \haoyuan{correct?}  Extensible
+unification, as the syntax is only built from extensions.
+%Moreover,
+%existing macros cannot be further changed after
+%definition. \haoyuan{correct?}
+Extensible
 compilers like JastAdd~\cite{Ekman2007} and
@@
 is mostly done using parser generators. They focus on the
-extensions to a host language. Our library is designed for type-safe 
+extensions to a host language. Our technique is designed for type-safe
 modular language parsers (not just language
@@
 is quite different from ours, since it uses parse
- tables, wheres we use parser combinators. 
+ tables, whereas we use parser combinators.
 Our approach supports both
@@
 high cost for backtracking and longest match composition issues,
-as we discussed in Section~\ref{}. Those limitations make Parsec 
+as we discussed in Section~\ref{subsec:challenges}. Those limitations make Parsec
 (and similar parsing techniques) inadequate for modular parsing.
@@
 %demonstrated that general left-recursion is supported theoretically.
-Despite our choice of Packrat parsing, as argued in Section~\ref{}, 
+Despite our choice of Packrat parsing, as argued in Section~\ref{sec:packrat},
 any parser combinator approach which supports left-recursion, longest
@@
 
-\paragraph{Extensibility} Various design patterns~\cite{}, in multiple
+\paragraph{Extensibility} Various design patterns~\cite{}\huang{???} in multiple
 languages, have been proposed over the years to address extensibility
-problems, such as the Expression Problem~\cite{}. However, to the best
+problems, such as the Expression Problem~\cite{wadler1998expression}. However, to the best
 of our knowledge, none of the existing extensibility techniques
@@
 object-oriented programming.  For instance, the famous ``Datatypes \`a
-la Carte'' (DTC)
-paper~\cite{swierstra2008} represents modular ASTs using co-products
+la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
 of every two functors, which forms a binary tree structure.
-Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}
+Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
 All of that work essentially covers how to traverse and
@@
 In OO languages, there are also many design patterns that achieve
-type-safe extensibility~\cite{torgerson,odersky,Oliveira09,Oliveira12,Wang16}. 
-We chose Object Algebras~\cite{Oliveira2012} because the pattern is 
+type-safe extensibility~\cite{torgerson,odersky,Oliveira09,Oliveira:2012,Wang16}.\huang{not in the bib file, pls fix}
+We chose Object Algebras~\cite{Oliveira:2012} because the pattern is
 relatively lightweight and makes good use of existing OO features,
@@
 
-It is worth mentioning that Scala case classes~\cite{} provide a near
+It is worth mentioning that Scala case classes~\cite{}\huang{???} provide a near
 solution to the Expression Problem. Case classes are similar to
@@
 exhaustiveness of pattern matching for extensible operations. In other
-words run-time pattern matching errors can happen when writting
-extensible code with case classes.  So full static type-safety is not
-ensured. Nevertheless case classes are a pragmatic approach, which is
-widely used in practice. Therefore modular parsing techniques may be
+words run-time pattern matching errors can happen when writing extensible code with case classes.
+So that full static type-safety is not ensured.
+Nevertheless case classes are a pragmatic approach, which is
+widely used in practice. Therefore, modular parsing techniques may be
 of value for extensible code using case classes.  The approach we
-presented in Section 3~\ref{} can readily be adapted to case classes:
+presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
 all that the users need to do is to use case classes, instead of
@@
 \begin{comment}
-Moreover, in~\cite{Oliveira2012} the authors have discussed the
+Moreover, in~\cite{Oliveira:2012} the authors have discussed the
 composition of algebras. In our parsing approach, a parser consumes an
@@
 
-\bruno{There's a problem with the text in 4.2. We need to first 
-fix the problem in S4.2 and then come back to the text below and 
-adapt it accordingly.}
 Another interesting observation, as we have mentioned
-in~\ref{subsec:objectalgebras}, is that transformation algebras are
+in Section~\ref{subsec:more-oa}, is that transformation algebras are
 frequently used in parsing, as concrete syntax may not always coincide

### 2017-01-14 paper/sections/Inheritance.tex
Fix refs, spell check
Fix some refs

@@
 
-We already have solutions such Object Algebras~\cite{Oliveira2012} for this problem.
+We already have solutions such Object Algebras~\cite{Oliveira:2012} for this problem.
 However, when ASTs evolve, the corresponding parsers which produce such structures should also change accordingly.
@@
 A nice feature of Scala is its support for a form of multiple inheritance.
-Scala has a linearized-style multiple inheritance for traits~\cite{}\huang{what's the proper citation here?}.
-This can be very helpful when composing several languages, which
+Scala has a linearized-style multiple inheritance for traits~\cite{odersky2004overview}. This can be very helpful when composing several languages, which
 may have been independently developed. Suppose now

### 2017-01-14 paper/sections/RelatedWork.tex
Fix some refs

@@
 In OO languages, there are also many design patterns that achieve
-type-safe extensibility~\cite{torgerson,odersky,Oliveira09,Oliveira:2012,Wang16}.\huang{not in the bib file, pls fix}
+type-safe extensibility~\cite{torgersen2004expression,odersky2005independently,oliveira2009modular,Oliveira:2012,wang2016expression}.
 We chose Object Algebras~\cite{Oliveira:2012} because the pattern is

### 2017-01-14 paper/sections/RelatedWork.tex
All refs are fixed except Introduction

@@
 
-\paragraph{Extensibility} Various design patterns~\cite{}\huang{???} in multiple
+\paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
 languages, have been proposed over the years to address extensibility
@@
 
-It is worth mentioning that Scala case classes~\cite{}\huang{???} provide a near
+It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
 solution to the Expression Problem. Case classes are similar to

### 2017-01-14 paper/paper.tex
Intro

@@
 \titlerunning{Type-Safe Modular Parsing}
+\author[1]{}
 
 %% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
-\author[1]{John Q. Open}
-\author[2]{Joan R. Access}
-\affil[1]{Dummy University Computing Laboratory, Address/City, Country\\
-  \texttt{open@dummyuniversity.org}}
-\affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
-  \texttt{access@dummycollege.org}}
-\authorrunning{J.\,Q. Open and J.\,R. Access} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
+%\author[1]{John Q. Open}
+%\author[2]{Joan R. Access}
+%\affil[1]{Dummy University Computing Laboratory, Address/City, Country\\
+%  \texttt{open@dummyuniversity.org}}
+%\affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
+ % \texttt{access@dummycollege.org}}
+%\authorrunning{J.\,Q. Open and J.\,R. Access} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
 
-\Copyright{John Q. Open and Joan R. Access}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+%\Copyright{John Q. Open and Joan R. Access}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 
-\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
+%\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
 E.g., cite as "F.1.1 Models of Computation".
-\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
+%\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
 % Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-01-14 paper/sections/Introduction.tex
Intro

@@
 practice, due to their simplicity of implementation and use. 
-Examples include many tools for developing Software-Product
-Lines~\cite{}, some Language Workbenches~\cite{}, or extensible parser
+Examples include many tools for developing Feature-Oriented Software-Product
+Lines (SPLs)~\cite{Kastner11road}, some Language Workbenches~\cite{}, or extensible parser
 generators~\cite{}.  Most syntactic approaches employ textual
 composition techniques such as \emph{superimposition}~\cite{} to
-enable the development modular program features. Such textual
+enable the development modular program features. Such 
 composition techniques collect the code for multiple features and
-merge it together when a concrete combination of features is needed
-for a particular program. As Kastner et. al~\cite{Kastner11road} note, the typical
-drawback of such techniques is that
-``\emph{most feature-oriented implementation mechanisms lack proper
+merge them together when a concrete combination of features is needed
+for a particular program. As Kastner et. al~\cite{Kastner11road} note, 
+a typical drawback of feature-oriented SPL implementations, which 
+more generally applies to syntactic modularity approaches, is that
+such ``\emph{implementation mechanisms lack proper
   interfaces and support neither modular type checking nor separate
@@
 their own syntax, such as SugarJ~\cite{Erdweg2011}, also offer a form
-of extensible parsing. However these syntactic approaches do not
-support separate compilation and/or modular type-checking of parsing
-code either.
+of extensible parsing. However since those approaches are syntactic
+they do not support separate compilation and/or modular type-checking
+of parsing code either.
 
@@
 \emph{virtual classes}~\cite{Ernst:2006}, as 
-well as several other techniques for solving the Expression
+well as various techniques for solving the Expression
 Problem~\cite{Oliveira:2012}\bruno{more here; yanlin; Torgersen; Odersky...}.
@@
 Semantic modularization techniques are less widely used in practice
-than syntactic techniques. This is partly due to the need for more
+than syntactic techniques. This is partly due to the perceived need for more
 sophisticated type systems, which are not available in mainstream
-languages and may require more knowledge from users. However, recently
+languages and may require more knowledge from users. However, recently,
 several lightweight modularization techniques have been shown to work
@@
 Algebras~\cite{Oliveira:2012} are one such technique, which works in
-Java-like languages and uses simple generics only. So far research on
-semantic modularization techniques have focused on operations that
-\emph{traverse} or or \emph{process} extensible datastructures, such
-as ASTs. Indeed many documented applications of semantic
-modularization techniques focus on modularizing various aspects of PL
-implementations.  However, as far as we know there is little work on
-operations that build/produce ASTs.  In particular the problem of how
-to modularize parsing has not been studied in semantic modularization
-approaches. This is a shame because parsing is a fundamental part of
-PL implementations, and it ought to be modularized as well.
+Java-like languages and uses simple generics only.
+
+So far research on semantic modularization techniques have focused on
+operations that \emph{traverse} or \emph{process} extensible
+datastructures, such as ASTs. Indeed many documented applications of
+semantic modularization techniques focus on modularizing various
+aspects of PL implementations.  However, as far as we know, there is
+little work on operations that build/produce ASTs.  In particular the
+problem of how to modularize parsing has not been studied in semantic
+modularization approaches. At best, techniques such as NOA~\cite{Gouseti2014}
+employ a syntactic modularity approach for parsing in combination with 
+a semantic modularity approach for defining other language components 
+that traverse or process ASTs. This is a shame because parsing is a
+fundamental part of PL implementations, and it ought to made
+semantically modular as well, so that the full benefits of semantic 
+modularity apply.
 
@@
 
-This paper investigates presents a technique for doing semantically
+This paper presents a technique for doing semantically
 modular parsing.  That is, our approach not only allows complete
@@
 those parsing components to be \emph{modularly type-checked} and
-\emph{separately compiled}. In developing our techniques we encontered 
-two different classes of challenges:
+\emph{separately compiled}. Developing techniques for modular parsing 
+is not without challenges. In developing our techniques we encontered 
+two different classes of challenges: algorithmic challenges; and
+typing/reuse challenges.
 
-\begin{itemize}
-
-\item {\bf Algorithmic challenges:} A first challenge was todo with
+\paragraph{Algorithmic Challenges} A first challenge was todo with
   the parsing algorithms themselves. Since many parsing formalisms
   were not designed with extensibility in mind, some of the employed
-  techniques were non-modular, or performed poorly in a modular setting.
+  techniques were non-modular, or performed poorly in a modular
+  setting.  The most widely use tools for parsing are parser
+  generators, but most algorithms employed by parser generators
+  require full information about the grammar to generate parsing
+  code. Moreover, actions associated to grammar productions are
+  typically only type-checked after the parser has been
+  generated. Both problems go against our goals of semantic
+  modularity.
 
-\item {\bf Typing and Reusability Challenges:} The second class of
+  An alternative to parser generators are \emph{parser
+    combinators}~\cite{burge1975,Wadler1985}.  At a first look, parser
+  combinators seem very suitable for our purpose. Each parser
+  combinator is represented by a piece of code directly in the
+  programming language. Thus, in a statically typed programming
+  language, such code is statically type-checked.  However many
+  techniques regularly employed by parser combinators cause difficulties in a
+  modular setting. In particular, many parser combinator approaches
+  (including Parsec~\cite{}) routinely use \emph{left-recursion
+    elimination}, \emph{priority-based matching}, and \emph{avoid
+    backtracking} as much as possible. All of these are problematic in
+  a modular setting as illustrated in Section~\ref{}.
+
+  To address such algorithmic challanges, we propose a methodology 
+  for implementing modular parsers built on top of an
+  existing Packrat~\cite{} parsing library for Scala~\cite{}. The Packrat parsing library
+  directly supports left-recursion, and a \emph{longest-match
+    composition} operator. The direct support for left-recursion
+  avoids the need for left-recursion elimination, which is very
+  cumbersome is a modular setting. The longest-match composition 
+  operator avoids the use of priorities, and it provides reasonably
+  efficient optimizations for backtracking.
+
+  \paragraph{Typing and Reusability Challenges} The second class of
   challenges were problems related to modularity, reusability and
-  typing.
+  typing of parsing code. An immediate concern is how to extend a
+  parser for an existing language or, more generally, how to compose
+  parsing code for two languages. It turns out that OO mechanisms that
+  provide some form of multiple inheritance,
+  such as traits/mixins~\cite{scalaoverview}\bruno{cite work on traits (Ducasse)
+  and mixins (Bracha)}, are very handy for this
+  purpose. Essentially, traits/mixins can act as modules for the parsing code
+  of different languages. This enables an approach where ASTs can be
+  modelled using standard OO techniques such as the {\sc Composite} 
+  pattern, while retaining the possibility of adding new language
+  constructs. Section~\ref{} gives the details of this approach.
 
-\end{itemize}
+  Our ultimate goal is to allow for full extensibility: it should be
+  possible to not only modularly add both new language constructs, but
+  also new operations. To accomplish this goal one final tweak on our
+  technique is to employ Object Algebras to allow fully extensible
+  ASTs. Thus a combination of Packrat parsing, multiple inheritance 
+  and Object Algebras enables a solution for semantically modular 
+  parsing. Section~\ref{} gives the details of the complete approach.
 
-\name is built on top of a Packrat parsing
-library, but adds new parsing combinators to enable modular
-parsing. The new parsing combinators employ \emph{delegation-based}
-techniques and \emph{Object Algebras} to support extensibility. The
-choice of Packrat parsing over other parsing techniques turns out to
-be important for achieving performance in a modular
-setting. \bruno{left recursion, and backtracking removal.}
-
+\begin{comment}
   By analising the \emph{full} grammar it is possible to remove
@@
   this has a severe impact on performance.
+\end{comment}
 
-  To evaluate \name we conduct a case study based on the ``Types and
-  Programming Languages'' (TAPL) interpreters.  The case study shows
-  that \name is effective at reusing parsing code from existing
-  interpreters, and the total parsing code is 60\% shorter than the
-  non-modular parsing code.\bruno{comment on the efficiency}
+To evaluate our approach we conduct a case study based on the ``Types
+and Programming Languages'' (TAPL) interpreters. The case study shows
+that our approach is effective at reusing parsing code from existing
+interpreters, and the total parsing code is 69\% shorter than an
+existing code base using non-modular parsing code~\cite{}.
 
@@
 
-\item {\name:} A parser combinator library that allows the development 
-of modular parser. The library uses \emph{delegation} and
-\emph{Object-Algebras} to achieve modularity and extensibility.
+\item {A technique for Modular Parsing} We present a technique 
+that allows the development semantically modular parsers.
+The technique relies on the combination of Packrat parsing, multiple inheritance 
+  and Object Algebras.
 
@@
 
-\item {{\bf Limitations of existing Parser Combinator Techniques:}}
-\bruno{Improve and write something here.}
+\item {{\bf A Methodology for Writing Modular Parsing Code:}} We
+  identify possible pitfalls using parser combinators. To avoid such
+  pitfalls, we propose guidelines for writing parsing code using 
+  \emph{left-recursion} and \emph{longest-match composition}.
 
@@
   from the TAPL book. The case study show the effectiveness of modular 
-  parsing in terms of reuse.
+  parsing in terms of reuse. The TAPL case study is available online
+  at:
+  \url{http://}
 
@@
 
-\haoyuan{I suggest this part to be moved to Introduction, and only discuss why Packrat is selected among
-different parser combinators in Overview.}
-
-\begin{comment}
-Although there are many parsing techniques, not all of them are
-suitable for type-safe modular parsing. In particular there are many
-techniques which fail to provide modular type-checking and separate
-compilation. Moreover, even if modular type-checking and separate
-compilation are supported, efficiency is another
-concern. A parsing technique should have low overhead when applied
-in a modular setting. In the remaider of this section, we eliminate
-various techniques that fail to satisfy our requirements, and argue
-that that Packrat parsing~\cite{Ford2002} is a suitable candidate for
-type-safe modular parsing.
-
-\paragraph{Parser Generators} The most widely use tools for parsing
-are parser generators. Parser generators help users generate parsers automatically or
-semi-automatically from a given grammar. There is no restriction on
-the algorithm, while most of them adopts table-based LL~\cite{lewis1968syntax} and LR~\cite{knuth1965translation} parsing
-algorithms.
-Although efficient, the main drawback of parser generators is that they do not support
-modular type-checking and separate compilation.
-
-Modular parsing based on parser generators is supported by many libraries~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.
-Users can separate the syntax definition and related parsing code into reusable components.
-Then the corresponding parsers are built by their library utilities.
-For example, NOA~\cite{Gouseti2014} uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers.
-However, such generation procedure requires a whole compilation after the collection of all grammar pieces.
-Once the grammar changed, even slightly, grammar information must be re-collected and the parser must be re-generated.
-Hence those libraries only have syntactic modularity.
-
-Generating parsers often requires full information of grammars, thus semantic modularity is difficult to achieve in this way.
-
-\paragraph{Parser Combinators}
-Comparing with the parser generators, a \textit{parser combinator}~\cite{burge1975,Wadler1985}
-takes several parsers and produce a new parser as its output. Parser combinators are
-popular in functional programming, where the parsers are represented
-by functions and parser combinators are higher-order functions accepts
-them.
-
-At a first look, parser combinators are very suitable for our purpose, because of two
-reasons. Firstly, they are naturally modular. The manner of using them
-is to write small parsers and use combinators to composed them
-together. The construction procedure is explicit and fully controlled
-by the programmer. Secondly, each parser combinator is represented by
-a piece of code, and also are the parsers it takes. Thus in a
-statically typed programming language they can be statically
-type-checked.
-\end{comment}
+All the code in the paper and the case study is written in Scala. But,
+other languages that support some form of multiple inheritance
+(including Java 8 with default methods~\cite{}) could in principle be
+used. Our choice for Scala had todo with the concise and elegant syntax  
+of the language which enable a clear presentation of our ideas.

### 2017-01-14 paper/sections/PackratParsing.tex
Section 2

@@
 This section discusses the algorithmic challenges introduced by
-modular parsing and argues that Packrat parser combinators
-are suitable to address those challenges. The algorithm challenges
+modular parsing and argues that Packrat parser combinators~\cite{Ford2002}
+are suitable to address those challenges. The algorithmic challenges
 are important because they rule out various common techniques
@@
 expression $1-2-3$ is parsed as $(1-2)-3$ in the left-recursive
-grammar, but after rewrite the result is $1((-2)-3))$. The parse tree
-must be transformed to recover its structure.
+grammar, but after rewrite the result is $1((-2)-3))$\bruno{??This is
+  not a valid expression}. The parse tree
+must be transformed to recover the correct syntactic structure.
 
 \paragraph{Longest Match Composition} Another problematic issue
-in parser combinator libraries is the need for manually ordering
+in parser combinator libraries is the need for manually prioratizing/ordering
 alternatives in a grammar.
@@
 
-\paragraph{Backtracking} The need for backtracking is also problematic
+\paragraph{Backtracking} The need for backtracking can also be problematic
 in a modular setting. For example, consider a grammar that includes
@@
 
-\noindent However, this results in the worst case exponential time complexity in Parsec, because it does not have related optimization.
+\noindent However, this results in the worst case exponential time
+complexity in Parsec, because it does not have proive optimizations
+for backtracking.
 
@@
 \cite{moors2008parser} which implements Packrat parsers.
-The library provides a number of parser combinators, including the longeset match alternative combinator.
-Below we present an example to illustrate Scala parsers.
+The library provides a number of parser combinators, including the longest match alternative combinator.
+%Below we present an example to illustrate Scala parsers.
 
@@
 Suppose we want to parse a simple language with literals and
-additions. The concrete syntax is as follows:
+additions. The concrete syntax is:
 
@@
 
-It is straightforward to model the abstract syntax by classes as below. The ASTs support pretty-printing via the \lstinline{print} method.
+It is straightforward to model the abstract syntax by classes. The ASTs support pretty-printing via the \lstinline{print} method.
 
@@
 Our parser also employs longest match composition by using the combinator \lstinline{|||} in the library.
-Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees the efficiency.
+Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees reasonable efficiency.
 

### 2017-01-14 paper/paper.tex
Abstract. Introduction: halfway.

@@
   \emph{modularly type-checked} and \emph{separately compiled}. We
-  present a technique based parsing combinators that enables modular
+  present a technique based on parsing combinators that enables modular
   parsing. Interestingly, the modularity requirements for modular
@@
   the approach we conduct a case study based on the ``Types and
-  Programming Languages'' interpreters. The case study the
+  Programming Languages'' interpreters. The case study shows the
   effectiveness at reusing parsing code from existing interpreters,
   and the total parsing code is 69\% shorter than an existing code
-  based using non-modular parsing code.
+  base using non-modular parsing code.
 

### 2017-01-14 paper/sections/Introduction.tex
Abstract. Introduction: halfway.

@@
 
-To address the modularity limitations of Programming Languages several
+To address the modularity limitations of Programming Languages, several
 different approaches have been proposed in the past. Existing
@@
 well as various techniques for solving the Expression
-Problem~\cite{Oliveira:2012}\bruno{more here; yanlin; Torgersen; Odersky...}.
-
+Problem~\cite{Oliveira:2012,wang2016expression,torgersen2004expression,odersky2005independently}.
 Semantic modularization techniques are less widely used in practice
@@
 
-So far research on semantic modularization techniques have focused on
+So far research on semantic modularization techniques has focused on
 operations that \emph{traverse} or \emph{process} extensible
-datastructures, such as ASTs. Indeed many documented applications of
+data structures, such as ASTs. Indeed many documented applications of
 semantic modularization techniques focus on modularizing various
@@
 that traverse or process ASTs. This is a shame because parsing is a
-fundamental part of PL implementations, and it ought to made
+fundamental part of PL implementations, and it ought to be made
 semantically modular as well, so that the full benefits of semantic
@@
 \emph{separately compiled}. Developing techniques for modular parsing
-is not without challenges. In developing our techniques we encontered 
+is not without challenges. In developing our techniques we encountered
 two different classes of challenges: algorithmic challenges; and
@@
 
-\paragraph{Algorithmic Challenges} A first challenge was todo with
+\paragraph{Algorithmic Challenges} A first challenge was to do with
   the parsing algorithms themselves. Since many parsing formalisms
@@
   modular setting. In particular, many parser combinator approaches
-  (including Parsec~\cite{}) routinely use \emph{left-recursion
+  (including Parsec~\cite{Leijen2001}) routinely use \emph{left-recursion
     elimination}, \emph{priority-based matching}, and \emph{avoid
     backtracking} as much as possible. All of these are problematic in
-  a modular setting as illustrated in Section~\ref{}.
+  a modular setting as illustrated in Section~\ref{subsec:challenges}.
 
@@
   for implementing modular parsers built on top of an
-  existing Packrat~\cite{} parsing library for Scala~\cite{}. The Packrat parsing library
+  existing Packrat~\cite{Ford2002} parsing library for Scala~\cite{}. The Packrat parsing library
   directly supports left-recursion, and a \emph{longest-match
@@
   \paragraph{Typing and Reusability Challenges} The second class of
-  challenges were problems related to modularity, reusability and
+  challenges was problems related to modularity, reusability and
   typing of parsing code. An immediate concern is how to extend a
@@
   pattern, while retaining the possibility of adding new language
-  constructs. Section~\ref{} gives the details of this approach.
+  constructs. Section~\ref{sec:inheritance} gives the details of this approach.
 
   Our ultimate goal is to allow for full extensibility: it should be
-  possible to not only modularly add both new language constructs, but
+  possible to modularly add not only new language constructs, but
   also new operations. To accomplish this goal one final tweak on our
@@
   and Object Algebras enables a solution for semantically modular
-  parsing. Section~\ref{} gives the details of the complete approach.
+  parsing. Section~\ref{sec:algebrasandparsing} gives the details of the complete approach.
 
@@
 
-\item {A technique for Modular Parsing} We present a technique 
-that allows the development semantically modular parsers.
+\item {{\bf A technique for Modular Parsing:}} We present a technique
+that allows the development of semantically modular parsers.
 The technique relies on the combination of Packrat parsing, multiple inheritance
@@
 (including Java 8 with default methods~\cite{}) could in principle be
-used. Our choice for Scala had todo with the concise and elegant syntax  
-of the language which enable a clear presentation of our ideas.
+used. Our choice for Scala had to do with the concise and elegant syntax
+of the language which enables a clear presentation of our ideas.

### 2017-01-14 paper/sections/Conclusion.tex

--- 2017-01-09 paper/sections/Conclusion.tex
@@
 
-Huang: to write a couple of sentences for a summary.
+This paper presents a solution for semantically modular parsing. It not only
+enables parsers to evolve with the syntax together, but also allows them to be modularly type-checked and separately compiled.
 
-\noindent Haoyuan: to discuss future work. (some boilerplate can be avoided via metaprogramming.)
+Based on parser combinators, we identify the the algorithmic challenges of building modular parsers, and show that Packrat parsing is suitable as the underlying parsing technique.
+Our solution does not require advanced language features.
+We show that with standard OO techniques including inheritance and overriding, it is practicable to build modular parsers for OO ASTs.
+However, the extensibility issue of traditional OO ASTs motivates us to adopt Object Algebras for full extensibility and more useful features.
+
+Abstracting language features as reuseable components can be achieved based on our solution, which is useful for rapid language prototyping.
+Our solution does not rely on particular techniques.
+On the contrary, it is a general framework customizable by the user.
+The TAPL case study shows that we can obtain considerable code reuse (69\%) than a non-modular implementation, using our modular parsing approach and language feature abstraction.
+
+There are certainly some aspects can be improved.
+We observed that the glue code of composition appears to be boilerplate.
+Such an issue refers to family polymorphism, and we could solve it by language features or meta-programming techniques.
+Moreover, we can possibly adopt the Shy framework~\cite{Zhang2015} and algebra composition patterns~\cite{oliveira2013feature}, to improve the usage of Object Algebras.
+
+For future work, we will experiment more on how open recursion contributes to extensible parsing in functional languages, by making use of laziness.
+It is also interesting to see that parsing, to some extent, can be viewed as a special example of \textit{unfolds}.
+So it is worthwhile considering to generalize our approach under certain circumstances.

### 2017-01-14 paper/sections/Inheritance.tex

@@
 the two dimensions of extensibility. On the one hand it is naturally hard to
-add new operations to the AST, as usual in the OO style. On the hand
+add new operations to the AST, as usual in the OO style~\cite{wadler1998expression}. On the hand
 grammars are not modularly extensible, thus the addition of new
@@
 
-It is actually quite simple to let \inlinecode{pExpr} cover the newly extended case without modifying existing code.
-Method overriding is a standard feature which often comes with inheritance, and it allows us to explain inherited method again, such as \inlinecode{pExpr}.
-We can build the new parser which correctly parses \inlinecode{"1 + x"} as below.
+It is actually quite simple to let \inlinecode{pExpr} cover the newly
+extended case without modifying existing code. Method overriding is a
+standard feature which often comes with inheritance, and it allows us
+to redefined an inherited method, such as \inlinecode{pExpr}. We can
+build the new parser which correctly parses \inlinecode{"1 + x"} 
+through overriding:
 
@@
 \subsection{Multiple Extensions and Independent Extensibility}
-A nice feature of Scala is its support for a form of multiple inheritance.
-Scala has a linearized-style multiple inheritance for traits~\cite{odersky2004overview}. This can be very helpful when composing several languages, which
-may have been independently developed. Suppose now
-we want to compose the parsers for expressions from pre-defined
-languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
-alternative. The new parser can be built by inheriting both parsers
-at the same time:
+A nice feature of Scala is its support for a form of multiple
+inheritance.  Scala has a linearized-style multiple inheritance for
+traits~\cite{odersky2004overview}. This can be very helpful when
+composing several languages, and to achieve independent
+extensibility~\cite{odersky2005independently}. 
+Suppose now we want to compose the parsers for expressions
+from pre-defined languages \inlinecode{LanguageA} and
+\inlinecode{LanguageB} using alternative. The new parser can be built
+by inheriting both parsers at the same time:
 

### 2017-01-14 paper/sections/Conclusion.tex
Revised 4.4

@@
 On the contrary, it is a general framework customizable by the user.
-The TAPL case study shows that we can obtain considerable code reuse (69\%) than a non-modular implementation, using our modular parsing approach and language feature abstraction.
+The TAPL case study shows that we can obtain considerable code reuse (69\% shorter) comparing with a non-modular implementation, using our modular parsing approach and language feature abstraction.
 

### 2017-01-14 paper/sections/ObjectAlgebras.tex
Revised 4.4

@@
 
-\subsection{Composition of Algebras}\label{subsec:more-oa}
-\huang{to revise}
-By adopting Object Algebras, we also inherit some advanced features from the pattern.
-Specifically, in the original paper~\cite{Oliveira:2012}, there are two kinds of Object Algebras: \textit{queries} and \textit{transformations}.
+\subsection{Discussion of Object Algebras}\label{subsec:more-oa}
+
+In Section~\ref{subsec:objectalgebras}, \lstinline{makeExp} represents the AST of expression \inlinecode{1 + x} as a method.
+Parsers with Object Algebras generate such ``implicit objects'', which are actually functions of type \lstinline{Alg[E] => } \lstinline{E} for generic \lstinline{E}.
+Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
+Especially, in the design of a compiler there will be a lot of desugarings (or AST transformations).
+
+However, we argue that this is not an issue, because functions can be used
+in a very clever way. Continuing with the previous example, suppose we want to transform an expression of \inlinecode{VarExprAlg} to an expression of \inlinecode{ExprAlg}, by
+replacing some variables with values based on a variable environment. A \textit{transformation} algebra can be implemented as follows:
+
+\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+Here the \inlinecode{VarExprAlg} expression \inlinecode{1 + x} is changed to \inlinecode{ExprAlg} expression \inlinecode{1 + 2}, before applying the pretty-printing algebra. The abstract field
+\inlinecode{alg} in \inlinecode{Refactor} could be any algebra simply for delegation.
+With an algebra instance of \inlinecode{alg}, \inlinecode{Refactor} is able to transform a \inlinecode{VarExprAlg} expression before producing the result.
+Furthermore, transformations can be composed in a linear pipeline-style beforehand~\cite{Zhang2015}.
+
+Another useful technique is the composition of \textit{query} algebras.
 A query traverses the abstract syntax tree, collects information and returns the result.
-The aforementioned \lstinline{Print} and \lstinline{Eval} are traditional queries.
+The aforementioned \lstinline{Print} and \lstinline{Eval} algebras are traditional queries.
 But as we discussed before, an algebra is first fed to the parser, then \lstinline{parse} is invoked to do the parsing.
 It is tedious to fed \lstinline{Print} and \lstinline{Eval} separately, as the same input is parsed twice.
-Instead, we can firstly compose two queries into one, then feed it to the parser.
-For space reasons we omit the code here.
-
-Another interesting observation, if we look back at the end of Section~\ref{subsec:objectalgebras}, is that \lstinline{makeExp} represents an ``object'' of the expression AST.
-Or rather, what OA parsers generate are ``implicit objects'', which are actually functions of type \lstinline{Alg[E]} \lstinline{=>} \lstinline{E} for generic \lstinline{E}.
-Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
-Especially, in the design of a compiler
-there will be a lot of desugarings (or transformations). However, we argue that this is not an issue, because functions can be used
-in a very clever way. Continuing from the old example, if one wants to transform a \inlinecode{VarExprAlg} expression to an \inlinecode{ExprAlg} expression by
-replacing some variables with values based on a variable environment, a transformation algebra can be implemented as follows:\haoyuan{mention in discussion that shy deals with boilerplate.}
-
-\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
-Here the \inlinecode{VarExprAlg} expression \inlinecode{"1 + x"} is changed to \inlinecode{ExprAlg} expression \inlinecode{"1 + 2"}, before the pretty-printing algebra is applied. The abstract method
-\inlinecode{alg} in \inlinecode{Refactor} can actually be any algebra simply for delegation. By passing a concrete implementation
-to \inlinecode{alg}, \inlinecode{Refactor} will be able to transform a \inlinecode{VarExprAlg} expression before we obtain the result.
-Furthermore, transformations can be composed in a linear pipeline-style, before a query algebra is fed ~\cite{Zhang2015}.
+Instead, we can firstly compose two queries into one~\cite{Oliveira:2012,oliveira2013feature}, before applying it for parsing.

### 2017-01-14 paper/sections/PackratParsing.tex
Revised 4.4

@@
 \noindent However, this results in the worst case exponential time
-complexity in Parsec, because it does not have proive optimizations
+complexity in Parsec, because it does not provide optimizations
 for backtracking.

### 2017-01-14 paper/sections/Introduction.tex
Small space tweak
Note for reviewers

@@
 
-\item {{\bf A technique for Modular Parsing:}} We present a technique
+\item {{\bf A Technique for Modular Parsing:}} We present a technique
 that allows the development of semantically modular parsers.
@@
   at:
-  \url{http://}
+  \url{http://}\footnote{{\bf Note to Reviewers:} Please find the url in the
+  suplementary material for the submission.}
 

### 2017-01-14 paper/sections/CaseStudy.tex
Fix refs, spell check
Down to 25 pages

@@
       \hline
+      \multicolumn{7}{c}{}
     \end{tabular}
-    \caption{Comparison of SLOC and execution time}
+    \caption{Comparison of SLOC and execution time.}
     \label{tab:comparison}
@@
 The overall result is that 69.2\% of code is reduced using our
-library, and our implementation is 42.7\% slower.
+approach, and our implementation is 42.7\% slower.
 
@@
       \hline
+        \multicolumn{8}{c}{}
     \end{tabular}
-    \caption{Execution time of four versions}
+    \caption{Execution time of four implementations.}
     \label{tab:ext-comparison}
@@
 Table \ref{tab:ext-comparison} shows part of the comparison results of these four versions,
-full table is available online []. It suggests that the difference of running time between
+full table is available online. It suggests that the difference of running time between
 using object algebra and class is little, roughly 1\%.

### 2017-01-14 paper/sections/Discussion.tex
Down to 25 pages

@@
 
-\begin{figure}
-\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
-\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
-\label{fig:multi}
-\end{figure}
-
 Figure~\ref{fig:multi} illustrates the corresponding Scala code
@@
 with \inlinecode{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
+
+\begin{figure}[ht]
+\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
+\label{fig:multi}
+\end{figure}
 
@@
 
+\begin{comment}
 It is natural to keep overriding existing parsers, whereas another potential use of this pattern is to reuse old versions of a parser. For instance,
@@
 
-\huang{i finally managed to make it work, with two overriding.
-i think this case may not be very common and the solution is kind of unly.
-not sure whether to keep this case or not}\haoyuan{I don't know if this works or makes sense.}
-
 \lstinputlisting[linerange=94-102]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
+\end{comment}
 
@@
 
+\begin{figure}[t]
 \lstinputlisting[linerange=160-170]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\caption{Reuse two language components \lstinline{VarExpr} and \lstinline{TypedLam}.}\label{fig:compose-components}
+\end{figure}
+
 Such composition shows perfect modularity. The only

### 2017-01-14 paper/sections/ObjectAlgebras.tex
Down to 25 pages

@@
 
-\lstinputlisting[linerange=23-33]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
+\lstinputlisting[linerange=23-30]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
@@
 Now let's try this pattern with the variables extension:
-\lstinputlisting[linerange=37-44]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+
+\lstinputlisting[linerange=34-39]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
 
@@
 
-\lstinputlisting[linerange=48-51]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
+\lstinputlisting[linerange=43-46]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
 

### 2017-01-14 paper/sections/PackratParsing.tex
Down to 25 pages

@@
 expression $1-2-3$ is parsed as $(1-2)-3$ in the left-recursive
-grammar, but after rewrite the result is $1((-2)-3))$\bruno{??This is
-  not a valid expression}. The parse tree
+grammar, but after rewrite the information of left associativity is lost. The parse tree
 must be transformed to recover the correct syntactic structure.

### 2017-01-14 paper/sections/Inheritance.tex
Section 3

@@
 the two dimensions of extensibility. On the one hand it is naturally hard to
-add new operations to the AST, as usual in the OO style~\cite{wadler1998expression}. On the hand
+add new operations to the AST, as usual in the OO
+style~\cite{wadler1998expression}. On the other hand
 grammars are not modularly extensible, thus the addition of new
@@
 style mechanisms such as subtyping and inheritance, as well as,
-Packrat parsing combinators.
+Packrat parsing combinators, and enables a better integration bettwen 
+parsing and OO ASTs.
 
@@
 A nice feature of Scala is its support for a form of multiple
-inheritance.  Scala has a linearized-style multiple inheritance for
+inheritance.  Scala has a linearized-style mixin inheritance for
 traits~\cite{odersky2004overview}. This can be very helpful when
@@
 
-Then the new parser, which contains both integer and boolean expressions, is able to parse hybrid expressions such as \inlinecode{"if true then 1 + 2 else 3 + 4"}.
+Then the new parser, which contains both integer and boolean
+expressions, is able to parse hybrid expressions such as
+\inlinecode{"if true then 1 + 2 else 3 + 4"}.
 
-As we demonstrated, inheritance with method overriding is the key technique to obtain semantic modularity.
-It enables type-safe code reuse and separate compilation for parsing OO style ASTs.
+\paragraph{Conflicts and/or ambiguity}
+
+As demonstrated, inheritance with method overriding is the key
+technique to obtain semantic modularity.  It enables type-safe code
+reuse and separate compilation for parsing OO style ASTs.

### 2017-01-14 paper/sections/Inheritance.tex
References

@@
 it is perhaps not a surprise that parser generators that also generate
-AST code (such as ANTLR~\cite{parr2013definitive}), normally generate code based on
+AST code (such as ANTLR~\cite{antlr1995}), normally generate code based on
 visitors. Clearly this seems a better choice since generating code
@@
 style mechanisms such as subtyping and inheritance, as well as,
-Packrat parsing combinators, and enables a better integration bettwen 
+Packrat parsing combinators, and enables a better integration between
 parsing and OO ASTs.

### 2017-01-14 paper/sections/ObjectAlgebras.tex
Section 4

@@
 
-The Expression Problem~\cite{wadler1998expression} illustrates the difficulty of extending data structures or ASTs in two dimensions.
-For traditional OO ASTs, adding new language constructs is easy, but adding new operations is hard.
-In the last section, we have a simple language containing literals, additions and variables.
-The corresponding ASTs are built using classes with a pretty-printing operation.
-Now suppose we want to collect free variables over the ASTs.
-By traditional OO approach, one attempt would be extending \inlinecode{Expr} with the new operation to obtain a new abstract type for ASTs:
+The Expression Problem~\cite{wadler1998expression} illustrates the
+difficulty of extending data structures or ASTs in two dimensions. For
+traditional OO ASTs, adding new language constructs is easy, but
+adding new operations is hard. In the last section, we have a simple
+language containing literals, additions and variables. The
+corresponding ASTs are built using classes with a pretty-printing
+operation. Now suppose we want to collect free variables over the
+ASTs. Using a traditional OO approach, one attempt would be extending
+\inlinecode{Expr} with the new operation to obtain a new abstract type
+for ASTs:
 
@@
 
-But then programmers have to define a new class for each construct by extending the old class with \lstinline{FVarsExpr}, and implementing the corresponding \lstinline{fVars} method:
+\noindent Then all classes represented language constructs could be 
+extended to implement the operation. A first well-known problem is that such
+approach is problematic in terms of type-safety (but see recent work 
+by Wang and Oliveira~\cite{wang2016expression}, which show a technique that is type-safe
+in many cases). More importantly, a second problem is that
+\begin{comment}
+But then programmers have to define a new class for each construct by
+extending the old class with \lstinline{FVarsExpr}, and implementing
+the corresponding \lstinline{fVars} method:
 
@@
 
-Besides introducing boilerplate code, more classes have to be defined along with future extensions.
-They pollutes the namespace heavily.
-More importantly, the \lstinline{VarExprParser} we defined is no longer suitable for parsing the new ASTs!
-All the methods in the previous parser has type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
+Besides introducing boilerplate code, more classes have to be defined
+along with future extensions. They pollutes the namespace heavily.
+More importantly, 
+\end{comment}
+is that even if that approach would work, the parsing code would not
+be reusable.
+\lstinline{VarExprParser} is no longer suitable for parsing the new ASTs!
+All the methods in the previous parser has type \inlinecode{Parser[Expr]}, which indicates the result have type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
 
-Therefore, in order to generate new AST structures, we have to substitute \lstinline{FVarsExpr}, \lstinline{FVarsLit}, \lstinline{FVarsAdd} and \lstinline{FVarsVar} for the old \lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var} correspondingly.
-Again, to avoid modifying existing code, we have to sacrifice code reuse.
-In fact, there is a trade-off between separate compilation and type-safe code reuse.
-For semantic modularity, we need a different approach for building ASTs.
+Therefore, in order to generate new AST structures, we have to
+substitute \lstinline{FVarsExpr}, \lstinline{FVarsLit},
+\lstinline{FVarsAdd} and \lstinline{FVarsVar} for the old
+\lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var}
+correspondingly. Thus we have to either modify existing code or 
+use casts. For semantic modularity, we need a different approach for building ASTs.
 
@@
 
-Fortunately, Object Algebras~\cite{Oliveira:2012} enables us to solve this problem. As a nice solution, it separates data variants
-and operations, and offers high flexibility in the choice of operations to be performed over ASTs. Before adopting it for parsing, the next section introduces the background of Object Algebras.
+Fortunately, Object Algebras~\cite{Oliveira:2012} enable us to solve
+this problem. Object Algebras separate data variants and
+operations, and offer high flexibility in the choice of operations to
+be performed over ASTs. The next section provides a brief introduction 
+to Object Algebras.
 
 \subsection{Object Algebras}\label{subsec:objectalgebras}
-Object Algebras captures a design pattern to address the Expression Problem, achieving two dimensions of extensibility (language constructs and operations) in a modular and type-safe way.
-The definition of data structures is separated from their behaviours, and future extensions on both dimensions no longer require existing code to be modified, supporting separate compilation.
+
+Object Algebras captures a design pattern that address the Expression
+Problem, achieving two dimensions of extensibility (language
+constructs and operations) in a modular and type-safe way. The
+definition of data structures is separated from their behaviours, and
+future extensions on both dimensions no longer require existing code
+to be modified, supporting separate compilation.
 
@@
 
-\subsection{Discussion of Object Algebras}\label{subsec:more-oa}
+\begin{comment}
+\subsection{ASTs and Object Algebras}\label{subsec:more-oa}
 
-In Section~\ref{subsec:objectalgebras}, \lstinline{makeExp} represents the AST of expression \inlinecode{1 + x} as a method.
-Parsers with Object Algebras generate such ``implicit objects'', which are actually functions of type \lstinline{Alg[E] => } \lstinline{E} for generic \lstinline{E}.
-Comparing with traditional objects, one may have concern that if a parser produces function structures, it could be hard to further process them.
-Especially, in the design of a compiler there will be a lot of desugarings (or AST transformations).
-
-However, we argue that this is not an issue, because functions can be used
-in a very clever way. Continuing with the previous example, suppose we want to transform an expression of \inlinecode{VarExprAlg} to an expression of \inlinecode{ExprAlg}, by
-replacing some variables with values based on a variable environment. A \textit{transformation} algebra can be implemented as follows:
+In Section~\ref{subsec:objectalgebras}, \lstinline{makeExp} represents
+the AST of expression \inlinecode{1 + x} as a method. Parsers with
+Object Algebras generate such ``implicit objects'', which are actually
+functions of type \lstinline{Alg[E] => } \lstinline{E} for generic
+\lstinline{E}. Comparing with traditional objects, one may have worry
+that if a parser produces function structures, it could be hard to
+further process them. Especially, in the design of a compiler there
+will be a lot of desugarings (or AST transformations). As previous
+work on Object Alegbras has shown~\cite{Gouseti2014,Zhang2015}, AST
+transformations can be implemented using \emph{transformation
+  algebras}. Continuing with the previous example, suppose we want to
+transform an expression of \inlinecode{VarExprAlg} to an expression of
+\inlinecode{ExprAlg}, by replacing some variables with values based on
+a variable environment. A \textit{transformation} algebra is
+implemented as follows:
 
@@
 Instead, we can firstly compose two queries into one~\cite{Oliveira:2012,oliveira2013feature}, before applying it for parsing.
+\end{comment}

### 2017-01-14 paper/sections/Introduction.tex
More space tweak
S1.

@@
 Examples include many tools for developing Feature-Oriented Software-Product
-Lines (SPLs)~\cite{Kastner11road}, some Language Workbenches~\cite{}, or extensible parser
-generators~\cite{}.  Most syntactic approaches employ textual
-composition techniques such as \emph{superimposition}~\cite{} to
+Lines (SPLs)~\cite{Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
+generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
+composition techniques such as \emph{superimposition}~\cite{Wagelaar2010} to
 enable the development modular program features. Such
@@
   for implementing modular parsers built on top of an
-  existing Packrat~\cite{Ford2002} parsing library for Scala~\cite{}. The Packrat parsing library
+  existing Packrat~\cite{Ford2002} parsing library for Scala~\cite{odersky2004overview}. The Packrat parsing library
   directly supports left-recursion, and a \emph{longest-match
@@
   provide some form of multiple inheritance,
-  such as traits/mixins~\cite{scalaoverview}\bruno{cite work on traits (Ducasse)
-  and mixins (Bracha)}, are very handy for this
+  such as traits/mixins~\cite{Bracha1990,Scharli2003}, are very handy for this
   purpose. Essentially, traits/mixins can act as modules for the parsing code
@@
 interpreters, and the total parsing code is 69\% shorter than an
-existing code base using non-modular parsing code~\cite{}.
+existing code base using non-modular parsing code\footnote{https://github.com/ilya-klyuchnikov/tapl-scala/}.
 
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep,nolistsep]
 
@@
 other languages that support some form of multiple inheritance
-(including Java 8 with default methods~\cite{}) could in principle be
+(including Java 8 with default methods~\cite{Goetz2012}) could in principle be
 used. Our choice for Scala had to do with the concise and elegant syntax

### 2017-01-14 paper/sections/Discussion.tex
Section 5

@@
 
-The use of inheritance-based approach and Object Algebras enables us to build modular parsers, which are able to evolve with syntax together.
-This section explores more interesting features, including parsing multi-sorted syntax, overriding existing parsing rules, language components for abstracting language features, and alternative techniques under the whole framework.
+The use of inheritance-based approach and Object Algebras enables us
+to build modular parsers, which are able to evolve with together with
+syntax. This section explores more interesting features, including
+parsing multi-sorted syntax, overriding existing parsing rules,
+language components for abstracting language features, and alternative
+techniques under the whole framework.
 
-\subsection{Parsing Multi-sorted Syntax}\label{subsec:differentsyntax}
+\subsection{Parsing Multi-Sorted Syntax}\label{subsec:differentsyntax}
 
@@
 
-Another advantage of using Object Algebras is that it is easy to
+Using Object Algebras is that it is easy to
 model multi-sorted languages.
@@
 
-As many syntactically extensible parsers, our approach also supports changing part of existing parsers, but in a type-safe way.
-We can easily update existing parsing rules with new implementations, or eliminate them in extended parsers.
+As many syntactically extensible parsers, our approach also supports
+changing part of existing parsers, but in a type-safe way. We can
+easily update existing parsing rules with new implementations, or
+eliminate them in extended parsers. This can be useful in many
+situations, including when conflicts or ambiguities arize when
+composing languages.
 As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below.
@@
 
-Then we want to replace the untyped lambda abstractions by typed lambdas.
-With inheritance and method overriding, it is easy to only change the implementation of \lstinline{pLam} in the extended parser.
-The new \inlinecode{pLam} parses a colon and a type in addition, between the identifier and the dot.
-Due to dynamic dispatch, our new explanation of lambdas will make a difference without affecting the other parts of the parser.
+Then we want to replace the untyped lambda abstractions by typed
+lambdas. With inheritance and method overriding, it is easy to only
+change the implementation of \lstinline{pLam} in the extended parser.
+The new \inlinecode{pLam} parses a colon and a type in addition,
+between the identifier and the dot. Due to dynamic dispatch, our new
+implementation of lambdas will be different without affecting the other parts of the parser.
 
@@
 
-\subsection{Language Component}\label{subsec:language-component}
+\subsection{Language Components}\label{subsec:language-component}
 
-Modular parsing not only enables us to build a corresponding parser which evolves with the language together, but also allows us to abstract language features as reusable, independent components.
-Generally, a language feature includes related abstract syntax, methods to \textit{build} the syntax (parsing), and methods to \textit{process} the syntax (evaluation, pretty-printing, etc.).
-From this perspective, not only one language, but many languages can be developed in a modular way, with common language features reused.
+Modular parsing not only enables us to build a corresponding parser
+which evolves with the language together, but also allows us to
+abstract language features as reusable, independent components.
+Generally, a language feature includes related abstract syntax,
+methods to \textit{build} the syntax (parsing), and methods to
+\textit{process} the syntax (evaluation, pretty-printing, etc.). From
+this perspective, not only one language, but many languages can be
+developed in a modular way, with common language features reused.
 
-Instead of design and build a language from scratch, we can easily add a new feature by reusing the corresponding language component.
-For example, if a language is composed with a component of boolean expressions, including if-then-else, it immediately knows how to parse, traverse, and pretty-print if-then-else structure!
-It could be very useful for rapid developing DSLs.
+Instead of designing and building a language from scratch, we can
+easily add a new feature by reusing the corresponding language
+component. For example, if a language is composed from a component of
+boolean expressions, including if-then-else, it immediately knows how
+to parse, traverse, and pretty-print the if-then-else structure.
+Grouping language features in this way can
+be very useful for rapid development of DSLs.
 
@@
 
-Such composition shows perfect modularity. The only
-drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty-printers of
-\lstinline{VarExpr} and \lstinline{TypedLam} respectively.
-Such a pattern refers to \textit{family polymorphism}~\cite{ernst01FP} which is unfortunately not supported in Scala.
-Nonetheless, one can avoid such boilerplate by using metaprogramming techniques.
+The only drawback is that the glue code of composition appears to be
+boilerplate. As shown above, we are combining ASTs, parsers and
+pretty-printers of \lstinline{VarExpr} and \lstinline{TypedLam}
+respectively. Such a pattern refers to \textit{family
+  polymorphism}~\cite{ernst01FP} which is unfortunately not fully supported
+in Scala, since nested classes/traits have to be manually composed. 
+%Nonetheless, one can avoid such boilerplate by using metaprogramming techniques.
 
@@
 Under our modular parsing framework, we use Packrat parsing as the underlying parsing technique, OO inheritance with method overriding for composing and extending parsers, and Object Algebras for parsing extensible ASTs.
-However, our framework is more general and thus more powerful, because those aspects are orthogonal to each other, and hence can have their alternatives.
+However, our framework is more general and thus more powerful, because those aspects are orthogonal to each other, and hence can have alternatives.
 
@@
 
-Our approach does not depend on a particular parser combinator library or parsing algorithm.
-We demonstrated that Packrat parsing has some advantages and is suitable in a modular setting.
-However, under certain circumstances or in other programming languages, it may not be the best choice.
-One can use other parser combinator libraries as long as it meets his requirements.
+Our approach does not depend on a particular parser combinator library
+or parsing algorithm. We demonstrated that Packrat parsing has some
+advantages and is suitable in a modular setting. However, under
+certain circumstances or in other programming languages, it may not be
+the best choice. One can use other parser combinator libraries as long
+as th guidelines proposed in Section~\ref{sec:packrat} are met.
 
@@
 
-We use traits to model parsers, OO inheritance to compose them, and method overriding for their extensibility.
-These language features may not be supported in other programming languages, especially in functional languages such as Haskell.
-Nevertheless, \textit{open recursion} could be used as an alternative.
-An explicit ``self-reference'' parameter is able to explain the recursive calls dynamically, by the real argument passed at runtime.
+  We use traits to model parsers, OO inheritance to compose them, and
+  method overriding for their extensibility. These language features
+  may not be supported in other programming languages, especially in
+  functional languages such as Haskell. Nevertheless, \textit{open
+    recursion}~\cite{CookThesis} could be used as an alternative. An explicit
+  ``self-reference'' parameter is able to explain the recursive calls
+  dynamically, by the real argument passed at runtime.
 
@@
 
-Besides Object Algebras, many other techniques including \textit{Data types à la carte} (DTC)~\cite{swierstra2008data} and \textit{Modular reifiable matching} (MRM)~\cite{Oliveira2015} also support extensible data structures.
-These alternatives can also be adopted to build ASTs and corresponding parsers.
+Besides Object Algebras, many other techniques including \textit{Data
+  types à la carte} (DTC)~\cite{swierstra2008data} ot the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
+These alternatives could also be adopted to build ASTs and corresponding parsers.
 
@@
 
-As we demonstrated, the whole modular parsing framework itself is modular, which means it can be customized easily.
-Actually, we conducted some previous experiments of modular paring in Haskell, based on open recursion, MRM and Parsec.
+As we demonstrated, the whole modular parsing framework itself is
+modular, which means it can be customized easily. 
+%We conducted some previous experiments of modular parsing in Haskell, based on open recursion, MRM and Parsec.
 However, we suffered from issues of Parsec, as illustrated in Section~\ref{subsec:challenges}.

### 2017-01-14 paper/sections/RelatedWork.tex
Added the last paragraph in S7.

@@
 separate compilation.
-\textit{Parse table composition}~\cite{Bravenboer2009}
+\textit{Parse table composition}~\cite{bravenboer2008parse}
 is an approach where grammars are compiled to
@@
 
-Another interesting observation, as we have mentioned
-in Section~\ref{subsec:more-oa}, is that transformation algebras are
-frequently used in parsing, as concrete syntax may not always coincide
-with its abstract syntax, for instance, the bruijn indices in lambda
-calculus, as well as other syntactic sugars. In the Shy
-framework~\cite{Zhang2015}, the pattern of transformations is captured
-into a template for Java annotation processing, and most of
-boilerplate code is automatically generated, hence users only need to
-override a few interesting cases, conveniently. It is potentially
-useful in our case. The idea of overriding existing cases was also
-proposed in MRM. Our approach allows existing parsing code to be
-overridden during inheritance.
+Another interesting observation from Section~\ref{4}, is that parsers with
+Object Algebras generate ``implicit objects'', which are actually functions of
+type \lstinline{Alg[E]} \lstinline{=>} \lstinline{E} for generic \lstinline{E}.
+Patterns of operations on such functions are captured by the original paper~\cite{Oliveira:2012}, called
+\textit{queries} and \textit{transformations}, together with the composition of algebras.
+They are quite useful, for instance, we have two operations (like the aforementioned \lstinline{Print} and \lstinline{Eval}) to be fed to the parser.
+If we feed them separately, it is tedious to parse the same input twice.
+Instead we can compose them in advance before the feeding.
+On the other hand, one would think this pattern is too limited to use against objects that traditional parsers generate.
+For example, there are a lot of transformations (or desugarings) on intermediate ASTs in the design of a compiler.
+Transformation algebras are helpful in that case.
+They can even be composed in a linear pipeline-style before a query algebra is delegated.
+The Shy framework~\cite{Zhang2015} has captured those patterns, and it generates templates for them, so that users only need to override a few interesting cases, conveniently.
+The idea of overriding existing cases was also
+proposed in MRM~\cite{Oliveira2015}. They are potentially useful to our approach. Regarding overriding, our approach allows existing parsing code to be overridden during inheritance.

### 2017-01-14 paper/sections/Inheritance.tex
A paragraph in 3.4 for ambiguity

@@
 
-\paragraph{Conflicts and/or ambiguity}
+\paragraph{Conflicts and/or ambiguity} As we demonstrated before, the longest match alternative combinator is useful when composing parsers.
+However, ambiguity grammars still cause subtle issues in modular parsing.
+For example, if \inlinecode{parserA} and \inlinecode{parserB} accept a same string but generate different results, then the result of \inlinecode{parserA ||| parserB} is really depends on the implementation of the alternative combinator, because the results can not be prioritized by length.
+Since the composition is handled explicitly by user, conflicts could be resolved by overriding ambiguous parsing code, using the pattern discussed in Section~\ref{subsec:overriding-rules}.
 

### 2017-01-14 paper/sections/PackratParsing.tex
Section 3
A paragraph in 3.4 for ambiguity

@@
 
-Consider we extended the language by adding subtractions to the grammar.
+Consider extending the language by adding subtractions to the grammar.
 If we still write its parser by directly following the grammar structure, the new left-recursive case will get the parser into an infinite loop.
@@
 recursively. Finally \lstinline{pExpr} composes \lstinline{pLit} and \lstinline{pAdd} using the longest match alternative combinator \inlinecode{|||}.
-Table~\ref{tab:packrat} gives a closer look at the parser combinators used in our work.
+Table~\ref{tab:packrat} shows common parser combinators from the library.
 
@@
 \end{tabular}
-\caption{Combinators used in our work, from the Scala standard parser combinator library.}\label{tab:packrat}
+\caption{Common combinators from the Scala standard parser combinator library.}\label{tab:packrat}
 \end{table}

### 2017-01-14 paper/sections/Introduction.tex
Intro

@@
 Examples include many tools for developing Feature-Oriented Software-Product
-Lines (SPLs)~\cite{Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
+Lines (SPLs)~\cite{AK:JOT09,Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
 generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
-composition techniques such as \emph{superimposition}~\cite{Wagelaar2010} to
+composition techniques such as \emph{superimposition}~\cite{AK:JOT09} to
 enable the development modular program features. Such
@@
 well as various techniques for solving the Expression
-Problem~\cite{Oliveira:2012,wang2016expression,torgersen2004expression,odersky2005independently}.
+Problem~\cite{torgersen2004expression,odersky2005independently,Oliveira:2012,wang2016expression}.
 Semantic modularization techniques are less widely used in practice

### 2017-01-14 paper/sections/RelatedWork.tex
Small fix.

@@
 
-Another interesting observation from Section~\ref{4}, is that parsers with
+Another interesting observation from Section~\ref{sec:algebrasandparsing}, is that parsers with
 Object Algebras generate ``implicit objects'', which are actually functions of

### 2017-01-14 paper/sections/RelatedWork.tex
removed paragraph in RW

--- 2017-01-14 paper/sections/RelatedWork.tex
@@
 
-Another interesting observation from Section~\ref{4}, is that parsers with
+\begin{comment}
+Another interesting observation from Section~\ref{sec:algebrasandparsing}, is that parsers with
 Object Algebras generate ``implicit objects'', which are actually functions of
@@
 proposed in MRM~\cite{Oliveira2015}. They are potentially useful to our approach. Regarding overriding, our approach allows existing parsing code to be overridden during inheritance.
+\end{comment}

### 2017-01-14 paper/sections/Discussion.tex
A paragraph in 3.4 for ambiguity
Fix

@@
 The use of inheritance-based approach and Object Algebras enables us
-to build modular parsers, which are able to evolve with together with
-syntax. This section explores more interesting features, including
+to build modular parsers, which are able to evolve with
+syntax together. This section explores more interesting features, including
 parsing multi-sorted syntax, overriding existing parsing rules,
@@
 
-Using Object Algebras is that it is easy to
+Using Object Algebras, it is easy to
 model multi-sorted languages.
 If the syntax contains multiple sorts, we can distinguish them by different type parameters.
-As an example, we extend the language of literals, additions and variables by introducing types as a new kind of type argument, and we add lambda abstractions to expressions.
+As an example, we extend the language of literals, additions and variables by introducing types, including primitive type \inlinecode{int} and function (arrow) type, and we add lambda abstractions to expressions.
 Now the language has two sorts in the syntax, types and expressions:\\
@@
 
-\subsection{Overriding Existing Rules}
+\subsection{Overriding Existing Rules}\label{subsec:overriding-rules}
 
 As many syntactically extensible parsers, our approach also supports
-changing part of existing parsers, but in a type-safe way. We can
+modifying part of existing parsers, but in a type-safe way. We can
 easily update existing parsing rules with new implementations, or
@@
 
-As shown below, those two language components can be merged into one to obtain the language we want. Furthermore, the new language is still a modular
+As shown in Figure~\ref{fig:compose-components}, those two language components can be merged into one to obtain the language we want. Furthermore, the new language is still a modular
 component ready for future composition. In that case modularity is realized over higher-order hierarchies.
@@
 Besides Object Algebras, many other techniques including \textit{Data
-  types à la carte} (DTC)~\cite{swierstra2008data} ot the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
+  types à la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
 These alternatives could also be adopted to build ASTs and corresponding parsers.

### 2017-01-14 paper/sections/Inheritance.tex
small typo
Typos in S1234.

@@
 language constructs is problematic. Therefore,
-it is perhaps not a surprise that parser generators that also generate
-AST code (such as ANTLR~\cite{antlr1995}), normally generate code based on
+it is perhaps not a surprise that parser generators also generate
+AST code (such as ANTLR~\cite{antlr1995}), normally the generated code is based on
 visitors. Clearly this seems a better choice since generating code
@@
 standard feature which often comes with inheritance, and it allows us
-to redefined an inherited method, such as \inlinecode{pExpr}. We can
+to redefine an inherited method, such as \inlinecode{pExpr}. We can
 build the new parser which correctly parses \inlinecode{"1 + x"}
@@
 The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}.
-Under multiple inheritance, it can be used to distinguish the method of the same name.
+Under multiple inheritance, it can be used to distinguish the methods of the same name.
 Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.
@@
 \paragraph{Conflicts and/or ambiguity} As we demonstrated before, the longest match alternative combinator is useful when composing parsers.
-However, ambiguity grammars still cause subtle issues in modular parsing.
-For example, if \inlinecode{parserA} and \inlinecode{parserB} accept a same string but generate different results, then the result of \inlinecode{parserA ||| parserB} is really depends on the implementation of the alternative combinator, because the results can not be prioritized by length.
+However, ambiguous grammars still cause subtle issues in modular parsing.
+For example, if \inlinecode{parserA} and \inlinecode{parserB} accept a same string but generate different results, then the result of \inlinecode{parserA ||| parserB} really depends on the implementation of the alternative combinator, because the results can not be prioritized by length.
 Since the composition is handled explicitly by user, conflicts could be resolved by overriding ambiguous parsing code, using the pattern discussed in Section~\ref{subsec:overriding-rules}.

### 2017-01-14 paper/sections/ObjectAlgebras.tex
Typos in S1234.

@@
 In this section, we
-show how to support both the forms of extensibility for ASTs (easy
+show how to support both forms of extensibility on ASTs (easy
 addition of language constructs, and easy addition of operations)
@@
 
-\noindent Then all classes represented language constructs could be 
+\noindent Then all classes representing language constructs could be 
 extended to implement the operation. A first well-known problem is that such
 approach is problematic in terms of type-safety (but see recent work 
-by Wang and Oliveira~\cite{wang2016expression}, which show a technique that is type-safe
+by Wang and Oliveira~\cite{wang2016expression}, which shows a technique that is type-safe
 in many cases). More importantly, a second problem is that
@@
 \lstinline{VarExprParser} is no longer suitable for parsing the new ASTs!
-All the methods in the previous parser has type \inlinecode{Parser[Expr]}, which indicates the result have type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
+All the methods in the previous parser has type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
 
@@
 
-Object Algebras captures a design pattern that address the Expression
+Object Algebras capture a design pattern that addresses the Expression
 Problem, achieving two dimensions of extensibility (language
@@
 the use of factory methods, instead of constructors.
-To pretty print the expression,
+To pretty-print the expression,
 we pass an instance of \inlinecode{VarExprPrint} to \inlinecode{makeExp}:
@@
 For example, we have demonstrated in Section~\ref{subsec:overriding} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
-However, trying to do a similar method override for \inlinecode{pExpr} would require a type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}, where the type of the \emph{extended} Object Algebra interface appears in \emph{contravariant} position.
+However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}, where the type of the \emph{extended} Object Algebra interface appears in \emph{contravariant} position.
 This difference prohibits method/field overriding.
@@
 The reason why we use an extra and seemingly redundant field, is due to a subtle issue caused by Scala language and its parser combinator library.
-There is a restriction of \inlinecode{super} keyword in Scala that \inlinecode{super} can only use fields defined by keyword \inlinecode{def}, can not use those defined by \inlinecode{val}, while the parser combinator library suggests using \inlinecode{val} to define parsers, especially for left-recursive ones.
+There is a restriction of \inlinecode{super} keyword in Scala that \inlinecode{super} can only use methods defined by keyword \inlinecode{def}, but cannot access fields defined by \inlinecode{val}, while the parser combinator library suggests using \inlinecode{val} to define parsers, especially for left-recursive ones.
 Our workaround is that we use different synonyms for \inlinecode{pE} in different traits, so that we can directly distinguish them by names without using \inlinecode{super}.
@@
 
-\noindent The type of the object algebra field \lstinline{alg} is first refined
+\noindent The type of the Object Algebra field \lstinline{alg} is first refined
 to \lstinline{VarExpAlg[E]}, to allow calling the additional factory method
@@
 call \lstinline{alg.varE}. The following code illustrates how to use
-the parser from a client perspective:
+the parser from a client's perspective:
 
@@
 In the client code shown above, we first need to choose a concrete
-object algebra to initialize the \lstinline{alg} field in \lstinline{VarExprOAParser}. Here we pick the
-pretty-printing algebra \lstinline{VarExprPrint}, but any other object
-algebra that implements \lstinline{VarExprAlg} would work.
+Object Algebra to initialize the \lstinline{alg} field in \lstinline{VarExprOAParser}. Here we pick the
+pretty-printing algebra \lstinline{VarExprPrint}, but any other Object
+Algebra that implements \lstinline{VarExprAlg} would work.
 With an instance of \lstinline{VarExprOAParser} in hand, we can call
@@
 
-Note that, similarly to the approach in Section~\ref{sec:inheritance} independent extensibility is also supported via multiple trait inheritance.
+Note that, similar to the approach in Section~\ref{sec:inheritance} independent extensibility is also supported via multiple trait inheritance.
 Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance} we omit that code here.

### 2017-01-14 paper/sections/PackratParsing.tex
Typos in S1234.

@@
 Firstly, they are naturally modular.
-The manner of using them is to write small parsers and use combinators to composed them together.
+The manner of using them is to write small parsers and use combinators to compose them together.
 The construction procedure is explicit and fully controlled by the programmer.
-Secondly, each parser combinator is represented by a piece of code, also are the parsers it takes.
+Secondly, each parser combinator is represented by a piece of code, and so are the parsers it takes.
 Thus in a statically typed programming language they can be statically type-checked.
@@
 \paragraph{Left-Recursion Elimination} The top-down, recursive descent parsing strategy adopted by those parser combinator libraries cannot support left-recursive grammars directly.
-The common solution is to rewrite the grammar into an equivalent but not left-recursive one, so called left-recursion elimination.
+The common solution is to rewrite the grammar into an equivalent but non-left-recursive one, called left-recursion elimination.
 
@@
 expression $1-2-3$ is parsed as $(1-2)-3$ in the left-recursive
-grammar, but after rewrite the information of left associativity is lost. The parse tree
+grammar, but after rewrite the information of left-associativity is lost. The parse tree
 must be transformed to recover the correct syntactic structure.
@@
 theory, supports both direct and indirect left-recursion~\cite{warth2008}.
-All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in for modular parsing.
+All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique for modular parsing.
 \begin{comment}

### 2017-01-14 paper/sections/RelatedWork.tex
Fix

@@
 \paragraph{Syntactically Extensible Parsing} Extensible parsing is
-used in many different areas. Extensible parser generators are a
+used in many different areas. Extensible parser generator is a
 mainstream area specially designed for modular syntax and
@@
 annotation processing to gather information about grammars in Java
-code. After all information is collected NOA merges all grammar
+code. After all information is collected, NOA merges all grammar
 components, produce an ANTLR~\cite{antlr1995} grammar, and generates
@@
 implementation was based on SDF~\cite{Heering1989} and
-Stratego~\cite{Visser2001}, the approach does not support separate
+Stratego~\cite{Visser2001}, it does not support separate
 compilation. Racket adopts a macro system for library-based language
@@
 
-\paragraph{Extensible Parsing Algorithms}
-Extensible parsing algorithms are another area that relates to
+\paragraph{Extensible Parsing Algorithm}
+Extensible parsing algorithm is another area that relates to
 separate compilation.

### 2017-01-14 paper/sections/Inheritance.tex
Typos in 1234.

@@
 For example, if \inlinecode{parserA} and \inlinecode{parserB} accept a same string but generate different results, then the result of \inlinecode{parserA ||| parserB} really depends on the implementation of the alternative combinator, because the results can not be prioritized by length.
-Since the composition is handled explicitly by user, conflicts could be resolved by overriding ambiguous parsing code, using the pattern discussed in Section~\ref{subsec:overriding-rules}.
+Since the composition is handled explicitly by a user, conflicts could be resolved by overriding ambiguous parsing code, using the pattern discussed in Section~\ref{subsec:overriding-rules}.
 

### 2017-01-14 paper/sections/PackratParsing.tex
Typos in 1234.

@@
 possible but unnecessarily complicated to analyse the grammar and
-rewrite it when doing extension. Anticipating that every non-terminal
+rewrite it when doing extensions. Anticipating that every non-terminal
 has left-recursive rules is helpful for extensibility but overkill,
 since it is inconvenient and introduces extra complexity for
-representation of grammar and implementation of parser.
+representation of a grammar and implementation of a parser.
 
@@
 \paragraph{Longest Match Composition} Another problematic issue
-in parser combinator libraries is the need for manually prioratizing/ordering
+in parser combinator libraries is the need for manually prioritizing/ordering
 alternatives in a grammar.
@@
 
-\noindent However, this results in the worst case exponential time
+\noindent However, this results in the worst-case exponential time
 complexity in Parsec, because it does not provide optimizations

### 2017-01-14 paper/sections/Discussion.tex
Typo.

@@
 eliminate them in extended parsers. This can be useful in many
-situations, including when conflicts or ambiguities arize when
+situations, including when conflicts or ambiguities arise when
 composing languages.
@@
 the best choice. One can use other parser combinator libraries as long
-as th guidelines proposed in Section~\ref{sec:packrat} are met.
+as the guidelines proposed in Section~\ref{sec:packrat} are met.
 

### 2017-03-21 paper/response.txt
Response draft.

We thank all reviewers for their thoughtful comments. Below we would like to talk about our contributions, and
some questions raised by reviewers.

As we discussed in the paper, almost all previous attempts on extensible/modular parsing stop at syntactical
level of modularity for various reasons. We propose a pattern that nicely achieves semantic modularity and
simplicity for parsing by using standard OO techniques. It’s a notable improvement, since compiling and
generating the parser again during extension every time is obviously a worse practice.

Our framework is indeed based on existing techniques, because a concrete algorithm is required for parsing.
In particular, we use Packrat (PEG) parsing and Object Algebras in our Scala implementation. However, we show
in our case study that they integrate well as expected; the parsing code is elegant and simplified to a great
extent, as boilerplate would explode along with a series of compositions. As discussed in Section 5.4 of the
paper, our approach can also be generalized thus not limited to any specific techniques. What we want to propose
is really a simple, powerful and practical framework, or a flexible platform, instead of a library, to achieve
easy and type-safe composition.

We are sorry that we missed the material of paper [1], as modular parsing and the pattern is not explained
in detail in the paper. Their code online seems to be different with ours as well. Based on parser
combinators, our pattern is more concise and easy to use.

Language components are an important contribution of our work, where we encapsulate a language feature with
three parts: abstract syntax, parser, and operations; everything is modular. With semantically modular parsing,
such components are fully type-safe. Just as SugarJ work and so on, we want languages to be small components,
that can be composed in a flexible way into modular languages again. They are reusable for further composition.
We don't need to rely on a base language to perform extensions as many extensible compilers do.

Rev B:
> However, object algebras solve a problem that is orthogonal to parsing...

We think Object Algebras are not a total orthogonal direction from modular parsing.
Since for parsing, the extension could happen in two cases:
-    When syntax changed, the parser should be extended easily to cooperate with the new syntax.
-    When ASTs changed (a new data variant/new operation added), the old parser should produce
     compatible ASTs without modification.
Regarding the second type of extension, Object Algebras or similar techniques are needed for
modular parsing. Our intuition of using Object Algebras, is not only it encodes multiple syntax
nicely, but also it adds a different dimension to obtain full extensibility of parsing,
supporting separate compilation.

Rev C:
> I would have liked to see it used for more realistic programming language grammars.

We agree that it would be better to show a more realistic language for demonstration. Thanks for your suggestion.

Rev C:
> - Unclear to me if indirect left recursion is supported, which would be needed
> for practical grammars.

Packrat parsing theoretically supports both direct and indirect left recursion [2],
although the implementation in Scala’s standard parser combinator library may be buggy
for some indirect left recursion cases.

Rev C:
> - Does not provide a sufficiently complete solution to language componentization,
> since object algebras are too restricted (only support bottom up computations).

Indeed Object Algebras only support bottom-up traversals, but we have known that some
other researchers have worked out some framework that transforms Object Algebra ASTs
to different ones, which also supports top-down traversals.

Rev E:
> Why is this "type-safety" of the parser so important, after all?  Yes, one may have
> bugs or missing cases in the parser itself, but I would not expect that to be a
> problem, because one should have tests that cover all the grammar anyway, so it is
> little likely that the type-safety of the parser itself would bring much practical benefits.

We think type-safe composition and separate compilation really matter, since it is not a good
idea to modify existing code; we are using OO and want to achieve better code reuse. As mentioned
before, we hope project code to be modular and easily composable, instead of being compiled and
generating code for every round of composition. It is better for debugging.

[1]: Rendel, Tillmann, Jonathan Immanuel Brachthäuser, and Klaus Ostermann. From object algebras to attribute grammars. OOPSLA 2014.
[2]: Alessandro Warth, James R Douglass, and Todd D Millstein. Packrat parsers can support left recursion. PEPM 2008.

### 2017-03-23 paper/response.txt
Response

@@
-We thank all reviewers for their thoughtful comments. Below we would like to talk about our contributions, and
-some questions raised by reviewers.
+We thank all reviewers for their thoughtful comments. Below we would
+like to talk about our contributions, and some questions raised by
+reviewers.
 
-As we discussed in the paper, almost all previous attempts on extensible/modular parsing stop at syntactical
-level of modularity for various reasons. We propose a pattern that nicely achieves semantic modularity and
-simplicity for parsing by using standard OO techniques. It’s a notable improvement, since compiling and
-generating the parser again during extension every time is obviously a worse practice.
+The essential criticism seems to be well summarized by reviewer D:
 
-Our framework is indeed based on existing techniques, because a concrete algorithm is required for parsing.
-In particular, we use Packrat (PEG) parsing and Object Algebras in our Scala implementation. However, we show
-in our case study that they integrate well as expected; the parsing code is elegant and simplified to a great
-extent, as boilerplate would explode along with a series of compositions. As discussed in Section 5.4 of the
-paper, our approach can also be generalized thus not limited to any specific techniques. What we want to propose
-is really a simple, powerful and practical framework, or a flexible platform, instead of a library, to achieve
-easy and type-safe composition.
+"Object algebras can be used to modularize tree traversals. Parser
+combinators combined with a Packrat algorithm and open recursion allow
+modular parser definitions. To people that know both techniques,
+object algebras and Packrat parsing, it also appears obvious to
+combine the two."
 
-We are sorry that we missed the material of paper [1], as modular parsing and the pattern is not explained
-in detail in the paper. Their code online seems to be different with ours as well. Based on parser
-combinators, our pattern is more concise and easy to use.
+We should also add, on top of the 2 techniques above, there's a third
+thing: *good knowledge of some of Scala's more advanced features.*
 
-Language components are an important contribution of our work, where we encapsulate a language feature with
-three parts: abstract syntax, parser, and operations; everything is modular. With semantically modular parsing,
-such components are fully type-safe. Just as SugarJ work and so on, we want languages to be small components,
-that can be composed in a flexible way into modular languages again. They are reusable for further composition.
-We don't need to rely on a base language to perform extensions as many extensible compilers do.
+Is it true that the paper is essentially combining 2 techniques + good
+use of Scala to achieve modular parsing, and that people that know
+about all of this could possibly recreate this work?
+
+Yes, it is true.
+
+Is that a bad thing?
+
+We are not sure about this. There are at least two arguments
+for why the fact that the technique actually ends up being quite
+simple is actually a good thing:
+
+1) *It shows that semantically modular parsing can be
+seriously considered for practical use*.
+
+Semantic modularization techniques are still far behind syntactic
+modularization techniques in terms of proving their
+worth/applicability in practice. Without question syntactic modularization
+approaches such as those currently used by Language Workbenches,
+are currently much more well-accepted in practice. This is in part
+because semantic modularization techniques are underdeveloped, and
+in part because there is an "established reputation" that semantic
+modularization techniques are just simply too hard to use in practice
+(and don't scale well).
+
+We believe the simplicity in our work/techniques makes a good
+argument for semantic modularization techniques.
+
+2) *We would dare to say that the set of people that knows
+Packrat parsing, Object Algebras and Scala well-enough to
+know that type-safe modular parsing can be easily done
+(before seeing our paper!) is less than 5*
+
+Honestly we didn't know this when we started our work (so we were
+not in that set). Although we knew Object Algebras and Scala
+well, we were not well-versed with Packrat parsing. We spent quite
+a bit of time playing with other parsing combinator libraries, and
+getting into trouble before we tried Packrat parsing and solved the
+issues we encountered.
+
+Finally, language components and our case study are important
+contributions of our work. Language components
+encapsulate a language feature with three parts:
+abstract syntax, parser, and operations; everything is modular. With
+semantically modular parsing, such components are fully
+type-safe. Just as SugarJ work and so on, we want languages to be
+small components, that can be composed in a flexible way into modular
+languages again. They are reusable for further composition.  We don't
+need to rely on a base language to perform extensions as many
+extensible compilers do. Our case study is also a good benchmark
+for other researchers and we believe a good contribution of this work.
 
@@
 
-We think Object Algebras are not a total orthogonal direction from modular parsing.
-Since for parsing, the extension could happen in two cases:
--    When syntax changed, the parser should be extended easily to cooperate with the new syntax.
--    When ASTs changed (a new data variant/new operation added), the old parser should produce
-     compatible ASTs without modification.
-Regarding the second type of extension, Object Algebras or similar techniques are needed for
-modular parsing. Our intuition of using Object Algebras, is not only it encodes multiple syntax
-nicely, but also it adds a different dimension to obtain full extensibility of parsing,
-supporting separate compilation.
+If the goal is to modularize languages (both parsing and semantics),
+and have libraries of language components, we don't see how the issues
+are orthogonal. The techniques must be combined.
 
@@
 
-We agree that it would be better to show a more realistic language for demonstration. Thanks for your suggestion.
+We agree that it would be better to show a more realistic language for
+demonstration. Thanks for your suggestion.
 
@@
 Packrat parsing theoretically supports both direct and indirect left recursion [2],
-although the implementation in Scala’s standard parser combinator library may be buggy
+although the implementation in Scala's standard parser combinator library may be buggy
 for some indirect left recursion cases.
@@
 
-Indeed Object Algebras only support bottom-up traversals, but we have known that some
-other researchers have worked out some framework that transforms Object Algebra ASTs
-to different ones, which also supports top-down traversals.
+Indeed Object Algebras only support bottom-up traversals, but there's
+ongoing work on variants of Object Algebras that remove this
+limitation. We think our approach will work just fine with such
+variants.
+
+Rev D:
+
+We are sorry that we missed the material of paper [1], as modular
+parsing and the pattern is not explained in detail in the paper. Their
+code online seems to be different with ours as well. Based on parser
+combinators, our pattern seems to be more concise and simpler to use.
 
@@
 
-We think type-safe composition and separate compilation really matter, since it is not a good
-idea to modify existing code; we are using OO and want to achieve better code reuse. As mentioned
-before, we hope project code to be modular and easily composable, instead of being compiled and
-generating code for every round of composition. It is better for debugging.
+We think type-safe composition and separate compilation really matter,
+since it is not a good idea to modify existing code; we are using OO
+and want to achieve better code reuse. As mentioned before, we hope
+project code to be modular and easily composable, instead of being
+compiled and generating code for every round of composition. It is
+better for debugging.
 

### 2017-04-24 ECOOP_17/paper.tex
Move ‘paper’ to ‘ECOOP_17’

--- 2017-01-14 paper/paper.tex
@@
 \usepackage{multirow}
-
+\usepackage{enumitem}
 \usepackage{listings}
@@
 
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Editor-only macros:: begin (do not touch as
+%author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment}
 \EventEditors{John Q. Open and Joan R. Acces}
@@
 \ArticleNo{23}
+\end{comment}
 % Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \emph{modularly type-checked} and \emph{separately compiled}. We
-  present a technique based parsing combinators that enables modular
+  present a technique based on parser combinators that enables modular
   parsing. Interestingly, the modularity requirements for modular
-  parsing rule out several existing parsing combinator approaches,
+  parsing rule out several existing parser combinator approaches,
   which rely on some non-modular techniques. We show that Packrat
@@
   the approach we conduct a case study based on the ``Types and
-  Programming Languages'' interpreters. The case study the
+  Programming Languages'' interpreters. The case study shows the
   effectiveness at reusing parsing code from existing interpreters,
   and the total parsing code is 69\% shorter than an existing code
-  based using non-modular parsing code.
+  base using a non-modular parsing approach.
 

### 2017-04-24 ECOOP_17/sections/CaseStudy.tex
Move ‘paper’ to ‘ECOOP_17’

--- 2017-01-14 paper/sections/CaseStudy.tex
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \lstinputlisting[linerange=37-64]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
@@
 the amount of code reuse using our modular parsing approach.
-For this propose, we measured source lines of code (SLOC) of two implementations.
+For this purpose, we measured source lines of code (SLOC) of two implementations.
 Second, we are interested to assess the performance penalty caused by modularity.
@@
         Untyped & 770 & 1018 & +32.2 & 821 & +6.6 & 1007 & +30.8 \\
-      \hline
-        \multicolumn{1}{|c|}{\dots} & \multicolumn{7}{c|}{\dots} \\
-      \hline
+        FullUntyped & 1297 & 1854 & +42.9 & 1343 & +3.5 & 1767 & +36.2 \\
+        TyArith & 746 & 888 & +19.0 & 772 & +3.5 & 918 & +23.1 \\
+        SimpleBool & 1376 & 1782 & +29.5 & 1494 & +8.6 & 1824 & +32.6 \\
+        FullSimple & 1441 & 2270 & +57.5 & 1574 & +9.2 & 2226 & +54.5 \\
+        Bot & 1080 & 1287 & +19.2 & 1078 & -0.2 & 1306 & +20.9 \\
+      %\hline
+        %\multicolumn{1}{|c|}{\dots} & \multicolumn{7}{c|}{\dots} \\
+      %\hline
+        FullRef & 1438 & 2291 & +59.3 & 1544 & +7.4 & 2142 & +49.0 \\
+        FullError & 1410 & 1946 & +38.0 & 1524 & +8.1 & 1981 & +40.5 \\
+        RcdSubBot & 1247 & 1524 & +22.2 & 1285 & +3.0 & 1612 & +29.3 \\
+        FullSub & 1320 & 1979 & +49.9 & 1393 & +5.5 & 1899 & +43.9 \\
+        FullEquiRec & 1407 & 2200 & +56.4 & 1561 & +10.9 & 2156 & +53.2 \\
+        FullIsoRec & 1492 & 2253 & +51.0 & 1648 & +10.5 & 2236 & +49.9 \\
+        EquiRec & 994 & 1254 & +26.2 & 1048 & +5.4 & 1304 & +31.2 \\
+        Recon & 1044 & 1482 & +42.0 & 1128 & +8.0 & 1506 & +44.3 \\
         FullRecon & 1094 & 1645 & +50.4 & 1161 & +6.1 & 1652 & +51.0 \\
@@
 
-Table \ref{tab:ext-comparison} shows part of the comparison results of these four versions,
-full table is available online. It suggests that the difference of running time between
+Table \ref{tab:ext-comparison} shows comparison results of these four versions. It suggests that the difference of running time between
 using object algebra and class is little, roughly 1\%.

### 2017-04-24 ECOOP_17/sections/Introduction.tex
Move ‘paper’ to ‘ECOOP_17’

--- 2017-01-14 paper/sections/Introduction.tex
@@
   require full information about the grammar to generate parsing
-  code. Moreover, actions associated to grammar productions are
+  code. Moreover, actions associated with grammar productions are
   typically only type-checked after the parser has been
@@
 
-  To address such algorithmic challanges, we propose a methodology
+  To address such algorithmic challenges, we propose a methodology
   for implementing modular parsers built on top of an
@@
   at:
-  \url{http://}\footnote{{\bf Note to Reviewers:} Please find the url in the
-  suplementary material for the submission.}
+  \url{http://}\footnote{{\bf Note to Reviewers:} Please find the URL in the
+  supplementary material for the submission.}
 

### 2017-04-24 ECOOP_17/sections/PackratParsing.tex
Move ‘paper’ to ‘ECOOP_17’

--- 2017-01-14 paper/sections/PackratParsing.tex
@@
 are important because they rule out various common techniques
-used by non-modular code using parsing combinators.
+used by non-modular code using parser combinators.
 To avoid pitfalls related to those algorithmic challenges,
@@
 since it is inconvenient and introduces extra complexity for
-representation of a grammar and implementation of a parser.
+representation of grammars and implementation of parsers.
 

### 2017-04-24 SLE_17/paper.tex
Copy content

%% For double-blind review submission
\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=false}
%% For single-blind review submission
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[sigplan,10pt]{acmart}\settopmatter{}

%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format should change 'sigplan,10pt' to
%% 'acmlarge'.

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{listings}
\usepackage{syntax}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{multirow}

\lstset{
  language=Scala,                % choose the language of the code
  columns=flexible,
  lineskip=-1pt,
  basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
  numbers=none,                   % where to put the line-numbers
  numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
  stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  morekeywords={var, lazy},
  %  frame=single,                   % adds a frame around the code
  tabsize=2,                  % sets default tabsize to 2 spaces
  captionpos=none,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
  escapeinside={(**}{**)},          % if you want to add a comment within your code
  keywordstyle=\ttfamily\bfseries,
}

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

\lstdefinelanguage{PlainCode}{}
\newcommand\inlinecode[1]{\lstinline[language=PlainCode]{#1}}

\newcommand{\mynote}[3]{{\color{#2} {#1}: #3}}
\newcommand\bruno[1]{\mynote{bruno}{red}{#1}}
\newcommand\huang[1]{\mynote{huang}{blue}{#1}}
\newcommand\haoyuan[1]{\mynote{haoyuan}{cyan}{#1}}

\begin{document}

%% Title information
\title{Type-Safe Modular Parsing}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%%\subtitle{Subtitle}                     %% \subtitle is optional
%%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Paper note
%% The \thanks command may be used to create a "paper note" ---
%% similar to a title note or an author note, but not explicitly
%% associated with a particular element.  It will appear immediately
%% above the permission/copyright statement.
\thanks{with paper note}                %% \thanks is optional
                                        %% can be repeated if necesary
                                        %% contents suppressed with 'anonymous'

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Over the years a lot of effort has been put on solving
  extensibility problems, while retaining important software
  engineering properties such as modular type-safety and
  separate compilation. Most previous work focused on
  operations that traverse and process extensible Abstract Syntax Tree
  (AST) structures. However, there is almost no work on operations that
  \emph{build} such extensible ASTs, including \emph{parsing}.

  This paper investigates solutions for the problem of \emph{modular
    parsing}. We focus on \emph{semantic} modularity and not just
  \emph{syntactic} modularity. That is, the solutions should not only
  allow complete parsers to be built out of modular parsing
  components, but also enable the parsing components to be
  \emph{modularly type-checked} and \emph{separately compiled}. We
  present a technique based on parser combinators that enables modular
  parsing. Interestingly, the modularity requirements for modular
  parsing rule out several existing parser combinator approaches,
  which rely on some non-modular techniques. We show that Packrat
  parsing techniques, provide solutions for such modularity problems,
  and enable reasonable performance in a modular setting.
  Extensibility is achieved using \emph{multiple inheritance} and
  \emph{Object Algebras}.  To evaluate
  the approach we conduct a case study based on the ``Types and
  Programming Languages'' interpreters. The case study shows the
  effectiveness at reusing parsing code from existing interpreters,
  and the total parsing code is 69\% shorter than an existing code
  base using a non-modular parsing approach.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\input{sections/Introduction.tex}

\input{sections/PackratParsing.tex}

\input{sections/Inheritance.tex}

\input{sections/ObjectAlgebras.tex}

\input{sections/Discussion.tex}

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{paper}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2017-04-24 SLE_17/sections/Discussion.tex
Copy content

--- 2017-01-14 paper/sections/Discussion.tex
@@
 
-\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<type> ::= `int' \alt <type> `->' <type>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::=  ... \alt `\\' <ident> `:' <type> `.' <expr>
-\end{grammar}
-\end{tabular}
+\huang{todo}
+%% \begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
+%% \setlength{\grammarindent}{5em}
+%% \begin{grammar}
+%% <type> ::= `int' \alt <type> `->' <type>
+%% \end{grammar}
+%% &
+%% \setlength{\grammarindent}{5em}
+%% \begin{grammar}
+%% <expr> ::=  ... \alt `\\' <ident> `:' <type> `.' <expr>
+%% \end{grammar}
+%% \end{tabular}
 
@@
 
-\begin{figure}[t]
-\begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
-\lstinputlisting[linerange=113-132]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
-&
-\lstinputlisting[linerange=136-156]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
-\end{tabular}
-\caption{Language components \lstinline{VarExpr} (left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
-\end{figure}
+\huang{todo}
+%% \begin{figure}[t]
+%% \begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
+%% \lstinputlisting[linerange=113-132]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+%% &
+%% \lstinputlisting[linerange=136-156]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+%% \end{tabular}
+%% \caption{Language components \lstinline{VarExpr} (left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
+%% \end{figure}
 

### 2017-04-24 SLE_17/sections/Introduction.tex
Copy content

--- 2017-04-24 ECOOP_17/sections/Introduction.tex
@@
 (classes), while adding new operations is more difficult.
-
-\begin{comment}
-A reason why a solution to the Expression Problem is important in
-practice is that it is necessary for the development of
-\emph{Software-Product Lines} (SPLs)~\cite{}. A software-product line
-is a reusable set of components, which can be combined in multiple ways
-to obtain different programs. Programming languages offer a concrete
-example for SPLs. A SPL for programming languages would allow us to
-model various typical operations of programming languages (such as
-evaluation, compilation, or parsing) for various different language
-constructs (such as binding, arithmetic, conditional or loops)
-independently and separately. For example, evaluation components could be defined
-independently for binding and arithmetic constructs. If the language
-to be implemented is the pure lambda calculus, only evaluation of
-binding constructs is necessary. However, more realistic programming
-languages will include arithmetic constructs, and will require
-evaluation for such constructs as well. In this case
-both the component for evaluation of binders and arithmetic
-expressions can be combined to implement the desired functionality.
-\end{comment}
-
-\begin{comment}
-Most programming languages share alot of features in
-common.
-
-For example, most languages have language constructs for:
-binding (such as variables, functions, and function applications);
-basic arithmetic operations; basic logic and conditional operations;
-loops; as well as various other features. For each language construct,
-various operations (such as evaluation, compilation, or parsing) need
-to be implemented. It is reasonable to wonder whether we can simply
-implement those features independently of a particular implementation
-of a programming language. Evaluation could be defined independently
-for binding and arithmetic constructs. If the language to be
-implemented is the pure lambda calculus, only evaluation of binding
-constructs is necessary. Thus only the component that implements
-evaluation for binding needs to be used in such an implementation.
-However, more realistic programming languages
-will include arithmetic constructs, and will require an evaluation
-function for those.
-
-Then it would be possible to \emph{reuse}
-some of those features in \emph{multiple} different implementations of
-programming languages. Essentially, this would enable a SPL for
-programming languages, where all
-
-A solution to the Expression Problem could ena
-
-A concrete
-example that illustrates this issue is
-\end{comment}
 
@@
 
-\begin{comment}
-  By analising the \emph{full} grammar it is possible to remove
-  backtracking, which would otherwise increase parsing times. Many
-  parsing combinator libraries routinely use backtracting elimination
-  to achieve performance. However, in a modular setting this technique
-  cannot be used, because the full grammar is not known. Thus we have
-  to be very conservative at eliminating backtracting. Unfortunatelly,
-  this has a severe impact on performance.
-\end{comment}
-
 To evaluate our approach we conduct a case study based on the ``Types
@@
 
-\begin{itemize}[noitemsep,nolistsep]
+\begin{itemize}
 
-\item {{\bf A Technique for Modular Parsing:}} We present a technique
-that allows the development of semantically modular parsers.
+\item \textbf{A Technique for Modular Parsing:}
+  We present a technique that allows the development of semantically modular parsers.
   The technique relies on the combination of Packrat parsing, multiple inheritance
@@
 
-\item {{\bf A Parsing Technique for OO ASTs:}} A simplified version of
+\item \textbf{A Parsing Technique for OO ASTs:} A simplified version of
   our technique also enables parsing OO-style ASTs, where new language
@@
 
-\item {{\bf A Methodology for Writing Modular Parsing Code:}} We
+\item \textbf{A Methodology for Writing Modular Parsers:} We
   identify possible pitfalls using parser combinators. To avoid such
@@
 
-\item {{\bf TAPL case study:}} We conduct a case study with 18 interpreters
+\item \textbf{TAPL case study:} We conduct a case study with 18 interpreters
   from the TAPL book. The case study show the effectiveness of modular
@@
   at:
-  \url{http://}\footnote{{\bf Note to Reviewers:} Please find the URL in the
+  \url{http://}\footnote{\textbf{Note to Reviewers:} Please find the URL in the
     supplementary material for the submission.}

### 2017-04-24 SLE_17/sections/ObjectAlgebras.tex
Copy content

--- 2017-01-14 paper/sections/ObjectAlgebras.tex
@@
 \lstinline{VarExprParser} is no longer suitable for parsing the new ASTs!
-All the methods in the previous parser has type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
+All the methods in the previous parser have type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
 
@@
 \paragraph{A Solution}
-A solution to this problem is instead have a field with the Object Algebra
+A solution to this problem is instead having a field with the Object Algebra
 in \lstinline{ExprOAParser}. This approach supports overriding as well.

### 2017-04-24 SLE_17/sections/OpenRecursion.tex
Copy content

--- 2017-01-03 paper/sections/OpenRecursion.tex
@@
 
-\lstinputlisting[linerange=8-12]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PVAR
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_PVAR
 
@@
 
-\lstinputlisting[linerange=16-17]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PAPP
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_PAPP
 
@@
 
-\lstinputlisting[linerange=29-32]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PAPP2
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_PAPP2
 
@@
 
-\lstinputlisting[linerange=44-49]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PLIT
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_PLIT
 \bruno{END: Text to be deleted/moved/rewritten in S2}
@@
 
-\lstinputlisting[linerange=63-63]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PAPPFIX
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_PAPPFIX
 
@@
 
-\lstinputlisting[linerange=67-75]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_COMBINATOR
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_COMBINATOR
 
@@
 
-\lstinputlisting[linerange=79-79]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_PVARAPP
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_PVARAPP
 
@@
 
-\lstinputlisting[linerange=83-102]{../Scala/Parser/src/PaperCode/ParsingWithOR.scala}% APPLY:linerange=PARSING_TYPEDEXPR
+\lstinputlisting[linerange=-]{}% APPLY:linerange=PARSING_TYPEDEXPR
 

### 2017-04-24 SLE_17/sections/PackratParsing.tex
Copy content

--- 2017-04-24 ECOOP_17/sections/PackratParsing.tex
@@
 
-\item {\bf Modular parsers should be left-recursive;}
+\item \textbf{Modular parsers should be left-recursive;}
 
-\item {\bf Modular parsers should use a longest match composition operator.}
+\item \textbf{Modular parsers should use a longest match composition operator.}
 
@@
 
-\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int>
-\end{grammar}
-&
-\begin{lstlisting}[language=PlainCode]
-parseExpr = parseInt
-\end{lstlisting}
-\end{tabular}
+\huang{todo}
+
+%% \begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
+%% \setlength{\grammarindent}{5em}
+%% \begin{grammar}
+%% <expr> ::= <int>
+%% \end{grammar}
+%% &
+%% \begin{lstlisting}[language=PlainCode]
+%% parseExpr = parseInt
+%% \end{lstlisting}
+%% \end{tabular}
 
@@
 
-\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> \alt <expr> `-' <int>
-\end{grammar}
-&
-\begin{lstlisting}[language=PlainCode]
-parseExpr = parseSub <|> parseInt
-parseSub = do
-  e <- parseExpr
-  ...
-\end{lstlisting}
-\end{tabular}
+\huang{todo}
+%% \begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
+%% \setlength{\grammarindent}{5em}
+%% \begin{grammar}
+%% <expr> ::= <int> \alt <expr> `-' <int>
+%% \end{grammar}
+%% &
+%% \begin{lstlisting}[language=PlainCode]
+%% parseExpr = parseSub <|> parseInt
+%% parseSub = do
+%%   e <- parseExpr
+%%   ...
+%% \end{lstlisting}
+%% \end{tabular}
 
@@
 
-\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int> <expr'>
+\huang{todo}
+%% \begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
+%% \setlength{\grammarindent}{5em}
+%% \begin{grammar}
+%% <expr> ::= <int> <expr'>
 
-<expr'> ::= <empty> \alt `-' <int> <expr'>
-\end{grammar}
-&
-\end{tabular}
+%% <expr'> ::= <empty> \alt `-' <int> <expr'>
+%% \end{grammar}
+%% &
+%% \end{tabular}
 
@@
 
-\begin{table}[t]
-\begin{tabular}{l}
-\hline
-\begin{lstlisting}
-def ~[U](q: => Parser[U]): Parser[~[T, U]]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition. \\
-\hline
-\begin{lstlisting}
-def ^^[U](f: (T) => U): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for function application. \\
-\hline
-\begin{lstlisting}
-def ^^^[U](v: => U): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
-\hline
-\begin{lstlisting}
-def <~[U](q: => Parser[U]): Parser[T]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
-\hline
-\begin{lstlisting}
-def ~>[U](q: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
-\hline
-\begin{lstlisting}
-def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
-\end{lstlisting} \\
-\hspace{.2in}- A parser generator for interleaved repetitions. \\
-\hline
-\begin{lstlisting}
-def ident: Parser[String]
-\end{lstlisting} \\
-\hspace{.2in}- A parser which matches an identifier. \\
-\hline
-\begin{lstlisting}
-def numericLit: Parser[String]
-\end{lstlisting} \\
-\hspace{.2in}- A parser which matches a numeric literal. \\
-\hline
-\begin{lstlisting}
-def |[U >: T](q: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for alternative composition. \\
-\hline
-\begin{lstlisting}
-def |||[U >: T](q0: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for alternative with longest match composition. \\
-\hline \\
-\end{tabular}
-\caption{Common combinators from the Scala standard parser combinator library.}\label{tab:packrat}
-\end{table}
+\huang{todo}
+
+%% \begin{table}[t]
+%% \begin{tabular}{l}
+%% \hline
+%% \begin{lstlisting}
+%% def ~[U](q: => Parser[U]): Parser[~[T, U]]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator for sequential composition. \\
+%% \hline
+%% \begin{lstlisting}
+%% def ^^[U](f: (T) => U): Parser[U]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator for function application. \\
+%% \hline
+%% \begin{lstlisting}
+%% def ^^^[U](v: => U): Parser[U]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
+%% \hline
+%% \begin{lstlisting}
+%% def <~[U](q: => Parser[U]): Parser[T]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
+%% \hline
+%% \begin{lstlisting}
+%% def ~>[U](q: => Parser[U]): Parser[U]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
+%% \hline
+%% \begin{lstlisting}
+%% def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser generator for interleaved repetitions. \\
+%% \hline
+%% \begin{lstlisting}
+%% def ident: Parser[String]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser which matches an identifier. \\
+%% \hline
+%% \begin{lstlisting}
+%% def numericLit: Parser[String]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser which matches a numeric literal. \\
+%% \hline
+%% \begin{lstlisting}
+%% def |[U >: T](q: => Parser[U]): Parser[U]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator for alternative composition. \\
+%% \hline
+%% \begin{lstlisting}
+%% def |||[U >: T](q0: => Parser[U]): Parser[U]
+%% \end{lstlisting} \\
+%% \hspace{.2in}- A parser combinator for alternative with longest match composition. \\
+%% \hline \\
+%% \end{tabular}
+%% \caption{Common combinators from the Scala standard parser combinator library.}\label{tab:packrat}
+%% \end{table}
 

### 2017-04-24 SLE_17/run.sh

@@ -1,0 +1,4 @@
+pdflatex paper.tex
+bibtex paper
+pdflatex paper.tex
+pdflatex paper.tex

### 2017-04-24 SLE_17/sections/CaseStudy.tex
Code & content

--- 2017-04-24 ECOOP_17/sections/CaseStudy.tex
@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=15-33]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
+\lstinputlisting[linerange=15-33]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
 \caption{The \inlinecode{VarApp} component in case study.}\label{fig:casestudy-varapp}
@@
 
-\lstinputlisting[linerange=9-11]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
+\lstinputlisting[linerange=9-11]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
 
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=37-64]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
+\lstinputlisting[linerange=37-64]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
 \caption{Build the \inlinecode{Untyped} calculus by composing to language components.}\label{fig:casestudy-untyped}
@@
 
-\lstinputlisting[linerange=68-78]{../Scala/Parser/src/PaperCode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_DEMO
+\lstinputlisting[linerange=68-78]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_DEMO
 

### 2017-04-24 SLE_17/sections/Discussion.tex
Code & content

--- 2017-01-14 paper/sections/Discussion.tex
@@
 As an example, we extend the language of literals, additions and variables by introducing types, including primitive type \inlinecode{int} and function (arrow) type, and we add lambda abstractions to expressions.
-Now the language has two sorts in the syntax, types and expressions:\\
+Now the language has two sorts in the syntax, types and expressions:
 
-\begin{tabular}{m{0.45\linewidth}m{0.45\linewidth}}
 \setlength{\grammarindent}{5em}
@@
 <type> ::= `int' \alt <type> `->' <type>
-\end{grammar}
-&
-\setlength{\grammarindent}{5em}
-\begin{grammar}
+
 <expr> ::=  ... \alt `\\' <ident> `:' <type> `.' <expr>
 \end{grammar}
-\end{tabular}
 
@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=31-59]{code/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 \caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
@@
 
-\lstinputlisting[linerange=63-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=63-66]{code/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 The code works as expected to parse and pretty-print the result.
@@
 
-\lstinputlisting[linerange=76-81]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=76-81]{code/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
 
@@
 
-\lstinputlisting[linerange=85-90]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\lstinputlisting[linerange=85-90]{code/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
 
 One can even ``eliminate'' a production rule in the extension, by overriding it with a failure parser. The lexer can also be updated, since keywords and delimiters are represented by sets of strings.
-
-\begin{comment}
-It is natural to keep overriding existing parsers, whereas another potential use of this pattern is to reuse old versions of a parser. For instance,
-we have a \lstinline{NewParser} that extends \lstinline{ExtParser}, with some more cases added, but for the lambda case we want to go back to the untyped one.
-It can be achieved by instantiating an instance of \lstinline{BaseParser} to obtain the old \lstinline{pLam}.
-However, we must have \inlinecode{lazy} modifier for all the \inlinecode{pLam} starting from \inlinecode{BaseParser}.
-
-\lstinputlisting[linerange=94-102]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM
-\end{comment}
 
@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
     \item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax. The type parameters represent multiple sorts of syntax, and the methods are constructs.
@@
 \begin{figure}[t]
-\begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
-\lstinputlisting[linerange=113-132]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
-&
-\lstinputlisting[linerange=136-156]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
-\end{tabular}
-\caption{Language components \lstinline{VarExpr} (left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
+\lstinputlisting[linerange=113-132]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=136-156]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\caption{Language components \lstinline{VarExpr} (\huang{todo}left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
 \end{figure}
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=160-170]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=160-170]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 \caption{Reuse two language components \lstinline{VarExpr} and \lstinline{TypedLam}.}\label{fig:compose-components}
@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
 
-\item {\bf Parsing Technique}
-
+\item {\bf Parsing Technique:}
 Our approach does not depend on a particular parser combinator library
@@
 
-\item {\bf Composing and Extending Parsers}
-
+\item {\bf Composing and Extending Parsers:}
   We use traits to model parsers, OO inheritance to compose them, and
@@
 
-\item {\bf Extensible ASTs}
-
+\item {\bf Extensible ASTs:}
 Besides Object Algebras, many other techniques including \textit{Data
@@
 modular, which means it can be customized easily.
-%We conducted some previous experiments of modular parsing in Haskell, based on open recursion, MRM and Parsec.
-However, we suffered from issues of Parsec, as illustrated in Section~\ref{subsec:challenges}.

### 2017-04-24 SLE_17/sections/Inheritance.tex
Code & content

--- 2017-01-14 paper/sections/Inheritance.tex
@@
 style mechanisms such as subtyping and inheritance, as well as,
-Packrat parsing combinators, and enables a better integration between
+Packrat parser combinators, and enables a better integration between
 parsing and OO ASTs.
@@
 
-\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
+\lstinputlisting[linerange=29-31]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
 
@@
 
-\lstinputlisting[linerange=35-38]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
+\lstinputlisting[linerange=35-38]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
 
@@
 
-\lstinputlisting[linerange=42-47]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
+\lstinputlisting[linerange=42-47]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
 
@@
 
-\lstinputlisting[linerange=67-73]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
+\lstinputlisting[linerange=67-73]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 
@@
 
-\lstinputlisting[linerange=79-95]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
+\lstinputlisting[linerange=79-95]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
 

### 2017-04-24 SLE_17/sections/ObjectAlgebras.tex
Code & content

@@
 
-\lstinputlisting[linerange=23-23]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=FVARSEXPR
+\lstinputlisting[linerange=23-23]{code/Sec4OA/Code1.scala}% APPLY:linerange=FVARSEXPR
 
@@
 
-\lstinputlisting[linerange=27-37]{../Scala/Parser/src/PaperCode/Sec4OA/Code1.scala}% APPLY:linerange=FVAR_CLASSES
+\lstinputlisting[linerange=27-37]{code/Sec4OA/Code1.scala}% APPLY:linerange=FVAR_CLASSES
 
@@
 
-\lstinputlisting[linerange=5-8]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
+\lstinputlisting[linerange=5-8]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
 We have two constructors \inlinecode{lit} and \inlinecode{add}, corresponding to literals and additions.
@@
 
-\lstinputlisting[linerange=12-15]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
+\lstinputlisting[linerange=12-15]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
 Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
@@
 
-\lstinputlisting[linerange=19-22]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
+\lstinputlisting[linerange=19-22]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
@@
 
-\lstinputlisting[linerange=26-28]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+\lstinputlisting[linerange=26-28]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
 Now pretty-printing on the new language can be realized without modifying existing code:
 
-\lstinputlisting[linerange=32-34]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+\lstinputlisting[linerange=32-34]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
 An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
@@
 
-\lstinputlisting[linerange=41-41]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+\lstinputlisting[linerange=41-41]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
 
@@
 
-\lstinputlisting[linerange=8-16]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+\lstinputlisting[linerange=8-16]{code/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
 Such a parser works fine, but it is not extensible!
@@
 
-\lstinputlisting[linerange=23-30]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
+\lstinputlisting[linerange=23-30]{code/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
@@
 
-\lstinputlisting[linerange=34-39]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+\lstinputlisting[linerange=34-39]{code/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
 
@@
 
-\lstinputlisting[linerange=43-46]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
+\lstinputlisting[linerange=43-46]{code/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
 
@@
 
-\lstinputlisting[linerange=45-56]{../Scala/Parser/src/PaperCode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
+\lstinputlisting[linerange=45-56]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
 Here the \inlinecode{VarExprAlg} expression \inlinecode{1 + x} is changed to \inlinecode{ExprAlg} expression \inlinecode{1 + 2}, before applying the pretty-printing algebra. The abstract field

### 2017-04-24 SLE_17/sections/PackratParsing.tex
Code & content

--- 2017-04-24 ECOOP_17/sections/PackratParsing.tex
@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
 
-\item {\bf Modular parsers should be left-recursive;}
+\item \textbf{Modular parsers should support left-recursion.}
 
-\item {\bf Modular parsers should use a longest match composition operator.}
+\item \textbf{Modular parsers should use a longest match composition operator.}
 
@@
 
-\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
+\begin{minipage}{10em}
 \setlength{\grammarindent}{5em}
@@
 \end{grammar}
-&
+\end{minipage}
+\begin{minipage}{10em}
 \begin{lstlisting}[language=PlainCode]
@@
 \end{lstlisting}
-\end{tabular}
+\end{minipage}
 
@@
 
-\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
-\setlength{\grammarindent}{5em}
+\begin{minipage}{12em}
+\setlength{\grammarindent}{4em}
 \begin{grammar}
@@
 \end{grammar}
-&
+\end{minipage}
+\begin{minipage}{15em}
 \begin{lstlisting}[language=PlainCode]
-parseExpr = parseSub <|> parseInt
+parseExpr =
+  parseSub <|> parseInt
+
 parseSub = do
@@
 \end{lstlisting}
-\end{tabular}
+\end{minipage}
 
-To solve this issue, we have to rewrite the grammar as below to eliminate left-recursion and build the parser based on this new grammar.\\
+To solve this issue, we have to rewrite the grammar as below to eliminate left-recursion and build the parser based on this new grammar.
 
-\begin{tabular}{m{0.4\linewidth}m{0.5\linewidth}}
+\vspace{0.4\baselineskip}
+\begin{minipage}{15em}
 \setlength{\grammarindent}{5em}
@@
 \end{grammar}
-&
-\end{tabular}
+\end{minipage}
+\vspace{0.4\baselineskip}
 
@@
 \centering
-\lstinputlisting[linerange=42-53]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_PAPERCODE
+\lstinputlisting[linerange=42-53]{code/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_PAPERCODE
 \caption{Helper object for code demonstration in this paper.}\label{fig:papercode}
@@
 
-\lstinputlisting[linerange=9-17]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
+\lstinputlisting[linerange=9-17]{code/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
 
@@
 
-\lstinputlisting[linerange=21-29]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
+\lstinputlisting[linerange=21-29]{code/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
 
@@
 
-\begin{table}[t]
-\begin{tabular}{l}
+\begin{table}
+\begin{tabularx}{\linewidth}{X}
 \hline
-\begin{lstlisting}
-def ~[U](q: => Parser[U]): Parser[~[T, U]]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition. \\
-\hline
-\begin{lstlisting}
-def ^^[U](f: (T) => U): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for function application. \\
-\hline
-\begin{lstlisting}
-def ^^^[U](v: => U): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator that changes a successful result into the specified value. \\
-\hline
-\begin{lstlisting}
-def <~[U](q: => Parser[U]): Parser[T]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition which keeps only the left result. \\
-\hline
-\begin{lstlisting}
-def ~>[U](q: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for sequential composition which keeps only the right result. \\
-\hline
-\begin{lstlisting}
-def repsep[T](p: => Parser[T], q: => Parser[Any]): Parser[List[T]]
-\end{lstlisting} \\
-\hspace{.2in}- A parser generator for interleaved repetitions. \\
-\hline
-\begin{lstlisting}
-def ident: Parser[String]
-\end{lstlisting} \\
-\hspace{.2in}- A parser which matches an identifier. \\
-\hline
-\begin{lstlisting}
-def numericLit: Parser[String]
-\end{lstlisting} \\
-\hspace{.2in}- A parser which matches a numeric literal. \\
-\hline
-\begin{lstlisting}
-def |[U >: T](q: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for alternative composition. \\
-\hline
-\begin{lstlisting}
-def |||[U >: T](q0: => Parser[U]): Parser[U]
-\end{lstlisting} \\
-\hspace{.2in}- A parser combinator for alternative with longest match composition. \\
-\hline \\
-\end{tabular}
+\lstinline{def ~[U](q: => Parser[U]): Parser[~[T, U]]} \\
+- A parser combinator for sequential composition. \\ \hline
+\lstinline{def \^\^[U](f: (T) => U): Parser[U]} \\
+- A parser combinator for function application. \\ \hline
+\lstinline{def \^\^\^[U](v: => U): Parser[U]} \\
+- A parser combinator that changes a successful result into the specified value. \\ \hline
+\lstinline{def <~[U](q: => Parser[U]): Parser[T]} \\
+- A parser combinator for sequential composition which keeps only the left result. \\ \hline
+\lstinline{def ~>[U](q: => Parser[U]): Parser[U]} \\
+- A parser combinator for sequential composition which keeps only the right result. \\ \hline
+\lstinline{def ident: Parser[String]} \\
+- A parser which matches an identifier. \\ \hline
+\lstinline{def numericLit: Parser[String]} \\
+- A parser which matches a numeric literal. \\ \hline
+\lstinline{def |[U >: T](q: => Parser[U]): Parser[U]} \\
+- A parser combinator for alternative composition. \\ \hline
+\lstinline{def |||[U >: T](q0: => Parser[U]): Parser[U]} \\
+- A parser combinator for alternative with longest match composition. \\ \hline
+\end{tabularx}
 \caption{Common combinators from the Scala standard parser combinator library.}\label{tab:packrat}
@@
 
-\lstinputlisting[linerange=33-33]{../Scala/Parser/src/PaperCode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+\lstinputlisting[linerange=33-33]{code/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER

### 2017-04-24 SLE_17/code/Common.scala

@@ -1,0 +1,19 @@
+package PaperCode
+
+import scala.util.parsing.combinator.syntactical._
+import scala.util.parsing.combinator._
+
+
+trait Common extends StandardTokenParsers with PackratParsers {
+
+  type Parser[E] = PackratParser[E]
+
+  val lcid: Parser[String] = ident ^? { case id if id.charAt(0).isLower => id }
+  val ucid: Parser[String] = ident ^? { case id if id.charAt(0).isUpper => id }
+
+  def parse[E](p: Parser[E]): String => E = in => {
+    val t = phrase(p)(new lexical.Scanner(in))
+    if (t.successful) t.get else scala.sys.error(t.toString)
+  }
+
+}

### 2017-04-24 SLE_17/code/Sec3Inheritance/Code1.scala

@@ -1,0 +1,126 @@
+package PaperCode.Sec3Inheritance
+
+import PaperCode.Common
+
+
+object Code1 extends Common {
+
+  trait Expr {
+    def print: String
+  }
+  class Lit(x: Int) extends Expr {
+    def print = x.toString
+  }
+  class Add(e1: Expr, e2: Expr) extends Expr {
+    def print = "(" + e1.print + " + " + e2.print + ")"
+  }
+
+  trait ExprParser {
+    lexical.delimiters += "+"
+    val pLit: PackratParser[Expr] =
+      numericLit ^^ { x => new Lit(x.toInt) }
+    val pAdd: PackratParser[Expr] =
+      pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
+    def pExpr: PackratParser[Expr] =
+      pLit ||| pAdd
+  }
+
+//BEGIN_INHERITANCE_SIMPLE_LAM
+class Var(x: String) extends Expr {
+  def print = x
+}
+//END_INHERITANCE_SIMPLE_LAM
+
+//BEGIN_INHERITANCE_BAD_ATTEMPT
+trait Attempt extends ExprParser {
+  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
+  val pVarExpr: Parser[Expr] = pExpr ||| pVar
+}
+//END_INHERITANCE_BAD_ATTEMPT
+
+//BEGIN_INHERITANCE_APPROACH
+trait VarExprParser extends ExprParser {
+  val pVar: Parser[Expr] = ident ^^ { x => new Var(x) }
+  override def pExpr: Parser[Expr] = super.pExpr ||| pVar
+}
+
+val r = parse(new VarExprParser {}.pExpr)("1 + x").print // "(1 + x)"
+//END_INHERITANCE_APPROACH
+
+
+  // ---------------------------------------------------
+
+  trait LanguageA {
+    def pExpr: Parser[String] = ident
+  }
+
+  trait LanguageB {
+    def pExpr: Parser[String] = ident
+  }
+
+  trait LanguageC extends LanguageA with LanguageB {
+    override def pExpr = super[LanguageA].pExpr ||| super[LanguageB].pExpr
+  }
+
+/*
+//BEGIN_MULTIPLE_INHERITANCE
+trait LanguageA {...}
+
+trait LanguageB {...}
+
[…]

### 2017-04-24 SLE_17/code/Sec4OA/Code1.scala

@@ -1,0 +1,40 @@
+package PaperCode.Sec4OA
+
+
+object Code1 {
+
+  trait Expr {
+    def print: String
+  }
+
+  class Lit(x: Int) extends Expr {
+    def print = x.toString
+  }
+
+  class Add(e1: Expr, e2: Expr) extends Expr {
+    def print = "(" + e1.print + " + " + e2.print + ")"
+  }
+
+  class Var(x: String) extends Expr {
+    def print = x
+  }
+
+//BEGIN_FVARSEXPR
+trait FVarsExpr extends Expr { def fVars: Set[String] }
+//END_FVARSEXPR
+
+//BEGIN_FVAR_CLASSES
+class FVarsLit(x: Int) extends Lit(x) with FVarsExpr {
+  def fVars = Set()
+}
+
+class FVarsAdd(e1: FVarsExpr, e2: FVarsExpr) extends Add(e1, e2) with FVarsExpr {
+  def fVars = e1.fVars ++ e2.fVars
+}
+
+class FVarsVar(x: String) extends Var(x) with FVarsExpr {
+  def fVars = Set(x)
+}
+//END_FVAR_CLASSES
+
+}

### 2017-04-24 SLE_17/code/Sec4OA/Code3.scala

@@ -1,0 +1,52 @@
+package PaperCode.Sec4OA
+
+import PaperCode.Common
+
+
+object Code30 extends Common {
+//BEGIN_BASE_OA_PARSER_BAD
+trait ExprOAParser[E] {
+  lexical.delimiters += "+"
+  val pLit: ExprAlg[E] => Parser[E] = alg => numericLit ^^ { x => alg.lit(x.toInt) }
+  val pAdd: ExprAlg[E] => Parser[E] = alg => {
+    val p = pExpr(alg)
+    p ~ ("+" ~> p) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
+  }
+  val pExpr: ExprAlg[E] => Parser[E] = alg => pLit(alg) ||| pAdd(alg)
+}
+//END_BASE_OA_PARSER_BAD
+}
+
+
+object Code3 extends Common {
+//BEGIN_BASE_OA_PARSER
+trait ExprOAParser[E] {
+  lexical.delimiters += "+"
+  val alg: ExprAlg[E]
+  val pLit: Parser[E] = numericLit ^^ { x => alg.lit(x.toInt) }
+  val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
+  val pExpr: Parser[E] = pLit ||| pAdd
+  val pE: Parser[E] = pExpr
+}
+//END_BASE_OA_PARSER
+
+//BEGIN_EXT_OA_PARSER
+trait VarExprOAParser[E] extends ExprOAParser[E] {
+  override val alg: VarExprAlg[E]
+  val pVar: Parser[E] = ident ^^ alg.varE
+  val pVarExpr: Parser[E] = pExpr ||| pVar
+  override val pE: Parser[E] = pVarExpr
+}
+//END_EXT_OA_PARSER
+
+//BEGIN_EXT_OA_PARSER_CLIENT
+val p = new VarExprOAParser[String] {
+  override val alg = new VarExprPrint {}
+}
+val r = parse(p.pE)("1 + x") // "(1 + x)"
+//END_EXT_OA_PARSER_CLIENT
+
+  def main(args: Array[String]): Unit = {
+    println(r)
+  }
+}

### 2017-04-24 SLE_17/code/Sec5/Code1.scala

@@ -1,0 +1,174 @@
+package PaperCode.Sec5
+
+import PaperCode.Common
+
+
+object Code1 extends Common {
+  trait Alg[E] {
+    def varE(x: String): E
+
+    def lam(x: String, e: E): E
+  }
+
+  trait ExtAlg[E, T] extends Alg[E] {
+    def tyInt(): T
+
+    def tyLam(x: String, t: T, e: E): E
+  }
+
+  trait Print extends ExtAlg[String, String] {
+    override def varE(x: String) = x
+
+    override def lam(x: String, e: String) = "\\" + x + "." + e
+
+    override def tyInt() = "Int"
+
+    override def tyLam(x: String, t: String, e: String) = "\\" + x + ":" + t + "." + e
+  }
+
+  trait BaseParser[E] {
+    lexical.delimiters += ("\\", ".")
+    val alg: Alg[E]
+    val pVar: Parser[E] = ident ^^ alg.varE
+    lazy val pLam: Parser[E] = ("\\" ~> ident) ~ ("." ~> pE) ^^ { case x ~ e => alg.lam(x, e) }
+    val pE: Parser[E] = pLam ||| pVar
+  }
+
+  trait ExtParser[E, T] extends BaseParser[E] {
+    lexical.delimiters += ("\\", ":", ".")
+    lexical.reserved += "Int"
+    override val alg: ExtAlg[E, T]
+    override lazy val pLam: Parser[E] = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ { case x ~ t ~ e => alg.tyLam(x, t, e) }
+    val pT: Parser[T] = "Int" ^^^ alg.tyInt()
+  }
+
+  trait NewAlg[E, T] extends ExtAlg[E, T] {
+    def num(x: Int): E
+  }
+
+  trait NewPrint extends NewAlg[String, String] with Print {
+    def num(x: Int): String = x.toString
+  }
+
+  trait NewParser[E, T] extends ExtParser[E, T] {
+    override val alg: NewAlg[E, T]
+    override lazy val pLam: Parser[E] = new BaseParser[E] {
+      override val alg = NewParser.this.alg
+      override val pE: Parser[E] = NewParser.this.pE
+    }.pLam
[…]

### 2017-04-24 SLE_17/computePositions.rb

@@ -1,0 +1,66 @@
+#require 'tempfile'
+require 'fileutils'
+
+# Example 1 - Read File and close
+def scan(fileName, beginpos, endpos, fileloc)
+  file = File.new(fileName, "r")
+  counter = 1
+  while (line = file.gets)
+    if line =~ /BEGIN_(\w+)/
+      throw "Symbol '#{$1}' already defined" if fileloc[$1]
+      fileloc[$1] = fileName
+      beginpos[$1] = counter+1
+      puts "BEGIN #{$1}: #{beginpos[$1]}"
+    end
+    if line =~ /END_(\w+)/
+      endpos[$1] = counter-1
+      puts "END #{$1}: #{endpos[$1]}"
+    end
+    counter = counter + 1
+  end
+  file.close
+end
+
+def process(file, out, beginpos, endpos, fileloc)
+  while (line = file.gets)
+    if line =~ /APPLY:(\w+)=(\w+)/
+      var = $1
+      label = $2
+      line = line.sub( /#{var}=([0-9-]*)/, "#{var}=#{beginpos[label]}-#{endpos[label]}" )
+      line = line.sub( /\{[^{]*\}/, "{#{fileloc[label]}}" )
+    end
+    out.write(line)
+  end
+  file.close
+  out.close
+end
+
+fileloc = {}
+beginpos = {}
+endpos = {}
+
+
+Dir['code/*/*.scala'].each do |file|
+  scan(file, beginpos, endpos, fileloc)
+end
+
+
+
+Dir['*.tex'].each do |file|
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+
+Dir['*/*.tex'].each do |file|
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+

### 2017-04-24 SLE_17/run.sh

@@ -1,3 +1,4 @@
+ruby computePositions.rb
 pdflatex paper.tex
 bibtex paper
 pdflatex paper.tex

### 2017-04-24 SLE_17/sections/CaseStudy.tex
Revise code (up to section 4) to fit the narrower column

@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=15-33]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
+\lstinputlisting[linerange=15-33]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
 \caption{The \inlinecode{VarApp} component in case study.}\label{fig:casestudy-varapp}
@@
 
-\lstinputlisting[linerange=9-11]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
+\lstinputlisting[linerange=9-11]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
 
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=37-64]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
+\lstinputlisting[linerange=37-64]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
 \caption{Build the \inlinecode{Untyped} calculus by composing to language components.}\label{fig:casestudy-untyped}
@@
 
-\lstinputlisting[linerange=68-78]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_DEMO
+\lstinputlisting[linerange=68-78]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_DEMO
 

### 2017-04-24 SLE_17/sections/Discussion.tex
Revise code (up to section 4) to fit the narrower column

@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=31-59]{code/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=31-59]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 \caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
@@
 
-\lstinputlisting[linerange=63-66]{code/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=63-66]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 The code works as expected to parse and pretty-print the result.
@@
 
-\lstinputlisting[linerange=76-81]{code/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=76-81]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
 
@@
 
-\lstinputlisting[linerange=85-90]{code/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\lstinputlisting[linerange=85-90]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
 
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=113-132]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
-\lstinputlisting[linerange=136-156]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\lstinputlisting[linerange=113-132]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=136-156]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
 \caption{Language components \lstinline{VarExpr} (\huang{todo}left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=160-170]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=160-170]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 \caption{Reuse two language components \lstinline{VarExpr} and \lstinline{TypedLam}.}\label{fig:compose-components}

### 2017-04-24 SLE_17/sections/Inheritance.tex
Revise code (up to section 4) to fit the narrower column

@@
 
-\lstinputlisting[linerange=29-31]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
+\lstinputlisting[linerange=7-9]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
 
@@
 
-\lstinputlisting[linerange=35-38]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
+\lstinputlisting[linerange=13-16]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
 
@@
 
-\lstinputlisting[linerange=42-47]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
+\lstinputlisting[linerange=20-26]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
 
-Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
+Now \inlinecode{VarParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
 method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
@@
 
-\lstinputlisting[linerange=67-73]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
+\lstinputlisting[linerange=46-50]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
 
@@
 
-\lstinputlisting[linerange=79-95]{code/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
+\lstinputlisting[linerange=56-73]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
 

### 2017-04-24 SLE_17/sections/ObjectAlgebras.tex
Revise code (up to section 4) to fit the narrower column

@@
 
-\lstinputlisting[linerange=23-23]{code/Sec4OA/Code1.scala}% APPLY:linerange=FVARSEXPR
+\lstinputlisting[linerange=23-23]{code/src/papercode/Sec4OA/Code1.scala}% APPLY:linerange=FVARSEXPR
 
@@
 in many cases). More importantly, a second problem is that
-\begin{comment}
-But then programmers have to define a new class for each construct by
-extending the old class with \lstinline{FVarsExpr}, and implementing
-the corresponding \lstinline{fVars} method:
-
-\lstinputlisting[linerange=27-37]{code/Sec4OA/Code1.scala}% APPLY:linerange=FVAR_CLASSES
-
-Besides introducing boilerplate code, more classes have to be defined
-along with future extensions. They pollutes the namespace heavily.
-More importantly, 
-\end{comment}
-is that even if that approach would work, the parsing code would not
+even if that approach would work, the parsing code would not
 be reusable.
-\lstinline{VarExprParser} is no longer suitable for parsing the new ASTs!
-All the methods in the previous parser have type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{FVarsExpr} now.
+\lstinline{VarParser} is no longer suitable for parsing the new ASTs!
+All the methods in the previous parser have type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{NewExpr} now.
 
 Therefore, in order to generate new AST structures, we have to
-substitute \lstinline{FVarsExpr}, \lstinline{FVarsLit},
-\lstinline{FVarsAdd} and \lstinline{FVarsVar} for the old
+substitute \lstinline{NewExpr}, \lstinline{NewLit},
+\lstinline{NewAdd} and \lstinline{NewVar} for the old
 \lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var}
@@
 use casts. For semantic modularity, we need a different approach for building ASTs.
-
-%\bruno{The following text is not very good, needs to be revised. I think
-%we need to show the full code and where it breaks. Since I removed text
-%about the Expression problem in S3, I think we need to talk more about it %here.}
-%It refers to one Scala solution to the Expression Problem.\huang{what's the %solution?
-does it have the later "critical issue"?
-if so, why is it a %"solution"} But then programmers have to define new classes for literals, %additions and variables
-%by extending the old ones together with \inlinecode{FreeVarsExpr}. The %critical issue is that, the existing parsing code only generates objects of %the old
-%types, again it is a modularity issue, since it is a bad idea to modify %existing code again and again with extensions.
-Other solutions including %modelling constructors as functions introduce unnecessary complexity.
 
@@
 
-\lstinputlisting[linerange=5-8]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
+\lstinputlisting[linerange=5-8]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
 We have two constructors \inlinecode{lit} and \inlinecode{add}, corresponding to literals and additions.
-The trait \inlinecode{ExprAlg} is called an \textit{Object Algebra interface}.
+The trait \inlinecode{Alg} is called an \textit{Object Algebra interface}.
 The trait is parametrized by the type
@@
 
-\lstinputlisting[linerange=12-15]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
+\lstinputlisting[linerange=12-16]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
 Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
@@
 
-\lstinputlisting[linerange=19-22]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
+\lstinputlisting[linerange=20-23]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
 \paragraph{Adding New AST Constructs}
-Furthermore, new language constructs can be added by extending \inlinecode{ExprAlg} and adding new cases only. Now we extend the language
+Furthermore, new language constructs can be added by extending \inlinecode{Alg} and adding new cases only. Now we extend the language
 with variables. A new Object Algebra interface \inlinecode{VarAlg} is defined as follows:
 
-\lstinputlisting[linerange=26-28]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+\lstinputlisting[linerange=27-29]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
 Now pretty-printing on the new language can be realized without modifying existing code:
 
-\lstinputlisting[linerange=32-34]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+\lstinputlisting[linerange=33-35]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
 An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
@@
 
-\lstinputlisting[linerange=41-41]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+\lstinputlisting[linerange=42-43]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
 
@@
 To pretty-print the expression,
-we pass an instance of \inlinecode{VarExprPrint} to \inlinecode{makeExp}:
+we pass an instance of \inlinecode{VarPrint} to \inlinecode{makeExp}:
 
 \begin{lstlisting}
-makeExp(new VarExprPrint {})
+makeExp(new VarPrint {})
 \end{lstlisting}
@@
 
-\begin{comment}
-\inlinecode{ExprAlg} is defined for a small language using Object Algebras.
-Since Object Algebras represent ``objects'' implicitly as functions like \inlinecode{ExprAlg[E] =>} \inlinecode{E} for abstract \inlinecode{E}, we first try to build the corresponding parser by type \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.
-\end{comment}
-
-\lstinputlisting[linerange=8-16]{code/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
-Such a parser works fine, but it is not extensible!
+\lstinputlisting[linerange=8-17]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+Such a parser looks fine, but it is not extensible.
 For example, we have demonstrated in Section~\ref{subsec:overriding} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
-However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}, where the type of the \emph{extended} Object Algebra interface appears in \emph{contravariant} position.
+However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]}, where the type of the \emph{extended} Object Algebra interface appears in \emph{contravariant} position.
 This difference prohibits method/field overriding.
@@
 field has a type which is a \emph{subtype} of the original field.
-Unfortunately, in this case, although \inlinecode{VarExprAlg} is a subtype
-of \inlinecode{ExprAlg}, the function type \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} is a supertype of \inlinecode{ExprAlg[E] =>} \inlinecode{Parser[E]}.
-Thus it is not possible to have overriding.
-
-% because Object Algebra interfaces extended from \inlinecode{VarExprAlg} %are subtypes of it, but the types of their parsers are supertypes of %\inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]}.
+Unfortunately, in this case, although \inlinecode{VarAlg} is a subtype
+of \inlinecode{Alg}, the function type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]} is a supertype of \inlinecode{Alg[E] =>} \inlinecode{Parser[E]}. Thus it is not possible to have overriding.
 
@@
 A solution to this problem is instead having a field with the Object Algebra
-in \lstinline{ExprOAParser}. This approach supports overriding as well.
+in \lstinline{OAParser}. This approach supports overriding as well.
 The parsing code is as follows:
 
-\lstinputlisting[linerange=23-30]{code/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
+\lstinputlisting[linerange=21-30]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
 This is precisely the pattern that we advocate for modular parsing.
@@
 
-\lstinputlisting[linerange=34-39]{code/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+\lstinputlisting[linerange=34-39]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
 
 \noindent The type of the Object Algebra field \lstinline{alg} is first refined
-to \lstinline{VarExpAlg[E]}, to allow calling the additional factory method
-for variables. Note that, unlike the previous attempt, \lstinline{VarExpAlg[E]}
-is a subtype of \lstinline{ExpAlg[E]}, thus the type-refinement of the
+to \lstinline{VarAlg[E]}, to allow calling the additional factory method
+for variables. Note that, unlike the previous attempt, \lstinline{VarAlg[E]}
+is a subtype of \lstinline{Alg[E]}, thus the type-refinement of the
 field is allowed. Now, the code for parsing variables (\lstinline{pVar}) can
@@
 
-\lstinputlisting[linerange=43-46]{code/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
+\lstinputlisting[linerange=43-46]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
 
 In the client code shown above, we first need to choose a concrete
-Object Algebra to initialize the \lstinline{alg} field in \lstinline{VarExprOAParser}. Here we pick the
-pretty-printing algebra \lstinline{VarExprPrint}, but any other Object
-Algebra that implements \lstinline{VarExprAlg} would work.
-With an instance of \lstinline{VarExprOAParser} in hand, we can call
+Object Algebra to initialize the \lstinline{alg} field in \lstinline{VarOAParser}. Here we pick the
+pretty-printing algebra \lstinline{VarPrint}, but any other Object
+Algebra that implements \lstinline{VarAlg} would work.
+With an instance of \lstinline{VarOAParser} in hand, we can call
 \lstinline{pE} to obtain the parser to feed to the \lstinline{parse} method.
@@
 Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance} we omit that code here.
-
-\begin{comment}
-\subsection{ASTs and Object Algebras}\label{subsec:more-oa}
-
-In Section~\ref{subsec:objectalgebras}, \lstinline{makeExp} represents
-the AST of expression \inlinecode{1 + x} as a method. Parsers with
-Object Algebras generate such ``implicit objects'', which are actually
-functions of type \lstinline{Alg[E] => } \lstinline{E} for generic
-\lstinline{E}. Comparing with traditional objects, one may have worry
-that if a parser produces function structures, it could be hard to
-further process them. Especially, in the design of a compiler there
-will be a lot of desugarings (or AST transformations). As previous
-work on Object Alegbras has shown~\cite{Gouseti2014,Zhang2015}, AST
-transformations can be implemented using \emph{transformation
-  algebras}. Continuing with the previous example, suppose we want to
-transform an expression of \inlinecode{VarExprAlg} to an expression of
-\inlinecode{ExprAlg}, by replacing some variables with values based on
-a variable environment. A \textit{transformation} algebra is
-implemented as follows:
-
-\lstinputlisting[linerange=45-56]{code/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_REFACTOR
-Here the \inlinecode{VarExprAlg} expression \inlinecode{1 + x} is changed to \inlinecode{ExprAlg} expression \inlinecode{1 + 2}, before applying the pretty-printing algebra. The abstract field
-\inlinecode{alg} in \inlinecode{Refactor} could be any algebra simply for delegation.
-With an algebra instance of \inlinecode{alg}, \inlinecode{Refactor} is able to transform a \inlinecode{VarExprAlg} expression before producing the result.
-Furthermore, transformations can be composed in a linear pipeline-style beforehand~\cite{Zhang2015}.
-
-Another useful technique is the composition of \textit{query} algebras.
-A query traverses the abstract syntax tree, collects information and returns the result.
-The aforementioned \lstinline{Print} and \lstinline{Eval} algebras are traditional queries.
-But as we discussed before, an algebra is first fed to the parser, then \lstinline{parse} is invoked to do the parsing.
-It is tedious to fed \lstinline{Print} and \lstinline{Eval} separately, as the same input is parsed twice.
-Instead, we can firstly compose two queries into one~\cite{Oliveira:2012,oliveira2013feature}, before applying it for parsing.
-\end{comment}

### 2017-04-24 SLE_17/sections/PackratParsing.tex
Revise code (up to section 4) to fit the narrower column

@@
 \centering
-\lstinputlisting[linerange=42-53]{code/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_PAPERCODE
+\lstinputlisting[linerange=16-27]{code/src/papercode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_PAPERCODE
 \caption{Helper object for code demonstration in this paper.}\label{fig:papercode}
@@
 
-\lstinputlisting[linerange=9-17]{code/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
+\lstinputlisting[linerange=9-17]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
 
@@
 
-\lstinputlisting[linerange=21-29]{code/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
+\lstinputlisting[linerange=21-28]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
 
-In the trait \lstinline{ExprParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} parses an integer for the literal case.
+In the trait \lstinline{AParser}, \lstinline{lexical} is used for lexing. \lstinline{pLit} parses an integer for the literal case.
 \lstinline{pAdd} handles the addition case and creates an object of \lstinline{Add}. It parses two sub-expressions by calling \lstinline{pExpr}
@@
 
-The last line demonstrates how we parse a valid expression \inlinecode{1 + 2} using our parser.
+The code below demonstrates how to parse a valid expression \inlinecode{1 + 2} using our parser.
 
-\lstinputlisting[linerange=33-33]{code/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+\lstinputlisting[linerange=6-7]{code/src/papercode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER

### 2017-04-24 SLE_17/code/Sec2Packrat/Code1.scala

@@ -1,31 +1,0 @@
-package PaperCode.Sec2Packrat
-
-import PaperCode.Common
-
-
-object Code1 extends Common {
-
-//BEGIN_PACKRAT_OO_AST
-trait Expr {
-  def print: String
-}
-class Lit(x: Int) extends Expr {
-  def print = x.toString
-}
-class Add(e1: Expr, e2: Expr) extends Expr {
-  def print = "(" + e1.print + " + " + e2.print + ")"
-}
-//END_PACKRAT_OO_AST
-
-//BEGIN_PACKRAT_SIMPLE_EXPR
-trait ExprParser {
-  lexical.delimiters += "+"
-  val pLit: Parser[Expr] =
-    numericLit ^^ { x => new Lit(x.toInt) }
-  val pAdd: Parser[Expr] =
-    pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-  def pExpr: Parser[Expr] =
-    pLit ||| pAdd
-}
-//END_PACKRAT_SIMPLE_EXPR
-}

### 2017-04-24 SLE_17/code/Sec2Packrat/Code2.scala

@@ -1,54 +1,0 @@
-package PaperCode.Sec2Packrat
-
-import PaperCode.Common
-
-
-object Code2 extends Common {
-
-  // Abstract syntax
-  trait Expr {
-    def print: String
-  }
-
-  class Lit(x: Int) extends Expr {
-    def print = x.toString
-  }
-
-  class Add(e1: Expr, e2: Expr) extends Expr {
-    def print = "(" + e1.print + " + " + e2.print + ")"
-  }
-
-  // Parser with lexing
-  trait ExprParser {
-    lexical.delimiters += "+"
-    val pLit: PackratParser[Expr] =
-      numericLit ^^ { x => new Lit(x.toInt) }
-    val pAdd: PackratParser[Expr] =
-      pExpr ~ ("+" ~> pExpr) ^^ { case e1 ~ e2 => new Add(e1, e2) }
-    def pExpr: PackratParser[Expr] =
-      pLit ||| pAdd
-  }
-
-//BEGIN_PACKRAT_RUNPARSER
-val r = parse(new ExprParser {}.pExpr)("1 + 2").print // "(1 + 2)"
-//END_PACKRAT_RUNPARSER
-
-  def main(args: Array[String]): Unit = {
-    println(r)
-  }
-}
-
-//BEGIN_PACKRAT_PAPERCODE
-import scala.util.parsing.combinator.syntactical.StandardTokenParsers
-import scala.util.parsing.combinator.PackratParsers
-
-object Code extends StandardTokenParsers with PackratParsers {
-  type Parser[E] = PackratParser[E]
-
-  def parse[E](p: Parser[E]): String => E = in => {
-    val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) t.get else scala.sys.error(t.toString)
-  }
-  // Any Scala code in the paper comes here
-}
-//END_PACKRAT_PAPERCODE

### 2017-04-24 SLE_17/code/Sec4OA/Code1.scala

@@ -1,40 +1,0 @@
-package PaperCode.Sec4OA
-
-
-object Code1 {
-
-  trait Expr {
-    def print: String
-  }
-
-  class Lit(x: Int) extends Expr {
-    def print = x.toString
-  }
-
-  class Add(e1: Expr, e2: Expr) extends Expr {
-    def print = "(" + e1.print + " + " + e2.print + ")"
-  }
-
-  class Var(x: String) extends Expr {
-    def print = x
-  }
-
-//BEGIN_FVARSEXPR
-trait FVarsExpr extends Expr { def fVars: Set[String] }
-//END_FVARSEXPR
-
-//BEGIN_FVAR_CLASSES
-class FVarsLit(x: Int) extends Lit(x) with FVarsExpr {
-  def fVars = Set()
-}
-
-class FVarsAdd(e1: FVarsExpr, e2: FVarsExpr) extends Add(e1, e2) with FVarsExpr {
-  def fVars = e1.fVars ++ e2.fVars
-}
-
-class FVarsVar(x: String) extends Var(x) with FVarsExpr {
-  def fVars = Set(x)
-}
-//END_FVAR_CLASSES
-
-}

### 2017-04-24 SLE_17/code/Sec4OA/Code3.scala

@@ -1,52 +1,0 @@
-package PaperCode.Sec4OA
-
-import PaperCode.Common
-
-
-object Code30 extends Common {
-//BEGIN_BASE_OA_PARSER_BAD
-trait ExprOAParser[E] {
-  lexical.delimiters += "+"
-  val pLit: ExprAlg[E] => Parser[E] = alg => numericLit ^^ { x => alg.lit(x.toInt) }
-  val pAdd: ExprAlg[E] => Parser[E] = alg => {
-    val p = pExpr(alg)
-    p ~ ("+" ~> p) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
-  }
-  val pExpr: ExprAlg[E] => Parser[E] = alg => pLit(alg) ||| pAdd(alg)
-}
-//END_BASE_OA_PARSER_BAD
-}
-
-
-object Code3 extends Common {
-//BEGIN_BASE_OA_PARSER
-trait ExprOAParser[E] {
-  lexical.delimiters += "+"
-  val alg: ExprAlg[E]
-  val pLit: Parser[E] = numericLit ^^ { x => alg.lit(x.toInt) }
-  val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
-  val pExpr: Parser[E] = pLit ||| pAdd
-  val pE: Parser[E] = pExpr
-}
-//END_BASE_OA_PARSER
-
-//BEGIN_EXT_OA_PARSER
-trait VarExprOAParser[E] extends ExprOAParser[E] {
-  override val alg: VarExprAlg[E]
-  val pVar: Parser[E] = ident ^^ alg.varE
-  val pVarExpr: Parser[E] = pExpr ||| pVar
-  override val pE: Parser[E] = pVarExpr
-}
-//END_EXT_OA_PARSER
-
-//BEGIN_EXT_OA_PARSER_CLIENT
-val p = new VarExprOAParser[String] {
-  override val alg = new VarExprPrint {}
-}
-val r = parse(p.pE)("1 + x") // "(1 + x)"
-//END_EXT_OA_PARSER_CLIENT
-
-  def main(args: Array[String]): Unit = {
-    println(r)
-  }
-}

### 2017-04-24 SLE_17/code/src/papercode/Common.scala

@@ -1,19 +1,17 @@
-package PaperCode
+package papercode
 
-import scala.util.parsing.combinator.syntactical._
-import scala.util.parsing.combinator._
+import util.parsing.combinator.PackratParsers
+import util.parsing.combinator.syntactical.StandardTokenParsers
 
 
 trait Common extends StandardTokenParsers with PackratParsers {
-
-  type Parser[E] = PackratParser[E]
-
   val lcid: Parser[String] = ident ^? { case id if id.charAt(0).isLower => id }
   val ucid: Parser[String] = ident ^? { case id if id.charAt(0).isUpper => id }
 
+  type Parser[E] = PackratParser[E]
+
   def parse[E](p: Parser[E]): String => E = in => {
     val t = phrase(p)(new lexical.Scanner(in))
-    if (t.successful) t.get else scala.sys.error(t.toString)
+    t.getOrElse(sys.error(t.toString))
   }
-
 }

### 2017-04-24 SLE_17/code/src/papercode/Sec2Packrat/Code1.scala

@@ -1,0 +1,30 @@
+package papercode.Sec2Packrat
+
+import papercode.Common
+
+
+trait Code1 extends Common {
+
+//BEGIN_PACKRAT_OO_AST
+trait Expr {
+  def print: String
+}
+class Lit(x: Int) extends Expr {
+  def print = x.toString
+}
+class Add(e1: Expr, e2: Expr) extends Expr {
+  def print = "(" + e1.print + "+" + e2.print + ")"
+}
+//END_PACKRAT_OO_AST
+
+//BEGIN_PACKRAT_SIMPLE_EXPR
+trait AParser {
+  lexical.delimiters += "+"
+  val pLit: Parser[Expr] = numericLit ^^
+    { x => new Lit(x.toInt) }
+  val pAdd: Parser[Expr] = pExpr ~ ("+" ~> pExpr) ^^
+    { case e1 ~ e2 => new Add(e1, e2) }
+  def pExpr: Parser[Expr] = pLit ||| pAdd
+}
+//END_PACKRAT_SIMPLE_EXPR
+}

### 2017-04-24 SLE_17/code/src/papercode/Sec4OA/Code1.scala

@@ -1,0 +1,26 @@
+package papercode.Sec4OA
+
+
+object Code1 {
+
+  trait Expr {
+    def print: String
+  }
+
+  class Lit(x: Int) extends Expr {
+    def print = x.toString
+  }
+
+  class Add(e1: Expr, e2: Expr) extends Expr {
+    def print = "(" + e1.print + " + " + e2.print + ")"
+  }
+
+  class Var(x: String) extends Expr {
+    def print = x
+  }
+
+//BEGIN_FVARSEXPR
+trait NewExpr extends Expr { def free: Set[String] }
+//END_FVARSEXPR
+
+}

### 2017-04-24 SLE_17/code/src/papercode/Sec4OA/Code2.scala

@@ -1,35 +1,36 @@
-package PaperCode.Sec4OA
+package papercode.Sec4OA
 
 
 //BEGIN_OVERVIEW_OA_ALG
-trait ExprAlg[E] {
+trait Alg[E] {
   def lit(n: Int): E
   def add(e1: E, e2: E): E
 }
 //END_OVERVIEW_OA_ALG
 
 //BEGIN_OVERVIEW_OA_PRINT
-trait Print extends ExprAlg[String] {
+trait Print extends Alg[String] {
   def lit(n: Int) = n.toString
-  def add(e1: String, e2: String) = "(" + e1 + " + " + e2 + ")"
+  def add(e1: String, e2: String) =
+    "(" + e1 + " + " + e2 + ")"
 }
 //END_OVERVIEW_OA_PRINT
 
 //BEGIN_OVERVIEW_OA_EVAL
-trait Eval extends ExprAlg[Int] {
+trait Eval extends Alg[Int] {
   def lit(n: Int) = n
   def add(e1: Int, e2: Int) = e1 + e2
 }
 //END_OVERVIEW_OA_EVAL
 
 //BEGIN_OVERVIEW_OA_ALGEXT
-trait VarExprAlg[E] extends ExprAlg[E] {
+trait VarAlg[E] extends Alg[E] {
   def varE(x: String): E
 }
 //END_OVERVIEW_OA_ALGEXT
 
 //BEGIN_OVERVIEW_OA_EXTPRINT
-trait VarExprPrint extends VarExprAlg[String] with Print {
+trait VarPrint extends VarAlg[String] with Print {
   def varE(x: String) = x
 }
 //END_OVERVIEW_OA_EXTPRINT
@@ -38,12 +39,13 @@ trait VarExprPrint extends VarExprAlg[String] with Print {
 object Code2 {
 
 //BEGIN_OVERVIEW_OA_MAKEEXP
-def makeExp[E](alg: VarExprAlg[E]): E = alg.add(alg.lit(1), alg.varE("x"))
+def makeExp[E](alg: VarAlg[E]): E =
+  alg.add(alg.lit(1), alg.varE("x"))
 //END_OVERVIEW_OA_MAKEEXP
 
 //BEGIN_OVERVIEW_OA_REFACTOR
-trait Refactor[E] extends VarExprAlg[E] {
-  val alg: ExprAlg[E]
+trait Refactor[E] extends VarAlg[E] {
+  val alg: Alg[E]
   val env: Map[String, Int]
[…]

### 2017-04-24 SLE_17/code/src/papercode/Sec4OA/Code3.scala

@@ -1,0 +1,52 @@
+package papercode.Sec4OA
+
+import papercode.Common
+
+
+object Code3 extends Common {
+//BEGIN_BASE_OA_PARSER_BAD
+trait Attempt[E] {
+  lexical.delimiters += "+"
+  val pLit: Alg[E] => Parser[E] = alg =>
+    numericLit ^^ { x => alg.lit(x.toInt) }
+  val pAdd: Alg[E] => Parser[E] = alg =>
+    pExpr(alg) ~ ("+" ~> pExpr(alg)) ^^
+      { case e1 ~ e2 => alg.add(e1, e2) }
+  val pExpr: Alg[E] => Parser[E] = alg =>
+    pLit(alg) ||| pAdd(alg)
+}
+//END_BASE_OA_PARSER_BAD
+
+//BEGIN_BASE_OA_PARSER
+trait OAParser[E] {
+  lexical.delimiters += "+"
+  val alg: Alg[E]
+  val pLit: Parser[E] = numericLit ^^
+    { x => alg.lit(x.toInt) }
+  val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^
+    { case e1 ~ e2 => alg.add(e1, e2) }
+  val pExpr: Parser[E] = pLit ||| pAdd
+  val pE: Parser[E] = pExpr
+}
+//END_BASE_OA_PARSER
+
+//BEGIN_EXT_OA_PARSER
+trait VarOAParser[E] extends OAParser[E] {
+  override val alg: VarAlg[E]
+  val pVar: Parser[E] = ident ^^ alg.varE
+  val pVarExpr: Parser[E] = pExpr ||| pVar
+  override val pE: Parser[E] = pVarExpr
+}
+//END_EXT_OA_PARSER
+
+//BEGIN_EXT_OA_PARSER_CLIENT
+val p = new VarOAParser[String] {
+  override val alg = new VarPrint {}
+}
+val r = parse(p.pE)("1 + x") // "(1 + x)"
+//END_EXT_OA_PARSER_CLIENT
+
+  def main(args: Array[String]): Unit = {
+    println(r)
+  }
+}

### 2017-04-24 SLE_17/code/src/papercode/Sec6CaseStudy/Code1.scala

@@ -1,6 +1,6 @@
-package PaperCode.Sec6CaseStudy
+package papercode.Sec6CaseStudy
 
-import PaperCode.Common
+import papercode.Common
 
 
 object Code1 extends Common {

### 2017-04-24 SLE_17/sections/CaseStudy.tex
Format

--- 2017-04-24 SLE_17/sections/CaseStudy.tex
@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=15-33]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
+\lstinputlisting[linerange=13-29]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
 \caption{The \inlinecode{VarApp} component in case study.}\label{fig:casestudy-varapp}
@@
 
-\lstinputlisting[linerange=9-11]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
+\lstinputlisting[linerange=9-9]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
 
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=37-64]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
+\lstinputlisting[linerange=33-60]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
 \caption{Build the \inlinecode{Untyped} calculus by composing to language components.}\label{fig:casestudy-untyped}
 \end{figure}
-
-\paragraph{Demo code}
-We have demo code for every calculus in order to test and evaluate our implementation.
-We feed the pretty-printing algebra to the parser and print out the result.
-The parsing result can be changed as long as the concrete algebra is available.
-
-\lstinputlisting[linerange=68-78]{code/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_DEMO
 

### 2017-04-24 SLE_17/sections/Discussion.tex
Format

--- 2017-04-24 SLE_17/sections/Discussion.tex
@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=31-59]{code/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=7-33]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 \caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
@@
 
-\lstinputlisting[linerange=63-66]{code/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
+\lstinputlisting[linerange=40-43]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
 The code works as expected to parse and pretty-print the result.
@@
 
-\lstinputlisting[linerange=76-81]{code/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+\lstinputlisting[linerange=76-82]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
 
@@
 
-\lstinputlisting[linerange=85-90]{code/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\lstinputlisting[linerange=86-92]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
 
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=113-132]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
-\lstinputlisting[linerange=136-156]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\lstinputlisting[linerange=115-130]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=134-150]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
 \caption{Language components \lstinline{VarExpr} (\huang{todo}left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
@@
 \begin{figure}[t]
-\lstinputlisting[linerange=160-170]{code/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=154-165]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 \caption{Reuse two language components \lstinline{VarExpr} and \lstinline{TypedLam}.}\label{fig:compose-components}

### 2017-04-24 SLE_17/code/src/papercode/Sec4OA/Code3.scala

@@ -3,7 +3,7 @@ package papercode.Sec4OA
 import papercode.Common
 
 
-object Code3 extends Common {
+trait Code3 extends Common {
 //BEGIN_BASE_OA_PARSER_BAD
 trait Attempt[E] {
   lexical.delimiters += "+"
@@ -38,7 +38,10 @@ trait VarOAParser[E] extends OAParser[E] {
   override val pE: Parser[E] = pVarExpr
 }
 //END_EXT_OA_PARSER
+}
 
+
+object Code30 extends Code3 {
 //BEGIN_EXT_OA_PARSER_CLIENT
 val p = new VarOAParser[String] {
   override val alg = new VarPrint {}

### 2017-04-24 SLE_17/code/src/papercode/Sec4OA/Code4.scala

@@ -1,69 +1,46 @@
 package papercode.Sec4OA
 
-import papercode.Common
 
-
-object Code4 extends Common {
-
-  trait ExprOAParser[E] {
-    lexical.delimiters += "+"
-
-    val alg: Alg[E]
-
-    val pLit: Parser[E] = numericLit ^^ { x => alg.lit(x.toInt) }
-    val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^ { case e1 ~ e2 => alg.add(e1, e2) }
-    val pExpr: Parser[E] = pLit ||| pAdd
-
-    val pE: Parser[E] = pExpr
-  }
-
-  trait VarExprOAParser[E] extends ExprOAParser[E] {
-    override val alg: VarAlg[E]
-
-    val pVar: Parser[E] = ident ^^ alg.varE
-    val pVarExpr: Parser[E] = pExpr ||| pVar
-
-    override val pE = pVarExpr
-  }
-
+trait Code4 extends Code3 {
 
 //BEGIN_OVERVIEW_OA_MULTI_SYNTAX
-trait TypedLamAlg[E, T] extends VarAlg[E] {
+trait LamAlg[E, T] extends VarAlg[E] {
   def intT(): T
   def arrT(t1: T, t2: T): T
   def lam(x: String, t: T, e: E): E
 }
-
-trait TypedLamPrint extends TypedLamAlg[String, String] with VarPrint {
+trait LamPrint extends LamAlg[String, String]
+  with VarPrint {
   def intT() = "int"
-  def arrT(t1: String, t2: String) = t1 + " -> " + t2
-  def lam(x: String, t: String, e: String) = "\\" + x + " : " + t + ". " + e
+  def arrT(t1: String, t2: String) = t1 + "->" + t2
+  def lam(x: String, t: String, e: String) =
+    "\\" + x + " : " + t + ". " + e
 }
-
-trait TypedLamOAParser[E, T] extends VarExprOAParser[E] {
+trait LamOAParser[E, T] extends VarOAParser[E] {
   lexical.reserved += "int"
   lexical.delimiters += ("->", "\\", ":", ".")
-
-  override val alg: TypedLamAlg[E, T]
-
+  override val alg: LamAlg[E, T]
   val pIntT: Parser[T] = "int" ^^^ alg.intT
[…]

### 2017-04-24 SLE_17/code/src/papercode/Sec5/Code1.scala

@@ -76,7 +76,8 @@ object Code1 extends Common {
 trait BaseParser[E] {
   ...
   val alg: ...
-  val pLam: Parser[E] = ("\\" ~> ident) ~ ("." ~> pE) ^^ ...
+  val pLam: Parser[E] =
+    ("\\" ~> ident) ~ ("." ~> pE) ^^ ...
   val pE: Parser[E] = ...
 }
 //END_BASEPARSER_UNTYPEDLAM
@@ -85,7 +86,8 @@ trait BaseParser[E] {
 trait ExtParser[E, T] extends BaseParser[E] {
   ...
   override val alg: ...
-  override val pLam: Parser[E] = ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ ...
+  override val pLam: Parser[E] =
+    ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ ..
   val pT: Parser[T] = ...
 }
 //END_EXTPARSER_TYPEDLAM
@@ -111,22 +113,18 @@ object Code2 extends Common {
 /*
 //BEGIN_LANGUAGE_COMPONENTS_VAREXPR
 object VarExpr {
-  // Abstract syntax
-  trait Alg[E] {
+  trait Alg[E] {    // Abstract syntax
     def lit(n: Int): E
     def add(e1: E, e2: E): E
     def varE(x: String): E
   }
-
-  // Parser
-  trait Parse[E] {
+  trait Parse[E] {   // Parser
     val alg: Alg[E]
     val pE: Parser[E] = ...
     ...
   }
-
-  // Pretty-printer
   trait Print extends Alg[String] {
+    // Pretty-printer
     ...
   }
 }
@@ -134,23 +132,19 @@ object VarExpr {
 
 //BEGIN_LANGUAGE_COMPONENTS_TYPEDLAM
 object TypedLam {
-  // Abstract syntax
-  trait Alg[E, T] {
+  trait Alg[E, T] {    // Abstract syntax
     def intT(): T
     def arrT(t1: T, t2: T): T
     def lam(x: String, t: T, e: E): E
   }
-
-  // Parser
-  trait Parse[E, T] {
+  trait Parse[E, T] {   // Parser
     val alg: Alg[E, T]
     val pE: Parser[E] = ...
     val pT: Parser[T] = ...
     ...
   }
-
-  // Pretty-printer
   trait Print extends Alg[String, String] {
[…]

### 2017-04-24 SLE_17/code/src/papercode/Sec6CaseStudy/Code1.scala

@@ -6,9 +6,7 @@ import papercode.Common
 object Code1 extends Common {
 
 //BEGIN_CASESTUDY_EPARSER
-trait EParser[E] {
-  val pE: Parser[E]
-}
+trait EParser[E] { val pE: Parser[E] }
 //END_CASESTUDY_EPARSER
 
 //BEGIN_CASESTUDY_VARAPP
@@ -17,18 +15,16 @@ object VarApp {
     def TmVar(x: String): E
     def TmApp(e1: E, e2: E): E
   }
-
   trait Print extends Alg[String] {
     def TmVar(x: String): String = x
-    def TmApp(e1: String, e2: String): String = "[" + e1 + " " + e2 + "]"
+    def TmApp(e1: String, e2: String): String =
+      "(" + e1 + " " + e2 + ")"
   }
-
   trait Parse[E] extends EParser[E] {
     lexical.delimiters += ("(", ")")
     val alg: Alg[E]
     val pVarAppE: Parser[E] = lcid ^^ alg.TmVar |||
-      pE ~ pE ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
-      "(" ~> pE <~ ")"
+      pE ~ pE ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } ||| "(" ~> pE <~ ")"
   }
 }
 //END_CASESTUDY_VARAPP
@@ -38,25 +34,25 @@ object UntypedAbs {
   trait Alg[E] {
     def TmAbs(x: String, e: E): E
   }
-
   trait Print extends Alg[String] {
-    def TmAbs(x: String, e: String): String = "\\" + x + "." + e
+    def TmAbs(x: String, e: String): String =
+      "\\" + x + "." + e
   }
-
   trait Parse[E] extends EParser[E] {
     lexical.delimiters += ("\\", ".")
     val alg: Alg[E]
     val pUntypedAbsE: Parser[E] =
-      ("\\" ~> lcid) ~ ("." ~> pE) ^^ { case x ~ e0 => alg.TmAbs(x, e0) }
+      ("\\" ~> lcid) ~ ("." ~> pE) ^^
+        { case x ~ e0 => alg.TmAbs(x, e0) }
   }
 }
-
 object Untyped {
-  trait Alg[E] extends UntypedAbs.Alg[E] with VarApp.Alg[E]
-
-  trait Print extends Alg[String] with UntypedAbs.Print with VarApp.Print
-
[…]

### 2017-04-27 SLE_17/sections/CaseStudy.tex
Some changes

@@
 example, the \lstinline{VarApp} component below represents variables and
-function applications, shown in Figure~\ref{fig:casestudy-varapp}.
+function applications.
 
-\begin{figure}[ht]
 \lstinputlisting[linerange=13-29]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
-\caption{The \inlinecode{VarApp} component in case study.}\label{fig:casestudy-varapp}
-\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \label{fig:dependency}
-\end{figure}
+\end{figure*}
 
@@
 created from scratch, while others such as \lstinline{Typed} are
-extended from existing components. Since calculi and components have the similar signature, each calculus
+extended from existing components. Since calculi and components have similar signatures, each calculus
 can also be extended and reused directly. For example, calculus \lstinline{FullRef} extends from
@@
 lots of calculi. Such reuse could shorten the code considerably. We
-will show this advantage and exam the possible performance penalty in
+will show this advantage and inspect the possible performance penalty in
 a quantitative way in the next subsection.
@@
 
-\begin{table}
+\begin{table*}
     \centering
@@
     \label{tab:comparison}
-\end{table}
+\end{table*}
 
@@
     \label{tab:ext-comparison}
-\end{table}
+\end{table*}
 

### 2017-04-27 SLE_17/sections/Discussion.tex
Some changes

@@
 
-We take the example in Section~\ref{subsec:parsingwithoa} again, which is a language of literals, additions and variables. It is defined as a
-language component on the left side of Figure~\ref{fig:lng-components}.
-Then for the extension of types and lambda abstractions in Section~\ref{subsec:differentsyntax}, instead of inheriting from the previous language directly, we define it as another independent language component.
-The code is presented on the right side of Figure~\ref{fig:lng-components}.
+We take the example in Section~\ref{subsec:parsingwithoa} again, which is a language of literals, additions and variables. It can be defined as a language component \inlinecode{VarExpr}.
 
-\begin{figure}[t]
 \lstinputlisting[linerange=115-130]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+
+For the extension of types and lambda abstractions in Section~\ref{subsec:differentsyntax}, instead of inheriting from the previous language directly, we can define it as another independent language component \inlinecode{TypedLam}.
+
 \lstinputlisting[linerange=134-150]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
-\caption{Language components \lstinline{VarExpr} (\huang{todo}left) and \lstinline{TypedLam} (right).}\label{fig:lng-components}
-\end{figure}
 
-As shown in Figure~\ref{fig:compose-components}, those two language components can be merged into one to obtain the language we want. Furthermore, the new language is still a modular
+The code below shows how we merge those two components together to obtain the language we want. Furthermore, the new language is still a modular
 component ready for future composition. In that case modularity is realized over higher-order hierarchies.
 
-\begin{figure}[t]
 \lstinputlisting[linerange=154-165]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
-\caption{Reuse two language components \lstinline{VarExpr} and \lstinline{TypedLam}.}\label{fig:compose-components}
-\end{figure}
 

### 2017-04-27 SLE_17/sections/Introduction.tex
Code & content
Some changes

@@
 employ a syntactic modularity approach for parsing in combination with
-a semantic modularity approach for defining other language components
+a semantic modularity approach for defining operations
 that traverse or process ASTs. This is a shame because parsing is a
@@
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
 

### 2017-04-27 SLE_17/sections/PackratParsing.tex
Some changes

@@
 modular parsing and argues that Packrat parser combinators~\cite{Ford2002}
-are suitable to address those challenges. The algorithmic challenges
+are suitable to address them. The algorithmic challenges
 are important because they rule out various common techniques

### 2017-05-11 SLE_17/paper.tex
Code & content
Some changes.

@@
 \usepackage{multirow}
+\usepackage{enumitem}
+\usepackage{tabularx}
+
+\lstdefinelanguage{JavaScala}{
+  morekeywords={public,int,interface,implements,default,
+    abstract,case,catch,class,def,static,%
+    do,else,extends,false,final,finally,%
+    for,if,implicit,import,match,mixin,%
+    new,null,object,override,package,%
+    private,protected,requires,return,sealed,%
+    super,this,throw,trait,true,try,%
+    type,var,while,yield,with,val},
+  otherkeywords={=>,<\%,<:,>:,\#,@},
+  sensitive=true,
+  morecomment=[l]{//},
+  morecomment=[n]{/*}{*/},
+  morestring=[b]",
+  morestring=[b]',
+  morestring=[b]"""
+}
 
 \lstset{
-  language=Scala,                % choose the language of the code
+  language=JavaScala,                % choose the language of the code
   columns=flexible,

### 2017-05-11 SLE_17/sections/Discussion.tex
Some changes.

@@
 They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
-Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions.
+Besides lexing, the trait \inlinecode{LamOAParser} also introduces parsers for types, and the new case for expressions.
 We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue
@@
 express multiple syntax. The isolation for different syntax
-is guaranteed by the type system. The following illustrates some client code
-for the new parser:
+is guaranteed by the type system. The following client code
 
 \lstinputlisting[linerange=40-43]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-The code works as expected to parse and pretty-print the result.
+works as expected to parse and pretty-print the result.
 
@@
 As many syntactically extensible parsers, our approach also supports
-modifying part of existing parsers, but in a type-safe way. We can
-easily update existing parsing rules with new implementations, or
-eliminate them in extended parsers. This can be useful in many
-situations, including when conflicts or ambiguities arise when
-composing languages.
-As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below.
-\inlinecode{pLam} parses a lambda symbol, an identifier, a dot and an expression in sequence.
-Some irrelevant details are omitted.
+modifying part of existing parsers, including updating or eliminating existing rules,
+but in a type-safe way. This can be useful in many
+situations, for instance when conflicts or ambiguities arise upon composing languages.
+As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below. Some irrelevant details are omitted.
 
 \lstinputlisting[linerange=76-82]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
-
+Here \inlinecode{pLam} parses a lambda symbol, an identifier, a dot and an expression in sequence.
 Then we want to replace the untyped lambda abstractions by typed
@@
 change the implementation of \lstinline{pLam} in the extended parser.
-The new \inlinecode{pLam} parses a colon and a type in addition,
-between the identifier and the dot. Due to dynamic dispatch, our new
+Due to dynamic dispatch, our new
 implementation of lambdas will be different without affecting the other parts of the parser.
@@
 \begin{itemize}[leftmargin=*]
-    \item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax. The type parameters represent multiple sorts of syntax, and the methods are constructs.
-    \item \textbf{Parser:} corresponding parser of the abstract syntax, written in a modular way as we demonstrated in previous sections.
+    \item \textbf{Object Algebra interface:} defined as a trait for the abstract syntax. The type parameters represent multiple sorts of syntax, and  methods are constructs.
+    \item \textbf{Parser:} corresponding parser of the abstract syntax, written in a modular way as we demonstrated before.
     \item \textbf{Object Algebras (optional):} concrete operations on ASTs, such as pretty-printing.
@@
 Under our modular parsing framework, we use Packrat parsing as the underlying parsing technique, OO inheritance with method overriding for composing and extending parsers, and Object Algebras for parsing extensible ASTs.
-However, our framework is more general and thus more powerful, because those aspects are orthogonal to each other, and hence can have alternatives.
+However, such a pattern is more general and thus more powerful, because those aspects are orthogonal to each other, and hence can have alternatives.
 
@@
 \item {\bf Parsing Technique:}
-Our approach does not depend on a particular parser combinator library
+The pattern does not depend on a particular parser combinator library
 or parsing algorithm. We demonstrated that Packrat parsing has some
 advantages and is suitable in a modular setting. However, under
-certain circumstances or in other programming languages, it may not be
-the best choice. One can use other parser combinator libraries as long
+other circumstances it may not be
+the best choice. One can use other parser combinators as long
 as the guidelines proposed in Section~\ref{sec:packrat} are met.
@@
   We use traits to model parsers, OO inheritance to compose them, and
-  method overriding for their extensibility. These language features
+  method overriding for their extensibility. These features
   may not be supported in other programming languages, especially in
@@
   types à la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
-These alternatives could also be adopted to build ASTs and corresponding parsers.
+These alternatives could also be adopted to build ASTs and their parsers.
 
@@
 
-As we demonstrated, the whole modular parsing framework itself is
-modular, which means it can be customized easily.
+As we demonstrated, such a pattern for modular parsing is 
+modular itself, and can be customized easily.

### 2017-05-11 SLE_17/sections/Introduction.tex
Some changes.

@@
 \paragraph{Algorithmic Challenges} A first challenge was to do with
-the parsing algorithms themselves. Since many parsing formalisms
-were not designed with extensibility in mind, some of the employed
-techniques were non-modular, or performed poorly in a modular
-setting.  The most widely use tools for parsing are parser
-generators, but most algorithms employed by parser generators
-require full information about the grammar to generate parsing
+the parsing algorithms themselves, since they 
+were usually not designed with extensibility in mind. The most widely used tools for parsing are parser
+generators, but they mostly require full information about the grammar to generate parsing
 code. Moreover, actions associated with grammar productions are
@@
 for implementing modular parsers built on top of an
-existing Packrat~\cite{Ford2002} parsing library for Scala~\cite{odersky2004overview}. The Packrat parsing library
-directly supports left-recursion, and a \emph{longest-match
-  composition} operator. The direct support for left-recursion
-avoids the need for left-recursion elimination, which is very
-cumbersome is a modular setting. The longest-match composition
-operator avoids the use of priorities, and it provides reasonably
-efficient optimizations for backtracking.
+existing Packrat~\cite{Ford2002} parsing library for Scala~\cite{odersky2004overview}. Such a library
+directly supports left-recursion, memoization, and a \emph{longest-match
+  composition} operator. We will see some examples in Section \ref{subsec:packratparsing}.
 
@@
 
-All the code in the paper and the case study is written in Scala. But,
+All the code in the paper and the case study is written in Scala, since its concise and elegant
+syntax gives better presentation. Clearly
 other languages that support some form of multiple inheritance
 (including Java 8 with default methods~\cite{Goetz2012}) could in principle be
-used. Our choice for Scala had to do with the concise and elegant syntax
-of the language which enables a clear presentation of our ideas.
+used.

### 2017-05-11 SLE_17/sections/ObjectAlgebras.tex
Format
Some changes.

@@
 corresponding ASTs are built using classes with a pretty-printing
-operation. Now suppose we want to collect free variables over the
-ASTs. Using a traditional OO approach, one attempt would be extending
+operation.
+To modularly add an operation like collecting free variables, one attempt
+would be extending
 \inlinecode{Expr} with the new operation to obtain a new abstract type
@@
 even if that approach would work, the parsing code would not
-be reusable.
-\lstinline{VarParser} is no longer suitable for parsing the new ASTs!
-All the methods in the previous parser have type \inlinecode{Parser[Expr]}, which indicates the result has type \inlinecode{Expr}, whereas the type should be \inlinecode{NewExpr} now.
-
-Therefore, in order to generate new AST structures, we have to
-substitute \lstinline{NewExpr}, \lstinline{NewLit},
-\lstinline{NewAdd} and \lstinline{NewVar} for the old
-\lstinline{Expr}, \lstinline{Lit}, \lstinline{Add} and \lstinline{Var}
-correspondingly. Thus we have to either modify existing code or
-use casts. For semantic modularity, we need a different approach for building ASTs.
+be reusable. \lstinline{VarParser} is no longer suitable for parsing the new ASTs! All the methods in the previous parser have type \inlinecode{Parser[Expr]},
+which generates ASTs that only support pretty-printing. Now we have to substitute \inlinecode{NewExpr} for \inlinecode{Expr} (the
+same for \inlinecode{Lit}, \inlinecode{Add}, and so on). It requires either modification on existing code or casts.
+The goal of semantic modularity motivates us to find a different approach for building ASTs.
 
@@
 \lstinputlisting[linerange=5-8]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
-We have two constructors \inlinecode{lit} and \inlinecode{add}, corresponding to literals and additions.
-The trait \inlinecode{Alg} is called an \textit{Object Algebra interface}.
-The trait is parametrized by the type
+Here \inlinecode{Alg} is called an \textit{Object Algebra interface}, parameterized by the type
 \inlinecode{E}, which abstracts over the concrete type of the AST.
@@
 To realize an operation on expressions, we simply instantiate the type parameter by a concrete type and
-provides implementations for all cases. Below is an example of pretty-printing an expression:
+provides implementations for all cases. Below is an example of pretty-printing:
 
@@
 An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
-In other words, existing code was reused and was not modified!
+Thus existing code was reused and was not modified!
 
@@
 the use of factory methods, instead of constructors.
-To pretty-print the expression,
-we pass an instance of \inlinecode{VarPrint} to \inlinecode{makeExp}:
+To pretty-print the expression, the code
 
@@
 \end{lstlisting}
-which results in \inlinecode{"(1 + x)"} as expected.
+results in \inlinecode{"(1 + x)"} as expected.
 
@@
 For example, we have demonstrated in Section~\ref{subsec:overriding} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
-However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]}, where the type of the \emph{extended} Object Algebra interface appears in \emph{contravariant} position.
-This difference prohibits method/field overriding.
-Scala supports type-refinement of fields if the overriding
-field has a type which is a \emph{subtype} of the original field.
-Unfortunately, in this case, although \inlinecode{VarAlg} is a subtype
-of \inlinecode{Alg}, the function type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]} is a supertype of \inlinecode{Alg[E] =>} \inlinecode{Parser[E]}. Thus it is not possible to have overriding.
+However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]}, which is a \emph{supertype} of the old type \inlinecode{Alg[E] =>} \inlinecode{Parser[E]}, since
+the \emph{extended} Object Algebra interface appears in \emph{contravariant} position. This violates
+overriding in Scala.
 
@@
 to \lstinline{VarAlg[E]}, to allow calling the additional factory method
-for variables. Note that, unlike the previous attempt, \lstinline{VarAlg[E]}
-is a subtype of \lstinline{Alg[E]}, thus the type-refinement of the
-field is allowed. Now, the code for parsing variables (\lstinline{pVar}) can
+for variables. Unlike the previous attempt, such a type-refinement is allowed.
+Now, the code for parsing variables (\lstinline{pVar}) can
 call \lstinline{alg.varE}. The following code illustrates how to use
@@
 
-\lstinputlisting[linerange=43-46]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
+\lstinputlisting[linerange=46-49]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
 
-In the client code shown above, we first need to choose a concrete
-Object Algebra to initialize the \lstinline{alg} field in \lstinline{VarOAParser}. Here we pick the
-pretty-printing algebra \lstinline{VarPrint}, but any other Object
+In the client code above,
+we pick the pretty-printing algebra \lstinline{VarPrint} to initialize the \lstinline{alg} field, but any other Object
 Algebra that implements \lstinline{VarAlg} would work.
@@
 \lstinline{pE} to obtain the parser to feed to the \lstinline{parse} method.
-The result shows that such a pattern provides modular parsing.
+Such a pattern provides modular parsing as expected.
 
-Note that, similar to the approach in Section~\ref{sec:inheritance} independent extensibility is also supported via multiple trait inheritance.
-Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance} we omit that code here.
+Note that, similar to the approach in Section~\ref{sec:inheritance}, independent extensibility is also supported via multiple trait inheritance.
+Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance}, we omit that code here.

### 2017-05-11 SLE_17/sections/PackratParsing.tex
Some changes.

@@
 \subsection{Algorithmic Challenges of Modularity}\label{subsec:challenges}
-At a first look, parser combinators are very suitable for modular parsing, because of two reasons.
-Firstly, they are naturally modular.
-The manner of using them is to write small parsers and use combinators to compose them together.
-The construction procedure is explicit and fully controlled by the programmer.
-Secondly, each parser combinator is represented by a piece of code, and so are the parsers it takes.
-Thus in a statically typed programming language they can be statically type-checked.
+At a first look, parser combinators are very suitable for modular parsing.
+One reason is that they are naturally modular for parser composition, providing full control to
+the programmer. The other one is that they are implemented by a piece of code, so that it ensures
+type safety in a statically typed programming language.
 Unfortunately many parser combinators have important limitations.
@@
 Consider extending the language by adding subtractions to the grammar.
-If we still write its parser by directly following the grammar structure, the new left-recursive case will get the parser into an infinite loop.
-Namely, \inlinecode{parseExpr} and \inlinecode{parseSub} call each other and never stop.
+If we directly extend the grammar, the new left-recursive case will cause an infinite loop.
+Below, \inlinecode{parseExpr} and \inlinecode{parseSub} call each other and never stop.
 
@@
 
-To solve this issue, we have to rewrite the grammar as below to eliminate left-recursion and build the parser based on this new grammar.
+Instead, we have to rewrite the grammar as below to eliminate left-recursion.
 
@@
 
-In Parsec, for instance, the parser below will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
+In Parsec, for instance, the parser \inlinecode{"parseInt <|> parseAdd"} will only parse the input \inlinecode{"1 + 2"} to \inlinecode{"1"}, as \inlinecode{parseInt} successfully parses \inlinecode{"1"}
 and terminates parsing.
 
-\begin{lstlisting}[language=PlainCode]
-parser = parseInt <|> parseAdd
-\end{lstlisting}
-
-Using traditional alternative composition, when a preceding parser successfully parses a prefix of the input, it will finish parsing and return the result, in spite that subsequent parsers may be able to parse the whole input.
-In contrast with the parser above, \inlinecode{parser = parseAdd <|> parseInt} works as expected with the two cases swapped.
+Traditional alternative composition will only find the first parser that succeeds on a prefix of the input,
+in spite that subsequent parsers may parse the whole input. In contrast with the parser above, \inlinecode{"parseAdd <|> parseInt"} works as expected with the two cases swapped.
 
@@
 \paragraph{Backtracking} The need for backtracking can also be problematic
-in a modular setting. For example, consider a grammar that includes
-``import'' statements including \inlinecode{import..from}.
-Now we want to extend the grammar with an \inlinecode{import..as} case, as shown in the third line of grammar below.
+in a modular setting. Consider a grammar that includes \inlinecode{"import..from"}.
+It is extended with an \inlinecode{"import..as"} case:
 
@@
 \end{grammar}
-
-Since the \inlinecode{import..from} case shares a prefix with the new case \inlinecode{import..as}, when the former case fails, we must backtrack to the beginning.
-Take Parsec as an example, its choice
-combinator \inlinecode{<|>} only tries the second alternative if the first fails
-without any token consumption. An auxiliary function \inlinecode{try} is used for explicit backtracking.
+Since the two cases share a common prefix, when the former fails, we must backtrack to the beginning.
+For example, the choice combinator in Parsec only tries the second alternative if the first fails
+without any token consumption. We have to use \inlinecode{try} for explicit backtracking.
 
 \begin{lstlisting}[language=PlainCode]
-oldParser = parseImpFrom <|> parseA <|> parseB <|> ...
-newParser = try parseImpFrom <|> parseA <|> parseB <|> ... <|> parseImpAs
+oldParser = parseImpFrom <|> ...
+newParser = try parseImpFrom <|> ... <|> parseImpAs
 \end{lstlisting}
 
-Given the full grammar, we can decide where to put \inlinecode{try} for backtracking.
-However, with modular parsing we are unable to have a global view of the full grammar.
+Similarly, this violates a modular setting because we cannot have a global view of the full grammar.
 Hence the worst case should be considered that all alternatives may share common prefixes with future cases.
 In that case we need to backtrack for all the branches.
-To avoid failures in the future, we have to add \inlinecode{try} everywhere:
-
-\begin{lstlisting}[language=PlainCode]
-parser = try parseImpFrom <|> try parseA <|> try parseB <|> ... <|> try parseImpAs
-\end{lstlisting}
-
-\noindent However, this results in the worst-case exponential time
-complexity in Parsec, because it does not provide optimizations
-for backtracking.
+To avoid failures in the future, we have to add \inlinecode{try} everywhere.
+However this results in the worst-case exponential time
+complexity.
 
@@
 parsing~\cite{Ford2002} have been developed to address limitations of
-simple parser combinators. Packrat parsers use memoization to record
+simple parser combinators. Packrat parsing uses memoization to record
 the result of applying each parser at each position of the input, so
-that repeated computation is eliminated. Moreover, Packrat parsing, in
-theory, supports both direct and indirect left-recursion~\cite{warth2008}.
+that repeated computation is eliminated. Moreover, it supports both direct left-recursion
+and (in theory) indirect left-recursion~\cite{warth2008}.
 All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique for modular parsing.
@@
 Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
-\cite{moors2008parser} which implements Packrat parsers.
+\cite{moors2008parser} for Packrat parsers.
 The library provides a number of parser combinators, including the longest match alternative combinator.
@@
 
-It is worth mentioning that the grammar is left-recursive in the addition case.
-Since left-recursion is well supported by Packrat parsers, we do not need extra code for it.
-Our parser also employs longest match composition by using the combinator \lstinline{|||} in the library.
+It is worth mentioning that the left-recursive grammar above is well supported without extra code.
+ The longest match composition is also employed by using the combinator \lstinline{|||}.
 Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees reasonable efficiency.

### 2017-05-15 SLE_17/sections/Discussion.tex
More.

@@
 Using Object Algebras, it is easy to
-model multi-sorted languages.
-If the syntax contains multiple sorts, we can distinguish them by different type parameters.
-As an example, we extend the language of literals, additions and variables by introducing types, including primitive type \inlinecode{int} and function (arrow) type, and we add lambda abstractions to expressions.
-Now the language has two sorts in the syntax, types and expressions:
+model multi-sorted languages. If the syntax contains multiple sorts, we can distinguish them by different type parameters.
+For instance, now we extend the expression language from the end of Section~\ref{sec:algebrasandparsing}, with a
+primitive type \inlinecode{int} type and typed lambda abstractions:
 
@@
 \begin{grammar}
-<type> ::= `int' \alt <type> `->' <type>
+<type> ::= `int'
 
@@
 We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
-They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
+The type system guarantees that invalid terms such as \inlinecode{int + int} will be rejected.
 Besides lexing, the trait \inlinecode{LamOAParser} also introduces parsers for types, and the new case for expressions.
@@
 \begin{figure}[ht]
-\lstinputlisting[linerange=7-33]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+\lstinputlisting[linerange=7-23]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
 \caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
@@
 \end{figure}
-
-From the code, we observe that the multi-sorts of Object Algebra interface
-express multiple syntax. The isolation for different syntax
-is guaranteed by the type system. The following client code
-
-\lstinputlisting[linerange=40-43]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
-works as expected to parse and pretty-print the result.
 
@@
 situations, for instance when conflicts or ambiguities arise upon composing languages.
-As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below. Some irrelevant details are omitted.
-
-\lstinputlisting[linerange=76-82]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM
+As an illustration, suppose we have an untyped lambda abstraction case in a base parser, defined as a value:
+\begin{lstlisting}
+val pLam: Parser[E] =
+  ("\\" ~> ident) ~ ("." ~> pE) ^^ ...
+\end{lstlisting}
 Here \inlinecode{pLam} parses a lambda symbol, an identifier, a dot and an expression in sequence.
@@
 implementation of lambdas will be different without affecting the other parts of the parser.
-
-\lstinputlisting[linerange=86-92]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM
+\begin{lstlisting}
+override val pLam: Parser[E] =
+  ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^ ...
+\end{lstlisting}
 
@@
 
-We take the example in Section~\ref{subsec:parsingwithoa} again, which is a language of literals, additions and variables. It can be defined as a language component \inlinecode{VarExpr}.
+We take the example in Section~\ref{subsec:parsingwithoa} again. It can be defined as a language component \inlinecode{VarExpr}.
+For space reasons we omit some detailed code.
 
-\lstinputlisting[linerange=115-130]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
+\lstinputlisting[linerange=115-124]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
 
@@
 
-\lstinputlisting[linerange=134-150]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
+\lstinputlisting[linerange=128-137]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
 
@@
 
-\lstinputlisting[linerange=154-165]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
+\lstinputlisting[linerange=141-152]{code/src/papercode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
 
@@
 
-Under our modular parsing framework, we use Packrat parsing as the underlying parsing technique, OO inheritance with method overriding for composing and extending parsers, and Object Algebras for parsing extensible ASTs.
-However, such a pattern is more general and thus more powerful, because those aspects are orthogonal to each other, and hence can have alternatives.
-
-\begin{itemize}[leftmargin=*]
-
-\item {\bf Parsing Technique:}
-The pattern does not depend on a particular parser combinator library
-or parsing algorithm. We demonstrated that Packrat parsing has some
-advantages and is suitable in a modular setting. However, under
-other circumstances it may not be
-the best choice. One can use other parser combinators as long
-as the guidelines proposed in Section~\ref{sec:packrat} are met.
-
-\item {\bf Composing and Extending Parsers:}
-  We use traits to model parsers, OO inheritance to compose them, and
-  method overriding for their extensibility. These features
-  may not be supported in other programming languages, especially in
-  functional languages such as Haskell. Nevertheless, \textit{open
-    recursion}~\cite{CookThesis} could be used as an alternative. An explicit
-  ``self-reference'' parameter is able to explain the recursive calls
-  dynamically, by the real argument passed at runtime.
-
-\item {\bf Extensible ASTs:}
-Besides Object Algebras, many other techniques including \textit{Data
-  types à la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
-These alternatives could also be adopted to build ASTs and their parsers.
-
-\end{itemize}
-
-As we demonstrated, such a pattern for modular parsing is 
-modular itself, and can be customized easily.
+Our prototype uses Packrat parsing as the underlying parsing technique, OO inheritance for composing and extending parsers, and Object Algebras for parsing extensible ASTs.
+Yet such a framework is itself flexible and modular, because those techniques can have alternatives.
+For example, as we mentioned before, any parsing library
+that resolves the algorithmic challenges in modular parsing can work well.
+Regarding OO inheritance for the extensibility, an alternative approach, called \textit{open recursion}~\cite{CookThesis} can be used in other languages, by introducing explicit ``self-reference'' parameters for the recursion.
+Furthermore, besides Object Algebras, v\textit{Data types à la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
+For the goal of modular parsing a custom combination of those alternatives can be adopted.

### 2017-05-15 SLE_17/sections/Inheritance.tex
More.

@@
 
-\subsection{The Mismatch Between OO ASTs and Grammars}
+\subsection{The Mismatch Between OO ASTs and Grammars}\label{subsec:mismatch}
 Object-oriented ASTs are ASTs encoded via standard OO techniques,
@@
 fix the operations required in the AST, but allow the easy addition of new
-language constructs. This is in contrast with ASTs modelled using
-functional programming algebraic datatypes or OO style Visitors, where
-adding new operations is easy, but adding new language constructs is hard.
-In Section~\ref{subsec:packratparsing} we provide a simple example of OO ASTs. The same abstract syntax could be encoded using algebraic datatypes (or alternatively Visitors)
-as:
+language constructs. We have seen an example of OO ASTs in Section~\ref{subsec:packratparsing}.
+In contrast, ASTs modelled using functional programming algebraic datatypes (or alternatively OO style Visitors) are as follows:
 
@@
 obvious similarities in declaring a grammar, and declaring an
-algebraic datatype. It is difficult to modularly add new language
-constructs to both grammars and ASTs based on algebraic datatypes. However,
-ASTs modelled with algebraic datatypes naturally allow for the easy
-addition of operations. So using such ASTs at least preserves one
-dimension of extensibility.
+algebraic datatype. Unlike OO ASTs, they naturally allow for the easy
+addition of operations, whereas modularly adding new constructs to grammars
+or ASTs becomes hard. Using this style of ASTs again preserves one dimension of extensibility.
 
@@
 \lstinputlisting[linerange=46-50]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
-
 The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}.
@@
 
-As a concrete example, we compose the language of integers and additions with a language of boolean values and if-then-else expressions, to build a new language.
+%As a concrete example, we compose the language of integers and additions with a language of boolean values and if-then-else expressions, to build a new language.
 
-\lstinputlisting[linerange=56-73]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
+%\lstinputlisting[linerange=56-68]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE
 
@@
 
-Then the new parser, which contains both integer and boolean
-expressions, is able to parse hybrid expressions such as
-\inlinecode{"if true then 1 + 2 else 3 + 4"}.
+%Then the new parser, which contains both integer and boolean
+%expressions, is able to parse hybrid expressions such as
+%\inlinecode{"if true then 1 + 2 else 3 + 4"}.
 

### 2017-05-15 SLE_17/sections/ObjectAlgebras.tex
More.

@@
 The Expression Problem~\cite{wadler1998expression} illustrates the
-difficulty of extending data structures or ASTs in two dimensions. For
-traditional OO ASTs, adding new language constructs is easy, but
-adding new operations is hard. In the last section, we have a simple
-language containing literals, additions and variables. The
-corresponding ASTs are built using classes with a pretty-printing
-operation.
+difficulty of extending data structures or ASTs in two dimensions.
+As we have mentioned in Section~\ref{subsec:mismatch}, the use of traditional ASTs
+makes it hard to add new operations. In the last section we have seen a language
+that supports pretty-printing (in \inlinecode{Expr}).
 To modularly add an operation like collecting free variables, one attempt
@@
 in many cases). More importantly, a second problem is that
-even if that approach would work, the parsing code would not
-be reusable. \lstinline{VarParser} is no longer suitable for parsing the new ASTs! All the methods in the previous parser have type \inlinecode{Parser[Expr]},
-which generates ASTs that only support pretty-printing. Now we have to substitute \inlinecode{NewExpr} for \inlinecode{Expr} (the
-same for \inlinecode{Lit}, \inlinecode{Add}, and so on). It requires either modification on existing code or casts.
+even if that approach would work, the parsing code in \lstinline{VarParser} is no longer reusable!
+The types \inlinecode{Expr, Lit, Add}, and so on, are all old types without the free variables operation.
+To match the new ASTs, we have to substitute \inlinecode{NewExpr} for \inlinecode{Expr} (the
+same for \inlinecode{Lit}, \inlinecode{Add}, ...). It requires either modification on existing code or casts.
 The goal of semantic modularity motivates us to find a different approach for building ASTs.
 
-Fortunately, Object Algebras~\cite{Oliveira:2012} enable us to solve
-this problem. Object Algebras separate data variants and
-operations, and offer high flexibility in the choice of operations to
-be performed over ASTs. The next section provides a brief introduction
-to Object Algebras.
+%Object Algebras separate data variants and
+%operations, and offer high flexibility in the choice of operations to
+%be performed over ASTs. The next section provides a brief introduction
+%to Object Algebras.
 
@@
 
-Object Algebras capture a design pattern that addresses the Expression
+Fortunately, Object Algebras~\cite{Oliveira:2012} enable us to solve
+this problem.
+They capture a design pattern that addresses the Expression
 Problem, achieving two dimensions of extensibility (language
@@
 Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
-One can define evaluation operation as a new trait on this expression language:
+One can also define an evaluation operation as a new trait that extends \inlinecode{Alg[Int]}. Hence adding new operations is modular. We omit that code
+due to space reasons.
 
-\lstinputlisting[linerange=20-23]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
+%\lstinputlisting[linerange=20-23]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
@@
 the use of factory methods, instead of constructors.
-To pretty-print the expression, the code
-
-\begin{lstlisting}
-makeExp(new VarPrint {})
-\end{lstlisting}
+To pretty-print the expression, the code \inlinecode{"makeExp(new VarPrint \{\})"}
 results in \inlinecode{"(1 + x)"} as expected.
@@
 \paragraph{Extensions}
-Now let's try this pattern with the variables extension:
+Now let's try on the variables extension:
 

### 2017-05-15 SLE_17/sections/PackratParsing.tex
More.

@@
 \subsection{Algorithmic Challenges of Modularity}\label{subsec:challenges}
-At a first look, parser combinators are very suitable for modular parsing.
-One reason is that they are naturally modular for parser composition, providing full control to
-the programmer. The other one is that they are implemented by a piece of code, so that it ensures
-type safety in a statically typed programming language.
+For the goal of modular parsing, parser combinators seem suitable because they are
+naturally modular for parser composition, but also they ensure type safety.
 Unfortunately many parser combinators have important limitations.
@@
 \paragraph{Left-Recursion Elimination} The top-down, recursive descent parsing strategy adopted by those parser combinator libraries cannot support left-recursive grammars directly.
-The common solution is to rewrite the grammar into an equivalent but non-left-recursive one, called left-recursion elimination.
-
-The grammar below represents a simple arithmetic language containing only integers. Its corresponding parser is shown on the right side, written in Parsec.
-
-\begin{minipage}{10em}
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= <int>
-\end{grammar}
-\end{minipage}
-\begin{minipage}{10em}
-\begin{lstlisting}[language=PlainCode]
-parseExpr = parseInt
-\end{lstlisting}
-\end{minipage}
-
-Consider extending the language by adding subtractions to the grammar.
-If we directly extend the grammar, the new left-recursive case will cause an infinite loop.
-Below, \inlinecode{parseExpr} and \inlinecode{parseSub} call each other and never stop.
+For instance, we start with a simple arithmetic language containing only integers and subtractions.
+The
+grammar with concrete syntax and part of the parsing code in Parsec are presented below:
 
@@
   parseSub <|> parseInt
-
 parseSub = do
-  e <- parseExpr
-  ...
+  e <- parseExpr ...
 \end{lstlisting}
 \end{minipage}
-
-Instead, we have to rewrite the grammar as below to eliminate left-recursion.
+Such a left-recursive implementation will cause an infinite loop, since \inlinecode{parseExpr} and \inlinecode{parseSub} call each other and never stop.  
+A common solution is to rewrite the grammar into an equivalent but non-left-recursive one, called left-recursion elimination:
 
@@
 
-After left-recursion elimination, the structure of grammar is changed,
+\noindent After left-recursion elimination, the structure of grammar is changed,
 as well as its corresponding parser. In a modular setting, it is
@@
 When the grammar is extended with new rules, programmers are supposed to \emph{manually} adjust
-the order of parsers, which requires rewriting previously written code.
+the order of parsers, by rewriting previously written code.
 
@@
 
-Similarly, this violates a modular setting because we cannot have a global view of the full grammar.
+Similarly, this violates a modular setting because it also requires a global view of the full grammar.
 Hence the worst case should be considered that all alternatives may share common prefixes with future cases.
@@
 
-\lstinputlisting[linerange=9-17]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
+\lstinputlisting[linerange=9-15]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
 
@@
 
-\lstinputlisting[linerange=21-28]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
+\lstinputlisting[linerange=19-26]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
 
@@
 - A parser combinator for function application. \\ \hline
-\lstinline{def \^\^\^[U](v: => U): Parser[U]} \\
-- A parser combinator that changes a successful result into the specified value. \\ \hline
+%\lstinline{def \^\^\^[U](v: => U): Parser[U]} \\
+%- A parser combinator that changes a successful result into the specified value. \\ \hline
 \lstinline{def <~[U](q: => Parser[U]): Parser[T]} \\

### 2017-05-15 SLE_17/code/src/papercode/Sec2Packrat/Code1.scala

@@ -6,9 +6,7 @@ import papercode.Common
 trait Code1 extends Common {
 
 //BEGIN_PACKRAT_OO_AST
-trait Expr {
-  def print: String
-}
+trait Expr { def print: String }
 class Lit(x: Int) extends Expr {
   def print = x.toString
 }

### 2017-05-15 SLE_17/code/src/papercode/Sec4OA/Code4.scala

@@ -6,24 +6,14 @@ trait Code4 extends Code3 {
 //BEGIN_OVERVIEW_OA_MULTI_SYNTAX
 trait LamAlg[E, T] extends VarAlg[E] {
   def intT(): T
-  def arrT(t1: T, t2: T): T
   def lam(x: String, t: T, e: E): E
-}
-trait LamPrint extends LamAlg[String, String]
-  with VarPrint {
-  def intT() = "int"
-  def arrT(t1: String, t2: String) = t1 + "->" + t2
-  def lam(x: String, t: String, e: String) =
-    "\\" + x + " : " + t + ". " + e
 }
 trait LamOAParser[E, T] extends VarOAParser[E] {
   lexical.reserved += "int"
   lexical.delimiters += ("->", "\\", ":", ".")
   override val alg: LamAlg[E, T]
-  val pIntT: Parser[T] = "int" ^^^ alg.intT
-  val pArrT: Parser[T] = pT ~ ("->" ~> pT) ^^
-    { case t1 ~ t2 => alg.arrT(t1, t2) }
-  val pTypedLamT: Parser[T] = pIntT ||| pArrT
+  val pIntT: Parser[T] = "int" ^^ { _ => alg.intT }
+  val pTypedLamT: Parser[T] = pIntT
   val pLam: Parser[E] =
     ("\\" ~> ident) ~ (":" ~> pT) ~ ("." ~> pE) ^^
       { case x ~ t ~ e => alg.lam(x, t, e) }

### 2017-05-15 SLE_17/code/src/papercode/Sec5/Code1.scala

@@ -115,17 +115,11 @@ object Code2 extends Common {
 object VarExpr {
   trait Alg[E] {    // Abstract syntax
     def lit(n: Int): E
-    def add(e1: E, e2: E): E
-    def varE(x: String): E
-  }
-  trait Parse[E] {   // Parser
-    val alg: Alg[E]
-    val pE: Parser[E] = ...
     ...
   }
+  trait Parse[E] { ... } // Parser
   trait Print extends Alg[String] {
-    // Pretty-printer
-    ...
+    ... // Pretty-printer
   }
 }
 //END_LANGUAGE_COMPONENTS_VAREXPR
@@ -134,18 +128,11 @@ object VarExpr {
 object TypedLam {
   trait Alg[E, T] {    // Abstract syntax
     def intT(): T
-    def arrT(t1: T, t2: T): T
-    def lam(x: String, t: T, e: E): E
-  }
-  trait Parse[E, T] {   // Parser
-    val alg: Alg[E, T]
-    val pE: Parser[E] = ...
-    val pT: Parser[T] = ...
     ...
   }
+  trait Parse[E, T] { ... } // Parser
   trait Print extends Alg[String, String] {
-    // Pretty-printer
-    ...
+    ... // Pretty-printer
   }
 }
 //END_LANGUAGE_COMPONENTS_TYPEDLAM

### 2017-05-15 SLE_17/sections/RelatedWork.tex
related work

--- 2017-01-14 paper/sections/RelatedWork.tex
@@
 
-\paragraph{Syntactically Extensible Parsing} Extensible parsing is
-used in many different areas. Extensible parser generator is a
-mainstream area specially designed for modular syntax and
-parsing. Many parser
-generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
-modular grammars. They allow users to create new modules where new
+\paragraph{Syntactically Extensible Parsing}
+Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}
+are a mainstream area of modular syntax and parsing. They allow users to write
+modular grammars, where new
 non-terminals and production rules can be introduced, some can even
@@
 \textit{Rats!}~\cite{Grimm2006} constructs its own module system for
-the collection of grammars.  NOA~\cite{Gouseti2014} uses Java
-annotation processing to gather information about grammars in Java
-code. After all information is collected, NOA merges all grammar
-components, produce an ANTLR~\cite{antlr1995} grammar, and generates
-corresponding parsing code from the ANTLR grammar. Those parser
+the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
+annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
+and the parsing code. Those parser
 generators focus on the \textit{syntactic extensibility} of grammars:
@@
 Java using library imports. Composition of syntactic sugar is easy for
-users, but it requires many rounds of parsing and adaption. This
-significantly affects efficiency of compilation. Since the
+users, but it requires many rounds of parsing and adaption, hence
+significantly affects the efficiency of compilation. Since the
 implementation was based on SDF~\cite{Heering1989} and
@@
 compilation. Racket adopts a macro system for library-based language
-extensibility~\cite{Tobin-Hochstadt2011}. This is quite powerful and it uses
+extensibility~\cite{Tobin-Hochstadt2011}. It uses
 attributed ASTs for contextual
-information. Moreover extensions can be integrated in a modular
+information, and extensions can be integrated in a modular
 way. However such modularity is not flexible enough for language
@@
 is mostly done using parser generators. They focus on the
-extensions to a host language. Our technique is designed for type-safe
-modular language parsers (not just language
-extensions), with very flexible parser composition.
+extensions to a host language. Those techniques are short of type safety in a modular
+setting as well.
 
@@
 \paragraph{Extensible Parsing Algorithm}
-Extensible parsing algorithm is another area that relates to
-separate compilation.
 \textit{Parse table composition}~\cite{bravenboer2008parse}
@@
 
-\paragraph{Parser Combinators} Since Burge~\cite{burge1975} firstly
-introduced parser generators, and after Wadler popularized them
-in functional programming~\cite{Wadler1985}, parser combinators have been
-more and more popularly developed and used in the research area of
-parsing. Many parsing libraries produce recursive descent
+\paragraph{Parser Combinators} Parser combinators have become more and more
+popular since~\cite{burge1975,Wadler1985}. Many parsing libraries produce recursive descent
 parsers by introducing functional monadic
@@
 novel parsing techniques that address the issue of
-left-recursion. We have selected Packrat parsing for our approach,
-as the combinators are convenient to use in Scala, and
-integrate well with multiple trait inheritance and
-Object Algebras.
-%Also,
-%its support for direct left-recursion is already helpful for designing
-%real-world parsers practically, though~\cite{warth2008} further
-%demonstrated that general left-recursion is supported theoretically.
-Despite our choice of Packrat parsing, as argued in Section~\ref{sec:packrat},
-any parser combinator approach which supports left-recursion, longest
-match composition, and makes the cost of backtracking reasonable,
-should be applicable to modular parsing.
-%Yet we have mentioned that the components used in our approach can
-%have alternatives to achieve better modular parsing.
+left-recursion. We have selected Packrat parsing due to its simplicity in Scala,
+but in general it can also have alternatives.
 
@@
 languages, have been proposed over the years to address extensibility
-problems, such as the Expression Problem~\cite{wadler1998expression}. However, to the best
-of our knowledge, none of the existing extensibility techniques
+problems, such as the Expression Problem~\cite{wadler1998expression}.
+The famous ``Datatypes \`a
+la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
+of every two functors. Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
+All of that work essentially covers how to traverse and
+consume extensible ASTs. However they do not
 address the problem of \emph{modularly
-parsing extensible ASTs}.  There has been
-a lot of work on extensible ASTs, in both functional and
-object-oriented programming.  For instance, the famous ``Datatypes \`a
-la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
-of every two functors, which forms a binary tree structure.
-Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
-All of that work essentially covers how to traverse and
-consume extensible ASTs.
-Only in Bahr's~\cite{Bahr2011} work a brief mention of \emph{unfolds}
-(which are a class of algorithms that \emph{build} ASTs) are mentioned.
-However Bahr's work does not cover parsing.
+parsing extensible ASTs}. Only in Bahr's~\cite{Bahr2011} work \emph{unfolds} is briefly mentioned,
+yet it does not cover parsing.
 
-In OO languages, there are also many design patterns that achieve
+There are also many design patterns in OO languages that achieve
 type-safe extensibility~\cite{torgersen2004expression,odersky2005independently,oliveira2009modular,Oliveira:2012,wang2016expression}.
@@
 the parsing code is concise and expressive using Object Algebras.
-In principle our approach is not coupled to the use of Object
-Algebras, and variants of our approach could be used with other
-extensibility techniques. Similarly to the work in functional
-programming, the extensibility techniques in OOP have focused
-on traversing and processing ASTs. We are unware of any work on OOP
-that covers how to do modular parsing for extensible ASTs.
+On the other hand we are unware of any work on OOP
+that has covered how to do modular parsing for extensible ASTs.
 
 It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
-solution to the Expression Problem. Case classes are similar to
-algebraic datatypes but, unlike algebraic datatypes, new cases can be
-modularly added. The reason why case classes are not a complete
-solution to the Expression Problem is that they do not enforce
+solution to the Expression Problem. Case classes can be
+modularly added, but they do not enforce
 exhaustiveness of pattern matching for extensible operations. In other
@@
 all that the users need to do is to use case classes, instead of
-standard OO classes in their code.
+standard OO classes in their code.\haoyuan{To simplify; conclusion}
 

### 2017-05-15 SLE_17/sections/CaseStudy.tex
Case study

@@
-\section{Case Studies}\label{sec:casestudy}
+\section{Case Study}\label{sec:casestudy}
 
@@
 TAPL introduces several calculi from simple to complex, by gradually adding new features to syntax.
-These calculi are suitable for our case study for mainly two reasons.
-Firstly, they capture many of the language features
-required in realistic programming languages, such as lambdas, records and variants. Secondly, the evolution of
-calculi in the book reveals the advantages of modular representation
+These calculi are suitable for our case study for mainly two reasons. Firstly, they capture many of the language features
+required in realistic programming languages, such as lambdas, records and polymorphism.
+Secondly, the evolution of calculi in the book reveals the advantages of modular representation
 of abstract syntax and modular parsing, which is the key functionality
@@
 
-\paragraph{Extracting Language Components}
-Using the pattern demonstrated in Section~\ref{subsec:language-component}, we extract
-reusable components from all the calculi. Each
+%% \paragraph{Extracting Language Components}
+We extract reusable components from all the calculi using the pattern
+demonstrated in Section~\ref{subsec:language-component}. Each
 component, which may contain several syntactical structures,
-represents a certain feature of the language. For
-example, the \lstinline{VarApp} component below represents variables and
-function applications.
+represents a certain feature. They are combined together as needed to
+build a calculus. For example, the calculus \lstinline{Untyped} in our case study,
+representing the famous untyped lambda calculus, consists of component
+\lstinline{VarApp} (for variables and applications)
+and component \lstinline{UntypedAbs} (for untyped lambdas).
 
-\lstinputlisting[linerange=13-29]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
+%% For example, the \lstinline{VarApp} component below represents variables and
+%% function applications.
 
-Similar as before, each component is represented by a Scala object which includes \lstinline{Alg} for the abstract syntax, \lstinline{Print} for pretty-printing, and \lstinline{Parse} for parsing.
+%% \lstinputlisting[linerange=13-29]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_VARAPP
 
-We have some naming conventions in our code, \lstinline{E} represents
-expressions, \lstinline{T} represents types and \lstinline{K}
-represents kinds. They are the three sorts of syntax in our case study.
-In the component \lstinline{VarApp} we only have expressions.
-We use some helper traits for eliminating duplicate definitions, such as \inlinecode{EParser} containing \inlinecode{pE} for parsing expressions.
-
-\lstinputlisting[linerange=9-9]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
-
-\paragraph{Composing Language Components}
-Each calculus could be composed directly from components and other
-calculi as needed. For example, the calculus \lstinline{Untyped} in our case study,
-representing the famous untyped lambda calculus, is constructed from components \lstinline{VarApp}
-and untyped lambda abstraction \lstinline{UntypedAbs}.
-
-The code of building \lstinline{Untyped} is presented in Figure~\ref{fig:casestudy-untyped}. Note that in the parser \inlinecode{Parse}, we need to override the object algebra interface and
-parsing functions accordingly.
-
-\begin{figure}[t]
-\lstinputlisting[linerange=33-60]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
-\caption{Build the \inlinecode{Untyped} calculus by composing to language components.}\label{fig:casestudy-untyped}
-\end{figure}
-
-\paragraph{Dependency Overview}
 Figure \ref{fig:dependency} shows the
@@
     \includegraphics[width=\textwidth]{resources/depGraph.pdf}
-    \caption{Dependency graph of all calculi and components in case study.}
+    \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
     \label{fig:dependency}
@@
 
-As shown in the graph, some components such as \lstinline{VarApp} are
-created from scratch, while others such as \lstinline{Typed} are
-extended from existing components. Since calculi and components have similar signatures, each calculus
+Each component or language is represented by a Scala object which includes \lstinline{Alg}
+for the abstract syntax, \lstinline{Print} for pretty-printing, and \lstinline{Parse} for parsing.
+Since calculi and components have similar signatures, each calculus
 can also be extended and reused directly. For example, calculus \lstinline{FullRef} extends from
@@
 
-From the dependency graph, we know
-that common components such as \lstinline{VarApp} are reused in
-lots of calculi. Such reuse could shorten the code considerably. We
-will show this advantage and inspect the possible performance penalty in
-a quantitative way in the next subsection.
+%% We have some naming conventions in our code, \lstinline{E} represents
+%% expressions, \lstinline{T} represents types and \lstinline{K}
+%% represents kinds. They are the three sorts of syntax in our case study.
+%% In the component \lstinline{VarApp} we only have expressions.
+We use some helper traits for eliminating duplicate definitions, such as \inlinecode{EParser} containing \inlinecode{pE} for parsing expressions.
+
+%% \lstinputlisting[linerange=9-9]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_EPARSER
+
+%% \paragraph{Composing Language Components}
+
+%% The code of building \lstinline{Untyped} is presented in Figure~\ref{fig:casestudy-untyped}. Note that in the parser \inlinecode{Parse}, we need to override the object algebra interface and
+%% parsing functions accordingly.
+
+%% \begin{figure}[t]
+%% \lstinputlisting[linerange=33-60]{code/src/papercode/Sec6CaseStudy/Code1.scala}% APPLY:linerange=CASESTUDY_UNTYPED
+%% \caption{Build the \inlinecode{Untyped} calculus by composing to language components.}\label{fig:casestudy-untyped}
+%% \end{figure}
+
+%% \paragraph{Dependency Overview}
+
+%% As shown in the graph, some components such as \lstinline{VarApp} are
+%% created from scratch, while others such as \lstinline{Typed} are
+%% extended from existing components.
+
+From the graph, we know that common components such as \lstinline{VarApp} are reused in
+lots of calculi. Such reuse could shorten the code considerably.
+Next subsection examines amount of code as well as performance.
 
@@
 We compared our implementation (named \ourimpl{}) with an implementation
-by Ilya Klyuchnikov (named \ilyaimpl{}), available online\footnote{https://github.com/ilya-klyuchnikov/tapl-scala/}.
+available online\footnote{https://github.com/ilya-klyuchnikov/tapl-scala/} (named \ilyaimpl{}).
 \ilyaimpl{} is suitable for comparison, because it is also
@@
 \paragraph{Standard of Comparison}
-In the SLOC comparison, all blank lines and comment lines are excluded,
+In the SLOC comparison, all blank lines and comments are excluded,
 and we formatted the code of both implementations to guarantee that
@@
 For the comparison of execution time, we built a generator to randomly
-generate valid expressions of each calculus, according to the syntax. These expressions are
+generate valid expressions for each calculus, according to its syntax. These expressions are
 written to test files, one file per calculus. Each test file consists of 500
 expressions randomly generated, and the size of test files varies from 20KB to 100KB.
-Then we run the corresponding parser to parse the file and the pretty-printer to print the result.
+We run the corresponding parser to parse the file and the pretty-printer to print the result.
 The average execution time of 5 runs excluding reading input file was calculated, in milliseconds.
@@
     \centering
-    \begin{tabular}{|l|r|r|r|r|r|r|}
+    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|}
       \hline
-        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{3}{ c| }{\bfseries Time (ms)} \\ \cline{2-7}
-        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% \\
+        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{7}{ c| }{\bfseries Time (ms)} \\ \cline{2-11}
+        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyalongest{} & \bfseries (+/-)\% & \ourclass{} & \bfseries (+/-)\% \\
       \hline
@@
 Table \ref{tab:comparison} shows results of the comparison.
+Let's only check \ourimpl{} and \ilyaimpl{} for now.
 The overall result is that 69.2\% of code is reduced using our
@@
 The good SLOC result is because of that the code of common language features
-such as variables, lambda abstractions, etc., are reused lots of times in
-the whole case study. We can see that in the first two calculi
-\lstinline{Arith} and \lstinline{Untyped} we are not better than \ilyaimpl{}, because in such two cases we do not reuse any existing
-components. However in the following 16 calculi, we can reuse
-language components, resulting considerably code reduction. In particular,
-the calculi \inlinecode{EquiRec}, \inlinecode{Recon} and some others are only 22 lines
-in our implementation, because we only compose existing codes in such cases.
+are reused lots of times in the whole case study. We can see that in the first two calculi
+\lstinline{Arith} and \lstinline{Untyped} we are not better than \ilyaimpl{},
+because in such two cases we do not reuse anything.
+However in the following 16 calculi, we indeed reuse language components.
+In particular, the calculi \inlinecode{EquiRec} and some others are only 22 lines
+in our implementation, because we only compose existing codes.
 
-To discover the reasons of slower execution time, we made some more experiments
-on two possible factors which could affect the performance.
-They are object algebra and the longest match alternative combinator.
-We use object algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
+To discover the reasons of slower execution time, we made experiments
+on two possible factors, which are Object Algebra and the longest match alternative combinator.
+We use Object Algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
 while \ilyaimpl{} uses case class and the ordinary alternative combinator.
+Therefore, we implemented two more versions. One is a modified version of our implementation,
+named \ourclass{}, with object algebra replaced by case class for the ASTs.
+The other is a modified version of \ilyaimpl{}, named \ilyalongest{},
+using the longest match alternative combinator instead of the ordinary one.
 
-We implemented two more versions. One is a modified version of our implementation, named \ourclass{}, with object algebra replaced by case class for the ASTs.
-The other is a modified version of \ilyaimpl{}, named \ilyalongest{}, using the longest match alternative combinator instead of the ordinary one.
+%% \begin{table*}
+%%     \centering
+%%     \begin{tabular}{|l|r|r|r|r|r|r|r|}
+%%       \hline
+%%         \multirow{2}{*}{\bfseries Calculus Name} & \ilyaimpl{} & \multicolumn{2}{ c| }{\ourimpl{}} & \multicolumn{2}{ c| }{\ilyalongest{}} & \multicolumn{2}{ c| }{\ourclass{}} \\ \cline{2-8}
+%%         \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{\bfseries Time} & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% \\
+%%       \hline
+%%         Arith & 741 & 913 & +23.2 & 793 & +7.0 & 932 & +25.8 \\
+%%         Untyped & 770 & 1018 & +32.2 & 821 & +6.6 & 1007 & +30.8 \\
+%%         FullUntyped & 1297 & 1854 & +42.9 & 1343 & +3.5 & 1767 & +36.2 \\
+%%         TyArith & 746 & 888 & +19.0 & 772 & +3.5 & 918 & +23.1 \\
+%%         SimpleBool & 1376 & 1782 & +29.5 & 1494 & +8.6 & 1824 & +32.6 \\
+%%         FullSimple & 1441 & 2270 & +57.5 & 1574 & +9.2 & 2226 & +54.5 \\
+%%         Bot & 1080 & 1287 & +19.2 & 1078 & -0.2 & 1306 & +20.9 \\
+%%       %\hline
+%%         %\multicolumn{1}{|c|}{\dots} & \multicolumn{7}{c|}{\dots} \\
+%%       %\hline
+%%         FullRef & 1438 & 2291 & +59.3 & 1544 & +7.4 & 2142 & +49.0 \\
+%%         FullError & 1410 & 1946 & +38.0 & 1524 & +8.1 & 1981 & +40.5 \\
+%%         RcdSubBot & 1247 & 1524 & +22.2 & 1285 & +3.0 & 1612 & +29.3 \\
+%%         FullSub & 1320 & 1979 & +49.9 & 1393 & +5.5 & 1899 & +43.9 \\
+%%         FullEquiRec & 1407 & 2200 & +56.4 & 1561 & +10.9 & 2156 & +53.2 \\
+%%         FullIsoRec & 1492 & 2253 & +51.0 & 1648 & +10.5 & 2236 & +49.9 \\
+%%         EquiRec & 994 & 1254 & +26.2 & 1048 & +5.4 & 1304 & +31.2 \\
+%%         Recon & 1044 & 1482 & +42.0 & 1128 & +8.0 & 1506 & +44.3 \\
+%%         FullRecon & 1094 & 1645 & +50.4 & 1161 & +6.1 & 1652 & +51.0 \\
+%%         FullPoly & 1398 & 2086 & +49.2 & 1511 & +8.1 & 2019 & +44.4 \\
+%%         FullOmega & 1451 & 2352 & +62.1 & 1582 & +9.0 & 2308 & +59.1 \\
+%%       \hline
+%%         Total & 21746 & 31024 & +42.7 & 23260 & +7.0 & 30795 & +41.6 \\
+%%       \hline
+%%         \multicolumn{8}{c}{}
+%%     \end{tabular}
+%%     \caption{Execution time of four implementations.}
+%%     \label{tab:ext-comparison}
+%% \end{table*}
 
-\begin{table*}
-    \centering
-    \begin{tabular}{|l|r|r|r|r|r|r|r|}
-      \hline
-        \multirow{2}{*}{\bfseries Calculus Name} & \ilyaimpl{} & \multicolumn{2}{ c| }{\ourimpl{}} & \multicolumn{2}{ c| }{\ilyalongest{}} & \multicolumn{2}{ c| }{\ourclass{}} \\ \cline{2-8}
-        \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{\bfseries Time} & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% & \bfseries Time & \bfseries (+/-)\% \\
-      \hline
-        Arith & 741 & 913 & +23.2 & 793 & +7.0 & 932 & +25.8 \\
-        Untyped & 770 & 1018 & +32.2 & 821 & +6.6 & 1007 & +30.8 \\
-        FullUntyped & 1297 & 1854 & +42.9 & 1343 & +3.5 & 1767 & +36.2 \\
-        TyArith & 746 & 888 & +19.0 & 772 & +3.5 & 918 & +23.1 \\
-        SimpleBool & 1376 & 1782 & +29.5 & 1494 & +8.6 & 1824 & +32.6 \\
-        FullSimple & 1441 & 2270 & +57.5 & 1574 & +9.2 & 2226 & +54.5 \\
-        Bot & 1080 & 1287 & +19.2 & 1078 & -0.2 & 1306 & +20.9 \\
-      %\hline
-        %\multicolumn{1}{|c|}{\dots} & \multicolumn{7}{c|}{\dots} \\
-      %\hline
-        FullRef & 1438 & 2291 & +59.3 & 1544 & +7.4 & 2142 & +49.0 \\
-        FullError & 1410 & 1946 & +38.0 & 1524 & +8.1 & 1981 & +40.5 \\
-        RcdSubBot & 1247 & 1524 & +22.2 & 1285 & +3.0 & 1612 & +29.3 \\
-        FullSub & 1320 & 1979 & +49.9 & 1393 & +5.5 & 1899 & +43.9 \\
-        FullEquiRec & 1407 & 2200 & +56.4 & 1561 & +10.9 & 2156 & +53.2 \\
-        FullIsoRec & 1492 & 2253 & +51.0 & 1648 & +10.5 & 2236 & +49.9 \\
-        EquiRec & 994 & 1254 & +26.2 & 1048 & +5.4 & 1304 & +31.2 \\
-        Recon & 1044 & 1482 & +42.0 & 1128 & +8.0 & 1506 & +44.3 \\
-        FullRecon & 1094 & 1645 & +50.4 & 1161 & +6.1 & 1652 & +51.0 \\
-        FullPoly & 1398 & 2086 & +49.2 & 1511 & +8.1 & 2019 & +44.4 \\
-        FullOmega & 1451 & 2352 & +62.1 & 1582 & +9.0 & 2308 & +59.1 \\
-      \hline
-        Total & 21746 & 31024 & +42.7 & 23260 & +7.0 & 30795 & +41.6 \\
-      \hline
-        \multicolumn{8}{c}{}
-    \end{tabular}
-    \caption{Execution time of four implementations.}
-    \label{tab:ext-comparison}
-\end{table*}
-
-Table \ref{tab:ext-comparison} shows comparison results of these four versions. It suggests that the difference of running time between
+The right part of table \ref{tab:comparison} suggests that the difference of running time between
 using object algebra and class is little, roughly 1\%.

### 2017-05-18 SLE_17/sections/Conclusion.tex
Shorten conclusion. Resize table.

\section{Conclusion}\label{sec:conclusion}

This paper presents a solution for type-safe modular parsing. Our solution
not only enables parsers to evolve together with the abstract syntax,
but also allows parsing code to be modularly type-checked and separately compiled.

We identify the algorithmic challenges of building modular parsers,
and use standard OO techniques including inheritance and overriding
for our goal. However, the extensibility issue of
traditional OO ASTs motivates us to adopt Object Algebras for full
extensibility and more useful features. Then language feature abstraction further enhances code reuse and modularity.
The TAPL case study demonstrates that a lot of boilerplate can be reduced by
modular parsing.

There are certainly some aspects that can be improved. We observed that the
glue code of composition appears to be boilerplate, for which family polymorphism~\cite{ernst01FP}
is a potential solution. Moreover, we can possibly
adopt the Shy framework~\cite{Zhang2015} and algebra composition
patterns~\cite{oliveira2013feature}, to improve the usage of Object
Algebras. For future work, it will be interesting to see how modular parsing
appears in functional programming languages, as they usually do not support subtyping
or inheritance. Potentially open recursion~\cite{CookThesis} can contribute.
%It is also interesting to see that parsing, to some
%extent, can be viewed as a special example of \textit{unfolds}. So it
%is worthwhile considering to generalize our approach under certain
%circumstances.

### 2017-05-18 SLE_17/sections/Introduction.tex
Shorten conclusion. Resize table.

@@
 Syntactic modularization techniques have also been applied to the
-problem of \emph{extensible parsing}. There are several approaches
-that enable the development of \emph{syntactically} modular parsers or
-grammars. Many parser
+problem of \emph{extensible parsing}. Many parser
 generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
@@
 their own syntax, such as SugarJ~\cite{Erdweg2011}, also offer a form
-of extensible parsing. However since those approaches are syntactic
-they do not support separate compilation and/or modular type-checking
+of extensible parsing. However those syntactic approaches do not
+support separate compilation and/or modular type-checking
 of parsing code either.

### 2017-05-18 SLE_17/sections/RelatedWork.tex
Shorten conclusion. Resize table.

@@
 all that the users need to do is to use case classes, instead of
-standard OO classes in their code.\haoyuan{To simplify; conclusion}
+standard OO classes in their code.
 

### 2017-05-29 SLE_17/sections/Inheritance.tex
A bit.

@@
 Scala via traits.
+
+\begin{comment}
 
@@
 
+\end{comment}
+
 \begin{comment}
@@
 
-To illustrate how to modularly add new language constructs and the
-corresponding parsing code, we continue with the expression language
-of literals and additions in Section~\ref{subsec:packratparsing}. We
-introduce variables as a new case:
+Part of the modular parsing problem is how to obtain an extensible parser. It
+is natural to make use of OO ASTs because adding new data constructs is cheap for
+them. Hence we have used OO traits and inheritance to represent the AST in the last
+section. Furthermore, we would like to write extensible parsing code on extensions of
+a grammar. That is to say, new extensions would not require modification on the existing code,
+and we can even reuse the old code.
 
-\setlength{\grammarindent}{5em}
-\begin{grammar}
-<expr> ::= ...
-   \alt <ident>
-\end{grammar}
-
+To illustrate such extensibility, we continue with the old
+example, and introduce variables as a new case.
 It is easy to extend the corresponding OO AST in a modular way:
@@
 
-Since we already have the parser for literals and additions, we would
-like to build the new parser by reusing the old one. In Scala we can take advantage of inheritance for such reuse. Specifically,
-the new parser can be defined in an enclosing trait that extends
+For code reuse, we want to define a new parser that extends
 the old one. Here one may quickly come up
@@
 
-\subsection{Multiple Extensions and Independent Extensibility}
-A nice feature of Scala is its support for a form of multiple
-inheritance.  Scala has a linearized-style mixin inheritance for
-traits~\cite{odersky2004overview}. This can be very helpful when
+\vspace{-.05in}
+\paragraph{Independent Extensibility}
+A nice feature of Scala is its support for the linearized-style multiple
+inheritance on traits~\cite{odersky2004overview}. This can be very helpful when
 composing several languages, and to achieve independent
-extensibility~\cite{odersky2005independently}.
-Suppose now we want to compose the parsers for expressions
-from pre-defined languages \inlinecode{LanguageA} and
-\inlinecode{LanguageB} using alternative. The new parser can be built
-by inheriting both parsers at the same time:
+extensibility~\cite{odersky2005independently}. Later we introduce language components
+in Section~\ref{subsec:language-component} as a nice pattern.
 
-\lstinputlisting[linerange=46-50]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE
-The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}.
-Under multiple inheritance, it can be used to distinguish the methods of the same name.
-Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.
+In the above code, we use the \lstinline{super} keyword to refer to the old \inlinecode{pExpr}.
+With multiple inheritance, Scala has a special syntax
+\lstinline{super[T].x}, called \textit{static super reference}. It refers to the type or
+method \inlinecode{x} in the specified parent trait \inlinecode{T}. With this we can reuse
+the old parsers even if they have the same name.
 
@@
 
-\paragraph{Conflicts and/or ambiguity} As we demonstrated before, the longest match alternative combinator is useful when composing parsers.
-However, ambiguous grammars still cause subtle issues in modular parsing.
-For example, if \inlinecode{parserA} and \inlinecode{parserB} accept a same string but generate different results, then the result of \inlinecode{parserA ||| parserB} really depends on the implementation of the alternative combinator, because the results can not be prioritized by length.
-Since the composition is handled explicitly by a user, conflicts could be resolved by overriding ambiguous parsing code, using the pattern discussed in Section~\ref{subsec:overriding-rules}.
+\vspace{-.05in}
+\paragraph{Conflicts and/or ambiguity} In a modular setting, conflicts and
+ambiguity could be introduced to the grammar. In that case, the help parser combinators
+can offer is quite restricted. Yet users can override those problematic methods to resolve
+such conflicts, and rely on dynamic dispatch. We will discuss it in Section~\ref{subsec:overriding-rules}.
 

### 2017-05-30 SLE_17/sections/Discussion.tex
Fit into 12 pages

@@
 Using Object Algebras, it is easy to
-model multi-sorted languages. If the syntax contains multiple sorts, we can distinguish them by different type parameters.
-For instance, now we extend the expression language from the end of Section~\ref{sec:algebrasandparsing}, with a
+model multi-sorted languages. If the syntax has multiple sorts, we can distinguish them by different type parameters.
+For instance, we extend the expression language from the end of Section~\ref{sec:algebrasandparsing}, with a
 primitive type \inlinecode{int} type and typed lambda abstractions:
+
+\vspace{0.8mm}
 
@@
 
-Figure~\ref{fig:multi} illustrates the corresponding Scala code
+\vspace{0.8mm}
+
+The code below illustrates the corresponding Scala code
 that extends the Object Algebra interface, pretty-printing operation and parser.
-We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types.
-The type system guarantees that invalid terms such as \inlinecode{int + int} will be rejected.
+
+\lstinputlisting[linerange=7-23]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
+
+We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types. The type system guarantees that invalid terms such as \inlinecode{int + int} will be rejected.
 Besides lexing, the trait \inlinecode{LamOAParser} also introduces parsers for types, and the new case for expressions.
 We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue
-with \inlinecode{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.
-
-\begin{figure}[ht]
-\lstinputlisting[linerange=7-23]{code/src/papercode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
-\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
-\label{fig:multi}
-\end{figure}
+with \inlinecode{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are used for recursion.
 

### 2017-05-30 SLE_17/sections/ObjectAlgebras.tex
Fit into 12 pages

@@
 difficulty of extending data structures or ASTs in two dimensions.
-As we have mentioned in Section~\ref{subsec:mismatch}, the use of traditional ASTs
-makes it hard to add new operations. In the last section we have seen a language
+In brief, it is hard to add new operations with traditional OO ASTs.
+In the last section we have seen a language
 that supports pretty-printing (in \inlinecode{Expr}).
@@
 \paragraph{A Solution}
-A solution to this problem is instead having a field with the Object Algebra
-in \lstinline{OAParser}. This approach supports overriding as well.
-The parsing code is as follows:
+A solution to this problem is to declare a field of Object Algebra interface
+in the parser. Figure~\ref{fig:oa-parser} shows the code of true modular parser,
+whose methods can be overridden for future extension.
 
+\begin{figure}
   \lstinputlisting[linerange=21-30]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
-This is precisely the pattern that we advocate for modular parsing.
+  \caption{Pattern of modular parsing using Object Algebras.}
+  \label{fig:oa-parser}
+  \vspace{-0.2cm}
+\end{figure}
+
+That is precisely the pattern that we advocate for modular parsing.
 One important remark is we introduce \inlinecode{pE} for recursive calls.
-The reason why we use an extra and seemingly redundant field, is due to a subtle issue caused by Scala language and its parser combinator library.
+The reason why we use it as an extra and seemingly redundant field, is due to a subtle issue caused by Scala language and its parser combinator library.
 There is a restriction of \inlinecode{super} keyword in Scala that \inlinecode{super} can only use methods defined by keyword \inlinecode{def}, but cannot access fields defined by \inlinecode{val}, while the parser combinator library suggests using \inlinecode{val} to define parsers, especially for left-recursive ones.
@@
 Note that, similar to the approach in Section~\ref{sec:inheritance}, independent extensibility is also supported via multiple trait inheritance.
-Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance}, we omit that code here.
+Since it is achieved using essentially the same technique as in Section~\ref{sec:inheritance}, we omit the code here.

### 2017-05-30 SLE_17/sections/RelatedWork.tex
Fit into 12 pages

@@
 compilers like JastAdd~\cite{Ekman2007} and
-Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
+Polyglot~\cite{Nystrom2003} also support extensible parsing, but it
 is mostly done using parser generators. They focus on the
@@
 
-Some more recent work on parser
+Some recent work on parser
 combinators~\cite{Ford2002,Might2011,Frost2008} proposed a series of
 novel parsing techniques that address the issue of
-left-recursion. We have selected Packrat parsing due to its simplicity in Scala,
-but in general it can also have alternatives.
+left-recursion. We chose Packrat parsing due to its simplicity in Scala,
+but in general there are alternatives to it.
 

### 2017-06-02 ECOOP_17/sections/Conclusion.tex

\section{Conclusion}\label{sec:conclusion}

This paper presents a solution for type-safe modular parsing. Our solution 
not only enables parsers to evolve together with the abstract syntax, 
but also allows parsing code to be modularly type-checked and separately compiled.

Based on parser combinators, we identify algorithmic
challenges of building modular parsers, and show that Packrat parsing
is suitable as the underlying parsing technique. Our solution does not
require advanced language features. We show that with standard OO
techniques including inheritance and overriding, it is practical to
build modular parsers for OO ASTs. However, the extensibility issue of
traditional OO ASTs motivates us to adopt Object Algebras for full
extensibility and more useful features.

Abstracting language features as reuseable components can be achieved
based on our solution, which is useful for rapid language prototyping.
The TAPL case study shows that we can obtain considerable code reuse
(69\% shorter) compared with a non-modular implementation, using our
modular parsing approach and language feature abstraction.

There are certainly some aspects that can be improved. We observed that the
glue code of composition appears to be boilerplate. Such an issue
is addressed by family polymorphism~\cite{ernst01FP}, and we could solve it by language
features or meta-programming techniques. Moreover, we can possibly
adopt the Shy framework~\cite{Zhang2015} and algebra composition
patterns~\cite{oliveira2013feature}, to improve the usage of Object
Algebras.

For future work, we want to experiment more with semantically modular
parsing techniques for functional programming. Since functional
programming languages like Haskell and ML do not support OO features 
such as subtyping or inheritance, it will be interesting to see
whether other functional features can be used instead. 
Open recursion~\cite{CookThesis} seems to be an alternative to inheritance, and may 
help in doing extensible parsing in functional languages. 
%It is also interesting to see that parsing, to some
%extent, can be viewed as a special example of \textit{unfolds}. So it
%is worthwhile considering to generalize our approach under certain
%circumstances.

### 2017-06-05 SLE_17/paper.tex
More.
A bit.
Fit into 12 pages
Revised abstract and intro

@@
 \documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=false}
+\settopmatter{printacmref=false}
 %% For single-blind review submission
@@
 \usepackage{tabularx}
+
+\setlength{\abovecaptionskip}{1.5mm}
+\setlength{\belowcaptionskip}{1.5mm}
+
+\setlength{\textfloatsep}{2mm}
 
@@
 %% above the permission/copyright statement.
-\thanks{with paper note}                %% \thanks is optional
+%%\thanks{with paper note}                %% \thanks is optional
                                         %% can be repeated if necesary
@@
   present a technique based on parser combinators that enables modular
-  parsing. Interestingly, the modularity requirements for modular
-  parsing rule out several existing parser combinator approaches,
-  which rely on some non-modular techniques. We show that Packrat
+  parsing. 
+  %Interestingly, the modularity requirements for modular
+  %parsing rule out several existing parser combinator approaches,
+  %which rely on some non-modular techniques. 
+  We show that Packrat
   parsing techniques, provide solutions for such modularity problems,
@@
 %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+
 \begin{CCSXML}
@@
 
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+%\ccsdesc[500]{Software and its engineering~General programming languages}
+%\ccsdesc[300]{Social and professional topics~History of programming languages}
 %% End of generated code
@@
 %% comma separated list
-\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional
+%\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional
 
@@
 %% Appendix
-\appendix
-\section{Appendix}
+%% \appendix
+%% \section{Appendix}
 
-Text of appendix \ldots
+%% Text of appendix \ldots
 

### 2017-06-05 SLE_17/sections/Introduction.tex
Paper
Revised abstract and intro

@@
 composition techniques such as \emph{superimposition}~\cite{AK:JOT09} to
-enable the development modular program features. Such
-composition techniques collect the code for multiple features and
-merge them together when a concrete combination of features is needed
-for a particular program. As Kastner et. al~\cite{Kastner11road} note,
+enable the development modular program features. 
+%Such
+%composition techniques collect the code for multiple features and
+%merge them together when a concrete combination of features is needed
+%for a particular program. 
+As Kastner et. al~\cite{Kastner11road} note,
 a typical drawback of feature-oriented SPL implementations, which
@@
   at:
-  \url{http://}\footnote{\textbf{Note to Reviewers:} Please find the URL in the
-    supplementary material for the submission.}
+  \url{http://}\bruno{Please add the link here!}
 
@@
 All the code in the paper and the case study is written in Scala, since its concise and elegant
-syntax gives better presentation. Clearly
-other languages that support some form of multiple inheritance
-(including Java 8 with default methods~\cite{Goetz2012}) could in principle be
+syntax gives better presentation. Other languages that support some form of multiple inheritance
+(including C++ or Java 8 with default methods~\cite{Goetz2012}) could in principle be
 used.

### 2017-06-05 SLE_17/sections/Inheritance.tex
Fit into 12 pages
Section 3

@@
 
-\section{Parsing OO ASTs with Multiple Inheritance}\label{sec:inheritance}
+\section{\hspace{-10pt}OO AST Parsing with Multiple Inheritance}\label{sec:inheritance}
 
@@
 section. Furthermore, we would like to write extensible parsing code on extensions of
-a grammar. That is to say, new extensions would not require modification on the existing code,
+a grammar. That is to say, new extensions would not require modifying the existing code,
 and we can even reuse the old code.
@@
 hence the parser does not work as expected. It is possible to build the correct parser by replacing the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr}.
-However, modification on existing code sacrifices separate compilation, and hence breaks semantic modularity.
+However, modifying existing code sacrifices separate compilation, and hence breaks semantic modularity.
 
@@
 With multiple inheritance, Scala has a special syntax
-\lstinline{super[T].x}, called \textit{static super reference}. It refers to the type or
-method \inlinecode{x} in the specified parent trait \inlinecode{T}. With this we can reuse
+\lstinline{super[T].x}, called \textit{static super reference}. It refers to the name
+\inlinecode{x} in the specified parent trait \inlinecode{T}. With this we can reuse
 the old parsers even if they have the same name.
+\bruno{Independent extensibility cannot be explained like this. 
+We need to see some code here! You need to expand the explanation. 
+I suggest instead cutting on the ``wrong'' attempt in the beggining of this section, and instead 
+have some brief discussion in the \emph{correct} code where we need to be careful.}
 

### 2017-06-05 SLE_17/sections/PackratParsing.tex
Fit into 12 pages
Section 3

@@
 Traditional alternative composition will only find the first parser that succeeds on a prefix of the input,
-in spite that subsequent parsers may parse the whole input. In contrast with the parser above, \inlinecode{"parseAdd <|> parseInt"} works as expected with the two cases swapped.
-
+even if subsequent parsers may parse the whole input. In contrast to the previous parser, \inlinecode{"parseAdd <|> parseInt"} works as expected with because the two cases are swapped.
 In this case, reordering the alternatives ensures that
@@
 However, manual reordering for the longest match is inconvenient, and worst still, it is essentially non-modular.
-When the grammar is extended with new rules, programmers are supposed to \emph{manually} adjust
+When the grammar is extended with new rules, programmers should \emph{manually} adjust
 the order of parsers, by rewriting previously written code.
@@
 \paragraph{Backtracking} The need for backtracking can also be problematic
-in a modular setting. Consider a grammar that includes \inlinecode{"import..from"}.
-It is extended with an \inlinecode{"import..as"} case:
+in a modular setting. Consider a grammar that includes \inlinecode{"import..from"}, 
+and is extended with an \inlinecode{"import..as"} case:
 
@@
 Similarly, this violates a modular setting because it also requires a global view of the full grammar.
-Hence the worst case should be considered that all alternatives may share common prefixes with future cases.
-In that case we need to backtrack for all the branches.
+Hence the worst case where all alternatives may share common prefixes with future cases
+should always be antecipated.
+Therefore we need to backtrack for all the branches.
 To avoid failures in the future, we have to add \inlinecode{try} everywhere.
@@
  The longest match composition is also employed by using the combinator \lstinline{|||}.
-Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees reasonable efficiency.
+Furthermore, the parser does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees reasonable efficiency.
 

### 2017-06-05 SLE_17/sections/CaseStudy.tex
Shorten conclusion. Resize table.
Fit into 12 pages
Revision

@@
     \centering
-    \includegraphics[width=\textwidth]{resources/depGraph.pdf}
+    \includegraphics[width=0.8\textwidth]{resources/depGraph.pdf}
     \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
@@
 
-From the graph, we know that common components such as \lstinline{VarApp} are reused in
-lots of calculi. Such reuse could shorten the code considerably.
-Next subsection examines amount of code as well as performance.
+%% From the graph, we know that common components such as \lstinline{VarApp} are reused in
+%% lots of calculi. Such reuse could shorten the code considerably.
+%% Next subsection examines amount of code as well as performance.
 
@@
     \centering
+    \resizebox{0.7\textheight}{!}{
     \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|}
@@
       \multicolumn{7}{c}{}
-    \end{tabular}
+    \end{tabular}}
     \caption{Comparison of SLOC and execution time.}
     \label{tab:comparison}
+    \vspace{-0.3cm}
 \end{table*}
@@
 To discover the reasons of slower execution time, we made experiments
-on two possible factors, which are Object Algebra and the longest match alternative combinator.
-We use Object Algebra for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
+on two possible factors, which are Object Algebras and the longest match alternative combinator.
+We use Object Algebras for ASTs and the longest match alternative combinator \inlinecode{|||} for parsing,
 while \ilyaimpl{} uses case class and the ordinary alternative combinator.
 Therefore, we implemented two more versions. One is a modified version of our implementation,
-named \ourclass{}, with object algebra replaced by case class for the ASTs.
+named \ourclass{}, with Object Algebras replaced by case class for the ASTs.
 The other is a modified version of \ilyaimpl{}, named \ilyalongest{},
@@
 The right part of table \ref{tab:comparison} suggests that the difference of running time between
-using object algebra and class is little, roughly 1\%.
+using Object Algebras and class is little, roughly 1\%.
 The usage of longest match combinator slows the performance by 7\%. The main reason of slower

### 2017-06-05 SLE_17/sections/Discussion.tex
Revision

@@
 Regarding OO inheritance for the extensibility, an alternative approach, called \textit{open recursion}~\cite{CookThesis} can be used in other languages, by introducing explicit ``self-reference'' parameters for the recursion.
-Furthermore, besides Object Algebras, v\textit{Data types à la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
+Furthermore, besides Object Algebras, \textit{Data types \`a la carte} (DTC)~\cite{swierstra2008data} and the Cake pattern~\cite{odersky2005independently} also support extensible data structures.
 For the goal of modular parsing a custom combination of those alternatives can be adopted.

### 2017-06-05 SLE_17/paper.tex
Author info and a footnote in case study

@@
 %% For double-blind review submission
-\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=false}
-\settopmatter{printacmref=false}
+%% \documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=false}
 %% For single-blind review submission
-%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
+\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
 %% For final camera-ready submission
@@
 %% 'acmlarge'.
+
+\settopmatter{printacmref=false}
 
@@
 %% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
+\author{Haoyuan Zhang}
+\email{hyzhang@cs.hku.hk}          %% \email is recommended
+
+\affiliation{
+  \institution{The University of Hong Kong}            %% \institution is required
+}
+
+\author{Huang Li}
+\email{hli@cs.hku.hk}          %% \email is recommended
+%% \authornote{with author1 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
 \affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}
+  %% \position{Position1}
+  %% \department{Department1}              %% \department is recommended
+  \institution{The University of Hong Kong}            %% \institution is required
+  %% \streetaddress{Street1 Address1}
+  %% \city{City1}
+  %% \state{State1}
+  %% \postcode{Post-Code1}
+  %% \country{Country1}
 }
-\email{first1.last1@inst1.edu}          %% \email is recommended
 
+\author{Bruno C. d. S. Oliveira}
+\email{bruno@cs.hku.hk}          %% \email is recommended
+
+\affiliation{
+  \institution{The University of Hong Kong}            %% \institution is required
+}
 %% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
+%% \author{First2 Last2}
+%% \authornote{with author2 note}          %% \authornote is optional;
+%%                                         %% can be repeated if necessary
+%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+%% \affiliation{
+%%   \position{Position2a}
+%%   \department{Department2a}             %% \department is recommended
+%%   \institution{Institution2a}           %% \institution is required
+%%   \streetaddress{Street2a Address2a}
+%%   \city{City2a}
+%%   \state{State2a}
+%%   \postcode{Post-Code2a}
+%%   \country{Country2a}
+%% }
+%% \email{first2.last2@inst2a.com}         %% \email is recommended
+%% \affiliation{
+%%   \position{Position2b}
+%%   \department{Department2b}             %% \department is recommended
+%%   \institution{Institution2b}           %% \institution is required
+%%   \streetaddress{Street3b Address2b}
+%%   \city{City2b}
+%%   \state{State2b}
+%%   \postcode{Post-Code2b}
+%%   \country{Country2b}
+%% }
+%% \email{first2.last2@inst2b.org}         %% \email is recommended
 

### 2017-06-07 SLE_17/paper.tex
12 pages.

@@
 
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
 %% Bibliography

### 2017-06-07 SLE_17/sections/Inheritance.tex
12 pages.

@@
 
-\subsection{Extensible Parsing via Inheritance}
-
 Part of the modular parsing problem is how to obtain an extensible parser. It
@@
 example, and introduce variables as a new case.
-It is easy to extend the corresponding OO AST in a modular way:
+It is easy to extend the corresponding OO AST together with its parser in a modular way:
 
-\lstinputlisting[linerange=7-9]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM
+\begin{lstlisting}
+class Var(x: String) extends Expr {
+  def print = x
+}
+\end{lstlisting}
+Here one may quickly define a new parser \lstinline|pVar| in \lstinline|AParser| for variables, and parse new expressions
+with \lstinline{"pExpr ||| pVar"}.
+Unfortunately, even \inlinecode{"1 + x"} cannot be parsed, which is obviously valid in the new grammar.
+The reason is that \inlinecode{pAdd} makes two recursive calls to \lstinline|pExpr| for parsing sub-expressions, whereas the newly added \inlinecode{pVar} is not observed, unless we replace all the occurrences of \lstinline|pExpr| with \lstinline{"pExpr ||| pVar"}.
+Yet modifying existing code breaks semantic modularity.
 
-For code reuse, we want to define a new parser that extends
-the old one. Here one may quickly come up
-with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
-with \inlinecode{pExpr}:
-
-\lstinputlisting[linerange=13-16]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT
-
-Unfortunately, this fails to parse some expressions like \inlinecode{"1 + x"}, which are obviously valid in the new grammar.
-The reason is that \inlinecode{pAdd} makes two recursive calls to parse sub-expressions, by using \inlinecode{pExpr}, which
-covers both cases in the old grammar. But the newly added case \inlinecode{pVar} is not observed by the recursive \inlinecode{pExpr},
-hence the parser does not work as expected. It is possible to build the correct parser by replacing the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr}.
-However, modifying existing code sacrifices separate compilation, and hence breaks semantic modularity.
-
-\subsection{Overriding for Extensibility}\label{subsec:overriding}
-
+\vspace{-.05in}
+\paragraph{Overriding for Extensibility}
 It is actually quite simple to let \inlinecode{pExpr} cover the newly
@@
 composing several languages, and to achieve independent
-extensibility~\cite{odersky2005independently}. Later we introduce language components
-in Section~\ref{subsec:language-component} as a nice pattern.
+extensibility~\cite{odersky2005independently}. Suppose now we want to compose the parsers
+for expressions from pre-defined languages \lstinline|LanguageA| and \lstinline|LanguageB| using alternative.
+The new parser can be built by inheriting both parsers at the same time:
+\begin{lstlisting}
+trait LanguageA {...}
+trait LanguageB {...}
+trait LanguageC extends LanguageA with LanguageB {
+  override def pExpr = super[LanguageA].pExpr ||| super[LanguageB].pExpr
+}
+\end{lstlisting}
 
-In the above code, we use the \lstinline{super} keyword to refer to the old \inlinecode{pExpr}.
-With multiple inheritance, Scala has a special syntax
-\lstinline{super[T].x}, called \textit{static super reference}. It refers to the name
-\inlinecode{x} in the specified parent trait \inlinecode{T}. With this we can reuse
-the old parsers even if they have the same name.
-\bruno{Independent extensibility cannot be explained like this. 
-We need to see some code here! You need to expand the explanation. 
-I suggest instead cutting on the ``wrong'' attempt in the beggining of this section, and instead 
-have some brief discussion in the \emph{correct} code where we need to be careful.}
+The \lstinline{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method
+\inlinecode{x} in the parent trait \inlinecode{T}. Under multiple inheritance, it can be used to distinguish
+the methods of the same name. Therefore in the new parser, we use \lstinline{super} to specify which
+\lstinline{pExpr} we are referring to.
 
@@
 \vspace{-.05in}
-\paragraph{Conflicts and/or ambiguity} In a modular setting, conflicts and
+\paragraph{Conflicts and/or Ambiguity} In a modular setting, conflicts and
 ambiguity could be introduced to the grammar. In that case, the help parser combinators

### 2017-06-08 SLE_17/sections/CaseStudy.tex

@@
 To demonstrate the utility of our modular parsing approach, we
-implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages} (TAPL) \cite{pierce2002types}. We compared our implementation with a non-modular implementation
+implemented parsers of the first 18 calculi~\footnote{There are some more calculi in the book,
+  but they are either not ported by the implementation we compare with, or just repeating the syntax of former ones.}
+from book \textit{Types and Programming Languages} (TAPL) \cite{pierce2002types}.
+We compared our implementation with a non-modular implementation
 we found online, which is also written in Scala and uses the same Packrat parsing library.

### 2017-06-08 SLE_17/sections/CaseStudy.tex
revision

--- 2017-06-05 SLE_17/sections/CaseStudy.tex
@@
 To demonstrate the utility of our modular parsing approach, we
-implemented parsers of the first 18 calculi from book \textit{Types and Programming Languages} (TAPL) \cite{pierce2002types}. We compared our implementation with a non-modular implementation
-we found online, which is also written in Scala and uses the same Packrat parsing library.
+implemented parsers for the first 18 calculi~\footnote{There are some more calculi in the book,
+  but they are either not ported by the implementation we compare with, or just repeats the syntax of former ones.}
+from the \textit{Types and Programming Languages} (TAPL)
+\cite{pierce2002types} book.
+We compared our implementation with a non-modular implementation
+available online, which is also written in Scala and uses the same Packrat parsing library.
 We counted source lines of code (SLOC) and measured execution time for both implementations.
-The result suggests that our implementation saves 69\% code comparing with that non-modular one.
+The result suggests that our implementation saves 69\% code comparing
+with that non-modular one, but there is a 43\% slowdown due to code modularity.
 
@@
 written in Scala using the same parser combinator library.
-Furthermore, it includes parsers of all the 18 calculi we have, but
-written in a non-modular way. Thus it is not able to reuse existing
+\ilyaimpl{} implements parsers 18 calculi in TAPL in a non-modular
+way. Thus \ilyaimpl{} is not able to reuse existing
 code when those calculi share common features.
+\ourimpl{} implements the same 18 calculi, but reuse 
+is possible due to modularity.
 
@@
 Table \ref{tab:comparison} shows results of the comparison.
-Let's only check \ourimpl{} and \ilyaimpl{} for now.
+Let us only check \ourimpl{} and \ilyaimpl{} for now.
 The overall result is that 69.2\% of code is reduced using our
@@
 The good SLOC result is because of that the code of common language features
-are reused lots of times in the whole case study. We can see that in the first two calculi
+are reused many times in the whole case study. We can see that in the first two calculi
 \lstinline{Arith} and \lstinline{Untyped} we are not better than \ilyaimpl{},
@@
 In particular, the calculi \inlinecode{EquiRec} and some others are only 22 lines
-in our implementation, because we only compose existing codes.
+in our implementation, because we only compose existing code.
 
@@
 slower, the algorithmic complexity is still the same.
+Since the slowdown seems to be caused by extra method
+dispatching, in future work we wish to investigate techniques such as 
+partial evaluation or meta-programming to eliminate such cost. The
+work by B{\'e}guet and Manohar~\cite{Beguet:2014} is an interesting starting point.

### 2017-06-08 SLE_17/sections/ObjectAlgebras.tex
revision

@@
 To match the new ASTs, we have to substitute \inlinecode{NewExpr} for \inlinecode{Expr} (the
-same for \inlinecode{Lit}, \inlinecode{Add}, ...). It requires either modification on existing code or casts.
+same for \inlinecode{Lit}, \inlinecode{Add}, ...). This requires either code modification or type casts.
 The goal of semantic modularity motivates us to find a different approach for building ASTs.
@@
 Such a parser looks fine, but it is not extensible.
-For example, we have demonstrated in Section~\ref{subsec:overriding} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
+For example, we have demonstrated in Section~\ref{sec:inheritance} that method overriding is essential to update \inlinecode{pExpr} for an extended syntax.
 However, trying to do a similar method overriding for \inlinecode{pExpr} would require a type \inlinecode{VarAlg[E] =>} \inlinecode{Parser[E]}, which is a \emph{supertype} of the old type \inlinecode{Alg[E] =>} \inlinecode{Parser[E]}, since

### 2017-06-08 SLE_17/sections/Introduction.tex
deleted one paragraph

@@
 
+\begin{comment}
 Syntactic modularization techniques have also been applied to the
@@
 of parsing code either.
+\end{comment}
 
@@
 problem of how to modularize parsing has not been studied in semantic
-modularization approaches. At best, techniques such as NOA~\cite{Gouseti2014}
+modularization approaches. Techniques such as NOA~\cite{Gouseti2014}
 employ a syntactic modularity approach for parsing in combination with
@@
   at:
-  \url{http://}\bruno{Please add the link here!}
+  \url{https://github.com/lihuanglx/modular-parsing}
 

### 2017-06-08 SLE_17/sections/RelatedWork.tex
Revision
Related work.

@@
 \paragraph{Syntactically Extensible Parsing}
-Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}
+Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016}
 are a mainstream area of modular syntax and parsing. They allow users to write
@@
 far as we are aware, no extensible parser generators support separate
-compilation or modular type-checking.
+compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
+grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
+by a typed parser generator. Nevertheless, it only support separate compilation
+on grammar fragments; a whole compilation is still needed for an extended grammar. And it requires
+a lot of advanced language features, making client code less expressive. Our approach is simple from a
+straightforward use of OO programming, and makes parsing code directly reusable.
 
@@
 
-\paragraph{Extensible Parsing Algorithm}
+\paragraph{Extensible Parsing Algorithms}
 \textit{Parse table composition}~\cite{bravenboer2008parse}

### 2017-06-10 SLE_17/sections/PackratParsing.tex
Revision

@@
 
-Moreover, the underlying parsing formalism should make backtracking
-cheap, since \emph{backtracking is pervasive in modular parsing}.
-Although we chose Packrat parsing (due to its immediate availability
-and support), any other parsing formalism that provides similar
-features should be adequate for modular parsing.
+Moreover, the underlying parsing formalism should make \emph{backtracking
+cheap}, due to its pervasiveness in modular parsing.
+Although we chose Packrat parsing, any other parsing formalism that provides similar
+features should be ok.
 

### 2017-06-10 SLE_17/sections/RelatedWork.tex
Revision

@@
 Our work touches upon several topics including extensible parsing,
-parser combinators and extensibility techniques. However, as far as we
-know there's no work that discusses how to do statically type-safe and
-separately compilable modular parsing.
+parser combinators and extensibility techniques. 
+%However, as far as we
+%know there's no work that discusses how to do statically type-safe and
+%separately compilable modular parsing.
 
@@
 grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
-by a typed parser generator. Nevertheless, it only support separate compilation
-on grammar fragments; a whole compilation is still needed for an extended grammar. And it requires
-a lot of advanced language features, making client code less expressive. Our approach is simple from a
-straightforward use of OO programming, and makes parsing code directly reusable.
+by a typed parser generator. Nevertheless this requires
+a lot of advanced language features, making client complex. Our
+approach is simple and a straightforward use of OO programming, 
+and makes parsing code directly reusable.
 

### 2017-06-14 SLE_17/sections/CaseStudy.tex
Spellcheck

@@
 more intermediate function calls and method overriding. However, it is worth mentioning that
-because of the memoization technique of Packrat parsers, we are only constant times
+because of the memorization technique of Packrat parsers, we are only constant times
 slower, the algorithmic complexity is still the same.

### 2017-06-14 SLE_17/sections/Introduction.tex
Revision
Spellcheck

@@
 \item \textbf{TAPL case study:} We conduct a case study with 18 interpreters
-  from the TAPL book. The case study show the effectiveness of modular
+  from the TAPL book. The case study shows the effectiveness of modular
   parsing in terms of reuse. The TAPL case study is available online
   at:
+
   \url{https://github.com/lihuanglx/modular-parsing}
@@
 All the code in the paper and the case study is written in Scala, since its concise and elegant
-syntax gives better presentation. Other languages that support some form of multiple inheritance
+syntax is good for presentation. Other languages that support some form of multiple inheritance
 (including C++ or Java 8 with default methods~\cite{Goetz2012}) could in principle be

### 2017-06-14 SLE_17/sections/PackratParsing.tex
Spellcheck

@@
 Hence the worst case where all alternatives may share common prefixes with future cases
-should always be antecipated.
+should always be anticipated.
 Therefore we need to backtrack for all the branches.
@@
 parsing~\cite{Ford2002} have been developed to address limitations of
-simple parser combinators. Packrat parsing uses memoization to record
+simple parser combinators. Packrat parsing uses memorization to record
 the result of applying each parser at each position of the input, so
@@
  The longest match composition is also employed by using the combinator \lstinline{|||}.
-Furthermore, the parser does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees reasonable efficiency.
+Furthermore, the parser does not suffer from the backtracking problem, as the memorization technique of Packrat parsing guarantees reasonable efficiency.
 

### 2017-06-15 SLE_17/sections/CaseStudy.tex
Spell check. Bibtex.

--- 2017-06-08 SLE_17/sections/CaseStudy.tex
@@
 
-The right part of table \ref{tab:comparison} suggests that the difference of running time between
+The right part of Table \ref{tab:comparison} suggests that the difference of running time between
 using Object Algebras and class is little, roughly 1\%.

### 2017-09-04 SLE_17/paper.tex
Working on final version.

@@
 %% For single-blind review submission
-\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
+\documentclass[sigplan,screen]{acmart}\settopmatter{}
 %% For final camera-ready submission
@@
 %% 'acmlarge'.
-
-\settopmatter{printacmref=false}
 
@@
 
-\makeatletter\if@ACM@journal\makeatother
-%% Journal information (used by PACMPL format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{1}
-\acmArticle{1}
+%%% The following is specific to SLE'17 and the paper
+%%% 'Type-Safe Modular Parsing'
+%%% by Haoyuan Zhang, Huang Li, and Bruno C. d. S. Oliveira.
+%%%
+%\setcopyright{ACMCT}
+\setcopyright{acmcopyright}
+\acmPrice{15.00}
+\acmDOI{10.1145/3136014.3136016}
 \acmYear{2017}
-\acmMonth{1}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\else\makeatother
-%% Conference information (used by SIGPLAN proceedings format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
-\acmYear{2017}
-\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\fi
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission
-\setcopyright{none}             %% For review submission
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2017}           %% If different from \acmYear
+\copyrightyear{2017}
+\acmISBN{978-1-4503-5525-4/17/10}
+\acmConference[SLE'17]{2017 ACM SIGPLAN International Conference on Software Language Engineering}{October 23--24, 2017}{Vancouver, Canada}
 
@@
 	<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
+	<concept_id>10011007.10011006.10011008.10011009.10011011</concept_id>
+	<concept_desc>Software and its engineering~Object oriented languages</concept_desc>
 	<concept_significance>500</concept_significance>
@@
 	<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+	<concept_id>10003752.10010124.10010125.10010128</concept_id>
+	<concept_desc>Theory of computation~Object oriented constructs</concept_desc>
 	<concept_significance>300</concept_significance>
@@
 
-%\ccsdesc[500]{Software and its engineering~General programming languages}
-%\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
+\ccsdesc[500]{Software and its engineering~Object oriented languages}
+\ccsdesc[300]{Theory of computation~Object oriented constructs}
 
@@
 %% comma separated list
-%\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional
+\keywords{modular parsing, object algebras, semantic modularity}  %% \keywords is optional
 
@@
 
+\section*{Acknowledgements}
+We would like to thank the anonymous reviewers for their helpful comments. This work has been sponsored by the Hong Kong Research Grant Council projects number 27200514 and 17258816.
+
 %% Bibliography

### 2017-09-04 SLE_17/sections/Introduction.tex
Working on final version.

@@
 Lines (SPLs)~\cite{AK:JOT09,Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
-generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
+generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016,schwerdfeger09}.  Most syntactic approaches employ textual
 composition techniques such as \emph{superimposition}~\cite{AK:JOT09} to

### 2017-09-04 SLE_17/sections/RelatedWork.tex
Working on final version.

@@
 \paragraph{Syntactically Extensible Parsing}
-Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016}
+Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016,schwerdfeger09}
 are a mainstream area of modular syntax and parsing. They allow users to write
@@
 
+Besides, type-specific languages~\cite{omar14} do support
+separate compilation and type safety. The modularity of parsing is incorporated in its type
+system, which allows users to introduce new syntax for parsing. Whereas in our approach, parsing 
+strategy is an orthogonal issue since we rely on an external parsing library.
+Our approach does not rely on an elaborated semantics extension, but simply uses a pattern
+that is applicable in mainstream OO languages, where the type system helps ensure
+type safety.
+
 %\bruno{Have you read
@@
 \paragraph{Extensible Parsing Algorithms}
-\textit{Parse table composition}~\cite{bravenboer2008parse}
+\textit{Parse table composition}~\cite{bravenboer2008parse, schwerdfeger2010}
 is an approach where grammars are compiled to
@@
 \emph{separate compilation} as well as \emph{modular
-  type-checking}. Moreover, the extensibility of parsing is further
-available at language composition and lexical level.
+  type-checking}, and is applied in common OO languages. Moreover, the extensibility of parsing is further
+available at language composition.
 

### 2017-09-04 SLE_17/sections/Inheritance.tex
shortening.

@@
 
+\vspace{-4pt}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\vspace{-4pt}
+
 Here one may quickly define a new parser \lstinline|pVar| in \lstinline|AParser| for variables, and parse new expressions
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=20-26]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
+\vspace{-4pt}
 
@@
 The new parser can be built by inheriting both parsers at the same time:
+\vspace{-4pt}
 \begin{lstlisting}
@@
 }
-\end{lstlisting}
+\end{lstlisting}\vspace{-4pt}
 

### 2017-09-04 SLE_17/sections/ObjectAlgebras.tex
shortening.

@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=23-23]{code/src/papercode/Sec4OA/Code1.scala}% APPLY:linerange=FVARSEXPR
+\vspace{-4pt}
 
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=5-8]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
+\vspace{-4pt}
 Here \inlinecode{Alg} is called an \textit{Object Algebra interface}, parameterized by the type
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=12-16]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
+\vspace{-4pt}
 Here \inlinecode{Print} is called an \textit{Object Algebra}. It traverses an expression bottom-up, and returns a string as the result.
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=27-29]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+\vspace{-4pt}
 Now pretty-printing on the new language can be realized without modifying existing code:
 
+\vspace{-4pt}
 \lstinputlisting[linerange=33-35]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+\vspace{-4pt}
 An observation is that only the new case is implemented for pretty-printing, and the others have been inherited.
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=42-43]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+\vspace{-4pt}
 
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=8-17]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+\vspace{-4pt}
 Such a parser looks fine, but it is not extensible.
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=34-39]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+\vspace{-4pt}
 
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=46-49]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
+\vspace{-4pt}
 

### 2017-09-04 SLE_17/sections/PackratParsing.tex
Spell check. Bibtex.
shortening.

--- 2017-06-10 SLE_17/sections/PackratParsing.tex
@@
+\vspace{-2pt}
 \section{Packrat Parsing for Modularity}\label{sec:packrat}
@@
 Consider the grammar:
+\vspace{-2pt}
 \setlength{\grammarindent}{5em}
@@
 \end{grammar}
+\vspace{-2pt}
 
@@
 \paragraph{Backtracking} The need for backtracking can also be problematic
-in a modular setting. Consider a grammar that includes \inlinecode{"import..from"}, 
+in a modular setting. Consider a grammar with \inlinecode{"import..from"}, 
 and is extended with an \inlinecode{"import..as"} case:
 
+\vspace{-2pt}
 \setlength{\grammarindent}{5em}
@@
 \end{grammar}
+\vspace{-2pt}
 Since the two cases share a common prefix, when the former fails, we must backtrack to the beginning.
@@
 
+\vspace{-4pt}
 \begin{lstlisting}[language=PlainCode]
@@
 \end{lstlisting}
+\vspace{-4pt}
 
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=9-15]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
+\vspace{-4pt}
 
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=19-26]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
+\vspace{-4pt}
 
@@
 
+\vspace{-4pt}
 \lstinputlisting[linerange=6-7]{code/src/papercode/Sec2Packrat/Code2.scala}% APPLY:linerange=PACKRAT_RUNPARSER
+\vspace{-10pt}

### 2017-09-04 SLE_17/sections/RelatedWork.tex
shortening.

@@
+\vspace{-3pt}
 \section{Related Work}\label{sec:relatedwork}
@@
 system, which allows users to introduce new syntax for parsing. Whereas in our approach, parsing 
-strategy is an orthogonal issue since we rely on an external parsing library.
+strategy is based on an external parsing library.
 Our approach does not rely on an elaborated semantics extension, but simply uses a pattern
-that is applicable in mainstream OO languages, where the type system helps ensure
+that is applicable in mainstream OO languages, where the type system ensures
 type safety.
@@
 
+\vspace{-6pt}
 \paragraph{Extensible Parsing Algorithms}
@@
 \emph{separate compilation} as well as \emph{modular
-  type-checking}, and is applied in common OO languages. Moreover, the extensibility of parsing is further
+  type-checking}, and is commonly applicable OO languages. Moreover, the extensibility of parsing is further
 available at language composition.
@@
 
+\vspace{-3pt}
 \paragraph{Parser Combinators} Parser combinators have become more and more
@@
 
+\vspace{-3pt}
 \paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
@@
 standard OO classes in their code.
+\vspace{-5pt}
 

### 2017-09-04 SLE_17/sections/CaseStudy.tex
shortening.
Shorten more.

@@
 
+\vspace{-3pt}
 \subsection{Implementation}\label{subsec:implementation}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{resources/depGraph.pdf}
+    \includegraphics[width=0.7\textwidth]{resources/depGraph.pdf}
     \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
@@
 
+\vspace{-3pt}
 \subsection{Comparison}\label{subsec:cs-comparison}
@@
 \paragraph{Standard of Comparison}
-In the SLOC comparison, all blank lines and comments are excluded,
-and we formatted the code of both implementations to guarantee that
+In terms of SLOC, all blank lines and comments are excluded,
+and we formatted the code of both implementations to ensure that
 the length of each line does not exceed 120 characters. Furthermore,
-because \ilyaimpl{} has extra code such as semantics,
-we removed all irrelevant code and only keep abstract
+because \ilyaimpl{} has extra code like semantics,
+we removed all irrelevant code, only kept abstract
 syntax definition, parser and pretty-printer for each calculus, to
@@
 
-\begin{table*}
+\begin{table*}\tiny
     \centering
@@
     \end{tabular}}
+ 	\vspace{-5pt}
     \caption{Comparison of SLOC and execution time.}
@@
 
+\vspace{-4pt}
 \paragraph{Comparison Results}
@@
 using Object Algebras and class is little, roughly 1\%.
-The usage of longest match combinator slows the performance by 7\%. The main reason of slower
+The use of longest match combinator slows the performance by 7\%. The main reason of slower
 execution time may be the overall structure of the modular parsing approach, because we indeed have
@@
 Since the slowdown seems to be caused by extra method
-dispatching, in future work we wish to investigate techniques such as 
+dispatching, in future work we wish to investigate techniques like
 partial evaluation or meta-programming to eliminate such cost. The

### 2017-09-04 SLE_17/sections/RelatedWork.tex
Shorten more.

@@
 
+\vspace{-4pt}
 \paragraph{Syntactically Extensible Parsing}
@@
 Besides, type-specific languages~\cite{omar14} do support
-separate compilation and type safety. The modularity of parsing is incorporated in its type
-system, which allows users to introduce new syntax for parsing. Whereas in our approach, parsing 
-strategy is based on an external parsing library.
-Our approach does not rely on an elaborated semantics extension, but simply uses a pattern
+separate compilation and type safety. The modularity of parsing is incorporated in its elaborated type
+system. Whereas our approach does not rely on an extension to semantics, but simply uses a pattern
 that is applicable in mainstream OO languages, where the type system ensures
@@
 
-\vspace{-3pt}
+\vspace{-5pt}
 \paragraph{Parser Combinators} Parser combinators have become more and more
@@
 
-\vspace{-3pt}
+\vspace{-5pt}
 \paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
@@
 standard OO classes in their code.
-\vspace{-5pt}
+\vspace{-7pt}
 

### 2017-09-05 SLE_17/sections/ObjectAlgebras.tex
Under 12 pages.

@@
 \vspace{-4pt}
-\lstinputlisting[linerange=5-8]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
+\lstinputlisting[linerange=5-7]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALG
 \vspace{-4pt}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=12-16]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
+\lstinputlisting[linerange=11-14]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_PRINT
 \vspace{-4pt}
@@
 
-%\lstinputlisting[linerange=20-23]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
+%\lstinputlisting[linerange=18-20]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EVAL
 
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=27-29]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
+\lstinputlisting[linerange=24-25]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_ALGEXT
 \vspace{-4pt}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=33-35]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
+\lstinputlisting[linerange=29-30]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_EXTPRINT
 \vspace{-4pt}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=42-43]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
+\lstinputlisting[linerange=37-38]{code/src/papercode/Sec4OA/Code2.scala}% APPLY:linerange=OVERVIEW_OA_MAKEEXP
 \vspace{-4pt}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=8-17]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
+\lstinputlisting[linerange=8-16]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
 \vspace{-4pt}
@@
 \begin{figure}
-  \lstinputlisting[linerange=21-30]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
+  \lstinputlisting[linerange=20-28]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
   \caption{Pattern of modular parsing using Object Algebras.}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=34-39]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
+\lstinputlisting[linerange=32-36]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
 \vspace{-4pt}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=46-49]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
+\lstinputlisting[linerange=43-46]{code/src/papercode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER_CLIENT
 \vspace{-4pt}

### 2017-09-05 SLE_17/sections/PackratParsing.tex
Shorten more.
Under 12 pages.

@@
 
+\vspace{-2pt}
 \begin{itemize}[leftmargin=*]
@@
 \end{itemize}
+\vspace{-2pt}
 
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=9-15]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
+\lstinputlisting[linerange=9-13]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_OO_AST
 \vspace{-4pt}
@@
 \vspace{-4pt}
-\lstinputlisting[linerange=19-26]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
+\lstinputlisting[linerange=17-24]{code/src/papercode/Sec2Packrat/Code1.scala}% APPLY:linerange=PACKRAT_SIMPLE_EXPR
 \vspace{-4pt}
@@
 - A parser which matches an identifier. \\ \hline
-\lstinline{def numericLit: Parser[String]} \\
-- A parser which matches a numeric literal. \\ \hline
 \lstinline{def |[U >: T](q: => Parser[U]): Parser[U]} \\

### 2017-09-05 SLE_17/code/src/papercode/Sec2Packrat/Code1.scala

@@ -8,11 +8,9 @@ trait Code1 extends Common {
 //BEGIN_PACKRAT_OO_AST
 trait Expr { def print: String }
 class Lit(x: Int) extends Expr {
-  def print = x.toString
-}
+  def print = x.toString }
 class Add(e1: Expr, e2: Expr) extends Expr {
-  def print = "(" + e1.print + "+" + e2.print + ")"
-}
+  def print = "(" + e1.print + "+" + e2.print + ")" }
 //END_PACKRAT_OO_AST
 
 //BEGIN_PACKRAT_SIMPLE_EXPR

### 2017-09-05 SLE_17/code/src/papercode/Sec4OA/Code2.scala

@@ -4,35 +4,30 @@ package papercode.Sec4OA
 //BEGIN_OVERVIEW_OA_ALG
 trait Alg[E] {
   def lit(n: Int): E
-  def add(e1: E, e2: E): E
-}
+  def add(e1: E, e2: E): E }
 //END_OVERVIEW_OA_ALG
 
 //BEGIN_OVERVIEW_OA_PRINT
 trait Print extends Alg[String] {
   def lit(n: Int) = n.toString
   def add(e1: String, e2: String) =
-    "(" + e1 + " + " + e2 + ")"
-}
+    "(" + e1 + " + " + e2 + ")" }
 //END_OVERVIEW_OA_PRINT
 
 //BEGIN_OVERVIEW_OA_EVAL
 trait Eval extends Alg[Int] {
   def lit(n: Int) = n
-  def add(e1: Int, e2: Int) = e1 + e2
-}
+  def add(e1: Int, e2: Int) = e1 + e2 }
 //END_OVERVIEW_OA_EVAL
 
 //BEGIN_OVERVIEW_OA_ALGEXT
 trait VarAlg[E] extends Alg[E] {
-  def varE(x: String): E
-}
+  def varE(x: String): E }
 //END_OVERVIEW_OA_ALGEXT
 
 //BEGIN_OVERVIEW_OA_EXTPRINT
 trait VarPrint extends VarAlg[String] with Print {
-  def varE(x: String) = x
-}
+  def varE(x: String) = x }
 //END_OVERVIEW_OA_EXTPRINT
 
 

### 2017-09-05 SLE_17/code/src/papercode/Sec4OA/Code3.scala

@@ -13,8 +13,7 @@ trait Attempt[E] {
     pExpr(alg) ~ ("+" ~> pExpr(alg)) ^^
       { case e1 ~ e2 => alg.add(e1, e2) }
   val pExpr: Alg[E] => Parser[E] = alg =>
-    pLit(alg) ||| pAdd(alg)
-}
+    pLit(alg) ||| pAdd(alg) }
 //END_BASE_OA_PARSER_BAD
 
 //BEGIN_BASE_OA_PARSER
@@ -26,8 +25,7 @@ trait OAParser[E] {
   val pAdd: Parser[E] = pE ~ ("+" ~> pE) ^^
     { case e1 ~ e2 => alg.add(e1, e2) }
   val pExpr: Parser[E] = pLit ||| pAdd
-  val pE: Parser[E] = pExpr
-}
+  val pE: Parser[E] = pExpr }
 //END_BASE_OA_PARSER
 
 //BEGIN_EXT_OA_PARSER
@@ -35,8 +33,7 @@ trait VarOAParser[E] extends OAParser[E] {
   override val alg: VarAlg[E]
   val pVar: Parser[E] = ident ^^ alg.varE
   val pVarExpr: Parser[E] = pExpr ||| pVar
-  override val pE: Parser[E] = pVarExpr
-}
+  override val pE: Parser[E] = pVarExpr }
 //END_EXT_OA_PARSER
 }
 

### 2017-09-05 SLE_17/paper.tex
shortening.
Change code font size.

@@
   lineskip=-1pt,
-  basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
+  basicstyle=\ttfamily\footnotesize,       % the size of the fonts that are used for the code
   numbers=none,                   % where to put the line-numbers
@@
 %% comma separated list
-\keywords{modular parsing, object algebras, semantic modularity}  %% \keywords is optional
+\keywords{modular parsing, Object Algebras, semantic modularity}  %% \keywords is optional
 
@@
 
+\vspace{-6pt}
 \section*{Acknowledgements}
 We would like to thank the anonymous reviewers for their helpful comments. This work has been sponsored by the Hong Kong Research Grant Council projects number 27200514 and 17258816.
+\vspace{-6pt}
 

### 2017-09-05 SLE_17/sections/RelatedWork.tex
Change code font size.
Improve text on wyvern

@@
 section will try to lead a discussion on what difference we have made.
+\end{comment}
+
+\paragraph{Safely Composable Type-Specific Languages} There is almost no work on 
+semantically modular parsing. A notable exception is the work 
+on safely composable type-specific languages~\cite{omar14}. In this
+work the extensible language Wyvern supports the addition of new
+syntax and semantics, while preserving type-safety and separate
+compilation. However this approach and our work have different goals: 
+their approach is aimed at supporting extensibility of Wyvern with
+new syntax; whereas our approach is a general technique aimed at 
+modular parsing of any languages. In contrast to their modular parsing 
+approach, which is directly built-in to the Wyvern language, our approach is 
+library-based and can be used by many  mainstream OO languages.
+
+\begin{comment}
+Besides, type-specific languages do support
+separate compilation and type safety. The modularity of parsing is incorporated in its elaborated type
+system, allowing users to introduce new syntax. Whereas our approach does not rely on an extension to semantics, but simply uses a pattern
+that is applicable in mainstream OO languages, where the type system ensures
+type safety.
 \end{comment}
@@
 setting as well.
-
-Besides, type-specific languages~\cite{omar14} do support
-separate compilation and type safety. The modularity of parsing is incorporated in its elaborated type
-system. Whereas our approach does not rely on an extension to semantics, but simply uses a pattern
-that is applicable in mainstream OO languages, where the type system ensures
-type safety.
 

### 2017-09-07 SLE_17/sections/Introduction.tex
Add the reference.

@@
 aspects of PL implementations.  However, as far as we know, there is
-little work on operations that build/produce ASTs.  In particular the
-problem of how to modularize parsing has not been studied in semantic
-modularization approaches. Techniques such as NOA~\cite{Gouseti2014}
+little work on operations that build/produce ASTs. \cite{SLOANE201520}
+provides a near solution to modular parsing by using case classes, but
+it could be potentially unsafe since exhaustiveness of pattern matching
+is not guaranteed. In particular the
+problem of how to modularize parsing, including its algorithmic challenges,
+has not been studied in other approaches. Techniques such as NOA~\cite{Gouseti2014}
 employ a syntactic modularity approach for parsing in combination with
@@
 
-\item \textbf{A Parsing Technique for OO ASTs:} A simplified version of
-  our technique also enables parsing OO-style ASTs, where new language
-  constructs can be easily added.
-
 \item \textbf{A Methodology for Writing Modular Parsers:} We

### 2017-09-07 SLE_17/sections/RelatedWork.tex
Add the reference.

@@
 
-\paragraph{Safely Composable Type-Specific Languages} There is almost no work on 
+\paragraph{Modular Parsing with Case Classes}
+
+Our work is close to~\cite{SLOANE201520}, in the sense that they also make use of Packrat parsers
+and standard OO inheritance achieving modular parsing. The difference is that case
+classes are used in their work for extensibility. Case classes~\cite{emir2007matching} provide a near solution
+to the Expression Problem since they can be modular added, and hence are widely used in
+practice, but they do not enforce exhaustiveness
+of pattern matching for extensible operations. In other words run-time pattern matching errors
+can happen when writing extensible code with case classes, so that full static type-safety is not
+ensured. Our work uses Object Algebras for full extensibility and type safety, and we have well
+studied the algorithmic challenges of parsing in a modular setting.
+
+\paragraph{Safely Composable Type-Specific Languages}
+
+There is almost no work on 
 semantically modular parsing. A notable exception is the work 
@@
 the parsing code is concise and expressive using Object Algebras.
-On the other hand we are unware of any work on OOP
-that has covered how to do modular parsing for extensible ASTs.
-
-It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
-solution to the Expression Problem. Case classes can be
-modularly added, but they do not enforce
-exhaustiveness of pattern matching for extensible operations. In other
-words run-time pattern matching errors can happen when writing extensible code with case classes.
-So that full static type-safety is not ensured.
-Nevertheless case classes are a pragmatic approach, which is
-widely used in practice. Therefore, modular parsing techniques may be
-of value for extensible code using case classes.  The approach we
-presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
-all that the users need to do is to use case classes, instead of
-standard OO classes in their code.
 \vspace{-7pt}

### 2017-09-08 SLE_17/sections/Introduction.tex
Revised

--- 2017-09-04 SLE_17/sections/Introduction.tex
@@
 aspects of PL implementations.  However, as far as we know, there is
-little work on operations that build/produce ASTs.  In particular the
-problem of how to modularize parsing has not been studied in semantic
-modularization approaches. Techniques such as NOA~\cite{Gouseti2014}
+little work on operations that build/produce extensible ASTs, and
+guarantee type-safety. 
+Sloane and Roberts~\cite{SLOANE201520}
+briefly mention a near solution to modular parsing by using case classes. However
+the solution could be potentially unsafe since exhaustiveness of
+pattern matching for extensible case classes
+is not guaranteed. The
+problem of how to modularize parsing, including its algorithmic challenges,
+has not been studied well-studied yet. Other techniques, such as NOA~\cite{Gouseti2014},
 employ a syntactic modularity approach for parsing in combination with
 a semantic modularity approach for defining operations
-that traverse or process ASTs. This is a shame because parsing is a
-fundamental part of PL implementations, and it ought to be made
+that traverse or process ASTs. Because parsing is a
+fundamental part of PL implementations, it ought to be made
 semantically modular as well, so that the full benefits of semantic

### 2017-09-08 SLE_17/sections/RelatedWork.tex
Revised

@@
 \end{comment}
-
-\paragraph{Modular Parsing with Case Classes}
-
-Our work is close to~\cite{SLOANE201520}, in the sense that they also make use of Packrat parsers
-and standard OO inheritance achieving modular parsing. The difference is that case
-classes are used in their work for extensibility. Case classes~\cite{emir2007matching} provide a near solution
-to the Expression Problem since they can be modular added, and hence are widely used in
-practice, but they do not enforce exhaustiveness
-of pattern matching for extensible operations. In other words run-time pattern matching errors
-can happen when writing extensible code with case classes, so that full static type-safety is not
-ensured. Our work uses Object Algebras for full extensibility and type safety, and we have well
-studied the algorithmic challenges of parsing in a modular setting.
 
@@
 the parsing code is concise and expressive using Object Algebras.
+
+Case classes in Scala can encode algebraic datatypes that allow the
+addition of new constructors. However such ``open'' case classes 
+do not enforce  exhaustiveness
+of pattern matching for extensible operations, and thus do not provide 
+a full solution to the Expression Problem. Nevertheless case classes 
+are widely used in practice, and a solution for parsing open case
+classes (and composing such parsers) is quite relevant in practice. 
+The techniques in Section~\ref{sec:inheritance} can be readily adapted to work with case classes. 
+The work by Sloane and Roberts~\cite{SLOANE201520} on a modular Oberon 
+compiler applied similar techniques with packrat parsers and case classes.
+In our work we use Object Algebras for full extensibility and type safety, and we have well
+studied the algorithmic challenges of parsing in a modular setting.
 \vspace{-7pt}
